### 2024-05-01 paper/main.tex
Updates abstract

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024

% ready for submission
\usepackage{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\title{Expert-in-the-Loop Causal Discovery}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
	Causal discovery is a fundamental problem in causal inference, and numerous algorithms have been developed to address it. However, their adoption in applied disciplines  as been limited, possibly due to a lack of confidence in these algorithms stemming from them making obvious mistakes, and the difficulties in choosing the right algorithm for specific datasets. Consequently, in applied disciplines, these causal Directed Acyclic Graphs (DAGs) are predominantly constructed manually based on domain expertise. However, some researchers do validate and modify their DAGs by testing them against data through Conditional Independence (CI) tests. To assist researchers in this task, we have developed a web-based tool to aid researchers in constructing, validating, and modifying DAGs. The tool uses effect size measures from CI testing to iteratively suggest potential additions or removals of edges between variables, while allowing experts to choose between recommendations and specify the directionality of the edges. Furthermore, we introduce a novel effect size measure for mixed data, enabling the tool to work for any combination of variable types. Empirically, we show that taking a greedy approach on the tool's recommendations can perform comparably to causal discovery algorithms, provided that the expert correctly determines the direction of the edges in at least one out of three cases, with performance improving as accuracy increases.
\end{abstract}

\section{Submission of papers to NeurIPS 2024}

Please read the instructions below carefully and follow them faithfully.

\subsection{Style}

Papers to be submitted to NeurIPS 2024 must be prepared according to the
instructions presented here. Papers may only be up to {\bf nine} pages long,
including figures. Additional pages \emph{containing only acknowledgments and
references} are allowed. Papers that exceed the page limit will not be
reviewed, or in any other way considered for presentation at the conference.

The margins in 2024 are the same as those in previous years.

Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
NeurIPS website as indicated below. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for NeurIPS and other conference information are available on
the website at
\begin{center}
  \url{http://www.neurips.cc/}
\end{center}
The file \verb+neurips_2024.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The only supported style file for NeurIPS 2024 is \verb+neurips_2024.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
  Microsoft Word, and RTF are no longer supported!}

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit, as long as you do not say which conference it was submitted to. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to NeurIPS.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+neurips_2024.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2024+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2024}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous'' and include a copy of the anonymized paper in the supplementary material.

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Math}
Note that display math in bare TeX commands will not create correct line numbers for submission. Please use LaTeX (or AMSTeX) commands for unnumbered display math. (You really shouldn't be using \$\$ anyway; see \url{https://tex.stackexchange.com/questions/503/why-is-preferable-to} and \url{https://tex.stackexchange.com/questions/40492/what-are-the-differences-between-align-equation-and-displaymath} for more information.)

\subsection{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2024/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
\end{ack}

\section*{References}

References follow the acknowledgments in the camera-ready paper. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
Note that the Reference section does not count towards the page limit.
\medskip

{
\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Appendix / supplemental material}

Optionally include supplemental material (complete proofs, additional experiments and plots) in appendix.
All such materials \textbf{SHOULD be included in the main submission.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{NeurIPS Paper Checklist}

%%% BEGIN INSTRUCTIONS %%%
The checklist is designed to encourage best practices for responsible machine learning research, addressing issues of reproducibility, transparency, research ethics, and societal impact. Do not remove the checklist: {\bf The papers not including the checklist will be desk rejected.} The checklist should follow the references and precede the (optional) supplemental material.  The checklist does NOT count towards the page
limit.

Please read the checklist guidelines carefully for information on how to answer these questions. For each question in the checklist:
\begin{itemize}
    \item You should answer \answerYes{}, \answerNo{}, or \answerNA{}.
    \item \answerNA{} means either that the question is Not Applicable for that particular paper or the relevant information is Not Available.
    \item Please provide a short (1–2 sentence) justification right after your answer (even for NA). 
   % \item {\bf The papers not including the checklist will be desk rejected.}
\end{itemize}

{\bf The checklist answers are an integral part of your paper submission.} They are visible to the reviewers, area chairs, senior area chairs, and ethics reviewers. You will be asked to also include it (after eventual revisions) with the final version of your paper, and its final version will be published with the paper.

The reviewers of your paper will be asked to use the checklist as one of the factors in their evaluation. While "\answerYes{}" is generally preferable to "\answerNo{}", it is perfectly acceptable to answer "\answerNo{}" provided a proper justification is given (e.g., "error bars are not reported because it would be too computationally expensive" or "we were unable to find the license for the dataset we used"). In general, answering "\answerNo{}" or "\answerNA{}" is not grounds for rejection. While the questions are phrased in a binary way, we acknowledge that the true answer is often more nuanced, so please just use your best judgment and write a justification to elaborate. All supporting evidence can appear either in the main paper or the supplemental material, provided in appendix. If you answer \answerYes{} to a question, in the justification please point to the section(s) where related material for the question can be found.

IMPORTANT, please:
\begin{itemize}
    \item {\bf Delete this instruction block, but keep the section heading ``NeurIPS paper checklist"},
    \item  {\bf Keep the checklist subsection headings, questions/answers and guidelines below.}
    \item {\bf Do not modify the questions and only use the provided macros for your answers}.
\end{itemize}

%%% END INSTRUCTIONS %%%

\begin{enumerate}

\item {\bf Claims}
    \item[] Question: Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the abstract and introduction do not include the claims made in the paper.
        \item The abstract and/or introduction should clearly state the claims made, including the contributions made in the paper and important assumptions and limitations. A No or NA answer to this question will not be perceived well by the reviewers. 
        \item The claims made should match theoretical and experimental results, and reflect how much the results can be expected to generalize to other settings. 
        \item It is fine to include aspirational goals as motivation as long as it is clear that these goals are not attained by the paper. 
    \end{itemize}

\item {\bf Limitations}
    \item[] Question: Does the paper discuss the limitations of the work performed by the authors?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper has no limitation while the answer No means that the paper has limitations, but those are not discussed in the paper. 
        \item The authors are encouraged to create a separate "Limitations" section in their paper.
        \item The paper should point out any strong assumptions and how robust the results are to violations of these assumptions (e.g., independence assumptions, noiseless settings, model well-specification, asymptotic approximations only holding locally). The authors should reflect on how these assumptions might be violated in practice and what the implications would be.
        \item The authors should reflect on the scope of the claims made, e.g., if the approach was only tested on a few datasets or with a few runs. In general, empirical results often depend on implicit assumptions, which should be articulated.
        \item The authors should reflect on the factors that influence the performance of the approach. For example, a facial recognition algorithm may perform poorly when image resolution is low or images are taken in low lighting. Or a speech-to-text system might not be used reliably to provide closed captions for online lectures because it fails to handle technical jargon.
        \item The authors should discuss the computational efficiency of the proposed algorithms and how they scale with dataset size.
        \item If applicable, the authors should discuss possible limitations of their approach to address problems of privacy and fairness.
        \item While the authors might fear that complete honesty about limitations might be used by reviewers as grounds for rejection, a worse outcome might be that reviewers discover limitations that aren't acknowledged in the paper. The authors should use their best judgment and recognize that individual actions in favor of transparency play an important role in developing norms that preserve the integrity of the community. Reviewers will be specifically instructed to not penalize honesty concerning limitations.
    \end{itemize}

\item {\bf Theory Assumptions and Proofs}
    \item[] Question: For each theoretical result, does the paper provide the full set of assumptions and a complete (and correct) proof?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include theoretical results. 
        \item All the theorems, formulas, and proofs in the paper should be numbered and cross-referenced.
        \item All assumptions should be clearly stated or referenced in the statement of any theorems.
        \item The proofs can either appear in the main paper or the supplemental material, but if they appear in the supplemental material, the authors are encouraged to provide a short proof sketch to provide intuition. 
        \item Inversely, any informal proof provided in the core of the paper should be complemented by formal proofs provided in appendix or supplemental material.
        \item Theorems and Lemmas that the proof relies upon should be properly referenced. 
    \end{itemize}

    \item {\bf Experimental Result Reproducibility}
    \item[] Question: Does the paper fully disclose all the information needed to reproduce the main experimental results of the paper to the extent that it affects the main claims and/or conclusions of the paper (regardless of whether the code and data are provided or not)?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include experiments.
        \item If the paper includes experiments, a No answer to this question will not be perceived well by the reviewers: Making the paper reproducible is important, regardless of whether the code and data are provided or not.
        \item If the contribution is a dataset and/or model, the authors should describe the steps taken to make their results reproducible or verifiable. 
        \item Depending on the contribution, reproducibility can be accomplished in various ways. For example, if the contribution is a novel architecture, describing the architecture fully might suffice, or if the contribution is a specific model and empirical evaluation, it may be necessary to either make it possible for others to replicate the model with the same dataset, or provide access to the model. In general. releasing code and data is often one good way to accomplish this, but reproducibility can also be provided via detailed instructions for how to replicate the results, access to a hosted model (e.g., in the case of a large language model), releasing of a model checkpoint, or other means that are appropriate to the research performed.
        \item While NeurIPS does not require releasing code, the conference does require all submissions to provide some reasonable avenue for reproducibility, which may depend on the nature of the contribution. For example
        \begin{enumerate}
            \item If the contribution is primarily a new algorithm, the paper should make it clear how to reproduce that algorithm.
            \item If the contribution is primarily a new model architecture, the paper should describe the architecture clearly and fully.
            \item If the contribution is a new model (e.g., a large language model), then there should either be a way to access this model for reproducing the results or a way to reproduce the model (e.g., with an open-source dataset or instructions for how to construct the dataset).
            \item We recognize that reproducibility may be tricky in some cases, in which case authors are welcome to describe the particular way they provide for reproducibility. In the case of closed-source models, it may be that access to the model is limited in some way (e.g., to registered users), but it should be possible for other researchers to have some path to reproducing or verifying the results.
        \end{enumerate}
    \end{itemize}

\item {\bf Open access to data and code}
    \item[] Question: Does the paper provide open access to the data and code, with sufficient instructions to faithfully reproduce the main experimental results, as described in supplemental material?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that paper does not include experiments requiring code.
        \item Please see the NeurIPS code and data submission guidelines (\url{https://nips.cc/public/guides/CodeSubmissionPolicy}) for more details.
        \item While we encourage the release of code and data, we understand that this might not be possible, so “No” is an acceptable answer. Papers cannot be rejected simply for not including code, unless this is central to the contribution (e.g., for a new open-source benchmark).
        \item The instructions should contain the exact command and environment needed to run to reproduce the results. See the NeurIPS code and data submission guidelines (\url{https://nips.cc/public/guides/CodeSubmissionPolicy}) for more details.
        \item The authors should provide instructions on data access and preparation, including how to access the raw data, preprocessed data, intermediate data, and generated data, etc.
        \item The authors should provide scripts to reproduce all experimental results for the new proposed method and baselines. If only a subset of experiments are reproducible, they should state which ones are omitted from the script and why.
        \item At submission time, to preserve anonymity, the authors should release anonymized versions (if applicable).
        \item Providing as much information as possible in supplemental material (appended to the paper) is recommended, but including URLs to data and code is permitted.
    \end{itemize}

\item {\bf Experimental Setting/Details}
    \item[] Question: Does the paper specify all the training and test details (e.g., data splits, hyperparameters, how they were chosen, type of optimizer, etc.) necessary to understand the results?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include experiments.
        \item The experimental setting should be presented in the core of the paper to a level of detail that is necessary to appreciate the results and make sense of them.
        \item The full details can be provided either with the code, in appendix, or as supplemental material.
    \end{itemize}

\item {\bf Experiment Statistical Significance}
    \item[] Question: Does the paper report error bars suitably and correctly defined or other appropriate information about the statistical significance of the experiments?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include experiments.
        \item The authors should answer "Yes" if the results are accompanied by error bars, confidence intervals, or statistical significance tests, at least for the experiments that support the main claims of the paper.
        \item The factors of variability that the error bars are capturing should be clearly stated (for example, train/test split, initialization, random drawing of some parameter, or overall run with given experimental conditions).
        \item The method for calculating the error bars should be explained (closed form formula, call to a library function, bootstrap, etc.)
        \item The assumptions made should be given (e.g., Normally distributed errors).
        \item It should be clear whether the error bar is the standard deviation or the standard error of the mean.
        \item It is OK to report 1-sigma error bars, but one should state it. The authors should preferably report a 2-sigma error bar than state that they have a 96\% CI, if the hypothesis of Normality of errors is not verified.
        \item For asymmetric distributions, the authors should be careful not to show in tables or figures symmetric error bars that would yield results that are out of range (e.g. negative error rates).
        \item If error bars are reported in tables or plots, The authors should explain in the text how they were calculated and reference the corresponding figures or tables in the text.
    \end{itemize}

\item {\bf Experiments Compute Resources}
    \item[] Question: For each experiment, does the paper provide sufficient information on the computer resources (type of compute workers, memory, time of execution) needed to reproduce the experiments?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not include experiments.
        \item The paper should indicate the type of compute workers CPU or GPU, internal cluster, or cloud provider, including relevant memory and storage.
        \item The paper should provide the amount of compute required for each of the individual experimental runs as well as estimate the total compute. 
        \item The paper should disclose whether the full research project required more compute than the experiments reported in the paper (e.g., preliminary or failed experiments that didn't make it into the paper). 
    \end{itemize}

\item {\bf Code Of Ethics}
    \item[] Question: Does the research conducted in the paper conform, in every respect, with the NeurIPS Code of Ethics \url{https://neurips.cc/public/EthicsGuidelines}?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the authors have not reviewed the NeurIPS Code of Ethics.
        \item If the authors answer No, they should explain the special circumstances that require a deviation from the Code of Ethics.
        \item The authors should make sure to preserve anonymity (e.g., if there is a special consideration due to laws or regulations in their jurisdiction).
    \end{itemize}

\item {\bf Broader Impacts}
    \item[] Question: Does the paper discuss both potential positive societal impacts and negative societal impacts of the work performed?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that there is no societal impact of the work performed.
        \item If the authors answer NA or No, they should explain why their work has no societal impact or why the paper does not address societal impact.
        \item Examples of negative societal impacts include potential malicious or unintended uses (e.g., disinformation, generating fake profiles, surveillance), fairness considerations (e.g., deployment of technologies that could make decisions that unfairly impact specific groups), privacy considerations, and security considerations.
        \item The conference expects that many papers will be foundational research and not tied to particular applications, let alone deployments. However, if there is a direct path to any negative applications, the authors should point it out. For example, it is legitimate to point out that an improvement in the quality of generative models could be used to generate deepfakes for disinformation. On the other hand, it is not needed to point out that a generic algorithm for optimizing neural networks could enable people to train models that generate Deepfakes faster.
        \item The authors should consider possible harms that could arise when the technology is being used as intended and functioning correctly, harms that could arise when the technology is being used as intended but gives incorrect results, and harms following from (intentional or unintentional) misuse of the technology.
        \item If there are negative societal impacts, the authors could also discuss possible mitigation strategies (e.g., gated release of models, providing defenses in addition to attacks, mechanisms for monitoring misuse, mechanisms to monitor how a system learns from feedback over time, improving the efficiency and accessibility of ML).
    \end{itemize}

\item {\bf Safeguards}
    \item[] Question: Does the paper describe safeguards that have been put in place for responsible release of data or models that have a high risk for misuse (e.g., pretrained language models, image generators, or scraped datasets)?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper poses no such risks.
        \item Released models that have a high risk for misuse or dual-use should be released with necessary safeguards to allow for controlled use of the model, for example by requiring that users adhere to usage guidelines or restrictions to access the model or implementing safety filters. 
        \item Datasets that have been scraped from the Internet could pose safety risks. The authors should describe how they avoided releasing unsafe images.
        \item We recognize that providing effective safeguards is challenging, and many papers do not require this, but we encourage authors to take this into account and make a best faith effort.
    \end{itemize}

\item {\bf Licenses for existing assets}
    \item[] Question: Are the creators or original owners of assets (e.g., code, data, models), used in the paper, properly credited and are the license and terms of use explicitly mentioned and properly respected?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not use existing assets.
        \item The authors should cite the original paper that produced the code package or dataset.
        \item The authors should state which version of the asset is used and, if possible, include a URL.
        \item The name of the license (e.g., CC-BY 4.0) should be included for each asset.
        \item For scraped data from a particular source (e.g., website), the copyright and terms of service of that source should be provided.
        \item If assets are released, the license, copyright information, and terms of use in the package should be provided. For popular datasets, \url{paperswithcode.com/datasets} has curated licenses for some datasets. Their licensing guide can help determine the license of a dataset.
        \item For existing datasets that are re-packaged, both the original license and the license of the derived asset (if it has changed) should be provided.
        \item If this information is not available online, the authors are encouraged to reach out to the asset's creators.
    \end{itemize}

\item {\bf New Assets}
    \item[] Question: Are new assets introduced in the paper well documented and is the documentation provided alongside the assets?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not release new assets.
        \item Researchers should communicate the details of the dataset/code/model as part of their submissions via structured templates. This includes details about training, license, limitations, etc. 
        \item The paper should discuss whether and how consent was obtained from people whose asset is used.
        \item At submission time, remember to anonymize your assets (if applicable). You can either create an anonymized URL or include an anonymized zip file.
    \end{itemize}

\item {\bf Crowdsourcing and Research with Human Subjects}
    \item[] Question: For crowdsourcing experiments and research with human subjects, does the paper include the full text of instructions given to participants and screenshots, if applicable, as well as details about compensation (if any)? 
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not involve crowdsourcing nor research with human subjects.
        \item Including this information in the supplemental material is fine, but if the main contribution of the paper involves human subjects, then as much detail as possible should be included in the main paper. 
        \item According to the NeurIPS Code of Ethics, workers involved in data collection, curation, or other labor should be paid at least the minimum wage in the country of the data collector. 
    \end{itemize}

\item {\bf Institutional Review Board (IRB) Approvals or Equivalent for Research with Human Subjects}
    \item[] Question: Does the paper describe potential risks incurred by study participants, whether such risks were disclosed to the subjects, and whether Institutional Review Board (IRB) approvals (or an equivalent approval/review based on the requirements of your country or institution) were obtained?
    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
    \item[] Justification: \justificationTODO{}
    \item[] Guidelines:
    \begin{itemize}
        \item The answer NA means that the paper does not involve crowdsourcing nor research with human subjects.
        \item Depending on the country in which research is conducted, IRB approval (or equivalent) may be required for any human subjects research. If you obtained IRB approval, you should clearly state this in the paper. 
        \item We recognize that the procedures for this may vary significantly between institutions and locations, and we expect authors to adhere to the NeurIPS Code of Ethics and the guidelines for their institution. 
        \item For initial submissions, do not include any information that would break anonymity (if applicable), such as the institution conducting the review.
    \end{itemize}

\end{enumerate}

\end{document}

### 2024-05-02 paper/main.tex
Typo fix

@@
 	Causal discovery is a fundamental problem in causal inference, and numerous algorithms have been developed to address it.
-However, their adoption in applied disciplines  as been limited, possibly due to a lack of confidence in these algorithms stemming from them making obvious mistakes, and the difficulties in choosing the right algorithm for specific datasets.
+However, their adoption in applied disciplines  has been limited, possibly due to a lack of confidence in these algorithms stemming from them making obvious mistakes, and the difficulties in choosing the right algorithm for specific datasets.
 Consequently, in applied disciplines, these causal Directed Acyclic Graphs (DAGs) are predominantly constructed manually based on domain expertise.

### 2024-05-06 paper/main.tex
Updates

@@
 
-\title{Expert-in-the-Loop Causal Discovery}
+% Packages added by me.
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{todonotes}
+
+\title{Title}
 
@@
 \begin{abstract}
-	Causal discovery is a fundamental problem in causal inference, and numerous algorithms have been developed to address it.
-However, their adoption in applied disciplines  has been limited, possibly due to a lack of confidence in these algorithms stemming from them making obvious mistakes, and the difficulties in choosing the right algorithm for specific datasets.
-Consequently, in applied disciplines, these causal Directed Acyclic Graphs (DAGs) are predominantly constructed manually based on domain expertise.
-However, some researchers do validate and modify their DAGs by testing them against data through Conditional Independence (CI) tests.
-To assist researchers in this task, we have developed a web-based tool to aid researchers in constructing, validating, and modifying DAGs.
-The tool uses effect size measures from CI testing to iteratively suggest potential additions or removals of edges between variables, while allowing experts to choose between recommendations and specify the directionality of the edges.
-Furthermore, we introduce a novel effect size measure for mixed data, enabling the tool to work for any combination of variable types.
-Empirically, we show that taking a greedy approach on the tool's recommendations can perform comparably to causal discovery algorithms, provided that the expert correctly determines the direction of the edges in at least one out of three cases, with performance improving as accuracy increases.
+	Causal discovery is a fundamental problem in causal inference, and
+	numerous algorithms have been developed to address it. However, their
+	adoption in applied disciplines  has been limited, possibly due to a
+	lack of confidence in these algorithms stemming from them making
+	obvious mistakes, and the difficulties in choosing the right algorithm
+	for specific datasets. Consequently, in applied disciplines, these
+	causal Directed Acyclic Graphs (DAGs) are predominantly constructed
+	manually based on domain expertise. This makes it vital to test and
+	potentially modify these models against data before using them for any
+	analysis. Local testing based on Conditional Independence tests and
+	their effect sizes have been commonly used for this purpose. However,
+	one of the current limitations of this approach is that there are no
+	effect sizes for mixed data. To address this, we introduce a novel
+	effect size measure for mixed data based on canonical correlations.
+	This effect size is a generalization of some of the effect sizes used
+	in special cases to mixed variables. Moreover, to assist researchers in
+	constructing DAGs manually, we use the proposed mixed data effect size
+	measure to build a web-based tool that can help researchers in
+	constructing, validating, and modifying their DAGs with the help of
+	available data. The tool iteratively suggests modifications to DAGs
+	based on correlations that are not explained/over-explained by the DAG,
+	while the expert chooses the direction of the edge between those
+	variables. Empirically, we show that taking a greedy approach on the
+	tool's recommendations can perform comparably to causal discovery
+	algorithms, provided that the expert correctly determines the direction
+	of the edges in at least one out of three cases, with performance
+	improving as accuracy increases.
 \end{abstract}
 
-\section{Submission of papers to NeurIPS 2024}
+\section{Introduction}
+Para 1: Many Causal discovery algorithms but commonly constructed manually.
 
-Please read the instructions below carefully and follow them faithfully.
+To perform any causal effect identification or estimation in Directed Acyclic
+Graph(DAG) based causal inference usually requires to come up with the DAG
+structure. Learning these DAG structures from data is known as \emph{causal
+discovery}. Plenty of algorithms have been developed for these in the
+literature such as constraint-based methods such as PC, FCI, score based
+methods such as Hill-Climb Search, GES, and constrained continuous optimization
+based such as No TEARS.
 
-\subsection{Style}
+Para 2: Reasons behind this
 
-Papers to be submitted to NeurIPS 2024 must be prepared according to the
-instructions presented here. Papers may only be up to {\bf nine} pages long,
-including figures. Additional pages \emph{containing only acknowledgments and
-references} are allowed. Papers that exceed the page limit will not be
-reviewed, or in any other way considered for presentation at the conference.
+Even with the existence of all these algorithms for constructing DAGs from
+data, in applied fields researchers prefer to construct DAGs by hand. We
+hypothesize that the adaption of these algorithms in applied literature is
+low due to the following reasons:
+\begin{enumerate}
+	\item Algorithms can make very obvious mistakes at times making it harder
+		for the user to trust their outputs.
+	\item Difficult to choose a single algorithm and their hyper-parameters.
+\end{enumerate}
 
-The margins in 2024 are the same as those in previous years.
+Para 3: Importance of testing
 
-Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
-NeurIPS website as indicated below. Please make sure you use the current files
-and not previous versions. Tweaking the style files may be grounds for
-rejection.
+When these models are constructed manually, it is extremely important that they
+are tested before they are used for making any inference. Local testing methods
+have been used for this purpose, where the implied conditional independences of
+the model are testing in the data using statistical tests. These tests can then
+be used to modify our models.
 
-\subsection{Retrieval of style files}
+Para 4: Mixed data testing
 
-The style files for NeurIPS and other conference information are available on
-the website at
-\begin{center}
-  \url{http://www.neurips.cc/}
-\end{center}
-The file \verb+neurips_2024.pdf+ contains these instructions and illustrates the
-various formatting requirements your NeurIPS paper must satisfy.
+Para 5: Contributions
 
-The only supported style file for NeurIPS 2024 is \verb+neurips_2024.sty+,
-rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
-  Microsoft Word, and RTF are no longer supported!}
+\section{Background}
+\subsection{Continuous Variables}
+\subsection{Discrete Variables}
+\subsection{Ordinal Variables}
+\subsection{Ordinal and Continuous Variables}
 
-The \LaTeX{} style file contains three optional arguments: \verb+final+, which
-creates a camera-ready copy, \verb+preprint+, which creates a preprint for
-submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
-\verb+natbib+ package for you in case of package clash.
+\section{Effect size for mixed data}
 
-\paragraph{Preprint option}
-If you wish to post a preprint of your work online, e.g., on arXiv, using the
-NeurIPS style, please use the \verb+preprint+ option. This will create a
-nonanonymized version of your work with the text ``Preprint. Work in progress.''
-in the footer. This version may be distributed as you see fit, as long as you do not say which conference it was submitted to. Please \textbf{do
-  not} use the \verb+final+ option, which should \textbf{only} be used for
-papers accepted to NeurIPS.
+\section{Web Tool}
+As most of the DAGs in practice are built manually, we use our mixed data effect size measure to build a web-tool that aids researches in building DAGs.
 
-At submission time, please omit the \verb+final+ and \verb+preprint+
-options. This will anonymize your submission and add line numbers to aid
-review. Please do \emph{not} refer to these line numbers in your paper as they
-will be removed during generation of camera-ready copies.
+\section{Empirical Analysis}
+We performed empirical analysis to compare how well does creating a graph manually based on effect sizes and p-value compare to algorithmic approaches.
+\subsection{Simulating an expert}
+To simulate an expert building a model manually, we wrote a simple greedy simulation as shown in the Figure. The expert always starts with a graph with
+no edges and chooses the pair of variables with the highest correlation between them. The expert based on some accuracy measure decides the direction
+of the edge between these variables such that the new added edge does not create a cycle in the DAG:
 
-The file \verb+neurips_2024.tex+ may be used as a ``shell'' for writing your
-paper. All you have to do is replace the author, title, abstract, and text of
-the paper with your own.
+\todo{Summarize this section and equaiton in an algorithm}
+\begin{equation}
+	\begin{split}
+		x &= \textnormal{rand}([0, 1]) \\
+		O(\alpha) &= \begin{cases} 
+			M \rightarrow Y, & \textnormal{if  } x <= \alpha \\
+			\textnormal{rand}(M \rightarrow Y, M \leftarrow Y, None) & \textnormal{otherwise} \\
+			     \end{cases} \\
+	\end{split}
+\end{equation}
 
-The formatting instructions contained in these style files are summarized in
-Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.
+After adding each edge, we also check if the p-value of any existing edge shows independence. If that happens we remove that edge and blacklist that edge. 
+Blacklisted edges are not suggested to the expert in future iterations. We repeat this till we have explained all the correlations in the dataset.
 
-\section{General formatting instructions}
-\label{gen_inst}
+This is a possibility that this greedy expert gets stuck at incorrect graph structures as it does not do any backtracking to fix its earlier mistakes.
 
-The text must be confined within a rectangle 5.5~inches (33~picas) wide and
-9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
-type with a vertical spacing (leading) of 11~points.  Times New Roman is the
-preferred typeface throughout, and will be selected for you by default.
-Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
-indentation.
-
-The paper title should be 17~point, initial caps/lower case, bold, centered
-between two horizontal rules. The top rule should be 4~points thick and the
-bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
-below the title to rules. All pages should start at 1~inch (6~picas) from the
-top of the page.
-
-For the final version, authors' names are set in boldface, and each name is
-centered above the corresponding address. The lead author's name is to be listed
-first (left-most), and the co-authors' names (if different address) are set to
-follow. If there is only one co-author, list both author and co-author side by
-side.
-
-Please pay special attention to the instructions in Section \ref{others}
-regarding figures, tables, acknowledgments, and references.
-
-\section{Headings: first level}
-\label{headings}
-
-All headings should be lower case (except for first word and proper nouns),
-flush left, and bold.
-
-First-level headings should be in 12-point type.
-
-\subsection{Headings: second level}
-
-Second-level headings should be in 10-point type.
-
-\subsubsection{Headings: third level}
-
-Third-level headings should be in 10-point type.
-
-\paragraph{Paragraphs}
-
-There is also a \verb+\paragraph+ command available, which sets the heading in
-bold, flush left, and inline with the text, with the heading followed by 1\,em
-of space.
-
-\section{Citations, figures, tables, references}
-\label{others}
-
-These instructions apply to everyone.
-
-\subsection{Citations within the text}
-
-The \verb+natbib+ package will be loaded for you by default.  Citations may be
-author/year or numeric, as long as you maintain internal consistency.  As to the
-format of the references themselves, any style is acceptable as long as it is
-used consistently.
-
-The documentation for \verb+natbib+ may be found at
-\begin{center}
-  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
-\end{center}
-Of note is the command \verb+\citet+, which produces citations appropriate for
-use in inline text.  For example,
-\begin{verbatim}
-   \citet{hasselmo} investigated\dots
-\end{verbatim}
-produces
-\begin{quote}
-  Hasselmo, et al.\ (1995) investigated\dots
-\end{quote}
-
-If you wish to load the \verb+natbib+ package with options, you may add the
-following before loading the \verb+neurips_2024+ package:
-\begin{verbatim}
-   \PassOptionsToPackage{options}{natbib}
-\end{verbatim}
-
-If \verb+natbib+ clashes with another package you load, you can add the optional
-argument \verb+nonatbib+ when loading the style file:
-\begin{verbatim}
-   \usepackage[nonatbib]{neurips_2024}
-\end{verbatim}
-
-As submission is double blind, refer to your own published work in the third
-person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
-previous work [4].'' If you cite your other papers that are not widely available
-(e.g., a journal paper under review), use anonymous author names in the
-citation, e.g., an author of the form ``A.\ Anonymous'' and include a copy of the anonymized paper in the supplementary material.
-
-\subsection{Footnotes}
-
-Footnotes should be used sparingly.  If you do require a footnote, indicate
-footnotes with a number\footnote{Sample of the first footnote.} in the
-text. Place the footnotes at the bottom of the page on which they appear.
-Precede the footnote with a horizontal rule of 2~inches (12~picas).
-
-Note that footnotes are properly typeset \emph{after} punctuation
-marks.\footnote{As in this example.}
-
-\subsection{Figures}
-
-\begin{figure}
-  \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
-\end{figure}
-
-All artwork must be neat, clean, and legible. Lines should be dark enough for
-purposes of reproduction. The figure number and caption always appear after the
-figure. Place one line space before the figure caption and one line space after
-the figure. The figure caption should be lower case (except for first word and
-proper nouns); figures are numbered consecutively.
-
-You may use color figures.  However, it is best for the figure captions and the
-paper body to be legible if the paper is printed in either black/white or in
-color.
-
-\subsection{Tables}
-
-All tables must be centered, neat, clean and legible.  The table number and
-title always appear before the table.  See Table~\ref{sample-table}.
-
-Place one line space before the table title, one line space after the
-table title, and one line space after the table. The table title must
-be lower case (except for first word and proper nouns); tables are
-numbered consecutively.
-
-Note that publication-quality tables \emph{do not contain vertical rules.} We
-strongly suggest the use of the \verb+booktabs+ package, which allows for
-typesetting high-quality, professional tables:
-\begin{center}
-  \url{https://www.ctan.org/pkg/booktabs}
-\end{center}
-This package was used to typeset Table~\ref{sample-table}.
-
-\begin{table}
-  \caption{Sample table title}
-  \label{sample-table}
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule(r){1-2}
-    Name     & Description     & Size ($\mu$m) \\
-    \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
-  \end{tabular}
-\end{table}
-
-\subsection{Math}
-Note that display math in bare TeX commands will not create correct line numbers for submission.
-Please use LaTeX (or AMSTeX) commands for unnumbered display math.
-(You really shouldn't be using \$\$ anyway; see \url{https://tex.stackexchange.com/questions/503/why-is-preferable-to} and \url{https://tex.stackexchange.com/questions/40492/what-are-the-differences-between-align-equation-and-displaymath} for more information.)
-
-\subsection{Final instructions}
-
-Do not change any aspects of the formatting parameters in the style files.  In
-particular, do not modify the width or length of the rectangle the text should
-fit into, and do not change font sizes (except perhaps in the
-\textbf{References} section; see below). Please note that pages should be
-numbered.
-
-\section{Preparing PDF files}
-
-Please prepare submission files with paper size ``US Letter,'' and not, for
-example, ``A4.''
-
-Fonts were the main cause of problems in the past years. Your PDF file must only
-contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
-achieve this.
-
-\begin{itemize}
-
-\item You should directly generate PDF files using \verb+pdflatex+.
-
-\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
-  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
-  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
-  available out-of-the-box on most Linux machines.
-
-\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
-  "solid" shapes instead.
-
-\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
-  the equivalent AMS Fonts:
-\begin{verbatim}
-   \usepackage{amsfonts}
-\end{verbatim}
-followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
-for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
-workaround for reals, natural and complex:
-\begin{verbatim}
-   \newcommand{\RR}{I\!\!R} %real numbers
-   \newcommand{\Nat}{I\!\!N} %natural numbers
-   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
-\end{verbatim}
-Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.
-
-\end{itemize}
-
-If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
-you to fix it.
-
-\subsection{Margins in \LaTeX{}}
-
-Most of the margin problems come from figures positioned by hand using
-\verb+\special+ or other commands. We suggest using the command
-\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
-figure width as a multiple of the line width as in the example below:
-\begin{verbatim}
-   \usepackage[pdftex]{graphicx} ...
-   \includegraphics[width=0.8\linewidth]{myfile.pdf}
-\end{verbatim}
-See Section 4.4 in the graphics bundle documentation
-(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})
-
-A number of width problems arise when \LaTeX{} cannot properly hyphenate a
-line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
-necessary.
-
-\begin{ack}
-Use unnumbered first level headings for the acknowledgments. All acknowledgments
-go at the end of the paper before the list of references. Moreover, you are required to declare
-funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
-More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2024/PaperInformation/FundingDisclosure}.
-
-Do {\bf not} include this section in the anonymized submission, only in the final paper.
-You can use the \texttt{ack} environment provided in the style file to automatically hide this section in the anonymized submission.
-\end{ack}
+\subsection{Results}
 
 \section*{References}
-
-References follow the acknowledgments in the camera-ready paper. Use unnumbered first-level heading for
-the references. Any choice of citation style is acceptable as long as you are
-consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
-when listing the references.
-Note that the Reference section does not count towards the page limit.
-\medskip
-
-{
-\small
-
-[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
-connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
-(eds.), {\it Advances in Neural Information Processing Systems 7},
-pp.\ 609--616. Cambridge, MA: MIT Press.
-
-[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
-  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
-TELOS/Springer--Verlag.
-
-[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
-recall at excitatory recurrent synapses and cholinergic modulation in rat
-hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
-}
+\bibliographystyle{plain}
+\bibliography{references.bib}
 

### 2024-05-06 paper/checklist.tex
Separate out checklist file

--- 2024-05-06 paper/main.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2024
-
-% ready for submission
-\usepackage{neurips_2024}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2024}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2024}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2024}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-
-% Packages added by me.
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{todonotes}
-
-\title{Title}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{%
-  David S.~Hippocampus\thanks{Use footnote for providing further information
-    about author (webpage, alternative address)---\emph{not} for acknowledging
-    funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  % examples of more authors
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \AND
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-}
-
-\begin{document}
-
-\maketitle
-
-\begin{abstract}
-	Causal discovery is a fundamental problem in causal inference, and
-	numerous algorithms have been developed to address it. However, their
-	adoption in applied disciplines  has been limited, possibly due to a
-	lack of confidence in these algorithms stemming from them making
-	obvious mistakes, and the difficulties in choosing the right algorithm
-	for specific datasets. Consequently, in applied disciplines, these
-	causal Directed Acyclic Graphs (DAGs) are predominantly constructed
-	manually based on domain expertise. This makes it vital to test and
-	potentially modify these models against data before using them for any
-	analysis. Local testing based on Conditional Independence tests and
-	their effect sizes have been commonly used for this purpose. However,
-	one of the current limitations of this approach is that there are no
-	effect sizes for mixed data. To address this, we introduce a novel
-	effect size measure for mixed data based on canonical correlations.
-	This effect size is a generalization of some of the effect sizes used
-	in special cases to mixed variables. Moreover, to assist researchers in
-	constructing DAGs manually, we use the proposed mixed data effect size
-	measure to build a web-based tool that can help researchers in
-	constructing, validating, and modifying their DAGs with the help of
-	available data. The tool iteratively suggests modifications to DAGs
-	based on correlations that are not explained/over-explained by the DAG,
-	while the expert chooses the direction of the edge between those
-	variables. Empirically, we show that taking a greedy approach on the
-	tool's recommendations can perform comparably to causal discovery
-	algorithms, provided that the expert correctly determines the direction
-	of the edges in at least one out of three cases, with performance
-	improving as accuracy increases.
-\end{abstract}
-
-\section{Introduction}
-Para 1: Many Causal discovery algorithms but commonly constructed manually.
-
-To perform any causal effect identification or estimation in Directed Acyclic
-Graph(DAG) based causal inference usually requires to come up with the DAG
-structure. Learning these DAG structures from data is known as \emph{causal
-discovery}. Plenty of algorithms have been developed for these in the
-literature such as constraint-based methods such as PC, FCI, score based
-methods such as Hill-Climb Search, GES, and constrained continuous optimization
-based such as No TEARS.
-
-Para 2: Reasons behind this
-
-Even with the existence of all these algorithms for constructing DAGs from
-data, in applied fields researchers prefer to construct DAGs by hand. We
-hypothesize that the adaption of these algorithms in applied literature is
-low due to the following reasons:
-\begin{enumerate}
-	\item Algorithms can make very obvious mistakes at times making it harder
-		for the user to trust their outputs.
-	\item Difficult to choose a single algorithm and their hyper-parameters.
-\end{enumerate}
-
-Para 3: Importance of testing
-
-When these models are constructed manually, it is extremely important that they
-are tested before they are used for making any inference. Local testing methods
-have been used for this purpose, where the implied conditional independences of
-the model are testing in the data using statistical tests. These tests can then
-be used to modify our models.
-
-Para 4: Mixed data testing
-
-Para 5: Contributions
-
-\section{Background}
-\subsection{Continuous Variables}
-\subsection{Discrete Variables}
-\subsection{Ordinal Variables}
-\subsection{Ordinal and Continuous Variables}
-
-\section{Effect size for mixed data}
-
-\section{Web Tool}
-As most of the DAGs in practice are built manually, we use our mixed data effect size measure to build a web-tool that aids researches in building DAGs.
-
-\section{Empirical Analysis}
-We performed empirical analysis to compare how well does creating a graph manually based on effect sizes and p-value compare to algorithmic approaches.
-\subsection{Simulating an expert}
-To simulate an expert building a model manually, we wrote a simple greedy simulation as shown in the Figure. The expert always starts with a graph with
-no edges and chooses the pair of variables with the highest correlation between them. The expert based on some accuracy measure decides the direction
-of the edge between these variables such that the new added edge does not create a cycle in the DAG:
-
-\todo{Summarize this section and equaiton in an algorithm}
-\begin{equation}
-	\begin{split}
-		x &= \textnormal{rand}([0, 1]) \\
-		O(\alpha) &= \begin{cases} 
-			M \rightarrow Y, & \textnormal{if  } x <= \alpha \\
-			\textnormal{rand}(M \rightarrow Y, M \leftarrow Y, None) & \textnormal{otherwise} \\
-			     \end{cases} \\
-	\end{split}
-\end{equation}
-
-After adding each edge, we also check if the p-value of any existing edge shows independence. If that happens we remove that edge and blacklist that edge. 
-Blacklisted edges are not suggested to the expert in future iterations. We repeat this till we have explained all the correlations in the dataset.
-
-This is a possibility that this greedy expert gets stuck at incorrect graph structures as it does not do any backtracking to fix its earlier mistakes.
-
-\subsection{Results}
-
-\section*{References}
-\bibliographystyle{plain}
-\bibliography{references.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix / supplemental material}
-
-Optionally include supplemental material (complete proofs, additional experiments and plots) in appendix.
-All such materials \textbf{SHOULD be included in the main submission.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newpage
 \section*{NeurIPS Paper Checklist}
@@
 \end{enumerate}
-
-\end{document}

### 2024-05-06 paper/main.tex
Separate out checklist file

@@
 
+One of the limitations here is there are no effect size measures available for
+
 Para 5: Contributions
@@
 
-\section*{References}
+\section{Conclusions}
+\subsection{Using LLMs as experts}
+
 \bibliographystyle{plain}
@@
 
-\newpage
-\section*{NeurIPS Paper Checklist}
-
-%%% BEGIN INSTRUCTIONS %%%
-The checklist is designed to encourage best practices for responsible machine learning research, addressing issues of reproducibility, transparency, research ethics, and societal impact.
-Do not remove the checklist: {\bf The papers not including the checklist will be desk rejected.} The checklist should follow the references and precede the (optional) supplemental material.
-The checklist does NOT count towards the page
-limit.
-
-Please read the checklist guidelines carefully for information on how to answer these questions. For each question in the checklist:
-\begin{itemize}
-    \item You should answer \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item \answerNA{} means either that the question is Not Applicable for that particular paper or the relevant information is Not Available.
-    \item Please provide a short (1–2 sentence) justification right after your answer (even for NA). 
-   % \item {\bf The papers not including the checklist will be desk rejected.}
-\end{itemize}
-
-{\bf The checklist answers are an integral part of your paper submission.} They are visible to the reviewers, area chairs, senior area chairs, and ethics reviewers.
-You will be asked to also include it (after eventual revisions) with the final version of your paper, and its final version will be published with the paper.
-
-The reviewers of your paper will be asked to use the checklist as one of the factors in their evaluation.
-While "\answerYes{}" is generally preferable to "\answerNo{}", it is perfectly acceptable to answer "\answerNo{}" provided a proper justification is given (e.g., "error bars are not reported because it would be too computationally expensive" or "we were unable to find the license for the dataset we used").
-In general, answering "\answerNo{}" or "\answerNA{}" is not grounds for rejection.
-While the questions are phrased in a binary way, we acknowledge that the true answer is often more nuanced, so please just use your best judgment and write a justification to elaborate.
-All supporting evidence can appear either in the main paper or the supplemental material, provided in appendix.
-If you answer \answerYes{} to a question, in the justification please point to the section(s) where related material for the question can be found.
-
-IMPORTANT, please:
-\begin{itemize}
-    \item {\bf Delete this instruction block, but keep the section heading ``NeurIPS paper checklist"},
-    \item  {\bf Keep the checklist subsection headings, questions/answers and guidelines below.}
-    \item {\bf Do not modify the questions and only use the provided macros for your answers}.
-\end{itemize}
-
-%%% END INSTRUCTIONS %%%
-
-\begin{enumerate}
-
-\item {\bf Claims}
-    \item[] Question: Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the abstract and introduction do not include the claims made in the paper.
-        \item The abstract and/or introduction should clearly state the claims made, including the contributions made in the paper and important assumptions and limitations.
-A No or NA answer to this question will not be perceived well by the reviewers.
-        \item The claims made should match theoretical and experimental results, and reflect how much the results can be expected to generalize to other settings. 
-        \item It is fine to include aspirational goals as motivation as long as it is clear that these goals are not attained by the paper. 
-    \end{itemize}
-
-\item {\bf Limitations}
-    \item[] Question: Does the paper discuss the limitations of the work performed by the authors?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper has no limitation while the answer No means that the paper has limitations, but those are not discussed in the paper. 
-        \item The authors are encouraged to create a separate "Limitations" section in their paper.
-        \item The paper should point out any strong assumptions and how robust the results are to violations of these assumptions (e.g., independence assumptions, noiseless settings, model well-specification, asymptotic approximations only holding locally).
-The authors should reflect on how these assumptions might be violated in practice and what the implications would be.
-        \item The authors should reflect on the scope of the claims made, e.g., if the approach was only tested on a few datasets or with a few runs.
-In general, empirical results often depend on implicit assumptions, which should be articulated.
-        \item The authors should reflect on the factors that influence the performance of the approach.
-For example, a facial recognition algorithm may perform poorly when image resolution is low or images are taken in low lighting.
-Or a speech-to-text system might not be used reliably to provide closed captions for online lectures because it fails to handle technical jargon.
-        \item The authors should discuss the computational efficiency of the proposed algorithms and how they scale with dataset size.
-        \item If applicable, the authors should discuss possible limitations of their approach to address problems of privacy and fairness.
-        \item While the authors might fear that complete honesty about limitations might be used by reviewers as grounds for rejection, a worse outcome might be that reviewers discover limitations that aren't acknowledged in the paper.
-The authors should use their best judgment and recognize that individual actions in favor of transparency play an important role in developing norms that preserve the integrity of the community.
-Reviewers will be specifically instructed to not penalize honesty concerning limitations.
-    \end{itemize}
-
-\item {\bf Theory Assumptions and Proofs}
-    \item[] Question: For each theoretical result, does the paper provide the full set of assumptions and a complete (and correct) proof?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include theoretical results. 
-        \item All the theorems, formulas, and proofs in the paper should be numbered and cross-referenced.
-        \item All assumptions should be clearly stated or referenced in the statement of any theorems.
-        \item The proofs can either appear in the main paper or the supplemental material, but if they appear in the supplemental material, the authors are encouraged to provide a short proof sketch to provide intuition.
-        \item Inversely, any informal proof provided in the core of the paper should be complemented by formal proofs provided in appendix or supplemental material.
-        \item Theorems and Lemmas that the proof relies upon should be properly referenced. 
-    \end{itemize}
-
-    \item {\bf Experimental Result Reproducibility}
-    \item[] Question: Does the paper fully disclose all the information needed to reproduce the main experimental results of the paper to the extent that it affects the main claims and/or conclusions of the paper (regardless of whether the code and data are provided or not)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include experiments.
-        \item If the paper includes experiments, a No answer to this question will not be perceived well by the reviewers: Making the paper reproducible is important, regardless of whether the code and data are provided or not.
-        \item If the contribution is a dataset and/or model, the authors should describe the steps taken to make their results reproducible or verifiable. 
-        \item Depending on the contribution, reproducibility can be accomplished in various ways.
-For example, if the contribution is a novel architecture, describing the architecture fully might suffice, or if the contribution is a specific model and empirical evaluation, it may be necessary to either make it possible for others to replicate the model with the same dataset, or provide access to the model.
-In general.
-releasing code and data is often one good way to accomplish this, but reproducibility can also be provided via detailed instructions for how to replicate the results, access to a hosted model (e.g., in the case of a large language model), releasing of a model checkpoint, or other means that are appropriate to the research performed.
-        \item While NeurIPS does not require releasing code, the conference does require all submissions to provide some reasonable avenue for reproducibility, which may depend on the nature of the contribution.
-For example
-        \begin{enumerate}
-            \item If the contribution is primarily a new algorithm, the paper should make it clear how to reproduce that algorithm.
-            \item If the contribution is primarily a new model architecture, the paper should describe the architecture clearly and fully.
-            \item If the contribution is a new model (e.g., a large language model), then there should either be a way to access this model for reproducing the results or a way to reproduce the model (e.g., with an open-source dataset or instructions for how to construct the dataset).
-            \item We recognize that reproducibility may be tricky in some cases, in which case authors are welcome to describe the particular way they provide for reproducibility.
-In the case of closed-source models, it may be that access to the model is limited in some way (e.g., to registered users), but it should be possible for other researchers to have some path to reproducing or verifying the results.
-        \end{enumerate}
-    \end{itemize}
-
-\item {\bf Open access to data and code}
-    \item[] Question: Does the paper provide open access to the data and code, with sufficient instructions to faithfully reproduce the main experimental results, as described in supplemental material?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that paper does not include experiments requiring code.
-        \item Please see the NeurIPS code and data submission guidelines (\url{https://nips.cc/public/guides/CodeSubmissionPolicy}) for more details.
-        \item While we encourage the release of code and data, we understand that this might not be possible, so “No” is an acceptable answer.
-Papers cannot be rejected simply for not including code, unless this is central to the contribution (e.g., for a new open-source benchmark).
-        \item The instructions should contain the exact command and environment needed to run to reproduce the results.
-See the NeurIPS code and data submission guidelines (\url{https://nips.cc/public/guides/CodeSubmissionPolicy}) for more details.
-        \item The authors should provide instructions on data access and preparation, including how to access the raw data, preprocessed data, intermediate data, and generated data, etc.
-        \item The authors should provide scripts to reproduce all experimental results for the new proposed method and baselines.
-If only a subset of experiments are reproducible, they should state which ones are omitted from the script and why.
-        \item At submission time, to preserve anonymity, the authors should release anonymized versions (if applicable).
-        \item Providing as much information as possible in supplemental material (appended to the paper) is recommended, but including URLs to data and code is permitted.
-    \end{itemize}
-
-\item {\bf Experimental Setting/Details}
-    \item[] Question: Does the paper specify all the training and test details (e.g., data splits, hyperparameters, how they were chosen, type of optimizer, etc.) necessary to understand the results?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include experiments.
-        \item The experimental setting should be presented in the core of the paper to a level of detail that is necessary to appreciate the results and make sense of them.
-        \item The full details can be provided either with the code, in appendix, or as supplemental material.
-    \end{itemize}
-
-\item {\bf Experiment Statistical Significance}
-    \item[] Question: Does the paper report error bars suitably and correctly defined or other appropriate information about the statistical significance of the experiments?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include experiments.
-        \item The authors should answer "Yes" if the results are accompanied by error bars, confidence intervals, or statistical significance tests, at least for the experiments that support the main claims of the paper.
-        \item The factors of variability that the error bars are capturing should be clearly stated (for example, train/test split, initialization, random drawing of some parameter, or overall run with given experimental conditions).
-        \item The method for calculating the error bars should be explained (closed form formula, call to a library function, bootstrap, etc.)
-        \item The assumptions made should be given (e.g., Normally distributed errors).
-        \item It should be clear whether the error bar is the standard deviation or the standard error of the mean.
-        \item It is OK to report 1-sigma error bars, but one should state it.
-The authors should preferably report a 2-sigma error bar than state that they have a 96\% CI, if the hypothesis of Normality of errors is not verified.
-        \item For asymmetric distributions, the authors should be careful not to show in tables or figures symmetric error bars that would yield results that are out of range (e.g.
-negative error rates).
-        \item If error bars are reported in tables or plots, The authors should explain in the text how they were calculated and reference the corresponding figures or tables in the text.
-    \end{itemize}
-
-\item {\bf Experiments Compute Resources}
-    \item[] Question: For each experiment, does the paper provide sufficient information on the computer resources (type of compute workers, memory, time of execution) needed to reproduce the experiments?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not include experiments.
-        \item The paper should indicate the type of compute workers CPU or GPU, internal cluster, or cloud provider, including relevant memory and storage.
-        \item The paper should provide the amount of compute required for each of the individual experimental runs as well as estimate the total compute. 
-        \item The paper should disclose whether the full research project required more compute than the experiments reported in the paper (e.g., preliminary or failed experiments that didn't make it into the paper).
-    \end{itemize}
-
-\item {\bf Code Of Ethics}
-    \item[] Question: Does the research conducted in the paper conform, in every respect, with the NeurIPS Code of Ethics \url{https://neurips.cc/public/EthicsGuidelines}?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the authors have not reviewed the NeurIPS Code of Ethics.
-        \item If the authors answer No, they should explain the special circumstances that require a deviation from the Code of Ethics.
-        \item The authors should make sure to preserve anonymity (e.g., if there is a special consideration due to laws or regulations in their jurisdiction).
-    \end{itemize}
-
-\item {\bf Broader Impacts}
-    \item[] Question: Does the paper discuss both potential positive societal impacts and negative societal impacts of the work performed?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that there is no societal impact of the work performed.
-        \item If the authors answer NA or No, they should explain why their work has no societal impact or why the paper does not address societal impact.
-        \item Examples of negative societal impacts include potential malicious or unintended uses (e.g., disinformation, generating fake profiles, surveillance), fairness considerations (e.g., deployment of technologies that could make decisions that unfairly impact specific groups), privacy considerations, and security considerations.
-        \item The conference expects that many papers will be foundational research and not tied to particular applications, let alone deployments.
-However, if there is a direct path to any negative applications, the authors should point it out.
-For example, it is legitimate to point out that an improvement in the quality of generative models could be used to generate deepfakes for disinformation.
-On the other hand, it is not needed to point out that a generic algorithm for optimizing neural networks could enable people to train models that generate Deepfakes faster.
-        \item The authors should consider possible harms that could arise when the technology is being used as intended and functioning correctly, harms that could arise when the technology is being used as intended but gives incorrect results, and harms following from (intentional or unintentional) misuse of the technology.
-        \item If there are negative societal impacts, the authors could also discuss possible mitigation strategies (e.g., gated release of models, providing defenses in addition to attacks, mechanisms for monitoring misuse, mechanisms to monitor how a system learns from feedback over time, improving the efficiency and accessibility of ML).
-    \end{itemize}
-
-\item {\bf Safeguards}
-    \item[] Question: Does the paper describe safeguards that have been put in place for responsible release of data or models that have a high risk for misuse (e.g., pretrained language models, image generators, or scraped datasets)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper poses no such risks.
-        \item Released models that have a high risk for misuse or dual-use should be released with necessary safeguards to allow for controlled use of the model, for example by requiring that users adhere to usage guidelines or restrictions to access the model or implementing safety filters.
-        \item Datasets that have been scraped from the Internet could pose safety risks. The authors should describe how they avoided releasing unsafe images.
-        \item We recognize that providing effective safeguards is challenging, and many papers do not require this, but we encourage authors to take this into account and make a best faith effort.
-    \end{itemize}
-
-\item {\bf Licenses for existing assets}
-    \item[] Question: Are the creators or original owners of assets (e.g., code, data, models), used in the paper, properly credited and are the license and terms of use explicitly mentioned and properly respected?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not use existing assets.
-        \item The authors should cite the original paper that produced the code package or dataset.
-        \item The authors should state which version of the asset is used and, if possible, include a URL.
-        \item The name of the license (e.g., CC-BY 4.0) should be included for each asset.
-        \item For scraped data from a particular source (e.g., website), the copyright and terms of service of that source should be provided.
-        \item If assets are released, the license, copyright information, and terms of use in the package should be provided.
-For popular datasets, \url{paperswithcode.com/datasets} has curated licenses for some datasets.
-Their licensing guide can help determine the license of a dataset.
-        \item For existing datasets that are re-packaged, both the original license and the license of the derived asset (if it has changed) should be provided.
-        \item If this information is not available online, the authors are encouraged to reach out to the asset's creators.
-    \end{itemize}
-
-\item {\bf New Assets}
-    \item[] Question: Are new assets introduced in the paper well documented and is the documentation provided alongside the assets?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not release new assets.
-        \item Researchers should communicate the details of the dataset/code/model as part of their submissions via structured templates.
-This includes details about training, license, limitations, etc.
-        \item The paper should discuss whether and how consent was obtained from people whose asset is used.
-        \item At submission time, remember to anonymize your assets (if applicable). You can either create an anonymized URL or include an anonymized zip file.
-    \end{itemize}
-
-\item {\bf Crowdsourcing and Research with Human Subjects}
-    \item[] Question: For crowdsourcing experiments and research with human subjects, does the paper include the full text of instructions given to participants and screenshots, if applicable, as well as details about compensation (if any)?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not involve crowdsourcing nor research with human subjects.
-        \item Including this information in the supplemental material is fine, but if the main contribution of the paper involves human subjects, then as much detail as possible should be included in the main paper.
-        \item According to the NeurIPS Code of Ethics, workers involved in data collection, curation, or other labor should be paid at least the minimum wage in the country of the data collector. 
-    \end{itemize}
-
-\item {\bf Institutional Review Board (IRB) Approvals or Equivalent for Research with Human Subjects}
-    \item[] Question: Does the paper describe potential risks incurred by study participants, whether such risks were disclosed to the subjects, and whether Institutional Review Board (IRB) approvals (or an equivalent approval/review based on the requirements of your country or institution) were obtained?
-    \item[] Answer: \answerTODO{} % Replace by \answerYes{}, \answerNo{}, or \answerNA{}.
-    \item[] Justification: \justificationTODO{}
-    \item[] Guidelines:
-    \begin{itemize}
-        \item The answer NA means that the paper does not involve crowdsourcing nor research with human subjects.
-        \item Depending on the country in which research is conducted, IRB approval (or equivalent) may be required for any human subjects research.
-If you obtained IRB approval, you should clearly state this in the paper.
-        \item We recognize that the procedures for this may vary significantly between institutions and locations, and we expect authors to adhere to the NeurIPS Code of Ethics and the guidelines for their institution.
-        \item For initial submissions, do not include any information that would break anonymity (if applicable), such as the institution conducting the review.
-    \end{itemize}
-
-\end{enumerate}
+\include{checklist}
 

### 2024-05-06 paper/main.tex
updates

@@
 \section{Introduction}
-Para 1: Many Causal discovery algorithms but commonly constructed manually.
+\textbf{Para 1: Many Causal discovery algorithms but commonly constructed manually.}
 
-To perform any causal effect identification or estimation in Directed Acyclic
-Graph(DAG) based causal inference usually requires to come up with the DAG
-structure. Learning these DAG structures from data is known as \emph{causal
-discovery}. Plenty of algorithms have been developed for these in the
-literature such as constraint-based methods such as PC, FCI, score based
+To perform any causal effect identification or estimation in Pearl's Directed
+Acyclic Graphs (DAGs) based causal inference usually requires to come up with
+the DAG structure. Learning these DAG structures from data is known as
+\emph{causal discovery}. Plenty of algorithms have been developed for these in
+the literature such as constraint-based methods such as PC, FCI, score based
 methods such as Hill-Climb Search, GES, and constrained continuous optimization
@@
 
-Para 2: Reasons behind this
+\textbf{Para 2: Reasons behind this}
 
@@
 
-Para 3: Importance of testing
+\textbf{Para 3: Importance of testing, include modification indices}
 
@@
 
-Para 4: Mixed data testing
+\textbf{Para 4: Mixed data testing}
 
-One of the limitations here is there are no effect size measures available for
+One of the limitations here is there are no effect size measures available for mixed data.
 
-Para 5: Contributions
+\textbf{Para 5: Contributions}
+\begin{enumerate}
+	\item An effect size measure for mixed data. 
+	\item A web-based tool to help with both model testing and construction. Combining both domain knowledge and data.
+\end{enumerate}
 
 \section{Background}
+\todo[inline]{Add notation used here}
+
 \subsection{Continuous Variables}
+Correlation coefficient and partial correlation coefficient.
+
 \subsection{Discrete Variables}
+Chi-Squared and Cram\'er's V
+
+\subsection{Ordinal and Continuous Variables}
+Polyserial Correlation
+
 \subsection{Ordinal Variables}
-\subsection{Ordinal and Continuous Variables}
+Polychoric Correlation
+
+\subsection{Mixed data}
+Background on CI tests for mixed data; but no effect size measure.
 

### 2024-05-06 paper/main.tex
Updates

@@
 
+\todo[inline]{Add citations}
+
 To perform any causal effect identification or estimation in Pearl's Directed
-Acyclic Graphs (DAGs) based causal inference usually requires to come up with
+Acyclic Graphs (DAGs) based causal framework usually requires to come up with
 the DAG structure. Learning these DAG structures from data is known as
@@
 methods such as Hill-Climb Search, GES, and constrained continuous optimization
-based such as No TEARS.
+based such as No TEARS. All these algorithms take different approaches to
+causal discovery, constraint-based algorithms utilize statistical conditional
+independence (CI) tests to come up with a model structure, the score-based
+methods iteratively perform local modifications to the model such that some
+score function is maximized, the continuous optimization methods reformulate
+the problem statement as a constrained optimization problem and use
+optimization algorithms to come up with a graph.
 
@@
 
-Even with the existence of all these algorithms for constructing DAGs from
-data, in applied fields researchers prefer to construct DAGs by hand. We
-hypothesize that the adaption of these algorithms in applied literature is
-low due to the following reasons:
+Even though we have various algorithms for doing casual discovery, yet
+researchers in applied fields prefer to construct these DAGs manually based on
+their domain knowledge. We hypothesize the following two reasons for low
+adaption of these methods in applied sciences:
+
 \begin{enumerate}
-	\item Algorithms can make very obvious mistakes at times making it harder
-		for the user to trust their outputs.
-	\item Difficult to choose a single algorithm and their hyper-parameters.
+	\item Algorithms making obvious mistakes: These algorithms often 
+		make very obvious mistakes, making it harder for the user
+		to trust them.
+	\item Difficulties in choosing an algorithm: For any given dataset, the
+		user need to make a choice of the algorithm. Given that these
+		algorithms work very differently, it can be difficult to make a
+		choice. Additionally, each of these algorithms have
+		hyperparameters and choices to make for them, making it
+		difficult. As there is no way to evaluate the performance of 
+		these algorithms on our dataset.
 \end{enumerate}
@@
 
-When these models are constructed manually, it is extremely important that they
-are tested before they are used for making any inference. Local testing methods
-have been used for this purpose, where the implied conditional independences of
-the model are testing in the data using statistical tests. These tests can then
-be used to modify our models.
+Both in the cases when the DAG is an output of an algorithm or when constructed
+by hand, it is vital to test these models against data before using them for
+making any inference. In the field of Structural Equation Models (SEMs)
+modification indices have been used to do this. However, as modification index
+is an evaluation of the global fit of the model. An alternative is local fit of
+models using CI tests where the implied conditional independences of the model
+can be tested using statistical tests. Similar to modification indices, these
+tests can then be used to modify our models. Figure shows an example of this
+testing.
 

### 2024-05-06 paper/main.tex
Updates

@@
 tests can then be used to modify our models. Figure shows an example of this
-testing.
+testing. We can not also only rely on the p-values, effect size can help in
+deciding.
 
@@
 
-One of the limitations here is there are no effect size measures available for mixed data.
+As these CI based testing uses p-values and effect size measures from CI tests
+for testing. And as mixed data is common, we need both a test and effect size
+measure for mixed data. Recently, a few CI tests for mixed data have been
+proposed but an effect size is still missing.
 
@@
 
-\subsection{Continuous Variables}
-Correlation coefficient and partial correlation coefficient.
+We need to do both conditional and non-conditional tests. We would assume that
+we are testing the statement $ X \ci Y \rvert \bm{Z} $, where $ \bm{Z} $ is a
+set of conditional variables possibly null.
 
-\subsection{Discrete Variables}
+\subsection{Effect Size Measures}
+
+\subsubsection{Continuous Variables}
+In the case when $ \bm{Z} = \emptyset $, and when both $ X $ and $ Y $ are
+continuous variables, Pearson's correlation coefficient based test can be used
+which also gives an effect size measure. The p-value computation requires the
+data to be normally distributed but other statistics can be used such as
+Spearman's rank correlation, or Kendall's Tau.
+
+In the case when $ \bm{Z} \neq \emptyset $, then partial correlations can be used.
+In this case we first compute the residuals of $ X $ and $ Y $ using $ \bm{Z} $ as 
+the predictor. We start by training two model $ X \sim \bm{Z} $ and $ Y \sim \bm{Z} $.
+Then the residuals are computed for both as $ R_X = X - \hat{X} $ and $ R_Y = Y - \hat{Y} $.
+After this we can do a normal non-conditional test on $ R_X $ and $ R_Y $.
+
+\subsubsection{Discrete Variables}
+
 Chi-Squared and Cram\'er's V
 
-\subsection{Ordinal and Continuous Variables}
+\subsubsection{Ordinal and Continuous Variables}
 Polyserial Correlation
 
-\subsection{Ordinal Variables}
+\subsubsection{Ordinal Variables}
 Polychoric Correlation
 
-\subsection{Mixed data}
-Background on CI tests for mixed data; but no effect size measure.
+\subsubsection{Mixed data}
+Recently multiple tests for mixed data have been proposed but they do not have
+an effect size measure.
+
+\subsection{Canonical Correlations}
 
 \section{Effect size for mixed data}
+\subsection{Non-conditional Case}
+\subsection{Conditional Case}
 

### 2024-05-07 paper/main.tex
Updates

@@
 \usepackage{todonotes}
+\usepackage{bm}
+\usepackage{subcaption}
+
+\def\ci{\perp\!\!\!\!\!\perp}
 
@@
 
+\begin{figure}
+		\centering
+		\begin{tikzpicture}
+			\begin{scope}[yshift=0.7cm, scale=0.8]
+			\tikzstyle{every node}=[align=center, inner sep=1pt]
+				\node (sex) at (-0.7, -0.8) {Sex};
+				\node (age) at (-0.3, 0) {Age};
+				\node (ed) at (-1.8, -1.6) {Education};
+				\node (occ) at (1.8, 0) {Occupation};
+				\node (hrpw) at (1.3, -1.6) {Hours Per \\ Week};
+				\node (income) at (-0.7, -2.8) {Income};
+
+				\draw[->]  (age) to[bend right=20] (ed);
+				\draw[->]  (sex) to (ed);
+				\draw[->]  (age) to (hrpw);
+				\draw[->]  (ed) to (hrpw);
+				\draw[->]  (sex) to (hrpw);
+				\draw[->]  (ed) to (income);
+				\draw[->]  (hrpw) to (income);
+				\draw[->]  (occ) to[out=300, in=0, looseness=1.4] (income.east);
+				\draw[->]  (sex) to (income);
+				\draw[->]  (ed) to[out=90, in=150, looseness=1.3] (occ);
+				\draw[->]  (sex) to (occ);	
+			\end{scope}
+			\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
+			\node[rectangle, align=center, inner sep=1pt] at (6, 0) {
+				\begin{minipage}{\textwidth}
+					\begin{equation*}
+						\begin{split}
+							\textnormal{Age} &\ci \textnormal{Incm} \rvert \textnormal{Edct, Hrpw, Sex} \\
+							\textnormal{Age} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
+							\textnormal{Age} &\ci \textnormal{Sex} \\
+							\textnormal{Hrpw} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
+						\end{split}
+					\end{equation*}
+				\end{minipage}
+				};
+		\end{tikzpicture}
+		\caption{\todo[inline]{Show a full example of testing using this model}}
+\end{figure}
+
 \section{Background}
@@
 
-Chi-Squared and Cram\'er's V
+When $ \bm{Z} = \emptyset $, in the case of discrete variables, chi-squared test can be used. Effect size 
+measures such as Cram\'er's V are available for it.
+
+When $ \bm{Z} \neq \emptyset $, we do stratification.
 
@@
 Polyserial Correlation
+
+When $ \bm{Z} = \emptyset $, ordinal variable can be assumed to be coming from a thresholded normal distribution, estimate the thresholds and latent variable to compute 
+the correlation coefficient.
 
@@
 Recently multiple tests for mixed data have been proposed but they do not have
-an effect size measure.
+an effect size measure. We propose one based on canonical correlations.
 
 \subsection{Canonical Correlations}
+Canonical correlations are generalization of correlations to set of variables.
 
@@
 \subsection{Conditional Case}
+\subsection{Connection to special cases}
 
@@
 As most of the DAGs in practice are built manually, we use our mixed data effect size measure to build a web-tool that aids researches in building DAGs.
+
+\begin{figure}
+	\centering
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.25]{../../presentations/2024_05_das/1.png}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.25]{../../presentations/2024_05_das/2.png}
+	\end{subfigure}
+	\caption{\todo[inline]{Insert screenshots of the web-tool}}
+\end{figure}
 
@@
 
-This is a possibility that this greedy expert gets stuck at incorrect graph structures as it does not do any backtracking to fix its earlier mistakes.
+There is a possibility that this greedy expert gets stuck at incorrect graph structures as it does not do any backtracking to fix its earlier mistakes.
 
 \subsection{Results}
+\begin{figure}
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics{../code/plots/shd_ribbon.pdf}
+	\end{subfigure}%
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics{../code/plots/sid_ribbon.pdf}
+	\end{subfigure}
+\end{figure}
 

### 2024-05-07 code/make_plots.R

@@ -98,7 +98,7 @@ make_plots <- function(filename, plot_type){
 			labs(color = "Algorithm") +
 			ylim(0, 40)
 			
-		ggsave("plots/shd_ribbon.pdf", p_shd, height=3.1, width=4.8, units='in')
+		ggsave("plots/shd_ribbon.pdf", p_shd, height=2.5, width=2.6, units='in')
 
 		p_sid <- ggplot(d_human, aes(x=edge_prob, y=sid_mean, group=alg, color=color)) +
 			guides(color='none') +
@@ -115,7 +115,7 @@ make_plots <- function(filename, plot_type){
 			labs(color = "Algorithm") +
 			ylim(0, 100)
 			
-		ggsave("plots/sid_ribbon.pdf", p_sid, height=3.1, width=4.8, units='in')
+		ggsave("plots/sid_ribbon.pdf", p_sid, height=2.5, width=2.6, units='in')
 	}
 }
 

### 2024-05-07 paper/main.tex
Updates

@@
 \section{Empirical Analysis}
-We performed empirical analysis to compare how well does creating a graph manually based on effect sizes and p-value compare to algorithmic approaches.
+To compare how well does creating a graph manually based on effect sizes and
+p-value compare to algorithmic approaches, we performed some emperical
+analysis.
+
 \subsection{Simulating an expert}
@@
 no edges and chooses the pair of variables with the highest correlation between them. The expert based on some accuracy measure decides the direction
-of the edge between these variables such that the new added edge does not create a cycle in the DAG:
+of the edge between these variables such that the new added edge does not create a cycle in the DAG.
 
-\todo{Summarize this section and equaiton in an algorithm}
+\todo[inline]{Summarize this section and equaiton in an algorithm}
+\todo[inline]{Show a consistency proof for this expert simulation}
+
 \begin{equation}

### 2024-05-07 paper/main.tex
Updates

@@
 \begin{abstract}
-	Causal discovery is a fundamental problem in causal inference, and
-	numerous algorithms have been developed to address it. However, their
+	Constructing a Directed Acyclic Graph (DAG) from observational data,
+	called causal discovery, is a fundamental problem in causal inference.
+	Numerous algorithms have been developed to address it. However, their
 	adoption in applied disciplines has been limited, possibly due to a
@@
 	potentially modify these models against data before using them for any
-	analysis. Local testing based on Conditional Independence tests and
-	their effect sizes have been commonly used for this purpose. However,
+	analysis. Local testing based on Conditional Independence (CI) tests and
+	effect sizes have been commonly used for this purpose. However,
 	one of the current limitations of this approach is that there are no
@@
 \begin{enumerate}
+	\item Algorithsm can be biased towards sparse graphs. Ref: Malinsky's paper.
 	\item Algorithms making obvious mistakes: These algorithms often 
@@
 \subsection{Canonical Correlations}
-Canonical correlations are generalization of correlations to set of variables.
+Canonical correlations are generalization of correlations to multi-dimensional case.
+
+Multiple effect sizes measures based on canonical correations are available.
 
 \section{Effect size for mixed data}
-\subsection{Non-conditional Case}
+In this section, we use canonical correlations to define an effect size for mixed data CI tests. The main idea is to transform any CI test into j
+We consider the following two cases separately:
+
+\subsection{No conditional variables}
+When we have no conditonal variables, we are trying
+
 \subsection{Conditional Case}
+
 \subsection{Connection to special cases}

### 2024-05-07 paper/main.tex
Updates

@@
 \begin{enumerate}
-	\item Algorithsm can be biased towards sparse graphs. Ref: Malinsky's paper.
+	\item Biased towards sparse graphs.
 	\item Algorithms making obvious mistakes: These algorithms often 
@@
 		\begin{tikzpicture}
-			\begin{scope}[yshift=0.7cm, scale=0.8]
+			\begin{scope}[yshift=0.7cm, scale=1]
 			\tikzstyle{every node}=[align=center, inner sep=1pt]
@@
 \section{Background}
-\todo[inline]{Add notation used here}
+We consider one-dimensional variables $ X $, and $ Y $ and a possibly
+multi-dimensional co$ Z = Z_1, \cdots, Z_k $, potentially being an empty set $
+Z = \emptyset $. We consider various scenarios where the variables $ X $, $ Y $
+and $ Z $ could be either discrete, ordinal, or continuous. With mixed data, we
+mean an arbitrary mix of variable types among $ X $, $ Y $, and $ Z $.
 
-We need to do both conditional and non-conditional tests. We would assume that
-we are testing the statement $ X \ci Y \rvert \bm{Z} $, where $ \bm{Z} $ is a
-set of conditional variables possibly null.
+We consider the problem of estimating the partial (or conditional) association
+between $ X $ and $ Y $ in the presence of the conditioning set $ Z $. When $ Z
+= \emptyset $, this is equivalent to the marginal association between $ X $ and
+$ Y $.
+
+\todo[inline]{Add more notation as they come}
 
@@
 
-\subsubsection{Continuous Variables}
+\subsubsection{$ X $ and $ Y $ are continuous}
 In the case when $ \bm{Z} = \emptyset $, and when both $ X $ and $ Y $ are
@@
 
-\subsubsection{Discrete Variables}
+\subsubsection{All $ X $, $ Y $, $ Z $ are discrete}
 
@@
 
-When $ \bm{Z} \neq \emptyset $, we do stratification.
+When $ \bm{Z} \neq \emptyset $, we can do stratification.
 
-\subsubsection{Ordinal and Continuous Variables}
+\subsubsection{$ X $ is ordinal and $ Y $ is continuous}
 Polyserial Correlation
@@
 
-\subsubsection{Ordinal Variables}
+\subsubsection{$ X $ and $ Y $ are ordinal}
 Polychoric Correlation

### 2024-05-07 paper/main.tex
Updates

@@
 \usepackage{subcaption}
+\usepackage[ruled]{algorithm2e}
 
 \def\ci{\perp\!\!\!\!\!\perp}
+
+\newtheorem{definition}{Definition}
+\newtheorem{proposition}{Proposition}
 
@@
 \begin{enumerate}
-	\item An effect size measure for mixed data. 
-	\item A web-based tool to help with both model testing and construction. Combining both domain knowledge and data.
+	\item We present an effect size measure based on canonical correlations
+		for mixed data that generalizes some of the commonly used
+		effect size in special cases such as partial correlation
+		coefficient for continuous variables, and Cram\'er's V for
+		discrete variables.
+	\item We also developed a web-based tool to help researches in
+		constructing and testing and modifying their DAGs using CI tests and
+		effect size measures.
 \end{enumerate}
@@
 \section{Background}
-We consider one-dimensional variables $ X $, and $ Y $ and a possibly
-multi-dimensional co$ Z = Z_1, \cdots, Z_k $, potentially being an empty set $
-Z = \emptyset $. We consider various scenarios where the variables $ X $, $ Y $
-and $ Z $ could be either discrete, ordinal, or continuous. With mixed data, we
-mean an arbitrary mix of variable types among $ X $, $ Y $, and $ Z $.
+We represent uni-dimensional variables as $ X $ and multi-dimensional variables
+as $ \bm{Z} $. We consider the case of estimating association between $ X $,
+and $ Y $ is the presence of a set of conditioning variables $ \bm{Z} = Z_1,
+\cdots, Z_k $, potentially being an empty set $ \bm{Z} = \emptyset $. We use $
+\rvert \bm{Z} \rvert $ to denote the cardinality of the variable $ \bm{Z} $. We
+consider various scenarios where the variables $ X $, $ Y $ and $ \bm{Z} $
+could be either discrete, ordinal, or continuous. With mixed data, we mean an
+arbitrary mix of variable types among $ X $, $ Y $, and $ \bm{Z} $.
 
@@
 
-\subsection{Effect Size Measures}
+\subsection{Effect size Measures}
+In this section, we give a background on some of the commonly used effect size
+measures in the context of CI testing.
 
-\subsubsection{$ X $ and $ Y $ are continuous}
+\paragraph{$ X $ and $ Y $ are continuous}
 In the case when $ \bm{Z} = \emptyset $, and when both $ X $ and $ Y $ are
@@
 
-\subsubsection{All $ X $, $ Y $, $ Z $ are discrete}
+\paragraph{All variables are discrete}
 
@@
 
-\subsubsection{$ X $ is ordinal and $ Y $ is continuous}
+\paragraph{$ X $ is ordinal and $ Y $ is continuous}
 Polyserial Correlation
@@
 
-\subsubsection{$ X $ and $ Y $ are ordinal}
+\paragraph{$ X $ and $ Y $ are ordinal}
 Polychoric Correlation
 
-\subsubsection{Mixed data}
+\paragraph{Mixed data}
 Recently multiple tests for mixed data have been proposed but they do not have
@@
 
-\subsection{Canonical Correlations}
-Canonical correlations are generalization of correlations to multi-dimensional case.
+\subsection{Canonical Correlation}
+Canonical correlations are generalization of correlations to multi-dimensional variables.
 
-Multiple effect sizes measures based on canonical correations are available.
+\begin{definition}
+	Given two sets of random variables $ \bm{U} = (U_1, U_2, \cdots, U_p) $
+	and $ \bm{V} = (V_1, V_2, \cdots, V_q) $, canonical correlation between
+	them, $\rho(\bm{U}, \bm{V}) $ is defined as:
+
+	\begin{equation}
+		\rho(\bm{U}, \bm{V}) = \sup_{a, b} \frac{a^T \Sigma_{\bm{U}\bm{V}} b}{\sqrt{a^T \Sigma_{\bm{U}\bm{U}} a} \sqrt{b^T \Sigma_{\bm{V}\bm{V}} b}}
+	\end{equation}
+
+\end{definition}
+
+	Canonical correlations try to find linear transformations $ a $ and $ b
+	$ of $ \bm{U} $ and $ \bm{V} $ such that the correlations between the
+	columns of $ a^T \bm{U} $ and $ b^T \bm{V} $ is maximized. The
+	individual columns of $ a^T \bm{U} $ and $ b^T \bm{V} $ are
+	independent. In the case when $ \rvert U \rvert = \rvert V \rvert = 1$,
+	canonical correlations are equivalent to Pearson's correlation
+	coefficient.
+
+Multiple effect sizes measures based on canonical correlation are available.
+\begin{itemize}
+	\item Wilks' Lambda: $ \Lambda = \prod (1 - \phi_i^2) $
+	\item Roy's Largest Root: $ \theta = \phi_{\max}^2 $
+	\item Pillai's Trace: $ \tau = \sum \phi_i^2 $
+\end{itemize}
+We use Pillai's Trace in this paper. \todo[inline]{Why?}.
 
 \section{Effect size for mixed data}
-In this section, we use canonical correlations to define an effect size for mixed data CI tests. The main idea is to transform any CI test into j
-We consider the following two cases separately:
+In this section, we use canonical correlations to define an effect size measure
+for mixed data CI tests. Our approach is similar to the approach used for the
+continuous variable case.
 
 \subsection{No conditional variables}
-When we have no conditonal variables, we are trying
+When we have no conditonal variables, we need to convert the variables into
 
@@
 
-\begin{figure}
-	\centering
-	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.25]{../../presentations/2024_05_das/1.png}
-	\end{subfigure}%
-	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.25]{../../presentations/2024_05_das/2.png}
-	\end{subfigure}
-	\caption{\todo[inline]{Insert screenshots of the web-tool}}
-\end{figure}
+% \begin{figure}
+% 	\centering
+% 	\begin{subfigure}{0.5\textwidth}
+% 		\includegraphics[scale=0.25]{../../presentations/2024_05_das/2.png}
+% 	\end{subfigure}%
+% 	\begin{subfigure}{0.5\textwidth}
+% 		\includegraphics[scale=0.25]{../../presentations/2024_05_das/5.png}
+% 	\end{subfigure}
+% 	\caption{Screenshots of the web-tool. \todo[inline]{Insert screenshots of the web-tool}}
+% \end{figure}
 
@@
 To compare how well does creating a graph manually based on effect sizes and
-p-value compare to algorithmic approaches, we performed some emperical
+p-value compare to algorithmic approaches, we performed some empirical
 analysis.
@@
 
-\todo[inline]{Summarize this section and equaiton in an algorithm}
+\todo[inline]{Summarize this section and equation in an algorithm}
 \todo[inline]{Show a consistency proof for this expert simulation}
@@
 \end{equation}
+
+\begin{algorithm}
+	\caption{Expert Simulation}
+	\While{True}{
+	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $ \;
+	   \If{$ \textnormal{len}(e) $}{
+		   $ \textnormal{return}(G) $ \;
+	   }
+	   $ X, Y \gets \max_{X, Y} e $\;
+	   $ dir \gets Expert(X, Y) $ \;
+	}
+
+\end{algorithm}
 

### 2024-05-10 paper/main.tex
Updates

@@
 \begin{abstract}
-	Constructing a Directed Acyclic Graph (DAG) from observational data,
-	called causal discovery, is a fundamental problem in causal inference.
-	Numerous algorithms have been developed to address it. However, their
-	adoption in applied disciplines has been limited, possibly due to a
-	lack of confidence in these algorithms stemming from them making
-	obvious mistakes, and the difficulties in choosing the right algorithm
-	for specific datasets. Consequently, in applied disciplines, these
-	causal Directed Acyclic Graphs (DAGs) are predominantly constructed
-	manually based on domain expertise. This makes it vital to test and
-	potentially modify these models against data before using them for any
-	analysis. Local testing based on Conditional Independence (CI) tests and
-	effect sizes have been commonly used for this purpose. However,
-	one of the current limitations of this approach is that there are no
-	effect sizes for mixed data. To address this, we introduce a novel
-	effect size measure for mixed data based on canonical correlations.
-	This effect size is a generalization of some of the effect sizes used
-	in special cases to mixed variables. Moreover, to assist researchers in
-	constructing DAGs manually, we use the proposed mixed data effect size
-	measure to build a web-based tool that can help researchers in
-	constructing, validating, and modifying their DAGs with the help of
-	available data. The tool iteratively suggests modifications to DAGs
-	based on correlations that are not explained/over-explained by the DAG,
-	while the expert chooses the direction of the edge between those
-	variables. Empirically, we show that taking a greedy approach on the
-	tool's recommendations can perform comparably to causal discovery
-	algorithms, provided that the expert correctly determines the direction
-	of the edges in at least one out of three cases, with performance
-	improving as accuracy increases.
+	% Constructing a Directed Acyclic Graph (DAG) from observational data,
+	% called causal discovery, is a fundamental problem in causal inference.
+	% Numerous algorithms have been developed to address it. However, their
+	% adoption in applied disciplines has been limited, possibly due to a
+	% lack of confidence in these algorithms stemming from them making
+	% obvious mistakes, and the difficulties in choosing the right algorithm
+	% for specific datasets. Consequently, in applied disciplines, these
+	% causal Directed Acyclic Graphs (DAGs) are predominantly constructed
+	% manually based on domain expertise. This makes it vital to test and
+	% potentially modify these models against data before using them for any
+	% analysis. Local testing based on Conditional Independence (CI) tests and
+	% effect sizes have been commonly used for this purpose. However,
+	% one of the current limitations of this approach is that there are no
+	% effect sizes for mixed data. To address this, we introduce a novel
+	% effect size measure for mixed data based on canonical correlations.
+	% This effect size is a generalization of some of the effect sizes used
+	% in special cases to mixed variables. Moreover, to assist researchers in
+	% constructing DAGs manually, we use the proposed mixed data effect size
+	% measure to build a web-based tool that can help researchers in
+	% constructing, validating, and modifying their DAGs with the help of
+	% available data. The tool iteratively suggests modifications to DAGs
+	% based on correlations that are not explained/over-explained by the DAG,
+	% while the expert chooses the direction of the edge between those
+	% variables. Empirically, we show that taking a greedy approach on the
+	% tool's recommendations can perform comparably to causal discovery
+	% algorithms, provided that the expert correctly determines the direction
+	% of the edges in at least one out of three cases, with performance
+	% improving as accuracy increases.
+
+	Causal Discovery that is learning the causal graph from observational
+	data is a fundamental problem in causal inference. The problem has been
+	extensively studied and many algorithms have been developed to address
+	it. However, their adaption in applied fields have been limited,
+	possibly due to a lack of confidence in these algorithms stemming from
+	them making obvious mistakes, and the difficulties in choosing the
+	right algorithm for specific datasets. Moreover, these algorithms can
+	usuallly only get us to markov equivance class where the remaining
+	edges need to be somehow oriented before inference can be done from
+	them. Consequently, in applied disciplines, these Directed Acyclic
+	Graphs (DAGs) are predominantly constructed manually based on domain
+	expertise. Due to being constructed by hand, testing these models is
+	vital before making any inference using them. Local testing in these
+	models can be done by checking whether the implied conditional
+	independences (CIs) by the model hold in the data or not. (Shipley
+	2000) also proposed a global test based on combining these local tests.
+	Based on these local tests, in the case of continuous or discrete
+	variables, we can start by modifying the worst offending test.
 \end{abstract}
@@
 	\While{True}{
-	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $ \;
-	   \If{$ \textnormal{len}(e) $}{
-		   $ \textnormal{return}(G) $ \;
+	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $\;
+	   $ e \gets e - \textnormal{blacklist} $\;
+	   \If {$ e = 0 $}{
+		$ G \gets G - e $ \;
+		$ \textnormal{blacklist} \gets \textnormal{blacklist} \cup e $\;
+	   } \;
+	   \If{$ \textnormal{len}(e) = 0$}{
+		   $ \textnormal{return}(G) $\;
 	   }
 	   $ X, Y \gets \max_{X, Y} e $\;
-	   $ dir \gets Expert(X, Y) $ \;
+	   $ dir \gets Expert(X, Y, \alpha) $\;
+	   $ G \gets G \cup dir $\;
 	}

### 2024-05-11 paper/main.tex
Udpates

@@
 	right algorithm for specific datasets. Moreover, these algorithms can
-	usuallly only get us to markov equivance class where the remaining
-	edges need to be somehow oriented before inference can be done from
-	them. Consequently, in applied disciplines, these Directed Acyclic
-	Graphs (DAGs) are predominantly constructed manually based on domain
-	expertise. Due to being constructed by hand, testing these models is
-	vital before making any inference using them. Local testing in these
-	models can be done by checking whether the implied conditional
-	independences (CIs) by the model hold in the data or not. (Shipley
-	2000) also proposed a global test based on combining these local tests.
-	Based on these local tests, in the case of continuous or discrete
-	variables, we can start by modifying the worst offending test.
+	usuallly only get us to markov equivance class (MEC) where the
+	remaining edges need to be somehow oriented before inference can be
+	done from them. Consequently, in applied disciplines, these Directed
+	Acyclic Graphs (DAGs) are predominantly constructed manually based on
+	domain expertise. Due to being constructed by hand, testing these
+	models is vital before making any inference using them. Local testing
+	in these models can be done by checking whether the implied conditional
+	independences (CIs) by the model hold in the data or not. In the field
+	on Structural Equation Models (SEMs), modification indices have been
+	used for this purpose where a global fit index is computed relative to
+	each possible modification in the model. These modification indices can
+	be used to select the appropriate modification backed by expert
+	knowledge. In the case of DAGs, we do have global fit index based on
+	combining local implied conditional independence tests. Based on this
+	test, we can make modifications to our model. This can for example be
+	done by using a conditional association measure
 \end{abstract}

### 2024-05-12 code/ci_test.R
Updates

@@ -27,7 +27,7 @@ p2r <- function( x, preddf ){
 		]
 }
 
-residual.test <- function(x, y, x_parents, y_parents, predictor, test) {
+residual.test <- function(x, y, z, predictor, test) {
   # Random Forest test v1. Return the p-value for the independence
   # test X _|_ Y | Z
   #
@@ -95,51 +95,32 @@ residual.test <- function(x, y, x_parents, y_parents, predictor, test) {
     x_prob <- FALSE
   }
 
-  if (length(x_parents) == 0){
-	  Rx <- x
+  if (predictor == 'rf'){
+      forest.x <- ranger::ranger(as.formula(paste("x ~", paste(colnames(z), collapse = "+"))),
+        data.frame(x, z),
+        num.trees = 50,
+        probability = x_prob
+      )
+      predict.x <- predict(forest.x, data = data.frame(z))$predictions
   }
-  else{
-  	if (predictor == 'rf'){
-  	    forest.x <- ranger::ranger(as.formula(paste("x ~", paste(colnames(z), collapse = "+"))),
-  	      data.frame(x, x_parents),
-  	      num.trees = 50,
-  	      probability = x_prob
-  	    )
-  	    predict.x <- predict(forest.x, data = data.frame(x_parents))$predictions
-  	}
-  	else if (predictor == 'glm'){
-  	    if (x_type == 'cont'){
-  	            lm.x <- lm(as.formula(paste("x ~", paste(colnames(x_parents), collapse = "+"))), data=data.frame(x, x_parents))
-  	            predict.x <- predict(lm.x, data.frame(x_parents))
-  	    }
-  	    else if (x_type == 'ord'){
-  	            # prop.x <- VGAM::vglm(as.formula(paste("x ~", paste(colnames(z), collapse="+"))), data=data.frame(x, z), family=VGAM::propodds, model=TRUE)
-  	  	      # predict.x <- as.data.frame(predict(prop.x, newdata=data.frame(z), type='response'))
-  	            prop.x <- ordinal::clm(as.formula(paste("x ~", paste(colnames(x_parents), collapse="+"))), data=data.frame(x, x_parents))
-  	            predict.x <- predict(prop.x, data.frame(x_parents))$fit
-  	    }
-  	    else{
[…]

### 2024-05-12 code/sl.R

@@ -204,7 +204,7 @@ run_single_exp_pc <- function(n_nodes, edge_prob){
 	true_adj <- d$true_adj
 	sim_data <- d$sim_data
 
-    	suffStat <- list(dm = sim_data, adaptDF = FALSE, test = 'mxm')
+    	suffStat <- list(dm = sim_data, adaptDF = FALSE, test = 'glm_q3')
     	alpha <- 0.05
 
 	pc.cpdag <- pcalg::pc(

### 2024-05-12 paper/main.tex
Updates

--- 2024-05-07 paper/main.tex
@@
 \begin{abstract}
-	Constructing a Directed Acyclic Graph (DAG) from observational data,
-	called causal discovery, is a fundamental problem in causal inference.
-	Numerous algorithms have been developed to address it. However, their
-	adoption in applied disciplines has been limited, possibly due to a
-	lack of confidence in these algorithms stemming from them making
-	obvious mistakes, and the difficulties in choosing the right algorithm
-	for specific datasets. Consequently, in applied disciplines, these
-	causal Directed Acyclic Graphs (DAGs) are predominantly constructed
-	manually based on domain expertise. This makes it vital to test and
-	potentially modify these models against data before using them for any
-	analysis. Local testing based on Conditional Independence (CI) tests and
-	effect sizes have been commonly used for this purpose. However,
-	one of the current limitations of this approach is that there are no
-	effect sizes for mixed data. To address this, we introduce a novel
-	effect size measure for mixed data based on canonical correlations.
-	This effect size is a generalization of some of the effect sizes used
-	in special cases to mixed variables. Moreover, to assist researchers in
-	constructing DAGs manually, we use the proposed mixed data effect size
-	measure to build a web-based tool that can help researchers in
-	constructing, validating, and modifying their DAGs with the help of
-	available data. The tool iteratively suggests modifications to DAGs
-	based on correlations that are not explained/over-explained by the DAG,
-	while the expert chooses the direction of the edge between those
-	variables. Empirically, we show that taking a greedy approach on the
-	tool's recommendations can perform comparably to causal discovery
-	algorithms, provided that the expert correctly determines the direction
-	of the edges in at least one out of three cases, with performance
-	improving as accuracy increases.
+	\emph{Causal Discovery} is the process of learning the causal structure
+	either in form of a Directed Acyclic Graph (DAG) or a Structural
+	Equation Model (SEM) from observational data. This problem has been
+	extensively studied for both DAGs and SEMs.
+	In the field of DAGs, many automated algorithms
+	have been developed for constructing a DAG from a given dataset. However, the
+	adaption of these algorithms in applied fields have been limited,
+	possibly due to a lack of confidence in these algorithms stemming from
+	them making obvious mistakes, and the difficulties in choosing the
+	right algorithm for a given dataset. Consequently, in applied
+	disciplines, these DAGs are predominantly constructed manually based on
+	domain expertise. Due to being constructed by hand, it is vital that
+	the fit of these models are tested against data and the models are
+	potentially modified to acheive a better fit. On the other hand, in the
+	field of SEMs models are typically constructed manually through an
+	iterative process of assessing fit and modifying the model according to
+	that. This process is commonly known as \emph{Specification Search} 
+	where things like \emph{modification indices} can be used to guide this
+	modification process. In the case of DAGs, although there are tests
+	such as Shipley 2000 that can assess the global fit of model by
+	combining the implied conditional independencies of the model, but
+	there are no methods that can guide our modification process. In this
+	paper, we present a simple modification process based on using a
+	measure of association for these individual local CI tests and
+	combining it with expert knowledge. We also present a measure of
+	association based on canonical correlations for mixed data. We also
+	present a graphical web-tool that can assist researchers in iteratively
+	modifying and constructing their DAGs. We theoretically show that in
+	the presence of an oracle, this iterative modification process is
+	consistent in recovering the true DAG. Emprically, we show that this
+	process performs comparably to PC and Hill-Climb Search algorithms if
+	the user is able to correctly identify the direction of the edge in one
+	out of three cases.
 \end{abstract}

### 2024-05-12 paper/main.tex
udpates

@@
 	that. This process is commonly known as \emph{Specification Search} 
-	where things like \emph{modification indices} can be used to guide this
+	where methods like \emph{modification indices} can be used to guide this
 	modification process. In the case of DAGs, although there are tests
 	such as Shipley 2000 that can assess the global fit of model by
-	combining the implied conditional independencies of the model, but
+	combining the tests using implied conditional independencies of the model, but
 	there are no methods that can guide our modification process. In this
-	paper, we present a simple modification process based on using a
-	measure of association for these individual local CI tests and
+	paper, we present a simple modification process aimed at helping researchers
+	in manually constructing or modifying the DAGs outputted by an algorithm.
+	This modification process is based on using a measure of association for 
+	these individual local CI tests to identify the worst offending CI test and
 	combining it with expert knowledge. We also present a measure of

### 2024-05-12 paper/main.tex
Updates

@@
 \begin{abstract}
-	\emph{Causal Discovery} is the process of learning the causal structure
-	either in form of a Directed Acyclic Graph (DAG) or a Structural
-	Equation Model (SEM) from observational data. This problem has been
-	extensively studied for both DAGs and SEMs.
-	In the field of DAGs, many automated algorithms
-	have been developed for constructing a DAG from a given dataset. However, the
-	adaption of these algorithms in applied fields have been limited,
-	possibly due to a lack of confidence in these algorithms stemming from
-	them making obvious mistakes, and the difficulties in choosing the
-	right algorithm for a given dataset. Consequently, in applied
-	disciplines, these DAGs are predominantly constructed manually based on
-	domain expertise. Due to being constructed by hand, it is vital that
-	the fit of these models are tested against data and the models are
-	potentially modified to acheive a better fit. On the other hand, in the
-	field of SEMs models are typically constructed manually through an
-	iterative process of assessing fit and modifying the model according to
-	that. This process is commonly known as \emph{Specification Search} 
-	where methods like \emph{modification indices} can be used to guide this
-	modification process. In the case of DAGs, although there are tests
-	such as Shipley 2000 that can assess the global fit of model by
-	combining the tests using implied conditional independencies of the model, but
-	there are no methods that can guide our modification process. In this
-	paper, we present a simple modification process aimed at helping researchers
-	in manually constructing or modifying the DAGs outputted by an algorithm.
-	This modification process is based on using a measure of association for 
-	these individual local CI tests to identify the worst offending CI test and
-	combining it with expert knowledge. We also present a measure of
-	association based on canonical correlations for mixed data. We also
+	\emph{Causal Discovery} involves learning causal structures
+	from data, typically represented as  Directed Acyclic Graphs
+	(DAGs) or Structural Equation Models (SEMs). This problem has been
+	extensively studied for both DAGs and SEMs. In the field of DAGs, many
+	automated algorithms have been developed for constructing a DAG from a
+	given dataset. However, the adaption of these algorithms in applied
+	fields have been limited, possibly due to a lack of confidence in these
+	algorithms stemming from them making obvious mistakes, and the
+	difficulties in choosing the right algorithm for a given dataset.
+	Consequently, in applied disciplines, these DAGs are predominantly
+	constructed manually based on domain expertise. Due to being
+	constructed by hand, it is vital that the fit of these models are
+	tested against data and the models are potentially modified to acheive
+	a better fit. On the other hand, in the field of SEMs models are
+	typically constructed manually through an iterative process of
+	assessing fit and modifying the model according to that. This process
+	is commonly known as \emph{Specification Search} where methods like
+	\emph{modification indices} can be used to guide this modification
+	process. In the case of DAGs, although there are tests such as Shipley
+	2000 that can assess the global fit of model by combining the tests
+	using implied conditional independencies of the model, but there are no
+	methods that can guide our modification process. In this paper, we
+	present a simple modification process aimed at helping researchers in
+	manually constructing or modifying the DAGs outputted by an algorithm.
+	This modification process is based on using a measure of association
+	for these individual local CI tests to identify the worst offending CI
+	test and combining it with expert knowledge. We also present a measure
+	of association based on canonical correlations for mixed data. We also
 	present a graphical web-tool that can assist researchers in iteratively
@@
 
-\textbf{Para 2: Reasons behind this}
+\textbf{Para 2: Reasons behind this and importance of verifying against data}
 
@@
 \end{enumerate}
-
-\textbf{Para 3: Importance of testing, include modification indices}
-
 Both in the cases when the DAG is an output of an algorithm or when constructed
 by hand, it is vital to test these models against data before using them for
-making any inference. In the field of Structural Equation Models (SEMs)
+making any inference.
+
+\textbf{Para 3: Structural Equation Models, testing, include modification indices}
+
+In the field of Structural Equation Models (SEMs)
 modification indices have been used to do this. However, as modification index
@@
 
-\textbf{Para 4: Mixed data testing}
-
+\textbf{Para 4: Measure of Association}
 As these CI based testing uses p-values and effect size measures from CI tests

### 2024-05-13 code/sl.R
Updates

@@ -255,7 +255,7 @@ run_sim <- function(R, n_nodes, edge_probs, oracle_accs){
 		hc_mean <- apply(hc_dist, mean, MARGIN=2)
 		hc_sd <- apply(hc_dist, sd, MARGIN=2)/sqrt(R)
 
-		pc_dist <- t(replicate(R, run_single_exp_pc(n_nodes=n_nodes, edge_prob=edge_prob)))
+		pc_dist <- t(future_replicate(R, run_single_exp_pc(n_nodes=n_nodes, edge_prob=edge_prob)))
 		pc_mean <- apply(pc_dist, function(x) mean(x, na.rm=T), MARGIN=2)
 		pc_sd <- apply(pc_dist, function(x) sd(x, na.rm=T), MARGIN=2)/sqrt(sum(!is.na(pc_dist[, 1])))
 

### 2024-05-13 paper/main.tex
udpates

@@
 	\emph{Causal Discovery} involves learning causal structures
-	from data, typically represented as  Directed Acyclic Graphs
-	(DAGs) or Structural Equation Models (SEMs). This problem has been
+	from data, and has been studied both in Directed Acyclic Graphs
+	(DAGs) and Structural Equation Models (SEMs). This problem has been
 	extensively studied for both DAGs and SEMs. In the field of DAGs, many
@@
 
-In the field of Structural Equation Models (SEMs)
-modification indices have been used to do this. However, as modification index
-is an evaluation of the global fit of the model. An alternative is local fit of
-models using CI tests where the implied conditional independences of the model
-can be tested using statistical tests. Similar to modification indices, these
-tests can then be used to modify our models. Figure shows an example of this
-testing. We can not also only rely on the p-values, effect size can help in
-deciding.
+In the field of Structural Equation Models (SEMs), models are typically
+constructed by hand. Researchers typically come up with an initial model, which
+then they use the data to make modifications to. This process is known as
+\emph{Specification Search}. One of the common ways to do this is through
+\emph{modification indices} where a global fit of the model is measured with
+different modification to the model. Utilizing expert knowledge to select the
+appropriate modification that gives the maximum improvement to the global fit.
+
+% modification indices have been used to do this. However, as modification index
+% is an evaluation of the global fit of the model. An alternative is local fit of
+% models using CI tests where the implied conditional independences of the model
+% can be tested using statistical tests. Similar to modification indices, these
+% tests can then be used to modify our models. Figure shows an example of this
+% testing. We can not also only rely on the p-values, effect size can help in
+% deciding.
 
 \textbf{Para 4: Measure of Association}
+In the case of DAGs however, we do not have any way to 
 As these CI based testing uses p-values and effect size measures from CI tests

### 2024-05-13 paper/main.tex
udpates

@@
 \begin{enumerate}
-	\item We present an effect size measure based on canonical correlations
-		for mixed data that generalizes some of the commonly used
-		effect size in special cases such as partial correlation
-		coefficient for continuous variables, and Cram\'er's V for
-		discrete variables.
-	\item We also developed a web-based tool to help researches in
-		constructing and testing and modifying their DAGs using CI tests and
-		effect size measures.
+	\item We present a procedure to construct and modify DAGs using domain
+		knowledge.
+	\item The presented procedure uses a (conditional) measure of
+		association between variables. In the simple case of continuous
+		or discrete variables, effect size measures can be used,
+		however as no such measure exists for mixed data, we present
+		one.
+	\item Lastly, to assist researchers in using the DAG construction
+		approach, we developed a web-based tool to help researches in
+		constructing and modifying their DAGs.
 \end{enumerate}

### 2024-05-13 paper/main.tex
Updates

@@
 \section{Introduction}
-\textbf{Para 1: Many Causal discovery algorithms but commonly constructed manually.}
-
-\todo[inline]{Add citations}
 
@@
 optimization algorithms to come up with a graph.
-
-\textbf{Para 2: Reasons behind this and importance of verifying against data}
 
@@
 
-\textbf{Para 3: Structural Equation Models, testing, include modification indices}
-
 In the field of Structural Equation Models (SEMs), models are typically
@@
 
-% modification indices have been used to do this. However, as modification index
-% is an evaluation of the global fit of the model. An alternative is local fit of
-% models using CI tests where the implied conditional independences of the model
-% can be tested using statistical tests. Similar to modification indices, these
-% tests can then be used to modify our models. Figure shows an example of this
-% testing. We can not also only rely on the p-values, effect size can help in
-% deciding.
-
-\textbf{Para 4: Measure of Association}
 In the case of DAGs however, we do not have any way to 
@@
 
-\textbf{Para 5: Contributions}
 \begin{enumerate}

### 2024-05-13 paper/main.tex
Updates

@@
 \begin{abstract}
-	\emph{Causal Discovery} involves learning causal structures
-	from data, and has been studied both in Directed Acyclic Graphs
-	(DAGs) and Structural Equation Models (SEMs). This problem has been
-	extensively studied for both DAGs and SEMs. In the field of DAGs, many
-	automated algorithms have been developed for constructing a DAG from a
-	given dataset. However, the adaption of these algorithms in applied
-	fields have been limited, possibly due to a lack of confidence in these
+	\emph{Causal Discovery} involves learning causal structures from data,
+	and has been studied both in Directed Acyclic Graphs (DAGs) and
+	Structural Equation Models (SEMs). This problem has been extensively
+	studied for both DAGs and SEMs. In the field of DAGs, many automated
+	algorithms have been developed for constructing a DAG from a given
+	dataset. However, the adaption of these algorithms in applied fields
+	have been limited, possibly due to a lack of confidence in these
 	algorithms stemming from them making obvious mistakes, and the
@@
 	\item Biased towards sparse graphs.
-	\item Algorithms making obvious mistakes: These algorithms often 
-		make very obvious mistakes, making it harder for the user
-		to trust them.
+	\item Algorithms making obvious mistakes: These algorithms often make
+		very obvious mistakes, making it harder for the user to trust
+		them.
 	\item Difficulties in choosing an algorithm: For any given dataset, the
@@
 \end{enumerate}
+
 Both in the cases when the DAG is an output of an algorithm or when constructed
@@
 
-In the case of DAGs however, we do not have any way to 
-As these CI based testing uses p-values and effect size measures from CI tests
-for testing. And as mixed data is common, we need both a test and effect size
-measure for mixed data. Recently, a few CI tests for mixed data have been
-proposed but an effect size is still missing.
+In the case of DAGs however, we do not have any way to As these CI based
+testing uses p-values and effect size measures from CI tests for testing. And
+as mixed data is common, we need both a test and effect size measure for mixed
+data. Recently, a few CI tests for mixed data have been proposed but an effect
+size is still missing.
 

### 2024-05-13 paper/main.tex
Updates

@@
 	is commonly known as \emph{Specification Search} where methods like
-	\emph{modification indices} can be used to guide this modification
-	process. In the case of DAGs, although there are tests such as Shipley
-	2000 that can assess the global fit of model by combining the tests
+	\emph{modification indices} combined with expert knowledge can be used
+	to guide this modification process. In the case of DAGs, although there
+	are tests to assess the global fit of a model by combining the tests
 	using implied conditional independencies of the model, but there are no
@@
 	present a simple modification process aimed at helping researchers in
-	manually constructing or modifying the DAGs outputted by an algorithm.
-	This modification process is based on using a measure of association
-	for these individual local CI tests to identify the worst offending CI
-	test and combining it with expert knowledge. We also present a measure
-	of association based on canonical correlations for mixed data. We also
-	present a graphical web-tool that can assist researchers in iteratively
-	modifying and constructing their DAGs. We theoretically show that in
-	the presence of an oracle, this iterative modification process is
-	consistent in recovering the true DAG. Emprically, we show that this
-	process performs comparably to PC and Hill-Climb Search algorithms if
-	the user is able to correctly identify the direction of the edge in one
-	out of three cases.
+	manually constructing or modifying DAGs outputted by an algorithm. This
+	modification process is based on using a (conditional) measure of
+	association between variables in the model to assess potential edges
+	that best explains unexplained correlation between variables which the
+	researcher specifies the direction of this edge. In the case of
+	continuous or discrete variables, the effect size measures of local
+	tests can be used as this measure of association, however for mixed
+	data there is no such measure. We present a measure of association
+	based on canonical correlations for mixed data. We also present a
+	graphical web-tool that can assist researchers in iteratively modifying
+	and constructing their DAGs. We theoretically show that in the presence
+	of an oracle, this iterative modification process is consistent in
+	recovering the true DAG. Emprically, we show that using the mixed data
+	measure of association, the modification process performs comparably to
+	PC and Hill-Climb Search algorithms if the user is able to correctly
+	identify the direction of the edge in one out of three cases.
 \end{abstract}
@@
 
-To perform any causal effect identification or estimation in Pearl's Directed
-Acyclic Graphs (DAGs) based causal framework usually requires to come up with
-the DAG structure. Learning these DAG structures from data is known as
-\emph{causal discovery}. Plenty of algorithms have been developed for these in
-the literature such as constraint-based methods such as PC, FCI, score based
-methods such as Hill-Climb Search, GES, and constrained continuous optimization
-based such as No TEARS. All these algorithms take different approaches to
-causal discovery, constraint-based algorithms utilize statistical conditional
-independence (CI) tests to come up with a model structure, the score-based
-methods iteratively perform local modifications to the model such that some
-score function is maximized, the continuous optimization methods reformulate
-the problem statement as a constrained optimization problem and use
-optimization algorithms to come up with a graph.
+To perform any causal effect identification or estimation using Directed
+Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
+first construct a causal structure. This process of constructing the casual
+structure from data is known as \emph{Causal Discovery}. In the field of DAGs,
+plenty of algorithms have been developed to automatically construct DAGs from
+data. These algorithm takes different approaches to causal discovery such as
+Constraint-based algorithms such as PC and Fast Causal Infenrece (FCI) where
+the algorithm tries to match the conditional independences in the data to the
+ones impled by the model, score-based methods such as Hill-Climb Search, Greedy
+Equivalence Search (GES)  where the algorithm tries to find a model that has
+the best score given some scoring metric, and continuous optimization where the
+problem is formulated as a constrained optimization problem such as No Tears.
 
@@
 Both in the cases when the DAG is an output of an algorithm or when constructed
-by hand, it is vital to test these models against data before using them for
-making any inference.
+by hand, it is vital to test these models against data and potentially modifying 
+them for a better fit before using them for making any inference. Both global
+and local tests for testing the fit of these models are available. However, when
+the fit fails, we don't have a procedure to modify these models.
 
-In the field of Structural Equation Models (SEMs), models are typically
+In the field of Structural Equation Models (SEMs), models are also typically
 constructed by hand. Researchers typically come up with an initial model, which
-then they use the data to make modifications to. This process is known as
-\emph{Specification Search}. One of the common ways to do this is through
-\emph{modification indices} where a global fit of the model is measured with
-different modification to the model. Utilizing expert knowledge to select the
-appropriate modification that gives the maximum improvement to the global fit.
+then they use the data and expert knowledge to make modifications to. This
+process is known as \emph{Specification Search}. One of the common ways to do
+this is through \emph{modification indices} where a global fit of the model is
+measured with different modification to the model. Utilizing expert knowledge
+to select the appropriate modification that gives the maximum improvement to
+the global fit.
 
-In the case of DAGs however, we do not have any way to As these CI based
-testing uses p-values and effect size measures from CI tests for testing. And
-as mixed data is common, we need both a test and effect size measure for mixed
-data. Recently, a few CI tests for mixed data have been proposed but an effect
-size is still missing.
+We present a similar procedure for modifying DAGs based on expert knowledge and
+measure of association. We first use the measure of association to test whether
+the association observed in the data is explained by the model structure. Based 
+on this testing, we suggest adding edges by ranking them by the most unexplained
+association between variables. The researcher can then use expert knowledge to 
+decide the direction of the edge between the variables.
 
+Our main contributions in this papers are as follows:
 \begin{enumerate}
-	\item We present a procedure to construct and modify DAGs using domain
-		knowledge.
-	\item The presented procedure uses a (conditional) measure of
-		association between variables. In the simple case of continuous
-		or discrete variables, effect size measures can be used,
-		however as no such measure exists for mixed data, we present
-		one.
-	\item Lastly, to assist researchers in using the DAG construction
-		approach, we developed a web-based tool to help researches in
-		constructing and modifying their DAGs.
+	\item We present a measure of (conditional) assocation between mixed variables using
+		canonical correlations. This mixed variable association is a generalization 
+		of partial correlation to mixed variables.
+	\item We present a procedure to iteratively construct and modify DAGs using expert 
+		knowledge such that it improves the explained association between variables.
+	\item Lastly, to assist researchers in manully constructing and modifying DAGs,
+		we developed a web-based tool to iteratively modify their DAGs.
 \end{enumerate}

### 2024-05-13 paper/main.tex
Updates

--- 2024-05-13 paper/main.tex
@@
 \begin{abstract}
-	\emph{Causal Discovery} involves learning causal structures from data,
-	and has been studied both in Directed Acyclic Graphs (DAGs) and
-	Structural Equation Models (SEMs). This problem has been extensively
-	studied for both DAGs and SEMs. In the field of DAGs, many automated
-	algorithms have been developed for constructing a DAG from a given
-	dataset. However, the adaption of these algorithms in applied fields
-	have been limited, possibly due to a lack of confidence in these
+	The initial step in a causal analysis involves determining the causal
+	structure between the variables from data typically in the form of a
+	Directed Acyclic Graph (DAG) or an Structural Equation Model (SEM).
+	This procedure, known as \emph{causal discovery}, has been extensively
+	studied for both DAGs and SEMs. For DAGs, numerous automated algorithms
+	have been developed that can automatically estimate the structure from
+	a given dataset. However, the adaption of these algorithms in applied
+	fields remain limited, possibly due to a lack of confidence in these
 	algorithms stemming from them making obvious mistakes, and the
-	difficulties in choosing the right algorithm for a given dataset.
-	Consequently, in applied disciplines, these DAGs are predominantly
-	constructed manually based on domain expertise. Due to being
-	constructed by hand, it is vital that the fit of these models are
-	tested against data and the models are potentially modified to acheive
-	a better fit. On the other hand, in the field of SEMs models are
+	difficulties in choosing the right algorithm for a given dataset. As a
+	result, in applied disciplines, DAGs are predominantly constructed
+	manually based on domain expertise which makes it vital to test the fit
+	of these models to the data and potentially modifying them to improve
+	the fit. On the other hand, in the field of SEMs, models are
 	typically constructed manually through an iterative process of
 	assessing fit and modifying the model according to that. This process
-	is commonly known as \emph{Specification Search} where methods like
-	\emph{modification indices} can be used to guide this modification
-	process. In the case of DAGs, although there are tests such as Shipley
-	2000 that can assess the global fit of model by combining the tests
-	using implied conditional independencies of the model, but there are no
-	methods that can guide our modification process. In this paper, we
-	present a simple modification process aimed at helping researchers in
-	manually constructing or modifying the DAGs outputted by an algorithm.
-	This modification process is based on using a measure of association
-	for these individual local CI tests to identify the worst offending CI
-	test and combining it with expert knowledge. We also present a measure
-	of association based on canonical correlations for mixed data. We also
-	present a graphical web-tool that can assist researchers in iteratively
-	modifying and constructing their DAGs. We theoretically show that in
-	the presence of an oracle, this iterative modification process is
-	consistent in recovering the true DAG. Emprically, we show that this
-	process performs comparably to PC and Hill-Climb Search algorithms if
-	the user is able to correctly identify the direction of the edge in one
-	out of three cases.
+	is commonly known as \emph{specification search} where expert knowledge
+	combined with methods like \emph{modification indices} can be used to 
+	guide the modification process. In the case of DAGs, although
+	statistical tests are available that can assess the global fit of
+	model, but there are no methods that can guide our modification
+	process. In this paper, we present a simple modification process aimed
+	at helping researchers in manually constructing or modifying the DAGs
+	outputted by an algorithm to improve its fit. This modification process
+	identifies pairs of variables whose association is not fully explained
+	by the model. Once such an unexplained association is found, an expert
+	can start by adding edges between the worst offending pair of variables 
+	while using their expertise to decide the direction between
+	the variables. For the measure of association, in the case of
+	continuous and discrete variables we can use effect sizes, however no
+	such measure of association is present for mixed data. We propose a
+	novel (conditional) measure of association for mixed data based on
+	canonical correlations. We also present a graphical web-tool that can
+	assist researchers in iteratively modifying and constructing their
+	DAGs. We theoretically show that in the presence of an oracle, this
+	iterative modification process is consistent in recovering the true
+	DAG. Emprically, we show that this process performs comparably to PC
+	and Hill-Climb Search algorithms if the user is able to correctly
+	identify the direction of the edge in one out of three cases.
 \end{abstract}

### 2024-05-13 paper/main.tex
Updates

@@
 	This procedure, known as \emph{causal discovery}, has been extensively
-	studied for both DAGs and SEMs. For DAGs, numerous automated algorithms
-	have been developed that can automatically estimate the structure from
-	a given dataset. However, the adaption of these algorithms in applied
-	fields remain limited, possibly due to a lack of confidence in these
+	studied for both DAGs and SEMs. While numerous automated algorithms
+	exist for estimating DAGs from datasets, their application in applied
+	remain limited, possibly due to a lack of confidence in these
 	algorithms stemming from them making obvious mistakes, and the
@@
 	result, in applied disciplines, DAGs are predominantly constructed
-	manually based on domain expertise which makes it vital to test the fit
-	of these models to the data and potentially modifying them to improve
-	the fit. On the other hand, in the field of SEMs, models are
-	typically constructed manually through an iterative process of
-	assessing fit and modifying the model according to that. This process
-	is commonly known as \emph{specification search} where expert knowledge
-	combined with methods like \emph{modification indices} can be used to 
-	guide the modification process. In the case of DAGs, although
-	statistical tests are available that can assess the global fit of
+	manually based on domain expertise making it vital to test the fit of
+	these models to the data and potentially modifying them to improve the
+	fit. Conversely, SEMs are typically constructed manually where
+	researchers come up with an initial model based on domain knowledge and
+	then iteratively assess its fit and modify the model according to that.
+	This process is commonly known as \emph{specification search} where
+	expert knowledge combined with methods like \emph{modification indices}
+	is used to guide the modification process. In the case of DAGs,
+	although statistical tests are available that can assess the fit of the
 	model, but there are no methods that can guide our modification
-	process. In this paper, we present a simple modification process aimed
+	process.
+
+	In this paper, we present a method to guide the modification process aimed
 	at helping researchers in manually constructing or modifying the DAGs
 	outputted by an algorithm to improve its fit. This modification process
-	identifies pairs of variables whose association is not fully explained
+	identifies pairs of variables whose association in data is not fully explained
 	by the model. Once such an unexplained association is found, an expert
-	can start by adding edges between the worst offending pair of variables 
+	can iteratively add edges between the worst offending pair of variables 
 	while using their expertise to decide the direction between
@@
 	iterative modification process is consistent in recovering the true
-	DAG. Emprically, we show that this process performs comparably to PC
+	DAG. Empirically, we show that this process performs comparably to PC
 	and Hill-Climb Search algorithms if the user is able to correctly

### 2024-05-13 paper/main.tex
Updates

@@
 	exist for estimating DAGs from datasets, their application in applied
-	remain limited, possibly due to a lack of confidence in these
+	fields remain limited, possibly due to a lack of confidence in these
 	algorithms stemming from them making obvious mistakes, and the
@@
 	This process is commonly known as \emph{specification search} where
-	expert knowledge combined with methods like \emph{modification indices}
-	is used to guide the modification process. In the case of DAGs,
-	although statistical tests are available that can assess the fit of the
-	model, but there are no methods that can guide our modification
-	process.
+	expert knowledge combined with data-driven methods like
+	\emph{modification indices} is used to guide the modification process.
+	In the case of DAGs, although statistical tests are available that can
+	assess the fit of the model, there are no methods that can guide our
+	modification process to improve the fit.
 
-	In this paper, we present a method to guide the modification process aimed
-	at helping researchers in manually constructing or modifying the DAGs
-	outputted by an algorithm to improve its fit. This modification process
-	identifies pairs of variables whose association in data is not fully explained
-	by the model. Once such an unexplained association is found, an expert
-	can iteratively add edges between the worst offending pair of variables 
-	while using their expertise to decide the direction between
-	the variables. For the measure of association, in the case of
-	continuous and discrete variables we can use effect sizes, however no
-	such measure of association is present for mixed data. We propose a
-	novel (conditional) measure of association for mixed data based on
-	canonical correlations. We also present a graphical web-tool that can
-	assist researchers in iteratively modifying and constructing their
-	DAGs. We theoretically show that in the presence of an oracle, this
-	iterative modification process is consistent in recovering the true
-	DAG. Empirically, we show that this process performs comparably to PC
-	and Hill-Climb Search algorithms if the user is able to correctly
+	Inspired by the modification process in SEMs, in this paper, we
+	introduce a data-driven modification process aimed at assisting
+	researchers in manually constructing or modifying DAGs generated by an
+	algorithm. The first step of this process involves identifying pairs of
+	variables whose observed association is not fully explained by the
+	model. The user can then make modifications to the model by adding
+	edges between such variables while using their domain expertise to
+	decide the direction of the edge. For the measure of association, in
+	the case of continuous and discrete variables we can use effect sizes,
+	however no such measure of association is present for mixed data. We
+	propose a novel (conditional) measure of association for mixed data
+	based on canonical correlations. We also present a graphical web-tool
+	that can assist researchers in iteratively modifying and constructing
+	their DAGs. We theoretically show that in the presence of an oracle,
+	this iterative modification process is consistent in recovering the
+	true DAG. Empirically, we show that this process performs comparably to
+	PC and Hill-Climb Search algorithms if the user is able to correctly
 	identify the direction of the edge in one out of three cases.

### 2024-05-13 paper/main.tex
Updates

@@
 
-	Inspired by the modification process in SEMs, in this paper, we
+	Inspired by the specification search in SEMs, in this paper, we
 	introduce a data-driven modification process aimed at assisting

### 2024-05-28 paper/main.tex
Updates

@@
 	algorithms stemming from them making obvious mistakes, and the
-	difficulties in choosing the right algorithm for a given dataset. As a
-	result, in applied disciplines, DAGs are predominantly constructed
-	manually based on domain expertise making it vital to test the fit of
-	these models to the data and potentially modifying them to improve the
-	fit. Conversely, SEMs are typically constructed manually where
-	researchers come up with an initial model based on domain knowledge and
-	then iteratively assess its fit and modify the model according to that.
-	This process is commonly known as \emph{specification search} where
-	expert knowledge combined with data-driven methods like
-	\emph{modification indices} is used to guide the modification process.
-	In the case of DAGs, although statistical tests are available that can
-	assess the fit of the model, there are no methods that can guide our
-	modification process to improve the fit.
-
-	Inspired by the specification search in SEMs, in this paper, we
-	introduce a data-driven modification process aimed at assisting
-	researchers in manually constructing or modifying DAGs generated by an
-	algorithm. The first step of this process involves identifying pairs of
-	variables whose observed association is not fully explained by the
-	model. The user can then make modifications to the model by adding
-	edges between such variables while using their domain expertise to
-	decide the direction of the edge. For the measure of association, in
-	the case of continuous and discrete variables we can use effect sizes,
-	however no such measure of association is present for mixed data. We
-	propose a novel (conditional) measure of association for mixed data
-	based on canonical correlations. We also present a graphical web-tool
-	that can assist researchers in iteratively modifying and constructing
-	their DAGs. We theoretically show that in the presence of an oracle,
-	this iterative modification process is consistent in recovering the
-	true DAG. Empirically, we show that this process performs comparably to
+	difficulties in choosing the right algorithm for a given dataset.
+	Consequently, in applied disciplines, these DAGs are predominantly
+	constructed manually based on domain expertise. Due to being
+	constructed by hand, it is vital that the fit of these models are
+	tested against data and the models are potentially modified to acheive
+	a better fit. On the other hand, in the field of SEMs models are
+	typically constructed manually through an iterative process of
+	assessing fit and modifying the model according to that. This process
+	is commonly known as \emph{Specification Search} where methods like
+	\emph{modification indices} combined with expert knowledge can be used
+	to guide this modification process. In the case of DAGs, although there
+	are tests to assess the global fit of a model by combining the tests
+	using implied conditional independencies of the model, but there are no
+	methods that can guide our modification process. In this paper, we
+	present a simple modification process aimed at helping researchers in
+	manually constructing or modifying DAGs outputted by an algorithm. This
+	modification process is based on using a (conditional) measure of
+	association between variables in the model to assess potential edges
+	that best explains unexplained correlation between variables which the
+	researcher specifies the direction of this edge. In the case of
+	continuous or discrete variables, the effect size measures of local
+	tests can be used as this measure of association, however for mixed
+	data there is no such measure. We present a measure of association
+	based on canonical correlations for mixed data. We also present a
+	graphical web-tool that can assist researchers in iteratively modifying
+	and constructing their DAGs. We theoretically show that in the presence
+	of an oracle, this iterative modification process is consistent in
+	recovering the true DAG. Emprically, we show that using the mixed data
+	measure of association, the modification process performs comparably to
 	PC and Hill-Climb Search algorithms if the user is able to correctly

### 2024-05-31 paper/main.tex
Updates format

@@
-\documentclass{article}
+%File: anonymous-submission-latex-2024.tex
+\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
+\usepackage[submission]{aaai24}  % DO NOT CHANGE THIS
+\usepackage{times}  % DO NOT CHANGE THIS
+\usepackage{helvet}  % DO NOT CHANGE THIS
+\usepackage{courier}  % DO NOT CHANGE THIS
+\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
+\usepackage{graphicx} % DO NOT CHANGE THIS
+\urlstyle{rm} % DO NOT CHANGE THIS
+\def\UrlFont{\rm}  % DO NOT CHANGE THIS
+\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
+\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
+\frenchspacing  % DO NOT CHANGE THIS
+\setlength{\pdfpagewidth}{8.5in} % DO NOT CHANGE THIS
+\setlength{\pdfpageheight}{11in} % DO NOT CHANGE THIS
+%
+% These are recommended to typeset algorithms but not required. See the subsubsection on algorithms. Remove them if you don't have algorithms in your paper.
+\usepackage{algorithm}
+\usepackage{algorithmic}
 
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2024
+%
+% These are are recommended to typeset listings but not required. See the subsubsection on listing. Remove this block if you don't have listings in your paper.
+\usepackage{newfloat}
+\usepackage{listings}
+\DeclareCaptionStyle{ruled}{labelfont=normalfont,labelsep=colon,strut=off} % DO NOT CHANGE THIS
+\lstset{%
+	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
+	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
+	aboveskip=0pt,belowskip=0pt,%
+	showstringspaces=false,tabsize=2,breaklines=true}
+\floatstyle{ruled}
+\newfloat{listing}{tb}{lst}{}
+\floatname{listing}{Listing}
+%
+% Keep the \pdfinfo as shown here. There's no need
+% for you to add the /Title and /Author tags.
+\pdfinfo{
+/TemplateVersion (2024.1)
+}
 
-% ready for submission
-\usepackage{neurips_2024}
+% DISALLOWED PACKAGES
+% \usepackage{authblk} -- This package is specifically forbidden
+% \usepackage{balance} -- This package is specifically forbidden
+% \usepackage{color (if used in text)
+% \usepackage{CJK} -- This package is specifically forbidden
+% \usepackage{float} -- This package is specifically forbidden
+% \usepackage{flushend} -- This package is specifically forbidden
+% \usepackage{fontenc} -- This package is specifically forbidden
+% \usepackage{fullpage} -- This package is specifically forbidden
+% \usepackage{geometry} -- This package is specifically forbidden
+% \usepackage{grffile} -- This package is specifically forbidden
+% \usepackage{hyperref} -- This package is specifically forbidden
+% \usepackage{navigator} -- This package is specifically forbidden
+% (or any other package that embeds links such as navigator or hyperref)
+% \indentfirst} -- This package is specifically forbidden
+% \layout} -- This package is specifically forbidden
+% \multicol} -- This package is specifically forbidden
+% \nameref} -- This package is specifically forbidden
+% \usepackage{savetrees} -- This package is specifically forbidden
+% \usepackage{setspace} -- This package is specifically forbidden
+% \usepackage{stfloats} -- This package is specifically forbidden
+% \usepackage{tabu} -- This package is specifically forbidden
+% \usepackage{titlesec} -- This package is specifically forbidden
+% \usepackage{tocbibind} -- This package is specifically forbidden
+% \usepackage{ulem} -- This package is specifically forbidden
+% \usepackage{wrapfig} -- This package is specifically forbidden
+% DISALLOWED COMMANDS
+% \nocopyright -- Your paper will not be published if you use this command
+% \addtolength -- This command may not be used
+% \balance -- This command may not be used
+% \baselinestretch -- Your paper will not be published if you use this command
+% \clearpage -- No page breaks of any kind may be used for the final version of your paper
+% \columnsep -- This command may not be used
+% \newpage -- No page breaks of any kind may be used for the final version of your paper
+% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
+% \pagestyle -- This command may not be used
+% \tiny -- This is not an acceptable font size.
+% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
+% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference
 
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2024}
+\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.
 
-% to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2024}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2024}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
-
-% Packages added by me.
 \usepackage{amsmath}
@@
 \usepackage{subcaption}
-\usepackage[ruled]{algorithm2e}
+% \usepackage[ruled]{algorithm2e}
 
@@
 
+% The file aaai24.sty is the style file for AAAI Press
+% proceedings, working notes, and technical reports.
+%
+
+% Title
+
+% Your title must be in mixed case, not sentence case.
+% That means all verbs (including short verbs like be, is, using,and go),
+% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
+% articles, conjunctions, and prepositions are lower case unless they
+% directly follow a colon or long dash
 \title{Title}
+\author{
+    %Authors
+    % All authors must be in the same font size and format.
+    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
+    AAAI Style Contributions by Pater Patel Schneider,
+    Sunil Issar,\\
+    J. Scott Penberthy,
+    George Ferguson,
+    Hans Guesgen,
+    Francisco Cruz\equalcontrib,
+    Marc Pujol-Gonzalez\equalcontrib
+}
+\affiliations{
+    %Afiliations
+    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
+    % If you have multiple authors and multiple affiliations
+    % use superscripts in text and roman font to identify them.
+    % For example,
 
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
+    % Sunil Issar\textsuperscript{\rm 2},
+    % J. Scott Penberthy\textsuperscript{\rm 3},
+    % George Ferguson\textsuperscript{\rm 4},
+    % Hans Guesgen\textsuperscript{\rm 5}
+    % Note that the comma should be placed after the superscript
+
+    1900 Embarcadero Road, Suite 101\\
+    Palo Alto, California 94303-3310 USA\\
+    % email address must be in roman text type, not monospace or sans serif
+    proceedings-questions@aaai.org
 %
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
+% See more examples next
+}
 
-\author{%
-  David S.~Hippocampus\thanks{Use footnote for providing further information
-    about author (webpage, alternative address)---\emph{not} for acknowledging
-    funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  % examples of more authors
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \AND
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
+%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
+\iffalse
+\title{My Publication Title --- Single Author}
+\author {
+    Author Name
 }
+\affiliations{
+    Affiliation\\
+    Affiliation Line 2\\
+    name@example.com
+}
+\fi
+
+\iffalse
+%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
+\title{My Publication Title --- Multiple Authors}
+\author {
+    % Authors
+    First Author Name\textsuperscript{\rm 1},
+    Second Author Name\textsuperscript{\rm 2},
+    Third Author Name\textsuperscript{\rm 1}
+}
+\affiliations {
+    % Affiliations
+    \textsuperscript{\rm 1}Affiliation 1\\
+    \textsuperscript{\rm 2}Affiliation 2\\
+    firstAuthor@affiliation1.com, secondAuthor@affilation2.com, thirdAuthor@affiliation1.com
+}
+\fi
 
@@
 
+\emph{Causal discovery} is the process of identifying and understanding causal
+relationships among variables in a dataset.
+
 To perform any causal effect identification or estimation using Directed
@@
 first construct a causal structure. This process of constructing the casual
-structure from data is known as \emph{Causal Discovery}. In the field of DAGs,
-plenty of algorithms have been developed to automatically construct DAGs from
-data. These algorithm takes different approaches to causal discovery such as
-Constraint-based algorithms such as PC and Fast Causal Infenrece (FCI) where
-the algorithm tries to match the conditional independences in the data to the
-ones impled by the model, score-based methods such as Hill-Climb Search, Greedy
-Equivalence Search (GES)  where the algorithm tries to find a model that has
-the best score given some scoring metric, and continuous optimization where the
-problem is formulated as a constrained optimization problem such as No Tears.
+structure from data is known as \emph{causal discovery}. In the field of DAGs,
+a lot of effort has been put on developing automated algorithms that can
+construct these DAGs from data. These algorithm takes different approaches to
+causal discovery such as Constraint-based algorithms such as PC and Fast Causal
+Infenrece (FCI) where the algorithm tries to match the conditional
+independences in the data to the ones impled by the model, score-based methods
+such as Hill-Climb Search, Greedy Equivalence Search (GES)  where the algorithm
+tries to find a model that has the best score given some scoring metric, and
+continuous optimization where the problem is formulated as a constrained
+optimization problem such as No Tears.
 
@@
 		\begin{tikzpicture}
-			\begin{scope}[yshift=0.7cm, scale=1]
+			\begin{scope}[xshift=-2.5cm, yshift=0.7cm, scale=0.7]
 			\tikzstyle{every node}=[align=center, inner sep=1pt]
@@
 			\end{scope}
+			\begin{scope}[xshift=-3cm]
 				\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
-			\node[rectangle, align=center, inner sep=1pt] at (6, 0) {
-				\begin{minipage}{\textwidth}
+				\node[rectangle, align=center, inner sep=1pt] at (5, 0) {
+					\begin{minipage}{.2\textwidth}
 						\begin{equation*}
@@
 					};
+			\end{scope}
 		\end{tikzpicture}
@@
 
-\begin{algorithm}
-	\caption{Expert Simulation}
-	\While{True}{
-	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $\;
-	   $ e \gets e - \textnormal{blacklist} $\;
-	   \If {$ e = 0 $}{
-		$ G \gets G - e $ \;
-		$ \textnormal{blacklist} \gets \textnormal{blacklist} \cup e $\;
-	   } \;
-	   \If{$ \textnormal{len}(e) = 0$}{
-		   $ \textnormal{return}(G) $\;
-	   }
-	   $ X, Y \gets \max_{X, Y} e $\;
-	   $ dir \gets Expert(X, Y, \alpha) $\;
-	   $ G \gets G \cup dir $\;
-	}
-
-\end{algorithm}
+% \begin{algorithm}
+% 	\caption{Expert Simulation}
+% 	\While{True}{
+% 	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $\;
+% 	   $ e \gets e - \textnormal{blacklist} $\;
+% 	   \If {$ e = 0 $}{
+% 		$ G \gets G - e $ \;
+% 		$ \textnormal{blacklist} \gets \textnormal{blacklist} \cup e $\;
+% 	   } \;
+% 	   \If{$ \textnormal{len}(e) = 0$}{
+% 		   $ \textnormal{return}(G) $\;
+% 	   }
+% 	   $ X, Y \gets \max_{X, Y} e $\;
+% 	   $ dir \gets Expert(X, Y, \alpha) $\;
+% 	   $ G \gets G \cup dir $\;
+% 	}
+% 
+% \end{algorithm}
 
@@
 		\includegraphics{../code/plots/shd_ribbon.pdf}
-	\end{subfigure}%
+	\end{subfigure}
 	\begin{subfigure}{0.5\textwidth}
@@
 
-\bibliographystyle{plain}
-\bibliography{references.bib}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-\section{Appendix / supplemental material}
-
-Optionally include supplemental material (complete proofs, additional experiments and plots) in appendix.
-All such materials \textbf{SHOULD be included in the main submission.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\include{checklist}
-
+\bibliography{references}
 \end{document}

### 2024-05-31 paper/main.tex
Updates

@@
 \emph{Causal discovery} is the process of identifying and understanding causal
-relationships among variables in a dataset.
+relationships among variables in a dataset. Causal discovery techniques are
+used typically to understand and estimate causal effects between variables. In
+the To perform any causal effect estimation using Directed Acyclic Graphs
+(DAGs) or Structural Equation Models (SEMs), researchers need to first
+construct a causal structure. In the field of DAGs, a lot of attention has been
+put on developing automated algorithms that can construct these DAGs from data.
+These algorithm takes different approaches to causal discovery such as
+constraint-based algorithms such as PC and Fast Causal Infenrece (FCI) where
+the algorithm tries to match the conditional independences in the data to the
+ones impled by the model, score-based methods such as Hill-Climb Search, Greedy
+Equivalence Search (GES)  where the algorithm tries to find a model that has
+the best score given some scoring metric, and continuous optimization where the
+problem is formulated as a constrained optimization problem such as No Tears, and
+some hybrid approaches.
 
-To perform any causal effect identification or estimation using Directed
-Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
-first construct a causal structure. This process of constructing the casual
-structure from data is known as \emph{causal discovery}. In the field of DAGs,
-a lot of effort has been put on developing automated algorithms that can
-construct these DAGs from data. These algorithm takes different approaches to
-causal discovery such as Constraint-based algorithms such as PC and Fast Causal
-Infenrece (FCI) where the algorithm tries to match the conditional
-independences in the data to the ones impled by the model, score-based methods
-such as Hill-Climb Search, Greedy Equivalence Search (GES)  where the algorithm
-tries to find a model that has the best score given some scoring metric, and
-continuous optimization where the problem is formulated as a constrained
-optimization problem such as No Tears.
-
-Even though we have various algorithms for doing casual discovery, yet
-researchers in applied fields prefer to construct these DAGs manually based on
-their domain knowledge. We hypothesize the following two reasons for low
-adaption of these methods in applied sciences:
+Even though we have various algorithms for casual discovery, the adaption of these
+algorithms in applied fields remain limited. In these applied fields, researchers 
+typically prefer to construct these DAGs manually based on their domain knowledge.
+We hypothesize that this preference towards manual DAG construction stems from the 
+following difficulties in applying these methods to real dataset.
 
 \begin{enumerate}
-	\item Biased towards sparse graphs.
-	\item Algorithms making obvious mistakes: These algorithms often make
-		very obvious mistakes, making it harder for the user to trust
-		them.
-	\item Difficulties in choosing an algorithm: For any given dataset, the
-		user need to make a choice of the algorithm. Given that these
-		algorithms work very differently, it can be difficult to make a
-		choice. Additionally, each of these algorithms have
-		hyperparameters and choices to make for them, making it
-		difficult. As there is no way to evaluate the performance of
-		these algorithms on our dataset.
+	\item Algorithms output Markov Equivalence Class (MEC): Using observational
+		data, any causal discovery algorithm can only output the MEC.
+		MEC do not have all the edges oriented and to do further
+		hypothesis testing or causal effect estimation, researchers
+		typically have to rely on domain expertise to orient the
+		remaining edges in the MEC.
+	\item Difficulties in choosing an algorithm: To use any of these algorithms on 
+		any practical dataset, users need to make a lot of choices. For
+		example, for using constraint-based method, users need to
+		choose a suitable condtional independence (CI) test, and
+		hyperparameters for these tests. Similarly for score-based
+		methods users need to choose the scoring metrics and their
+		hyperparameters. Moreover, for practical implementation with
+		reasonable runtime, users also need to choose the algorithm
+		hyperparameters such as maximum number of conditioning
+		variables to consider, max indegree of variables in the model. With
+		no easy guides or method to evaluate these choices, it can be difficult.
+	\item Lack of trust: Most of these automated algorithms can make very obvious 
+		mistakes. Typically users have to test the outputs of the models and 
+		manually make corrections to the output to get a usable model.
+		This can lead to researchers not trusting these algorithms.
 \end{enumerate}
 
-Both in the cases when the DAG is an output of an algorithm or when constructed
+On the other hand, in the field of Structural Equation Models (SEMs), models
+are also typically constructed by hand. And a lot of focus has been put on
+developing tools to help researchers in constructing these models. Researchers
+typically come up with an initial model, which then they use the data and
+expert knowledge to make modifications to. This process is commonly known as
+\emph{Specification Search}. These tools can then be used to guide the
+researchers in suggesting best ways to modify the model. For example,
+\emph{modification indices} take the initial drawn model, and make all possible
+modifications to the model to check which modification improves the model fit
+the best. Based on this information, the researcher can make modifications to
+the model and reiterate this process till they get something that they are
+satisfied with.
+
+In the case of DAGs, global testing methodology does exist such as Shipley
+test, that can combine local structure tests into a single p-value test.
+However, this does not give us a way to make modifications to our model. A
+possible way to do this by using check which of these local tests are violated
+based on the
+
+In the cases when the DAG is an output of an algorithm or when constructed
 by hand, it is vital to test these models against data and potentially modifying 
@@
 the fit fails, we don't have a procedure to modify these models.
-
-In the field of Structural Equation Models (SEMs), models are also typically
-constructed by hand. Researchers typically come up with an initial model, which
-then they use the data and expert knowledge to make modifications to. This
-process is known as \emph{Specification Search}. One of the common ways to do
-this is through \emph{modification indices} where a global fit of the model is
-measured with different modification to the model. Utilizing expert knowledge
-to select the appropriate modification that gives the maximum improvement to
-the global fit.
 
@@
 
-\subsection{Effect size Measures}
+\subsection{Association Measures}
 In this section, we give a background on some of the commonly used effect size

### 2024-06-08 paper/main.tex
udpates

@@
 \emph{Causal discovery} is the process of identifying and understanding causal
-relationships among variables in a dataset. Causal discovery techniques are
-used typically to understand and estimate causal effects between variables. In
-the To perform any causal effect estimation using Directed Acyclic Graphs
-(DAGs) or Structural Equation Models (SEMs), researchers need to first
-construct a causal structure. In the field of DAGs, a lot of attention has been
-put on developing automated algorithms that can construct these DAGs from data.
-These algorithm takes different approaches to causal discovery such as
-constraint-based algorithms such as PC and Fast Causal Infenrece (FCI) where
-the algorithm tries to match the conditional independences in the data to the
-ones impled by the model, score-based methods such as Hill-Climb Search, Greedy
-Equivalence Search (GES)  where the algorithm tries to find a model that has
-the best score given some scoring metric, and continuous optimization where the
-problem is formulated as a constrained optimization problem such as No Tears, and
-some hybrid approaches.
+relationships among variables in a dataset.  To perform any causal effect
+estimation using Directed Acyclic Graphs (DAGs) or Structural Equation Models
+(SEMs), researchers need to first construct a causal structure. In the field of
+DAGs, a lot of attention has been put on developing automated algorithms that
+can construct these DAGs from data. These algorithm takes different approaches
+to causal discovery such as constraint-based algorithms such as PC and Fast
+Causal Infenrece (FCI) where the algorithm tries to match the conditional
+independences in the data to the ones impled by the model, score-based methods
+such as Hill-Climb Search, Greedy Equivalence Search (GES)  where the algorithm
+tries to find a model that has the best score given some scoring metric, and
+continuous optimization where the problem is formulated as a constrained
+optimization problem such as No Tears, and some hybrid approaches.
 
@@
 \emph{Specification Search}. These tools can then be used to guide the
-researchers in suggesting best ways to modify the model. For example,
-\emph{modification indices} take the initial drawn model, and make all possible
-modifications to the model to check which modification improves the model fit
-the best. Based on this information, the researcher can make modifications to
-the model and reiterate this process till they get something that they are
-satisfied with.
+researchers in suggesting best ways to modify the model combined with the
+expert knowledge. For example, \emph{modification indices} take the initial
+drawn model, and make all possible modifications to the model to check which
+modification improves the model fit the best. Based on this information, the
+researcher can make modifications to the model and reiterate this process till
+they get something that they are satisfied with.
 
@@
 
-\subsection{Association Measures}
+\subsection{Measures of Association}
 In this section, we give a background on some of the commonly used effect size
@@
 
-\section{Effect size for mixed data}
+\section{Measure of Association for mixed data}
 In this section, we use canonical correlations to define an effect size measure

### 2024-06-09 paper/main.tex
Updates

--- 2024-05-31 paper/main.tex
@@
-%File: anonymous-submission-latex-2024.tex
-\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
-\usepackage[submission]{aaai24}  % DO NOT CHANGE THIS
-\usepackage{times}  % DO NOT CHANGE THIS
-\usepackage{helvet}  % DO NOT CHANGE THIS
-\usepackage{courier}  % DO NOT CHANGE THIS
-\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
-\usepackage{graphicx} % DO NOT CHANGE THIS
-\urlstyle{rm} % DO NOT CHANGE THIS
-\def\UrlFont{\rm}  % DO NOT CHANGE THIS
-\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
-\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
-\frenchspacing  % DO NOT CHANGE THIS
-\setlength{\pdfpagewidth}{8.5in} % DO NOT CHANGE THIS
-\setlength{\pdfpageheight}{11in} % DO NOT CHANGE THIS
-%
-% These are recommended to typeset algorithms but not required. See the subsubsection on algorithms. Remove them if you don't have algorithms in your paper.
-\usepackage{algorithm}
-\usepackage{algorithmic}
+\documentclass{article}
 
-%
-% These are are recommended to typeset listings but not required. See the subsubsection on listing. Remove this block if you don't have listings in your paper.
-\usepackage{newfloat}
-\usepackage{listings}
-\DeclareCaptionStyle{ruled}{labelfont=normalfont,labelsep=colon,strut=off} % DO NOT CHANGE THIS
-\lstset{%
-	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
-	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
-	aboveskip=0pt,belowskip=0pt,%
-	showstringspaces=false,tabsize=2,breaklines=true}
-\floatstyle{ruled}
-\newfloat{listing}{tb}{lst}{}
-\floatname{listing}{Listing}
-%
-% Keep the \pdfinfo as shown here. There's no need
-% for you to add the /Title and /Author tags.
-\pdfinfo{
-/TemplateVersion (2024.1)
-}
+% if you need to pass options to natbib, use, e.g.:
+%     \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading neurips_2024
 
-% DISALLOWED PACKAGES
-% \usepackage{authblk} -- This package is specifically forbidden
-% \usepackage{balance} -- This package is specifically forbidden
-% \usepackage{color (if used in text)
-% \usepackage{CJK} -- This package is specifically forbidden
-% \usepackage{float} -- This package is specifically forbidden
-% \usepackage{flushend} -- This package is specifically forbidden
-% \usepackage{fontenc} -- This package is specifically forbidden
-% \usepackage{fullpage} -- This package is specifically forbidden
-% \usepackage{geometry} -- This package is specifically forbidden
-% \usepackage{grffile} -- This package is specifically forbidden
-% \usepackage{hyperref} -- This package is specifically forbidden
-% \usepackage{navigator} -- This package is specifically forbidden
-% (or any other package that embeds links such as navigator or hyperref)
-% \indentfirst} -- This package is specifically forbidden
-% \layout} -- This package is specifically forbidden
-% \multicol} -- This package is specifically forbidden
-% \nameref} -- This package is specifically forbidden
-% \usepackage{savetrees} -- This package is specifically forbidden
-% \usepackage{setspace} -- This package is specifically forbidden
-% \usepackage{stfloats} -- This package is specifically forbidden
-% \usepackage{tabu} -- This package is specifically forbidden
-% \usepackage{titlesec} -- This package is specifically forbidden
-% \usepackage{tocbibind} -- This package is specifically forbidden
-% \usepackage{ulem} -- This package is specifically forbidden
-% \usepackage{wrapfig} -- This package is specifically forbidden
-% DISALLOWED COMMANDS
-% \nocopyright -- Your paper will not be published if you use this command
-% \addtolength -- This command may not be used
-% \balance -- This command may not be used
-% \baselinestretch -- Your paper will not be published if you use this command
-% \clearpage -- No page breaks of any kind may be used for the final version of your paper
-% \columnsep -- This command may not be used
-% \newpage -- No page breaks of any kind may be used for the final version of your paper
-% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
-% \pagestyle -- This command may not be used
-% \tiny -- This is not an acceptable font size.
-% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
-% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference
+% ready for submission
+\usepackage{neurips_2024}
 
-\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.
+% to compile a preprint version, e.g., for submission to arXiv, add add the
+% [preprint] option:
+%     \usepackage[preprint]{neurips_2024}
 
+% to compile a camera-ready version, add the [final] option, e.g.:
+%     \usepackage[final]{neurips_2024}
+
+% to avoid loading the natbib package, add option nonatbib:
+%    \usepackage[nonatbib]{neurips_2024}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{xcolor}         % colors
+
+% Packages added by me.
 \usepackage{amsmath}
@@
 \usepackage{subcaption}
-% \usepackage[ruled]{algorithm2e}
+\usepackage[ruled]{algorithm2e}
 
@@
 
-% The file aaai24.sty is the style file for AAAI Press
-% proceedings, working notes, and technical reports.
+\title{Title}
+
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
 %
+% Using \And between authors leaves it to LaTeX to determine where to break the
+% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
+% authors names on the first line, and the last on the second line, try using
+% \AND instead of \And before the third author name.
 
-% Title
-
-% Your title must be in mixed case, not sentence case.
-% That means all verbs (including short verbs like be, is, using,and go),
-% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
-% articles, conjunctions, and prepositions are lower case unless they
-% directly follow a colon or long dash
-\title{Title}
-\author{
-    %Authors
-    % All authors must be in the same font size and format.
-    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
-    AAAI Style Contributions by Pater Patel Schneider,
-    Sunil Issar,\\
-    J. Scott Penberthy,
-    George Ferguson,
-    Hans Guesgen,
-    Francisco Cruz\equalcontrib,
-    Marc Pujol-Gonzalez\equalcontrib
+\author{%
+  David S.~Hippocampus\thanks{Use footnote for providing further information
+    about author (webpage, alternative address)---\emph{not} for acknowledging
+    funding agencies.} \\
+  Department of Computer Science\\
+  Cranberry-Lemon University\\
+  Pittsburgh, PA 15213 \\
+  \texttt{hippo@cs.cranberry-lemon.edu} \\
+  % examples of more authors
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \AND
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
 }
-\affiliations{
-    %Afiliations
-    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
-    % If you have multiple authors and multiple affiliations
-    % use superscripts in text and roman font to identify them.
-    % For example,
-
-    % Sunil Issar\textsuperscript{\rm 2},
-    % J. Scott Penberthy\textsuperscript{\rm 3},
-    % George Ferguson\textsuperscript{\rm 4},
-    % Hans Guesgen\textsuperscript{\rm 5}
-    % Note that the comma should be placed after the superscript
-
-    1900 Embarcadero Road, Suite 101\\
-    Palo Alto, California 94303-3310 USA\\
-    % email address must be in roman text type, not monospace or sans serif
-    proceedings-questions@aaai.org
-%
-% See more examples next
-}
-
-%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
-\iffalse
-\title{My Publication Title --- Single Author}
-\author {
-    Author Name
-}
-\affiliations{
-    Affiliation\\
-    Affiliation Line 2\\
-    name@example.com
-}
-\fi
-
-\iffalse
-%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
-\title{My Publication Title --- Multiple Authors}
-\author {
-    % Authors
-    First Author Name\textsuperscript{\rm 1},
-    Second Author Name\textsuperscript{\rm 2},
-    Third Author Name\textsuperscript{\rm 1}
-}
-\affiliations {
-    % Affiliations
-    \textsuperscript{\rm 1}Affiliation 1\\
-    \textsuperscript{\rm 2}Affiliation 2\\
-    firstAuthor@affiliation1.com, secondAuthor@affilation2.com, thirdAuthor@affiliation1.com
-}
-\fi
 
@@
 \begin{abstract}
-	The initial step in a causal analysis involves determining the causal
-	structure between the variables from data typically in the form of a
-	Directed Acyclic Graph (DAG) or an Structural Equation Model (SEM).
-	This procedure, known as \emph{causal discovery}, has been extensively
-	studied for both DAGs and SEMs. While numerous automated algorithms
-	exist for estimating DAGs from datasets, their application in applied
-	fields remain limited, possibly due to a lack of confidence in these
-	algorithms stemming from them making obvious mistakes, and the
-	difficulties in choosing the right algorithm for a given dataset.
-	Consequently, in applied disciplines, these DAGs are predominantly
-	constructed manually based on domain expertise. Due to being
-	constructed by hand, it is vital that the fit of these models are
-	tested against data and the models are potentially modified to acheive
-	a better fit. On the other hand, in the field of SEMs models are
-	typically constructed manually through an iterative process of
+	In Pearl's framework, the initial step for performing a causal analysis
+	involves determining the causal structure between the variables from
+	data typically in the form of a Directed Acyclic Graph (DAG) or an
+	Structural Equation Model (SEM). This procedure, known as \emph{causal
+	discovery}, has been extensively studied for both DAGs and SEMs. While
+	numerous automated algorithms exist for estimating DAGs from datasets,
+	their application in applied fields remain limited, possibly due to a
+	lack of confidence in these algorithms stemming from them making
+	obvious mistakes, and the difficulties in choosing the right algorithm
+	for a given dataset. Consequently, in applied disciplines, these DAGs
+	are predominantly constructed manually based on domain expertise. Due
+	to being constructed by hand, it is vital that the fit of these models
+	are tested against data and the models are potentially modified to
+	acheive a better fit. On the other hand, in the field of SEMs models
+	are typically constructed manually through an iterative process of
 	assessing fit and modifying the model according to that. This process
@@
 
-\emph{Causal discovery} is the process of identifying and understanding causal
-relationships among variables in a dataset.
-
 To perform any causal effect identification or estimation using Directed
 Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
-first construct a causal structure. This process of constructing the casual
-structure from data is known as \emph{causal discovery}. In the field of DAGs,
-a lot of effort has been put on developing automated algorithms that can
-construct these DAGs from data. These algorithm takes different approaches to
-causal discovery such as Constraint-based algorithms such as PC and Fast Causal
-Infenrece (FCI) where the algorithm tries to match the conditional
-independences in the data to the ones impled by the model, score-based methods
-such as Hill-Climb Search, Greedy Equivalence Search (GES)  where the algorithm
-tries to find a model that has the best score given some scoring metric, and
-continuous optimization where the problem is formulated as a constrained
-optimization problem such as No Tears.
+first construct a causal structure commonly known as DAG or path diagram. This
+process of constructing the casual structure from data is known as \emph{Causal
+Discovery}. In the field of DAGs, plenty of algorithms have been developed to
+automatically construct DAGs from data. These algorithm takes different
+approaches to causal discovery such as Constraint-based algorithms such as PC
+and Fast Causal Infenrece (FCI) where the algorithm tries to match the
+conditional independences in the data to the ones impled by the model,
+score-based methods such as Hill-Climb Search, Greedy Equivalence Search (GES)
+where the algorithm tries to find a model that has the best score given some
+scoring metric, and continuous optimization where the problem is formulated as
+a constrained optimization problem such as No Tears.
 
@@
 \begin{enumerate}
+	\item Can only give a Markov Equivalence class (MEC), and then experts
+		need to rely on their domain expertise to orient these edges.
 	\item Biased towards sparse graphs.
@@
 
+\todo[inline]{Other methods: Modification indices and Lagrange Multiplier tests to
+selectively add edges And using Wald test or z statistics (also known as
+critical ratios) for removing edges.}
+
+One of the drawbacks of these global fit metrics is that they do not only test
+for the structure but also the fitted parameterization.
+
 We present a similar procedure for modifying DAGs based on expert knowledge and
 measure of association. We first use the measure of association to test whether
-the association observed in the data is explained by the model structure. Based 
+the association observed in the data is explained by the model structure. 
+We can not directly use the local CI tests to make modifications because that would give
+way too many independencies and we have no way to rank them.
+Based 
 on this testing, we suggest adding edges by ranking them by the most unexplained
@@
 		\begin{tikzpicture}
-			\begin{scope}[xshift=-2.5cm, yshift=0.7cm, scale=0.7]
+			\begin{scope}[yshift=0.7cm, scale=1]
 			\tikzstyle{every node}=[align=center, inner sep=1pt]
@@
 			\end{scope}
-			\begin{scope}[xshift=-3cm]
 			\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
-				\node[rectangle, align=center, inner sep=1pt] at (5, 0) {
-					\begin{minipage}{.2\textwidth}
+			\node[rectangle, align=center, inner sep=1pt] at (6, 0) {
+				\begin{minipage}{\textwidth}
 					\begin{equation*}
@@
 				};
-			\end{scope}
 		\end{tikzpicture}
@@
 
-% \begin{algorithm}
-% 	\caption{Expert Simulation}
-% 	\While{True}{
-% 	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $\;
-% 	   $ e \gets e - \textnormal{blacklist} $\;
-% 	   \If {$ e = 0 $}{
-% 		$ G \gets G - e $ \;
-% 		$ \textnormal{blacklist} \gets \textnormal{blacklist} \cup e $\;
-% 	   } \;
-% 	   \If{$ \textnormal{len}(e) = 0$}{
-% 		   $ \textnormal{return}(G) $\;
-% 	   }
-% 	   $ X, Y \gets \max_{X, Y} e $\;
-% 	   $ dir \gets Expert(X, Y, \alpha) $\;
-% 	   $ G \gets G \cup dir $\;
-% 	}
-% 
-% \end{algorithm}
+\begin{algorithm}
+	\caption{Expert Simulation}
+	\While{True}{
+	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $\;
+	   $ e \gets e - \textnormal{blacklist} $\;
+	   \If {$ e = 0 $}{
+		$ G \gets G - e $ \;
+		$ \textnormal{blacklist} \gets \textnormal{blacklist} \cup e $\;
+	   } \;
+	   \If{$ \textnormal{len}(e) = 0$}{
+		   $ \textnormal{return}(G) $\;
+	   }
+	   $ X, Y \gets \max_{X, Y} e $\;
+	   $ dir \gets Expert(X, Y, \alpha) $\;
+	   $ G \gets G \cup dir $\;
+	}
+
+\end{algorithm}
 
@@
 		\includegraphics{../code/plots/shd_ribbon.pdf}
-	\end{subfigure}
+	\end{subfigure}%
 	\begin{subfigure}{0.5\textwidth}
@@
 
-\bibliography{references}
+\bibliographystyle{plain}
+\bibliography{references.bib}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\appendix
+
+\section{Appendix / supplemental material}
+
+Optionally include supplemental material (complete proofs, additional experiments and plots) in appendix.
+All such materials \textbf{SHOULD be included in the main submission.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\include{checklist}
+
 \end{document}

### 2024-06-09 paper/main.tex

--- 2024-06-08 paper/main.tex
@@
 
-\emph{Causal discovery} is the process of identifying and understanding causal
-relationships among variables in a dataset.  To perform any causal effect
-estimation using Directed Acyclic Graphs (DAGs) or Structural Equation Models
-(SEMs), researchers need to first construct a causal structure. In the field of
-DAGs, a lot of attention has been put on developing automated algorithms that
-can construct these DAGs from data. These algorithm takes different approaches
-to causal discovery such as constraint-based algorithms such as PC and Fast
-Causal Infenrece (FCI) where the algorithm tries to match the conditional
-independences in the data to the ones impled by the model, score-based methods
-such as Hill-Climb Search, Greedy Equivalence Search (GES)  where the algorithm
-tries to find a model that has the best score given some scoring metric, and
-continuous optimization where the problem is formulated as a constrained
-optimization problem such as No Tears, and some hybrid approaches.
+To perform any causal effect identification or estimation using Directed
+Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
+first construct a causal structure commonly known as DAG or path diagram. This
+process of constructing the casual structure from data is known as \emph{Causal
+Discovery}. In the field of DAGs, plenty of algorithms have been developed to
+automatically construct DAGs from data. These algorithm takes different
+approaches to causal discovery such as Constraint-based algorithms such as PC
+and Fast Causal Infenrece (FCI) where the algorithm tries to match the
+conditional independences in the data to the ones impled by the model,
+score-based methods such as Hill-Climb Search, Greedy Equivalence Search (GES)
+where the algorithm tries to find a model that has the best score given some
+scoring metric, and continuous optimization where the problem is formulated as
+a constrained optimization problem such as No Tears.
 
@@
 
+In the field of Structural Equation Models (SEMs), models are also typically
+constructed by hand. Researchers typically come up with an initial model, which
+then they use the data and expert knowledge to make modifications to. This
+process is known as \emph{Specification Search}. One of the common ways to do
+this is through \emph{modification indices} where a global fit of the model is
+measured with different modification to the model.  Utilizing expert knowledge
+to select the appropriate modification that gives the maximum improvement to
+the global fit.
+
+\todo[inline]{Other methods: Modification indices and Lagrange Multiplier tests to
+selectively add edges And using Wald test or z statistics (also known as
+critical ratios) for removing edges.}
+
+One of the drawbacks of these global fit metrics is that they do not only test
+for the structure but also the fitted parameterization.
+
 We present a similar procedure for modifying DAGs based on expert knowledge and
 measure of association. We first use the measure of association to test whether
-the association observed in the data is explained by the model structure. Based 
+the association observed in the data is explained by the model structure. 
+We can not directly use the local CI tests to make modifications because that would give
+way too many independencies and we have no way to rank them.
+Based 
 on this testing, we suggest adding edges by ranking them by the most unexplained

### 2024-06-09 paper/main.tex
Updates

@@
 
+A lot of questions in science center around a causal understanding between 
+variables. \emph{Causal Discovery} methods aim to find these casual structure
+among variables from a given dataset.
+
 To perform any causal effect identification or estimation using Directed
@@
 
-Even though we have various algorithms for casual discovery, the adaption of these
+Even though we have many algorithms for casual discovery, the adaption of these
 algorithms in applied fields remain limited. In these applied fields, researchers 
@@
 However, this does not give us a way to make modifications to our model. A
-possible way to do this by using check which of these local tests are violated
-based on the
-
-In the cases when the DAG is an output of an algorithm or when constructed
-by hand, it is vital to test these models against data and potentially modifying 
-them for a better fit before using them for making any inference. Both global
-and local tests for testing the fit of these models are available. However, when
-the fit fails, we don't have a procedure to modify these models.
-
-In the field of Structural Equation Models (SEMs), models are also typically
-constructed by hand. Researchers typically come up with an initial model, which
-then they use the data and expert knowledge to make modifications to. This
-process is known as \emph{Specification Search}. One of the common ways to do
-this is through \emph{modification indices} where a global fit of the model is
-measured with different modification to the model.  Utilizing expert knowledge
-to select the appropriate modification that gives the maximum improvement to
-the global fit.
+possible way to do this by checking which of these local tests are violated.
+However, this approach can quickly become diffult as the size of the network
+increases. For example, the popular alarm network implies $x$ condition independence
+statements.
 
@@
 measure of association. We first use the measure of association to test whether
-the association observed in the data is explained by the model structure. 
-We can not directly use the local CI tests to make modifications because that would give
-way too many independencies and we have no way to rank them.
-Based 
-on this testing, we suggest adding edges by ranking them by the most unexplained
-association between variables. The researcher can then use expert knowledge to 
-decide the direction of the edge between the variables.
+the association observed in the data is explained by the model structure. We
+can not directly use the local CI tests to make modifications because that
+would give way too many independencies and we have no way to rank them. Based
+on this testing, we suggest adding edges by ranking them by the most
+unexplained association between variables. The researcher can then use expert
+knowledge to decide the direction of the edge between the variables.
 
@@
 		canonical correlations. This mixed variable association is a generalization 
-		of partial correlation to mixed variables.
-	\item We present a procedure to iteratively construct and modify DAGs using expert 
-		knowledge such that it improves the explained association between variables.
+		of some of the other association measures such as partial correlation to 
+		mixed variables.
+	\item We present a procedure to use this measure of association to
+		iteratively modify DAGs using expert knowledge such that it
+		improves the explained association between variables.
 	\item Lastly, to assist researchers in manully constructing and modifying DAGs,
@@
 \subsection{Measures of Association}
-In this section, we give a background on some of the commonly used effect size
-measures in the context of CI testing.
+As our DAG construction procedure relies on a measure of conditional
+association between variables, in this section we list out some of the already
+known measures for specific data types. These measure of associations are
+typically the effect size of various conditional indepdnence tests but they
+don't need to be. However, there are no generalized measure of association for
+mixed data, that we propose in the next section. This is by no means an
+exhaustive list, rather just to give an idea of what can be done.
 
-\paragraph{$ X $ and $ Y $ are continuous}
-In the case when $ \bm{Z} = \emptyset $, and when both $ X $ and $ Y $ are
-continuous variables, Pearson's correlation coefficient based test can be used
-which also gives an effect size measure. The p-value computation requires the
-data to be normally distributed but other statistics can be used such as
-Spearman's rank correlation, or Kendall's Tau.
+\paragraph{Both $ X $ and $ Y $ are continuous}
+Where there are no conditional variables, $ \bm{Z} = \emptyset $, Pearson's
+correlation coefficient can be used as a measure of association between the
+variables. This does make various assumptions such as linearity between the
+variables, normal distribution for $ X $ and $ Y $, and homoscedasticity.
 
-In the case when $ \bm{Z} \neq \emptyset $, then partial correlations can be used.
-In this case we first compute the residuals of $ X $ and $ Y $ using $ \bm{Z} $ as 
-the predictor. We start by training two model $ X \sim \bm{Z} $ and $ Y \sim \bm{Z} $.
-Then the residuals are computed for both as $ R_X = X - \hat{X} $ and $ R_Y = Y - \hat{Y} $.
-After this we can do a normal non-conditional test on $ R_X $ and $ R_Y $.
+In the case when $ \bm{Z} \neq \emptyset $, then partial correlations can be
+used. This requires using two linear regression models: $ E_X: X \sim \bm{Z} $
+and $ E_Y: Y \sim \bim{Z} $. Then computing the residuals using these two
+regressions as: $ R_X = X - E_X(\bm{Z}) $ and $ R_Y = Y - E_Y(\bm{Y}) $. After
+this a simple correlation coefficient can be computed between $ R_X $ and $ R_Y $.
 
-\paragraph{All variables are discrete}
+\paragraph{All $ X $, $ Y $, and $ \bm{Z} $ are discrete}
 
-When $ \bm{Z} = \emptyset $, in the case of discrete variables, chi-squared test can be used. Effect size 
-measures such as Cram\'er's V are available for it.
+In the case of discrete variables, Cramer\'s V has been typically used as a
+measure of association. When there are no conditional variables, $ \bm{Z} = \emptyset $,
+Cramer\'s V can be computed directly from the contingency table.
 
-When $ \bm{Z} \neq \emptyset $, we can do stratification.
+However, when conditional variables are present, we can iterate over all
+possible combinations of the conditional variables, compute the Cramer\'s V and
+then lastly combine them together.
+
+\todo[inline]{Is it even possible}
 
@@
 
-\paragraph{Mixed data}
-Recently multiple tests for mixed data have been proposed but they do not have
-an effect size measure. We propose one based on canonical correlations.
+\section{Measure of Association for mixed data}
+In this section, we use canonical correlations to define an effect size measure
+for mixed data CI tests. Our approach is similar to the approach used for the
+continuous variable case.
 
@@
 We use Pillai's Trace in this paper. \todo[inline]{Why?}.
-
-\section{Measure of Association for mixed data}
-In this section, we use canonical correlations to define an effect size measure
-for mixed data CI tests. Our approach is similar to the approach used for the
-continuous variable case.
 

### 2024-06-09 paper/main.tex
Updates

@@
 \subsection{Conditional Case}
+For the conditional case we use an approach similar to partial correlation. We
+combine the mixed data residualization approach proposed in Ankan 2023. We then
+use a canonical correlation based measure on these residuals to get an effect
+size.
 
 \subsection{Connection to special cases}
+This natually is a generalization of partial correlations. Furthermore, it is 
+also a generalization of discrete case.
+
+\section{Using measure of association for modifying DAGs}
+The idea behind using the measure of association for modifying DAGs is that,
+for any given model and a dataset, we can use the condtiional measure of
+association to compute the amount of correlation in the data that is still not
+explained by the model. Based on this, we can compute how much improvement in
+the explanation would happen if we were to add an edge between variables. This
+can then be ranked using the maximum improvement.
+
+Theorem - If we have access to an oracle that can give all correct results, 
+the algorithm above is consistent.
 
 \section{Web Tool}
-As most of the DAGs in practice are built manually, we use our mixed data effect size measure to build a web-tool that aids researches in building DAGs.
+Based on the approach outlined in the previous section, we built a web-based
+tool that researchers can upload their dataset to and apply this method to
+interactively draw models.
+
+\todo[inline]{List some of the features of the online platform}
 

### 2024-06-09 paper/main.tex
Updates

@@
 the explanation would happen if we were to add an edge between variables. This
-can then be ranked using the maximum improvement.
+can then be ranked using the maximum improvement. Give some examples here.
 
@@
 \section{Empirical Analysis}
+
 To compare how well does creating a graph manually based on effect sizes and

### 2024-06-10 paper/main.tex
Updates

@@
 
-To compare how well does creating a graph manually based on effect sizes and
-p-value compare to algorithmic approaches, we performed some empirical
-analysis.
-
-\subsection{Simulating an expert}
-To simulate an expert building a model manually, we wrote a simple greedy simulation as shown in the Figure. The expert always starts with a graph with
-no edges and chooses the pair of variables with the highest correlation between them. The expert based on some accuracy measure decides the direction
-of the edge between these variables such that the new added edge does not create a cycle in the DAG.
-
-\todo[inline]{Summarize this section and equation in an algorithm}
-\todo[inline]{Show a consistency proof for this expert simulation}
+To compare how well this manual appraoch assisted by the suggested modications
+perform compared to automated learning algorithms, we performed an analysis
+using simulated data and simulating an expert. To construct the model using
+suggestions, we start with a model with no edges, and take a greedy approach
+for adding edges. At any point, we select to add an edge between variables that
+have the highest conditional association between them. A simulated expert is
+then asked to specify the direction for the edge. Given an accuracy of the
+expert $ \alpha $, the expert either gives the correct direction, incorrect
+direction, or can specify do not know as:
 
@@
 
-% \begin{algorithm}
-% 	\caption{Expert Simulation}
-% 	\While{True}{
-% 	   $ e \gets \textnormal{Effects}(G) > \textnormal{THRESHOLD} $\;
-% 	   $ e \gets e - \textnormal{blacklist} $\;
-% 	   \If {$ e = 0 $}{
-% 		$ G \gets G - e $ \;
-% 		$ \textnormal{blacklist} \gets \textnormal{blacklist} \cup e $\;
-% 	   } \;
-% 	   \If{$ \textnormal{len}(e) = 0$}{
-% 		   $ \textnormal{return}(G) $\;
-% 	   }
-% 	   $ X, Y \gets \max_{X, Y} e $\;
-% 	   $ dir \gets Expert(X, Y, \alpha) $\;
-% 	   $ G \gets G \cup dir $\;
-% 	}
-% 
-% \end{algorithm}
-
-After adding each edge, we also check if the p-value of any existing edge shows independence. If that happens we remove that edge and blacklist that edge. 
-Blacklisted edges are not suggested to the expert in future iterations. We repeat this till we have explained all the correlations in the dataset.
-
-There is a possibility that this greedy expert gets stuck at incorrect graph structures as it does not do any backtracking to fix its earlier mistakes.
+If the expert says `do not know` for any of the edges, the edge is not prompted
+again to the expert. We also make sure that only edges which do not form a
+cyclic in the graph are used to select the next potential edge. After adding
+each edge, we also check if the p-value of any existing edge shows
+independence. If that happens we remove that edge and blacklist that edge,
+i.e., that edge is not prompted again during the run of the algorithm. We
+repeat this procedure till the model explains all correlations between the
+variables. There is a possibility that this greedy expert gets stuck at
+incorrect graph structures as it does not do any backtracking to fix its
+earlier mistakes. This procedure is outlined in the Algorithm.
+\todo[inline]{Add the algorithm}
 

### 2024-06-10 paper/main.tex
Updates

@@
 used. This requires using two linear regression models: $ E_X: X \sim \bm{Z} $
-and $ E_Y: Y \sim \bim{Z} $. Then computing the residuals using these two
+and $ E_Y: Y \sim \bm{Z} $. Then computing the residuals using these two
 regressions as: $ R_X = X - E_X(\bm{Z}) $ and $ R_Y = Y - E_Y(\bm{Y}) $. After
@@
 
-To compare how well this manual appraoch assisted by the suggested modications
-perform compared to automated learning algorithms, we performed an analysis
-using simulated data and simulating an expert. To construct the model using
-suggestions, we start with a model with no edges, and take a greedy approach
-for adding edges. At any point, we select to add an edge between variables that
-have the highest conditional association between them. A simulated expert is
-then asked to specify the direction for the edge. Given an accuracy of the
-expert $ \alpha $, the expert either gives the correct direction, incorrect
-direction, or can specify do not know as:
+We analyzed the performance of this manual approach by comparing it with two
+automated algorithms: PC and Hill-Climb Search. For the comparison we use
+simulated data from a known DAG and compare how well the algorithm is able to
+recover the original model using two metrics: Structural Hamming Distance (SHD)
+and Structural Intervention Distance (SID). To simulate the data, we start with
+a randomly generated DAG and use linear models with random effects to generate
+the data. We repeat the data generation $ 30 $ times for each density value for
+the DAG. As both PC and Hill-Climb Search can only give us a CPDAG, we compare
+the orientation of these CPDAG which result in the best and worst case scores
+on SHD and SID. This procedure is outlined in Algorithm 1.
+
+\todo[inline]{Add the algorithm}
+
+To simulate an expert using this assisted model construction approach, we start
+with an empty model and take a greedy approach to add edges to this model. At
+any point, we select the pair of variables that has significant correlation and
+the highest unexplained correlation, i.e., highest conditional association. We
+then use an oracle to decide the direction of the edge between the variables.
+Given an accuracy of the oracle, $ \alpha $, the oracle either gives the
+correct direction, incorrect direction, or returns None representing that it
+does not know the direction.
 
@@
 
-If the expert says `do not know` for any of the edges, the edge is not prompted
-again to the expert. We also make sure that only edges which do not form a
-cyclic in the graph are used to select the next potential edge. After adding
-each edge, we also check if the p-value of any existing edge shows
-independence. If that happens we remove that edge and blacklist that edge,
-i.e., that edge is not prompted again during the run of the algorithm. We
-repeat this procedure till the model explains all correlations between the
-variables. There is a possibility that this greedy expert gets stuck at
-incorrect graph structures as it does not do any backtracking to fix its
-earlier mistakes. This procedure is outlined in the Algorithm.
+If the oracle returns None for any pair of variables, an edge between this pair
+is not suggested to the oracle in future iterations. We also make sure that
+only edges which do not form a cyclic in the graph are used to select the next
+potential edge. After adding each edge, we also check if the p-value of any
+existing edge shows independence. If that happens we remove that edge and
+blacklist that edge, i.e., that edge is not prompted again during the run of
+the algorithm. We repeat this procedure till the model explains all
+correlations between the variables. There is a possibility that this greedy
+expert gets stuck at incorrect graph structures as it does not do any
+backtracking to fix its earlier mistakes. This procedure is outlined in the
+Algorithm.
 \todo[inline]{Add the algorithm}
@@
 	\begin{subfigure}{0.5\textwidth}
+		\centering
 		\includegraphics{../code/plots/shd_ribbon.pdf}
+		\caption{}
 	\end{subfigure}
 	\begin{subfigure}{0.5\textwidth}
+		\centering
 		\includegraphics{../code/plots/sid_ribbon.pdf}
+		\caption{}
 	\end{subfigure}
+	\caption{Comparison of PC and Hill-Climb Search algorithms against }
 \end{figure}

### 2024-06-10 code/sl.R

@@ -105,7 +105,7 @@ simulate_human_sl <- function(sim_data, true_dag, oracle_acc, max_iter=1e4){
 			new_dag <- remove_edges(new_dag, edges_to_prune)
 		}
 		
-		# Recompute effects after pruning: TODO: Do we need to do this?
+		# Recompute effects after pruning. We don't really need to do this.
 		effects <- compute_effects_v2(new_dag, sim_data) %>% filter(p < PVALUE_THRESHOLD)
 		effects <- effects[, c('X', 'A', 'Y', 'cor')]
 		unexplain_cor_sorted <- effects %>% filter(abs(as.double(effects$cor)) > EFFECT_THRESHOLD) %>% filter(A == '--') %>% arrange(desc(abs(cor)))

### 2024-06-11 paper/main.tex
Updates intro

@@
 \begin{abstract}
-	In Pearl's framework, the initial step for performing a causal analysis
-	involves determining the causal structure between the variables from
-	data typically in the form of a Directed Acyclic Graph (DAG) or an
-	Structural Equation Model (SEM). This procedure, known as \emph{causal
-	discovery}, has been extensively studied for both DAGs and SEMs. While
-	numerous automated algorithms exist for estimating DAGs from datasets,
-	their application in applied fields remain limited, possibly due to a
-	lack of confidence in these algorithms stemming from them making
-	obvious mistakes, and the difficulties in choosing the right algorithm
-	for a given dataset. Consequently, in applied disciplines, these DAGs
-	are predominantly constructed manually based on domain expertise. Due
-	to being constructed by hand, it is vital that the fit of these models
-	are tested against data and the models are potentially modified to
-	acheive a better fit. On the other hand, in the field of SEMs models
-	are typically constructed manually through an iterative process of
-	assessing fit and modifying the model according to that. This process
-	is commonly known as \emph{Specification Search} where methods like
-	\emph{modification indices} combined with expert knowledge can be used
-	to guide this modification process. In the case of DAGs, although there
-	are tests to assess the global fit of a model by combining the tests
-	using implied conditional independencies of the model, but there are no
-	methods that can guide our modification process. In this paper, we
-	present a simple modification process aimed at helping researchers in
-	manually constructing or modifying DAGs outputted by an algorithm. This
-	modification process is based on using a (conditional) measure of
-	association between variables in the model to assess potential edges
-	that best explains unexplained correlation between variables which the
-	researcher specifies the direction of this edge. In the case of
-	continuous or discrete variables, the effect size measures of local
-	tests can be used as this measure of association, however for mixed
-	data there is no such measure. We present a measure of association
-	based on canonical correlations for mixed data. We also present a
-	graphical web-tool that can assist researchers in iteratively modifying
-	and constructing their DAGs. We theoretically show that in the presence
-	of an oracle, this iterative modification process is consistent in
-	recovering the true DAG. Emprically, we show that using the mixed data
-	measure of association, the modification process performs comparably to
-	PC and Hill-Climb Search algorithms if the user is able to correctly
-	identify the direction of the edge in one out of three cases.
+	% TODO: Needs to be shortened.
+	% In Pearl's framework, the initial step for performing a causal analysis
+	% involves determining the causal structure between the variables from
+	% data typically in the form of a Directed Acyclic Graph (DAG) or an
+	% Structural Equation Model (SEM). This procedure, known as \emph{causal
+	% discovery}, has been extensively studied for both DAGs and SEMs. While
+	% numerous automated algorithms exist for estimating DAGs from datasets,
+	% their application in applied fields remain limited, possibly due to a
+	% lack of confidence in these algorithms stemming from them making
+	% obvious mistakes, and the difficulties in choosing the right algorithm
+	% for a given dataset. Consequently, in applied disciplines, these DAGs
+	% are predominantly constructed manually based on domain expertise. Due
+	% to being constructed by hand, it is vital that the fit of these models
+	% are tested against data and the models are potentially modified to
+	% acheive a better fit. On the other hand, in the field of SEMs models
+	% are typically constructed manually through an iterative process of
+	% assessing fit and modifying the model according to that. This process
+	% is commonly known as \emph{Specification Search} where methods like
+	% \emph{modification indices} combined with expert knowledge can be used
+	% to guide this modification process. In the case of DAGs, although there
+	% are tests to assess the global fit of a model by combining the tests
+	% using implied conditional independencies of the model, but there are no
+	% methods that can guide our modification process. In this paper, we
+	% present a simple modification process aimed at helping researchers in
+	% manually constructing or modifying DAGs outputted by an algorithm. This
+	% modification process is based on using a (conditional) measure of
+	% association between variables in the model to assess potential edges
+	% that best explains unexplained correlation between variables which the
+	% researcher specifies the direction of this edge. In the case of
+	% continuous or discrete variables, the effect size measures of local
+	% tests can be used as this measure of association, however for mixed
+	% data there is no such measure. We present a measure of association
+	% based on canonical correlations for mixed data. We also present a
+	% graphical web-tool that can assist researchers in iteratively modifying
+	% and constructing their DAGs. We theoretically show that in the presence
+	% of an oracle, this iterative modification process is consistent in
+	% recovering the true DAG. Emprically, we show that using the mixed data
+	% measure of association, the modification process performs comparably to
+	% PC and Hill-Climb Search algorithms if the user is able to correctly
+	% identify the direction of the edge in one out of three cases.
 \end{abstract}
@@
 
-A lot of questions in science center around a causal understanding between 
-variables. \emph{Causal Discovery} methods aim to find these casual structure
-among variables from a given dataset.
+A lot of questions in life sciences center around a causal understanding
+between variables. \emph{Causal Discovery} aims to find these casual structure
+among random variables from a given dataset. The problem of causal discovery
+has been studied both the in Directed Acyclic Graph (DAGs) and Structure
+Equation Models (SEM) literatures. In the field of DAGs, many automated
+algorithms have been developed to learn the casual structure form a given 
+dataset such as constraint-based methods, score-based methods, and continuous
+optimization based methods.
 
-To perform any causal effect identification or estimation using Directed
-Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
-first construct a causal structure commonly known as DAG or path diagram. This
-process of constructing the casual structure from data is known as \emph{Causal
-Discovery}. In the field of DAGs, plenty of algorithms have been developed to
-automatically construct DAGs from data. These algorithm takes different
-approaches to causal discovery such as Constraint-based algorithms such as PC
-and Fast Causal Infenrece (FCI) where the algorithm tries to match the
-conditional independences in the data to the ones impled by the model,
-score-based methods such as Hill-Climb Search, Greedy Equivalence Search (GES)
-where the algorithm tries to find a model that has the best score given some
-scoring metric, and continuous optimization where the problem is formulated as
-a constrained optimization problem such as No Tears.
+% To perform any causal effect identification or estimation using Directed
+% Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
+% first construct a causal structure commonly known as DAG or path diagram. This
+% process of constructing the casual structure from data is known as \emph{Causal
+% Discovery}. In the field of DAGs, plenty of algorithms have been developed to
+% automatically construct DAGs from data. These algorithm takes different
+% approaches to causal discovery such as Constraint-based algorithms such as PC
+% and Fast Causal Infenrece (FCI) where the algorithm tries to match the
+% conditional independences in the data to the ones impled by the model,
+% score-based methods such as Hill-Climb Search, Greedy Equivalence Search (GES)
+% where the algorithm tries to find a model that has the best score given some
+% scoring metric, and continuous optimization where the problem is formulated as
+% a constrained optimization problem such as No Tears.
 
-Even though we have many algorithms for casual discovery, the adaption of these
-algorithms in applied fields remain limited. In these applied fields, researchers 
-typically prefer to construct these DAGs manually based on their domain knowledge.
-We hypothesize that this preference towards manual DAG construction stems from the 
-following difficulties in applying these methods to real dataset.
+However, the adaption of these algorithms in applied fields remain extremely
+limited. In these applied fields, even though these algorithms exist,
+researchers typically prefer to construct these DAGs manually based on their
+domain knowledge. We hypothesize that this preference towards manual DAG
+construction stems from the following difficulties in applying causal discovery
+methods:
 
 \begin{enumerate}
-	\item Algorithms output Markov Equivalence Class (MEC): Using observational
-		data, any causal discovery algorithm can only output the MEC.
-		MEC do not have all the edges oriented and to do further
-		hypothesis testing or causal effect estimation, researchers
-		typically have to rely on domain expertise to orient the
-		remaining edges in the MEC.
-	\item Difficulties in choosing an algorithm: To use any of these algorithms on 
-		any practical dataset, users need to make a lot of choices. For
-		example, for using constraint-based method, users need to
-		choose a suitable condtional independence (CI) test, and
-		hyperparameters for these tests. Similarly for score-based
-		methods users need to choose the scoring metrics and their
-		hyperparameters. Moreover, for practical implementation with
-		reasonable runtime, users also need to choose the algorithm
-		hyperparameters such as maximum number of conditioning
-		variables to consider, max indegree of variables in the model. With
-		no easy guides or method to evaluate these choices, it can be difficult.
-	\item Lack of trust: Most of these automated algorithms can make very obvious 
-		mistakes. Typically users have to test the outputs of the models and 
-		manually make corrections to the output to get a usable model.
-		This can lead to researchers not trusting these algorithms.
+	\item Difficult to trust: Although asypotically these methods are 
+		consistent, in the finite sample scenario they can end up
+		making obvious mistakes. As a result of this, researchers
+		have to rely on posthoc modification of the output of these
+		algorithms using either their domain knowledge or assisted
+		by testing methods.
+	\item Algorithms output the Markov Equivalence Class (MEC): With any
+		causal discovery algorithm, we can only get to the MEC. However,
+		most of the methods developed for downstream tasks such 
+		as identification or casual effect estimation assume the
+		knowledge of a DAG. To orient these MECs to a DAG, researchers
+		have to either rely on their domain expertise, or make further
+		assumption to utilize a pairwise orientation method.
+	\item Difficulties in choosing an algorithm: All the algorithms
+		have the same goal with all of them having similar asympottic
+		properties. However, the finite sample properties of these
+		methods vary depending on the choices that we make. For
+		example, the choice of the conditional independence (CI) test
+		being used for constraint-based method depends on the dataset,
+		similarly for the scoring method for score-based methods. With
+		no obvious way to test the impact/performance of these choices
+		on a given dataset, it becomes quite hard to chose the optimal
+		algorithm for a given dataset.
 \end{enumerate}
 
-On the other hand, in the field of Structural Equation Models (SEMs), models
-are also typically constructed by hand. And a lot of focus has been put on
-developing tools to help researchers in constructing these models. Researchers
-typically come up with an initial model, which then they use the data and
-expert knowledge to make modifications to. This process is commonly known as
-\emph{Specification Search}. These tools can then be used to guide the
-researchers in suggesting best ways to modify the model combined with the
-expert knowledge. For example, \emph{modification indices} take the initial
-drawn model, and make all possible modifications to the model to check which
-modification improves the model fit the best. Based on this information, the
-researcher can make modifications to the model and reiterate this process till
-they get something that they are satisfied with.
+On the other hand, in the field of Structural Equation Models (SEMs), a lot
+more focus has been put on methods to help researchers draw models manually
+along with integrating their expertise. Researchers typically come up with an
+initial model based on their domain knowledge, and then use various tools to
+guide them in making modifications to this initial model. This process of
+constructing models is commonly known as \emph{Specification Search}. A
+commonly used method for suggesting potential modifications to the SEM is
+\emph{modification indices}. Given a specified model, modification indices
+tries to find potential modifications to the model that improve the model's fit
+to the data best. Based on the suggestions from this method along with their
+domain expertise, the researcher can then choose the most optimal modification
+for their model.
+
+\todo[inline]{Other methods: Modification indices and Lagrange Multiplier tests to
+selectively add edges And using Wald test or z statistics (also known as
+critical ratios) for removing edges.}
 
@@
 statements.
-
-\todo[inline]{Other methods: Modification indices and Lagrange Multiplier tests to
-selectively add edges And using Wald test or z statistics (also known as
-critical ratios) for removing edges.}
 

### 2024-06-11 paper/main.tex
Updates

@@
 
-\section{Measure of Association for mixed data}
-In this section, we use canonical correlations to define an effect size measure
-for mixed data CI tests. Our approach is similar to the approach used for the
-continuous variable case.
+\section{A Measure of Association for Mixed Data}
+In the previous section, we outlined measure of associations for different data
+types, however we do not have any measure of association for a mixture of
+categorical, ordinal, and continuous data. In this section, we develop a
+measure of association for mixed data as a generalization of partial
+correlation by combining a mixed data residualization approach with canonical
+correlations.
+
+\subsection{Mixed Data Residualization}
 
@@
 
-\subsection{Connection to special cases}
+\subsection{Connections to other measures}
 This natually is a generalization of partial correlations. Furthermore, it is 

### 2024-06-11 paper/main.tex
Updates

@@
 correlation by combining a mixed data residualization approach with canonical
-correlations.
+correlations. Similar to partial correlation, we first compute the residuals of
+both variables using some machine learning model, and then use a canoncial
+correlation based measure of association between the residuals instead of
+correlation coefficient.
 
 \subsection{Mixed Data Residualization}
+Residualization for mixed data is not as straight forwards as for continuous
+data. We use the residualization approach proposed in Ankan 2023. Given a variable
+$ X $ conditioned on $ \bm{Z} $, we compute the residuals as follows:
+
+If $ X $ is continuous, we simply take the difference between the true and the 
+predicted value.
+
+If $ X $ is ordinal, we take the LS-Residuals
+
+If $ X $ is categorical, we one hot encode the variable.
+
+As a result of this residualization we can get a matrix of continuous values for 
+the residuals. We need a measure of association for this set of variables.
 
@@
 size.
+
+Benefits: Indifferent to one-hot encoding of the categorical variable.
 

### 2024-06-11 web/adult-test.R
Updates

@@ -11,7 +11,7 @@ cors <- function(res) {
 #* @param dag
 #* @param threshold
 #* @get /simpletest
-my_endpoint <- function( dag, threshold){
+my_endpoint <- function( dag, threshold, pval){
 		d <- get("d")
 		g <- dagitty::dagitty(dag)
 		r <- c()
@@ -42,7 +42,7 @@ my_endpoint <- function( dag, threshold){
 
 			}
 		}
-		r <- r %>% filter(abs(cor) > as.double(threshold))
+		r <- r %>% filter(abs(cor) > as.double(threshold)) %>% filter(p <= as.double(pval))
 		return (r)
 }
 

### 2024-06-11 web/utils.js
Updates

@@ -1,0 +1,72 @@
+function getEdgeDOM( u , v, dir ){ let ekv = DAGitty.controllers[0].getView().edge_shapes.kv
+	let eid = `${u}\u0000${v}\u0000${dir}`
+	if( ekv[eid] ) {
+		return ekv[eid].dom.firstChild
+	} else {
+		return null
+	}
+}
+
+function dagOnly(){
+	let g = DAGitty.controllers[0].graph
+	let g2 = new Graph()
+	for( let v of g.getVertices() ){
+		g2.addVertex( new Graph.Vertex( v ) )
+	}
+	for( let e of g.getEdges() ){
+		if( e.directed == Graph.Edgetype.Directed ){
+			g2.addEdge( e.v1, e.v2, e.directed )
+		}
+	}
+	g.copyAllPropertiesTo( g2 )
+	g2.setBoundingBox( g.getBoundingBox() )	
+	return g2
+}
+
+async function send(){
+	let g  = DAGitty.controllers[0].graph
+	DAGitty.controllers[0].event_listeners["graphchange"] = []
+	// remove all undirected edges
+	g = dagOnly(g)
+	// send DAG to backend for testing
+	const a = await fetch('http://127.0.0.1:8000/simpletest?dag='+encodeURIComponent(g.toString())+'&threshold='+document.getElementById('thres_txt').value+'&pval='+document.getElementById('pval_txt').value)
+	const b = await a.json()
+	if( Array.isArray(b) ){
+		for( let e of b ){
+			e.edge = e.A == "->"
+			if( !e.edge ){
+				g.addEdge( e.X, e.Y, Graph.Edgetype.Undirected )
+			}
+		}
+	} else {
+		return
+	}
+	DAGitty.controllers[0].setGraph( g )
+	DAGitty.controllers[0].redraw() // creates new edge shapes*/
+	//return
+	for( let e of b ){
+		let edom = getEdgeDOM( e.X, e.Y, 0+e.edge )
+		if( !edom && e.edge ){
+			edom = getEdgeDOM( e.v, e.u, 0+e.edge )
+
+		}
+		if( edom ){
+			edom.style.strokeWidth = 5*parseFloat(e.cor)
+			if( e.edge ){
+				edom.setAttribute("stroke", "#0A0")
+			} else {
+				edom.setAttribute("stroke", "#A00")
+			}
+		} else {
+			console.log( e )
+		}
+	}
[…]

### 2024-06-11 web/utils.js
Updates

@@ -31,6 +31,7 @@ async function send(){
 	// send DAG to backend for testing
 	const a = await fetch('http://127.0.0.1:8000/simpletest?dag='+encodeURIComponent(g.toString())+'&threshold='+document.getElementById('thres_txt').value+'&pval='+document.getElementById('pval_txt').value)
 	const b = await a.json()
+
 	if( Array.isArray(b) ){
 		for( let e of b ){
 			e.edge = e.A == "->"

### 2024-06-11 paper/main.tex
Updates

@@
 	\end{subfigure}
-	\caption{Comparison of PC and Hill-Climb Search algorithms against }
+	\caption{Comparison of PC and Hill-Climb Search algorithms against
+		manually drawn DAGs using the assistance method. As PC and
+		Hill-Climb Search return the Markov Equivalence Class (MEC), we
+		use the best and worst scoring orientation of the MEC to get
+		the range of values. The human drawn values are done for
+		multiple accuracy values of the oracle.}
 \end{figure}

### 2024-06-21 paper/main.tex
Updates

@@
 
-A lot of questions in life sciences center around a causal understanding
-between variables. \emph{Causal Discovery} aims to find these casual structure
+Many questions in science center around a causal understanding between
+variables. \emph{Causal Discovery} aims to find the underlying causal structure
 among random variables from a given dataset. The problem of causal discovery
-has been studied both the in Directed Acyclic Graph (DAGs) and Structure
-Equation Models (SEM) literatures. In the field of DAGs, many automated
-algorithms have been developed to learn the casual structure form a given 
-dataset such as constraint-based methods, score-based methods, and continuous
-optimization based methods.
+has been studied both in Directed Acyclic Graphs (DAGs) and Structural Equation
+Models (SEM) literatures. In the DAG literature, the focus has been on
+developing automated algorithms that can learn the causal structure from a
+given observational or interventional dataset, whereas in the SEM literature
+the focus has on developing tools to help researchers construct or modify
+manually constructed models.
 
@@
 
-However, the adaption of these algorithms in applied fields remain extremely
-limited. In these applied fields, even though these algorithms exist,
-researchers typically prefer to construct these DAGs manually based on their
-domain knowledge. We hypothesize that this preference towards manual DAG
-construction stems from the following difficulties in applying causal discovery
-methods:
+In the DAG literature, many automated algorithms have been developed for causal
+discovery such as constraint based method (PC, FCI), score-based methods
+(Hill-Climb Search, Greedy Equivalence Search), and optimization based methods.
+However, the adaption of these algorithms in applied fields remain limited. In
+these applied fields, researchers typically prefer to construct these DAGs
+manually based on their domain knowledge. We hypothesize that this preference
+towards manual DAG construction stems from the following difficulties in
+applying causal discovery methods:
 
 \begin{enumerate}
-	\item Difficult to trust: Although asypotically these methods are 
-		consistent, in the finite sample scenario they can end up
-		making obvious mistakes. As a result of this, researchers
-		have to rely on posthoc modification of the output of these
-		algorithms using either their domain knowledge or assisted
-		by testing methods.
-	\item Algorithms output the Markov Equivalence Class (MEC): With any
-		causal discovery algorithm, we can only get to the MEC. However,
-		most of the methods developed for downstream tasks such 
-		as identification or casual effect estimation assume the
-		knowledge of a DAG. To orient these MECs to a DAG, researchers
-		have to either rely on their domain expertise, or make further
+	\item \textbf{Difficult to trust: } Even though most of these algorithms
+		have asymptotic theory proving them to be consistent, their
+		finite sample properties is not well understood. In real
+		setting with finite samples, these algorithms can make obvious
+		mistakes. As a result of this, researchers have to rely on
+		post-hoc modification of the output of these algorithms
+		using either their domain knowledge or assisted by model testing
+		methods.
+	\item \textbf{Algorithms output Markov Equivalence Class (MEC):} Given an 
+		observational dataset, multiple causal structures can be
+		faithful to it. As a result, using only observational datasets,
+		these algorithms can only get to the MEC such as Completed
+		Partially Directed Acyclic Graph (CPDAG) or a Partial Ancestral
+		Graph (PAG) where not all the edges are directed. However, as
+		most of the methods developed for downstream tasks such as
+		identification or casual effect estimation assume the knowledge
+		of a DAG. To orient these MECs to a DAG, researchers have to
+		either rely on their domain expertise, or make further
 		assumption to utilize a pairwise orientation method.
-	\item Difficulties in choosing an algorithm: All the algorithms
-		have the same goal with all of them having similar asympottic
-		properties. However, the finite sample properties of these
-		methods vary depending on the choices that we make. For
-		example, the choice of the conditional independence (CI) test
-		being used for constraint-based method depends on the dataset,
-		similarly for the scoring method for score-based methods. With
-		no obvious way to test the impact/performance of these choices
-		on a given dataset, it becomes quite hard to chose the optimal
-		algorithm for a given dataset.
+	\item \textbf{Difficult to choose appropriate algorithm and
+		hyperparameters:} All the algorithms have the same goal with
+		all of them having similar asymptotic properties. However, the
+		finite sample properties of these methods vary depending on the
+		choices that we make. For example, the choice of the
+		conditional independence (CI) test being used for
+		constraint-based method depends on the dataset, similarly for
+		the scoring method for score-based methods. With no obvious way
+		to test the impact/performance of these choices on a given
+		dataset, it becomes quite hard to chose the optimal algorithm
+		for a given dataset.
 \end{enumerate}
@@
 
-In the case of DAGs, global testing methodology does exist such as Shipley
-test, that can combine local structure tests into a single p-value test.
-However, this does not give us a way to make modifications to our model. A
-possible way to do this by checking which of these local tests are violated.
-However, this approach can quickly become diffult as the size of the network
-increases. For example, the popular alarm network implies $x$ condition independence
-statements.
+As it is evident that researchers also draw DAGs by hand and even if they use
+an algorithm, the need to make some post-hoc adjustments to the model, in this
+paper we present a tool to assist researchers in making these adjustments
+somewhat inspired by modification indices. In the case of DAGs, global testing
+methodology does exist such as Shipley test, that can combine local structure
+tests into a single p-value test. However, this does not give us a way to make
+modifications to our model. A possible way to do this by checking which of
+these local tests are violated. However, this approach can quickly become
+diffult as the size of the network increases. For example, the popular alarm
+network implies $x$ condition independence statements.
 
-One of the drawbacks of these global fit metrics is that they do not only test
-for the structure but also the fitted parameterization.
+% One of the drawbacks of these global fit metrics is that they do not only test
+% for the structure but also the fitted parameterization.
 
-We present a similar procedure for modifying DAGs based on expert knowledge and
-measure of association. We first use the measure of association to test whether
-the association observed in the data is explained by the model structure. We
-can not directly use the local CI tests to make modifications because that
-would give way too many independencies and we have no way to rank them. Based
-on this testing, we suggest adding edges by ranking them by the most
-unexplained association between variables. The researcher can then use expert
-knowledge to decide the direction of the edge between the variables.
+Our proposed approach uses a measure of association to find correlated variables
+in the data that are not explained by the current model. Using the measure, we 
+can rank the edge that can lead to the highest explanation of the observed 
+correlation. Based on this information, the user can select a potential edge
+and based on their expert knowledge decide the direction of the edge.
+
+% We present a similar procedure for modifying DAGs based on expert knowledge and
+% measure of association. We first use the measure of association to test whether
+% the association observed in the data is explained by the model structure. We
+% can not directly use the local CI tests to make modifications because that
+% would give way too many independencies and we have no way to rank them. Based
+% on this testing, we suggest adding edges by ranking them by the most
+% unexplained association between variables. The researcher can then use expert
+% knowledge to decide the direction of the edge between the variables.
 
@@
 \begin{enumerate}
-	\item We present a measure of (conditional) assocation between mixed variables using
-		canonical correlations. This mixed variable association is a generalization 
-		of some of the other association measures such as partial correlation to 
-		mixed variables.
-	\item We present a procedure to use this measure of association to
-		iteratively modify DAGs using expert knowledge such that it
-		improves the explained association between variables.
-	\item Lastly, to assist researchers in manully constructing and modifying DAGs,
-		we developed a web-based tool to iteratively modify their DAGs.
+	\item We present a measure of (conditional) association for mixed data
+		based on canonical correlations. This measure of association
+		generalizes the widely used partial correlation to mixed data
+		setting.
+	\item Using this association measure, we present a procedure to suggest
+		potential new edges, or edge removals in the model. The user
+		gets to decide the direction of the new edge based on their
+		expertise.
+	\item Lastly, we present a web-browser based tool to allow researchers
+		to apply this method to their own datasets.
 \end{enumerate}
@@
 
+The rest of the paper is structured as follows. In
+Section~\ref{sec:background}, we give a background on the measures of
+association for various data types. In Section~\ref{sec:mixed_association}, we
+present our method for computing mixed data association. In
+Section~\ref{sec:modification}, we present how to use this method to rank the
+modifications, and lastly in Section~\ref{sec:empirical}, we show some
+empirical analysis on how well this modification process works.
+
 \section{Background}
+\label{sec:background}
 We represent uni-dimensional variables as $ X $ and multi-dimensional variables
@@
 \section{A Measure of Association for Mixed Data}
+\label{sec:mixed_association}
 In the previous section, we outlined measure of associations for different data
@@
 \section{Using measure of association for modifying DAGs}
+\label{sec:modification}
 The idea behind using the measure of association for modifying DAGs is that,
@@
 \section{Web Tool}
+\label{sec:web}
 Based on the approach outlined in the previous section, we built a web-based
@@
 \section{Empirical Analysis}
+\label{sec:empirical}
 

### 2024-06-21 paper/main.tex
Updates

@@
 \label{sec:mixed_association}
-In the previous section, we outlined measure of associations for different data
-types, however we do not have any measure of association for a mixture of
-categorical, ordinal, and continuous data. In this section, we develop a
-measure of association for mixed data as a generalization of partial
-correlation by combining a mixed data residualization approach with canonical
-correlations. Similar to partial correlation, we first compute the residuals of
-both variables using some machine learning model, and then use a canoncial
-correlation based measure of association between the residuals instead of
-correlation coefficient.
+
+In this section, we develop a (conditional) measure of association for mixed
+data as a generalization of partial correlation by combining a mixed data
+residualization approach with canonical correlations. Assuming that we need to
+compute the association, $ \rho(X, Y; Z) $ between variables $ X $ and $ Y $
+conditioned on $ \bm{Z} $. Similar to partial correlations, we first start by
+computing the residuals for both $ X $ and $ Y $ using $ \bm{Z} $ as the
+predictor variables. When $ \bm{Z} = \emptyset $ we can use a constant value
+for $ \bm{Z} $. We use the mixed data residualization approach presented in
+Ankan and Textor 2023.
 
@@
 
-\subsection{No conditional variables}
-When we have no conditonal variables, we need to convert the variables into
-
-\subsection{Conditional Case}
 For the conditional case we use an approach similar to partial correlation. We
@@
 \label{sec:modification}
+
 The idea behind using the measure of association for modifying DAGs is that,
@@
 tool that researchers can upload their dataset to and apply this method to
-interactively draw models.
+interactively draw models. Users can start by uploading their datasets. The web
+interface should then automatically show them all the correlated variables in a
+graphical representation. The red colored edges represent potential new edge
+between the variables.
 
@@
 
+\subsection{Using Large Language Models as Experts}
+A lot of recent work has focused on finding pairwise causal direction among
+variables using Large Language Models. In this analysis, instead of using the
+oracle to decide the direction of edges, we used GPT4 to give us the direction
+of the edges based on description of the variables. As GPT4 has already
+memorized many of these models, we encoded the variable names using a random
+string and provided the description of the variable to let it decide the
+direction of the edge.
+
 \section{Conclusions}
-\subsection{Using LLMs as experts}
 

### 2024-06-22 paper/main.tex
Updates

@@
 
-When $ \bm{Z} = \emptyset $, ordinal variable can be assumed to be coming from a thresholded normal distribution, estimate the thresholds and latent variable to compute 
-the correlation coefficient.
+When $ \bm{Z} = \emptyset $, ordinal variable can be assumed to be coming from
+a thresholded normal distribution, estimate the thresholds and latent variable
+to compute the correlation coefficient.
 
@@
 In this section, we develop a (conditional) measure of association for mixed
-data as a generalization of partial correlation by combining a mixed data
-residualization approach with canonical correlations. Assuming that we need to
-compute the association, $ \rho(X, Y; Z) $ between variables $ X $ and $ Y $
-conditioned on $ \bm{Z} $. Similar to partial correlations, we first start by
-computing the residuals for both $ X $ and $ Y $ using $ \bm{Z} $ as the
-predictor variables. When $ \bm{Z} = \emptyset $ we can use a constant value
-for $ \bm{Z} $. We use the mixed data residualization approach presented in
-Ankan and Textor 2023.
+data. This measure is similar to partial correlation in the sense that it uses
+a residualization approach and takes the association between the residuals.
 
-\subsection{Mixed Data Residualization}
-Residualization for mixed data is not as straight forwards as for continuous
-data. We use the residualization approach proposed in Ankan 2023. Given a variable
-$ X $ conditioned on $ \bm{Z} $, we compute the residuals as follows:
+Given two variables $ X $ and $ Y $, and a set of conditioanl variables $
+\bm{Z} $, we are interested in computing the conditional association, $ \rho(X,
+Y; Z) $. Similar to partial correlation, we start by computing the residuals
+for $ X $ and $ Y $ using the $ \bm{Z} $ as the predictor varaibles and then
+compute the assocation between the residuals. As the residuals for mixed
+variable is not straightforward to define, we use the mixed data
+residualization approach presented in Ankan and Textor 2023.
 
-If $ X $ is continuous, we simply take the difference between the true and the 
-predicted value.
+Given a variable $ X $ with predictor variables $ \bm{Z} $, we first train an
+estimator, $ E_X $ for predicting $ X $ using $ \bm{Z} $. If the varialbes are
+ordinal or categorical, the estimator needs to be probability estimators for
+example a Random Forest Estimator or XGBoost. Given an estimator we can compute
+the residuals as follows:
+\begin{enumerate}
+	\item If $ X $ is continuous, we simply take the difference between the true and estimated value.
+		$$ R_X = X - E_X(\bm{Z}) $$
+	\item If $ X $ is ordinal, we take the Li-Sheperd residual.
+		$$ R_X = E_X(\bm{Z}) $$
+	\item If $ X $ is categorical, we one hot encode the variable and Li-Sheperd residuals as well.
+		$$ R_X = E_X(\bm{Z}) $$
+\end{enumerate}
+As a result of this residualization approach, the residuals are a matrix of
+continuous values. The shape of this matrix depends on the type of variable. If
+the variables are continuous or ordinal, we end up with a residual matrix of
+size $ n \times 1 $. If the the variable is categorical we end up with a
+residual matrix of size $ n \times (k-1) $, where $ k $ is the cardinality of
+the categorical variables. We have dropped one of the one hot encoded columns
+because it is redundant.
 
-If $ X $ is ordinal, we take the LS-Residuals
-
-If $ X $ is categorical, we one hot encode the variable.
-
-As a result of this residualization we can get a matrix of continuous values for 
-the residuals. We need a measure of association for this set of variables.
-
-\subsection{Canonical Correlation}
-Canonical correlations are generalization of correlations to multi-dimensional variables.
+Next we need a measure to quantify the association between these residuals. One
+of the general ways to have assocation between sets of variables is to use
+canonical correlations. Canonical correlations are generalization of
+correlations to multi-dimensional variables.
 
@@
 
-Multiple effect sizes measures based on canonical correlation are available.
+Multiple measures of association based on canonical correlation are available such as:
 \begin{itemize}
@@
 \end{itemize}
+
 We use Pillai's Trace in this paper. \todo[inline]{Why?}.
 
-For the conditional case we use an approach similar to partial correlation. We
-combine the mixed data residualization approach proposed in Ankan 2023. We then
-use a canonical correlation based measure on these residuals to get an effect
-size.
+This measure of association has many of the desired properties:
 
-Benefits: Indifferent to one-hot encoding of the categorical variable.
+\begin{enumerate}
+	\item Is bounded between $ 0 $ and $ 1 $.
+	\item Is independent of the sample size.
+	\item Is indifferent to the one-hot encoding scheme used for categorical variables.
+	\item In the case of both continuous variables, reduces to partial correlation.
+\end{enumerate}
 
 \subsection{Connections to other measures}
+In the continuous variable case, this is equivalent to partial correlation.
+
 This natually is a generalization of partial correlations. Furthermore, it is 
@@
 \label{sec:modification}
+
+Any two variables in a DAG that do not have an edge between them imply a
+conditional independence test that is they are independent given the parents of
+both the variables, $ X \ci Y | pa(X) \cup pa(Y) $. In the case of model
+testing, a CI statement like this can be used to check whether the assumption
+of the model holds in the data or is violated. For mixed data, a few different
+CI tests have been proposed such as likelihood-ratio based, and residualization
+based. However, if we just use the CI test to check the p-value we cannot use
+it to find the modification that is able to improve the score the most.
 

### 2024-06-22 paper/main.tex
Updates

@@
 based. However, if we just use the CI test to check the p-value we cannot use
-it to find the modification that is able to improve the score the most.
+it to find the modification that is able to improve the score the most. We can
+however use the measure of association to check which assocation between the
+variables in the data is not explained by the model. Using this the expert can
+decide to add an edge between the variables but will need to choose the
+direction of the edge.
 
-The idea behind using the measure of association for modifying DAGs is that,
-for any given model and a dataset, we can use the condtiional measure of
-association to compute the amount of correlation in the data that is still not
-explained by the model. Based on this, we can compute how much improvement in
-the explanation would happen if we were to add an edge between variables. This
-can then be ranked using the maximum improvement. Give some examples here.
+The idea behind using the measure of association for modifying DAGs is inspired
+by modification indices in SEM. For any given model and a dataset, we can use
+the condtiional measure of association to compute the amount of correlation in
+the data that is still not explained by the model. Based on this, we can
+compute how much improvement in the explanation would happen if we were to add
+an edge between variables. This can then be ranked using the maximum
+improvement. Give some examples here.
 

### 2024-06-22 paper/main.tex
Updates

@@
 The idea behind using the measure of association for modifying DAGs is inspired
-by modification indices in SEM. For any given model and a dataset, we can use
-the condtiional measure of association to compute the amount of correlation in
-the data that is still not explained by the model. Based on this, we can
-compute how much improvement in the explanation would happen if we were to add
-an edge between variables. This can then be ranked using the maximum
-improvement. Give some examples here.
+by modification indices in SEM.
 

### 2024-06-24 paper/main.tex
updates

@@
 We represent uni-dimensional variables as $ X $ and multi-dimensional variables
-as $ \bm{Z} $. We consider the case of estimating association between $ X $,
-and $ Y $ is the presence of a set of conditioning variables $ \bm{Z} = Z_1,
-\cdots, Z_k $, potentially being an empty set $ \bm{Z} = \emptyset $. We use $
-\rvert \bm{Z} \rvert $ to denote the cardinality of the variable $ \bm{Z} $. We
+as $ \bm{Z} $. All the variables are considered to be any mix of continuous,
+ordinal, or categorical unless specified explicitly.
+
+We consider the case of estimating association between $ X $ and $ Y $ is the
+presence of a set of conditioning variables $ \bm{Z} = Z_1, \cdots, Z_k $,
+potentially being an empty set $ \bm{Z} = \emptyset $. We use $ \rvert \bm{Z}
+\rvert $ to denote the cardinality of the set of variables $ \bm{Z} $. We
 consider various scenarios where the variables $ X $, $ Y $ and $ \bm{Z} $
@@
 
-\subsection{Measures of Association}
+\subsection{Measures of (Conditional) Association}
 As our DAG construction procedure relies on a measure of conditional
@@
 
-\section{A Measure of Association for Mixed Data}
+\section{Measure of Association for Mixed Data}
 \label{sec:mixed_association}
 
-In this section, we develop a (conditional) measure of association for mixed
-data. This measure is similar to partial correlation in the sense that it uses
-a residualization approach and takes the association between the residuals.
+In this section, we develop a novel (conditional) measure of association for
+mixed data. The presented measure extends partial correlation to mixed data by
+combining a mixed-data residualization method with a multivariate measure of
+association.
 
-Given two variables $ X $ and $ Y $, and a set of conditioanl variables $
+Given two variables $ X $ and $ Y $, and a set of conditional variables $
 \bm{Z} $, we are interested in computing the conditional association, $ \rho(X,
-Y; Z) $. Similar to partial correlation, we start by computing the residuals
-for $ X $ and $ Y $ using the $ \bm{Z} $ as the predictor varaibles and then
-compute the assocation between the residuals. As the residuals for mixed
-variable is not straightforward to define, we use the mixed data
-residualization approach presented in Ankan and Textor 2023.
+Y; Z) $. As our measure is based on residulaization, we start by computing the
+residuals for $ X $ and $ Y $ using the $ \bm{Z} $ as the predictor variables.
+We use the mixed data residualization approach from \citet{Ankan2023}. We
+consider each variable type separately and repeat the following steps for both
+$ X $ and $ Y $:
 
-Given a variable $ X $ with predictor variables $ \bm{Z} $, we first train an
-estimator, $ E_X $ for predicting $ X $ using $ \bm{Z} $. If the varialbes are
-ordinal or categorical, the estimator needs to be probability estimators for
-example a Random Forest Estimator or XGBoost. Given an estimator we can compute
-the residuals as follows:
 \begin{enumerate}
-	\item If $ X $ is continuous, we simply take the difference between the true and estimated value.
-		$$ R_X = X - E_X(\bm{Z}) $$
-	\item If $ X $ is ordinal, we take the Li-Sheperd residual.
-		$$ R_X = E_X(\bm{Z}) $$
-	\item If $ X $ is categorical, we one hot encode the variable and Li-Sheperd residuals as well.
-		$$ R_X = E_X(\bm{Z}) $$
+	\item $ X $ is continuous: We start by training a regression model, $
+		E_X: X \sim \bm{Z} $. Next we take the predictions from this
+		model $ \hat{x} = E_X(\bm{Z}) $ and lastly the residuals are
+		computed as the difference between the true and the predicted
+		values.
+		$$ R_X = X - \hat{x} $$
+	\item $ X $ is ordinal: We start by training a probability estimator, $
+		p_X: X \sim \bm{Z} $, and take probability predictions from it
+		$ \hat{p}_X: p_X(\bm{Z}) $. Then we compute the residuals as follows:
+		$$ R_X = $$
+	\item $ X $ is categorical: We again start by training a probability estimator
+		$ p_X: X \sim \bm{Z} $, and take the probability estimates from it $ \hat{p}_X: p_Z(\bm{Z}) $.
+		Next, we one-hot encode the categorical variable and then compute the residuals
+		as follows:
+		$$ R_X $$
 \end{enumerate}
-As a result of this residualization approach, the residuals are a matrix of
-continuous values. The shape of this matrix depends on the type of variable. If
-the variables are continuous or ordinal, we end up with a residual matrix of
-size $ n \times 1 $. If the the variable is categorical we end up with a
-residual matrix of size $ n \times (k-1) $, where $ k $ is the cardinality of
-the categorical variables. We have dropped one of the one hot encoded columns
-because it is redundant.
 
-Next we need a measure to quantify the association between these residuals. One
-of the general ways to have assocation between sets of variables is to use
-canonical correlations. Canonical correlations are generalization of
-correlations to multi-dimensional variables.
+Here, users can choose the type of estimator that they want to use. There are
+single estimators such as Random Forest or XGBoost that can work for both the
+probability estimates and regression estimates. Or users can also choose
+different estimators such as variants of linear regression for computing the
+residuals. After the above steps are repeated for both $ X $ and $ Y $, we end
+up with two residual matrices: $ R_X $ and $ R_Y $. The type of the variable
+determines the shape for these matrices, i.e., if the variable is continuous we
+get a residual matrix of shape $ ( n \times 1 ) $ and if the variable is
+categorical we have a residual matrix of shape $ ( n \times (k-1)) $ (where one
+of the dummy encoded variables is dropped to avoid multicollinearity)
+
+Now, we have the two residuals and we need a measure to quantify the
+association between these residuals. As the residuals are matrices, we can
+treat them as sets of random variables, and in multivariate analysis canonical
+correlations have been used to quantify association between two sets of
+variables. Canonical correlations are generalization of correlations to
+multi-dimensional variables where they try to find linear combinations of the
+given sets of variables that maximize the correlation between them.
 
@@
 
-Multiple measures of association based on canonical correlation are available such as:
+Many measures of association based on canonical correlation are also available such as:
 \begin{itemize}
@@
 \begin{enumerate}
-	\item Is bounded between $ 0 $ and $ 1 $.
+	\item Is bounded between $ 0 $ and $ 1 $, with $ 0 $ representing no association and $ 1 $ representing a perfect linear relationship.
 	\item Is independent of the sample size.
-	\item Is indifferent to the one-hot encoding scheme used for categorical variables.
-	\item In the case of both continuous variables, reduces to partial correlation.
+	\item As canonical correlation finds linear transformations, the measure is independent of the one-hot encoding scheme used for categorical variables.
+	\item In the case of both continuous variables, it is equivalent to partial correlations.
+	\item For both discrete variables, the measure is connected to Cramers\' V, a well know effect size measure for discrete variables.
 \end{enumerate}
-
-\subsection{Connections to other measures}
-In the continuous variable case, this is equivalent to partial correlation.
-
-This natually is a generalization of partial correlations. Furthermore, it is 
-also a generalization of discrete case.
 

### 2024-06-25 paper/main.tex
Updates

@@
 
-Many questions in science center around a causal understanding between
-variables. \emph{Causal Discovery} aims to find the underlying causal structure
+Many questions in science center around understanding the cause and effect
+relationships between variables, as it can reveal the underlying mechanisms for
+the observed phenomena and inform effective interventions or policy decisions.
+\emph{Causal Discovery} methods aim to find the underlying causal structure
 among random variables from a given dataset. The problem of causal discovery
-has been studied both in Directed Acyclic Graphs (DAGs) and Structural Equation
-Models (SEM) literatures. In the DAG literature, the focus has been on
-developing automated algorithms that can learn the causal structure from a
-given observational or interventional dataset, whereas in the SEM literature
-the focus has on developing tools to help researchers construct or modify
-manually constructed models.
+has been studied in both Directed Acyclic Graphs (DAGs) and Structural Equation
+Models (SEM) literatures with different approaches. In the DAG literature, the
+focus has been on developing automated algorithms that can learn the causal
+structure from a given observational or interventional dataset, whereas in the
+SEM literature the focus has on developing tools to help researchers construct 
+models manually or modify manually constructed models.
 
@@
 In the DAG literature, many automated algorithms have been developed for causal
-discovery such as constraint based method (PC, FCI), score-based methods
-(Hill-Climb Search, Greedy Equivalence Search), and optimization based methods.
-However, the adaption of these algorithms in applied fields remain limited. In
-these applied fields, researchers typically prefer to construct these DAGs
-manually based on their domain knowledge. We hypothesize that this preference
-towards manual DAG construction stems from the following difficulties in
-applying causal discovery methods:
+discovery such as constraint based methods like PC \citep{Spirtes2001} and FCI
+\citep{Spirtes2000}, score-based methods like Hill-Climb Search, Greedy
+Equivalence Search \citep{Chickering2002}, and optimization based methods.
+However, the adaption of these algorithms in applied fields remain limited
+\citep{Tennant2020}. In applied fields, researchers typically prefer to
+construct these DAGs manually based on their domain knowledge. We hypothesize
+that this preference towards manual DAG construction stems from the following
+difficulties in applying causal discovery methods:
 

### 2024-06-25 paper/main.tex
Updates

@@
 In this section, we develop a novel (conditional) measure of association for
-mixed data. The presented measure extends partial correlation to mixed data by
+mixed data by extending partial correlation to mixed data. We do this by
 combining a mixed-data residualization method with a multivariate measure of
-association.
-
-Given two variables $ X $ and $ Y $, and a set of conditional variables $
-\bm{Z} $, we are interested in computing the conditional association, $ \rho(X,
-Y; Z) $. As our measure is based on residulaization, we start by computing the
-residuals for $ X $ and $ Y $ using the $ \bm{Z} $ as the predictor variables.
-We use the mixed data residualization approach from \citet{Ankan2023}. We
-consider each variable type separately and repeat the following steps for both
-$ X $ and $ Y $:
+association. Given two variables $ X $ and $ Y $, and a set of conditional
+variables $ \bm{Z} $, we are interested in computing an estimate of the
+conditional association, $ \rho(X, Y; Z) $. Firstly, we compute the residuals
+for $ x $ and $ y $ using the $ \bm{z} $ as the predictor variables. We use the
+mixed data residualization approach from \citet{Ankan2023}. We consider each
+variable type separately and repeat the following steps for both variables $ X
+$ and $ Y $:
 
 \begin{enumerate}
-	\item $ X $ is continuous: We start by training a regression model, $
-		E_X: X \sim \bm{Z} $. Next we take the predictions from this
-		model $ \hat{x} = E_X(\bm{Z}) $ and lastly the residuals are
-		computed as the difference between the true and the predicted
-		values.
-		$$ R_X = X - \hat{x} $$
-	\item $ X $ is ordinal: We start by training a probability estimator, $
-		p_X: X \sim \bm{Z} $, and take probability predictions from it
-		$ \hat{p}_X: p_X(\bm{Z}) $. Then we compute the residuals as follows:
-		$$ R_X = $$
-	\item $ X $ is categorical: We again start by training a probability estimator
-		$ p_X: X \sim \bm{Z} $, and take the probability estimates from it $ \hat{p}_X: p_Z(\bm{Z}) $.
-		Next, we one-hot encode the categorical variable and then compute the residuals
-		as follows:
-		$$ R_X $$
+	\item \textbf{$ X $ is continuous:} We start by training a regression
+		model, $ E_X: x \sim \bm{z} $. Next we take the predictions
+		from this model $ \hat{x} = E_X(\bm{z}) $ and lastly the
+		residuals are computed as the difference between the true and
+		the predicted values.
+		$$ R_{x_i} = x_i - \hat{x}_i $$
+	\item \textbf{$ X $ is ordinal:} We start by training a probability
+		estimator, $ p_X: x \sim \bm{z} $, and take probability
+		predictions from it $ \hat{p}_X: \hat{p}_X(\bm{Z}) $. Then we
+		compute the residuals as follows:
+		$$ R_{x_i} = \hat{p}(\hat{x}_i > x_i) - \hat{p}(\hat{x}_i < x_i) $$
+	\item \textbf{$ X $ is categorical:} We again start by training a
+		probability estimator $ p_X: X \sim \bm{Z} $, and take the
+		probability estimates from it $ \hat{p}_X: p_Z(\bm{Z}) $. Next,
+		we one-hot encode the categorical variable and then compute the
+		residuals as follows: 
+		$$ R_X = x_i - \hat{p}(\bm{z}_i) $$
 \end{enumerate}
 
-Here, users can choose the type of estimator that they want to use. There are
-single estimators such as Random Forest or XGBoost that can work for both the
-probability estimates and regression estimates. Or users can also choose
-different estimators such as variants of linear regression for computing the
-residuals. After the above steps are repeated for both $ X $ and $ Y $, we end
-up with two residual matrices: $ R_X $ and $ R_Y $. The type of the variable
-determines the shape for these matrices, i.e., if the variable is continuous we
-get a residual matrix of shape $ ( n \times 1 ) $ and if the variable is
-categorical we have a residual matrix of shape $ ( n \times (k-1)) $ (where one
-of the dummy encoded variables is dropped to avoid multicollinearity)
+\todo[inline]{Fix notations above}
+
+Here, users have the option to choose the estimators based on the properties of
+their data. Non-parametric ensemble estimators such as Random Forest and
+XGBoost can work well on a large variety of data and any data type. For linear
+datasets users can also choose variants of linear regression depending on the
+type of data. After the above steps are repeated for both $ X $ and $ Y $, we
+end up with two residual matrices: $ R_X $ and $ R_Y $. The type of the
+variable determines the shape for these matrices, i.e., if the variable is
+continuous we get a residual matrix of shape $ ( n \times 1 ) $ and if the
+variable is categorical we have a residual matrix of shape $ ( n \times (k-1))
+$ (where one of the dummy encoded variables is dropped to avoid
+multicollinearity)
 

### 2024-06-25 paper/main.tex
Updates

@@
 In this section, we develop a novel (conditional) measure of association for
-mixed data by extending partial correlation to mixed data. We do this by
-combining a mixed-data residualization method with a multivariate measure of
-association. Given two variables $ X $ and $ Y $, and a set of conditional
-variables $ \bm{Z} $, we are interested in computing an estimate of the
-conditional association, $ \rho(X, Y; Z) $. Firstly, we compute the residuals
-for $ x $ and $ y $ using the $ \bm{z} $ as the predictor variables. We use the
-mixed data residualization approach from \citet{Ankan2023}. We consider each
-variable type separately and repeat the following steps for both variables $ X
-$ and $ Y $:
+mixed data by extending the idea of partial correlation to mixed data. We do
+this by combining a mixed-data residualization method with a multivariate
+measure of association. Given two variables $ X $ and $ Y $, and a set of
+conditional variables $ \bm{Z} $, we are interested in computing an estimate of
+the conditional association, $ \rho(X, Y; Z) $. If $ \bm{Z} = \emptyset $, we
+use a constant value as $ Z $. Firstly, we compute the residuals for $ x $ and
+$ y $ using the $ \bm{z} $ as the predictor variables. We use the mixed data
+residualization approach from \citet{Ankan2023}. We consider each variable type
+separately and repeat the following steps for both variables $ X $ and $ Y $:
 
@@
 
-Now, we have the two residuals and we need a measure to quantify the
-association between these residuals. As the residuals are matrices, we can
-treat them as sets of random variables, and in multivariate analysis canonical
-correlations have been used to quantify association between two sets of
-variables. Canonical correlations are generalization of correlations to
-multi-dimensional variables where they try to find linear combinations of the
-given sets of variables that maximize the correlation between them.
+We now have two continuous valued residual matrices and we need a measure to
+quantify the association between these residuals. In multivariate statistics,
+canonical correlation based measures have been used for measuring association
+between two sets of variables. We can treat the residuals as sets of variables
+in case of categorical variable and use canonical corrleations to quantify the
+association between them.
+\todo[inline]{Justify the argument}
 
@@
 
-	Canonical correlations try to find linear transformations $ a $ and $ b
-	$ of $ \bm{U} $ and $ \bm{V} $ such that the correlations between the
-	columns of $ a^T \bm{U} $ and $ b^T \bm{V} $ is maximized. The
-	individual columns of $ a^T \bm{U} $ and $ b^T \bm{V} $ are
-	independent. In the case when $ \rvert U \rvert = \rvert V \rvert = 1$,
-	canonical correlations are equivalent to Pearson's correlation
-	coefficient.
+Canonical correlations are generalization of correlations to multi-dimensional
+variables. It tries to find orthogonal linear transformations $ a $ and $ b $
+of $ \bm{U} $ and $ \bm{V} $ such that the correlation between the transformed
+variables $ a^T \bm{U} $ and $ b^T \bm{V} $ is maximized. It gives a vector of
+correlation values of size $ \min(\rvert U \rvert, \rvert V \rvert) $
+representing the correlation coefficient of each column of the transformed
+variables. Pearson's correlation coefficient is a special case of canonical
+correlations when $ \rvert U \rvert = \rvert V \rvert = 1$.
 
-Many measures of association based on canonical correlation are also available such as:
+Many measures of association based on canonical correlation have been used in
+multivariate statistics with each of them having slightly different properties.
 \begin{itemize}
@@
 
-We use Pillai's Trace in this paper. \todo[inline]{Why?}.
+We use Pillai's Trace for our purpose for two main reasons: 1) It uses all the
+canoncial correlation values, 2) Its interpretation is similar to correlation
+coefficient, i.e., $ 0 $ signifies no association and $ 1 $ signifies perfect
+linear relationship.
 
-This measure of association has many of the desired properties:
+\todo[inline]{Add the formula for our case here}
+
+This measure of association has many desired properties:
 
 \begin{enumerate}
-	\item Is bounded between $ 0 $ and $ 1 $, with $ 0 $ representing no association and $ 1 $ representing a perfect linear relationship.
-	\item Is independent of the sample size.
-	\item As canonical correlation finds linear transformations, the measure is independent of the one-hot encoding scheme used for categorical variables.
-	\item In the case of both continuous variables, it is equivalent to partial correlations.
-	\item For both discrete variables, the measure is connected to Cramers\' V, a well know effect size measure for discrete variables.
+	\item Bounded between $ 0 $ and $ 1 $.
+	\item Independent of sample size.
+	\item Invariant to the one-hot encoding scheme used fr categorical variables.
+	\item Reduces to partial correlations when both variables are continuous.
+	\item In case of discrete variables, connected to Cramer\'s V, a well
+		known effect size measure for discrete variables.
+		\todo[inline]{Show the connection}
 \end{enumerate}
 
-\section{Using measure of association for modifying DAGs}
+\section{Using measure of association for DAG construction}
 \label{sec:modification}

### 2024-06-25 code/llm.py
Adds llm script

@@ -1,0 +1,54 @@
+import pandas as pd
+from itertools import combinations
+
+from pgmpy.estimators.CITests import ci_pillai
+from pgmpy.base import DAG
+
+def test_all(dag, data):
+    cis = []
+    for u, v in combinations(list(dag.nodes()), 2):
+        u_parents = dag.get_parents(u)
+        v_parents = dag.get_parents(v)
+        cond_set = list(set(u_parents).union(v_parents))
+        import pdb; pdb.set_trace()
+        effect, p_value = ci_pillai(X=u, Y=v, Z=cond_set, data=data, boolean=False)
+        cis.append(u, v, cond_set, effect, p_value)
+
+    return pd.DataFrame(cis, columns=['u', 'v', 'z', 'effect', 'p_val'])
+
+df = pd.read_csv("adult_proc.csv", index_col=0)
+df.Age = pd.Categorical(
+    df.Age,
+    categories=["<21", "21-30", "31-40", "41-50", "51-60", "61-70", ">70"],
+    ordered=True,
+)
+df.Education = pd.Categorical(
+    df.Education,
+    categories=[
+        "Preschool",
+        "1st-4th",
+        "5th-6th",
+        "7th-8th",
+        "9th",
+        "10th",
+        "11th",
+        "12th",
+        "HS-grad",
+        "Some-college",
+        "Assoc-voc",
+        "Assoc-acdm",
+        "Bachelors",
+        "Masters",
+        "Doctorate",
+    ],
+    ordered=True,
+)
+df.HoursPerWeek = pd.Categorical(
+    df.HoursPerWeek, categories=["<=20", "21-30", "31-40", ">40"], ordered=True
+)
+
+nodes = list(df.columns)
+dag = DAG()
+dag.add_nodes_from(nodes)
+
+

### 2024-06-26 paper/main.tex
Updates

@@
 \newtheorem{proposition}{Proposition}
+\newtheorem{theorem}{Theorem}
 
@@
 
-Many questions in science center around understanding the cause and effect
-relationships between variables, as it can reveal the underlying mechanisms for
-the observed phenomena and inform effective interventions or policy decisions.
-\emph{Causal Discovery} methods aim to find the underlying causal structure
+Understanding cause-and-effect relationships between variables is a fundamental
+goal in many scientific questions as it reveals the underlying mechanisms of
+the observed phenomena and informs effective interventions or policy decisions.
+\emph{Causal Discovery} methods aim to uncover the underlying causal structure
 among random variables from a given dataset. The problem of causal discovery
-has been studied in both Directed Acyclic Graphs (DAGs) and Structural Equation
-Models (SEM) literatures with different approaches. In the DAG literature, the
-focus has been on developing automated algorithms that can learn the causal
-structure from a given observational or interventional dataset, whereas in the
-SEM literature the focus has on developing tools to help researchers construct 
-models manually or modify manually constructed models.
+has been explored in both the Directed Acyclic Graphs (DAGs) and Structural
+Equation Models (SEM) literatures, each with different approaches. In the DAG
+literature, the focus has been on developing automated algorithms that can
+learn the causal structure from observational or interventional dataset. In
+contrast, the SEM literature emphasizes tools that help researchers manually
+construct or modify modify.
 
@@
 However, the adaption of these algorithms in applied fields remain limited
-\citep{Tennant2020}. In applied fields, researchers typically prefer to
-construct these DAGs manually based on their domain knowledge. We hypothesize
-that this preference towards manual DAG construction stems from the following
-difficulties in applying causal discovery methods:
+\citep{Tennant2020}. Researchers in applied fields typically prefer to
+construct DAGs manually based on their domain knowledge. We hypothesize that
+this preference towards manual DAG construction stems from the following
+challenges in applying causal discovery algorithms:
 
 \begin{enumerate}
-	\item \textbf{Difficult to trust: } Even though most of these algorithms
-		have asymptotic theory proving them to be consistent, their
-		finite sample properties is not well understood. In real
-		setting with finite samples, these algorithms can make obvious
-		mistakes. As a result of this, researchers have to rely on
-		post-hoc modification of the output of these algorithms
-		using either their domain knowledge or assisted by model testing
-		methods.
-	\item \textbf{Algorithms output Markov Equivalence Class (MEC):} Given an 
-		observational dataset, multiple causal structures can be
-		faithful to it. As a result, using only observational datasets,
-		these algorithms can only get to the MEC such as Completed
-		Partially Directed Acyclic Graph (CPDAG) or a Partial Ancestral
-		Graph (PAG) where not all the edges are directed. However, as
-		most of the methods developed for downstream tasks such as
-		identification or casual effect estimation assume the knowledge
-		of a DAG. To orient these MECs to a DAG, researchers have to
-		either rely on their domain expertise, or make further
-		assumption to utilize a pairwise orientation method.
-	\item \textbf{Difficult to choose appropriate algorithm and
-		hyperparameters:} All the algorithms have the same goal with
-		all of them having similar asymptotic properties. However, the
-		finite sample properties of these methods vary depending on the
-		choices that we make. For example, the choice of the
-		conditional independence (CI) test being used for
-		constraint-based method depends on the dataset, similarly for
-		the scoring method for score-based methods. With no obvious way
-		to test the impact/performance of these choices on a given
-		dataset, it becomes quite hard to chose the optimal algorithm
-		for a given dataset.
+	\item \textbf{Trustworthiness: } Although most algorithms have
+		asymptotic theory proving them to be consistent, their finite
+		sample properties are not well understood. In real-world
+		settings with finite samples, these algorithms can make obvious
+		mistakes, for example, the constraint-based algorithms are
+		biased towards sparse graphs \citep{Malinsky2024}. Consequently,
+		researchers often have rely on post-hoc modifications based on 
+		their domain knowledge or model testing methods.
+	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple causal
+		structures can be faithful to a given observational dataset.
+		Thus, using only observational data, these algorithms can only
+		recover the MECs that can contain unoriented edges. However,
+		most methods for downstream tasks, such as identification or
+		causal effect estimation, assume knowledge of a fully oriented
+		DAG. As a results, researchers have to rely on their expertise
+		to orient these edges or make further assumptions to make use
+		of pairwise orientation techniques.
+	\item \textbf{Algorithm and hyperparameter selection:} Each algorithm
+		has the same goal but they can learn very different models
+		depending on the choice of algorithm and their hyperparameters.
+		For example, the choice of the conditional independence (CI)
+		test being used for constraint-based method depends on the
+		dataset, similarly for the scoring method for score-based
+		methods. With no obvious way to test the impact/performance of
+		these choices on a given dataset, it becomes quite hard to
+		chose the optimal algorithm for a given dataset.
 \end{enumerate}
 
-On the other hand, in the field of Structural Equation Models (SEMs), a lot
-more focus has been put on methods to help researchers draw models manually
-along with integrating their expertise. Researchers typically come up with an
-initial model based on their domain knowledge, and then use various tools to
-guide them in making modifications to this initial model. This process of
-constructing models is commonly known as \emph{Specification Search}. A
-commonly used method for suggesting potential modifications to the SEM is
-\emph{modification indices}. Given a specified model, modification indices
-tries to find potential modifications to the model that improve the model's fit
-to the data best. Based on the suggestions from this method along with their
-domain expertise, the researcher can then choose the most optimal modification
-for their model.
+As a result of these challenges, no matter if the researcher is manually
+constructing the DAG or using an algorithm, they need to utilize their domain
+expertise to construct and modify models. One potential way to through
+conditional independence (CI) based testing. As each missing edge in the DAG
+implies a CI statement, that can be tested in our data \ref{fig:ci_testing}.
+This however has two main drawbacks: 
+\begin{enumerate}
+	\item As CIs are implied by only missing edges, we can not test if an
+		incorrect edge is present in the model.
+	\item For larger graphs, the number of CI tests can be very large, for example,
+		the alarm network with $ 37 $ nodes has $ 287 $ implied CIs. As
+		there is no way to rank these CIs, it can be tedious to
+		consider all these statements.
+\end{enumerate}
 
+% As it is evident that researchers also draw DAGs by hand and even if they use
+% an algorithm, the need to make some post-hoc adjustments to the model, in this
+% paper we present a tool to assist researchers in making these adjustments
+% somewhat inspired by modification indices. In the case of DAGs, global testing
+% methodology does exist such as Shipley test, that can combine local structure
+% tests into a single p-value test. However, this does not give us a way to make
+% modifications to our model. A possible way to do this by checking which of
+% these local tests are violated. However, this approach can quickly become
+% diffult as the size of the network increases. For example, the popular alarm
+% network implies $x$ condition independence statements.
+
+On the other hand, in Structural Equation Models (SEMs) literature, more
+emphasis has been placed on helping researchers manually construct models while
+integrating their expertise. Researchers typically develop an initial model
+based on domain knowledge and use various tools to guide modifications to
+improve the fit of their model to data while it still making theoretical sense.
+This model construction process is known as \emph{Specification Search}. A
+common method for specification search is \emph{modification indices}, which
+identify potential changes to the model structure that improve the model fit
+and ranks them using the most improvement to the model fit. Researchers then
+use these suggestions and their expertise to choose the most optimal
+modifications.
+
+\todo[inline]{Show a figure of how the output looks like for modification indices}
 \todo[inline]{Other methods: Modification indices and Lagrange Multiplier tests to
@@
 
-As it is evident that researchers also draw DAGs by hand and even if they use
-an algorithm, the need to make some post-hoc adjustments to the model, in this
-paper we present a tool to assist researchers in making these adjustments
-somewhat inspired by modification indices. In the case of DAGs, global testing
-methodology does exist such as Shipley test, that can combine local structure
-tests into a single p-value test. However, this does not give us a way to make
-modifications to our model. A possible way to do this by checking which of
-these local tests are violated. However, this approach can quickly become
-diffult as the size of the network increases. For example, the popular alarm
-network implies $x$ condition independence statements.
-
 % One of the drawbacks of these global fit metrics is that they do not only test
 % for the structure but also the fitted parameterization.
+
+Inspired by this ranking approach, in this paper we present a way to rank the CI 
+testing using a measure of association. This can be interpreted as the edge
 
@@
 	\item Using this association measure, we present a procedure to suggest
-		potential new edges, or edge removals in the model. The user
-		gets to decide the direction of the new edge based on their
-		expertise.
+		modifications (i.e., new edges, and deleting edges) to a given
+		model such that the model can better explain the observed
+		associations in the data. 
 	\item Lastly, we present a web-browser based tool to allow researchers
@@
 		\caption{\todo[inline]{Show a full example of testing using this model}}
+		\label{fig:ci_testing}
 \end{figure}
@@
 
-\section{Background}
+\section{Background and Related Work}
 \label{sec:background}
-We represent uni-dimensional variables as $ X $ and multi-dimensional variables
-as $ \bm{Z} $. All the variables are considered to be any mix of continuous,
-ordinal, or categorical unless specified explicitly.
-
-We consider the case of estimating association between $ X $ and $ Y $ is the
-presence of a set of conditioning variables $ \bm{Z} = Z_1, \cdots, Z_k $,
-potentially being an empty set $ \bm{Z} = \emptyset $. We use $ \rvert \bm{Z}
-\rvert $ to denote the cardinality of the set of variables $ \bm{Z} $. We
-consider various scenarios where the variables $ X $, $ Y $ and $ \bm{Z} $
-could be either discrete, ordinal, or continuous. With mixed data, we mean an
-arbitrary mix of variable types among $ X $, $ Y $, and $ \bm{Z} $.
+We consider random variable $ X $ and $ Y $ with a possible multi-dimensional
+variable $ \bm{Z} = \{ Z_1, \cdots, Z_k \} $. We consider these variables in
+the mixed data setting where these variables can be any combination of
+continuous, categorical, or ordinal unless specified. We use $ \rvert \bm{Z}
+\rvert $ to represent the cardinality of the variable $ \bm{Z} $. We write $
+\bm{x} = (x_1, \cdots, x_n) $ for a sample from $ X $ of size $ n $. We write
+the expectation of a variable $ X $ as $ \mathbb{E}[X] $, conditional
+expectations as $ \mathbb{E}[X | \bm{Z}] $.
 
 We consider the problem of estimating the partial (or conditional) association
-between $ X $ and $ Y $ in the presence of the conditioning set $ Z $. When $ Z
-= \emptyset $, this is equivalent to the marginal association between $ X $ and
-$ Y $.
+between $ X $ and $ Y $ in the presence of a set of conditioning variables $
+\bm{Z} $. When $ Z = \emptyset $, this is equivalent to the marginal
+association between $ X $ and $ Y $.
 
@@
 
-In this section, we develop a novel (conditional) measure of association for
+In this section, we develop a novel partial measure of association for
 mixed data by extending the idea of partial correlation to mixed data. We do
@@
 conditional variables $ \bm{Z} $, we are interested in computing an estimate of
-the conditional association, $ \rho(X, Y; Z) $. If $ \bm{Z} = \emptyset $, we
-use a constant value as $ Z $. Firstly, we compute the residuals for $ x $ and
+the conditional association, $ \rho(X, Y; \bm{Z}) $. If $ \bm{Z} = \emptyset $, we
+use a constant value as $ Z $. Firstly, we compute the residuals for $ \bm{x} $ and
 $ y $ using the $ \bm{z} $ as the predictor variables. We use the mixed data
@@
 
+Using this mixed data partial association measure, we propose an iterative DAG
+modification / construction procedure. Users can start with an empty graph.
+Given a DAG $ G $ and a dataset $ D $, for each pair for variable $ X $ and $ Y
+$ in $ G $ that does not have an edge between them, we compute the partial
+association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. Additionally, for
+each pair of variable in $ G $, that has an edge between them in we compute the
+partial association measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup
+pa_{\underline{G}}(Y) $. In the case of no edge, this measure of association
+can be interpreted as the unexplained
+
 Any two variables in a DAG that do not have an edge between them imply a
@@
 
-The idea behind using the measure of association for modifying DAGs is inspired
-by modification indices in SEM.
-
-Theorem - If we have access to an oracle that can give all correct results, 
-the algorithm above is consistent.
+\begin{theorem}
+If we have access to an oracle that can give correct edge directions, the
+algorithm above is consistent.
+\end{theorem}
 

### 2024-06-26 paper/main.tex
Updates

@@
 
+% Add that we need a p-value as well.
+
 Using this mixed data partial association measure, we propose an iterative DAG
@@
 $ in $ G $ that does not have an edge between them, we compute the partial
-association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. Additionally, for
-each pair of variable in $ G $, that has an edge between them in we compute the
-partial association measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup
-pa_{\underline{G}}(Y) $. In the case of no edge, this measure of association
-can be interpreted as the unexplained
+association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. The measure of
+assocation in this case can be interpreted as the unexplained observed
+association between them given the current model. We can then use this to rank
+pair of variables which have very high unexplained association between them.
+Based on the domain expertise, the user can then choose one of these suggested
+pair of variables, decide the direction of the edge between the variables, and
+add it to the model. Once the new edge is added we can recompute the assocation
+of all other variables and $ X $ and $ Y $.
 
-Any two variables in a DAG that do not have an edge between them imply a
-conditional independence test that is they are independent given the parents of
-both the variables, $ X \ci Y | pa(X) \cup pa(Y) $. In the case of model
-testing, a CI statement like this can be used to check whether the assumption
-of the model holds in the data or is violated. For mixed data, a few different
-CI tests have been proposed such as likelihood-ratio based, and residualization
-based. However, if we just use the CI test to check the p-value we cannot use
-it to find the modification that is able to improve the score the most. We can
-however use the measure of association to check which assocation between the
-variables in the data is not explained by the model. Using this the expert can
-decide to add an edge between the variables but will need to choose the
-direction of the edge.
+Additionally, in each iteration of modification, for each pair of variable in $
+G $, that has an edge between them in we compute the partial association
+measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $. This
+measure of assocation can be interpreted as the strength of the edge between
+the variables. If the value of this assocation below a certain threshold, we
+can remove the edge.
 
 \begin{theorem}
-If we have access to an oracle that can give correct edge directions, the
-algorithm above is consistent.
+	If we have access to an oracle that can give correct edge directions, taking a
+	greedy approach with the produre above results in the correct graph
+	being recovered.
 \end{theorem}
@@
 \label{sec:web}
-Based on the approach outlined in the previous section, we built a web-based
-tool that researchers can upload their dataset to and apply this method to
-interactively draw models. Users can start by uploading their datasets. The web
-interface should then automatically show them all the correlated variables in a
-graphical representation. The red colored edges represent potential new edge
-between the variables.
-
-\todo[inline]{List some of the features of the online platform}
+Based on the procedure outlined in the previous section, we built a web-based
+tool that allows users to interactively create/modify their model. The user
+starts by uploading their dataset and optionally the initial DAG. If no initial
+DAG is provided an empty graph is initialized. The tool also computes the
+Shipley's C value at any point. The web-tool would automatically show all
+unexplained assocations in the DAG by adding red edges between them. The edge
+weight of the edge represents how high this strength is. Users can also set the
+threshold for p-value and association to control how many edges are shown at a
+time. Based on these shown potential edges, the user can select the edge that
+they want to add and decide the direction for it. If the tool detects an
+incorrect edge in the graph, it is shown in black. Once the user is happy with
+the graph they have they can export it into different formats to use it for
+further computation.
 

### 2024-06-26 code/api.R
Remove webapp related files

@@ -1,48 +1,0 @@
-source('test_dag.R')
-
-#* @filter cors
-cors <- function(res) {
-    res$setHeader("Access-Control-Allow-Origin", "*") # Or whatever
-    plumber::forward()
-}
-
-
-#* Return the sum of two numbers
-#* @param dag The dag to compute the marginals on.
-#* @param d:file
-#* @post /marginal
-#* @preempt cors
-#* @parser multi
-#* @parser csv
-function(dag, d){
-	dag <- dagitty::dagitty(dag)
-	marginal_cor <- compute_effects_marg(dag=dag, data=as.data.frame(d[[1]]))
-	marginal_cor$cor <- round(as.double(marginal_cor$cor), 3)
-	return(marginal_cor)
-}
-
-
-#* Return the sum of two numbers
-#* @param dag The dag to compute the conditional correlations on.
-#* @param d:file
-#* @post /conditional
-#* @parser multi
-#* @parser csv
-#* @preempt cors
-function(dag, d){
-	dag <- dagitty::dagitty(dag)
-	cond_cor <- compute_effects_cond(dag=dag, data=as.data.frame(d[[1]]))
-	cond_cor$cor <- round(as.double(cond_cor$cor), 3)
-	return(cond_cor)
-}
-
-
-#* @get /simpletest
-function(){
-	dag <- dagitty::dagitty("X->M->Y")
-	marginal_cor <- compute_effects_marg(dag=dag, 
-	data=simulateSEM("X->M->Y",b.default=.4,empirical=TRUE))
-	marginal_cor$cor <- round(as.double(marginal_cor$cor), 3)
-	return(marginal_cor)
-}
-

### 2024-06-26 code/run_api.R

@@ -1,5 +1,0 @@
-# Runs the web server to accept requests. Available at: http://localhost:8000
-
-library(plumber)
-
-pr("api.R") %>% pr_run(port=8000)

### 2024-06-27 paper/main.tex
Updates

@@
 
-In this section, we develop a novel partial measure of association for
-mixed data by extending the idea of partial correlation to mixed data. We do
-this by combining a mixed-data residualization method with a multivariate
-measure of association. Given two variables $ X $ and $ Y $, and a set of
-conditional variables $ \bm{Z} $, we are interested in computing an estimate of
-the conditional association, $ \rho(X, Y; \bm{Z}) $. If $ \bm{Z} = \emptyset $, we
-use a constant value as $ Z $. Firstly, we compute the residuals for $ \bm{x} $ and
-$ y $ using the $ \bm{z} $ as the predictor variables. We use the mixed data
-residualization approach from \citet{Ankan2023}. We consider each variable type
-separately and repeat the following steps for both variables $ X $ and $ Y $:
+In this section, we define a novel partial measure of association for mixed
+data by extending the idea of partial correlation to mixed data. We do this by
+combining a mixed-data residualization method with a multivariate measure of
+association. Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $,
+and $ \bm{Z} $, we are interested in computing an estimate of the conditional
+association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $ conditioned on $
+\bm{Z} $. As the residualization approach requires atleast one conditional
+variable to be present, if $ \bm{Z} = \emptyset $, we use a constant value for
+$ Z $. Firstly, we compute the residuals for $ X $ and $ Y $ using the $ \bm{Z}
+$ as the predictor variables. We use the mixed data residualization approach
+from \citet{Ankan2023}. The residual computation depends on the type of 
+variable:
 

### 2024-06-27 paper/main.tex
Updates

@@
 \emph{Causal Discovery} methods aim to uncover the underlying causal structure
-among random variables from a given dataset. The problem of causal discovery
-has been explored in both the Directed Acyclic Graphs (DAGs) and Structural
-Equation Models (SEM) literatures, each with different approaches. In the DAG
-literature, the focus has been on developing automated algorithms that can
-learn the causal structure from observational or interventional dataset. In
-contrast, the SEM literature emphasizes tools that help researchers manually
-construct or modify modify.
+among random variables from a given observational dataset. The problem of
+causal discovery has been explored in both the Directed Acyclic Graphs (DAGs)
+and Structural Equation Models (SEMs) literatures, each with different
+approaches. In the DAG literature, the focus has been on developing automated
+algorithms that can learn the causal structure from observational datasets. In
+contrast, the SEM literature focuses on tools that help researchers manually
+construct or modify models.
 
@@
 		mistakes, for example, the constraint-based algorithms are
-		biased towards sparse graphs \citep{Malinsky2024}. Consequently,
-		researchers often have rely on post-hoc modifications based on 
-		their domain knowledge or model testing methods.
+		biased towards sparse graphs \citep{Malinsky2024}.
+		Consequently, researchers often have rely on post-hoc
+		modifications based on model testing methods and their domain
+		knowledge.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple causal
@@
 		causal effect estimation, assume knowledge of a fully oriented
-		DAG. As a results, researchers have to rely on their expertise
+		DAG. As a result, researchers have to rely on their expertise
 		to orient these edges or make further assumptions to make use
@@
 
-As a result of these challenges, no matter if the researcher is manually
-constructing the DAG or using an algorithm, they need to utilize their domain
-expertise to construct and modify models. One potential way to through
-conditional independence (CI) based testing. As each missing edge in the DAG
-implies a CI statement, that can be tested in our data \ref{fig:ci_testing}.
-This however has two main drawbacks: 
+As a result of these challenges, doesn't matter if the researcher is manually
+constructing the DAG or using an algorithm, they need to do some manual
+modifications to the model before being able to use them for analysis. One of
+the common ways to guide this modification is to use Conditional Independence
+(CI) based model testing \cite{Ankan2021}. As DAGs imply CI statements that can
+be tested in the given data, and then model can be modified based on the
+results of this test. Figure~\ref{fig:ci_testing} shows an example of testing
+using CIs. This approach has two main drawbacks:
 \begin{enumerate}
-	\item As CIs are implied by only missing edges, we can not test if an
-		incorrect edge is present in the model.
-	\item For larger graphs, the number of CI tests can be very large, for example,
-		the alarm network with $ 37 $ nodes has $ 287 $ implied CIs. As
-		there is no way to rank these CIs, it can be tedious to
-		consider all these statements.
+	\item CIs are implied by missing edges only, and hence we can not test
+		whether an incorrect edge is present in the model.
+	\item For larger graphs, the number of CI tests can be very large, for
+		example, the alarm network with $ 37 $ nodes has $ 287 $
+		implied CIs. It can be difficult to look at all these tests for
+		violations.
 \end{enumerate}
@@
 identify potential changes to the model structure that improve the model fit
-and ranks them using the most improvement to the model fit. Researchers then
-use these suggestions and their expertise to choose the most optimal
-modifications.
+and ranks them using the most improvement to the model fit. This ranking allows
+the researchers to focus first on the suggestions that can improve the model
+the most and can use their expertise to choose the most optimal modifications.
 

### 2024-06-27 code/llm.py
Updates

@@ -35,7 +35,7 @@ def query_direction(u, v, descriptions):
 
 
 def preprocess_data():
-    df = pd.read_csv("../work/2024-human-sl/code/adult_proc.csv", index_col=0)
+    df = pd.read_csv("adult_proc.csv", index_col=0)
     df.Age = pd.Categorical(
         df.Age,
         categories=["<21", "21-30", "31-40", "41-50", "51-60", "61-70", ">70"],
@@ -98,7 +98,7 @@ def test_all(dag, data):
 
     return pd.DataFrame(cis, columns=['u', 'v', 'z', 'edge_present', 'effect', 'p_val'])
 
-def simulate_human(data, descriptions, pval_thres=0.1, effect_thres=0.05):
+def simulate_human(data, descriptions, pval_thres=0.1, effect_thres=0.1):
     nodes = list(data.columns)
     dag = DAG()
     dag.add_nodes_from(nodes)

### 2024-06-27 code/llm.py
Updated llm script

@@ -98,7 +98,7 @@ def test_all(dag, data):
 
     return pd.DataFrame(cis, columns=['u', 'v', 'z', 'edge_present', 'effect', 'p_val'])
 
-def simulate_human(data, descriptions, pval_thres=0.1, effect_thres=0.1):
+def simulate_human(data, descriptions, pval_thres=0.05, effect_thres=0.05):
     nodes = list(data.columns)
     dag = DAG()
     dag.add_nodes_from(nodes)
@@ -115,23 +115,20 @@ def simulate_human(data, descriptions, pval_thres=0.1, effect_thres=0.1):
 
         nonedge_effects = all_effects[all_effects.edge_present == False]
         nonedge_effects = nonedge_effects[(nonedge_effects.effect >= effect_thres) & (nonedge_effects.p_val <= pval_thres)]
-        add_edges = list(nonedge_effects.loc[:, ('u', 'v')].to_records(index=False))
-        add_edges_directed = []
-        print(f"Adding: {add_edges}")
-
-        for edge in add_edges:
-            add_edges_directed.append(query_direction(edge[0], edge[1], descriptions))
-        dag.add_edges_from(add_edges_directed)
 
         if (edge_effects.shape[0] == 0) and (nonedge_effects.shape[0] == 0):
             break
+
+        selected_edge = nonedge_effects.iloc[nonedge_effects.effect.argmax()]
+        print(f"Adding: {selected_edge.u} -- {selected_edge.v}")
+        dag.add_edges_from([query_direction(selected_edge.u, selected_edge.v, descriptions)])
     return(dag)
 
 descriptions = {'Age': 'The age of a person',
                 'Workclass': 'The workplace where the person is employed such as Private industry, or self employed',
                 'Education': "The highest level of education the person has finished.",
                 'MaritalStatus': "The marital status of a person",
-                "Occupation": "The kind of job the person does. For example, sales, craft repair, clerkial",
+                "Occupation": "The kind of job the person does. For example, sales, craft repair, clerical",
                 "Relationship": "The relationship status of the person",
                 "Race": "The ethnicity of the person",
                 "Sex": "The sex or gender of the person",

### 2024-06-27 code/llm.py
Updates

@@ -98,7 +98,7 @@ def test_all(dag, data):
 
     return pd.DataFrame(cis, columns=['u', 'v', 'z', 'edge_present', 'effect', 'p_val'])
 
-def simulate_human(data, descriptions, pval_thres=0.05, effect_thres=0.05):
+def simulate_human(data, descriptions, pval_thres=0.05, effect_thres=0.1):
     nodes = list(data.columns)
     dag = DAG()
     dag.add_nodes_from(nodes)

### 2024-06-27 paper/main.tex
Updates

@@
 		asymptotic theory proving them to be consistent, their finite
-		sample properties are not well understood. In real-world
-		settings with finite samples, these algorithms can make obvious
-		mistakes, for example, the constraint-based algorithms are
-		biased towards sparse graphs \citep{Malinsky2024}.
-		Consequently, researchers often have rely on post-hoc
-		modifications based on model testing methods and their domain
-		knowledge.
+		sample properties are not well understood. Moreover, depending
+		on the selected hyperparameters of the algorithms, the output
+		can vary very significantly. In real-world settings with finite
+		samples, these algorithms can make obvious mistakes, for
+		example, the constraint-based algorithms are biased towards
+		sparse graphs \citep{Malinsky2024}. Consequently, researchers
+		often have rely on post-hoc modifications based on model
+		testing methods and their domain knowledge.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple causal
@@
 		of pairwise orientation techniques.
-	\item \textbf{Algorithm and hyperparameter selection:} Each algorithm
-		has the same goal but they can learn very different models
-		depending on the choice of algorithm and their hyperparameters.
-		For example, the choice of the conditional independence (CI)
-		test being used for constraint-based method depends on the
-		dataset, similarly for the scoring method for score-based
-		methods. With no obvious way to test the impact/performance of
-		these choices on a given dataset, it becomes quite hard to
-		chose the optimal algorithm for a given dataset.
+	% \item \textbf{Algorithm and hyperparameter selection:} Each algorithm
+	% 	has the same goal but they can learn very different models
+	% 	depending on the choice of algorithm and their hyperparameters.
+	% 	For example, the choice of the conditional independence (CI)
+	% 	test being used for constraint-based method depends on the
+	% 	dataset, similarly for the scoring method for score-based
+	% 	methods. With no obvious way to test the impact/performance of
+	% 	these choices on a given dataset, it becomes quite hard to
+	% 	chose the optimal algorithm for a given dataset.
 \end{enumerate}
@@
 be tested in the given data, and then model can be modified based on the
-results of this test. Figure~\ref{fig:ci_testing} shows an example of testing
+p-values of these test. Figure~\ref{fig:ci_testing} shows an example of testing
 using CIs. This approach has two main drawbacks:
@@
 the researchers to focus first on the suggestions that can improve the model
-the most and can use their expertise to choose the most optimal modifications.
+the most and while integrating their expertise to choose the most optimal
+modifications.
 
@@
 critical ratios) for removing edges.}
-
-% One of the drawbacks of these global fit metrics is that they do not only test
-% for the structure but also the fitted parameterization.
 
@@
 		\end{tikzpicture}
-		\caption{\todo[inline]{Show a full example of testing using this model}}
+		\caption{\todo[inline]{Finish figure to show an example of model testing}}
 		\label{fig:ci_testing}
+\end{figure}
+\begin{figure}
+	\caption{\todo[inline]{Placeholder for figure showing an example of modification indices}}
 \end{figure}
@@
 present our method for computing mixed data association. In
-Section~\ref{sec:modification}, we present how to use this method to rank the
-modifications, and lastly in Section~\ref{sec:empirical}, we show some
-empirical analysis on how well this modification process works.
+Section~\ref{sec:modification} and Section~\ref{sec:}, we present how to use
+this method to make modifications and the web tool, and lastly in
+Section~\ref{sec:empirical}, we show some empirical analysis on how well this
+modification process works.
 
@@
 
-\section{Measure of Association for Mixed Data}
+\section{A Measure of Association for Mixed Data}
 \label{sec:mixed_association}
@@
 
-\section{Using measure of association for DAG construction}
+\section{Using Measure of Association for DAG Modification and Construction}
 \label{sec:modification}
@@
 Using this mixed data partial association measure, we propose an iterative DAG
-modification / construction procedure. Users can start with an empty graph.
-Given a DAG $ G $ and a dataset $ D $, for each pair for variable $ X $ and $ Y
-$ in $ G $ that does not have an edge between them, we compute the partial
-association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. The measure of
-assocation in this case can be interpreted as the unexplained observed
-association between them given the current model. We can then use this to rank
-pair of variables which have very high unexplained association between them.
-Based on the domain expertise, the user can then choose one of these suggested
-pair of variables, decide the direction of the edge between the variables, and
-add it to the model. Once the new edge is added we can recompute the assocation
-of all other variables and $ X $ and $ Y $.
+modification procedure. 
+\todo[inline]{How similar is this to GES algorithm}
+If the user desires to construct the model manually, they can also start with
+an empty graph. Given a DAG $ G $ and a dataset $ D $, for each pair for
+variable $ X $ and $ Y $ in $ G $ that does not have an edge between them, we
+compute the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $.
+The measure of association in this case can be interpreted as the unexplained
+observed association between them given the current model. We can then use this
+to rank pair of variables which have very high unexplained association between
+them. Based on the domain expertise, the user can then choose one of these
+suggested pair of variables, decide the direction of the edge between the
+variables, and add it to the model. Once the new edge is added we can recompute
+the association of all other variables and $ X $ and $ Y $.
 
@@
 measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $. This
-measure of assocation can be interpreted as the strength of the edge between
-the variables. If the value of this assocation below a certain threshold, we
+measure of association can be interpreted as the strength of the edge between
+the variables. If the value of this association below a certain threshold, we
 can remove the edge.
@@
 	If we have access to an oracle that can give correct edge directions, taking a
-	greedy approach with the produre above results in the correct graph
+	greedy approach with the procedure above results in the correct graph
 	being recovered.
@@
 Shipley's C value at any point. The web-tool would automatically show all
-unexplained assocations in the DAG by adding red edges between them. The edge
+unexplained associations in the DAG by adding red edges between them. The edge
 weight of the edge represents how high this strength is. Users can also set the

### 2024-06-30 paper/main.tex
Updates

@@
 goal in many scientific questions as it reveals the underlying mechanisms of
-the observed phenomena and informs effective interventions or policy decisions.
+observed phenomena and informs effective interventions or policy decisions.
 \emph{Causal Discovery} methods aim to uncover the underlying causal structure
@@
 causal discovery has been explored in both the Directed Acyclic Graphs (DAGs)
-and Structural Equation Models (SEMs) literatures, each with different
-approaches. In the DAG literature, the focus has been on developing automated
-algorithms that can learn the causal structure from observational datasets. In
-contrast, the SEM literature focuses on tools that help researchers manually
-construct or modify models.
+and Structural Equation Models (SEMs) literatures, each adopting different
+approaches. In the DAG literature, the focus has primarily been on developing
+automated algorithms that can learn the causal structure from observational
+datasets. In contrast, the SEM literature has focused more towards developing
+tools that assist researchers in manually constructing or modifying models.
 
@@
 
-In the DAG literature, many automated algorithms have been developed for causal
-discovery such as constraint based methods like PC \citep{Spirtes2001} and FCI
-\citep{Spirtes2000}, score-based methods like Hill-Climb Search, Greedy
-Equivalence Search \citep{Chickering2002}, and optimization based methods.
-However, the adaption of these algorithms in applied fields remain limited
-\citep{Tennant2020}. Researchers in applied fields typically prefer to
-construct DAGs manually based on their domain knowledge. We hypothesize that
-this preference towards manual DAG construction stems from the following
-challenges in applying causal discovery algorithms:
+As a result of these efforts, numerous approaches to automated algorithms for
+causal discovery have been proposed in the DAG literature, including
+constraint-based method like the PC algorithm \citep{Spirtes2001} and Fast
+Causal Inference (FCI) \citep{Spirtes2000}), score-based methods like
+Hill-Climb Search and Greedy Equivalence Search (GES) \citep{Chickering2002},
+and continuous optimization-based methods like NO TEARS \citep{Zheng2018} and
+DAGMA \citep{Bello2022}. However, the adaption of these algorithms in applied
+fields remain limited, and researchers typically prefer to construct DAGs
+manually based on their domain expertise \citep{Tennant2020, Petersen2021}.
+Even when using automated algorithms, users often rely on manual post hoc
+modifications to the model learned by the algorithm because they:
 
 \begin{enumerate}
-	\item \textbf{Trustworthiness: } Although most algorithms have
-		asymptotic theory proving them to be consistent, their finite
-		sample properties are not well understood. Moreover, depending
-		on the selected hyperparameters of the algorithms, the output
-		can vary very significantly. In real-world settings with finite
-		samples, these algorithms can make obvious mistakes, for
-		example, the constraint-based algorithms are biased towards
-		sparse graphs \citep{Malinsky2024}. Consequently, researchers
-		often have rely on post-hoc modifications based on model
-		testing methods and their domain knowledge.
-	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple causal
-		structures can be faithful to a given observational dataset.
-		Thus, using only observational data, these algorithms can only
-		recover the MECs that can contain unoriented edges. However,
-		most methods for downstream tasks, such as identification or
-		causal effect estimation, assume knowledge of a fully oriented
-		DAG. As a result, researchers have to rely on their expertise
-		to orient these edges or make further assumptions to make use
-		of pairwise orientation techniques.
-	% \item \textbf{Algorithm and hyperparameter selection:} Each algorithm
-	% 	has the same goal but they can learn very different models
-	% 	depending on the choice of algorithm and their hyperparameters.
-	% 	For example, the choice of the conditional independence (CI)
-	% 	test being used for constraint-based method depends on the
-	% 	dataset, similarly for the scoring method for score-based
-	% 	methods. With no obvious way to test the impact/performance of
-	% 	these choices on a given dataset, it becomes quite hard to
-	% 	chose the optimal algorithm for a given dataset.
+	\item \textbf{Make Obvious Mistakes:} Although most algorithms are
+		theoretically consistent asymptotically, their finite sample
+		properties are not well understood. Moreover, depending on the
+		selected hyperparameters, their output can vary significantly.
+		These issues combined with no good method to test the
+		performance of these algorithms on a given dataset, in
+		real-world settings with finite samples, these algorithms can
+		make obvious mistakes. Consequently, users have to rely on
+		posthoc modifications to the model based on their domain
+		expertise.
+	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
+		DAGs can be faithful to a given observational dataset. Thus,
+		using only observational data, algorithms can only recover the
+		MECs that can contain a combination of directed and undirected
+		edges. However, most methods for downstream tasks, such as
+		identification or causal effect estimation, assume knowledge of
+		a fully oriented DAG. As a result, users have to manually
+		orient these edges before using them for analysis.
 \end{enumerate}
 
-As a result of these challenges, doesn't matter if the researcher is manually
-constructing the DAG or using an algorithm, they need to do some manual
-modifications to the model before being able to use them for analysis. One of
-the common ways to guide this modification is to use Conditional Independence
+Due to these challenges, whether constructing the model manually or using an
+algorithm, users must rely on their domain expertise to construct or modify the
+model. Hence, it is crucial to have tools to guide this modification process.
+One of the ways to guide this modification is to use Conditional Independence
 (CI) based model testing \cite{Ankan2021}. As DAGs imply CI statements that can
-be tested in the given data, and then model can be modified based on the
-p-values of these test. Figure~\ref{fig:ci_testing} shows an example of testing
-using CIs. This approach has two main drawbacks:
-\begin{enumerate}
-	\item CIs are implied by missing edges only, and hence we can not test
-		whether an incorrect edge is present in the model.
-	\item For larger graphs, the number of CI tests can be very large, for
-		example, the alarm network with $ 37 $ nodes has $ 287 $
-		implied CIs. It can be difficult to look at all these tests for
-		violations.
-\end{enumerate}
+be tested in the given data using statistical CI tests, we can verify whether
+the implied CIs are consistent in the data or not. If some of implied CIs are
+not valid in the given dataset, we can modify the model accordingly.
+Figure~\ref{fig:ci_testing} shows an example of model testing using CIs.
+However, this CI testing approach has a few drawbacks. Firstly, as CIs are
+implied only by missing edges in the model, we can not test whether an
+incorrect edge is already present in the model. Secondly, the number of implied
+CIs can be quite large. For example, in the moderately sized alarm network with 
+$ 37 $ node, there are $287$ implied CIs. If we rely only on the p-values of the
+CI tests, it can be difficult to figure which test to focus on.
 
@@
 On the other hand, in Structural Equation Models (SEMs) literature, more
-emphasis has been placed on helping researchers manually construct models while
-integrating their expertise. Researchers typically develop an initial model
-based on domain knowledge and use various tools to guide modifications to
-improve the fit of their model to data while it still making theoretical sense.
-This model construction process is known as \emph{Specification Search}. A
-common method for specification search is \emph{modification indices}, which
-identify potential changes to the model structure that improve the model fit
-and ranks them using the most improvement to the model fit. This ranking allows
-the researchers to focus first on the suggestions that can improve the model
-the most and while integrating their expertise to choose the most optimal
-modifications.
+emphasis has been placed on developing tools to help researchers manually
+construct models while integrating their domain expertise. Researchers
+typically develop an initial model based on domain knowledge and use various
+tools to guide modifications to improve the fit of their model to data. This
+model construction process is known as \emph{Specification Search}. A common
+method for specification search is using \emph{modification indices}, which
+identify potential changes to the model structure that improve the model fit.
+These potential changes are then ranked by the ones that improve the fit the
+most. This ranking allows the researchers to focus first on the suggestions
+that can improve the model the most and while integrating their expertise to
+choose the most optimal modifications.
 
-\todo[inline]{Show a figure of how the output looks like for modification indices}
-\todo[inline]{Other methods: Modification indices and Lagrange Multiplier tests to
-selectively add edges And using Wald test or z statistics (also known as
-critical ratios) for removing edges.}
-
-Inspired by this ranking approach, in this paper we present a way to rank the CI 
-testing using a measure of association. This can be interpreted as the edge
-
-Our proposed approach uses a measure of association to find correlated variables
-in the data that are not explained by the current model. Using the measure, we 
-can rank the edge that can lead to the highest explanation of the observed 
-correlation. Based on this information, the user can select a potential edge
-and based on their expert knowledge decide the direction of the edge.
-
-% We present a similar procedure for modifying DAGs based on expert knowledge and
-% measure of association. We first use the measure of association to test whether
-% the association observed in the data is explained by the model structure. We
-% can not directly use the local CI tests to make modifications because that
-% would give way too many independencies and we have no way to rank them. Based
-% on this testing, we suggest adding edges by ranking them by the most
-% unexplained association between variables. The researcher can then use expert
-% knowledge to decide the direction of the edge between the variables.
-
-Our main contributions in this papers are as follows:
+Inspired by this ranking approach in modification indices, in this paper we
+develop a method to rank the CI tests using a conditional measure of
+association. Our proposed approach uses the measure of association to find
+correlated variables in the data that are not explained by the current model.
+Using the measure, we can then rank the pair of variables between which adding
+an edge would improve the explanation the most. Based on this information, the
+user can select a potential pair of variables and based on their expert
+knowledge decide the direction of the edge. Our main contributions in this
+paper are as follows:
 \begin{enumerate}
@@
 \begin{figure}
-	\caption{\todo[inline]{Placeholder for figure showing an example of modification indices}}
+	\caption{\todo[inline]{Figure showing an example of modification indices}}
 \end{figure}
@@
 
-\todo[inline]{Add more notation as they come}
-
 \subsection{Measures of (Conditional) Association}
+\todo[inline]{Complete this section}
 As our DAG construction procedure relies on a measure of conditional
@@
 
-\todo[inline]{Is it even possible}
-
 \paragraph{$ X $ is ordinal and $ Y $ is continuous}
@@
 
-In this section, we define a novel partial measure of association for mixed
+In this section, we develop a novel partial measure of association for mixed
 data by extending the idea of partial correlation to mixed data. We do this by
 combining a mixed-data residualization method with a multivariate measure of
-association. Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $,
-and $ \bm{Z} $, we are interested in computing an estimate of the conditional
-association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $ conditioned on $
-\bm{Z} $. As the residualization approach requires atleast one conditional
-variable to be present, if $ \bm{Z} = \emptyset $, we use a constant value for
-$ Z $. Firstly, we compute the residuals for $ X $ and $ Y $ using the $ \bm{Z}
-$ as the predictor variables. We use the mixed data residualization approach
-from \citet{Ankan2023}. The residual computation depends on the type of 
-variable:
+association. Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y
+$, and $ \bm{Z} $, we are interested in computing an estimate of the
+conditional association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $
+conditioned on $ \bm{Z} $. As the residualization approach requires atleast one
+conditional variable to be present, if $ \bm{Z} = \emptyset $, we use a
+constant value for $ Z $. Similar to partial correlation, we first compute the
+residuals for $ X $ and $ Y $ using $ \bm{Z} $ as the predictor variables. We
+use the mixed data residualization approach from \citet{Ankan2023}. Depending
+on the type of dependent variable, the residuals are defined as follows:
 
 \begin{enumerate}
-	\item \textbf{$ X $ is continuous:} We start by training a regression
-		model, $ E_X: x \sim \bm{z} $. Next we take the predictions
-		from this model $ \hat{x} = E_X(\bm{z}) $ and lastly the
-		residuals are computed as the difference between the true and
-		the predicted values.
-		$$ R_{x_i} = x_i - \hat{x}_i $$
-	\item \textbf{$ X $ is ordinal:} We start by training a probability
+	\item \textbf{Continuous:} We train a regression
+		model, $ E_X: x \sim \bm{z} $. The residuals are then defined
+		as the difference between the true and the predicted value using
+		$ E_X $.
+		$$ R_{x_i} = x_i - E_X(\bm{z}_i) $$
+	\item \textbf{Ordinal:} We start by training a probability
 		estimator, $ p_X: x \sim \bm{z} $, and take probability
@@
 		$$ R_{x_i} = \hat{p}(\hat{x}_i > x_i) - \hat{p}(\hat{x}_i < x_i) $$
-	\item \textbf{$ X $ is categorical:} We again start by training a
+	\item \textbf{Categorical:} We again start by training a
 		probability estimator $ p_X: X \sim \bm{Z} $, and take the
@@
 their data. Non-parametric ensemble estimators such as Random Forest and
-XGBoost can work well on a large variety of data and any data type. For linear
-datasets users can also choose variants of linear regression depending on the
-type of data. After the above steps are repeated for both $ X $ and $ Y $, we
-end up with two residual matrices: $ R_X $ and $ R_Y $. The type of the
-variable determines the shape for these matrices, i.e., if the variable is
-continuous we get a residual matrix of shape $ ( n \times 1 ) $ and if the
-variable is categorical we have a residual matrix of shape $ ( n \times (k-1))
-$ (where one of the dummy encoded variables is dropped to avoid
+XGBoost can work on any combination of data types while performing reasonably
+well. For linear datasets users can also choose variants of linear regression
+depending on the type of data. After the above steps are repeated for both $ X
+$ and $ Y $, we end up with two residual matrices: $ R_X $ and $ R_Y $. The
+type of the variable determines the shape for these matrices, i.e., if the
+variable is continuous we get a residual matrix of shape $ ( n \times 1 ) $ and
+if the variable is categorical we have a residual matrix of shape $ ( n \times
+(k-1)) $ (where one of the dummy encoded variables is dropped to avoid
 multicollinearity)
 
-We now have two continuous valued residual matrices and we need a measure to
-quantify the association between these residuals. In multivariate statistics,
-canonical correlation based measures have been used for measuring association
-between two sets of variables. We can treat the residuals as sets of variables
-in case of categorical variable and use canonical corrleations to quantify the
-association between them.
+We now have two continuous valued residual matrices and the second step now is
+to quantify the association between these residuals. In multivariate
+statistics, canonical correlation based measures have been used for measuring
+association between two sets of variables. We can treat the residuals as sets
+of variables in case of categorical variable and use canonical correlations to
+quantify the association between them.
 \todo[inline]{Justify the argument}
@@
 	and $ \bm{V} = (V_1, V_2, \cdots, V_q) $, canonical correlation between
-	them, $\rho(\bm{U}, \bm{V}) $ is defined as:
+	them, $\phi(\bm{U}, \bm{V}) $ is defined as:
 
 	\begin{equation}
-		\rho(\bm{U}, \bm{V}) = \sup_{a, b} \frac{a^T \Sigma_{\bm{U}\bm{V}} b}{\sqrt{a^T \Sigma_{\bm{U}\bm{U}} a} \sqrt{b^T \Sigma_{\bm{V}\bm{V}} b}}
+		\phi(\bm{U}, \bm{V}) = \sup_{a, b} \frac{a^T \Sigma_{\bm{U}\bm{V}} b}{\sqrt{a^T \Sigma_{\bm{U}\bm{U}} a} \sqrt{b^T \Sigma_{\bm{V}\bm{V}} b}}
 	\end{equation}
@@
 coefficient, i.e., $ 0 $ signifies no association and $ 1 $ signifies perfect
-linear relationship.
+linear relationship. This gives us the following partial measure of association:
 
-\todo[inline]{Add the formula for our case here}
+$$ \rho(X, Y; \bm{Z}) = \sum_{i=1}^{min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\phi_i(R_X, R_Y))^2 $$
 
@@
 		known effect size measure for discrete variables.
-		\todo[inline]{Show the connection}
 \end{enumerate}
+\todo[inline]{Expand on these points}
 
@@
 
+\todo[inline]{Add screenshots here}
+
 % \begin{figure}
@@
 \label{sec:empirical}
+
+\begin{figure}
+	\begin{subfigure}{0.5\textwidth}
+		\centering
+		\includegraphics{../code/plots/shd_ribbon.pdf}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}{0.5\textwidth}
+		\centering
+		\includegraphics{../code/plots/sid_ribbon.pdf}
+		\caption{}
+	\end{subfigure}
+	\caption{Comparison of PC and Hill-Climb Search algorithms against
+		manually drawn DAGs using the assistance method. As PC and
+		Hill-Climb Search return the Markov Equivalence Class (MEC), we
+		use the best and worst scoring orientation of the MEC to get
+		the range of values. The human drawn values are done for
+	multiple accuracy values of the oracle. \todo[inline]{Replace Hill-Climb with GES}}
+\end{figure}
 
@@
 \subsection{Results}
-\begin{figure}
-	\begin{subfigure}{0.5\textwidth}
-		\centering
-		\includegraphics{../code/plots/shd_ribbon.pdf}
-		\caption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-		\centering
-		\includegraphics{../code/plots/sid_ribbon.pdf}
-		\caption{}
-	\end{subfigure}
-	\caption{Comparison of PC and Hill-Climb Search algorithms against
-		manually drawn DAGs using the assistance method. As PC and
-		Hill-Climb Search return the Markov Equivalence Class (MEC), we
-		use the best and worst scoring orientation of the MEC to get
-		the range of values. The human drawn values are done for
-		multiple accuracy values of the oracle.}
-\end{figure}
 
@@
 
+\begin{figure}[ht]
+	\centering
+	\includegraphics[scale=0.3]{../code/plots/llm_adult.png}
+	\caption{\todo[inline]{Change figure to have consistent style}}
+\end{figure}
+
 \section{Conclusions}
 
+\newpage
 \bibliography{references}

### 2024-06-30 paper/main.tex
Updates

@@
 
-Inspired by this ranking approach in modification indices, in this paper we
-develop a method to rank the CI tests using a conditional measure of
-association. Our proposed approach uses the measure of association to find
-correlated variables in the data that are not explained by the current model.
-Using the measure, we can then rank the pair of variables between which adding
-an edge would improve the explanation the most. Based on this information, the
-user can select a potential pair of variables and based on their expert
-knowledge decide the direction of the edge. Our main contributions in this
-paper are as follows:
-\begin{enumerate}
-	\item We present a measure of (conditional) association for mixed data
-		based on canonical correlations. This measure of association
-		generalizes the widely used partial correlation to mixed data
-		setting.
-	\item Using this association measure, we present a procedure to suggest
-		modifications (i.e., new edges, and deleting edges) to a given
-		model such that the model can better explain the observed
-		associations in the data. 
-	\item Lastly, we present a web-browser based tool to allow researchers
-		to apply this method to their own datasets.
-\end{enumerate}
-
 \begin{figure}
@@
 
+Inspired by this ranking approach in modification indices, in this paper we
+develop a method to rank the CI tests using a conditional measure of
+association. Our proposed approach uses the measure of association to find
+correlated variables in the data that are not explained by the current model.
+Using the measure, we can then rank the pair of variables between which adding
+an edge would improve the explanation the most. Based on this information, the
+user can select a potential pair of variables and based on their expert
+knowledge decide the direction of the edge. Our main contributions in this
+paper are as follows:
+\begin{enumerate}
+	\item We present a measure of (conditional) association for mixed data
+		based on canonical correlations. This measure of association
+		generalizes the widely used partial correlation to mixed data
+		setting.
+	\item Using this partial association measure, we present a procedure to
+		rank CI tests. This ranking can in turn then be used to
+		prioritize addition or deletion of edges from the model.
+	\item Lastly, we present a web-browser based tool to allow researchers
+		to apply this method to their own datasets.
+\end{enumerate}
+
 The rest of the paper is structured as follows. In
@@
 the mixed data setting where these variables can be any combination of
-continuous, categorical, or ordinal unless specified. We use $ \rvert \bm{Z}
-\rvert $ to represent the cardinality of the variable $ \bm{Z} $. We write $
-\bm{x} = (x_1, \cdots, x_n) $ for a sample from $ X $ of size $ n $. We write
-the expectation of a variable $ X $ as $ \mathbb{E}[X] $, conditional
-expectations as $ \mathbb{E}[X | \bm{Z}] $.
+continuous, categorical, or ordinal unless specified. We also assume all
+variables to be observed. We use $ \rvert \bm{Z} \rvert $ to represent the
+cardinality of the variable $ \bm{Z} $. We write $ \bm{x} = (x_1, \cdots, x_n)
+$ for a sample from $ X $ of size $ n $. We write the expectation of a variable
+$ X $ as $ \mathbb{E}[X] $, conditional expectations as $ \mathbb{E}[X |
+\bm{Z}] $.
 
@@
 
-Using this mixed data partial association measure, we propose an iterative DAG
-modification procedure. 
-\todo[inline]{How similar is this to GES algorithm}
-If the user desires to construct the model manually, they can also start with
-an empty graph. Given a DAG $ G $ and a dataset $ D $, for each pair for
-variable $ X $ and $ Y $ in $ G $ that does not have an edge between them, we
-compute the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $.
-The measure of association in this case can be interpreted as the unexplained
+Using this mixed data partial association measure, we propose a method to rank
+CI tests. If the user desires to construct the model manually, they can also
+start with an empty graph, or start with a DAG outputted by a causal discovery
+algorithm. Given a DAG $ G $ and a dataset $ D $, for each pair for variable
+$ X $ and $ Y $ in $ G $ that does not have an edge between them, we compute
+the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. The
+measure of association in this case can be interpreted as the unexplained
 observed association between them given the current model. We can then use this
@@
 	greedy approach with the procedure above results in the correct graph
-	being recovered.
+	being recovered. \todo[inline]{Reword and add proof}
 \end{theorem}
@@
 We analyzed the performance of this manual approach by comparing it with two
-automated algorithms: PC and Hill-Climb Search. For the comparison we use
-simulated data from a known DAG and compare how well the algorithm is able to
-recover the original model using two metrics: Structural Hamming Distance (SHD)
-and Structural Intervention Distance (SID). To simulate the data, we start with
-a randomly generated DAG and use linear models with random effects to generate
-the data. We repeat the data generation $ 30 $ times for each density value for
-the DAG. As both PC and Hill-Climb Search can only give us a CPDAG, we compare
-the orientation of these CPDAG which result in the best and worst case scores
-on SHD and SID. This procedure is outlined in Algorithm 1.
+automated algorithms: PC and Greedy Equivalence Search. For the comparison we
+use simulated data from a known DAG and compare how well the algorithm is able
+to recover the original model using two metrics: Structural Hamming Distance
+(SHD) and Structural Intervention Distance (SID). To simulate the data, we
+start with a randomly generated DAG and use linear models with random effects
+to generate the data. We repeat the data generation $ 30 $ times for each
+density value for the DAG. As both PC and Hill-Climb Search can only give us a
+CPDAG, we compare the orientation of these CPDAG which result in the best and
+worst case scores on SHD and SID. This procedure is outlined in Algorithm 1.
+best and worst case scores on SHD and SID. This procedure is outlined in
+Algorithm 1.
 
@@
 
-\begin{figure}[ht]
+\begin{figure}
+	\centering
+	\begin{subfigure}{0.5\textwidth}
+		\caption{\todo[inline]{Add the prompt used}}
+	\end{subfigure}
+	\begin{subfigure}{0.5\textwidth}
 		\centering
@@
 		\caption{\todo[inline]{Change figure to have consistent style}}
+	\end{subfigure}
 \end{figure}

### 2024-07-01 paper/main.tex
Udpates

@@
 manually based on their domain expertise \citep{Tennant2020, Petersen2021}.
-Even when using automated algorithms, users often rely on manual post hoc
+Even when using automated algorithms, we often have to rely on manual post hoc
 modifications to the model learned by the algorithm because they:
@@
 		real-world settings with finite samples, these algorithms can
-		make obvious mistakes. Consequently, users have to rely on
+		make obvious mistakes. Consequently, researchers have to rely on
 		posthoc modifications to the model based on their domain
@@
 		identification or causal effect estimation, assume knowledge of
-		a fully oriented DAG. As a result, users have to manually
+		a fully oriented DAG. As a result, researchers have to manually
 		orient these edges before using them for analysis.
@@
 Due to these challenges, whether constructing the model manually or using an
-algorithm, users must rely on their domain expertise to construct or modify the
+algorithm, we must rely on domain expertise to construct or modify the
 model. Hence, it is crucial to have tools to guide this modification process.
@@
 		\end{tikzpicture}
-		\caption{\todo[inline]{Finish figure to show an example of model testing}}
+		\caption{\todo[inline]{Finish figure. Show an example of model testing using CIs}}
 		\label{fig:ci_testing}
@@
 	\item Using this partial association measure, we present a procedure to
-		rank CI tests. This ranking can in turn then be used to
-		prioritize addition or deletion of edges from the model.
+		rank CI tests. This ranking can be used to prioritize addition
+		or deletion of edges from the model.
 	\item Lastly, we present a web-browser based tool to allow researchers
@@
 
-Here, users have the option to choose the estimators based on the properties of
+Here, we have the option to choose the estimators based on the properties of
 their data. Non-parametric ensemble estimators such as Random Forest and
 XGBoost can work on any combination of data types while performing reasonably
-well. For linear datasets users can also choose variants of linear regression
+well. For linear datasets we can also choose variants of linear regression
 depending on the type of data. After the above steps are repeated for both $ X
@@
 quantify the association between them.
-\todo[inline]{Justify the argument}
 
@@
 Using this mixed data partial association measure, we propose a method to rank
-CI tests. If the user desires to construct the model manually, they can also
-start with an empty graph, or start with a DAG outputted by a causal discovery
-algorithm. Given a DAG $ G $ and a dataset $ D $, for each pair for variable
-$ X $ and $ Y $ in $ G $ that does not have an edge between them, we compute
-the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. The
-measure of association in this case can be interpreted as the unexplained
-observed association between them given the current model. We can then use this
-to rank pair of variables which have very high unexplained association between
-them. Based on the domain expertise, the user can then choose one of these
-suggested pair of variables, decide the direction of the edge between the
-variables, and add it to the model. Once the new edge is added we can recompute
-the association of all other variables and $ X $ and $ Y $.
+CI tests, and use it for model construction and modification. If the user
+desires to construct the model manually, they can also start with an empty
+graph, or start with a DAG outputted by a causal discovery algorithm. Given a
+DAG $ G $ and a dataset $ D $, for each pair for variable $ X $ and $ Y $ in $
+G $ that does not have an edge between them, we compute the partial association
+measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. The measure of association in
+this case can be interpreted as the unexplained observed association between
+them given the current model. We can then use this to rank pair of variables
+which have very high unexplained association between them. Based on the domain
+expertise, we can then choose one of these suggested pair of variables, decide
+the direction of the edge between the variables, and add it to the model. Once
+the new edge is added we can recompute the association of all other variables
+and $ X $ and $ Y $.
 
@@
 Based on the procedure outlined in the previous section, we built a web-based
-tool that allows users to interactively create/modify their model. The user
+tool that allows us to interactively create/modify their model. The user
 starts by uploading their dataset and optionally the initial DAG. If no initial
@@
 		the range of values. The human drawn values are done for
-	multiple accuracy values of the oracle. \todo[inline]{Replace Hill-Climb with GES}}
+	multiple accuracy values of the oracle. \todo[inline]{Replace Hill-Climb with GES and improve the plot}}
 \end{figure}
@@
 We analyzed the performance of this manual approach by comparing it with two
-automated algorithms: PC and Greedy Equivalence Search. For the comparison we
-use simulated data from a known DAG and compare how well the algorithm is able
-to recover the original model using two metrics: Structural Hamming Distance
-(SHD) and Structural Intervention Distance (SID). To simulate the data, we
-start with a randomly generated DAG and use linear models with random effects
-to generate the data. We repeat the data generation $ 30 $ times for each
-density value for the DAG. As both PC and Hill-Climb Search can only give us a
-CPDAG, we compare the orientation of these CPDAG which result in the best and
-worst case scores on SHD and SID. This procedure is outlined in Algorithm 1.
-best and worst case scores on SHD and SID. This procedure is outlined in
-Algorithm 1.
+automated algorithms: PC and Greedy Equivalence Search (GES). For the
+comparison we use simulated data from a known DAG and compare how well the
+algorithm is able to recover the original model using two metrics: Structural
+Hamming Distance (SHD) and Structural Intervention Distance (SID). To simulate
+the data, we start with a randomly generated DAG on $ 10 $ nodes and use linear
+models with random effects to generate the data.
 
-\todo[inline]{Add the algorithm}
+We vary the edge probability of the DAG to generate DAGs with varying density.
+We repeat the data generation $ 30 $ times for each density value for the DAG.
+As both PC and Hill-Climb Search can only give us a CPDAG, we compare the
+orientation of these CPDAG which result in the best and worst case scores on
+SHD and SID.
 
-To simulate an expert using this assisted model construction approach, we start
-with an empty model and take a greedy approach to add edges to this model. At
-any point, we select the pair of variables that has significant correlation and
-the highest unexplained correlation, i.e., highest conditional association. We
-then use an oracle to decide the direction of the edge between the variables.
-Given an accuracy of the oracle, $ \alpha $, the oracle either gives the
-correct direction, incorrect direction, or returns None representing that it
-does not know the direction.
+To use our proposed approach, we need an expert to tell us the direction of the
+edges. We simulated an expert using this assisted model construction approach,
+we start with an empty model and take a greedy approach to add edges to this
+model. At any point, we select the pair of variables that has significant
+correlation and the highest unexplained correlation, i.e., highest conditional
+association. We then use an oracle to decide the direction of the edge between
+the variables. Given an accuracy of the oracle, $ \alpha $, the oracle either
+gives the correct direction, incorrect direction, or returns None representing
+that it does not know the direction.
 
@@
 backtracking to fix its earlier mistakes. This procedure is outlined in the
-Algorithm.
-\todo[inline]{Add the algorithm}
-
-\subsection{Results}
+Algorithm 1.
+\todo[inline]{Show an algorithm of how this expert method is working}
 

### 2024-07-01 code/llm.py

@@ -1,5 +1,6 @@
 import os
 import pandas as pd
+import networkx as nx
 from itertools import combinations
 
 import google.generativeai as genai
@@ -99,11 +100,12 @@ def test_all(dag, data):
 
     return pd.DataFrame(cis, columns=['u', 'v', 'z', 'edge_present', 'effect', 'p_val'])
 
-def simulate_human(data, descriptions, pval_thres=0.05, effect_thres=0.1):
+def simulate_human(data, descriptions, pval_thres=0.05, effect_thres=0.05):
     nodes = list(data.columns)
     dag = DAG()
     dag.add_nodes_from(nodes)
 
+    blacklisted_edges = []
     while(True):
         all_effects = test_all(dag, data)
 
@@ -120,9 +122,19 @@ def simulate_human(data, descriptions, pval_thres=0.05, effect_thres=0.1):
         if (edge_effects.shape[0] == 0) and (nonedge_effects.shape[0] == 0):
             break
 
+        if len(blacklisted_edges) > 0:
+            nonedge_effects = nonedge_effects.loc[((nonedge_effects.u in [edge[0] for edge in blacklisted_edges]) & (nonedge_effects.v in [edge[1] for edge in blacklisted_edges])) or
+                             ((nonedge_effects.u in [edge[1] for edge in blacklisted_edges]) & (nonedge_effects.v in [edge[0] for edge in blacklisted_edges])), :]
+        if len(blacklisted_edges) > 0:
+            import ipdb; ipdb.set_trace()
         selected_edge = nonedge_effects.iloc[nonedge_effects.effect.argmax()]
         print(f"Adding: {selected_edge.u} -- {selected_edge.v}")
-        dag.add_edges_from([query_direction(selected_edge.u, selected_edge.v, descriptions)])
+        edge_direction = query_direction(selected_edge.u, selected_edge.v, descriptions)
+        if nx.has_path(dag, edge_direction[1], edge_direction[0]):
+            print(f"Blacklisting: {edge_direction}")
+            blacklisted_edges.append(edge_direction)
+        else:
+            dag.add_edges_from([edge_direction])
     return(dag)
 
 descriptions = {'Age': 'The age of a person',

### 2024-07-02 paper/main.tex
Updates

@@
 \end{enumerate}
+\todo[inline]{The second point doesn't seem to be adding much}
 
@@
 present our method for computing mixed data association. In
-Section~\ref{sec:modification} and Section~\ref{sec:}, we present how to use
+Section~\ref{sec:modification} and Section~\ref{sec:web}, we present how to use
 this method to make modifications and the web tool, and lastly in

### 2024-08-01 paper/main.tex
updates

@@
 % directly follow a colon or long dash
-\title{Title}
+\title{Human-In-The-Loop Causal Discovery}
 \author{
@@
 
-Understanding cause-and-effect relationships between variables is a fundamental
-goal in many scientific questions as it reveals the underlying mechanisms of
-observed phenomena and informs effective interventions or policy decisions.
-\emph{Causal Discovery} methods aim to uncover the underlying causal structure
-among random variables from a given observational dataset. The problem of
-causal discovery has been explored in both the Directed Acyclic Graphs (DAGs)
-and Structural Equation Models (SEMs) literatures, each adopting different
-approaches. In the DAG literature, the focus has primarily been on developing
-automated algorithms that can learn the causal structure from observational
-datasets. In contrast, the SEM literature has focused more towards developing
-tools that assist researchers in manually constructing or modifying models.
+Understanding the cause-and-effect relationships between variables is a
+fundamental goal in many scientific questions as it reveals the underlying
+mechanisms of observed phenomena and informs effective interventions or policy
+decisions. \emph{Causal Discovery} methods aim to uncover the underlying causal
+structure among random variables from a given observational dataset. The
+problem of causal discovery has been explored in both the Directed Acyclic
+Graphs (DAGs) and Structural Equation Models (SEMs) literatures, each adopting
+different approaches. In the DAG literature, the focus has primarily been on
+developing automated algorithms that can learn the causal structure from
+observational datasets. In contrast, the SEM literature has focused more
+towards developing tools that assist researchers in manually constructing or
+modifying models.
 
@@
 		performance of these algorithms on a given dataset, in
-		real-world settings with finite samples, these algorithms can
-		make obvious mistakes. Consequently, researchers have to rely on
-		posthoc modifications to the model based on their domain
-		expertise.
+		real-world settings with finite samples, it can be difficult to
+		trust the output of these algorithms. Consequently, researchers
+		have to rely on posthoc modifications to the model based on
+		their domain expertise.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
@@
 \end{enumerate}
-\todo[inline]{The second point doesn't seem to be adding much}
+
+\todo[inline]{Add an example here showing these problems}
 
@@
 user can select a potential pair of variables and based on their expert
-knowledge decide the direction of the edge. Our main contributions in this
-paper are as follows:
+knowledge decide the direction of the edge. 
+\todo[inline]{Add about the score based methods somewhere here}
+
+Our main contributions in this paper are as follows:
 \begin{enumerate}

### 2024-08-02 paper/main.tex
Updates

@@
 % directly follow a colon or long dash
-\title{Human-In-The-Loop Causal Discovery}
+\title{Expert-In-The-Loop Causal Discovery: Integrating Expert Knowledge}
 \author{
@@
 mechanisms of observed phenomena and informs effective interventions or policy
-decisions. \emph{Causal Discovery} methods aim to uncover the underlying causal
-structure among random variables from a given observational dataset. The
+decisions. \emph{Causal Discovery} methods aim to uncover these underlying
+cause-effect relationships among random variables from observational data. The
 problem of causal discovery has been explored in both the Directed Acyclic
 Graphs (DAGs) and Structural Equation Models (SEMs) literatures, each adopting
-different approaches. In the DAG literature, the focus has primarily been on
+a different approach. In the DAG literature, the focus has primarily been on
 developing automated algorithms that can learn the causal structure from
-observational datasets. In contrast, the SEM literature has focused more
-towards developing tools that assist researchers in manually constructing or
-modifying models.
+datasets. In contrast, the SEM literature has focused more towards developing
+tools that assist researchers in manually constructing or modifying models.
 
-% To perform any causal effect identification or estimation using Directed
-% Acyclic Graphs (DAGs) or Structural Equation Models (SEMs), researchers need to
-% first construct a causal structure commonly known as DAG or path diagram. This
-% process of constructing the casual structure from data is known as \emph{Causal
-% Discovery}. In the field of DAGs, plenty of algorithms have been developed to
-% automatically construct DAGs from data. These algorithm takes different
-% approaches to causal discovery such as Constraint-based algorithms such as PC
-% and Fast Causal Infenrece (FCI) where the algorithm tries to match the
-% conditional independences in the data to the ones impled by the model,
-% score-based methods such as Hill-Climb Search, Greedy Equivalence Search (GES)
-% where the algorithm tries to find a model that has the best score given some
-% scoring metric, and continuous optimization where the problem is formulated as
-% a constrained optimization problem such as No Tears.
-
-As a result of these efforts, numerous approaches to automated algorithms for
-causal discovery have been proposed in the DAG literature, including
-constraint-based method like the PC algorithm \citep{Spirtes2001} and Fast
-Causal Inference (FCI) \citep{Spirtes2000}), score-based methods like
-Hill-Climb Search and Greedy Equivalence Search (GES) \citep{Chickering2002},
-and continuous optimization-based methods like NO TEARS \citep{Zheng2018} and
-DAGMA \citep{Bello2022}. However, the adaption of these algorithms in applied
-fields remain limited, and researchers typically prefer to construct DAGs
-manually based on their domain expertise \citep{Tennant2020, Petersen2021}.
-Even when using automated algorithms, we often have to rely on manual post hoc
-modifications to the model learned by the algorithm because they:
+As a result of these efforts, numerous algorithms for causal discovery have
+been proposed in the DAG literature, such as constraint-based methods like the
+PC algorithm \citep{Spirtes2001} and Fast Causal Inference (FCI)
+\citep{Spirtes2000}), score-based methods like Hill-Climb Search and Greedy
+Equivalence Search (GES) \citep{Chickering2002}, and continuous
+optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
+\citep{Bello2022}. However, the adaption of these algorithms in applied fields
+remain limited, and researchers typically prefer to construct DAGs manually
+based on their domain expertise \citep{Tennant2020, Petersen2021}. We believe this 
+is the case because of problems with these algorithms when applied in practical
+settings:
 
@@
 
-Due to these challenges, whether constructing the model manually or using an
+Due to these challenges, even when using automated algorithms, we often have to
+rely on manual post hoc testing and modifications to the model learned by the
+algorithm. Whether constructing the model manually or using an
 algorithm, we must rely on domain expertise to construct or modify the
@@
 CI tests, it can be difficult to figure which test to focus on.
-
-% As it is evident that researchers also draw DAGs by hand and even if they use
-% an algorithm, the need to make some post-hoc adjustments to the model, in this
-% paper we present a tool to assist researchers in making these adjustments
-% somewhat inspired by modification indices. In the case of DAGs, global testing
-% methodology does exist such as Shipley test, that can combine local structure
-% tests into a single p-value test. However, this does not give us a way to make
-% modifications to our model. A possible way to do this by checking which of
-% these local tests are violated. However, this approach can quickly become
-% diffult as the size of the network increases. For example, the popular alarm
-% network implies $x$ condition independence statements.
 

### 2024-08-05 paper/main.tex
Updates

@@
 rely on manual post hoc testing and modifications to the model learned by the
-algorithm. Whether constructing the model manually or using an
-algorithm, we must rely on domain expertise to construct or modify the
-model. Hence, it is crucial to have tools to guide this modification process.
-One of the ways to guide this modification is to use Conditional Independence
-(CI) based model testing \cite{Ankan2021}. As DAGs imply CI statements that can
-be tested in the given data using statistical CI tests, we can verify whether
-the implied CIs are consistent in the data or not. If some of implied CIs are
-not valid in the given dataset, we can modify the model accordingly.
-Figure~\ref{fig:ci_testing} shows an example of model testing using CIs.
-However, this CI testing approach has a few drawbacks. Firstly, as CIs are
-implied only by missing edges in the model, we can not test whether an
-incorrect edge is already present in the model. Secondly, the number of implied
-CIs can be quite large. For example, in the moderately sized alarm network with 
-$ 37 $ node, there are $287$ implied CIs. If we rely only on the p-values of the
-CI tests, it can be difficult to figure which test to focus on.
+algorithm. Whether constructing the model manually or using an algorithm, we
+must rely on domain expertise to construct or modify the model. Hence, it is
+crucial to have tools to guide this modification process. One of the ways to
+guide this modification is to use Conditional Independence (CI) based model
+testing \cite{Ankan2023}. As DAGs imply CI statements that can be tested in the
+given data using statistical CI tests, we can verify whether the implied CIs
+are consistent in the data or not. If some of implied CIs are not valid in the
+given dataset, we can modify the model accordingly. Figure~\ref{fig:ci_testing}
+shows an example of model testing using CIs. However, using p-values of the CI
+tests has a few drawbacks. Firstly, as CIs are implied only by missing edges in
+the model, we can not test whether an incorrect edge is already present in the
+model. For examples, in high sample size scenarios, we might get a significant
+association between a pair of variables with very small association. Secondly,
+the number of implied CIs can be quite large. For example, in the moderately
+sized alarm network with $ 37 $ node, there are $287$ implied CIs. If we rely
+only on the p-values of the CI tests, it can be difficult to figure which test
+to focus on.
 
-On the other hand, in Structural Equation Models (SEMs) literature, more
-emphasis has been placed on developing tools to help researchers manually
-construct models while integrating their domain expertise. Researchers
-typically develop an initial model based on domain knowledge and use various
-tools to guide modifications to improve the fit of their model to data. This
-model construction process is known as \emph{Specification Search}. A common
-method for specification search is using \emph{modification indices}, which
-identify potential changes to the model structure that improve the model fit.
-These potential changes are then ranked by the ones that improve the fit the
-most. This ranking allows the researchers to focus first on the suggestions
+To deal with both this issues, in this paper, we develop a measure of partial
+association which can be using in combination of the CI test to rank the
+violations in these CI tests. On the other hand, in Structural Equation Models
+(SEMs) literature, more emphasis has been placed on developing tools to help
+researchers manually construct models while integrating their domain expertise.
+Researchers typically develop an initial model based on domain knowledge and
+use various tools to guide modifications to improve the fit of their model to
+data. This model construction process is known as \emph{Specification Search}.
+A common method for specification search is using \emph{modification indices},
+which identify potential changes to the model structure that improve the model
+fit. These potential changes are then ranked by the ones that improve the fit
+the most. This ranking allows the researchers to focus first on the suggestions
 that can improve the model the most and while integrating their expertise to
@@
 In this section, we develop a novel partial measure of association for mixed
-data by extending the idea of partial correlation to mixed data. We do this by
-combining a mixed-data residualization method with a multivariate measure of
-association. Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y
-$, and $ \bm{Z} $, we are interested in computing an estimate of the
-conditional association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $
-conditioned on $ \bm{Z} $. As the residualization approach requires atleast one
-conditional variable to be present, if $ \bm{Z} = \emptyset $, we use a
-constant value for $ Z $. Similar to partial correlation, we first compute the
-residuals for $ X $ and $ Y $ using $ \bm{Z} $ as the predictor variables. We
-use the mixed data residualization approach from \citet{Ankan2023}. Depending
-on the type of dependent variable, the residuals are defined as follows:
+data by extending the idea of partial correlation to mixed data. As partial
+correlations require two things: 1) A residualization method, and 2) measure of
+associations between the residuals. We generalize this to mixed data by combining a
+mixed data residualization method \citep{Ankan2023} with a multivariate measure
+of association based on Pillai's Trace. Given $ n $ samples $ D = (x, y,
+\bm{z}) $ of variables $ X $, $ Y $, and $ \bm{Z} $, we are interested in
+computing an estimate of the conditional association $ \rho(X, Y; \bm{Z}) $
+between $ X $ and $ Y $ conditioned on $ \bm{Z} $. As the residualization
+approach requires atleast one conditional variable to be present, if $ \bm{Z} =
+\emptyset $, we use a constant value for $ Z $. Similar to partial correlation,
+we first compute the residuals for $ X $ and $ Y $ using $ \bm{Z} $ as the
+predictor variables. We use the mixed data residualization approach from
+\citet{Ankan2023}. Depending on the type of dependent variable, the residuals
+are defined as follows:
 
@@
 Using this mixed data partial association measure, we propose a method to rank
-CI tests, and use it for model construction and modification. If the user
-desires to construct the model manually, they can also start with an empty
-graph, or start with a DAG outputted by a causal discovery algorithm. Given a
-DAG $ G $ and a dataset $ D $, for each pair for variable $ X $ and $ Y $ in $
-G $ that does not have an edge between them, we compute the partial association
-measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $. The measure of association in
-this case can be interpreted as the unexplained observed association between
-them given the current model. We can then use this to rank pair of variables
-which have very high unexplained association between them. Based on the domain
-expertise, we can then choose one of these suggested pair of variables, decide
-the direction of the edge between the variables, and add it to the model. Once
-the new edge is added we can recompute the association of all other variables
-and $ X $ and $ Y $.
+CI tests, and use it for model construction and modification. Given a
+potentially empty DAG $ G $ and a dataset $ D $, for each pair for variable $ X
+$ and $ Y $ in $ G $:
+\begin{itemize}
+	\item If there is no edge between $ X $ and $ Y $, we compute p-value for the CI test $ X \ci Y \rvert pa_G(X) \cup pa_G(Y) $ and the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $.
+	\item If there is an edge between $ X $ and $ Y $, we compute p-value for the CI test $ X \ci Y \rvert pa_\underling{G}(X) \cup pa_{\underline{G}(Y)} $ and the partial association measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $.
+\end{itemize}
+
+The value of this partial measure of association can be interpreted as - given
+the current structure of the model how much of the observed correlation between
+the variables $ X $ and $ Y $ in data is not explained by the model. To modify
+the model based on this, we can look at the pair of variables that have an edge
+between them. If any of these pair have a p-value greater than partial
+association below our chosen threshold, we can remove that edge as it is
+redundant. In the case when there are
+
+We can then use this to rank pair of variables which have very high unexplained
+association between them. Based on the domain expertise, we can then choose one
+of these suggested pair of variables, decide the direction of the edge between
+the variables, and add it to the model. Once the new edge is added we can
+recompute the association of all other variables and $ X $ and $ Y $.
 

### 2024-08-05 paper/main.tex
Updates

@@
 	\item If there is no edge between $ X $ and $ Y $, we compute p-value for the CI test $ X \ci Y \rvert pa_G(X) \cup pa_G(Y) $ and the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $.
-	\item If there is an edge between $ X $ and $ Y $, we compute p-value for the CI test $ X \ci Y \rvert pa_\underling{G}(X) \cup pa_{\underline{G}(Y)} $ and the partial association measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $.
+	\item If there is an edge between $ X $ and $ Y $, we compute p-value for the CI test $ X \ci Y \rvert pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $ and the partial association measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $.
 \end{itemize}

### 2024-08-06 paper/main.tex
updates

@@
 % directly follow a colon or long dash
-\title{Expert-In-The-Loop Causal Discovery: Integrating Expert Knowledge}
+\title{Expert-In-The-Loop Causal Discovery}
 \author{
@@
 \begin{abstract}
-	% TODO: Needs to be shortened.
-	% In Pearl's framework, the initial step for performing a causal analysis
-	% involves determining the causal structure between the variables from
-	% data typically in the form of a Directed Acyclic Graph (DAG) or an
-	% Structural Equation Model (SEM). This procedure, known as \emph{causal
-	% discovery}, has been extensively studied for both DAGs and SEMs. While
-	% numerous automated algorithms exist for estimating DAGs from datasets,
-	% their application in applied fields remain limited, possibly due to a
-	% lack of confidence in these algorithms stemming from them making
-	% obvious mistakes, and the difficulties in choosing the right algorithm
-	% for a given dataset. Consequently, in applied disciplines, these DAGs
-	% are predominantly constructed manually based on domain expertise. Due
-	% to being constructed by hand, it is vital that the fit of these models
-	% are tested against data and the models are potentially modified to
-	% acheive a better fit. On the other hand, in the field of SEMs models
-	% are typically constructed manually through an iterative process of
-	% assessing fit and modifying the model according to that. This process
-	% is commonly known as \emph{Specification Search} where methods like
-	% \emph{modification indices} combined with expert knowledge can be used
-	% to guide this modification process. In the case of DAGs, although there
-	% are tests to assess the global fit of a model by combining the tests
-	% using implied conditional independencies of the model, but there are no
-	% methods that can guide our modification process. In this paper, we
-	% present a simple modification process aimed at helping researchers in
-	% manually constructing or modifying DAGs outputted by an algorithm. This
-	% modification process is based on using a (conditional) measure of
-	% association between variables in the model to assess potential edges
-	% that best explains unexplained correlation between variables which the
-	% researcher specifies the direction of this edge. In the case of
-	% continuous or discrete variables, the effect size measures of local
-	% tests can be used as this measure of association, however for mixed
-	% data there is no such measure. We present a measure of association
-	% based on canonical correlations for mixed data. We also present a
-	% graphical web-tool that can assist researchers in iteratively modifying
-	% and constructing their DAGs. We theoretically show that in the presence
-	% of an oracle, this iterative modification process is consistent in
-	% recovering the true DAG. Emprically, we show that using the mixed data
-	% measure of association, the modification process performs comparably to
-	% PC and Hill-Climb Search algorithms if the user is able to correctly
-	% identify the direction of the edge in one out of three cases.
+	Causal Discovery has received substantial attention in the DAG
+	literature and many automated algorithms have been proposed for
+	constructing DAGs from data. However, the application of these
+	algorithms in applied fields remain limited due to challenges in using
+	them in practice for example they make obvious mistakes, output CPDAGs
+	that need to be oriented using expert knowledge before estimation
+	methods can be applied. As a result of these challenges, researchers
+	typically construct these model by hand. In this paper, inspired by
+	specification search in SEMs where the method suggests changes in the
+	model that improves the fit the most, we present a method based on
+	Conditional Independence testing and a novel measure of partial
+	association. The method iteratively suggests addition and removal of
+	edges between pairs of variables, while the user can choose the
+	direction of the edge. Using empirical analysis we show that this
+	manual construction approach can outperform automated algorithms if the
+	user gets the direction correct in <x out of y> cases. We also show
+	that the user expert can be replaced with an LLM.
 \end{abstract}

### 2024-08-06 paper/main.tex
updates

@@
 developing automated algorithms that can learn the causal structure from
-datasets. In contrast, the SEM literature has focused more towards developing
-tools that assist researchers in manually constructing or modifying models.
+datasets. As a result of these efforts, numerous algorithms for causal
+discovery have been proposed in the DAG literature, such as constraint-based
+methods like the PC algorithm \citep{Spirtes2001} and Fast Causal Inference
+(FCI) \citep{Spirtes2000}), score-based methods like Hill-Climb Search and
+Greedy Equivalence Search (GES) \citep{Chickering2002}, and continuous
+optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
+\citep{Bello2022}. In contrast, the SEM literature has focused more towards
+developing tools that can assist researchers in manually constructing or
+modifying models by allowing them to integrate their expert knowledge into the
+model building process. Researchers typically develop an initial model based on
+domain knowledge and use various tools to guide modifications to improve the
+fit of their model to data. This model construction process is known as
+\emph{Specification Search}. A common method for specification search is using
+\emph{modification indices}, which identify potential changes to the model
+structure that improve the model fit. These potential changes are then ranked
+by the ones that improve the fit the most. This ranking allows the researchers
+to focus first on the suggestions that can improve the model the most and while
+integrating their expertise to choose the most optimal modifications.
 
-As a result of these efforts, numerous algorithms for causal discovery have
-been proposed in the DAG literature, such as constraint-based methods like the
-PC algorithm \citep{Spirtes2001} and Fast Causal Inference (FCI)
-\citep{Spirtes2000}), score-based methods like Hill-Climb Search and Greedy
-Equivalence Search (GES) \citep{Chickering2002}, and continuous
-optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
-\citep{Bello2022}. However, the adaption of these algorithms in applied fields
-remain limited, and researchers typically prefer to construct DAGs manually
-based on their domain expertise \citep{Tennant2020, Petersen2021}. We believe this 
-is the case because of problems with these algorithms when applied in practical
-settings:
+Even though so many automated algorithms have been proposed, the adaption of
+these algorithms in applied fields remain limited, and researchers typically
+prefer to construct DAGs manually based on their domain expertise
+\citep{Tennant2020, Petersen2021}. We believe this is the case because of
+problems with these algorithms when applied in practical settings:
 
@@
 researchers manually construct models while integrating their domain expertise.
-Researchers typically develop an initial model based on domain knowledge and
-use various tools to guide modifications to improve the fit of their model to
-data. This model construction process is known as \emph{Specification Search}.
-A common method for specification search is using \emph{modification indices},
-which identify potential changes to the model structure that improve the model
-fit. These potential changes are then ranked by the ones that improve the fit
-the most. This ranking allows the researchers to focus first on the suggestions
-that can improve the model the most and while integrating their expertise to
-choose the most optimal modifications.
 
@@
 		\end{tikzpicture}
-		\caption{\todo[inline]{Finish figure. Show an example of model testing using CIs}}
+		\caption{\todo[inline]{Finish figure. Show an example of model learnt using PC algorithm $\rightarrow$ Orienting the unoriented edges $\rightarrow$  Model testing}}
 		\label{fig:ci_testing}
@@
 \begin{figure}
-	\caption{\todo[inline]{Figure showing an example of modification indices}}
+	\caption{\todo[inline]{Take a model; show the modification indices for it.}}
 \end{figure}
@@
 
-% Add that we need a p-value as well.
-
 Using this mixed data partial association measure, we propose a method to rank
@@
 
-The value of this partial measure of association can be interpreted as - given
-the current structure of the model how much of the observed correlation between
-the variables $ X $ and $ Y $ in data is not explained by the model. To modify
-the model based on this, we can look at the pair of variables that have an edge
+For a pair of variables $ X $ and $ Y $, if there is an edge present between
+them, the interpretation of this partial association measure is similar to path
+coefficients in SEMs, i.e. how strong is the edge between the variables. Hence,
+we can use this to identify potential edges that can be removed from the model.
+In the case when there is no edge between the variables, the value of this
+partial measure of association can be interpreted as - given the current
+structure of the model how much of the observed correlation between the
+variables $ X $ and $ Y $ in data is not explained by the model. To modify the
+model based on this, we can look at the pair of variables that have an edge
 between them. If any of these pair have a p-value greater than partial
 association below our chosen threshold, we can remove that edge as it is
-redundant. In the case when there are
+redundant.
+
+The measure of association also gives us a way to rank the potential
+modifications. The smallest values (below the threshold) of the association
+measure are the worst offending ones and for when the edges are not present the
+largest values would add the most amount of explainability to the model. Using
+this ranking the user can decide which pair of variables to focus on.
 
@@
 
+\todo[inline]{Add a figure showing an example of how this modification works.}
+
 \begin{theorem}
@@
 \end{theorem}
+
+\section{Connection to score based methods}
+
+\begin{itemize}
+	\item Because they make only local changes, can get stuck in local minimas.
+\end{itemize}
 

### 2024-08-06 paper/main.tex
Updates abstract

@@
 % directly follow a colon or long dash
-\title{Expert-In-The-Loop Causal Discovery}
+\title{Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Using Expert Knowledge in Model Construction}
 \author{
@@
 \begin{abstract}
-	Causal Discovery has received substantial attention in the DAG
-	literature and many automated algorithms have been proposed for
-	constructing DAGs from data. However, the application of these
-	algorithms in applied fields remain limited due to challenges in using
-	them in practice for example they make obvious mistakes, output CPDAGs
-	that need to be oriented using expert knowledge before estimation
-	methods can be applied. As a result of these challenges, researchers
-	typically construct these model by hand. In this paper, inspired by
-	specification search in SEMs where the method suggests changes in the
-	model that improves the fit the most, we present a method based on
-	Conditional Independence testing and a novel measure of partial
-	association. The method iteratively suggests addition and removal of
-	edges between pairs of variables, while the user can choose the
-	direction of the edge. Using empirical analysis we show that this
-	manual construction approach can outperform automated algorithms if the
-	user gets the direction correct in <x out of y> cases. We also show
-	that the user expert can be replaced with an LLM.
+	Causal Discovery has received significant attention in the Directed
+	Acyclic Graphs (DAGs) literature, leading to the development of
+	numerous automated algorithms for constructing DAGs from data. Despite
+	these advancements, the application of these algorithms in applied
+	domains remain limited, with researchers often preferring to construct
+	DAGs manually based on their expert knowledge. We hypothesize that the
+	reluctance to use automated algorithms arise due to issues with these
+	algorithms such as their tendency to make obvious mistakes, and
+	outputting the Markov Equivalence Class (MEC) instead of a DAG, which
+	is often required for further analysis. Even when using these
+	algorithms, researchers have to rely on their expert knowledge to make
+	modifications to the learned model. Given this reliance on manual model
+	construction and modification, in this paper, we introduce a method to
+	assist researchers in this task inspired by specification search in
+	Structural Equation Models(SEMs). We present an iterative approach to
+	model modification utilizing Conditional Independence (CI) testing
+	combined with a novel measure of partial association for mixed data to
+	rank potential additions or removal of edges to the DAG such that the
+	improvement in the model's explanation of the data is improved the
+	most. Specifically, the method ranks violations of implied conditional
+	independences (CI) in the data using the measure of partial
+	association. The user can choose to fix these violations by adding
+	edges between the pair of variables while using their domain expertise
+	to decide the direction of edge between the variables. Through
+	empirical analysis, we demonstrate that taking a greedy approach (i.e.
+	fixing the worst ranked violation first) with this manual construction
+	approach can get comparable performance to automated algorithms if the
+	user are able to get the edge direction correctly between the variables
+	at least $ 1 $ out of $ 3 $ prompts (check if this is correct).
+	Alternatively, we show that a Large Language Model (LLM) can also be
+	used to help in deciding the direction between the variables.
 \end{abstract}

### 2024-08-06 paper/main.tex
Updates paper

--- 2024-08-06 paper/main.tex
@@
 % directly follow a colon or long dash
-\title{Expert-In-The-Loop Causal Discovery}
+\title{Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge in Model Modification}
 \author{
@@
 \begin{abstract}
-	Causal Discovery has received substantial attention in the DAG
-	literature and many automated algorithms have been proposed for
-	constructing DAGs from data. However, the application of these
-	algorithms in applied fields remain limited due to challenges in using
-	them in practice for example they make obvious mistakes, output CPDAGs
-	that need to be oriented using expert knowledge before estimation
-	methods can be applied. As a result of these challenges, researchers
-	typically construct these model by hand. In this paper, inspired by
-	specification search in SEMs where the method suggests changes in the
-	model that improves the fit the most, we present a method based on
-	Conditional Independence testing and a novel measure of partial
-	association. The method iteratively suggests addition and removal of
-	edges between pairs of variables, while the user can choose the
-	direction of the edge. Using empirical analysis we show that this
-	manual construction approach can outperform automated algorithms if the
-	user gets the direction correct in <x out of y> cases. We also show
-	that the user expert can be replaced with an LLM.
+	\emph{Causal Discovery} has received significant attention in the
+	Directed Acyclic Graphs (DAGs) literature, leading to the development
+	of numerous automated algorithms for constructing DAGs from data.
+	Despite these advancements, the application of these algorithms in
+	applied domains remain limited, with researchers often preferring to
+	construct DAGs manually based on their expert knowledge. We hypothesize
+	that the reluctance to use automated algorithms arises due to several
+	practical issues, such as difficulty in choosing the best algorithm and
+	their hyperparameters for a given dataset, algorithms can make obvious
+	mistakes, and they output the Markov Equivalence Class (MEC) instead of
+	a DAG, which is often required for further analysis. Even when using
+	these algorithms, researchers must rely on their expert knowledge to
+	make modifications, or orient edges of the learned model. Given this
+	reliance on manual model construction and modification, in this paper,
+	we present a method inspired by \emph{specification search} in
+	Structural Equation Models (SEMs) to assist researchers in iteratively
+	modifying their models. Specifically, we utilize implied Conditional
+	Independence (CI) based model testing combined with a novel measure of
+	partial association for mixed data to rank violations of implied CIs in
+	data. Using this ranking users can make modifications to their models
+	by either removing edges or adding edges between pair of variables
+	while deciding the direction of the edge based on their domain
+	expertise. The ranking allows users to prioritize the worst offending
+	violations first. Through empirical analysis, when applied to learning
+	the model, we demonstrate that taking a greedy approach (i.e., fixing
+	the worst violation first) this approach can get comparable performance
+	to automated algorithms if the users are able to get the edge direction
+	correctly with an accuracy higher than $ 0.4 $. (check if this is
+	correct). Alternatively, we show that a Large Language Model (LLM) can
+	also be used to help in deciding the direction between the variables.
 \end{abstract}

### 2024-08-06 paper/main.tex
Updates

--- 2024-08-06 paper/main.tex
@@
 % directly follow a colon or long dash
-\title{Expert-In-The-Loop Causal Discovery}
+\title{Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge for Model Modification}
 \author{
@@
 \begin{abstract}
-	Causal Discovery has received substantial attention in the DAG
-	literature and many automated algorithms have been proposed for
-	constructing DAGs from data. However, the application of these
-	algorithms in applied fields remain limited due to challenges in using
-	them in practice for example they make obvious mistakes, output CPDAGs
-	that need to be oriented using expert knowledge before estimation
-	methods can be applied. As a result of these challenges, researchers
-	typically construct these model by hand. In this paper, inspired by
-	specification search in SEMs where the method suggests changes in the
-	model that improves the fit the most, we present a method based on
-	Conditional Independence testing and a novel measure of partial
-	association. The method iteratively suggests addition and removal of
-	edges between pairs of variables, while the user can choose the
-	direction of the edge. Using empirical analysis we show that this
-	manual construction approach can outperform automated algorithms if the
-	user gets the direction correct in <x out of y> cases. We also show
-	that the user expert can be replaced with an LLM.
+	\emph{Causal Discovery} has received significant attention in the
+	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
+	development of numerous automated algorithms for constructing DAGs from
+	data. Despite these advancements, the application of these algorithms
+	in applied domains remain limited, with researchers often preferring to
+	construct DAGs manually based on their expert knowledge. We hypothesize
+	that the reluctance to use automated algorithms arises due to several
+	practical issues, such as algorithms often make obvious mistakes making
+	it difficult to trust, they output the Markov Equivalence Class (MEC)
+	instead of a DAG (which is often required for further analysis), and
+	difficulties in choosing the appropriate algorithm and hyperparameters
+	for a given dataset. Even when using these algorithms, researchers must
+	rely on their expert knowledge to make modifications and orient edges
+	of the learned model. Given this reliance on manual model modification,
+	in this paper, we present a method inspired by \emph{specification
+	search} in \emph{Structural Equation Models} (SEMs) to assist
+	researchers in iteratively modifying their models to improve their fit
+	to data. Specifically, we utilize implied Conditional Independence (CI)
+	based model testing combined with a novel measure of partial
+	association for mixed data to rank violations of implied CIs of the
+	model in data. Using this ranking users can choose to make
+	modifications to their models by either removing edges or adding edges
+	between pair of variables while using their expert knowledge to decide
+	the orientation of the edges. The ranking allows users to prioritize
+	the worst offending violations first. Through empirical analysis, when
+	applied to learning the model from scratch, we demonstrate that taking
+	a greedy approach (i.e., fixing the worst violation first) this
+	approach can get comparable performance to automated algorithms if the
+	users are able to decide the edge direction correctly with an accuracy
+	higher than $ 0.4 $. (TODO: Double check this value). We also show that
+	a Large Language Model (LLM) can also be used for the purpose of
+	deciding the edge orientation in the absence of an expert.
 \end{abstract}
@@
 		orient these edges before using them for analysis.
+	\item \textbf{Difficulty in Choosing Best Algorithm and Hyperparameters:}
 \end{enumerate}

### 2024-08-07 paper/main.tex
Updates

@@
 	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
-	development of numerous automated algorithms for constructing DAGs from
+	development of numerous automated algorithms for learning DAGs from
 	data. Despite these advancements, the application of these algorithms
@@
 	construct DAGs manually based on their expert knowledge. We hypothesize
-	that the reluctance to use automated algorithms arises due to several
+	that the reluctance to use these algorithms arises due to several
 	practical issues, such as algorithms often make obvious mistakes making

### 2024-08-07 paper/main.tex
udpates

@@
 % directly follow a colon or long dash
-\title{Expert-In-The-Loop Causal Discovery: An Iterative Approach for Efficiently Utilizing Expert Knowledge for Model Modification}
+\title{Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge}
 \author{
@@
 	difficulties in choosing the appropriate algorithm and hyperparameters
-	for a given dataset. Even when using these algorithms, researchers must
-	rely on their expert knowledge to make modifications and orient edges
-	of the learned model. Given this reliance on manual model modification,
-	in this paper, we present a method inspired by \emph{specification
-	search} in \emph{Structural Equation Models} (SEMs) to assist
-	researchers in iteratively modifying their models to improve their fit
-	to data. Specifically, we utilize implied Conditional Independence (CI)
-	based model testing combined with a novel measure of partial
-	association for mixed data to rank violations of implied CIs of the
-	model in data. Using this ranking users can choose to make
-	modifications to their models by either removing edges or adding edges
-	between pair of variables while using their expert knowledge to decide
-	the orientation of the edges. The ranking allows users to prioritize
-	the worst offending violations first. Through empirical analysis, when
-	applied to learning the model from scratch, we demonstrate that taking
-	a greedy approach (i.e., fixing the worst violation first) this
-	approach can get comparable performance to automated algorithms if the
-	users are able to decide the edge direction correctly with an accuracy
-	higher than $ 0.4 $. (TODO: Double check this value). We also show that
-	a Large Language Model (LLM) can also be used for the purpose of
-	deciding the edge orientation in the absence of an expert.
+	for a given dataset. Even if using these algorithms, researchers have
+	to rely on their expert knowledge to make modifications to the learned
+	model in terms of fixing any mistakes or orienting the edges before
+	these models can be used for any further analysis. Given this reliance
+	on manual model modification, in this paper, we present a method
+	inspired by \emph{specification search} in \emph{Structural Equation
+	Models} (SEMs) to assist researchers in iteratively modifying or
+	constructing their models to improve their fit to data. Specifically,
+	we utilize implied Conditional Independence (CI) based model testing
+	combined with a novel measure of partial association for mixed data to
+	rank violations of implied CIs of the model in data. Based on this
+	ranking, users can then choose to make modifications to the model by
+	either removing or adding edges while using their expertise to decide
+	the orientation of the edge. The ranking allows users to prioritize the
+	worst offending violations first. Through empirical analysis, when
+	applied to constructing the model from scratch, we demonstrate that
+	taking a greedy approach (i.e., fixing the worst violation first) gets
+	us comparable performance to automated algorithms if the user is able
+	to correctly provide the edge direction in $ 1 $ out of $ 3 $ prompts.
+	We further show that in the absence of an expert, a Large Language
+	Model (LLM) can also be used for getting the edge orientations.
 \end{abstract}

### 2024-08-07 paper/main.tex
updates

@@
 \begin{abstract}
-	\emph{Causal Discovery} has received significant attention in the
-	\emph{Directed Acyclic Graphs} (DAGs) literature, leading to the
-	development of numerous automated algorithms for learning DAGs from
-	data. Despite these advancements, the application of these algorithms
-	in applied domains remain limited, with researchers often preferring to
-	construct DAGs manually based on their expert knowledge. We hypothesize
-	that the reluctance to use these algorithms arises due to several
-	practical issues, such as algorithms often make obvious mistakes making
-	it difficult to trust, they output the Markov Equivalence Class (MEC)
-	instead of a DAG (which is often required for further analysis), and
-	difficulties in choosing the appropriate algorithm and hyperparameters
-	for a given dataset. Even if using these algorithms, researchers have
-	to rely on their expert knowledge to make modifications to the learned
-	model in terms of fixing any mistakes or orienting the edges before
-	these models can be used for any further analysis. Given this reliance
-	on manual model modification, in this paper, we present a method
-	inspired by \emph{specification search} in \emph{Structural Equation
-	Models} (SEMs) to assist researchers in iteratively modifying or
-	constructing their models to improve their fit to data. Specifically,
-	we utilize implied Conditional Independence (CI) based model testing
-	combined with a novel measure of partial association for mixed data to
-	rank violations of implied CIs of the model in data. Based on this
-	ranking, users can then choose to make modifications to the model by
-	either removing or adding edges while using their expertise to decide
-	the orientation of the edge. The ranking allows users to prioritize the
-	worst offending violations first. Through empirical analysis, when
-	applied to constructing the model from scratch, we demonstrate that
-	taking a greedy approach (i.e., fixing the worst violation first) gets
-	us comparable performance to automated algorithms if the user is able
-	to correctly provide the edge direction in $ 1 $ out of $ 3 $ prompts.
-	We further show that in the absence of an expert, a Large Language
-	Model (LLM) can also be used for getting the edge orientations.
+Causal discovery has received significant attention in the Directed Acyclic
+Graphs (DAGs) literature, leading to the development of numerous automated
+algorithms for learning DAGs from data. Despite these advancements, the
+application of these algorithms in applied domains remains limited, with
+researchers often preferring to manually construct DAGs based on their expert
+knowledge. Even when using these algorithms, researchers must rely on their
+expertise to correct any mistakes or orient edges before the models can be used
+for further analysis. Given this reliance on manual model construction and
+modification, we propose a method inspired by specification search in
+Structural Equation Models (SEMs) to assist researchers in iteratively
+modifying DAGs. Our method employs implied Conditional Independence (CI) based
+model testing, combined with a novel measure of partial association for mixed
+data, to rank violations of implied CIs of the model in the data. This ranking
+helps the user in choosing pairs of variables between which to add or remove
+edges. Based of their expertise, the user can decide to remove or add edge
+while deciding the correct orientation between the pair of variables.
+Empirically, we demonstrate that a greedy approach (i.e., fixing the worst
+violation first) achieves performance comparable to automated algorithms,
+provided the user can correctly determine the edge direction in $ 1 $ out of $
+3 $ edge additions. Additionally, we show that in the absence of an expert, a
+Large Language Model (LLM) can potentially be used to determine edge
+orientations. We provide the implementation of this approach as both a website
+and in the Python package <redacted>.
 \end{abstract}

### 2024-08-07 paper/main.tex
Updates

@@
 algorithms for learning DAGs from data. Despite these advancements, the
-application of these algorithms in applied domains remains limited, with
+application of these algorithms in practical domains remains limited, with
 researchers often preferring to manually construct DAGs based on their expert
 knowledge. Even when using these algorithms, researchers must rely on their
-expertise to correct any mistakes or orient edges before the models can be used
-for further analysis. Given this reliance on manual model construction and
+expertise to correct mistakes or orient edges before the models can be used for
+further analysis. Given this reliance on manual model construction and
 modification, we propose a method inspired by specification search in
@@
 data, to rank violations of implied CIs of the model in the data. This ranking
-helps the user in choosing pairs of variables between which to add or remove
-edges. Based of their expertise, the user can decide to remove or add edge
-while deciding the correct orientation between the pair of variables.
-Empirically, we demonstrate that a greedy approach (i.e., fixing the worst
-violation first) achieves performance comparable to automated algorithms,
-provided the user can correctly determine the edge direction in $ 1 $ out of $
-3 $ edge additions. Additionally, we show that in the absence of an expert, a
-Large Language Model (LLM) can potentially be used to determine edge
-orientations. We provide the implementation of this approach as both a website
-and in the Python package <redacted>.
+helps users to prioritize pair of variables between which to add or remove
+edges. Users can then decide to remove or add edges while determining the
+correct orientation of the edge based on their expertise. Empirically, we
+demonstrate that a greedy approach (i.e., fixing the worst violation first)
+achieves performance comparable to automated algorithms, provided the user can
+correctly determine the edge direction in $ 1 $ out of $ 3 $ edge additions.
+Additionally, we show that in the absence of an expert, a Large Language Model
+(LLM) can potentially be used to determine edge orientations. 
 \end{abstract}

### 2024-08-07 paper/main.tex
Updates

@@
 
-Understanding the cause-and-effect relationships between variables is a
-fundamental goal in many scientific questions as it reveals the underlying
-mechanisms of observed phenomena and informs effective interventions or policy
-decisions. \emph{Causal Discovery} methods aim to uncover these underlying
-cause-effect relationships among random variables from observational data. The
-problem of causal discovery has been explored in both the Directed Acyclic
-Graphs (DAGs) and Structural Equation Models (SEMs) literatures, each adopting
-a different approach. In the DAG literature, the focus has primarily been on
-developing automated algorithms that can learn the causal structure from
-datasets. As a result of these efforts, numerous algorithms for causal
-discovery have been proposed in the DAG literature, such as constraint-based
-methods like the PC algorithm \citep{Spirtes2001} and Fast Causal Inference
-(FCI) \citep{Spirtes2000}), score-based methods like Hill-Climb Search and
-Greedy Equivalence Search (GES) \citep{Chickering2002}, and continuous
-optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
-\citep{Bello2022}. In contrast, the SEM literature has focused more towards
-developing tools that can assist researchers in manually constructing or
-modifying models by allowing them to integrate their expert knowledge into the
-model building process. Researchers typically develop an initial model based on
-domain knowledge and use various tools to guide modifications to improve the
-fit of their model to data. This model construction process is known as
-\emph{Specification Search}. A common method for specification search is using
-\emph{modification indices}, which identify potential changes to the model
-structure that improve the model fit. These potential changes are then ranked
-by the ones that improve the fit the most. This ranking allows the researchers
-to focus first on the suggestions that can improve the model the most and while
-integrating their expertise to choose the most optimal modifications.
+Understanding cause-and-effect relationships between variables is a fundamental
+goal in many scientific inquiries, as it reveals the underlying mechanisms of
+observed phenomena and informs effective interventions or policy decisions.
+\emph{Causal Discovery} methods aim to uncover these relationships among random
+variables from observational data. This problem has been extensively explored
+in both the Directed Acyclic Graphs (DAGs) and Structural Equation Models
+(SEMs) literatures, each adopting a different approach.
 
-Even though so many automated algorithms have been proposed, the adaption of
-these algorithms in applied fields remain limited, and researchers typically
-prefer to construct DAGs manually based on their domain expertise
-\citep{Tennant2020, Petersen2021}. We believe this is the case because of
-problems with these algorithms when applied in practical settings:
+In the DAG literature, the focus has primarily been on developing automated
+algorithms to learn causal structures from datasets. As a result of these
+efforts, numerous algorithms for causal discovery have been proposed, such as
+constraint-based methods like the PC algorithm \citep{Spirtes2001} and Fast
+Causal Inference (FCI) \citep{Spirtes2000}), score-based methods like
+Hill-Climb Search and Greedy Equivalence Search (GES) \citep{Chickering2002},
+and continuous optimization-based methods like NO TEARS \citep{Zheng2018} and
+DAGMA \citep{Bello2022}.
+
+In contrast, the SEM literature has focused more on developing tools to assist
+researchers in manually constructing or modifying models while allowing them to
+integrate their expert knowledge into the model-building process. Researchers
+typically develop an initial model based on their expertise and then use
+various tools to guide modifications to improve the model's fit to data. This
+model modification process is commonly known as \emph{Specification Search}
+\citep{Long1983}, employing method like modification indices, and Wald-based
+tests \citep{Marcoulides2018}.
+
+Despite the many automated algorithms proposed for constructing DAGs, their
+adoption in applied research remains limited, with researchers often preferring
+to manually construct DAGs based on their domain expertise \citep{Tennant2020,
+Petersen2021}. We believe this preference persists due to several issues with
+these algorithms when applied in practical settings:
 
 \begin{enumerate}
-	\item \textbf{Make Obvious Mistakes:} Although most algorithms are
+	\item \textbf{Obvious Mistakes:} Although most algorithms are
 		theoretically consistent asymptotically, their finite sample
-		properties are not well understood. Moreover, depending on the
-		selected hyperparameters, their output can vary significantly.
-		These issues combined with no good method to test the
-		performance of these algorithms on a given dataset, in
-		real-world settings with finite samples, it can be difficult to
-		trust the output of these algorithms. Consequently, researchers
-		have to rely on posthoc modifications to the model based on
-		their domain expertise.
+		properties are not well understood. Moreover, their output can
+		vary significantly depending on the selected algorithm and
+		their hyperparameters. Combined with the lack of robust methods
+		to test these algorithms' performance on a given dataset, it
+		can be difficult to trust their outputs. Consequently,
+		researchers have to rely on post-hoc modifications to the model
+		based on their domain expertise.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
@@
 		orient these edges before using them for analysis.
-	\item \textbf{Difficulty in Choosing Best Algorithm and Hyperparameters:}
 \end{enumerate}
@@
 \end{figure}
-\begin{figure}
-	\caption{\todo[inline]{Take a model; show the modification indices for it.}}
-\end{figure}
 
-Inspired by this ranking approach in modification indices, in this paper we
-develop a method to rank the CI tests using a conditional measure of
-association. Our proposed approach uses the measure of association to find
-correlated variables in the data that are not explained by the current model.
-Using the measure, we can then rank the pair of variables between which adding
-an edge would improve the explanation the most. Based on this information, the
-user can select a potential pair of variables and based on their expert
-knowledge decide the direction of the edge. 
+A common method for specification search is using \emph{modification indices},
+which identify potential changes to the model structure that improve the model
+fit. These potential changes are then ranked by the ones that improve the fit
+the most. This ranking allows the researchers to focus first on the suggestions
+that can improve the model the most and while integrating their expertise to
+choose the most optimal modifications. Inspired by this ranking approach in
+modification indices, in this paper we develop a method to rank the CI tests
+using a conditional measure of association. Our proposed approach uses the
+measure of association to find correlated variables in the data that are not
+explained by the current model. Using the measure, we can then rank the pair of
+variables between which adding an edge would improve the explanation the most.
+Based on this information, the user can select a potential pair of variables
+and based on their expert knowledge decide the direction of the edge. 
 \todo[inline]{Add about the score based methods somewhere here}

### 2024-08-07 code/intro_example.R
Adult income example in introduction

@@ -1,0 +1,27 @@
+library(tidyverse)
+library(bnlearn)
+
+d <- read.csv('adult_unproc.csv')
+
+# Preprocess the data
+d$Fnlwgt <- NULL
+d$EducationNum <- NULL
+d$CapitalGain <- NULL
+d$CaptialLoss <- NULL
+
+d$Workclass <- factor(d$Workclass, ordered=F)
+d$MaritalStatus <- factor(d$MaritalStatus, ordered=F)
+d$Occupation <- factor(d$Occupation, ordered=F)
+d$Relationship <- factor(d$Relationship, ordered=F)
+d$Race <- factor(d$Race, ordered=F)
+d$Sex <- factor(d$Sex, ordered=F)
+d$NativeCountry <- factor(d$NativeCountry, ordered=F)
+d$Income <- factor(d$Income, ordered=F)
+
+d$Age <- as.double(d$Age)
+d$HoursPerWeek <- as.double(d$HoursPerWeek)
+
+education_levels = c( "Preschool", "1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th","12th", "HS-grad", "Some-college", "Assoc-voc", "Assoc-acdm", "Bachelors", "Masters", "Prof-school", "Doctorate" )
+d$Education <- factor(d$Education, levels=education_levels, ordered=T)
+
+dag <- pc.stable(d)

### 2024-08-07 code/intro_example.R
updates

@@ -1,5 +1,347 @@
-library(tidyverse)
-library(bnlearn)
+
+sqrtinv <- function(mA) {
+  ei <- eigen(mA)
+  d <- ei$values
+  d <- (d + abs(d)) / 2
+  d2 <- 1 / sqrt(d)
+  d2[d == 0] <- 0
+  return(ei$vectors %*% diag(d2) %*% t(ei$vectors))
+}
+
+pred2res <- function( pred ){
+	setNames( c( 0, cumsum( pred )[-length(pred)]) - 
+		c( rev( cumsum( rev( pred ) ) ), 0 )[-1], names(pred) )
+}
+
+p2r <- function( x, preddf ){
+	present.levels <- colnames( preddf )
+	missing.levels <-  setdiff( levels( x ), present.levels )
+	if( length( missing.levels ) > 0 ){
+		for( xx in missing.levels ){
+			preddf <- cbind( preddf, 0 )
+		}
+		colnames( preddf ) <- c( present.levels, missing.levels )
+		preddf <- preddf[ , levels( x ) ]
+	}
+	as.matrix( apply( preddf, 1, pred2res ), rownames=colnames(preddf) )[
+			cbind( x, seq_along( x ) )
+		]
+}
+
+residual.test <- function(x, y, z, predictor, test) {
+  # Random Forest test v1. Return the p-value for the independence
+  # test X _|_ Y | Z
+  #
+  # @param x: vector
+  # @param y: vector
+  # @param z: vector or data.frame
+  # @param predictor: 'glm' or 'rf'
+  # @param test: 'q3' or 'pillai'
+
+  n <- length(x)
+
+  # Step 1: Determine the variable type of x
+  if (all(class(x) == "factor")){
+    x <- droplevels(x)
+    k <- length(levels(x))
+
+    stopifnot(k > 1)
+    tx <- table(seq_len(n), x)[, -k, drop = FALSE]
+
+    x_type <- "cat"
+  }
+  else if(all(class(x) == c("ordered", "factor"))){
+    x <- droplevels(x)
+    k <- 1 
+    tx <- x
+
+    x_type <- "ord"
+  }
+  else{
+    k <- 1
+    tx <- x
+
+    x_type <- "cont"
+  }
+
+  # Step 2: Determine the variable type of y
+  if (all(class(y) == "factor")){
+    y <- droplevels(y)
+    r <- length(levels(y))
+
[…]

### 2024-08-08 paper/main.tex
Updates

@@
 achieves performance comparable to automated algorithms, provided the user can
-correctly determine the edge direction in $ 1 $ out of $ 3 $ edge additions.
+correctly determine the edge orientation in $ 1 $ out of $ 3 $ edge additions.
 Additionally, we show that in the absence of an expert, a Large Language Model

### 2024-08-08 code/intro_example.R
Updates

@@ -296,20 +296,19 @@ ci.test <- function(x, y, z, d, test){
 }
 
 sl.test <- function(x, y, z, suffStat){
-	# if (length(z) == 0){
-	# 	x_type = suffStat$var_types[[names(suffStat$dm)[x]]]
-	# 	y_type = suffStat$var_types[[names(suffStat$dm)[y]]]
-	# 	browser()
-	# 	if ((x_type == 'cont') & (y_type == 'cont')){
-	# 		return(bnlearn::ci.test(names(suffStat$dm)[x], names(suffStat$dm)[y], data=suffStat$dm, test='mc-mi-g')$p.value)
-	# 	}
-	# 	else if ((x_type %in% c('ord', 'cat')) & (y_type %in% c('ord', 'cat'))){
-	# 		return(bnlearn::ci.test(names(suffStat$dm)[x], names(suffStat$dm)[y], data=suffStat$dm, test='mc-mi')$p.value)
-	# 	}
-	# 	else{
-	# 		return(bnlearn::ci.test(names(suffStat$dm)[x], names(suffStat$dm)[y], data=suffStat$dm, test='mi-cg')$p.value)
-	# 	}
-	# }
+	if (length(z) == 0){
+		x_type = suffStat$var_types[[names(suffStat$dm)[x]]]
+		y_type = suffStat$var_types[[names(suffStat$dm)[y]]]
+		if ((x_type == 'cont') & (y_type == 'cont')){
+			return(bnlearn::ci.test(names(suffStat$dm)[x], names(suffStat$dm)[y], data=suffStat$dm, test='mc-mi-g')$p.value)
+		}
+		else if ((x_type %in% c('ord', 'cat')) & (y_type %in% c('ord', 'cat'))){
+			return(bnlearn::ci.test(names(suffStat$dm)[x], names(suffStat$dm)[y], data=suffStat$dm, test='mc-mi')$p.value)
+		}
+		else{
+			return(bnlearn::ci.test(names(suffStat$dm)[x], names(suffStat$dm)[y], data=suffStat$dm, test='mi-cg')$p.value)
+		}
+	}
 
 	if (suffStat$test=='rf_pillai'){
 		return(residual.test(suffStat$dm[, x], suffStat$dm[, y], suffStat$dm[z], 'rf', 'pillai'))
@@ -365,11 +364,13 @@ d$HoursPerWeek <- as.double(d$HoursPerWeek)
 
[…]

### 2024-08-08 paper/main.tex
Updates

@@
 
-\todo[inline]{Add an example here showing these problems}
-
 Due to these challenges, even when using automated algorithms, we often have to
@@
 and based on their expert knowledge decide the direction of the edge.
-\todo[inline]{Add about the score based methods somewhere here}
 
@@
 		or deletion of edges from the model.
-	\item Lastly, we present a web-browser based tool to allow researchers
-		to apply this method to their own datasets.
+	\item Lastly, we present a both a web-browser based tool and an
+		implementation in the Python package to allow researchers to
+		apply this method to their own datasets.
 \end{enumerate}

### 2024-08-08 paper/main.tex
Updates

@@
 various tools to guide modifications to improve the model's fit to data. This
-model modification process is commonly known as \emph{Specification Search}
-\citep{Long1983}, employing method like modification indices, and Wald-based
+model construction process is commonly known as \emph{Specification Search}
+\citep{Long1983} and uses method like modification indices, and Wald-based
 tests \citep{Marcoulides2018}.
@@
 \begin{enumerate}
-	\item \textbf{Obvious Mistakes:} Although most algorithms are
-		theoretically consistent asymptotically, their finite sample
-		properties are not well understood. Moreover, their output can
-		vary significantly depending on the selected algorithm and
-		their hyperparameters. Combined with the lack of robust methods
-		to test these algorithms' performance on a given dataset, it
-		can be difficult to trust their outputs. Consequently,
-		researchers have to rely on post-hoc modifications to the model
+	\item \textbf{Lack of Trust:} Although most algorithms are
+		asymptotically consistent, their finite sample properties are
+		not well understood. Moreover, their output can vary
+		significantly depending on the algorithm and their
+		hyperparameters. Combined with the lack of robust methods to
+		test these algorithms' performance on a given dataset, it can
+		be difficult to trust their outputs. Consequently, researchers
+		usually have to rely on post-hoc modifications to the model
 		based on their domain expertise.
@@
 
-Due to these challenges, even when using automated algorithms, we often have to
+Figure shows some of these issues when applied to the Adult Income dataset. Due
+to these challenges, even when using automated algorithms, we often have to
 rely on manual post hoc testing and modifications to the model learned by the

### 2024-08-08 paper/main.tex
Updates

@@
 
-Figure shows some of these issues when applied to the Adult Income dataset. Due
-to these challenges, even when using automated algorithms, we often have to
-rely on manual post hoc testing and modifications to the model learned by the
-algorithm. Whether constructing the model manually or using an algorithm, we
-must rely on domain expertise to construct or modify the model. Hence, it is
-crucial to have tools to guide this modification process. One of the ways to
-guide this modification is to use Conditional Independence (CI) based model
-testing \cite{Ankan2023}. As DAGs imply CI statements that can be tested in the
-given data using statistical CI tests, we can verify whether the implied CIs
-are consistent in the data or not. If some of implied CIs are not valid in the
-given dataset, we can modify the model accordingly. Figure~\ref{fig:ci_testing}
-shows an example of model testing using CIs. However, using p-values of the CI
-tests has a few drawbacks. Firstly, as CIs are implied only by missing edges in
-the model, we can not test whether an incorrect edge is already present in the
-model. For examples, in high sample size scenarios, we might get a significant
-association between a pair of variables with very small association. Secondly,
-the number of implied CIs can be quite large. For example, in the moderately
-sized alarm network with $ 37 $ node, there are $287$ implied CIs. If we rely
-only on the p-values of the CI tests, it can be difficult to figure which test
-to focus on.
+Figure~\ref{fig:intro} shows some of these issues when applied to the Adult
+Income dataset. Due to these challenges, whether constructing the model
+manually or using an algorithm, we typically have to rely on some expert
+knowledge to get a model that we can use for further analysis. Hence, it is
+crucial to have tools to guide this construction or modification process.
 
-To deal with both this issues, in this paper, we develop a measure of partial
-association which can be using in combination of the CI test to rank the
-violations in these CI tests. On the other hand, in Structural Equation Models
-(SEMs) literature, more emphasis has been placed on developing tools to help
-researchers manually construct models while integrating their domain expertise.
+A common method for specification search is using \emph{modification indices},
+which identify potential changes to the model structure that improve the model
+fit. These potential changes are then ranked by the ones that improve the fit
+the most. This ranking allows the researchers to focus first on the suggestions
+that can improve the model the most and while integrating their expertise to
+choose the most optimal modifications.
+
+One of the ways to guide this modification is to use Conditional Independence
+(CI) based model testing \cite{Ankan2023}. As DAGs imply CI statements that can
+be tested in the given data using statistical CI tests, we can verify whether
+the implied CIs are consistent with the data or not. If some of implied CIs are
+not valid in the given dataset, we can modify the model accordingly.
+Figure~\ref{fig:ci_testing} shows an example of model testing using CIs.
+However, using p-values of the CI tests has a few drawbacks. Firstly, as CIs
+are implied only by missing edges in the model, we can not test whether an
+incorrect edge is already present in the model. For examples, in high sample
+size scenarios, we might get a significant association between a pair of
+variables with very small association. Secondly, the number of implied CIs can
+be quite large. For example, in the moderately sized alarm network with $ 37 $
+node, there are $287$ implied CIs. If we rely only on the p-values of the CI
+tests, it can be difficult to figure which test to focus on.
+
+Inspired by this ranking approach in modification indices, in this paper we
+develop a method to rank the CI tests using a conditional measure of
+association.  To deal with both this issues, in this paper, we develop a
+measure of partial association which can be using in combination of the CI test
+to rank the violations in these CI tests. On the other hand, in Structural
+Equation Models (SEMs) literature, more emphasis has been placed on developing
+tools to help researchers manually construct models while integrating their
+domain expertise. Our proposed approach uses the measure of association to find
+correlated variables in the data that are not explained by the current model.
+Using the measure, we can then rank the pair of variables between which adding
+an edge would improve the explanation the most. Based on this information, the
+user can select a potential pair of variables and based on their expert
+knowledge decide the direction of the edge.
 
@@
 		\caption{\todo[inline]{Finish figure. Show an example of model learnt using PC algorithm $\rightarrow$ Orienting the unoriented edges $\rightarrow$  Model testing}}
-		\label{fig:ci_testing}
+		\label{fig:intro}
 \end{figure}
-
-A common method for specification search is using \emph{modification indices},
-which identify potential changes to the model structure that improve the model
-fit. These potential changes are then ranked by the ones that improve the fit
-the most. This ranking allows the researchers to focus first on the suggestions
-that can improve the model the most and while integrating their expertise to
-choose the most optimal modifications. Inspired by this ranking approach in
-modification indices, in this paper we develop a method to rank the CI tests
-using a conditional measure of association. Our proposed approach uses the
-measure of association to find correlated variables in the data that are not
-explained by the current model. Using the measure, we can then rank the pair of
-variables between which adding an edge would improve the explanation the most.
-Based on this information, the user can select a potential pair of variables
-and based on their expert knowledge decide the direction of the edge.
 
@@
 		or deletion of edges from the model.
-	\item Lastly, we present a both a web-browser based tool and an
+	\item Lastly, we present both a web-browser based tool and an
 		implementation in the Python package to allow researchers to
@@
 In this section, we develop a novel partial measure of association for mixed
-data by extending the idea of partial correlation to mixed data. As partial
-correlations require two things: 1) A residualization method, and 2) measure of
-associations between the residuals. We generalize this to mixed data by combining a
-mixed data residualization method \citep{Ankan2023} with a multivariate measure
-of association based on Pillai's Trace. Given $ n $ samples $ D = (x, y,
-\bm{z}) $ of variables $ X $, $ Y $, and $ \bm{Z} $, we are interested in
-computing an estimate of the conditional association $ \rho(X, Y; \bm{Z}) $
-between $ X $ and $ Y $ conditioned on $ \bm{Z} $. As the residualization
-approach requires atleast one conditional variable to be present, if $ \bm{Z} =
-\emptyset $, we use a constant value for $ Z $. Similar to partial correlation,
-we first compute the residuals for $ X $ and $ Y $ using $ \bm{Z} $ as the
-predictor variables. We use the mixed data residualization approach from
+data by extending partial correlations to mixed data. As partial correlations
+require two things: 1) A residualization method, and 2) measure of associations
+between the residuals. We generalize this to mixed data by combining a mixed
+data residualization method \citep{Ankan2023} with a multivariate measure of
+association based on Pillai's Trace. Given $ n $ samples $ D = (x, y, \bm{z}) $
+of variables $ X $, $ Y $, and $ \bm{Z} $, we are interested in computing an
+estimate of the conditional association $ \rho(X, Y; \bm{Z}) $ between $ X $
+and $ Y $ conditioned on $ \bm{Z} $. As the residualization approach requires
+atleast one conditional variable to be present, if $ \bm{Z} = \emptyset $, we
+use a constant value for $ Z $. Similar to partial correlation, we first
+compute the residuals for $ X $ and $ Y $ using $ \bm{Z} $ as the predictor
+variables. We use the mixed data residualization approach from
 \citet{Ankan2023}. Depending on the type of dependent variable, the residuals
@@
 
-\section{Connection to score based methods}
+\section{Connection to Score Based Methods}
+\begin{itemize}
+	\item The process of iteratively adding and removing edges in very similar to
+		Greedy Equivalence Search (GES).
+	\item But Because they make only local changes, can get stuck in local minimas.
+	\item This approach looks for modifications globally.
+\end{itemize}
 
-\begin{itemize}
-	\item Because they make only local changes, can get stuck in local minimas.
-\end{itemize}
+\todo[inline]{Show an example where local changes can get stuck}
 

### 2024-08-08 paper/main.tex
Updates

@@
 
-In this section, we develop a novel partial measure of association for mixed
-data by extending partial correlations to mixed data. As partial correlations
-require two things: 1) A residualization method, and 2) measure of associations
-between the residuals. We generalize this to mixed data by combining a mixed
-data residualization method \citep{Ankan2023} with a multivariate measure of
-association based on Pillai's Trace. Given $ n $ samples $ D = (x, y, \bm{z}) $
-of variables $ X $, $ Y $, and $ \bm{Z} $, we are interested in computing an
-estimate of the conditional association $ \rho(X, Y; \bm{Z}) $ between $ X $
-and $ Y $ conditioned on $ \bm{Z} $. As the residualization approach requires
-atleast one conditional variable to be present, if $ \bm{Z} = \emptyset $, we
-use a constant value for $ Z $. Similar to partial correlation, we first
-compute the residuals for $ X $ and $ Y $ using $ \bm{Z} $ as the predictor
-variables. We use the mixed data residualization approach from
-\citet{Ankan2023}. Depending on the type of dependent variable, the residuals
-are defined as follows:
+In this section, we develop a novel measure of partial association for mixed
+data by extending partial correlations \citep{Whittaker2009} \todo{Verify this
+citation} to mixed data.
+As partial correlations require two things: 1) A
+residualization method, and 2) measure of associations between the residuals.
+We generalize this to mixed data by combining a mixed data residualization
+method \citep{Ankan2023} with a multivariate measure of association based on
+Pillai's Trace. Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $,
+$ Y $, and $ \bm{Z} $, we are interested in computing an estimate of the
+conditional association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $
+conditioned on $ \bm{Z} $. As the residualization approach requires atleast one
+conditional variable to be present, if $ \bm{Z} = \emptyset $, we use a
+constant value for $ Z $. Similar to partial correlation, we first compute the
+residuals for $ X $ and $ Y $ using $ \bm{Z} $ as the predictor variables. We
+use the mixed data residualization approach from \citet{Ankan2023}. Depending
+on the type of dependent variable, the residuals are defined as follows:
 

### 2024-08-08 paper/main.tex
Updates

@@
 data by extending partial correlations \citep{Whittaker2009} \todo{Verify this
-citation} to mixed data.
-As partial correlations require two things: 1) A
-residualization method, and 2) measure of associations between the residuals.
-We generalize this to mixed data by combining a mixed data residualization
-method \citep{Ankan2023} with a multivariate measure of association based on
-Pillai's Trace. Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $,
-$ Y $, and $ \bm{Z} $, we are interested in computing an estimate of the
-conditional association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $
-conditioned on $ \bm{Z} $. As the residualization approach requires atleast one
-conditional variable to be present, if $ \bm{Z} = \emptyset $, we use a
-constant value for $ Z $. Similar to partial correlation, we first compute the
-residuals for $ X $ and $ Y $ using $ \bm{Z} $ as the predictor variables. We
-use the mixed data residualization approach from \citet{Ankan2023}. Depending
-on the type of dependent variable, the residuals are defined as follows:
+citation} to mixed data. To compute partial correlations we use: 1) A
+residualization method, and 2) a measure of associations between the residuals.
+In the case of continuous data, the residuals can be computed using a
+regression based method and Pearson's correlation coefficient can be used for
+computing the association between the residuals. We extend this idea to mixed
+data by combining a mixed data residualization method \citep{Ankan2023} with a
+multivariate measure of association based, Pillai's Trace \citep{Pillai1955}.
+Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $, and $
+\bm{Z} $, we are interested in computing an estimate of the conditional
+association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $ conditioned on $
+\bm{Z} $. As the residualization approach requires atleast one conditional
+variable to be present, if $ \bm{Z} = \emptyset $, we use a constant value for
+$ Z $. Similar to partial correlation, we first compute the residuals for $ X $
+and $ Y $ using $ \bm{Z} $ as the predictor variables. We use the mixed data
+residualization approach from \citet{Ankan2023}. Depending on the type of
+dependent variable, the residuals are defined as follows:
 

### 2024-08-08 paper/main.tex
Updates

@@
 residualization method, and 2) a measure of associations between the residuals.
-In the case of continuous data, the residuals can be computed using a
-regression based method and Pearson's correlation coefficient can be used for
-computing the association between the residuals. We extend this idea to mixed
-data by combining a mixed data residualization method \citep{Ankan2023} with a
-multivariate measure of association based, Pillai's Trace \citep{Pillai1955}.
+In the case when both $ X $ and $ Y $ are continuous variables, the residuals
+are typically computed using a regression based method and  Pearson's
+correlation coefficient is used as a measure of association between the
+residuals. To extend this idea to mixed data, we combining a mixed data
+residualization method \citep{Ankan2023} with a multivariate measure of
+association based, Pillai's Trace \citep{Pillai1955}.
+
 Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $, and $
@@
 association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $ conditioned on $
-\bm{Z} $. As the residualization approach requires atleast one conditional
-variable to be present, if $ \bm{Z} = \emptyset $, we use a constant value for
-$ Z $. Similar to partial correlation, we first compute the residuals for $ X $
-and $ Y $ using $ \bm{Z} $ as the predictor variables. We use the mixed data
-residualization approach from \citet{Ankan2023}. Depending on the type of
-dependent variable, the residuals are defined as follows:
+\bm{Z} $. To compute the residuals, we utilize the residualization approach
+from \citet{Ankan2023}. Depending on the type of dependent variable, the
+residuals are defined as follows:
 
@@
 
-Here, we have the option to choose the estimators based on the properties of
+We repeat the above step for the variable $ Y $ and compute its residual $ R_Y
+$. Here, we have the option to choose the estimators based on the properties of
 their data. Non-parametric ensemble estimators such as Random Forest and
@@
 \begin{definition}
+
 	Given two sets of random variables $ \bm{U} = (U_1, U_2, \cdots, U_p) $
@@
 
-	\begin{equation}
-		\phi(\bm{U}, \bm{V}) = \sup_{a, b} \frac{a^T \Sigma_{\bm{U}\bm{V}} b}{\sqrt{a^T \Sigma_{\bm{U}\bm{U}} a} \sqrt{b^T \Sigma_{\bm{V}\bm{V}} b}}
-	\end{equation}
+	% \begin{equation}
+	% 	\phi(\bm{U}, \bm{V}) = \sup_{a, b} \frac{a^T \Sigma_{\bm{U}\bm{V}} b}{\sqrt{a^T \Sigma_{\bm{U}\bm{U}} a} \sqrt{b^T \Sigma_{\bm{V}\bm{V}} b}}
+	% \end{equation}
+
+\begin{equation*}
+	\begin{split}
+		\bm{\rho}(\bm{U}, \bm{V}) &= \sqrt{\mathrm{E} \left( \Sigma_{\bm{UU}}^{-1} \Sigma_{\bm{UV}} \Sigma_{\bm{VV}}^{-1} \Sigma_{\bm{VU}} \right)}
+	\end{split}
+\end{equation*}
 
 \end{definition}
+
+where $ \mathrm{E}(\Sigma) $ denotes the eigenvalues of the matrix $ \Sigma $.
+$ \Sigma_{UU} $ and $ \Sigma_{VV} $ are the covariance matrices of $ \bm{U} $
+and $ \bm{V} $, and $ \Sigma_{UV} $ and $ \Sigma_{VU} $ are the
+cross-covariance matrices.
 

### 2024-08-08 paper/main.tex
Updates
updates

@@
 
-In this section, we develop a novel measure of partial association for mixed
-data by extending partial correlations \citep{Whittaker2009} \todo{Verify this
-citation} to mixed data. To compute partial correlations we use: 1) A
-residualization method, and 2) a measure of associations between the residuals.
-In the case when both $ X $ and $ Y $ are continuous variables, the residuals
-are typically computed using a regression based method and  Pearson's
-correlation coefficient is used as a measure of association between the
-residuals. To extend this idea to mixed data, we combining a mixed data
-residualization method \citep{Ankan2023} with a multivariate measure of
-association based, Pillai's Trace \citep{Pillai1955}.
+In this section, we introduce a novel measure of partial association for mixed
+data by extending the concept of partial correlations \citep{Whittaker2009}
+\todo{Verify this citation} to handle mixed data types. To compute partial
+correlations, we have two key steps: 1) a residualization method, and 2) a
+measure of associations between the residuals. When both $ X $ and $ Y $ are
+continuous variables, residuals are typically computed using a regression-based
+method, and Pearson's correlation coefficient is used to measure the association
+between these residuals. To extend this approach to mixed data, we combine a mixed
+data residualization method \citep{Ankan2023} with a multivariate measure of
+association based on canonical correlations, Pillai's Trace \citep{Pillai1955}.
 
 Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $, and $
-\bm{Z} $, we are interested in computing an estimate of the conditional
+\bm{Z} $, we aim to estimate the conditional
 association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $ conditioned on $
-\bm{Z} $. To compute the residuals, we utilize the residualization approach
-from \citet{Ankan2023}. Depending on the type of dependent variable, the
-residuals are defined as follows:
+\bm{Z} $. We use the residualization approach from \citet{Ankan2023} to compute 
+the residuals. The residuals are defined as follows, depending on the type of 
+dependent variable:
 
@@
 
-\begin{equation*}
+\begin{equation}
 	\begin{split}
@@
 	\end{split}
-\end{equation*}
+\end{equation}
 
@@
 \begin{itemize}
-	\item Wilks' Lambda: $ \Lambda = \prod (1 - \phi_i^2) $
-	\item Roy's Largest Root: $ \theta = \phi_{\max}^2 $
-	\item Pillai's Trace: $ \tau = \sum \phi_i^2 $
+	\item Wilks' Lambda:
+		$$ \Lambda = \prod (1 - \phi_i^2) $$
+	\item Roy's Largest Root: 
+		$$ \theta = \phi_{\max}^2 $$
+	\item Pillai's Trace: 
+		$$ \tau = \sum \phi_i^2 $$
 \end{itemize}
@@
 
-$$ \rho(X, Y; \bm{Z}) = \sum_{i=1}^{min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\phi_i(R_X, R_Y))^2 $$
+\begin{equation}
+	\rho(X, Y; \bm{Z}) = \frac{1}{\min(\rvert R_X \rvert, \rvert R_Y \rvert)}
+	\sum_{i=1}^{\min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\phi_i(R_X, R_Y))^2
+\end{equation}
 
@@
 \begin{enumerate}
-	\item Bounded between $ 0 $ and $ 1 $.
-	\item Independent of sample size.
-	\item Invariant to the one-hot encoding scheme used fr categorical variables.
-	\item Reduces to partial correlations when both variables are continuous.
-	\item In case of discrete variables, connected to Cramer\'s V, a well
-		known effect size measure for discrete variables.
+	\item \textbf{Bounded: } $ \rho $ is bounded between $ 0 $ and $ 1 $,
+		with $ 0 $ representing no association while $ 1 $ representing
+		perfect linear association. This helps in interpretation and
+		comparing values across multiple partial associations.
+	\item \textbf{Independent of sample size: }
+	\item \textbf{Invariant to the dummy encoding scheme: } As we use dummy
+		encoding to compute the residuals for categorical variables,
+		but as canonical correlations search for linear combinations
+		over the columns, the measure is invariant to the dummy
+		encoding scheme that we use.
+	\item \textbf{Generalizes partial correlations to mixed data: }
+		As canonical correlation is equivalent to Pearson's correlation
+		coefficient, when both $ X $ and $ Y $ are continuous, $ \rho $
+		is equivalent to canonical correlations. Hence, $ \rho $ is a
+		generalization of partial correlations to mixed data.
+	\item \textbf{Related to Cramer\'s V: }
 \end{enumerate}

### 2024-08-08 paper/main.tex
Updates

@@
 \begin{enumerate}
-	\item \textbf{Continuous:} We train a regression
-		model, $ E_X: x \sim \bm{z} $. The residuals are then defined
-		as the difference between the true and the predicted value using
-		$ E_X $.
+	\item \textbf{Continuous:} We train a regression model, $ E_X: x \sim
+		\bm{z} $. The residuals are then defined as the difference
+		between the true and the predicted values using $ E_X $. 
 		$$ R_{x_i} = x_i - E_X(\bm{z}_i) $$
-	\item \textbf{Ordinal:} We start by training a probability
-		estimator, $ p_X: x \sim \bm{z} $, and take probability
-		predictions from it $ \hat{p}_X: \hat{p}_X(\bm{Z}) $. Then we
-		compute the residuals as follows:
+	\item \textbf{Ordinal:} We start by training a probability estimator, $
+		p_X: x \sim \bm{z} $, and obtain probability predictions 
+		$ \hat{p}_X: \hat{p}_X(\bm{Z}) $. Then, we compute the residuals
+		as follows \citep{Li2012}:
 		$$ R_{x_i} = \hat{p}(\hat{x}_i > x_i) - \hat{p}(\hat{x}_i < x_i) $$
-	\item \textbf{Categorical:} We again start by training a
-		probability estimator $ p_X: X \sim \bm{Z} $, and take the
-		probability estimates from it $ \hat{p}_X: p_Z(\bm{Z}) $. Next,
-		we one-hot encode the categorical variable and then compute the
-		residuals as follows: 
+	\item \textbf{Categorical:} We again start by training a probability
+		estimator $ p_X: X \sim \bm{Z} $, and obtain probability
+		estimates $ \hat{p}_X: p_Z(\bm{Z}) $. Next, we dummy encode the
+		categorical variable and then compute the residuals as follows: 
 		$$ R_X = x_i - \hat{p}(\bm{z}_i) $$
@@
 $. Here, we have the option to choose the estimators based on the properties of
-their data. Non-parametric ensemble estimators such as Random Forest and
-XGBoost can work on any combination of data types while performing reasonably
-well. For linear datasets we can also choose variants of linear regression
-depending on the type of data. After the above steps are repeated for both $ X
-$ and $ Y $, we end up with two residual matrices: $ R_X $ and $ R_Y $. The
-type of the variable determines the shape for these matrices, i.e., if the
-variable is continuous we get a residual matrix of shape $ ( n \times 1 ) $ and
-if the variable is categorical we have a residual matrix of shape $ ( n \times
-(k-1)) $ (where one of the dummy encoded variables is dropped to avoid
-multicollinearity)
+the data. Non-parametric ensemble estimators such as Random Forest and XGBoost
+can handle any combination of data types while performing reasonably well. For
+linear datasets, we can also choose variants of linear regression depending on
+the type of data.
 
-We now have two continuous valued residual matrices and the second step now is
+After the above steps are repeated for both $ X $ and $ Y $, we end up with two
+residual matrices: $ R_X $ and $ R_Y $. The type of the variable determines the
+shape for these matrices. If the variable is continuous, we get a residual
+matrix of shape $ ( n \times 1 ) $ and if the variable is categorical, we have
+a residual matrix of shape $ ( x \times (k -1)) $ where one of the dummy encoded
+variables is dropped to avoid multicollinearity.
+
+We now have two continuous-valued residual matrices. The next step is
 to quantify the association between these residuals. In multivariate
-statistics, canonical correlation based measures have been used for measuring
-association between two sets of variables. We can treat the residuals as sets
-of variables in case of categorical variable and use canonical correlations to
-quantify the association between them.
+statistics, canonical correlation based measures have been used to measure
+the association between two sets of variables. We can treat our residuals
+as sets of variables and use canonical correlations to quantify the association
+between them.
 
@@
 	and $ \bm{V} = (V_1, V_2, \cdots, V_q) $, canonical correlation between
-	them, $\phi(\bm{U}, \bm{V}) $ is defined as:
+	them, $\rho(\bm{U}, \bm{V}) $ is defined as:
 
@@
 
+% \begin{equation}
+% 	\begin{split}
+% 		\bm{\rho}(\bm{U}, \bm{V}) &= \sqrt{\mathrm{E} \left( \Sigma_{\bm{UU}}^{-1} \Sigma_{\bm{UV}} \Sigma_{\bm{VV}}^{-1} \Sigma_{\bm{VU}} \right)}
+% 	\end{split}
+% \end{equation}
 	\begin{equation}
-	\begin{split}
-		\bm{\rho}(\bm{U}, \bm{V}) &= \sqrt{\mathrm{E} \left( \Sigma_{\bm{UU}}^{-1} \Sigma_{\bm{UV}} \Sigma_{\bm{VV}}^{-1} \Sigma_{\bm{VU}} \right)}
-	\end{split}
+		\rho(\bm{U}, \bm{V}) = \max_{a, b} Corr(a^T \bm{U}, b^T \bm{V})
 	\end{equation}
 
+	where $ a $ and $ b $ are vectors of coefficients that maximize the correlations
+	between the linear combinations of $ a^T \bm{U} $ and $ b^T \bm{V} $.
 \end{definition}
 
-where $ \mathrm{E}(\Sigma) $ denotes the eigenvalues of the matrix $ \Sigma $.
-$ \Sigma_{UU} $ and $ \Sigma_{VV} $ are the covariance matrices of $ \bm{U} $
-and $ \bm{V} $, and $ \Sigma_{UV} $ and $ \Sigma_{VU} $ are the
-cross-covariance matrices.
+% where $ \mathrm{E}(\Sigma) $ denotes the eigenvalues of the matrix $ \Sigma $.
+% $ \Sigma_{UU} $ and $ \Sigma_{VV} $ are the covariance matrices of $ \bm{U} $
+% and $ \bm{V} $, and $ \Sigma_{UV} $ and $ \Sigma_{VU} $ are the
+% cross-covariance matrices.
 
-Canonical correlations are generalization of correlations to multi-dimensional
-variables. It tries to find orthogonal linear transformations $ a $ and $ b $
-of $ \bm{U} $ and $ \bm{V} $ such that the correlation between the transformed
-variables $ a^T \bm{U} $ and $ b^T \bm{V} $ is maximized. It gives a vector of
-correlation values of size $ \min(\rvert U \rvert, \rvert V \rvert) $
-representing the correlation coefficient of each column of the transformed
-variables. Pearson's correlation coefficient is a special case of canonical
-correlations when $ \rvert U \rvert = \rvert V \rvert = 1$.
+Canonical correlations generalize correlations to multi-dimensional variables. They
+find orthogonal linear transformations $ a $ and $ b $ of $ \bm{U} $ and $ \bm{V} $
+such that the correlation between the transformed variables $ a^T \bm{U} $ and 
+$ b^T \bm{V} $ is maximized. This yields a vector of correlation values of 
+size $ \min(p, q) $, representing the correlation coefficient of each pair of 
+transformed variables. Pearson's correlation coefficient is a special case of 
+canonical correlations when $ p = q = 1 $.
 
 Many measures of association based on canonical correlation have been used in
-multivariate statistics with each of them having slightly different properties.
+multivariate statistics, each with slightly different properties.
 \begin{itemize}
 	\item Wilks' Lambda:
-		$$ \Lambda = \prod (1 - \phi_i^2) $$
+		$$ \Lambda = \prod_{i}^{\min(p, q)} (1 - \phi_i^2) $$
 	\item Roy's Largest Root: 
-		$$ \theta = \phi_{\max}^2 $$
+		$$ \theta = \max_i(\phi_i^2) $$
 	\item Pillai's Trace: 
-		$$ \tau = \sum \phi_i^2 $$
+		$$ \tau = \sum_{i=1}^{\min(p, q)} \phi_i^2 $$
 \end{itemize}
@@
 We use Pillai's Trace for our purpose for two main reasons: 1) It uses all the
-canoncial correlation values, 2) Its interpretation is similar to correlation
+canonical correlation values, 2) Its interpretation is similar to correlation
 coefficient, i.e., $ 0 $ signifies no association and $ 1 $ signifies perfect
-linear relationship. This gives us the following partial measure of association:
+linear relationship. This gives us the following partial measure of
+association:
 
@@
 	\item \textbf{Bounded: } $ \rho $ is bounded between $ 0 $ and $ 1 $,
-		with $ 0 $ representing no association while $ 1 $ representing
+		with $ 0 $ representing no association while $ 1 $ represents
 		perfect linear association. This helps in interpretation and
 		comparing values across multiple partial associations.
-	\item \textbf{Independent of sample size: }
-	\item \textbf{Invariant to the dummy encoding scheme: } As we use dummy
-		encoding to compute the residuals for categorical variables,
-		but as canonical correlations search for linear combinations
-		over the columns, the measure is invariant to the dummy
-		encoding scheme that we use.
-	\item \textbf{Generalizes partial correlations to mixed data: }
-		As canonical correlation is equivalent to Pearson's correlation
-		coefficient, when both $ X $ and $ Y $ are continuous, $ \rho $
-		is equivalent to canonical correlations. Hence, $ \rho $ is a
-		generalization of partial correlations to mixed data.
+	\item \textbf{Independent of Sample Size: } The measure remains consistent
+		regardless of the sample size ensuring reliability across 
+		studies with different sample sizes.
+	\item \textbf{Invariant to the Dummy Encoding Scheme: } Since canonical
+		correlations search for linear combinations over the columns,
+		the measure is invariant to the specific dummy encoding
+		scheme used for categorical variables.
+	\item \textbf{Generalizes Partial Correlations to Mixed Data: }
+		Canonical correlations is equivalent to Pearson's correlation coefficient
+		when both $ X $ and $ Y $ are continuous. Hence, $ \rho $ generalizes
+		partial correlations to mixed data types.
 	\item \textbf{Related to Cramer\'s V: } 
 \end{enumerate}
-\todo[inline]{Expand on these points}
 

### 2024-08-13 paper/main.tex
Updates

@@
 
-\todo[inline]{Add screenshots here}
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.4]{plots/web_tool_full.png}
+	\caption{A screenshot of the web tool for constructing the model. Users
+		can upload their dataset after which the tool creates an empty
+		graph and shows all pair of variables which are associated in
+		the model using undirected red edges with the strength of
+		association represented using edge width. Users can then add
+		edges to the model (shown in green). Unnecessary edges are shown in 
+		black.}
+\end{figure}
 

### 2024-09-10 paper/main.tex
Updates

@@
 algorithms for learning DAGs from data. Despite these advancements, the
-application of these algorithms in practical domains remains limited, with
+application of these algorithms in applied domains remains limited, with
 researchers often preferring to manually construct DAGs based on their expert
-knowledge. Even when using these algorithms, researchers must rely on their
-expertise to correct mistakes or orient edges before the models can be used for
-further analysis. Given this reliance on manual model construction and
-modification, we propose a method inspired by specification search in
-Structural Equation Models (SEMs) to assist researchers in iteratively
-modifying DAGs. Our method employs implied Conditional Independence (CI) based
-model testing, combined with a novel measure of partial association for mixed
-data, to rank violations of implied CIs of the model in the data. This ranking
-helps users to prioritize pair of variables between which to add or remove
-edges. Users can then decide to remove or add edges while determining the
-correct orientation of the edge based on their expertise. Empirically, we
-demonstrate that a greedy approach (i.e., fixing the worst violation first)
-achieves performance comparable to automated algorithms, provided the user can
-correctly determine the edge orientation in $ 1 $ out of $ 3 $ edge additions.
-Additionally, we show that in the absence of an expert, a Large Language Model
-(LLM) can potentially be used to determine edge orientations. 
+knowledge. This preference stems due to some practical issues with the
+automated algorithms such as they can make obvious mistakes, and they output
+the Markov Equivalence Class (MEC). Due to these issues, when using these
+algorithms, users typically have to rely on expert knowledge to fix the
+mistakes and to orient the MEC to a DAG. Given this reliance on manual model
+construction, we propose an iterative method to assist researchers in manually
+constructing DAGs by combining implied Conditional Independence (CI) based
+model testing and utilizing their expert knowledge. Our approach is inspired by
+specification search in Structural Equation Models (SEMs), where we utilize a
+novel measure of partial association for mixed data to rank violations of
+implied Conditional Independences in the model. Using this ranking users can
+prioritize the pair of variables which need to be corrected by either adding an
+edge between them or removing the edge based on their expertise to specify the
+correct orientation. Empirically, we demonstrate that a greedy approach (i.e.,
+fixing the worst violation first) achieves performance comparable to automated
+algorithms, provided the user can correctly determine the edge orientation in $
+1 $ out of $ 3 $ edge additions. Additionally, we show that in the absence of
+an expert, a Large Language Model (LLM) can potentially be used to determine
+edge orientations. 
 \end{abstract}
@@
 \emph{Causal Discovery} methods aim to uncover these relationships among random
-variables from observational data. This problem has been extensively explored
-in both the Directed Acyclic Graphs (DAGs) and Structural Equation Models
-(SEMs) literatures, each adopting a different approach.
+variables from observational data. Methods for causal discovery have been
+developed in both the Directed Acyclic Graphs (DAGs) and Structural Equation
+Models (SEMs) literatures, each adopting a different approach.
 

### 2024-09-10 code/llm.py
Updates code

@@ -12,17 +12,18 @@ model = genai.GenerativeModel(model_name="gemini-1.5-flash")
 
 
 def query_direction(u, v, descriptions):
-    prompt = f""" You are an expert in social sciences. You are given two variables with the following descriptions:
+    prompt = f""" You are an expert in Social Science. Following are the descriptions of two variables:
+
         <A>: {descriptions[u]}
         <B>: {descriptions[v]}
 
-        Which of the following two options is the most likely causal direction between them:
+        Which of the following two options is the most likely causal direction between these variables:
+
         1. <A> causes <B>
         2. <B> causes <A>
 
-        Return a single letter answer between the choices above. I do not need the reasoning behind it. Do not add any formatting in the answer.
+        Return a single letter answer between the choices above; Do not provide any reasoning in the answer; Do not add any text formatting to the answer.
         """
-
     response = model.generate_content([prompt])
     response_txt = response.text.strip().lower().replace('*', '')
     if response_txt in ('a', '1'):

### 2024-09-10 code/local_maximas.py

@@ -28,3 +28,11 @@ d3_score <- score(model2network(convert(d3, "bnlearn")), df3, "bic-g")
 d3_est <- hc(df3, start=model2network("[E][Y|X:E][Z|E][X|E]"), score='bic-g')
 # d3_est <- hc(df3, score='bic-g')
 d3_est_score <- score(d3_est, df3, 'bic-g')
+
+# Example 4:
+d4 <- dagitty("dag{X -> Y -> Z -> A}")
+df4 <- simulateSEM(d4, N=1e3)
+
+d4_score <- score(model2network(convert(d4, "bnlearn")), df4, "bic-g")
+d4_est <- hc(df4, start=model2network("[Y|X][Z|A][A][X]"), score="bic-g")
+d4_est_score <- score(d4_est, df4, 'bic-g')

### 2024-09-10 web/api.R

@@ -1,5 +1,32 @@
 library(jsonlite)
 
+preprocess_adult <- function(d){
+	d$X <- NULL
+	d$Fnlwgt <- NULL
+	d$EducationNum <- NULL
+	d$CapitalGain <- NULL
+	d$CaptialLoss <- NULL
+
+	d$Workclass <- factor(d$Workclass, ordered=F)
+	d$MaritalStatus <- factor(d$MaritalStatus, ordered=F)
+	d$Occupation <- factor(d$Occupation, ordered=F)
+	d$Relationship <- factor(d$Relationship, ordered=F)
+	d$Race <- factor(d$Race, ordered=F)
+	d$Sex <- factor(d$Sex, ordered=F)
+	d$NativeCountry <- factor(d$NativeCountry, ordered=F)
+	d$Income <- factor(d$Income, ordered=F)
+	
+	d$Age <- as.double(d$Age)
+	d$HoursPerWeek <- as.double(d$HoursPerWeek)
+	
+	education_levels = c( "Preschool", "1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th","12th", "HS-grad", "Some-college", "Assoc-voc", "Assoc-acdm", "Bachelors", "Masters", "Prof-school", "Doctorate" )
+	d$Education <- factor(d$Education, levels=education_levels, ordered=T)
+	d <- d[complete.cases(d), ]
+	d <- d[1:1000, ]
+	
+	return(d)
+}
+
 #* @filter cors
 cors <- function(res) {
     res$setHeader("Access-Control-Allow-Origin", "*") # Or whatever
@@ -20,7 +47,6 @@ function(req, res, file) {
   filename = names(file)
   # Read the uploaded CSV file
   csv_data <- read.csv(text=file[[filename]])
-  # dataset <<- csv_data
   dataset_name <- paste0(sample(c(LETTERS, tolower(LETTERS)), 20, T), collapse='')
   datasets[[dataset_name]] <<- csv_data
 
@@ -46,7 +72,7 @@ run_citests <- function( req, res, dag, threshold, pval ){
 		g <- dagitty::dagitty(dag)
 		r <- c()
 		nn <- names(g)
-		dataset <- datasets[[1]]
+		dataset <- preprocess_adult(datasets[[1]])
 		for( n1i in seq(1,length(nn)-1,by=1) ){
 			n1 <- nn[n1i]
[…]

### 2024-11-26 code/fig3/sl.R
Adds results

@@ -120,7 +120,7 @@ run_single_exp_hc <- function(n_nodes, edge_prob){
 	### Temp code block #################################################
 	shd <- causalDisco::shd(hc_adj, true_adj)
 	sid <- SID::structIntervDist(trueGraph=true_adj, estGraph=hc_adj)
-	return(shd, shd, sid$sidLowerBound, sid$sidUpperBound)
+	return(c(shd, shd, sid$sidLowerBound, sid$sidUpperBound))
 
 	#####################################################################
 
@@ -161,7 +161,7 @@ run_single_exp_pc <- function(n_nodes, edge_prob){
 	### Temp code block #################################################
 	shd <- causalDisco::shd(pc_adj, true_adj)
 	sid <- SID::structIntervDist(trueGraph=true_adj, estGraph=pc_adj)
-	return(shd, shd, sid$sidLowerBound, sid$sidUpperBound)
+	return(c(shd, shd, sid$sidLowerBound, sid$sidUpperBound))
 
 	#####################################################################
 
@@ -192,7 +192,7 @@ run_single_exp_ges <- function(n_nodes, edge_prob){
 	### Temp code block #################################################
 	shd <- causalDisco::shd(ges_adj, true_adj)
 	sid <- SID::structIntervDist(trueGraph=true_adj, estGraph=ges_adj)
-	return(shd, shd, sid$sidLowerBound, sid$sidUpperBound)
+	return(c(shd, shd, sid$sidLowerBound, sid$sidUpperBound))
 
 	#####################################################################
 	ges_pdag <- pcalg::dag2cpdag(ges_adj)

### 2024-11-26 paper/main.tex

@@
 goal in many scientific inquiries, as it reveals the underlying mechanisms of
-observed phenomena and informs effective interventions or policy decisions.
-\emph{Causal Discovery} methods aim to uncover these relationships among random
-variables from observational data. Methods for causal discovery have been
-developed in both the Directed Acyclic Graphs (DAGs) and Structural Equation
-Models (SEMs) literatures, each adopting a different approach.
+observed phenomena and help in determining effective interventions or policy
+decisions. \emph{Causal Discovery} methods aim to uncover these relationships
+among random variables from observational data. Methods for causal discovery
+have been developed in both the Directed Acyclic Graphs (DAGs) and Structural
+Equation Models (SEMs) literatures, each adopting a different approach.
 
@@
 algorithms to learn causal structures from datasets. As a result of these
-efforts, numerous algorithms for causal discovery have been proposed, such as
+efforts, many algorithms for causal discovery have been proposed, such as
 constraint-based methods like the PC algorithm \citep{Spirtes2001} and Fast
@@
 and continuous optimization-based methods like NO TEARS \citep{Zheng2018} and
-DAGMA \citep{Bello2022}.
-
-In contrast, the SEM literature has focused more on developing tools to assist
-researchers in manually constructing or modifying models while allowing them to
-integrate their expert knowledge into the model-building process. Researchers
-typically develop an initial model based on their expertise and then use
-various tools to guide modifications to improve the model's fit to data. This
-model construction process is commonly known as \emph{Specification Search}
-\citep{Long1983} and uses method like modification indices, and Wald-based
-tests \citep{Marcoulides2018}.
+DAGMA \citep{Bello2022}. In contrast, the SEM literature has focused more on
+developing tools to assist researchers in manually constructing or modifying
+models while allowing them to integrate their expert knowledge into the
+model-building process. Researchers typically develop an initial model based on
+their expertise and then use various tools to guide modifications to improve
+the model's fit to data. This model construction process is commonly known as
+\emph{Specification Search} \citep{Long1983} and uses method like modification
+indices, and Wald-based tests \citep{Marcoulides2018}.
 
@@
 		not well understood. Moreover, their output can vary
-		significantly depending on the algorithm and their
+		significantly depending on the chosen algorithm and their
 		hyperparameters. Combined with the lack of robust methods to
@@
 		usually have to rely on post-hoc modifications to the model
-		based on their domain expertise.
+		based on model testing and their domain expertise.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
@@
 
-Figure~\ref{fig:intro} shows some of these issues when applied to the Adult
-Income dataset. Due to these challenges, whether constructing the model
-manually or using an algorithm, we typically have to rely on some expert
-knowledge to get a model that we can use for further analysis. Hence, it is
-crucial to have tools to guide this construction or modification process.
+Figure~\ref{fig:intro} shows an example of application of a causal discovery
+algorithms namely PC on the Adult Income dataset. \todo[inline]{Explain}.
 
-A common method for specification search is using \emph{modification indices},
-which identify potential changes to the model structure that improve the model
-fit. These potential changes are then ranked by the ones that improve the fit
-the most. This ranking allows the researchers to focus first on the suggestions
-that can improve the model the most and while integrating their expertise to
-choose the most optimal modifications.
+Due to these challenges, whether constructing the model manually or using an
+algorithm, we typically have to rely on some expert knowledge to get a model
+that we can use for further analysis. Hence, it is crucial to have tools to
+guide this construction or modification process.
 
@@
 tests, it can be difficult to figure which test to focus on.
+
+A common method for specification search is using \emph{modification indices},
+which identify potential changes to the model structure that improve the model
+fit. These potential changes are then ranked by the ones that improve the fit
+the most. This ranking allows the researchers to focus first on the suggestions
+that can improve the model the most and while integrating their expertise to
+choose the most optimal modifications.
 
@@
 association between $ X $ and $ Y $.
+
+\subsection{Ranking using Correlation Coefficients}
 

### 2024-11-27 code/fig3/make_plots.R
Final plots

@@ -162,8 +162,8 @@ make_plots_v2 <- function(filename, plot_type){
 	d_shd <- as.data.frame(cbind(c(d_shd$edge_prob, d_shd$edge_prob, d_shd$edge_prob),
 	      	       c(d_shd$oracle_acc, d_shd$oracle_acc, d_shd$oracle_acc),
 	      	       c(d_shd$ges_mean, d_shd$hc_mean, d_shd$pc_mean, d_shd$human_mean),
-	      	       c(d_shd$ges_mean, d_shd$hc_lower, d_shd$pc_lower, d_shd$human_lower),
-	      	       c(d_shd$ges_mean, d_shd$hc_upper, d_shd$pc_upper, d_shd$human_upper),
+	      	       c(d_shd$ges_lower, d_shd$hc_lower, d_shd$pc_lower, d_shd$human_lower),
+	      	       c(d_shd$ges_upper, d_shd$hc_upper, d_shd$pc_upper, d_shd$human_upper),
 		       c(rep('GES', nrow(d_shd)), rep('Hill-Climb Search', nrow(d_shd)), rep('PC', nrow(d_shd)), rep('Expert', nrow(d_shd))),
 		       c(rep('ges', nrow(d_shd)), rep('hc', nrow(d_shd)), rep('pc', nrow(d_shd)), paste('human', d_shd$oracle_acc)),
 		       c(rep(NA, nrow(d_shd)), rep(NA, nrow(d_shd)), rep(NA, nrow(d_shd)), d_shd$oracle_acc)
@@ -184,7 +184,7 @@ make_plots_v2 <- function(filename, plot_type){
 		geom_line(alpha=0.4, show.legend=T) +
 		geom_point(data=subset(d_shd, !(algo %in% c('GES', 'PC', 'Hill-Climb Search'))), pch=19, size=1, alpha=0.4, show.legend=F) +
 		geom_ribbon(linetype=0, alpha=0.4, show.legend=F) + 
-		# geom_dl(aes(label=line_labels), color='black', method=list(dl.trans(x=x+.1), "last.points", cex=0.5)) + 
+		geom_dl(aes(label=line_labels), color='black', method=list(dl.trans(x=x+.1), "last.points", cex=0.5)) + 
       		theme_minimal(base_size = 8) + 
 		theme(legend.position='top',
 		      legend.title=element_blank(),

### 2024-11-27 code/fig3/sl.R

@@ -211,7 +211,7 @@ run_sim <- function(R, n_nodes, edge_probs, oracle_accs){
 	for (edge_prob in edge_probs){
 		ges_dist <- t(future_replicate(R, run_single_exp_ges(n_nodes=n_nodes, edge_prob=edge_prob)))
 		ges_mean <- apply(ges_dist, function(x) mean(x), MARGIN=2)
-		ges_sd <- apply(ges_dist, function(x) sd(x), MARGIN=2)
+		ges_sd <- apply(ges_dist, function(x) sd(x), MARGIN=2)/sqrt(R)
 
 		print("Done with GES")
 

### 2024-11-27 paper/main.tex
Updates plot and makefile

@@
 
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.4]{plots/web_tool_full.png}
-	\caption{A screenshot of the web tool for constructing the model. Users
-		can upload their dataset after which the tool creates an empty
-		graph and shows all pair of variables which are associated in
-		the model using undirected red edges with the strength of
-		association represented using edge width. Users can then add
-		edges to the model (shown in green). Unnecessary edges are shown in 
-		black.}
-\end{figure}
+% \begin{figure}
+% 	\centering
+% 	\includegraphics[scale=0.4]{plots/web_tool_full.png}
+% 	\caption{A screenshot of the web tool for constructing the model. Users
+% 		can upload their dataset after which the tool creates an empty
+% 		graph and shows all pair of variables which are associated in
+% 		the model using undirected red edges with the strength of
+% 		association represented using edge width. Users can then add
+% 		edges to the model (shown in green). Unnecessary edges are shown in 
+% 		black.}
+% \end{figure}
 
@@
 		\centering
-		\includegraphics{../code/plots/shd_ribbon.pdf}
+		\includegraphics{../code/fig3/shd_ribbon.pdf}
 		\caption{}
@@
 		\centering
-		\includegraphics{../code/plots/sid_ribbon.pdf}
+		\includegraphics{../code/fig3/sid_ribbon.pdf}
 		\caption{}

### 2024-11-27 code/utils/data.py
Fixes path in data.py

@@ -2,7 +2,7 @@ import pandas as pd
 
 
 def get_adult_df():
-    df = pd.read_csv("adult_proc.csv", index_col=0)
+    df = pd.read_csv("/home/ankur/work/2024-human-sl/code/utils/adult_proc.csv", index_col=0)
     df.Age = pd.Categorical(
         df.Age,
         categories=["<21", "21-30", "31-40", "41-50", "51-60", "61-70", ">70"],

### 2024-11-27 code/fig5/figure4.py
results

@@ -52,9 +52,9 @@ if __name__ == "__main__":
     with open('results/llm_edge_list.txt', 'w') as f:
         f.write(str(list(dag.edges())))
     with open('results/unexplained_effect.txt', 'w') as f:
-        f.write(str(list(total_unexplained_effect)))
+        f.write(str(total_unexplained_effect.tolist()))
         f.write('\n')
-        f.write(str(list(total_ll)))
+        f.write(list(total_ll.tolist()))
 
     make_plot()
 

### 2024-11-27 code/fig5/figure4.py
Plot for unexplained effect

@@ -52,9 +52,9 @@ if __name__ == "__main__":
     with open('results/llm_edge_list.txt', 'w') as f:
         f.write(str(list(dag.edges())))
     with open('results/unexplained_effect.txt', 'w') as f:
-        f.write(str(total_unexplained_effect.tolist()))
+        f.write(str(total_unexplained_effect.tolist())[1:-1])
         f.write('\n')
-        f.write(list(total_ll.tolist()))
+        f.write(str(total_ll.tolist())[1:-1])
 
     make_plot()
 

### 2024-11-27 code/fig5/make_plot.R

@@ -1,0 +1,20 @@
+library(ggplot2)
+
+
+d <- read.csv('results/unexplained_effect.txt', header=F)
+d <- t(d)
+colnames(d) <- c('pillai', 'log-like')
+rownames(d) <- seq(1, nrow(d))
+d <- as.data.frame(d)
+d <- cbind(d, x=seq(1, nrow(d)))
+
+p <- ggplot(d, aes(x=x, y=pillai)) +
+	geom_line(alpha=0.6, show.legend=F) +
+	geom_point() +
+	theme_minimal(base_size=8) +
+	labs(x="Iteration No.") +
+	labs(y="Unexplained Effect")
+
+ggsave('unexplained.pdf', p, height=2.5, width=3.2, units='in')
+
+

### 2024-11-27 code/fig5/GES.py
Adds GES for comparison

@@ -1,0 +1,264 @@
+from itertools import combinations
+
+import networkx as nx
+import numpy as np
+
+from pgmpy import config
+from pgmpy.base import DAG
+from pgmpy.estimators import (
+    AIC,
+    BIC,
+    K2,
+    AICCondGauss,
+    AICGauss,
+    BDeu,
+    BDs,
+    BICCondGauss,
+    BICGauss,
+    LogLikelihoodCondGauss,
+    LogLikelihoodGauss,
+    ScoreCache,
+    StructureEstimator,
+    StructureScore,
+)
+from pgmpy.global_vars import logger
+from pgmpy.estimators import ExpertInLoop
+
+
+class GES(StructureEstimator):
+    """
+    Implementation of Greedy Equivalence Search (GES) causal discovery / structure learning algorithm.
+
+    GES is a score-based casual discovery / structure learning algorithm that works in three phases:
+        1. Forward phase: New edges are added such that the model score improves.
+        2. Backward phase: Edges are removed from the model such that the model score improves.
+        3. Edge flipping phase: Edge orientations are flipped such that model score improves.
+
+    Parameters
+    ----------
+    data: pandas DataFrame object
+        dataframe object where each column represents one variable.
+        (If some values in the data are missing the data cells should be set to `numpy.nan`.
+        Note that pandas converts each column containing `numpy.nan`s to dtype `float`.)
+
+    use_caching: boolean
+        If True, uses caching of score for faster computation.
+        Note: Caching only works for scoring methods which are decomposable. Can
+        give wrong results in case of custom scoring methods.
+
+    References
+    ----------
+    Chickering, David Maxwell. "Optimal structure identification with greedy search." Journal of machine learning research 3.Nov (2002): 507-554.
+    """
+
+    def __init__(self, data, use_cache=True, **kwargs):
+        self.use_cache = use_cache
+
+        super(GES, self).__init__(data=data, **kwargs)
+
+    def _legal_edge_additions(self, current_model):
+        """
+        Returns a list of all edges that can be added to the graph such that it remains a DAG.
+        """
+        edges = []
+        for u, v in combinations(current_model.nodes(), 2):
+            if not (current_model.has_edge(u, v) or current_model.has_edge(v, u)):
+                if not nx.has_path(current_model, v, u):
+                    edges.append((u, v))
+                if not nx.has_path(current_model, u, v):
+                    edges.append((v, u))
+        return edges
+
+    def _legal_edge_flips(self, current_model):
+        """
[…]

### 2024-11-27 code/fig5/figure4.py

@@ -12,6 +12,7 @@ from pgmpy.models import BayesianNetwork
 from data import get_adult_df
 
 from expert import ExpertInLoop
+from GES import GES
 
 
 def build_network():
@@ -39,22 +40,38 @@ def build_network():
     )
     return(dag, total_unexplained_effect, total_ll)
 
+def build_ges():
+    est = GES(get_adult_df())
+    dag, total_unexplained_effect, total_ll = est.estimate(scoring_method='bic-cg')
+    return(dag, total_unexplained_effect, total_ll)
+
 def make_plot():
+    with open('results/ges_edge_list.txt', 'r') as f:
+        edge_list = literal_eval(f.readline())
+    bn = BayesianNetwork(edge_list)
+    bn.to_graphviz().draw('results/ges_adult.pdf', prog='dot')
+
     with open('results/llm_edge_list.txt', 'r') as f:
         edge_list = literal_eval(f.readline())
     bn = BayesianNetwork(edge_list)
-    bn.to_graphviz().draw('results/llm_adult.pdf', prog='dot')
-
+    bn.to_graphviz().draw('results/expert_llm_adult.pdf', prog='dot')
 
 if __name__ == "__main__":
-    dag, total_unexplained_effect, total_ll = build_network()
+    dag, total_unexplained_effect, total_ll = build_ges()
+    with open('results/ges_edge_list.txt', 'w') as f:
+        f.write(str(list(dag.edges())))
+    with open('results/ges_unexplained_effect.txt', 'w') as f:
+        f.write(str(np.array(total_unexplained_effect).tolist())[1:-1])
+        f.write('\n')
+        f.write(str(np.array(total_ll).tolist())[1:-1])
 
+    dag, total_unexplained_effect, total_ll = build_network()
     with open('results/llm_edge_list.txt', 'w') as f:
         f.write(str(list(dag.edges())))
-    with open('results/unexplained_effect.txt', 'w') as f:
-        f.write(str(total_unexplained_effect.tolist())[1:-1])
+    with open('results/expert_unexplained_effect.txt', 'w') as f:
+        f.write(str(np.array(total_unexplained_effect).tolist())[1:-1])
         f.write('\n')
-        f.write(str(total_ll.tolist())[1:-1])
+        f.write(str(np.array(total_ll).tolist())[1:-1])
 
     make_plot()
 

### 2024-11-27 code/fig5/expert.py
Minor changes

@@ -8,7 +8,7 @@ from pgmpy import config
 from pgmpy.base import DAG
 from pgmpy.estimators import StructureEstimator
 from pgmpy.estimators.CITests import pillai_trace
-from pgmpy.estimators import LogLikelihoodCondGauss
+from pgmpy.estimators import BICCondGauss
 from pgmpy.utils import llm_pairwise_orient, manual_pairwise_orient
 
 
@@ -139,7 +139,12 @@ class ExpertInLoop(StructureEstimator):
         total_unexplained_effect = []
         total_ll = []
 
-        score_method = LogLikelihoodCondGauss(self.data)
+        score_method = BICCondGauss(self.data)
+
+        temp_all_effects = self.test_all(dag)
+        total_unexplained_effect.append(temp_all_effects[temp_all_effects.edge_present == False].effect.sum())
+        total_ll.append(score_method.score(dag))
+
         while True:
             # Step 1: Compute effects and p-values between every combination of variables.
             all_effects = self.test_all(dag)

### 2024-11-27 code/fig5/figure4.py

@@ -7,6 +7,7 @@ from itertools import combinations
 
 import pandas as pd
 import networkx as nx
+import numpy as np
 from pgmpy.models import BayesianNetwork
 
 from data import get_adult_df

### 2024-11-27 code/fig5/make_plot.R

@@ -1,14 +1,20 @@
 library(ggplot2)
 
 
-d <- read.csv('results/unexplained_effect.txt', header=F)
-d <- t(d)
-colnames(d) <- c('pillai', 'log-like')
-rownames(d) <- seq(1, nrow(d))
-d <- as.data.frame(d)
-d <- cbind(d, x=seq(1, nrow(d)))
+d_expert <- read.csv('results/unexplained_effect.txt', header=F)
+d_expert <- t(d_expert)
+colnames(d_expert) <- c('expert_pillai', 'expert_ll')
+rownames(d_expert) <- seq(1, nrow(d_expert))
+d_expert <- as.data.frame(d_expert)
+d_expert <- cbind(d_expert, x=seq(1, nrow(d_expert)))
 
-p <- ggplot(d, aes(x=x, y=pillai)) +
+d_ges <- read.csv('results/ges_unexplained_effect.txt', header=F)
+d_ges <- as.data.frame(t(d_ges))
+colnames(d_ges) <- c('ges_pillai', 'ges_ll')
+rownames(d_ges) <- seq(1, nrow(d_ges))
+d_ges <- cbind(d_ges, x=seq(1, nrow(d_ges)))
+
+p <- ggplot(d_expert, aes(x=x, y=pillai)) +
 	geom_line(alpha=0.6, show.legend=F) +
 	geom_point() +
 	theme_minimal(base_size=8) +

### 2024-11-27 code/fig5/make_plots.R
Fix plot

@@ -3,20 +3,24 @@ library(ggplot2)
 
 d_expert <- read.csv('results/unexplained_effect.txt', header=F)
 d_expert <- t(d_expert)
-colnames(d_expert) <- c('expert_pillai', 'expert_ll')
+colnames(d_expert) <- c('pillai', 'll')
 rownames(d_expert) <- seq(1, nrow(d_expert))
 d_expert <- as.data.frame(d_expert)
 d_expert <- cbind(d_expert, x=seq(1, nrow(d_expert)))
+d_expert$algo <- 'Expert'
 
 d_ges <- read.csv('results/ges_unexplained_effect.txt', header=F)
 d_ges <- as.data.frame(t(d_ges))
-colnames(d_ges) <- c('ges_pillai', 'ges_ll')
+colnames(d_ges) <- c('pillai', 'll')
 rownames(d_ges) <- seq(1, nrow(d_ges))
 d_ges <- cbind(d_ges, x=seq(1, nrow(d_ges)))
+d_ges$algo <- 'GES'
 
-p <- ggplot(d_expert, aes(x=x, y=pillai)) +
-	geom_line(alpha=0.6, show.legend=F) +
-	geom_point() +
+d_results = rbind(d_expert, d_ges)
+
+p <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
+	geom_line(alpha=0.6) +
+	geom_point(alpha=0.6, show.legend=F) +
 	theme_minimal(base_size=8) +
 	labs(x="Iteration No.") +
 	labs(y="Unexplained Effect")

### 2024-11-27 code/fig5/llm.py
Adds code for fig5

@@ -22,8 +22,7 @@ def query_direction(u, v, descriptions):
         1. <A> causes <B>
         2. <B> causes <A>
 
-        Return a single letter answer between the choices above; Do not provide any reasoning in the answer; Do not add any text formatting to the answer.
-        """
+        Return a single letter answer between the choices above; Do not provide any reasoning in the answer; Do not add any text formatting to the answer."""
     response = model.generate_content([prompt])
     response_txt = response.text.strip().lower().replace('*', '')
     if response_txt in ('a', '1'):
@@ -38,7 +37,7 @@ def query_direction(u, v, descriptions):
 
 
 def preprocess_data():
-    df = pd.read_csv("adult_proc.csv", index_col=0)
+    df = pd.read_csv("../utils/adult_proc.csv", index_col=0)
     df.Age = pd.Categorical(
         df.Age,
         categories=["<21", "21-30", "31-40", "41-50", "51-60", "61-70", ">70"],
@@ -120,12 +119,15 @@ def simulate_human(data, descriptions, pval_thres=0.05, effect_thres=0.05):
         nonedge_effects = all_effects[all_effects.edge_present == False]
         nonedge_effects = nonedge_effects[(nonedge_effects.effect >= effect_thres) & (nonedge_effects.p_val <= pval_thres)]
 
+        if len(blacklisted_edges) > 0:
+            nonedge_effects = nonedge_effects.loc[
+                    ~(((nonedge_effects.u.isin([edge[0] for edge in blacklisted_edges])) &
+                       (nonedge_effects.v.isin([edge[1] for edge in blacklisted_edges]))) |
+                      ((nonedge_effects.u.isin([edge[1] for edge in blacklisted_edges])) &
+                       (nonedge_effects.v.isin([edge[0] for edge in blacklisted_edges])))), :]
+
         if (edge_effects.shape[0] == 0) and (nonedge_effects.shape[0] == 0):
             break
-
-        if len(blacklisted_edges) > 0:
-            nonedge_effects = nonedge_effects.loc[((nonedge_effects.u in [edge[0] for edge in blacklisted_edges]) & (nonedge_effects.v in [edge[1] for edge in blacklisted_edges])) or
-                             ((nonedge_effects.u in [edge[1] for edge in blacklisted_edges]) & (nonedge_effects.v in [edge[0] for edge in blacklisted_edges])), :]
 
[…]

### 2024-11-27 code/fig5/llm.py
Adds results for llm adult

@@ -153,4 +153,6 @@ descriptions = {'Age': 'The age of a person',
                 }
 
 dag = simulate_human(preprocess_data(), descriptions)
-import ipdb; ipdb.set_trace()
+with open('llm_adult.txt', 'w') as f:
+    f.write(str(list(dag.edges())))
+dag.to_graphviz().draw('llm_adult.png', prog='dot')

### 2024-11-28 code/fig5/llm.py
Adds code for generating the plot

@@ -155,4 +155,3 @@ descriptions = {'Age': 'The age of a person',
 dag = simulate_human(preprocess_data(), descriptions)
 with open('llm_adult.txt', 'w') as f:
     f.write(str(list(dag.edges())))
-dag.to_graphviz().draw('llm_adult.png', prog='dot')

### 2024-11-28 code/fig5/make_plot.py

@@ -1,0 +1,7 @@
+from pgmpy.models import BayesianNetwork
+
+with open('adult_llm.txt', 'r') as f:
+    edge_list = eval(f.readline())
+
+dag = BayesianNetwork(edge_list)
+dag.to_graphviz().draw('llm_adult.png', prog='dot')

### 2024-11-28 code/fig4/figure4.py
Clean up code

@@ -46,17 +46,6 @@ def build_ges():
     dag, total_unexplained_effect, total_ll = est.estimate(scoring_method='bic-cg')
     return(dag, total_unexplained_effect, total_ll)
 
-def make_plot():
-    with open('results/ges_edge_list.txt', 'r') as f:
-        edge_list = literal_eval(f.readline())
-    bn = BayesianNetwork(edge_list)
-    bn.to_graphviz().draw('results/ges_adult.pdf', prog='dot')
-
-    with open('results/llm_edge_list.txt', 'r') as f:
-        edge_list = literal_eval(f.readline())
-    bn = BayesianNetwork(edge_list)
-    bn.to_graphviz().draw('results/expert_llm_adult.pdf', prog='dot')
-
 if __name__ == "__main__":
     dag, total_unexplained_effect, total_ll = build_ges()
     with open('results/ges_edge_list.txt', 'w') as f:
@@ -73,6 +62,4 @@ if __name__ == "__main__":
         f.write(str(np.array(total_unexplained_effect).tolist())[1:-1])
         f.write('\n')
         f.write(str(np.array(total_ll).tolist())[1:-1])
-
-    make_plot()
 

### 2024-11-28 code/fig4/make_plots.R
Also make plots showing the change in log-likelihood over iterations

@@ -18,13 +18,22 @@ d_ges$algo <- 'GES'
 
 d_results = rbind(d_expert, d_ges)
 
-p <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
+
+p_unexplained_effect <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
 	geom_line(alpha=0.6) +
 	geom_point(alpha=0.6, show.legend=F) +
 	theme_minimal(base_size=8) +
 	labs(x="Iteration No.") +
-	labs(y="Unexplained Effect")
+	labs(y="Total Unexplained Effect")
 
-ggsave('unexplained.pdf', p, height=2.5, width=3.2, units='in')
+p_ll <- ggplot(d_results, aes(x=x, y=ll, color=algo)) +
+	geom_line(alpha=0.6) +
+	geom_point(alpha=0.6, show.legend=F) +
+	theme_minimal(base_size=8) +
+	labs(x="Iteration No.") +
+	labs(y="Log-Likelihood")
+
+ggsave('unexplained_effect.pdf', p_unexplained_effect, height=2.5, width=3.2, units='in')
+ggsave('ll.pdf', p_ll, height=2.5, width=3.2, units='in')
 
 

### 2024-11-28 code/fig4/figure4.py
Updates

@@ -35,7 +35,7 @@ def build_network():
             pval_threshold=0.05,
             effect_size_threshold=0.05,
             use_llm=True,
-            llm_model="gemini/gemini-1.5-pro",
+            llm_model="gemini/gemini-1.5-flash",
             variable_descriptions=descriptions,
             show_progress=True,
     )
@@ -43,7 +43,7 @@ def build_network():
 
 def build_ges():
     est = GES(get_adult_df())
-    dag, total_unexplained_effect, total_ll = est.estimate(scoring_method='bic-cg')
+    dag, total_unexplained_effect, total_ll = est.estimate(scoring_method='ll-cg', debug=True)
     return(dag, total_unexplained_effect, total_ll)
 
 if __name__ == "__main__":
@@ -54,6 +54,7 @@ if __name__ == "__main__":
         f.write(str(np.array(total_unexplained_effect).tolist())[1:-1])
         f.write('\n')
         f.write(str(np.array(total_ll).tolist())[1:-1])
+        f.write('\n')
 
     dag, total_unexplained_effect, total_ll = build_network()
     with open('results/llm_edge_list.txt', 'w') as f:

### 2024-11-28 paper/main.tex
Minor paper updates
Fix plots in paper
Updates
Updates paper

--- 2024-11-26 paper/main.tex
@@
 observed phenomena and help in determining effective interventions or policy
-decisions. \emph{Causal Discovery} methods aim to uncover these relationships
+decisions. \emph{Causal Discovery} methods aim to find these relationships
 among random variables from observational data. Methods for causal discovery
@@
 and continuous optimization-based methods like NO TEARS \citep{Zheng2018} and
-DAGMA \citep{Bello2022}. In contrast, the SEM literature has focused more on
-developing tools to assist researchers in manually constructing or modifying
-models while allowing them to integrate their expert knowledge into the
-model-building process. Researchers typically develop an initial model based on
-their expertise and then use various tools to guide modifications to improve
-the model's fit to data. This model construction process is commonly known as
-\emph{Specification Search} \citep{Long1983} and uses method like modification
-indices, and Wald-based tests \citep{Marcoulides2018}.
+DAGMA \citep{Bello2022}. In contrast, the SEM literature has focused more
+towards developing tools to assist researchers in manually constructing or
+modifying models. This approach allows experts to easily integrate their expert
+knowledge into the model-building process. Researchers typically develop an
+initial model based on their expertise and then use various tools to guide
+modifications to improve the model's fit to data. This model construction
+process is commonly known as \emph{Specification Search} \citep{Long1983} and
+uses method like modification indices, and Wald-based tests
+\citep{Marcoulides2018}.
 
@@
 		asymptotically consistent, their finite sample properties are
-		not well understood. Moreover, their output can vary
-		significantly depending on the chosen algorithm and their
-		hyperparameters. Combined with the lack of robust methods to
-		test these algorithms' performance on a given dataset, it can
-		be difficult to trust their outputs. Consequently, researchers
-		usually have to rely on post-hoc modifications to the model
-		based on model testing and their domain expertise.
+		not well understood \todo{Is there any citation for this}.
+		Moreover, their output can vary significantly depending on the
+		chosen algorithm and their hyperparameters. Combined with the
+		lack of robust methods to test these algorithms' performance on
+		a given dataset, it can be difficult to trust their outputs.
+		Consequently, researchers usually have to rely on post-hoc
+		modifications to the model based on model testing and their
+		domain expertise.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
@@
 
-Figure~\ref{fig:intro} shows an example of application of a causal discovery
-algorithms namely PC on the Adult Income dataset. \todo[inline]{Explain}.
+Figure~\ref{fig:intro} shows an example to highlight some of these issues using 
+Adult income dataset.
 
-Due to these challenges, whether constructing the model manually or using an
-algorithm, we typically have to rely on some expert knowledge to get a model
-that we can use for further analysis. Hence, it is crucial to have tools to
-guide this construction or modification process.
+% Due to these challenges, whether constructing the model manually or using an
+% algorithm, we typically have to rely on some expert knowledge to get a model
+% that we can use for further analysis. Hence, it is crucial to have tools to
+% guide this construction or modification process.
 
-One of the ways to guide this modification is to use Conditional Independence
-(CI) based model testing \cite{Ankan2023}. As DAGs imply CI statements that can
-be tested in the given data using statistical CI tests, we can verify whether
-the implied CIs are consistent with the data or not. If some of implied CIs are
-not valid in the given dataset, we can modify the model accordingly.
-Figure~\ref{fig:ci_testing} shows an example of model testing using CIs.
-However, using p-values of the CI tests has a few drawbacks. Firstly, as CIs
-are implied only by missing edges in the model, we can not test whether an
-incorrect edge is already present in the model. For examples, in high sample
-size scenarios, we might get a significant association between a pair of
-variables with very small association. Secondly, the number of implied CIs can
-be quite large. For example, in the moderately sized alarm network with $ 37 $
-node, there are $287$ implied CIs. If we rely only on the p-values of the CI
-tests, it can be difficult to figure which test to focus on.
+One of the common ways to guide this post-hoc modification of DAGs is to use
+Conditional Independence (CI) based model testing \cite{Ankan2023}. As DAGs
+imply CI statements that can be tested in the given data using statistical CI
+tests, we can verify whether the implied CIs are consistent with the data or
+not. If some of implied CIs are not valid in the given dataset, we can modify
+the model accordingly. Figure~\ref{fig:ci_testing} shows an example of model
+testing using CIs. However, this CI based testing has a few drawbacks: 1) CIs
+are implied only by missing edges in the model, so we can not test for already
+existing edges. 2) In high sample size scenarios, p-values can suggest a
+significant association between variables even if they have a very weak
+relationship. 3) Number of implied CIs can be large making it difficult to
+manually go through each of them; for example,  in the moderately sized alarm
+network with $ 37 $ node, there are $287$ implied CIs.
 
-A common method for specification search is using \emph{modification indices},
-which identify potential changes to the model structure that improve the model
-fit. These potential changes are then ranked by the ones that improve the fit
-the most. This ranking allows the researchers to focus first on the suggestions
-that can improve the model the most and while integrating their expertise to
-choose the most optimal modifications.
+On the other hand, a common method for specification search is using
+\emph{modification indices}, which identify potential changes to the model
+structure that improve the model fit. These potential changes are then ranked
+by the ones that improve the fit the most. This ranking allows researchers to
+focus first on the suggestions that can improve the model the most and while
+integrating their expertise to choose the most optimal modifications.
 
 Inspired by this ranking approach in modification indices, in this paper we
-develop a method to rank the CI tests using a conditional measure of
-association.  To deal with both this issues, in this paper, we develop a
-measure of partial association which can be using in combination of the CI test
-to rank the violations in these CI tests. On the other hand, in Structural
-Equation Models (SEMs) literature, more emphasis has been placed on developing
-tools to help researchers manually construct models while integrating their
-domain expertise. Our proposed approach uses the measure of association to find
-correlated variables in the data that are not explained by the current model.
-Using the measure, we can then rank the pair of variables between which adding
-an edge would improve the explanation the most. Based on this information, the
-user can select a potential pair of variables and based on their expert
-knowledge decide the direction of the edge.
+develop a method to rank the CI tests. We utilize a measure of conditional
+association to do this ranking. Additionally, this measure of conditional
+association also allows us tackle other issues with CI based testing outlined
+in the previous paragraphs.
+
+% using a conditional measure of
+% association.  To deal with both this issues, in this paper, we develop a
+% measure of partial association which can be using in combination of the CI test
+% to rank the violations in these CI tests. On the other hand, in Structural
+% Equation Models (SEMs) literature, more emphasis has been placed on developing
+% tools to help researchers manually construct models while integrating their
+% domain expertise. Our proposed approach uses the measure of association to find
+% correlated variables in the data that are not explained by the current model.
+% Using the measure, we can then rank the pair of variables between which adding
+% an edge would improve the explanation the most. Based on this information, the
+% user can select a potential pair of variables and based on their expert
+% knowledge decide the direction of the edge.
 
@@
 		\end{tikzpicture}
-		\caption{\todo[inline]{Finish figure. Show an example of model learnt using PC algorithm $\rightarrow$ Orienting the unoriented edges $\rightarrow$  Model testing}}
+		\caption{\todo[inline]{Generate this figure using a few different CI tests and also show the unoriented edges in some of them}}
 		\label{fig:intro}
@@
 		or deletion of edges from the model.
-	\item Lastly, we present both a web-browser based tool and an
-		implementation in the Python package to allow researchers to
-		apply this method to their own datasets.
+	\item We present both a web-browser based tool and an implementation in
+		the Python package to allow researchers to apply this method to
+		their own datasets.
 \end{enumerate}
@@
 	\centering
-	\includegraphics[scale=0.4]{plots/web_tool_full.png}
+	% \includegraphics[scale=0.4]{plots/web_tool_full.png}
 	\caption{A screenshot of the web tool for constructing the model. Users
@@
 	multiple accuracy values of the oracle. \todo[inline]{Replace Hill-Climb with GES and improve the plot}}
+\end{figure}
+
+\begin{figure}
+	\begin{subfigure}{0.5\textwidth}
+		\centering
+		\includegraphics{../code/fig4/unexplained_effect.pdf}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}{0.5\textwidth}
+		\centering
+		\includegraphics{../code/fig4/ll.pdf}
+		\caption{}
+	\end{subfigure}
+	\caption{Plots showing improvements in fit of the model over iterations
+	on the Adult income dataset. For the expert simulation an LLM is used
+	for the edge orientations. (a) Shows using total unexplained effect that is the
+	sum of effect size between variables that don't have an edge between them. (b)
+	Shows the log-likelihood of the overall model.
+	}
 \end{figure}
@@
 		\centering
-		\includegraphics[scale=0.3]{../code/plots/llm_adult.png}
+		\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
 		\caption{\todo[inline]{Change figure to have consistent style}}

### 2024-11-29 code/fig1/adult_sl.R
Additional ways to learn the adult income dataset

@@ -356,9 +356,9 @@ var_types <- list(
 		NativeCountry = "cat"
 		)
 
-suffStat <- list(dm = d, adaptDF = F, test = "rf_q3", var_types = var_types)
-pc.skel <- pcalg::pc(
-		suffStat,
+suffStat_q3 <- list(dm = d, adaptDF = F, test = "rf_q3", var_types = var_types)
+pc.cpdag.q3 <- pcalg::pc(
+		suffStat_q3,
 		indepTest = sl.test,
 		alpha = 0.05,
 		labels = colnames(d),
@@ -368,4 +368,33 @@ pc.skel <- pcalg::pc(
 		verbose = F
 		)
 
-adj_mat <- pcalg::dag2cpdag(as(pc.skel, "amat")) # Use this to get the final model.
+suffStat_mxm <- list(dm = d, adaptDF = F, test = "mxm", var_types = var_types)
+pc.cpdag.mxm <- pcalg::pc(
+		suffStat_mxm,
+		indepTest = sl.test,
+		alpha = 0.05,
+		labels = colnames(d),
+		u2pd = "relaxed",
+		skel.method = "stable.fast",
+		numCores = 8,
+		verbose = F
+		)
+
+suffStat_mi <- list(dm = d, adaptDF = F, test = "mi_cg", var_types = var_types)
+pc.cpdag.mi <- pcalg::pc(
+		suffStat_mi,
+		indepTest = sl.test,
+		alpha = 0.05,
+		labels = colnames(d),
+		u2pd = "relaxed",
+		skel.method = "stable.fast",
+		numCores = 8,
+		verbose = F
+		)
+
+hc.cpdag <- bnlearn::cpdag(bnlearn::hc(d, score='bic-cg'))
+
+adj_mat_q3 <- pcalg::dag2cpdag(as(pc.skel.q3, "amat")) # Use this to get the final model.
+adj_mat_mxm <- pcalg::dag2cpdag(as(pc.skel.mxm, "amat")) # Use this to get the final model.
+adj_mat_mi <- pcalg::dag2cpdag(as(pc.skel.mi, "amat")) # Use this to get the final model.
+

### 2024-11-29 paper/main.tex
Updates paper
Adds another plot

@@
 not. If some of implied CIs are not valid in the given dataset, we can modify
-the model accordingly. Figure~\ref{fig:ci_testing} shows an example of model
-testing using CIs. However, this CI based testing has a few drawbacks: 1) CIs
+the model accordingly. However, this CI based testing has a few drawbacks: 1) CIs
 are implied only by missing edges in the model, so we can not test for already
@@
 \centering
+\begin{subfigure}{0.5 \textwidth}
 	\begin{tikzpicture}
-		\begin{scope}[xshift=-2.5cm, yshift=0.7cm, scale=0.7]
+	\begin{scope}[xshift=-2.5cm, yshift=0.7cm]
 		\tikzstyle{every node}=[align=center, inner sep=1pt]
@@
 	\end{scope}
-		\begin{scope}[xshift=-3cm]
-			\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
-			\node[rectangle, align=center, inner sep=1pt] at (5, 0) {
-				\begin{minipage}{.2\textwidth}
-					\begin{equation*}
-						\begin{split}
-							\textnormal{Age} &\ci \textnormal{Incm} \rvert \textnormal{Edct, Hrpw, Sex} \\
-							\textnormal{Age} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
-							\textnormal{Age} &\ci \textnormal{Sex} \\
-							\textnormal{Hrpw} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
-						\end{split}
-					\end{equation*}
-				\end{minipage}
-				};
+	\end{tikzpicture}
+\end{subfigure}
+\begin{subfigure}{0.5 \textwidth}
+	\begin{tikzpicture}
+	\begin{scope}[xshift=-2.5cm, yshift=0.7cm]
+		\tikzstyle{every node}=[align=center, inner sep=1pt]
+		\node (wclass) at (0, 1) {WorkClass};
+		\node (sex) at (0, -0.8) {Sex};
+		\node (age) at (-0.3, 0) {Age};
+		\node (ed) at (-2, -1.6) {Education};
+		\node (occ) at (1.8, 0) {Occupation};
+		\node (hrpw) at (1, -2) {Hours Per \\ Week};
+		\node (income) at (-0.7, -2.3) {Income};
+		\node (marital) at (3, 1) {Marital \\ Status};
+		\node (reln) at (-3, 1) {Relationship};
+		\node (race) at (-4, 2) {Race};
+		\node (country) at (-4, 3) {Native \\ Country};
+
+		\draw[] (ed) to (income);
+		\draw[] (marital) to (reln);
+		\draw[] (occ) to (sex);
+		\draw[->] (reln) to (age);
+		\draw[->] (reln) to (marital);
+		\draw[] (reln) to (sex);
+		\draw[->] (reln) to (hrpw);
+		\draw[] (reln) to (income);
+		\draw[->] (income) to (ed);
 	\end{scope}
 	\end{tikzpicture}
-		\caption{\todo[inline]{Generate this figure using a few different CI tests and also show the unoriented edges in some of them}}
+\end{subfigure}
+\begin{subfigure}{0.5\textwidth}
+	\begin{tikzpicture}
+	\begin{scope}[xshift=-2.5cm, yshift=0.7cm]
+		\tikzstyle{every node}=[align=center, inner sep=1pt]
+		\node (wclass) at (0, 1) {WorkClass};
+		\node (sex) at (0, -0.8) {Sex};
+		\node (age) at (-0.3, 0) {Age};
+		\node (ed) at (-2, -1.6) {Education};
+		\node (occ) at (1.8, 0) {Occupation};
+		\node (hrpw) at (1, -2) {Hours Per \\ Week};
+		\node (income) at (-0.7, -2.3) {Income};
+		\node (marital) at (3, 1) {Marital \\ Status};
+		\node (reln) at (-3, 1) {Relationship};
+		\node (race) at (-4, 2) {Race};
+		\node (country) at (-4, 3) {Native \\ Country};
+
+		\draw[]   (wclass) to (ed);
+		\draw[]   (wclass) to (occ);
+		\draw[]   (wclass) to (race);
+		\draw[]   (wclass) to (income);
+		\draw[]   (ed) to (occ);
+		\draw[]   (ed) to (country);
+		\draw[]   (ed) to (income);
+		\draw[]   (marital) to (occ);
+		\draw[]   (marital) to (reln);
+		\draw[->] (marital) to (sex);
+		\draw[]   (marital) to (hrpw);
+		\draw[->] (occ) to (sex);
+		\draw[]   (occ) to (hrpw);
+		\draw[->] (reln) to (sex);
+		\draw[]   (reln) to (hrpw);
+		\draw[->] (income) to (occ);
+
+	\end{scope}
+	\end{tikzpicture}
+\end{subfigure}
+
+\caption{Markov Equivalence Class (MEC) learnt for Adult Income Dataset using (a) PC algorithm, (b) Hill-Climb Search with BIC score, (c) }
 \label{fig:intro}
 \end{figure}
+
+			% \begin{scope}[xshift=-3cm]
+			% 	\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
+			% 	\node[rectangle, align=center, inner sep=1pt] at (5, 0) {
+			% 		\begin{minipage}{.2\textwidth}
+			% 			\begin{equation*}
+			% 				\begin{split}
+			% 					\textnormal{Age} &\ci \textnormal{Incm} \rvert \textnormal{Edct, Hrpw, Sex} \\
+			% 					\textnormal{Age} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
+			% 					\textnormal{Age} &\ci \textnormal{Sex} \\
+			% 					\textnormal{Hrpw} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
+			% 				\end{split}
+			% 			\end{equation*}
+			% 		\end{minipage}
+			% 		};
+			% \end{scope}
 
@@
 		Greedy Equivalence Search (GES).
+	\item But not decomposible like scoring methods. 
 	\item But Because they make only local changes, can get stuck in local minimas.
-	\item This approach looks for modifications globally.
+	\item Both approach looks for modifications globally.
+	\item The sum of unexplained effect provides an absolute measure of model fit 
+		unlike scoring methods.
 \end{itemize}
-
-\todo[inline]{Show an example where local changes can get stuck}
 
@@
 		the range of values. The human drawn values are done for
-	multiple accuracy values of the oracle. \todo[inline]{Replace Hill-Climb with GES and improve the plot}}
+		multiple accuracy values of the oracle.}
 \end{figure}

### 2024-11-29 paper/main.tex
Updates figure 1

@@
 
-\caption{Markov Equivalence Class (MEC) learnt for Adult Income Dataset using (a) PC algorithm, (b) Hill-Climb Search with BIC score, (c) }
+\caption{Markov Equivalence Class (MEC) learnt for Adult Income Dataset using (a) PC algorithm, (b) Hill-Climb Search with BIC score, (c) PC algorithm with MXM test}
 \label{fig:intro}

### 2025-01-14 paper/main.tex
Update format for UAI

@@
-%File: anonymous-submission-latex-2024.tex
-\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
-\usepackage[submission]{aaai24}  % DO NOT CHANGE THIS
-\usepackage{times}  % DO NOT CHANGE THIS
-\usepackage{helvet}  % DO NOT CHANGE THIS
-\usepackage{courier}  % DO NOT CHANGE THIS
-\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
-\usepackage{graphicx} % DO NOT CHANGE THIS
-\urlstyle{rm} % DO NOT CHANGE THIS
-\def\UrlFont{\rm}  % DO NOT CHANGE THIS
-\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
-\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
-\frenchspacing  % DO NOT CHANGE THIS
-\setlength{\pdfpagewidth}{8.5in} % DO NOT CHANGE THIS
-\setlength{\pdfpageheight}{11in} % DO NOT CHANGE THIS
-%
-% These are recommended to typeset algorithms but not required. See the subsubsection on algorithms. Remove them if you don't have algorithms in your paper.
-\usepackage{algorithm}
-\usepackage{algorithmic}
+\documentclass{uai2025} % for initial submission
+%\documentclass[accepted]{uai2025} % after acceptance, for a revised version; 
+% also before submission to see how the non-anonymous paper would look like
 
-%
-% These are are recommended to typeset listings but not required. See the subsubsection on listing. Remove this block if you don't have listings in your paper.
-\usepackage{newfloat}
-\usepackage{listings}
-\DeclareCaptionStyle{ruled}{labelfont=normalfont,labelsep=colon,strut=off} % DO NOT CHANGE THIS
-\lstset{%
-	basicstyle={\footnotesize\ttfamily},% footnotesize acceptable for monospace
-	numbers=left,numberstyle=\footnotesize,xleftmargin=2em,% show line numbers, remove this entire line if you don't want the numbers.
-	aboveskip=0pt,belowskip=0pt,%
-	showstringspaces=false,tabsize=2,breaklines=true}
-\floatstyle{ruled}
-\newfloat{listing}{tb}{lst}{}
-\floatname{listing}{Listing}
-%
-% Keep the \pdfinfo as shown here. There's no need
-% for you to add the /Title and /Author tags.
-\pdfinfo{
-/TemplateVersion (2024.1)
-}
+%% There is a class option to choose the math font
+% \documentclass[mathfont=ptmx]{uai2025} % ptmx math instead of Computer
+                                         % Modern (has noticeable issues)
+% \documentclass[mathfont=newtx]{uai2025} % newtx fonts (improves upon
+                                          % ptmx; less tested, no support)
+% NOTE: Only keep *one* line above as appropriate, as it will be replaced
+%       automatically for papers to be published. Do not make any other
+%       change above this note for an accepted version.
 
-% DISALLOWED PACKAGES
-% \usepackage{authblk} -- This package is specifically forbidden
-% \usepackage{balance} -- This package is specifically forbidden
-% \usepackage{color (if used in text)
-% \usepackage{CJK} -- This package is specifically forbidden
-% \usepackage{float} -- This package is specifically forbidden
-% \usepackage{flushend} -- This package is specifically forbidden
-% \usepackage{fontenc} -- This package is specifically forbidden
-% \usepackage{fullpage} -- This package is specifically forbidden
-% \usepackage{geometry} -- This package is specifically forbidden
-% \usepackage{grffile} -- This package is specifically forbidden
-% \usepackage{hyperref} -- This package is specifically forbidden
-% \usepackage{navigator} -- This package is specifically forbidden
-% (or any other package that embeds links such as navigator or hyperref)
-% \indentfirst} -- This package is specifically forbidden
-% \layout} -- This package is specifically forbidden
-% \multicol} -- This package is specifically forbidden
-% \nameref} -- This package is specifically forbidden
-% \usepackage{savetrees} -- This package is specifically forbidden
-% \usepackage{setspace} -- This package is specifically forbidden
-% \usepackage{stfloats} -- This package is specifically forbidden
-% \usepackage{tabu} -- This package is specifically forbidden
-% \usepackage{titlesec} -- This package is specifically forbidden
-% \usepackage{tocbibind} -- This package is specifically forbidden
-% \usepackage{ulem} -- This package is specifically forbidden
-% \usepackage{wrapfig} -- This package is specifically forbidden
-% DISALLOWED COMMANDS
-% \nocopyright -- Your paper will not be published if you use this command
-% \addtolength -- This command may not be used
-% \balance -- This command may not be used
-% \baselinestretch -- Your paper will not be published if you use this command
-% \clearpage -- No page breaks of any kind may be used for the final version of your paper
-% \columnsep -- This command may not be used
-% \newpage -- No page breaks of any kind may be used for the final version of your paper
-% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
-% \pagestyle -- This command may not be used
-% \tiny -- This is not an acceptable font size.
-% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
-% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference
+%% Choose your variant of English; be consistent
+\usepackage[american]{babel}
+% \usepackage[british]{babel}
 
-\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.
+%% Some suggested packages, as needed:
+\usepackage{natbib} % has a nice set of citation styles and commands
+    \bibliographystyle{plainnat}
+    \renewcommand{\bibsection}{\subsubsection*{References}}
+\usepackage{mathtools} % amsmath with fixes and additions
+% \usepackage{siunitx} % for proper typesetting of numbers and units
+\usepackage{booktabs} % commands to create good-looking tables
+\usepackage{tikz} % nice language for creating drawings and diagrams
 
@@
 
-% The file aaai24.sty is the style file for AAAI Press
-% proceedings, working notes, and technical reports.
+\title{Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge}
+
+% The standard author block has changed for UAI 2025 to provide
+% more space for long author lists and allow for complex affiliations
 %
-
-% Title
-
-% Your title must be in mixed case, not sentence case.
-% That means all verbs (including short verbs like be, is, using,and go),
-% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
-% articles, conjunctions, and prepositions are lower case unless they
-% directly follow a colon or long dash
-\title{Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge}
-\author{
-    %Authors
-    % All authors must be in the same font size and format.
-    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
-    AAAI Style Contributions by Pater Patel Schneider,
-    Sunil Issar,\\
-    J. Scott Penberthy,
-    George Ferguson,
-    Hans Guesgen,
-    Francisco Cruz\equalcontrib,
-    Marc Pujol-Gonzalez\equalcontrib
+% All author information is authomatically removed by the class for the
+% anonymous submission version of your paper, so you can already add your
+% information below.
+%
+% Add authors
+\author[1]{\href{mailto:<jj@example.edu>?Subject=Your UAI 2025 paper}{Jane~J.~von~O'L\'opez}{}}
+\author[1]{Harry~Q.~Bovik}
+\author[1,2]{Further~Coauthor}
+\author[3]{Further~Coauthor}
+\author[1]{Further~Coauthor}
+\author[3]{Further~Coauthor}
+\author[3,1]{Further~Coauthor}
+% Add affiliations after the authors
+\affil[1]{%
+    Computer Science Dept.\\
+    Cranberry University\\
+    Pittsburgh, Pennsylvania, USA
 }
-\affiliations{
-    %Afiliations
-    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
-    % If you have multiple authors and multiple affiliations
-    % use superscripts in text and roman font to identify them.
-    % For example,
-
-    % Sunil Issar\textsuperscript{\rm 2},
-    % J. Scott Penberthy\textsuperscript{\rm 3},
-    % George Ferguson\textsuperscript{\rm 4},
-    % Hans Guesgen\textsuperscript{\rm 5}
-    % Note that the comma should be placed after the superscript
-
-    1900 Embarcadero Road, Suite 101\\
-    Palo Alto, California 94303-3310 USA\\
-    % email address must be in roman text type, not monospace or sans serif
-    proceedings-questions@aaai.org
-%
-% See more examples next
+\affil[2]{%
+    Second Affiliation\\
+    Address\\
+    …
 }
-
-%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
-\iffalse
-\title{My Publication Title --- Single Author}
-\author {
-    Author Name
+\affil[3]{%
+    Another Affiliation\\
+    Address\\
+    …
   }
-\affiliations{
-    Affiliation\\
-    Affiliation Line 2\\
-    name@example.com
-}
-\fi
-
-\iffalse
-%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
-\title{My Publication Title --- Multiple Authors}
-\author {
-    % Authors
-    First Author Name\textsuperscript{\rm 1},
-    Second Author Name\textsuperscript{\rm 2},
-    Third Author Name\textsuperscript{\rm 1}
-}
-\affiliations {
-    % Affiliations
-    \textsuperscript{\rm 1}Affiliation 1\\
-    \textsuperscript{\rm 2}Affiliation 2\\
-    firstAuthor@affiliation1.com, secondAuthor@affilation2.com, thirdAuthor@affiliation1.com
-}
-\fi
 

### 2025-01-14 paper/main.tex
Updates

@@
 % Add authors
-\author[1]{\href{mailto:<jj@example.edu>?Subject=Your UAI 2025 paper}{Jane~J.~von~O'L\'opez}{}}
-\author[1]{Harry~Q.~Bovik}
-\author[1,2]{Further~Coauthor}
-\author[3]{Further~Coauthor}
-\author[1]{Further~Coauthor}
-\author[3]{Further~Coauthor}
-\author[3,1]{Further~Coauthor}
+\author[1]{\href{mailto:<ankur.ankan@ru.nl>?Subject=Your UAI 2025 paper}{Ankur~Ankan}{}}
+\author[1]{Johannes~Textor}
+
 % Add affiliations after the authors
 \affil[1]{%
-    Computer Science Dept.\\
-    Cranberry University\\
-    Pittsburgh, Pennsylvania, USA
+    Institute for Computing and Information Sciences\\
+    Radboud University\\
+    Nijmegen, The Netherlands
 }
-\affil[2]{%
-    Second Affiliation\\
-    Address\\
-    …
-}
-\affil[3]{%
-    Another Affiliation\\
-    Address\\
-    …
-  }
-
 \begin{document}
@@
 \begin{abstract}
+% Causal discovery has received significant attention in the Directed Acyclic
+% Graphs (DAGs) literature, leading to the development of numerous automated
+% algorithms for learning DAGs from data. Despite these advancements, the
+% application of these algorithms in applied domains remains limited, with
+% researchers often preferring to manually construct DAGs based on their expert
+% knowledge. This preference stems due to some practical issues with these
+% algorithms such as making obvious mistakes, and outputting Markov Equivalence
+% Class (MEC). As a result of these issues users typically have to rely on post
+% hoc modifications to the model based on their expertise. Given this reliance on
+% manual model modification and construction, we propose an iterative method to
+% assist researchers in manually constructing DAGs by combining implied
+% Conditional Independence (CI) based model testing and utilizing their expert
+% knowledge. We present a novel measure of partial association for mixed data
+% that allows us to rank violations of implied CIs in the model. Using this
+% ranking users can prioritize modifications to the model while using their
+% expert knowledge to decide orientation of edges. Empirically, we show that
+% a greedy approach achieves performance comparable to automated algorithms, provided
+% the user can correctly determine the edge orientation in $ 1 $ out of $ 3 $ cases.
+% Additionally, we show that in the absence of an expert, a Large Language Model can
+% potentially be used for determining edge orientations.
+
 Causal discovery has received significant attention in the Directed Acyclic
 Graphs (DAGs) literature, leading to the development of numerous automated
-algorithms for learning DAGs from data. Despite these advancements, the
-application of these algorithms in applied domains remains limited, with
-researchers often preferring to manually construct DAGs based on their expert
-knowledge. This preference stems due to some practical issues with the
-automated algorithms such as they can make obvious mistakes, and they output
-the Markov Equivalence Class (MEC). Due to these issues, when using these
-algorithms, users typically have to rely on expert knowledge to fix the
-mistakes and to orient the MEC to a DAG. Given this reliance on manual model
-construction, we propose an iterative method to assist researchers in manually
-constructing DAGs by combining implied Conditional Independence (CI) based
-model testing and utilizing their expert knowledge. Our approach is inspired by
-specification search in Structural Equation Models (SEMs), where we utilize a
-novel measure of partial association for mixed data to rank violations of
-implied Conditional Independences in the model. Using this ranking users can
-prioritize the pair of variables which need to be corrected by either adding an
-edge between them or removing the edge based on their expertise to specify the
-correct orientation. Empirically, we demonstrate that a greedy approach (i.e.,
-fixing the worst violation first) achieves performance comparable to automated
-algorithms, provided the user can correctly determine the edge orientation in $
-1 $ out of $ 3 $ edge additions. Additionally, we show that in the absence of
-an expert, a Large Language Model (LLM) can potentially be used to determine
-edge orientations. 
+algorithms for learning DAGs from data. However, their adoption in applied
+domains remains limited, as researchers often prefer to construct DAGs manually
+based on their expert knowledge. This preference arises due to practical issues
+with these automated algorithms, such as making obvious errors and outputting
+Markov Equivalence Classes (MECs). To address these challenge, we propose an
+iterative method to assist researchers in manually constructing or modifying
+DAGs by combining implied Conditional Independence (CI) testing with expert
+knowledge. We introduce a novel measure of partial association for mixed data
+to rank violations of implied CIs, enabling users to prioritize model
+modifications while leveraging their expertise to decide the orientation of
+edges when making these modifications. Empirical results show that this
+approach achieves performance comparable to automated algorithms, assuming the
+user can correctly orient edges in one out of three cases. Additionally, we
+demonstrate that, in the absence of an expert, a Large Language Model can
+potentially determine edge orientations.
+
 \end{abstract}
@@
 Petersen2021}. We believe this preference persists due to several issues with
-these algorithms when applied in practical settings:
+these algorithms when applied in practical settings with finite sample size:
 
@@
 		a given dataset, it can be difficult to trust their outputs.
-		Consequently, researchers usually have to rely on post-hoc
-		modifications to the model based on model testing and their
-		domain expertise.
+		Consequently, even when using these algorithms, users have to
+		rely on post hoc manual modifications to the model based on model
+		testing and their domain expertise.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
@@
 		a fully oriented DAG. As a result, researchers have to manually
-		orient these edges before using them for analysis.
+		orient these edges based on their expert knowledge before using
+		them for analysis.
 \end{enumerate}
@@
 
-One of the common ways to guide this post-hoc modification of DAGs is to use
-Conditional Independence (CI) based model testing \cite{Ankan2023}. As DAGs
-imply CI statements that can be tested in the given data using statistical CI
-tests, we can verify whether the implied CIs are consistent with the data or
-not. If some of implied CIs are not valid in the given dataset, we can modify
-the model accordingly. However, this CI based testing has a few drawbacks: 1) CIs
-are implied only by missing edges in the model, so we can not test for already
+As a result of these issues, whether constructing the model manually or using
+an algorithm, the user has to rely on their expertise to make modifications to
+the model. One of the common ways to guide this post hoc modification is to use
+implied CI based model testing \citep{Ankan2021, Ankan2023}. In this testing
+procedure, users test whether the implied CIs of the DAG hold in the dataset or
+not using statistical CI tests. Based on the results of these statistical
+tests, we can accordingly make modifications to the current model to fix the
+violations. However, this CI based testing also has a few drawbacks: 1) CIs are
+implied only by missing edges in the model, so we can not test for already
 existing edges. 2) In high sample size scenarios, p-values can suggest a
@@
 
-On the other hand, a common method for specification search is using
-\emph{modification indices}, which identify potential changes to the model
-structure that improve the model fit. These potential changes are then ranked
-by the ones that improve the fit the most. This ranking allows researchers to
-focus first on the suggestions that can improve the model the most and while
-integrating their expertise to choose the most optimal modifications.
-
-Inspired by this ranking approach in modification indices, in this paper we
-develop a method to rank the CI tests. We utilize a measure of conditional
-association to do this ranking. Additionally, this measure of conditional
-association also allows us tackle other issues with CI based testing outlined
-in the previous paragraphs.
+To tackle these issues with CI based testing, we take inspiration from
+\emph{modification indices} in specification search. The modification indices
+method tries to identify and rank changes in the model such that modifications
+that improve the model fit the most are ranked higher.This ranking allows
+researchers to focus first on the suggestions that can improve the model the
+most and while integrating their expertise to choose the most optimal
+modifications. Analogous to this, in this paper we develop a measure for
+partial association for mixed data that allows us to rank the CI violations.
+Additionally, this measure of conditional association also allows us tackle
+other issues with CI based testing outlined in the previous paragraphs.
 
@@
 \begin{figure}
-\centering
 \begin{subfigure}{0.5 \textwidth}
-	\begin{tikzpicture}
-	\begin{scope}[xshift=-2.5cm, yshift=0.7cm]
-		\tikzstyle{every node}=[align=center, inner sep=1pt]
-		\node (sex) at (-0.7, -0.8) {Sex};
-		\node (age) at (-0.3, 0) {Age};
-		\node (ed) at (-1.8, -1.6) {Education};
-		\node (occ) at (1.8, 0) {Occupation};
-		\node (hrpw) at (1.3, -1.6) {Hours Per \\ Week};
-		\node (income) at (-0.7, -2.8) {Income};
-
-		\draw[->]  (age) to[bend right=20] (ed);
-		\draw[->]  (sex) to (ed);
-		\draw[->]  (age) to (hrpw);
-		\draw[->]  (ed) to (hrpw);
-		\draw[->]  (sex) to (hrpw);
-		\draw[->]  (ed) to (income);
-		\draw[->]  (hrpw) to (income);
-		\draw[->]  (occ) to[out=300, in=0, looseness=1.4] (income.east);
-		\draw[->]  (sex) to (income);
-		\draw[->]  (ed) to[out=90, in=150, looseness=1.3] (occ);
-		\draw[->]  (sex) to (occ);	
-	\end{scope}
-	\end{tikzpicture}
+	\includegraphics[page=1]{figures.pdf}
 \end{subfigure}
 \begin{subfigure}{0.5 \textwidth}
-	\begin{tikzpicture}
-	\begin{scope}[xshift=-2.5cm, yshift=0.7cm]
-		\tikzstyle{every node}=[align=center, inner sep=1pt]
-		\node (wclass) at (0, 1) {WorkClass};
-		\node (sex) at (0, -0.8) {Sex};
-		\node (age) at (-0.3, 0) {Age};
-		\node (ed) at (-2, -1.6) {Education};
-		\node (occ) at (1.8, 0) {Occupation};
-		\node (hrpw) at (1, -2) {Hours Per \\ Week};
-		\node (income) at (-0.7, -2.3) {Income};
-		\node (marital) at (3, 1) {Marital \\ Status};
-		\node (reln) at (-3, 1) {Relationship};
-		\node (race) at (-4, 2) {Race};
-		\node (country) at (-4, 3) {Native \\ Country};
-
-		\draw[] (ed) to (income);
-		\draw[] (marital) to (reln);
-		\draw[] (occ) to (sex);
-		\draw[->] (reln) to (age);
-		\draw[->] (reln) to (marital);
-		\draw[] (reln) to (sex);
-		\draw[->] (reln) to (hrpw);
-		\draw[] (reln) to (income);
-		\draw[->] (income) to (ed);
-	\end{scope}
-	\end{tikzpicture}
+	\includegraphics[page=2]{figures.pdf}
 \end{subfigure}
 \begin{subfigure}{0.5\textwidth}
-	\begin{tikzpicture}
-	\begin{scope}[xshift=-2.5cm, yshift=0.7cm]
-		\tikzstyle{every node}=[align=center, inner sep=1pt]
-		\node (wclass) at (0, 1) {WorkClass};
-		\node (sex) at (0, -0.8) {Sex};
-		\node (age) at (-0.3, 0) {Age};
-		\node (ed) at (-2, -1.6) {Education};
-		\node (occ) at (1.8, 0) {Occupation};
-		\node (hrpw) at (1, -2) {Hours Per \\ Week};
-		\node (income) at (-0.7, -2.3) {Income};
-		\node (marital) at (3, 1) {Marital \\ Status};
-		\node (reln) at (-3, 1) {Relationship};
-		\node (race) at (-4, 2) {Race};
-		\node (country) at (-4, 3) {Native \\ Country};
-
-		\draw[]   (wclass) to (ed);
-		\draw[]   (wclass) to (occ);
-		\draw[]   (wclass) to (race);
-		\draw[]   (wclass) to (income);
-		\draw[]   (ed) to (occ);
-		\draw[]   (ed) to (country);
-		\draw[]   (ed) to (income);
-		\draw[]   (marital) to (occ);
-		\draw[]   (marital) to (reln);
-		\draw[->] (marital) to (sex);
-		\draw[]   (marital) to (hrpw);
-		\draw[->] (occ) to (sex);
-		\draw[]   (occ) to (hrpw);
-		\draw[->] (reln) to (sex);
-		\draw[]   (reln) to (hrpw);
-		\draw[->] (income) to (occ);
-
-	\end{scope}
-	\end{tikzpicture}
+	\includegraphics[page=3]{figures.pdf}
 \end{subfigure}

### 2025-01-14 paper/main.tex
Updates

@@
 \section{Introduction}
+Understanding cause-and-effect relationships between variables is a fundamental
+objective in many scientific fields. These relationships reveal the mechanisms
+behind observed phenomena and guide effective interventions or policy
+decisions. \emph{Causal Discovery} methods aim to identify such relationships
+among random variables using observational data. Approaches to causal discovery
+have been developed within both the Directed Acyclic Graphs (DAGs) and
+Structural Equation Models (SEMs) frameworks, each adapting a different approach.
 
-Understanding cause-and-effect relationships between variables is a fundamental
-goal in many scientific inquiries, as it reveals the underlying mechanisms of
-observed phenomena and help in determining effective interventions or policy
-decisions. \emph{Causal Discovery} methods aim to find these relationships
-among random variables from observational data. Methods for causal discovery
-have been developed in both the Directed Acyclic Graphs (DAGs) and Structural
-Equation Models (SEMs) literatures, each adopting a different approach.
+In the DAG literature, the primary focus has been on developing automated
+algorithms to learn causal structures from datasets. These efforts have led to
+numerous causal discovery algorithms, including constraint-based methods like
+the PC algorithm \citep{Spirtes2001} and Fast Causal Inference (FCI)
+\citep{Spirtes2000}), score-based methods such as Hill-Climb Search and Greedy
+Equivalence Search (GES) \citep{Chickering2002}, and continuous
+optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
+\citep{Bello2022}. In contrast, the SEM literature has focused on developing
+tools to assist researchers in manually constructing or modifying models,
+enabling experts to incorporate their domain knowledge in the model building
+process. Researchers typically begin with an initial model based on their
+expertise and then use these tools to guide modifications that improve the
+model's fit to data. This process is commonly known as \emph{Specification
+Search} \citep{Long1983} and uses method such as modification indices, and
+Wald-based tests \citep{Marcoulides2018}.
 
-In the DAG literature, the focus has primarily been on developing automated
-algorithms to learn causal structures from datasets. As a result of these
-efforts, many algorithms for causal discovery have been proposed, such as
-constraint-based methods like the PC algorithm \citep{Spirtes2001} and Fast
-Causal Inference (FCI) \citep{Spirtes2000}), score-based methods like
-Hill-Climb Search and Greedy Equivalence Search (GES) \citep{Chickering2002},
-and continuous optimization-based methods like NO TEARS \citep{Zheng2018} and
-DAGMA \citep{Bello2022}. In contrast, the SEM literature has focused more
-towards developing tools to assist researchers in manually constructing or
-modifying models. This approach allows experts to easily integrate their expert
-knowledge into the model-building process. Researchers typically develop an
-initial model based on their expertise and then use various tools to guide
-modifications to improve the model's fit to data. This model construction
-process is commonly known as \emph{Specification Search} \citep{Long1983} and
-uses method like modification indices, and Wald-based tests
-\citep{Marcoulides2018}.
-
-Despite the many automated algorithms proposed for constructing DAGs, their
-adoption in applied research remains limited, with researchers often preferring
-to manually construct DAGs based on their domain expertise \citep{Tennant2020,
-Petersen2021}. We believe this preference persists due to several issues with
-these algorithms when applied in practical settings with finite sample size:
+Despite the availability of many automated algorithms for constructing DAGs,
+their adoption in applied research remains limited. Researchers often prefer to
+manually construct DAGs based on their domain expertise \citep{Tennant2020,
+Petersen2021}. We believe this preference persists due to several challenges
+with these algorithms when applied in practical settings:
 
 \begin{enumerate}
-	\item \textbf{Lack of Trust:} Although most algorithms are
+	\item \textbf{Lack of Trust:} While most algorithms are
 		asymptotically consistent, their finite sample properties are
 		not well understood \todo{Is there any citation for this}.
-		Moreover, their output can vary significantly depending on the
-		chosen algorithm and their hyperparameters. Combined with the
+		Additionally, their output can vary significantly depending on
+		the chosen algorithm and their hyperparameters. Combined with a
 		lack of robust methods to test these algorithms' performance on
-		a given dataset, it can be difficult to trust their outputs.
-		Consequently, even when using these algorithms, users have to
-		rely on post hoc manual modifications to the model based on model
-		testing and their domain expertise.
-	\item \textbf{Outputs Markov Equivalence Class (MEC):} Multiple
-		DAGs can be faithful to a given observational dataset. Thus,
-		using only observational data, algorithms can only recover the
-		MECs that can contain a combination of directed and undirected
-		edges. However, most methods for downstream tasks, such as
-		identification or causal effect estimation, assume knowledge of
-		a fully oriented DAG. As a result, researchers have to manually
-		orient these edges based on their expert knowledge before using
-		them for analysis.
+		a given dataset, it can be difficult to trust their outputs. As
+		a result, even when using these algorithms are used,
+		researchers often rely on post hoc manual modifications to the
+		model based on model testing and their domain expertise.
+	\item \textbf{Outputs Markov Equivalence Class (MEC):} As multiple
+		DAGs can be faithful to a given observational dataset, the
+		algorithms can only recover the MECs. These MECs can contain a
+		combination of directed and undirected edges. However, most
+		methods for downstream tasks, such as identification or causal
+		effect estimation, assume knowledge of a fully oriented DAG. As
+		a result, researchers have to manually orient these edges based
+		on their expert knowledge before using them for analysis.
 \end{enumerate}
@@
 
-As a result of these issues, whether constructing the model manually or using
-an algorithm, the user has to rely on their expertise to make modifications to
-the model. One of the common ways to guide this post hoc modification is to use
+As a result of these issues, expert knowledge integration is vital to get usable 
+results from these algorithms.
+% whether constructing the model manually or using
+% an algorithm, the user has to rely on their expertise to make modifications to
+% the model. 
+One way to do this to specify expert knowledge in terms of
+fixing edges, blacklisting edges, or specifying tiered knowledge.
+The other ways is to do post hoc modification to the learned model using
 implied CI based model testing \citep{Ankan2021, Ankan2023}. In this testing
@@
 	\centering
-	% \includegraphics[scale=0.4]{plots/web_tool_full.png}
+	\includegraphics[scale=0.4]{plots/web_tool_full.png}
 	\caption{A screenshot of the web tool for constructing the model. Users

### 2025-01-23 paper/main.tex
Updates

@@
 
-% Due to these challenges, whether constructing the model manually or using an
-% algorithm, we typically have to rely on some expert knowledge to get a model
-% that we can use for further analysis. Hence, it is crucial to have tools to
-% guide this construction or modification process.
-
-As a result of these issues, expert knowledge integration is vital to get usable 
-results from these algorithms.
-% whether constructing the model manually or using
-% an algorithm, the user has to rely on their expertise to make modifications to
-% the model. 
-One way to do this to specify expert knowledge in terms of
+These challenges highlight the need for methods that effectively combine
+automated tools with expert-driven model refinement.  As a result of these
+issues, expert knowledge integration is vital to get usable results from these
+algorithms. One way to do this to specify expert knowledge in terms of
 fixing edges, blacklisting edges, or specifying tiered knowledge.
@@
 \label{sec:background}
-We consider random variable $ X $ and $ Y $ with a possible multi-dimensional
+We consider random variables $ X $ and $ Y $ with a possible multi-dimensional
 variable $ \bm{Z} = \{ Z_1, \cdots, Z_k \} $. We consider these variables in

### 2025-01-25 code/fig1/adult_sl.R
Updates

@@ -1,313 +1,6 @@
 library(dplyr)
 
-sqrtinv <- function(mA) {
-  ei <- eigen(mA)
-  d <- ei$values
-  d <- (d + abs(d)) / 2
-  d2 <- 1 / sqrt(d)
-  d2[d == 0] <- 0
-  return(ei$vectors %*% diag(d2) %*% t(ei$vectors))
-}
-
-pred2res <- function(pred) {
-  setNames(c(0, cumsum(pred)[-length(pred)]) -
-    c(rev(cumsum(rev(pred))), 0)[-1], names(pred))
-}
-
-p2r <- function(x, preddf) {
-  present.levels <- colnames(preddf)
-  missing.levels <- setdiff(levels(x), present.levels)
-  if (length(missing.levels) > 0) {
-    for (xx in missing.levels) {
-      preddf <- cbind(preddf, 0)
-    }
-    colnames(preddf) <- c(present.levels, missing.levels)
-    preddf <- preddf[, levels(x)]
-  }
-  as.matrix(apply(preddf, 1, pred2res), rownames = colnames(preddf))[
-    cbind(x, seq_along(x))
-  ]
-}
-
-residual.test <- function(x, y, z, predictor, test) {
-  # Random Forest test v1. Return the p-value for the independence
-  # test X _|_ Y | Z
-  #
-  # @param x: vector
-  # @param y: vector
-  # @param z: vector or data.frame
-  # @param predictor: 'glm' or 'rf'
-  # @param test: 'q3' or 'pillai'
-
-  n <- length(x)
-
-  # Step 1: Determine the variable type of x
-  if (all(class(x) == "factor")) {
-    x <- droplevels(x)
-    k <- length(levels(x))
-
-    stopifnot(k > 1)
-    tx <- table(seq_len(n), x)[, -k, drop = FALSE]
-
-    x_type <- "cat"
-  } else if (all(class(x) == c("ordered", "factor"))) {
-    x <- droplevels(x)
-    k <- 1
-    tx <- x
-
-    x_type <- "ord"
-  } else {
-    k <- 1
-    tx <- x
-
-    x_type <- "cont"
-  }
-
-  # Step 2: Determine the variable type of y
-  if (all(class(y) == "factor")) {
-    y <- droplevels(y)
-    r <- length(levels(y))
-
-    stopifnot(r > 1)
-    ty <- table(seq_len(n), y)[, -r, drop = FALSE]
-
-    y_type <- "cat"
[…]

### 2025-01-25 code/fig1/ci_tests.R

@@ -1,0 +1,308 @@
+sqrtinv <- function(mA) {
+  ei <- eigen(mA)
+  d <- ei$values
+  d <- (d + abs(d)) / 2
+  d2 <- 1 / sqrt(d)
+  d2[d == 0] <- 0
+  return(ei$vectors %*% diag(d2) %*% t(ei$vectors))
+}
+
+pred2res <- function(pred) {
+  setNames(c(0, cumsum(pred)[-length(pred)]) -
+    c(rev(cumsum(rev(pred))), 0)[-1], names(pred))
+}
+
+p2r <- function(x, preddf) {
+  present.levels <- colnames(preddf)
+  missing.levels <- setdiff(levels(x), present.levels)
+  if (length(missing.levels) > 0) {
+    for (xx in missing.levels) {
+      preddf <- cbind(preddf, 0)
+    }
+    colnames(preddf) <- c(present.levels, missing.levels)
+    preddf <- preddf[, levels(x)]
+  }
+  as.matrix(apply(preddf, 1, pred2res), rownames = colnames(preddf))[
+    cbind(x, seq_along(x))
+  ]
+}
+
+residual.test <- function(x, y, z, predictor, test) {
+  # Random Forest test v1. Return the p-value for the independence
+  # test X _|_ Y | Z
+  #
+  # @param x: vector
+  # @param y: vector
+  # @param z: vector or data.frame
+  # @param predictor: 'glm' or 'rf'
+  # @param test: 'q3' or 'pillai'
+
+  n <- length(x)
+
+  # Step 1: Determine the variable type of x
+  if (all(class(x) == "factor")) {
+    x <- droplevels(x)
+    k <- length(levels(x))
+
+    stopifnot(k > 1)
+    tx <- table(seq_len(n), x)[, -k, drop = FALSE]
+
+    x_type <- "cat"
+  } else if (all(class(x) == c("ordered", "factor"))) {
+    x <- droplevels(x)
+    k <- 1
+    tx <- x
+
+    x_type <- "ord"
+  } else {
+    k <- 1
+    tx <- x
+
+    x_type <- "cont"
+  }
+
+  # Step 2: Determine the variable type of y
+  if (all(class(y) == "factor")) {
+    y <- droplevels(y)
+    r <- length(levels(y))
+
+    stopifnot(r > 1)
+    ty <- table(seq_len(n), y)[, -r, drop = FALSE]
+
+    y_type <- "cat"
[…]

### 2025-01-25 code/fig4/fig4.py

@@ -1,0 +1,69 @@
+# Simulate a human structure learning based on CI tests.
+simulate_human_sl <- function(sim_data, true_dag, oracle_acc, max_iter=1e4){
+	varnames <- colnames(sim_data)
+	new_dag <- dagitty(paste0("dag{ ", paste(varnames, collapse=" "), " }"))
+	counter <- 0
+	blacklist_edges <- matrix(nrow=0, ncol=2)
+
+	while (counter < max_iter){
+		counter <- counter + 1
+		
+		# Take only potential edges with significant correlation.
+		effects <- compute_effects_v2(new_dag, sim_data) %>% filter(p < PVALUE_THRESHOLD)
+		effects <- effects[, c('X', 'A', 'Y', 'cor')]
+
+		# If edges are present do pruning.
+		edges_to_prune <- effects %>% filter(A == '->') %>% filter(abs(cor) < EFFECT_THRESHOLD)
+		if (edges_to_prune %>% nrow() > 0){
+			new_dag <- remove_edges(new_dag, edges_to_prune)
+            browser()
+		}
+
+		# Recompute effects after pruning. We don't really need to do this.
+		effects <- compute_effects_v2(new_dag, sim_data) %>% filter(p < PVALUE_THRESHOLD)
+		effects <- effects[, c('X', 'A', 'Y', 'cor')]
+		unexplain_cor_sorted <- effects %>% filter(abs(as.double(effects$cor)) > EFFECT_THRESHOLD) %>% filter(A == '--') %>% arrange(desc(abs(cor)))
+
+		# Remove blacklisted edges
+		if (blacklist_edges %>% nrow() > 0){
+			unexplain_cor_sorted <- unexplain_cor_sorted %>%
+				filter(!((X %in% blacklist_edges[, 1]) &
+                      (Y %in% blacklist_edges[, 2])))
+		}
+
+		# If no unexplained correlation is remaining, exit loop.
+		if(unexplain_cor_sorted %>% nrow() == 0){
+			print(paste0(counter, ": All corr explained"))
+			break
+		}
+		# Else add the highest correlated edge.
+		else{
+			oracle_edge <- oracle(
+				u=unexplain_cor_sorted[1, 'X'],
+				v=unexplain_cor_sorted[1, 'Y'],
+				true_dag=true_dag,
+				accuracy=oracle_acc
+			)
+			# If oracle doesn't know the direction of this edge, add it to
[…]

### 2025-01-25 code/fig1/adult_sl.R
Figure 1 data

@@ -28,66 +28,80 @@ d$HoursPerWeek <- as.double(d$HoursPerWeek)
 education_levels <- c("Preschool", "1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th", "12th", "HS-grad", "Some-college", "Assoc-voc", "Assoc-acdm", "Bachelors", "Masters", "Prof-school", "Doctorate")
 d$Education <- factor(d$Education, levels = education_levels, ordered = T)
 
+set.seed(42)
 d <- d[complete.cases(d), ]
-d <- d[1:500, ]
 
-# Using bnlearn
-dag <- bnlearn::pc.stable(d)
+for (n_samples in c(400, 800)){
+	df <- d[sample(nrow(d), n_samples), ]
 
-# Using pcalg
-var_types <- list(
-		Age = "cont",
-		Workclass = "cat",
-		Education = "ord",
-		MaritalStatus = "cat",
-		Occupation = "cat",
-		Relationship = "cat",
-		Race = "cat",
-		Sex = "cat",
-		HoursPerWeek = "cont",
-		Income = "cat",
-		NativeCountry = "cat"
-		)
+	# Using pcalg
+	var_types <- list(
+			Age = "cont",
+			Workclass = "cat",
+			Education = "ord",
+			MaritalStatus = "cat",
+			Occupation = "cat",
+			Relationship = "cat",
+			Race = "cat",
+			Sex = "cat",
+			HoursPerWeek = "cont",
+			Income = "cat",
+			NativeCountry = "cat"
+			)
 
-suffStat_q3 <- list(dm = d, adaptDF = F, test = "rf_q3", var_types = var_types)
-pc.cpdag.q3 <- pcalg::pc(
-		suffStat_q3,
-		indepTest = sl.test,
-		alpha = 0.05,
-		labels = colnames(d),
-		u2pd = "relaxed",
-		skel.method = "stable.fast",
-		numCores = 8,
-		verbose = F
-		)
+	# Using our mixed data test
+	suffStat_q3 <- list(dm = df, adaptDF = F, test = "rf_q3", var_types = var_types)
+	cpdag.q3 <- pcalg::pc(
+			suffStat_q3,
+			indepTest = sl.test,
+			alpha = 0.05,
+			labels = colnames(df),
+			u2pd = "relaxed",
+			skel.method = "stable.fast",
+			numCores = 8,
[…]

### 2025-01-25 paper/main.tex
Updates paper

@@
 
-To tackle these issues with CI based testing, we take inspiration from
-\emph{modification indices} in specification search. The modification indices
-method tries to identify and rank changes in the model such that modifications
-that improve the model fit the most are ranked higher.This ranking allows
-researchers to focus first on the suggestions that can improve the model the
-most and while integrating their expertise to choose the most optimal
-modifications. Analogous to this, in this paper we develop a measure for
-partial association for mixed data that allows us to rank the CI violations.
-Additionally, this measure of conditional association also allows us tackle
-other issues with CI based testing outlined in the previous paragraphs.
+There has been work on trying to come up with a single test to combine all
+these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
+what is missing is the guide on how to how to fix these tests. To tackle these
+issues with CI based testing, we take inspiration from \emph{modification
+indices} in specification search. The modification indices method tries to
+identify and rank changes in the model such that modifications that improve the
+model fit the most are ranked higher.This ranking allows researchers to focus
+first on the suggestions that can improve the model the most and while
+integrating their expertise to choose the most optimal modifications. Analogous
+to this, in this paper we develop a measure for partial association for mixed
+data that allows us to rank the CI violations. Additionally, this measure of
+conditional association also allows us tackle other issues with CI based
+testing outlined in the previous paragraphs.
 

### 2025-01-25 code/fig3/make_plots.R
Updates figure size

@@ -195,7 +195,7 @@ make_plots_v2 <- function(filename, plot_type){
 		ylim(0, 40) +
 		guides(linetype='none', fill='none')
 
-	ggsave("shd_ribbon.pdf", p_shd, height=2.5, width=3.2, units='in')
+	ggsave("shd_ribbon.pdf", p_shd, height=2, width=3.2, units='in')
 
 
 	d_sid <- d %>% select(edge_prob, oracle_acc, ges_lower_sid_mean, ges_upper_sid_mean, ges_lower_sid_sd, ges_upper_sid_sd, hc_lower_sid_mean, hc_upper_sid_mean, hc_lower_sid_sd, hc_upper_sid_sd, pc_lower_sid_mean, pc_upper_sid_mean, pc_lower_sid_sd, pc_upper_sid_sd, human_sid_mean, human_sid_sd)
@@ -253,7 +253,7 @@ make_plots_v2 <- function(filename, plot_type){
 		ylim(0, 90) +
 		guides(linetype='none', fill='none')
 
-	ggsave("sid_ribbon.pdf", p_sid, height=2.5, width=3.2, units='in')
+	ggsave("sid_ribbon.pdf", p_sid, height=2, width=3.2, units='in')
 }
 
 make_plots_v2(filename='results/sl_results_mixed.csv', plot_type='ribbon')

### 2025-01-25 code/fig4/make_plots.R
Updates figure 4

@@ -18,22 +18,25 @@ d_ges$algo <- 'GES'
 
 d_results = rbind(d_expert, d_ges)
 
-
 p_unexplained_effect <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
 	geom_line(alpha=0.6) +
 	geom_point(alpha=0.6, show.legend=F) +
 	theme_minimal(base_size=8) +
 	labs(x="Iteration No.") +
-	labs(y="Total Unexplained Effect")
+	labs(y="Total Unexplained Effect") + 
+	theme(legend.position='top',
+	      legend.title=element_blank(),
+	      legned.text=element_blank())
 
 p_ll <- ggplot(d_results, aes(x=x, y=ll, color=algo)) +
 	geom_line(alpha=0.6) +
 	geom_point(alpha=0.6, show.legend=F) +
 	theme_minimal(base_size=8) +
 	labs(x="Iteration No.") +
-	labs(y="Log-Likelihood")
+	labs(y="Log-Likelihood") +
+	theme(legend.position='top',
+	      legend.title=element_blank(),
+	      legned.text=element_blank())
 
-ggsave('unexplained_effect.pdf', p_unexplained_effect, height=2.5, width=3.2, units='in')
-ggsave('ll.pdf', p_ll, height=2.5, width=3.2, units='in')
-
-
+ggsave('unexplained_effect.pdf', p_unexplained_effect, height=2.5, width=1.6, units='in')
+ggsave('ll.pdf', p_ll, height=2.5, width=1.6, units='in')

### 2025-01-25 code/plot_likelihood.py

@@ -1,3 +1,0 @@
-from data import get_adult_df
-
-df_adult = get_adult_df()

### 2025-01-25 paper/main.tex
Udpates
Updates figure 4
Updates

@@
 these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
-what is missing is the guide on how to how to fix these tests. To tackle these
+what is missing is the guide on how to how to fix these tests. Additionally
+there has been a lot of work to integrate expert knowledge into causal
+discovery algorithms which has shown significant improvement. To tackle these
 issues with CI based testing, we take inspiration from \emph{modification
@@
 \begin{figure}
-\begin{subfigure}{0.5 \textwidth}
+\begin{subfigure}{0.25 \textwidth}
 	\includegraphics[page=1]{figures.pdf}
-\end{subfigure}
-\begin{subfigure}{0.5 \textwidth}
+\end{subfigure}%
+\begin{subfigure}{0.25 \textwidth}
 	\includegraphics[page=2]{figures.pdf}
 \end{subfigure}
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.25 \textwidth}
 	\includegraphics[page=3]{figures.pdf}
+\end{subfigure}%
+\begin{subfigure}{0.25\textwidth}
+	\includegraphics[page=4]{figures.pdf}
 \end{subfigure}
@@
 	\centering
-	\includegraphics[scale=0.4]{plots/web_tool_full.png}
+	\includegraphics[scale=0.4]{../code/plots/web_tool_full.png}
 	\caption{A screenshot of the web tool for constructing the model. Users
@@
 \begin{figure}
+	\centering
 	\begin{subfigure}{0.5\textwidth}
@@
 \begin{figure}
-	\begin{subfigure}{0.5\textwidth}
+	\begin{subfigure}{0.25\textwidth}
 		\centering
@@
 		\caption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
+	\end{subfigure}%
+	\begin{subfigure}{0.25\textwidth}
 		\centering
@@
 		\centering
-		\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
+		\includegraphics[scale=0.3]{../code/plots/llm_adult.png}
 		\caption{\todo[inline]{Change figure to have consistent style}}

### 2025-01-28 paper/main.tex
updates
Updates

@@
 behind observed phenomena and guide effective interventions or policy
-decisions. \emph{Causal Discovery} methods aim to identify such relationships
+decisions. \emph{Causal Discovery} methods aim to discover such relationships
 among random variables using observational data. Approaches to causal discovery
 have been developed within both the Directed Acyclic Graphs (DAGs) and
-Structural Equation Models (SEMs) frameworks, each adapting a different approach.
+Structural Equation Models (SEMs) frameworks, each adapting a different
+approach.
 
@@
 Search} \citep{Long1983} and uses method such as modification indices, and
-Wald-based tests \citep{Marcoulides2018}.
+Wald-based tests \citep{Marcoulides2018}. While there has been some effort in
+DAG literature as well towards allowing for expert knowledge but these methods
+focus towards providing the expert knowledge to the automated algorithms
+\citep{Bang2023, Meek1995}. \todo[inline]{Whether to keep this sentence?}
 
@@
 Figure~\ref{fig:intro} shows an example to highlight some of these issues using
-Adult income dataset.
+the Adult income dataset \citep{Becker1996}. As can be seen in the figure
+depending on the sample size and the algorithm used, we get vastly different
+results.
 
-These challenges highlight the need for methods that effectively combine
-automated tools with expert-driven model refinement.  As a result of these
-issues, expert knowledge integration is vital to get usable results from these
-algorithms. One way to do this to specify expert knowledge in terms of
-fixing edges, blacklisting edges, or specifying tiered knowledge.
 The other ways is to do post hoc modification to the learned model using
@@
 
-There has been work on trying to come up with a single test to combine all
-these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
-what is missing is the guide on how to how to fix these tests. Additionally
-there has been a lot of work to integrate expert knowledge into causal
-discovery algorithms which has shown significant improvement. To tackle these
-issues with CI based testing, we take inspiration from \emph{modification
+There has been work on integrating expert knowledge into model building process
+and has been quite successful such as tpc and background knowledge
+specification. There has been work on trying to come up with a single test to
+combine all these test to give a single answer \citep{Eulig2023, Shipley2000}.
+However, what is missing is the guide on how to how to fix these tests.
+Additionally there has been a lot of work to integrate expert knowledge into
+causal discovery algorithms which has shown significant improvement. To tackle
+these issues with CI based testing, we take inspiration from \emph{modification
 indices} in specification search. The modification indices method tries to
@@
 
-			% \begin{scope}[xshift=-3cm]
-			% 	\draw[thick, ->, double, cyan] (2.5,0) -- (3,0);
-			% 	\node[rectangle, align=center, inner sep=1pt] at (5, 0) {
-			% 		\begin{minipage}{.2\textwidth}
-			% 			\begin{equation*}
-			% 				\begin{split}
-			% 					\textnormal{Age} &\ci \textnormal{Incm} \rvert \textnormal{Edct, Hrpw, Sex} \\
-			% 					\textnormal{Age} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
-			% 					\textnormal{Age} &\ci \textnormal{Sex} \\
-			% 					\textnormal{Hrpw} &\ci \textnormal{Occp} \rvert \textnormal{Edct, Sex} \\
-			% 				\end{split}
-			% 			\end{equation*}
-			% 		\end{minipage}
-			% 		};
-			% \end{scope}
-
 Our main contributions in this paper are as follows:
@@
 
-\subsection{Ranking using Correlation Coefficients}
-
-\subsection{Measures of (Conditional) Association}
-\todo[inline]{Complete this section}
+\subsection{Measures of Conditional Association}
 As our DAG construction procedure relies on a measure of conditional
@@
 known measures for specific data types. These measure of associations are
-typically the effect size of various conditional indepdnence tests but they
+typically the effect size of various conditional independence tests but they
 don't need to be. However, there are no generalized measure of association for
@@
 In the case of discrete variables, Cramer\'s V has been typically used as a
-measure of association. When there are no conditional variables, $ \bm{Z} = \emptyset $,
-Cramer\'s V can be computed directly from the contingency table.
+measure of association. When there are no conditional variables, $ \bm{Z} =
+\emptyset $, Cramer\'s V can be computed directly from the contingency table.
 
@@
 
-\paragraph{$ X $ is ordinal and $ Y $ is continuous}
-Polyserial Correlation
-
-When $ \bm{Z} = \emptyset $, ordinal variable can be assumed to be coming from
-a thresholded normal distribution, estimate the thresholds and latent variable
-to compute the correlation coefficient.
-
-\paragraph{$ X $ and $ Y $ are ordinal}
-Polychoric Correlation
+\paragraph{$ X $ is ordinal and $ Y $ is continuous or ordinal}
+Polyserial Correlation (when one is oridinal and the other is continuous) and
+polychoric correlation (when both are ordinal) can be use. Here the assumption
+is made that $ X $ is a thresholded version of an underlying continuous normal
+variable. The method then tries to learn the thresholds for this latent
+variable while trying to maximize the correlation between $ X_{lat} $ and $ Y
+$. This gives us values that are Pearson's correlation coefficient.
 

### 2025-01-28 paper/main.tex
Updates

@@
 \label{sec:background}
+\subsection{Notations}
 We consider random variables $ X $ and $ Y $ with a possible multi-dimensional
@@
 \end{enumerate}
+
+So, in a sense this measure generalized currently used measures of associations
+to mixed data.
 

### 2025-01-28 paper/main.tex
Updates
updates

@@
 domains remains limited, as researchers often prefer to construct DAGs manually
-based on their expert knowledge. This preference arises due to practical issues
-with these automated algorithms, such as making obvious errors and outputting
-Markov Equivalence Classes (MECs). To address these challenge, we propose an
-iterative method to assist researchers in manually constructing or modifying
-DAGs by combining implied Conditional Independence (CI) testing with expert
-knowledge. We introduce a novel measure of partial association for mixed data
-to rank violations of implied CIs, enabling users to prioritize model
-modifications while leveraging their expertise to decide the orientation of
-edges when making these modifications. Empirical results show that this
+based on their expert knowledge. This preference arises due to several
+practical issues with these automated algorithms, such as making obvious errors
+and outputting Markov Equivalence Classes (MECs). To address these challenges,
+we propose an iterative method to assist researchers in manually constructing
+or modifying DAGs by combining implied Conditional Independence (CI) testing
+with expert knowledge. We introduce a novel measure of partial association for
+mixed data to rank violations of implied CIs, enabling users to prioritize
+model modifications while leveraging their expertise to decide the orientation
+of edges when making these modifications. Empirical results show that this
 approach achieves performance comparable to automated algorithms, assuming the
@@
 
-Figure~\ref{fig:intro} shows an example to highlight some of these issues using
-the Adult income dataset \citep{Becker1996}. As can be seen in the figure
-depending on the sample size and the algorithm used, we get vastly different
-results.
+Fig.~\ref{fig:intro} highlights some of these issues using the Adult income
+dataset \citep{Becker1996}. As can be seen in the figure depending on the
+sample size and the algorithm used, we get vastly different results.
 
-The other ways is to do post hoc modification to the learned model using
-implied CI based model testing \citep{Ankan2021, Ankan2023}. In this testing
-procedure, users test whether the implied CIs of the DAG hold in the dataset or
-not using statistical CI tests. Based on the results of these statistical
-tests, we can accordingly make modifications to the current model to fix the
-violations. However, this CI based testing also has a few drawbacks: 1) CIs are
-implied only by missing edges in the model, so we can not test for already
-existing edges. 2) In high sample size scenarios, p-values can suggest a
-significant association between variables even if they have a very weak
-relationship. 3) Number of implied CIs can be large making it difficult to
-manually go through each of them; for example, in the moderately sized alarm
-network with $ 37 $ node, there are $287$ implied CIs.
+As a result of these challenges users have relied on model testing approaches
+along with their expert knowledge to construct these models. The other ways is
+to do post hoc modification to the learned model using implied CI based model
+testing \citep{Ankan2021, Ankan2023}. In this testing procedure, users test
+whether the implied CIs of the DAG hold in the dataset or not using statistical
+CI tests. Based on the results of these statistical tests, we can accordingly
+make modifications to the current model to fix the violations. However, this CI
+based testing also has a few drawbacks: 1) CIs are implied only by missing
+edges in the model, so we can not test for already existing edges. 2) In high
+sample size scenarios, p-values can suggest a significant association between
+variables even if they have a very weak relationship. 3) Number of implied CIs
+can be large making it difficult to manually go through each of them; for
+example, in the moderately sized alarm network with $ 37 $ node, there are
+$287$ implied CIs.
 
@@
 
-% using a conditional measure of
-% association.  To deal with both this issues, in this paper, we develop a
-% measure of partial association which can be using in combination of the CI test
-% to rank the violations in these CI tests. On the other hand, in Structural
-% Equation Models (SEMs) literature, more emphasis has been placed on developing
-% tools to help researchers manually construct models while integrating their
-% domain expertise. Our proposed approach uses the measure of association to find
-% correlated variables in the data that are not explained by the current model.
-% Using the measure, we can then rank the pair of variables between which adding
-% an edge would improve the explanation the most. Based on this information, the
-% user can select a potential pair of variables and based on their expert
-% knowledge decide the direction of the edge.
-
 \begin{figure}
@@
 	\includegraphics[page=1]{figures.pdf}
+	\caption{PC algorithm $ \chi^2 $ test. $ N=400$ }
 \end{subfigure}%
@@
 	\includegraphics[page=2]{figures.pdf}
+	\caption{PC algorithm $ \chi^2 $ test. $ N=800$ }
 \end{subfigure}
@@
 	\includegraphics[page=3]{figures.pdf}
+	\caption{PC algorithm residualization test. $ N=400$ }
 \end{subfigure}%
@@
 	\includegraphics[page=4]{figures.pdf}
+	\caption{PC algorithm residualization test. $ N=800$ }
+\end{subfigure}
+\begin{subfigure}{0.25 \textwidth}
+	\includegraphics[page=5]{figures.pdf}
+	\caption{Hill Climb Search BIC Score. $ N=400$ }
+\end{subfigure}%
+\begin{subfigure}{0.25\textwidth}
+	\includegraphics[page=6]{figures.pdf}
+	\caption{Hill Climb Search BIC Score. $ N=800$ }
 \end{subfigure}

### 2025-01-28 paper/main.tex
Updates

@@
 \usepackage{subcaption}
+\usepackage{fancyvrb}
 % \usepackage[ruled]{algorithm2e}
@@
 	\begin{subfigure}{0.5\textwidth}
-		\caption{\todo[inline]{Add the prompt used}}
+		\begin{Verbatim}[fontsize=\footnotesize]
+You are an expert in Social Science. Following
+are the descriptions of two variables:
+
+<A>: {description of variable A}
+<B>: {description of variable B}
+
+Which of the following two options is the most 
+likely causal direction between these variables:
+
+1. <A> causes <B>
+2. <B> causes <A>
+
+Return a single letter answer between the 
+choices above; Do not provide any reasoning 
+in the answer; Do not add any text formatting 
+to the answer.
+		\end{Verbatim}
+	\caption{Prompt used for the LLM}
+	\end{subfigure}
+	\begin{subfigure}{0.5 \textwidth}
+		\begin{Verbatim}[fontsize=\footnotesize]
+Age: The age of a person
+Workclass: The workplace where the person is 
+     employed such as Private industry, 
+     or self employed
+Education: The highest level of education the 
+     person has finished
+MaritalStatus: The marital status of the person
+Occupation: The kind of job the person does. 
+     For example, sales, craft repair, clerical.
+Relationship: The relationship status of 
+     the person
+Race: The ethnicity of the person
+Sex: The sex or gender of the person
+HoursPerWeek: The number of hours per week the 
+     person works
+NativeCountry: The native country of the person
+Income: The income i.e. amount of money the 
+     person makes
+		\end{Verbatim}
+	\caption{Variable descriptions used for prompting the LLM}
 	\end{subfigure}
@@
 		\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
-		\caption{\todo[inline]{Change figure to have consistent style}}
+		\caption{The learned DAG}
 	\end{subfigure}

### 2025-01-28 code/fig5/make_plots.py
Fixes figure 5

@@ -1,6 +1,6 @@
 from pgmpy.models import BayesianNetwork
 
-with open('adult_llm.txt', 'r') as f:
+with open('llm_adult.txt', 'r') as f:
     edge_list = eval(f.readline())
 
 dag = BayesianNetwork(edge_list)

### 2025-01-28 code/fig5/llm_v2.py
Udpated llm code with continuous variables

@@ -1,0 +1,60 @@
+import os
+import pandas as pd
+import networkx as nx
+from itertools import combinations
+
+import google.generativeai as genai
+from pgmpy.estimators import ExpertInLoop
+from pgmpy.base import DAG
+
+
+def preprocess_data():
+    df = pd.read_csv("../utils/adult_proc.csv", index_col=0)
+    df.Education = pd.Categorical(
+        df.Education,
+        categories=[
+            "Preschool",
+            "1st-4th",
+            "5th-6th",
+            "7th-8th",
+            "9th",
+            "10th",
+            "11th",
+            "12th",
+            "HS-grad",
+            "Some-college",
+            "Assoc-voc",
+            "Assoc-acdm",
+            "Bachelors",
+            "Prof-school",
+            "Masters",
+            "Doctorate",
+        ],
+        ordered=True,
+    )
+    df.Workclass = pd.Categorical(df.Workclass, ordered=False)
+    df.MaritalStatus = pd.Categorical(df.MaritalStatus, ordered=False)
+    df.Occupation = pd.Categorical(df.Occupation, ordered=False)
+    df.Relationship = pd.Categorical(df.Relationship, ordered=False)
+    df.Race = pd.Categorical(df.Race, ordered=False)
+    df.Sex = pd.Categorical(df.Sex, ordered=False)
+    df.NativeCountry = pd.Categorical(df.NativeCountry, ordered=False)
+    df.Income = pd.Categorical(df.Income, ordered=False)
+    return (df)
+
+descriptions = {'Age': 'The age of a person',
+                'Workclass': 'The workplace where the person is employed such as Private industry, or self employed',
+                'Education': "The highest level of education the person has finished",
+                'MaritalStatus': "The marital status of the person",
+                "Occupation": "The kind of job the person does. For example, sales, craft repair, clerical",
+                "Relationship": "The relationship status of the person",
+                "Race": "The ethnicity of the person",
+                "Sex": "The sex or gender of the person",
+                "HoursPerWeek": "The number of hours per week the person works",
+                "NativeCountry": "The native country of the person",
+                "Income": "The income i.e. amount of money the person makes"
+                }
+
+dag = ExpertInLoop(preprocess_data()).estimate(variable_descriptions=descriptions, effect_size_threshold=0.1)
+with open('llm_adult.txt', 'w') as f:
+    f.write(str(list(dag.edges())))

### 2025-01-28 code/fig5/llm_v2.py
Updates

@@ -9,7 +9,10 @@ from pgmpy.base import DAG
 
 
 def preprocess_data():
-    df = pd.read_csv("../utils/adult_proc.csv", index_col=0)
+    df = pd.read_csv("../utils/adult_unproc.csv")
+    df = df.dropna()
+    df = df.drop(['Fnlwgt', 'EducationNum', 'CapitalGain', 'CaptialLoss'], axis=1)
+    df = df.iloc[:1000, :]
     df.Education = pd.Categorical(
         df.Education,
         categories=[
@@ -55,6 +58,6 @@ descriptions = {'Age': 'The age of a person',
                 "Income": "The income i.e. amount of money the person makes"
                 }
 
-dag = ExpertInLoop(preprocess_data()).estimate(variable_descriptions=descriptions, effect_size_threshold=0.1)
+dag = ExpertInLoop(preprocess_data()).estimate(variable_descriptions=descriptions, effect_size_threshold=0.05)
 with open('llm_adult.txt', 'w') as f:
     f.write(str(list(dag.edges())))

### 2025-01-28 paper/main.tex
Updates

@@
 \begin{subfigure}{0.25 \textwidth}
+	\includegraphics[page=5]{figures.pdf}
+	\caption{Hill Climb Search BIC Score. $ N=400$ }
+\end{subfigure}%
+\begin{subfigure}{0.25\textwidth}
+	\includegraphics[page=6]{figures.pdf}
+	\caption{Hill Climb Search BIC Score. $ N=800$ }
+\begin{subfigure}{0.25 \textwidth}
 	\includegraphics[page=3]{figures.pdf}
@@
 \end{subfigure}
-\begin{subfigure}{0.25 \textwidth}
-	\includegraphics[page=5]{figures.pdf}
-	\caption{Hill Climb Search BIC Score. $ N=400$ }
-\end{subfigure}%
-\begin{subfigure}{0.25\textwidth}
-	\includegraphics[page=6]{figures.pdf}
-	\caption{Hill Climb Search BIC Score. $ N=800$ }
 \end{subfigure}
 
-\caption{Markov Equivalence Class (MEC) learnt for Adult Income Dataset using (a) PC algorithm, (b) Hill-Climb Search with BIC score, (c) PC algorithm with MXM test}
+\caption{Markov Equivalence Class (MEC) learned for Adult Income Dataset using (a) PC algorithm with chi-squared test, (b) Hill-Climb Search with BIC score, (c) PC algorithm with residualization based test.
+There is significant variation in the learned model depending on the sample size and the algorithm used.
+Additionally, all of them make some mistakes.}
 \label{fig:intro}

### 2025-01-28 code/fig5/llm_v2.py
Updates LLm plot

@@ -58,6 +58,6 @@ descriptions = {'Age': 'The age of a person',
                 "Income": "The income i.e. amount of money the person makes"
                 }
 
-dag = ExpertInLoop(preprocess_data()).estimate(variable_descriptions=descriptions, effect_size_threshold=0.05)
+dag = ExpertInLoop(preprocess_data()).estimate(variable_descriptions=descriptions, effect_size_threshold=0.1)
 with open('llm_adult.txt', 'w') as f:
     f.write(str(list(dag.edges())))

### 2025-01-30 paper/main.tex
Fixes
Updates

@@
 \usepackage{fancyvrb}
-% \usepackage[ruled]{algorithm2e}
+\usepackage[ruled]{algorithm2e}
 
@@
 	\caption{Hill Climb Search BIC Score. $ N=800$ }
+\end{subfigure}
 \begin{subfigure}{0.25 \textwidth}
@@
 	\caption{PC algorithm residualization test. $ N=800$ }
-\end{subfigure}
 \end{subfigure}
@@
 
+\begin{verbatim}
+	Input: Dataset $ D $ on variables $ \bm{X} $
+	Output: Learned DAG G.
+
+	function compute_effects(G):
+		pvalues <- dict()
+		effects <- dict()
+		for X_i, X_j in X:
+			if edge between X_i, X_j:
+				p_values[(X_i, X_j)] <- pvalue(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
+				effects[(X_i, X_j)] <- \phi(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
+			else:
+				p_values[(X_i, X_j)] <- pvalue(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
+				effects[(X_i, X_j)] <- \phi(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
+		return p_values, effects
+
+	$ G $ <- Empty Graph
+	p_values, effects <- compute_effects(G)
+
+\end{verbatim}
+
+\begin{algorithm}
+	\KwData{Data set $ D $ on variables $ \bm{X} $, pthres, effectthres}
+	\KwResult{The learned DAG: $ G $}
+	\BlankLine
+	$ G \leftarrow $ Empty Graph
+\end{algorithm}
+
 \subsection{Using Large Language Models as Experts}

### 2025-01-31 paper/main.tex
Updates

@@
 \subsection{Notations}
-We consider random variables $ X $ and $ Y $ with a possible multi-dimensional
-variable $ \bm{Z} = \{ Z_1, \cdots, Z_k \} $. We consider these variables in
-the mixed data setting where these variables can be any combination of
-continuous, categorical, or ordinal unless specified. We also assume all
-variables to be observed. We use $ \rvert \bm{Z} \rvert $ to represent the
-cardinality of the variable $ \bm{Z} $. We write $ \bm{x} = (x_1, \cdots, x_n)
-$ for a sample from $ X $ of size $ n $. We write the expectation of a variable
-$ X $ as $ \mathbb{E}[X] $, conditional expectations as $ \mathbb{E}[X |
-\bm{Z}] $.
+We denote random variables with uppercase letters $ X $, and a set of random
+variables as $ \bm{X} = \{X_1, \cdots, X_k\} $ with $ \rvert \bm{X} \rvert = k
+$. A sample from the random variable $ X $ is denoted as $ x $ and from a set
+of random variables $ \bm{X} $ is denoted as $ \bm{x} $. In this paper, we
+consider random variables in the mixed data setting where each of the random
+variables can be continuous, categorical, or ordinal unless specified.
 
-We consider the problem of estimating the partial (or conditional) association
-between $ X $ and $ Y $ in the presence of a set of conditioning variables $
-\bm{Z} $. When $ Z = \emptyset $, this is equivalent to the marginal
-association between $ X $ and $ Y $.
+We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $, conditional
+expectations as $ \mathbb{E}[X | \bm{Z}] $.
 
 \subsection{Measures of Conditional Association}
-As our DAG construction procedure relies on a measure of conditional
-association between variables, in this section we list out some of the already
-known measures for specific data types. These measure of associations are
-typically the effect size of various conditional independence tests but they
-don't need to be. However, there are no generalized measure of association for
-mixed data, that we propose in the next section. This is by no means an
-exhaustive list, rather just to give an idea of what can be done.
+Our proposed method for iterative DAG construction relies on a measure of
+conditional association between variables. We are interested in measuring the
+association between variables $ X $ and $ Y $ when conditioned on a set of
+variables $ \bm{Z} $ which is potentially an empty set, $ \bm{Z} = \emptyset $.
+Depending on the type of $ X $, $ Y $, and $ \bm{Z} $ various measures of
+associations have been used. Typically, these are the effect size measures of
+the conditional independence test $ X \ci Y \rvert \bm{Z} $. Here, we give an
+overview of some of the commonly used measures for different data types.
 
-\paragraph{Both $ X $ and $ Y $ are continuous}
-Where there are no conditional variables, $ \bm{Z} = \emptyset $, Pearson's
-correlation coefficient can be used as a measure of association between the
-variables. This does make various assumptions such as linearity between the
-variables, normal distribution for $ X $ and $ Y $, and homoscedasticity.
-
-In the case when $ \bm{Z} \neq \emptyset $, then partial correlations can be
-used. This requires using two linear regression models: $ E_X: X \sim \bm{Z} $
-and $ E_Y: Y \sim \bm{Z} $. Then computing the residuals using these two
-regressions as: $ R_X = X - E_X(\bm{Z}) $ and $ R_Y = Y - E_Y(\bm{Y}) $. After
-this a simple correlation coefficient can be computed between $ R_X $ and $ R_Y $.
+\subsubsection{Both $ X $ and $ Y $ are continuous}
+In the case when both $ X $ and $ Y $ are continuous variables with the
+assumptions of normality and linearly relationship, typically Pearson's
+correction coefficients are used. In the case when $ \bm{Z} = \emptyset $, the
+correlation coefficient can be directly computed. When $ \bm{Z} \neq \emptyset
+$, partial correlations are used. One of the ways to estimate it is to use two
+linear regression models: $ E_X: x \sim \bm{z} $, and $ E_Y: y \sim \bm{z} $.
+Then we take the compute the Pearson's correlation coefficient between the
+residuals of these two regressions: $ R_X = x - E_X(\bm{z}) $, $ R_Y = y -
+E_Y(\bm{z}) $.
 
@@
 
-\section{A Measure of Association for Mixed Data}
+\section{A Generalized Measure of Conditional Association}
 \label{sec:mixed_association}
 
-In this section, we introduce a novel measure of partial association for mixed
+As there are no generalized measures of association for mixed data, in this
+section, we first introduce a novel measure of partial association for mixed
 data by extending the concept of partial correlations \citep{Whittaker2009}
@@
 continuous variables, residuals are typically computed using a regression-based
-method, and Pearson's correlation coefficient is used to measure the association
-between these residuals. To extend this approach to mixed data, we combine a mixed
-data residualization method \citep{Ankan2023} with a multivariate measure of
-association based on canonical correlations, Pillai's Trace \citep{Pillai1955}.
+method, and Pearson's correlation coefficient is used to measure the
+association between these residuals. To extend this approach to mixed data, we
+combine a mixed data residualization method \citep{Ankan2023} with a
+multivariate measure of association based on canonical correlations, Pillai's
+Trace \citep{Pillai1955}.
 
 Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $, and $
-\bm{Z} $, we aim to estimate the conditional
-association $ \rho(X, Y; \bm{Z}) $ between $ X $ and $ Y $ conditioned on $
-\bm{Z} $. We use the residualization approach from \citet{Ankan2023} to compute 
-the residuals. The residuals are defined as follows, depending on the type of 
-dependent variable:
+\bm{Z} $, we aim to estimate the conditional association $ \rho(X, Y; \bm{Z}) $
+between $ X $ and $ Y $ conditioned on $ \bm{Z} $. We use the residualization
+approach from \citet{Ankan2023} to compute the residuals. For each of the
+variable $ X $ and $ Y $, we compute the residual $ R_X $ and $R_Y$
+respectively as follows depending on their type.
 
@@
 	p_values, effects <- compute_effects(G)
-
+	while (p_values > p_thres, 
 \end{verbatim}

### 2025-01-31 paper/main.tex
Updates
Adds example
Adds figure caption

@@
 
+\begin{figure}
+	\begin{subfigure}{0.17 \textwidth}
+		\includegraphics[page=1]{example.pdf}
+		\caption{True DAG}
+	\end{subfigure}%
+	\begin{subfigure}{0.17 \textwidth}
+		\includegraphics[page=2]{example.pdf}
+		\caption{No edges}
+	\end{subfigure}%
+	\begin{subfigure}{0.17 \textwidth}
+		\includegraphics[page=3]{example.pdf}
+		\caption{Orient $ X_1 \rightarrow X_3 $}
+	\end{subfigure}
+	\begin{subfigure}{0.17 \textwidth}
+		\includegraphics[page=4]{example.pdf}
+		\caption{Orient $ X_2 \rightarrow X_3 $}
+	\end{subfigure}%
+	\begin{subfigure}{0.17 \textwidth}
+		\includegraphics[page=5]{example.pdf}
+		\caption{Orient $ X_3 \rightarrow X_4 $}
+	\end{subfigure}%
+	\begin{subfigure}{0.17 \textwidth}
+		\includegraphics[page=6]{example.pdf}
+		\caption{Orient $ X_3 \rightarrow X_5 $}
+	\end{subfigure}
+	\caption{An example showing a greedy approach for causal discovery using the mixed data measure of association.
+We start with an empty graph (b) where all the variables in the model are associated with each other.
+The strength of association is denoted using the thickness of the dotted edges.
+We select the pair of variables with the highest association, for example, $X_1$ and $X_3$, and orient it based on our domain knowledge.
+We continue the process till we have a DAG.
+	}
+\end{figure}
+
 \section{Connection to Score Based Methods}
@@
 \section{Conclusions}
+Future work:
+\begin{itemize}
+	\item Add latent variables; extend to ADMGs.
+	\item Can potentially be combined with pairwise orientation methods.
+	\item Improvements in the web-interface.
+\end{itemize}
 

### 2025-01-31 paper/main.tex
Updates

@@
 		partial correlations to mixed data types.
+	\item \textbf{Generalizes polychoric and polyserial correlation: }
+		Under the assumption that the ordinal variable is generated by
+		discretizing an underlying gaussian continuous variable, this
+		effect size is equivalent to polychoric and polyserial
+		correlation. As both of them recover the Pearson correlation
+		coefficient.
 	\item \textbf{Related to Cramer\'s V: } 
@@
 
-\section{Using Measure of Association for DAG Modification and Construction}
+\section{Using Measure of Association for DAG Construction}
 \label{sec:modification}
@@
 can remove the edge.
-
-\todo[inline]{Add a figure showing an example of how this modification works.}
 

### 2025-01-31 paper/main.tex
Updates

@@
 algorithms for learning DAGs from data. However, their adoption in applied
-domains remains limited, as researchers often prefer to construct DAGs manually
-based on their expert knowledge. This preference arises due to several
-practical issues with these automated algorithms, such as making obvious errors
-and outputting Markov Equivalence Classes (MECs). To address these challenges,
-we propose an iterative method to assist researchers in manually constructing
-or modifying DAGs by combining implied Conditional Independence (CI) testing
-with expert knowledge. We introduce a novel measure of partial association for
-mixed data to rank violations of implied CIs, enabling users to prioritize
-model modifications while leveraging their expertise to decide the orientation
-of edges when making these modifications. Empirical results show that this
-approach achieves performance comparable to automated algorithms, assuming the
-user can correctly orient edges in one out of three cases. Additionally, we
-demonstrate that, in the absence of an expert, a Large Language Model can
-potentially determine edge orientations.
+domains remain limited, as researchers often prefer to construct DAGs manually
+based on their domain knowledge. This preference arises due to several
+practical issues with automated algorithms, such as they make obvious errors
+and out Markov Equivalence Classes (MECs). To address these challenges, we
+propose an iterative method to assist researchers in manually constructing or
+improving an existing DAG by combining implied Conditional Independence (CI)
+testing with expert knowledge. Our proposed method utilizes a novel measure of
+partial association for mixed data to rank violations of implied CIs in a given
+DAG, enabling users to prioritize modifications while utilizing their expertise
+when making these modifications. Empirical results show that this manual model
+construction approach achieves performance comparable to automated algorithms
+even if the user is able to decide the orientation of an edge correctly in one
+out of three cases. Additionally, we demonstrate that, in the absence of an
+expert, a Large Language Model (LLM) can potentially determine edge orientations.
 
@@
 Search} \citep{Long1983} and uses method such as modification indices, and
-Wald-based tests \citep{Marcoulides2018}. While there has been some effort in
+Wald-based tests \citep{Marcoulides2018}.
+
+While there has been some effort in
 DAG literature as well towards allowing for expert knowledge but these methods
 focus towards providing the expert knowledge to the automated algorithms
-\citep{Bang2023, Meek1995}. \todo[inline]{Whether to keep this sentence?}
+\citep{Bang2023, Meek1995}.
 
@@
 		scheme used for categorical variables.
-	\item \textbf{Generalizes Partial Correlations to Mixed Data: }
+	\item \textbf{Equivalent to Partial Correlations for Continuous variables}
 		Canonical correlations is equivalent to Pearson's correlation coefficient
@@
 		partial correlations to mixed data types.
-	\item \textbf{Generalizes polychoric and polyserial correlation: }
+	\item \textbf{Equivalent to polychoric and polyserial correlation for ordinal and continuous variables: }
 		Under the assumption that the ordinal variable is generated by

### 2025-02-01 paper/main.tex
Updates

@@
 
-While there has been some effort in
-DAG literature as well towards allowing for expert knowledge but these methods
-focus towards providing the expert knowledge to the automated algorithms
-\citep{Bang2023, Meek1995}.
-
 Despite the availability of many automated algorithms for constructing DAGs,
@@
 
-Fig.~\ref{fig:intro} highlights some of these issues using the Adult income
-dataset \citep{Becker1996}. As can be seen in the figure depending on the
-sample size and the algorithm used, we get vastly different results.
+Fig.~\ref{fig:intro} highlights some of these issues. To deal with these issues
+methods for integrating expert knowledge have been developed. The two main
+methods to integrate expert knowledge in the model construction process is to:
+1. Provide all the expert knowledge to the causal discovery algorithm for
+example in the form of required and forbidden edges \citep{Meek1995}, or
+temporal information \citep{Bang2023} and let the algorithm use it to construct
+the model. 2. Utilize conditional independence (CI) testing to identify issues in
+the model, and manually fix them based on expert knowledge \citep{Ankan2021}.
+We focus on this second approach in this paper.
 
-As a result of these challenges users have relied on model testing approaches
-along with their expert knowledge to construct these models. The other ways is
-to do post hoc modification to the learned model using implied CI based model
-testing \citep{Ankan2021, Ankan2023}. In this testing procedure, users test
-whether the implied CIs of the DAG hold in the dataset or not using statistical
-CI tests. Based on the results of these statistical tests, we can accordingly
-make modifications to the current model to fix the violations. However, this CI
-based testing also has a few drawbacks: 1) CIs are implied only by missing
-edges in the model, so we can not test for already existing edges. 2) In high
-sample size scenarios, p-values can suggest a significant association between
-variables even if they have a very weak relationship. 3) Number of implied CIs
-can be large making it difficult to manually go through each of them; for
-example, in the moderately sized alarm network with $ 37 $ node, there are
-$287$ implied CIs.
+In this CI testing procedure, users test whether the implied CIs of the DAG
+hold in the dataset or not using statistical CI tests. Based on the results of
+these statistical tests, we can accordingly make modifications to the current
+model to fix the violations. However, this CI based testing also has a few
+drawbacks: 1) CIs are implied only by missing edges in the model, so we can not
+test for already existing edges. 2) In high sample size scenarios, p-values can
+suggest a significant association between variables even if they have a very
+weak relationship. 3) Number of implied CIs can be large making it difficult to
+manually go through each of them; for example, in the moderately sized alarm
+network with $ 37 $ node, there are $287$ implied CIs.
 
-There has been work on integrating expert knowledge into model building process
-and has been quite successful such as tpc and background knowledge
-specification. There has been work on trying to come up with a single test to
-combine all these test to give a single answer \citep{Eulig2023, Shipley2000}.
-However, what is missing is the guide on how to how to fix these tests.
-Additionally there has been a lot of work to integrate expert knowledge into
-causal discovery algorithms which has shown significant improvement. To tackle
-these issues with CI based testing, we take inspiration from \emph{modification
+There has been work on trying to come up with a single test to combine all
+these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
+what is missing is the guide on how to how to fix these tests. Additionally
+there has been a lot of work to integrate expert knowledge into causal
+discovery algorithms which has shown significant improvement. To tackle these
+issues with CI based testing, we take inspiration from \emph{modification
 indices} in specification search. The modification indices method tries to
@@
 	\includegraphics[page=1]{figures.pdf}
-	\caption{PC algorithm $ \chi^2 $ test. $ N=400$ }
+	\caption{$ N = 400 $}
 \end{subfigure}%
@@
 	\includegraphics[page=2]{figures.pdf}
-	\caption{PC algorithm $ \chi^2 $ test. $ N=800$ }
+	\caption{$ N = 800$ }
 \end{subfigure}
@@
 	\includegraphics[page=5]{figures.pdf}
-	\caption{Hill Climb Search BIC Score. $ N=400$ }
+	\caption{$ N = 400 $ }
 \end{subfigure}%
@@
 	\includegraphics[page=6]{figures.pdf}
-	\caption{Hill Climb Search BIC Score. $ N=800$ }
+	\caption{ $ N = 800$ }
 \end{subfigure}
@@
 	\includegraphics[page=3]{figures.pdf}
-	\caption{PC algorithm residualization test. $ N=400$ }
+	\caption{ $ N = 400$ }
 \end{subfigure}%
@@
 	\includegraphics[page=4]{figures.pdf}
-	\caption{PC algorithm residualization test. $ N=800$ }
+	\caption{$ N = 800$ }
 \end{subfigure}
 
-\caption{Markov Equivalence Class (MEC) learned for Adult Income Dataset using (a) PC algorithm with chi-squared test, (b) Hill-Climb Search with BIC score, (c) PC algorithm with residualization based test.
-There is significant variation in the learned model depending on the sample size and the algorithm used.
-Additionally, all of them make some mistakes.}
+\caption{A comparison of Markov Equivalence Classes (MEC) learned using Adult Income Dataset \citep{Becker1996} using (a \& b) PC algorithm using a mutual information based test, (c \& d) Hill Climb Search with Bayesian Information Criterion (BIC) score, (e \& f) PC algorithm with a residualization based test \citep{Ankan2023}.
+The learned model structure varies significantly depending on the algorithm, the CI test, and the sample size used.}
 \label{fig:intro}

### 2025-02-01 paper/main.tex
Updates

@@
 optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
-\citep{Bello2022}. In contrast, the SEM literature has focused on developing
+\citep{Bello2022}. While DAG-based methods focus on automated discovery,
+SEM-based methods emphasize expert-driven model specification. This includes
 tools to assist researchers in manually constructing or modifying models,
@@
 
-Despite the availability of many automated algorithms for constructing DAGs,
-their adoption in applied research remains limited. Researchers often prefer to
-manually construct DAGs based on their domain expertise \citep{Tennant2020,
-Petersen2021}. We believe this preference persists due to several challenges
-with these algorithms when applied in practical settings:
+Despite significant progress in automated causal discovery, their adoption in
+applied research has been limited. Researchers often prefer to manually
+construct DAGs based on their domain expertise \citep{Tennant2020,
+Petersen2021}. We attribute this preference to several challenges with existing
+causal discovery algorithms in practical settings:
 
 \begin{enumerate}
-	\item \textbf{Lack of Trust:} While most algorithms are
-		asymptotically consistent, their finite sample properties are
-		not well understood \todo{Is there any citation for this}.
-		Additionally, their output can vary significantly depending on
-		the chosen algorithm and their hyperparameters. Combined with a
-		lack of robust methods to test these algorithms' performance on
-		a given dataset, it can be difficult to trust their outputs. As
-		a result, even when using these algorithms are used,
-		researchers often rely on post hoc manual modifications to the
-		model based on model testing and their domain expertise.
+	\item \textbf{Lack of Trust:} While most algorithms are asymptotically
+		consistent, their behavior on finite samples is not well
+		understood. Additionally, their output can vary significantly
+		depending on the choice of algorithm and hyperparameters,
+		making it difficult to assess reliability. The absence of
+		robust performance evaluation methods further reduce the
+		confidence in their outputs. As a result, even when using these
+		algorithms are used, researchers often rely on post hoc manual
+		modifications to the model based on model testing and their
+		domain expertise.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} As multiple
@@
 
-There has been work on trying to come up with a single test to combine all
-these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
-what is missing is the guide on how to how to fix these tests. Additionally
-there has been a lot of work to integrate expert knowledge into causal
-discovery algorithms which has shown significant improvement. To tackle these
-issues with CI based testing, we take inspiration from \emph{modification
-indices} in specification search. The modification indices method tries to
-identify and rank changes in the model such that modifications that improve the
-model fit the most are ranked higher.This ranking allows researchers to focus
-first on the suggestions that can improve the model the most and while
-integrating their expertise to choose the most optimal modifications. Analogous
-to this, in this paper we develop a measure for partial association for mixed
-data that allows us to rank the CI violations. Additionally, this measure of
-conditional association also allows us tackle other issues with CI based
-testing outlined in the previous paragraphs.
+% There has been work on trying to come up with a single test to combine all
+% these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
+% what is missing is the guide on how to how to fix these tests. Additionally
+% there has been a lot of work to integrate expert knowledge into causal
+% discovery algorithms which has shown significant improvement.
+
+To tackle these issues with CI based testing, we take inspiration from
+\emph{modification indices} in specification search. The modification indices
+method tries to identify and rank changes in the model such that modifications
+that improve the model fit the most are ranked higher.This ranking allows
+researchers to focus first on the suggestions that can improve the model the
+most and while integrating their expertise to choose the most optimal
+modifications. Analogous to this, we develop a method for ranking CI test
+violations. This ranking method utilizes both p-values and a measure of
+association that we develop in this paper. Using this ranking we present a
+procedure for iteratively modifying the DAG while integrating expert knowledge
+into it.
 

### 2025-02-04 paper/main.tex
Updates

@@
 
-Using this mixed data partial association measure, we propose a method to rank
-CI tests, and use it for model construction and modification. Given a
-potentially empty DAG $ G $ and a dataset $ D $, for each pair for variable $ X
-$ and $ Y $ in $ G $:
+Using this mixed data partial association measure, we now propose a method for
+construcing DAGs while incorporating expert knowledge iteratively. Given a
+potentially empty DAG $ G $ and a dataset $ D $, a p-value threshold of $
+\alpha $ and a measure of association threshold of $ \beta $, the construction
+procedure works as follows:
+
+\begin{enumerate}
+	\item For each pair of variables $ X $ and $ Y $ in the model:
 		\begin{itemize}
-	\item If there is no edge between $ X $ and $ Y $, we compute p-value for the CI test $ X \ci Y \rvert pa_G(X) \cup pa_G(Y) $ and the partial association measure: $ \rho(X, Y; pa_G(X) \cup pa_G(Y)) $.
-	\item If there is an edge between $ X $ and $ Y $, we compute p-value for the CI test $ X \ci Y \rvert pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $ and the partial association measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $.
+			\item If there is an edge between $ X $ and $ Y $, remove the edge to obtain the DAG $ \underline{G} $, and 
+				compute the measure of association $ \phi(X, Y \rvert \underline{G}_{pa}(X) \cup \underline{G}_{pa}(Y)) $
+			\item If there is no edge between $ X $ and $ Y $, compute the measure of association $ \phi(X, Y \rvert G_{pa}(X) \cup \underline{G}_{pa}(Y) $.
 		\end{itemize}
+	\item If for any pair $ X $ and $ Y $, if there is an edge between them with p-value $ > \alpha $ and $ \phi < \beta $, remove the edge between them.
+	\item For the remaining $ X $ and $ Y $ pairs with p-value $ \geq \alpha $, sort them by $ \beta $.
+Select the pair with highest $ \beta $, ask the user to specify the direction between $ X $ and $ Y $ and add the edge to $ G $.
+\end{enumerate}
 

### 2025-02-04 paper/main.tex
udpates

@@
 \section{Conclusions}
+\begin{itemize}
+	\item As researchers prefer to anyways build models by hand, this is like a companion tool to help them in deciding the method.
+	\item Solves all the $ 3 $ issues outlined in the introduction.
+	\item Compared to other expert knowledge specification methods, users don't need to guess what information then need to provide as exact questions are asked. Also enables us to ask an LLM for it.
+\end{itemize}
+
 Future work:

### 2025-02-04 paper/main.tex
Updates

@@
 partial association for mixed data to rank violations of implied CIs in a given
-DAG, enabling users to prioritize modifications while utilizing their expertise
-when making these modifications. Empirical results show that this manual model
-construction approach achieves performance comparable to automated algorithms
-even if the user is able to decide the orientation of an edge correctly in one
-out of three cases. Additionally, we demonstrate that, in the absence of an
-expert, a Large Language Model (LLM) can potentially determine edge orientations.
+DAG, enabling users to prioritize modifications while utilizing their domain
+knowledge to decide the orientation of new edges. Empirical results show that
+this manual model construction approach achieves performance comparable to
+automated algorithms even if the user is able to decide the edge orientation
+correctly in one out of three cases. Additionally, we demonstrate that, in the
+absence of an expert, a Large Language Model (LLM) can be potentially determine
+edge orientations. We provide the implementation of our approach in both a web 
+tool at: <redacted for review> and a Python package <redacted for review>.
 

### 2025-02-04 paper/main.tex
Updates

@@
 SEM-based methods emphasize expert-driven model specification. This includes
-tools to assist researchers in manually constructing or modifying models,
-enabling experts to incorporate their domain knowledge in the model building
-process. Researchers typically begin with an initial model based on their
-expertise and then use these tools to guide modifications that improve the
-model's fit to data. This process is commonly known as \emph{Specification
-Search} \citep{Long1983} and uses method such as modification indices, and
-Wald-based tests \citep{Marcoulides2018}.
+tools to assist researchers in manually constructing models, enabling them to
+incorporate their domain knowledge in the model building process. Researchers
+typically begin with an initial model based on their expertise and then use
+these tools to guide modifications that improve the model's fit to data. This
+process is commonly known as \emph{Specification Search} \citep{Long1983} and
+uses method such as modification indices, and Wald-based tests
+\citep{Marcoulides2018}.
 
@@
 		making it difficult to assess reliability. The absence of
-		robust performance evaluation methods further reduce the
-		confidence in their outputs. As a result, even when using these
-		algorithms are used, researchers often rely on post hoc manual
-		modifications to the model based on model testing and their
-		domain expertise.
+		robust performance evaluation methods for any given dataset
+		further reduce the confidence in their outputs. 
+		% As a result,
+		% even when using these algorithms are used, researchers often
+		% rely on post hoc manual modifications to the model based on
+		% model testing and their domain expertise.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} As multiple
-		DAGs can be faithful to a given observational dataset, the
+		DAGs can be faithful to a given observational dataset, automated 
 		algorithms can only recover the MECs. These MECs can contain a
@@
 		methods for downstream tasks, such as identification or causal
-		effect estimation, assume knowledge of a fully oriented DAG. As
-		a result, researchers have to manually orient these edges based
-		on their expert knowledge before using them for analysis.
+		effect estimation, assume knowledge of a fully oriented DAG. 
+		% As
+		% a result, researchers have to manually orient these edges based
+		% on their expert knowledge before using them for analysis.
 \end{enumerate}
@@
 
-\caption{A comparison of Markov Equivalence Classes (MEC) learned using Adult Income Dataset \citep{Becker1996} using (a \& b) PC algorithm using a mutual information based test, (c \& d) Hill Climb Search with Bayesian Information Criterion (BIC) score, (e \& f) PC algorithm with a residualization based test \citep{Ankan2023}.
+\caption{A comparison of Markov Equivalence Class (MEC) learned using Adult Income Dataset \citep{Becker1996} using (a \& b) PC algorithm using a mutual information based test, (c \& d) Hill Climb Search with Bayesian Information Criterion (BIC) score, (e \& f) PC algorithm with a residualization based test \citep{Ankan2023}.
 The learned model structure varies significantly depending on the algorithm, the CI test, and the sample size used.}
@@
 \begin{enumerate}
-	\item We present a measure of (conditional) association for mixed data
+	\item We present a measure of conditional association for mixed data
 		based on canonical correlations. This measure of association
-		generalizes the widely used partial correlation to mixed data
-		setting.
-	\item Using this partial association measure, we present a procedure to
-		rank CI tests. This ranking can be used to prioritize addition
-		or deletion of edges from the model.
-	\item We present both a web-browser based tool and an implementation in
-		the Python package to allow researchers to apply this method to
-		their own datasets.
+		generalizes some of the commonly used special case measures of 
+		association to mixed data.
+	\item Using this conditional association measure, we present a procedure to
+		manually construct DAGs by ranking violation of implied CIs in the 
+		model. This iterative procedure allows users to integrate their
+		domain knowledge in the model construction procedure.
+	\item Lastly, to allow researchers to easily apply this method on their
+		datasets, we present a web-browser based interactive tool as well
+		as implementation in a Python package.
 \end{enumerate}
@@
 The rest of the paper is structured as follows. In
-Section~\ref{sec:background}, we give a background on the measures of
-association for various data types. In Section~\ref{sec:mixed_association}, we
-present our method for computing mixed data association. In
-Section~\ref{sec:modification} and Section~\ref{sec:web}, we present how to use
-this method to make modifications and the web tool, and lastly in
-Section~\ref{sec:empirical}, we show some empirical analysis on how well this
-modification process works.
+Section~\ref{sec:background}, we give a background on the commonly used
+measures of association for various data types. In
+Section~\ref{sec:mixed_association}, we present our generalized measure of
+conditional association for mixed data. Section~\ref{sec:modification} provides
+details on the procedure for using the measure of association for constructing
+DAGs. Section~\ref{sec:web} presents our web-browser based tool and lastly in
+Section~\ref{sec:empirical}, we show some empirical results to compare this
+manual approach to automated algorithms.
 

### 2025-02-04 paper/main.tex
Updates

@@
 
+\todo[inline]{The following three paragraphs need work.}
+
 Fig.~\ref{fig:intro} highlights some of these issues. To deal with these issues
@@
 consider random variables in the mixed data setting where each of the random
-variables can be continuous, categorical, or ordinal unless specified.
+variables can be either of continuous, categorical, or ordinal unless
+specified.
 
@@
 Our proposed method for iterative DAG construction relies on a measure of
-conditional association between variables. We are interested in measuring the
-association between variables $ X $ and $ Y $ when conditioned on a set of
-variables $ \bm{Z} $ which is potentially an empty set, $ \bm{Z} = \emptyset $.
-Depending on the type of $ X $, $ Y $, and $ \bm{Z} $ various measures of
-associations have been used. Typically, these are the effect size measures of
-the conditional independence test $ X \ci Y \rvert \bm{Z} $. Here, we give an
-overview of some of the commonly used measures for different data types.
+conditional association also known as partial association between variables. We
+are interested in measuring the association between variables $ X $ and $ Y $
+when conditioned on a set of variables $ \bm{Z} $ which is potentially an empty
+set, $ \bm{Z} = \emptyset $. Many special case measures of conditional
+association have been used in literature depending on the type of $ X $, $ Y $,
+and $ \bm{Z} $. Typically, these are the effect size measures of the
+conditional independence test $ X \ci Y \rvert \bm{Z} $. In this section, we
+give an overview of some of the commonly used measures for different data
+types.
 

### 2025-02-04 paper/main.tex
Updates

@@
 \subsection{Notations}
+\todo[inline]{Improve this section once all the notations are in place}
 We denote random variables with uppercase letters $ X $, and a set of random
@@
 variables can be either of continuous, categorical, or ordinal unless
-specified.
-
-We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $, conditional
+specified. We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $, conditional
 expectations as $ \mathbb{E}[X | \bm{Z}] $.

### 2025-02-04 paper/main.tex
Updates

@@
 association have been used in literature depending on the type of $ X $, $ Y $,
-and $ \bm{Z} $. Typically, these are the effect size measures of the
-conditional independence test $ X \ci Y \rvert \bm{Z} $. In this section, we
-give an overview of some of the commonly used measures for different data
-types.
+and $ \bm{Z} $. Typically, these are the effect size measures of the CI test $
+X \ci Y \rvert \bm{Z} $. In this section, we give an overview of some of the
+commonly used measures for different data types.
 
 \subsubsection{Both $ X $ and $ Y $ are continuous}
-In the case when both $ X $ and $ Y $ are continuous variables with the
-assumptions of normality and linearly relationship, typically Pearson's
-correction coefficients are used. In the case when $ \bm{Z} = \emptyset $, the
-correlation coefficient can be directly computed. When $ \bm{Z} \neq \emptyset
-$, partial correlations are used. One of the ways to estimate it is to use two
-linear regression models: $ E_X: x \sim \bm{z} $, and $ E_Y: y \sim \bm{z} $.
-Then we take the compute the Pearson's correlation coefficient between the
-residuals of these two regressions: $ R_X = x - E_X(\bm{z}) $, $ R_Y = y -
-E_Y(\bm{z}) $.
+When both $ X $ and $ Y $ are continuous variables, Pearson's correlation
+coefficient is typically used.  In the case when $ \bm{Z} = \emptyset $, the
+correlation coefficient is:
+
+$$ \rho_{X, Y} = \frac{cov(X, Y)}{\sigma_X \sigma_Y} $$
+
+When $ \bm{Z} \neq \emptyset $, partial correlation coefficient can be used. 
+One of the ways to estimate it is to use two linear regression models:
+
+\begin{equation}
+	\begin{split}
+		E_X &: X \sim \bm{Z} \\
+		E_Y &: Y \sim \bm{Z} \\
+		\rho_{X, Y; \bm{Z}} &= \frac{cov((X - E_X), (Y - E_Y))}{\sigma_{(X - E_X)} \sigma_{(Y-E_Y)}}
+	\end{split}
+\end{equation}
+
+$ E_X: x \sim \bm{z} $,
+and $ E_Y: y \sim \bm{z} $. Then we take the compute the Pearson's correlation
+coefficient between the residuals of these two regressions: $ R_X = x -
+E_X(\bm{z}) $, $ R_Y = y - E_Y(\bm{z}) $.
 
@@
 
-\section{Connection to Score Based Methods}
+\subsection{Connection to Score Based Methods}
 \begin{itemize}

### 2025-02-04 paper/main.tex
Updates paper

@@
 practical issues with automated algorithms, such as they make obvious errors
-and out Markov Equivalence Classes (MECs). To address these challenges, we
+and output Markov Equivalence Classes (MECs). To address these challenges, we
 propose an iterative method to assist researchers in manually constructing or
@@
 this manual model construction approach achieves performance comparable to
-automated algorithms even if the user is able to decide the edge orientation
-correctly in one out of three cases. Additionally, we demonstrate that, in the
-absence of an expert, a Large Language Model (LLM) can be potentially determine
-edge orientations. We provide the implementation of our approach in both a web 
-tool at: <redacted for review> and a Python package <redacted for review>.
+automated algorithms if the user is able correctly orientation edges in one out
+of three cases. Additionally, we demonstrate that, in the absence of an expert,
+a Large Language Model (LLM) can be potentially determine edge orientations. We
+provide the implementation of our approach in both a web tool at: <redacted for
+review> and a Python package <redacted for review>.
 
@@
 		further reduce the confidence in their outputs. 
-		% As a result,
-		% even when using these algorithms are used, researchers often
-		% rely on post hoc manual modifications to the model based on
-		% model testing and their domain expertise.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} As multiple
@@
 		effect estimation, assume knowledge of a fully oriented DAG. 
-		% As
-		% a result, researchers have to manually orient these edges based
-		% on their expert knowledge before using them for analysis.
 \end{enumerate}
 
-\todo[inline]{The following three paragraphs need work.}
+Fig.~\ref{fig:intro} highlights some of these issues. When constructing models
+manually it is important to test whether the constructed DAG is representative
+of the dataset or not. Based on this testing and domain knowledge further
+modifications can be made to the DAG. One of the approaches to test and modify
+DAGs is by testing whether the implied CI tests of the DAG are true in the data
+or not \citep{Ankan2021}. If some of the tests are violated in the data,
+researchers can use their domain knowledge to make modifications to the DAG.
 
-Fig.~\ref{fig:intro} highlights some of these issues. To deal with these issues
-methods for integrating expert knowledge have been developed. The two main
-methods to integrate expert knowledge in the model construction process is to:
-1. Provide all the expert knowledge to the causal discovery algorithm for
-example in the form of required and forbidden edges \citep{Meek1995}, or
-temporal information \citep{Bang2023} and let the algorithm use it to construct
-the model. 2. Utilize conditional independence (CI) testing to identify issues in
-the model, and manually fix them based on expert knowledge \citep{Ankan2021}.
-We focus on this second approach in this paper.
+This CI testing
+
+% To deal with these issues
+% methods for integrating expert knowledge have been developed. The two main
+% methods to integrate expert knowledge in the model construction process is to:
+% 1. Provide all the expert knowledge to the causal discovery algorithm for
+% example in the form of required and forbidden edges \citep{Meek1995}, or
+% temporal information \citep{Bang2023} and let the algorithm use it to construct
+% the model. 2. Utilize conditional independence (CI) testing to identify issues in
+% the model, and manually fix them based on expert knowledge \citep{Ankan2021}.
+% We focus on this second approach in this paper.
 
@@
 commonly used measures for different data types.
+
+\todo[inline]{Remove or move equations to appendix if no space}
 
@@
 \label{sec:web}
-Based on the procedure outlined in the previous section, we built a web-based
-tool that allows us to interactively create/modify their model. The user
-starts by uploading their dataset and optionally the initial DAG. If no initial
-DAG is provided an empty graph is initialized. The tool also computes the
-Shipley's C value at any point. The web-tool would automatically show all
-unexplained associations in the DAG by adding red edges between them. The edge
-weight of the edge represents how high this strength is. Users can also set the
-threshold for p-value and association to control how many edges are shown at a
-time. Based on these shown potential edges, the user can select the edge that
+
+To allow users to apply this method to their own dataset, we also built an
+interactive web browser based tool to allow them to interactively construct
+models. The user starts by uploading their dataset which initializes an empty
+DAG on the variables in the dataset. Users can then set the desired p-value
+and measure of association threshold. The tools then shows a red edge between
+all the variables that are correlated in the data but the correlation is not
+explained by the current DAG. The thickness of the edge represents the strength
+of this correlation, which would help the user in prioritizing edges to add.
+
+The tool also computes the
+Shipley's C value at any point.
+
+The edge
+weight of the edge represents how high this strength is. 
+Based on these shown potential edges, the user can select the edge that
 they want to add and decide the direction for it. If the tool detects an

### 2025-02-04 paper/main.tex
Updates

@@
 
-This CI testing
+However, this CI testing approach also has a few drawbacks: 1) CIs are implied
+only by missing edges in the model, so we can not test whether the existing
+edges are correct or not. 2) The decision whether a CI holds in the data or not
+is based on a p-value threshold. But we can get significant p-values for even
+very weak relationship if the sample size is high. 3) Lastly, the number of
+implied CIs can be large making it difficult to go through all of them; for 
+example, in the moderately sized alarm network \citep{Beinlich1989} with $ 37 $
+nodes, there are $ 287 $ implied CIs.
 
@@
 
-In this CI testing procedure, users test whether the implied CIs of the DAG
-hold in the dataset or not using statistical CI tests. Based on the results of
-these statistical tests, we can accordingly make modifications to the current
-model to fix the violations. However, this CI based testing also has a few
-drawbacks: 1) CIs are implied only by missing edges in the model, so we can not
-test for already existing edges. 2) In high sample size scenarios, p-values can
-suggest a significant association between variables even if they have a very
-weak relationship. 3) Number of implied CIs can be large making it difficult to
-manually go through each of them; for example, in the moderately sized alarm
-network with $ 37 $ node, there are $287$ implied CIs.
+% In this CI testing procedure, users test whether the implied CIs of the DAG
+% hold in the dataset or not using statistical CI tests. Based on the results of
+% these statistical tests, we can accordingly make modifications to the current
+% model to fix the violations. However, this CI based testing also has a few
+% drawbacks: 1) CIs are implied only by missing edges in the model, so we can not
+% test for already existing edges. 2) In high sample size scenarios, p-values can
+% suggest a significant association between variables even if they have a very
+% weak relationship. 3) Number of implied CIs can be large making it difficult to
+% manually go through each of them; for example, in the moderately sized alarm
+% network with $ 37 $ node, there are $287$ implied CIs.
 
@@
 method tries to identify and rank changes in the model such that modifications
-that improve the model fit the most are ranked higher.This ranking allows
+that improve the model fit the most are ranked higher. This ranking allows
 researchers to focus first on the suggestions that can improve the model the
 most and while integrating their expertise to choose the most optimal
-modifications. Analogous to this, we develop a method for ranking CI test
-violations. This ranking method utilizes both p-values and a measure of
-association that we develop in this paper. Using this ranking we present a
-procedure for iteratively modifying the DAG while integrating expert knowledge
-into it.
+modifications. Analogous to this, we develop a procedure for ranking CI test
+violations allowing researchers to focus on the worst violations in the model.
+This ranking method utilizes both p-values and a measure of conditional
+association. Utilizing this procedure researchers can iteratively construct
+DAGs while integrating their domain knowledge in the model building procedure.
 
-\begin{figure}
+\begin{figure}[t!]
 \begin{subfigure}{0.25 \textwidth}
@@
 \begin{enumerate}
-	\item We present a measure of conditional association for mixed data
+	\item We propose a novel measure of conditional association for mixed data
 		based on canonical correlations. This measure of association
@@
 Section~\ref{sec:empirical}, we show some empirical results to compare this
-manual approach to automated algorithms.
+manual DAG construction approach to automated algorithms.
 
@@
 
-\begin{figure}
+\begin{figure}[t!]
 	\begin{subfigure}{0.17 \textwidth}
@@
 
-\begin{figure}
+\begin{figure}[t!]
 	\centering
@@
 
-\begin{figure}
+\begin{figure}[t!]
 	\begin{subfigure}{0.25\textwidth}

### 2025-02-04 paper/main.tex
Updates

@@
 coefficient is typically used. In the case when $ \bm{Z} = \emptyset $, the
-correlation coefficient is:
+correlation coefficient is given as:
 
-$$ \rho_{X, Y} = \frac{cov(X, Y)}{\sigma_X \sigma_Y} $$
+\begin{equation}
+\rho_{X, Y} = \frac{\mathrm{cov}(X, Y)}{\sigma_X \sigma_Y}
+\end{equation}
 
-When $ \bm{Z} \neq \emptyset $, partial correlation coefficient can be used. 
-One of the ways to estimate it is to use two linear regression models:
+where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, and $
+\sigma_X $ and $ \sigma_Y $ are the standard deviations of $ X $ and $ Y $
+respectively. When $ \bm{Z} \neq \emptyset $, partial correlation coefficient
+can be used. One of the ways to estimate it is to use two linear regression
+models:
 
@@
 
-$ E_X: x \sim \bm{z} $,
-and $ E_Y: y \sim \bm{z} $. Then we take the compute the Pearson's correlation
-coefficient between the residuals of these two regressions: $ R_X = x -
-E_X(\bm{z}) $, $ R_Y = y - E_Y(\bm{z}) $.
+$ E_X: x \sim \bm{z} $, and $ E_Y: y \sim \bm{z} $. Then we take the compute
+the Pearson's correlation coefficient between the residuals of these two
+regressions: $ R_X = x - E_X(\bm{z}) $, $ R_Y = y - E_Y(\bm{z}) $.
 
@@
 
-As there are no generalized measures of association for mixed data, in this
-section, we first introduce a novel measure of partial association for mixed
-data by extending the concept of partial correlations \citep{Whittaker2009}
-\todo{Verify this citation} to handle mixed data types. To compute partial
-correlations, we have two key steps: 1) a residualization method, and 2) a
-measure of associations between the residuals. When both $ X $ and $ Y $ are
-continuous variables, residuals are typically computed using a regression-based
-method, and Pearson's correlation coefficient is used to measure the
-association between these residuals. To extend this approach to mixed data, we
-combine a mixed data residualization method \citep{Ankan2023} with a
-multivariate measure of association based on canonical correlations, Pillai's
-Trace \citep{Pillai1955}.
+As we saw in the previous section, depending on the type of variable, different
+conditional measures of association are used. However, no generalized measure
+of association exists for mixed data. In this section, we present a novel
+measure of partial association for mixed data. This measure of association is
+analogous to partial correlation coefficient where a residualization method is
+combined with Pearson correlation coefficient. Here, we use a mixed data
+residualization method \citep{Ankan2023} and combine it with a multivariate
+measure of association based on canonical correlations called Pillai's Trace
+{Pillai1955}.
+
+% As there are no generalized measures of association for mixed data, in this
+% section, we first introduce a novel measure of partial association for mixed
+% data by extending the concept of partial correlations \citep{Whittaker2009}
+% \todo{Verify this citation} to handle mixed data types. To compute partial
+% correlations, we have two key steps: 1) a residualization method, and 2) a
+% measure of associations between the residuals. When both $ X $ and $ Y $ are
+% continuous variables, residuals are typically computed using a regression-based
+% method, and Pearson's correlation coefficient is used to measure the
+% association between these residuals. To extend this approach to mixed data, we
+% combine a mixed data residualization method \citep{Ankan2023} with a
+% multivariate measure of association based on canonical correlations, Pillai's
+% Trace \citep{Pillai1955}.
 
@@
 \bm{Z} $, we aim to estimate the conditional association $ \rho(X, Y; \bm{Z}) $
-between $ X $ and $ Y $ conditioned on $ \bm{Z} $. We use the residualization
-approach from \citet{Ankan2023} to compute the residuals. For each of the
-variable $ X $ and $ Y $, we compute the residual $ R_X $ and $R_Y$
-respectively as follows depending on their type.
+between $ X $ and $ Y $ conditioned on $ \bm{Z} $. We first use the
+residualization approach from \citet{Ankan2023} to compute the residual
+matrices for both $ X $ and $ Y $. For each of the variable $ X $ and $ Y $, we
+compute the residual $ R_X $ and $R_Y$ respectively as follows depending on
+their type.
 
@@
 
-We repeat the above step for the variable $ Y $ and compute its residual $ R_Y
-$. Here, we have the option to choose the estimators based on the properties of
-the data. Non-parametric ensemble estimators such as Random Forest and XGBoost
-can handle any combination of data types while performing reasonably well. For
-linear datasets, we can also choose variants of linear regression depending on
-the type of data.
+We repeat the above steps for both the variables $ X $ and  $ Y $ to get
+residuals $ R_X $ and $ R_Y $. Here, we have the option to choose the
+estimators based on the properties of the data. Non-parametric ensemble
+estimators such as Random Forest and XGBoost can handle any combination of data
+types while performing reasonably well. For linear datasets, we can also choose
+variants of linear regression depending on the type of data.
 
@@
 \begin{itemize}
-	\item Wilks' Lambda:
-		$$ \Lambda = \prod_{i}^{\min(p, q)} (1 - \phi_i^2) $$
-	\item Roy's Largest Root: 
-		$$ \theta = \max_i(\phi_i^2) $$
-	\item Pillai's Trace: 
-		$$ \tau = \sum_{i=1}^{\min(p, q)} \phi_i^2 $$
+	\item Wilks' Lambda: $\Lambda = \prod_{i}^{\min(p, q)} (1 - \phi_i^2) $
+	\item Roy's Largest Root: $ \theta = \max_i(\phi_i^2) $
+	\item Pillai's Trace: $ \tau = \sum_{i=1}^{\min(p, q)} \phi_i^2 $
 \end{itemize}

### 2025-02-04 paper/main.tex
Updates

@@
 
-As we saw in the previous section, depending on the type of variable, different
-conditional measures of association are used. However, no generalized measure
-of association exists for mixed data. In this section, we present a novel
-measure of partial association for mixed data. This measure of association is
-analogous to partial correlation coefficient where a residualization method is
-combined with Pearson correlation coefficient. Here, we use a mixed data
-residualization method \citep{Ankan2023} and combine it with a multivariate
-measure of association based on canonical correlations called Pillai's Trace
-{Pillai1955}.
+In the previous section, we discussed how different types of variables require
+different conditional measures of association. However, no unified measure
+exists for mixed data. In this section, we introduce a novel measure of partial
+association for mixed data, by extending the idea of partial correlation
+coefficient to mixed data. Specifically, our approach combines a mixed data
+residualization method \citep{Ankan2023} with Pillai's Trace
+\citep{Pillai1955}, a multivariate measure of association based on canonical
+correlations.
 
@@
 \label{sec:web}
+To enable users to apply this method to their own datasets, we developed an
+interactive web-based tool (shown in Fig.~\ref{fig:web}) for constructing
+models. Users begin by uploading their dataset, which initializes an empty DAG
+with nodes corresponding to the dataset\'s variables. They can then specify a
+p-value threshold and a measure of association threshold.
 
-To allow users to apply this method to their own dataset, we also built an
-interactive web browser based tool to allow them to interactively construct
-models. The user starts by uploading their dataset which initializes an empty
-DAG on the variables in the dataset. Users can then set the desired p-value
-and measure of association threshold. The tools then shows a red edge between
-all the variables that are correlated in the data but the correlation is not
-explained by the current DAG. The thickness of the edge represents the strength
-of this correlation, which would help the user in prioritizing edges to add.
-
-The tool also computes the
-Shipley's C value at any point.
-
-The edge
-weight of the edge represents how high this strength is. 
-Based on these shown potential edges, the user can select the edge that
-they want to add and decide the direction for it. If the tool detects an
-incorrect edge in the graph, it is shown in black. Once the user is happy with
-the graph they have they can export it into different formats to use it for
-further computation.
+Using the specified thresholds, the tool visually highlights unexplained
+correlations by displaying red edges between variables where correlation exists
+in the data but the current DAG is not able to explain. The thickness of these
+edges represents the strength of the correlation, helping users prioritize
+which edges to add. Similarly, if an edge in the graph is detected to be
+unnecessary, it is highlighted in black. Based on this information, users can
+select to remove unnecessary edge or select a potential edge to add and specify
+the orientation of the edge. The tool computes Shipley’s C \citep{Shipley2000}
+value at each change to determine the overall fit of the model to the data.
+Once satisfied with the constructed DAG, users can export the model for further
+analysis.
 
@@
 		black.}
+	\label{fig:web}
 \end{figure}

### 2025-02-04 paper/appendix.tex
Updates

\onecolumn
\title{Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge \\ (Supplementary Material)}
\maketitle
\appendix

\section{Prompts Used for LLM}

\begin{figure}[ht!]
	\centering
	\begin{Verbatim}
You are an expert in Social Science. Following are the descriptions of two variables:

<A>: {description of variable A}
<B>: {description of variable B}

Which of the following two options is the most likely causal direction between these 
variables:

1. <A> causes <B>
2. <B> causes <A>

Return a single letter answer between the choices above; Do not provide any reasoning 
in the answer; Do not add any text formatting to the answer.
	\end{Verbatim}
	\caption{Prompt used for the LLM. Here the variable descriptions are replaced with description provided in Fig.~\ref{fig:var_description}}
	\label{fig:prompt}
\end{figure}

\begin{figure}[ht!]
	\begin{Verbatim}
          Age: The age of a person
    Workclass: The workplace where the person is employed such as Private industry, 
     	       or self employed
    Education: The highest level of education the person has finished.
MaritalStatus: The marital status of the person
   Occupation: The kind of job the person does. For example, sales, craft repair, 
   		clerical.
 Relationship: The relationship status of the person.
         Race: The ethnicity of the person.
          Sex: The sex or gender of the person.
 HoursPerWeek: The number of hours per week the person works.
NativeCountry: The native country of the person.
       Income: The income i.e. amount of money the person makes.
	\end{Verbatim}
	\caption{Variable descriptions used for prompting the LLM}
	\label{fig:var_description}
\end{figure}

### 2025-02-04 paper/main.tex
Updates

@@
 	\centering
-	\begin{subfigure}{0.5\textwidth}
-		\begin{Verbatim}[fontsize=\footnotesize]
-You are an expert in Social Science. Following
-are the descriptions of two variables:
-
-<A>: {description of variable A}
-<B>: {description of variable B}
-
-Which of the following two options is the most 
-likely causal direction between these variables:
-
-1. <A> causes <B>
-2. <B> causes <A>
-
-Return a single letter answer between the 
-choices above; Do not provide any reasoning 
-in the answer; Do not add any text formatting 
-to the answer.
-		\end{Verbatim}
-	\caption{Prompt used for the LLM}
-	\end{subfigure}
-	\begin{subfigure}{0.5 \textwidth}
-		\begin{Verbatim}[fontsize=\footnotesize]
-Age: The age of a person
-Workclass: The workplace where the person is 
-     employed such as Private industry, 
-     or self employed
-Education: The highest level of education the 
-     person has finished
-MaritalStatus: The marital status of the person
-Occupation: The kind of job the person does. 
-     For example, sales, craft repair, clerical.
-Relationship: The relationship status of 
-     the person
-Race: The ethnicity of the person
-Sex: The sex or gender of the person
-HoursPerWeek: The number of hours per week the 
-     person works
-NativeCountry: The native country of the person
-Income: The income i.e. amount of money the 
-     person makes
-		\end{Verbatim}
-	\caption{Variable descriptions used for prompting the LLM}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-		\centering
 	\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
 	\caption{The learned DAG}
-	\end{subfigure}
 \end{figure}
@@
 \section{Conclusions}
+Points:
 \begin{itemize}
@@
 
+\bibliography{references}
+
 \newpage
-\bibliography{references}
+
+\include{appendix}
 \end{document}

### 2025-02-04 paper/main.tex
Uupdates

@@
 
-\subsection{Using Large Language Models as Experts}
-A lot of recent work has focused on finding pairwise causal direction among
-variables using Large Language Models. In this analysis, instead of using the
-oracle to decide the direction of edges, we used GPT4 to give us the direction
-of the edges based on description of the variables. As GPT4 has already
-memorized many of these models, we encoded the variable names using a random
-string and provided the description of the variable to let it decide the
-direction of the edge.
+\subsection{Using LLMs as Experts}
+Recently, there has been a lot of work towards utilizing LLMs for causal
+discovery including approaches ranging from determining pairwise edge
+orientation \citep{Kiciman2023, Jin2024} to full causal discovery
+\citep{Naik2023, Vashishtha2023}, and counterfactual reasoning
+tasks\citep{Kiciman2023} (see \citet{Liu2024} for an overview) As our approach
+utilizes expert knowledge for determining the orientation of the edges, this
+opened up an opportunity to use an LLM for determining this. We tried applying
+our procedure for causal discovery with using GEMINI as the expert. We used a
+greedy approach where the pair of variables with the highest unexplained
+correlation is oriented first. The prompt used to determine the orientation is
+provided in Appendix~\ref{section:llms}. Our LLM prompt utilizes variable
+description to determine the edge orientation. We applied this approach to the
+adult income dataset and the output is shown in Fig.~\ref{fig:adult_llm}.
 
@@
 	\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
-	\caption{The learned DAG}
+	\caption{DAG learned using the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert.}
+	\label{fig:adult_llm}
 \end{figure}

### 2025-02-04 paper/main.tex
Updates

@@
 \subsection{Using LLMs as Experts}
-Recently, there has been a lot of work towards utilizing LLMs for causal
-discovery including approaches ranging from determining pairwise edge
-orientation \citep{Kiciman2023, Jin2024} to full causal discovery
-\citep{Naik2023, Vashishtha2023}, and counterfactual reasoning
-tasks\citep{Kiciman2023} (see \citet{Liu2024} for an overview) As our approach
-utilizes expert knowledge for determining the orientation of the edges, this
-opened up an opportunity to use an LLM for determining this. We tried applying
-our procedure for causal discovery with using GEMINI as the expert. We used a
-greedy approach where the pair of variables with the highest unexplained
+Recently, there has been a lot of work towards using LLMs for causal discovery,
+with tasks ranging from determining pairwise edge orientation
+\citep{Kiciman2023, Jin2024} to full causal structure learning \citep{Naik2023,
+Vashishtha2023} and counterfactual reasoning tasks\citep{Kiciman2023} (see
+\citet{Liu2024} for an overview).
+
+Since our approach utilizes expert knowledge to determine edge orientations, we
+investigated the potential of using an LLM for this task. We applied our causal
+discovery procedure with using Gemini 1.5 flash model as the expert. Our method
+uses a greedy approach where the pair of variables with the highest unexplained
 correlation is oriented first. The prompt used to determine the orientation is
@@
 adult income dataset and the output is shown in Fig.~\ref{fig:adult_llm}.
+
+Unlike the other approaches to utilize LLM for doing full causal discovery, we 
+used a combination of statistical tests and LLMs. This greatly reduces the
+amount of information that we need from the LLM. Our approach is also able to 
+ask exact questions about the variables unlike some of the other approaches.
 

### 2025-02-04 paper/main.tex
Adds tikz file for figure 5
Updates

@@
 \begin{enumerate}
-	\item We propose a novel measure of conditional association for mixed data
-		based on canonical correlations. This measure of association
-		generalizes some of the commonly used special case measures of 
-		association to mixed data.
-	\item Using this conditional association measure, we present a procedure to
-		manually construct DAGs by ranking violation of implied CIs in the 
-		model. This iterative procedure allows users to integrate their
-		domain knowledge in the model construction procedure.
-	\item Lastly, to allow researchers to easily apply this method on their
+	\item We propose a novel measure of conditional association for mixed
+		data based on canonical correlations
+		(Section~\ref{sec:mixed_association}. This measure of
+		association generalizes some of the commonly used special case
+		measures of association to mixed data.
+	\item Using this conditional association measure, we present a
+		procedure to manually construct DAGs by ranking violation of
+		implied CIs in the model (Section~\ref{sec:modification}. This
+		iterative procedure allows users to integrate their domain
+		knowledge in the model construction procedure.
+	\item To allow researchers to easily apply this method on their
 		datasets, we present a web-browser based interactive tool as well
-		as implementation in a Python package.
+		as implementation in a Python package (Section~\ref{sec:web}).
+	\item Lastly, we compare the manual DAG construction method with automated
+		causal discovery algorithms (Section~\ref{sec:empirical}).
+
 \end{enumerate}
 
-The rest of the paper is structured as follows. In
-Section~\ref{sec:background}, we give a background on the commonly used
-measures of association for various data types. In
-Section~\ref{sec:mixed_association}, we present our generalized measure of
-conditional association for mixed data. Section~\ref{sec:modification} provides
-details on the procedure for using the measure of association for constructing
-DAGs. Section~\ref{sec:web} presents our web-browser based tool and lastly in
-Section~\ref{sec:empirical}, we show some empirical results to compare this
-manual DAG construction approach to automated algorithms.
+% The rest of the paper is structured as follows. In
+% Section~\ref{sec:background}, we give a background on the commonly used
+% measures of association for various data types. In
+% Section~\ref{sec:mixed_association}, we present our generalized measure of
+% conditional association for mixed data. Section~\ref{sec:modification} provides
+% details on the procedure for using the measure of association for constructing
+% DAGs. Section~\ref{sec:web} presents our web-browser based tool and lastly in
+% Section~\ref{sec:empirical}, we show some empirical results to compare this
+% manual DAG construction approach to automated algorithms.
 
@@
 	\centering
-	\includegraphics[scale=0.3]{../code/fig5/llm_adult.png}
+	\includegraphics[page=1]{fig5.pdf}
 	\caption{DAG learned using the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert.}

### 2025-02-05 paper/main.tex
Updates

@@
 variables can be either of continuous, categorical, or ordinal unless
-specified. We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $, conditional
-expectations as $ \mathbb{E}[X | \bm{Z}] $.
+specified. We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $,
+conditional expectations as $ \mathbb{E}[X | \bm{Z}] $. We denote a graph $ G =
+(V, E) $ which nodes $ V$ and edges $ E $. In this paper we focus on all
+observed variables, and linear measure of association.
 
@@
 Our proposed method for iterative DAG construction relies on a measure of
-conditional association also known as partial association between variables. We
-are interested in measuring the association between variables $ X $ and $ Y $
-when conditioned on a set of variables $ \bm{Z} $ which is potentially an empty
-set, $ \bm{Z} = \emptyset $. Many special case measures of conditional
-association have been used in literature depending on the type of $ X $, $ Y $,
-and $ \bm{Z} $. Typically, these are the effect size measures of the CI test $
-X \ci Y \rvert \bm{Z} $. In this section, we give an overview of some of the
-commonly used measures for different data types.
+conditional association, also commonly known as partial association, between
+variables. Specifically, we are interested in quantifying the association
+between variables $ X $ and $ Y $ when conditioned on a set of variables $
+\bm{Z} $, which may be empty ($ \bm{Z} = \emptyset $). Various measures of
+conditional association have been used in the literature, depending on the type
+of $ X $, $ Y $, and $ \bm{Z} $. These measures are often based on the effect
+size of the CI test $ X \ci Y \rvert \bm{Z} $. In this section, we provide an
+overview of some of the commonly used measures for different data types.
 
@@
 respectively. When $ \bm{Z} \neq \emptyset $, partial correlation coefficient
-can be used. One of the ways to estimate it is to use two linear regression
-models:
+can be used. One of the ways to estimate it is to use two regression
+models $ E_X: X \sim \bm{Z} $, $ E_Y: Y \sim \bm{Z} $, and compute the Pearson's 
+correlation coefficient between the residuals of these regression models:
 
@@
 	\begin{split}
-		E_X &: X \sim \bm{Z} \\
-		E_Y &: Y \sim \bm{Z} \\
-		\rho_{X, Y; \bm{Z}} &= \frac{cov((X - E_X), (Y - E_Y))}{\sigma_{(X - E_X)} \sigma_{(Y-E_Y)}}
+		R_X &= X - E_X(\bm{Z})  \;\;\;\;\;\;\;\;  R_Y = Y - E_Y(\bm{Z}) \\
+		\rho_{X, Y; \bm{Z}} &= \frac{\mathrm{cov}(R_X, R_Y)}{\sigma_{R_X} \sigma_{R_Y}}
 	\end{split}
@@
 
-$ E_X: x \sim \bm{z} $, and $ E_Y: y \sim \bm{z} $. Then we take the compute
-the Pearson's correlation coefficient between the residuals of these two
-regressions: $ R_X = x - E_X(\bm{z}) $, $ R_Y = y - E_Y(\bm{z}) $.
-
 \paragraph{All $ X $, $ Y $, and $ \bm{Z} $ are discrete}
 
-In the case of discrete variables, Cramer\'s V has been typically used as a
-measure of association. When there are no conditional variables, $ \bm{Z} =
-\emptyset $, Cramer\'s V can be computed directly from the contingency table.
+In the case of all discrete variables, Cram\'er's V can used as a measure of
+association. When $ \bm{Z} = \emptyset $, Cram\'er's V can be computed from the 
+chi-squared test statistic as:
 
-However, when conditional variables are present, we can iterate over all
-possible combinations of the conditional variables, compute the Cramer\'s V and
-then lastly combine them together.
+\begin{equation}
+	\mathbb{V_{X, Y}} = \sqrt{\frac{\chi^2 / n}{\mathrm{min}(k-1, r-1)}}
+\end{equation}
+
+where $ \chi^2 $ is the chi-squared test statistic for $ X $ and $ Y $, $ n $ 
+is the sample size and $ k $ and $ r $ are the number of categories of $ X $ and 
+$ Y $.
+
+However, when $ \bm{Z} \neq \emptyset $, we first split the dataset by unique
+values of $ \bm{Z} $, $ D = \{D_{\bm{Z} = \bm{Z_1}}, D_{\bm{Z} = \bm{Z_2}},
+\cdots, \} $  and then combine them as:
+
+\begin{equation}
+	\mathbb{V_{X, Y; Z}} = \sum_{D_{\bm{Z} = \bm{Z_i}}} \mathbb{V}_{X_{D_i}, Y_{D_i}}
+\end{equation}
 
 \paragraph{$ X $ is ordinal and $ Y $ is continuous or ordinal}
-Polyserial Correlation (when one is oridinal and the other is continuous) and
-polychoric correlation (when both are ordinal) can be use. Here the assumption
-is made that $ X $ is a thresholded version of an underlying continuous normal
-variable. The method then tries to learn the thresholds for this latent
-variable while trying to maximize the correlation between $ X_{lat} $ and $ Y
-$. This gives us values that are Pearson's correlation coefficient.
+When one of the variables is ordinal and the other is continuous polyserial
+correlation can be used. When both the variables are ordinal polychoric
+correlation can be used. Both polychoric and polyserial correlation make the
+assumption that the ordinal variable has an underlying normal distributed
+continuous variable, and the ordinal variable has been generated by
+thresholding this latent variable. The method then tries to maximize the
+likelihood of data given polychoric or polyserial correlation coefficient. This
+method gives us values that are Pearson's correlation coefficient.
 
@@
 		coefficient.
-	\item \textbf{Related to Cramer\'s V: } 
+	\item \textbf{Related to Cram\'ers V: } 
 \end{enumerate}

### 2025-02-05 paper/main.tex
Updates

@@
 different conditional measures of association. However, no unified measure
-exists for mixed data. In this section, we introduce a novel measure of partial
-association for mixed data, by extending the idea of partial correlation
-coefficient to mixed data. Specifically, our approach combines a mixed data
-residualization method \citep{Ankan2023} with Pillai's Trace
+exists for mixed data. In this section, we introduce a novel measure of
+conditional association for mixed data by extending the idea of partial
+correlation coefficient to mixed data. Specifically, our approach combines a
+mixed data residualization method \citep{Ankan2023} with Pillai's Trace
 \citep{Pillai1955}, a multivariate measure of association based on canonical
@@
 
-Given $ n $ samples $ D = (x, y, \bm{z}) $ of variables $ X $, $ Y $, and $
-\bm{Z} $, we aim to estimate the conditional association $ \rho(X, Y; \bm{Z}) $
-between $ X $ and $ Y $ conditioned on $ \bm{Z} $. We first use the
-residualization approach from \citet{Ankan2023} to compute the residual
-matrices for both $ X $ and $ Y $. For each of the variable $ X $ and $ Y $, we
-compute the residual $ R_X $ and $R_Y$ respectively as follows depending on
-their type.
+Given a dataset $ D = (x, y, \bm{z}) $ on variables $ X $, $ Y $, and $ \bm{Z}
+$, we aim to estimate the conditional association $ \rho(X, Y; \bm{Z}) $
+between $ X $ and $ Y $ conditioned on $ \bm{Z} $. We first use a mixed data
+residualization approach \citep{Ankan2023} to compute residuals $R_X$ and $R_Y$
+for $ X $ and $ Y $. The method for computing these residuals depend on the 
+type of variable:
 
 \begin{enumerate}
-	\item \textbf{Continuous:} We train a regression model, $ E_X: x \sim
-		\bm{z} $. The residuals are then defined as the difference
+	\item \textbf{Continuous:} We train a model, $ E_X: x \sim
+		\bm{z} $. The residuals are then computed by taking the difference
 		between the true and the predicted values using $ E_X $. 
@@
 	\item \textbf{Ordinal:} We start by training a probability estimator, $
-		p_X: x \sim \bm{z} $, and obtain probability predictions 
-		$ \hat{p}_X: \hat{p}_X(\bm{Z}) $. Then, we compute the residuals
-		as follows \citep{Li2012}:
-		$$ R_{x_i} = \hat{p}(\hat{x}_i > x_i) - \hat{p}(\hat{x}_i < x_i) $$
+		p_X: x \sim \bm{z} $, and then use the estimated probabilities, 
+		$ \hat{p}_X(x) $ to compute the residuals:
+		$$ R_{x_i} = \hat{p}_X(X < x_i) - \hat{p}_X(X > x_i) $$
 	\item \textbf{Categorical:} We again start by training a probability
 		estimator $ p_X: X \sim \bm{Z} $, and obtain probability
-		estimates $ \hat{p}_X: p_Z(\bm{Z}) $. Next, we dummy encode the
+		estimates $ \hat{p}_X: p_X(\bm{Z}) $. Next, we dummy encode the
 		categorical variable and then compute the residuals as follows: 
@@
 
-\todo[inline]{Fix notations above}
-
 We repeat the above steps for both the variables $ X $ and  $ Y $ to get
 residuals $ R_X $ and $ R_Y $. Here, we have the option to choose the
-estimators based on the properties of the data. Non-parametric ensemble
-estimators such as Random Forest and XGBoost can handle any combination of data
-types while performing reasonably well. For linear datasets, we can also choose
-variants of linear regression depending on the type of data.
+estimators based on the properties (distribution, type of relationship, and so
+on) of the data. Non-parametric ensemble estimators such as Random Forest and
+XGBoost can handle any combination of data types while performing reasonably
+well on a wide range of distributions and relationship types. For linear
+relationships, we can also choose simpler models such as linear regression.
 
-After the above steps are repeated for both $ X $ and $ Y $, we end up with two
-residual matrices: $ R_X $ and $ R_Y $. The type of the variable determines the
-shape for these matrices. If the variable is continuous, we get a residual
-matrix of shape $ ( n \times 1 ) $ and if the variable is categorical, we have
-a residual matrix of shape $ ( x \times (k -1)) $ where one of the dummy encoded
-variables is dropped to avoid multicollinearity.
+After the above residualization steps are repeated for both $ X $ and $ Y $, we
+end up with two residual matrices: $ R_X $ and $ R_Y $. The type of the
+variable determines the shape for these matrices. If the variable is continuous
+or ordinal, we get a residual matrix of shape $ ( n \times 1 ) $ and if the
+variable is categorical, we have a residual matrix of shape $ ( x \times (k-1))
+$ where one of the dummy encoded variables is dropped to avoid
+multicollinearity.
 
 We now have two continuous-valued residual matrices. The next step is
-to quantify the association between these residuals. In multivariate
-statistics, canonical correlation based measures have been used to measure
-the association between two sets of variables. We can treat our residuals
-as sets of variables and use canonical correlations to quantify the association
-between them.
+to quantify the association between these residuals. For this purpose
+we use canonical correlations that have been used to measure the association
+between two sets of random variables.
 
@@
 
-	% \begin{equation}
-	% 	\phi(\bm{U}, \bm{V}) = \sup_{a, b} \frac{a^T \Sigma_{\bm{U}\bm{V}} b}{\sqrt{a^T \Sigma_{\bm{U}\bm{U}} a} \sqrt{b^T \Sigma_{\bm{V}\bm{V}} b}}
-	% \end{equation}
-
-% \begin{equation}
-% 	\begin{split}
-% 		\bm{\rho}(\bm{U}, \bm{V}) &= \sqrt{\mathrm{E} \left( \Sigma_{\bm{UU}}^{-1} \Sigma_{\bm{UV}} \Sigma_{\bm{VV}}^{-1} \Sigma_{\bm{VU}} \right)}
-% 	\end{split}
-% \end{equation}
 	\begin{equation}
@@
 
-% where $ \mathrm{E}(\Sigma) $ denotes the eigenvalues of the matrix $ \Sigma $.
-% $ \Sigma_{UU} $ and $ \Sigma_{VV} $ are the covariance matrices of $ \bm{U} $
-% and $ \bm{V} $, and $ \Sigma_{UV} $ and $ \Sigma_{VU} $ are the
-% cross-covariance matrices.
-
-Canonical correlations generalize correlations to multi-dimensional variables. They
-find orthogonal linear transformations $ a $ and $ b $ of $ \bm{U} $ and $ \bm{V} $
-such that the correlation between the transformed variables $ a^T \bm{U} $ and 
-$ b^T \bm{V} $ is maximized. This yields a vector of correlation values of 
-size $ \min(p, q) $, representing the correlation coefficient of each pair of 
-transformed variables. Pearson's correlation coefficient is a special case of 
-canonical correlations when $ p = q = 1 $.
+Canonical correlations generalize correlation coefficients to multi-dimensional
+variables. They find orthogonal linear transformations $ a $ and $ b $ of $
+\bm{U} $ and $ \bm{V} $ such that the correlation coefficient between the
+transformed variables $ a^T \bm{U} $ and $ b^T \bm{V} $ is maximized. This
+yields a vector of correlation coefficient values of size $ \min(p, q) $,
+representing the correlation coefficient of each pair of transformed variables.
+Pearson's correlation coefficient is a special case of canonical correlations
+when $ p = q = 1 $.
 
@@
 
-This measure of association has many desired properties:
+This measure of association has many desired properties for our application:
 

### 2025-02-05 paper/main.tex
Updates

@@
 different conditional measures of association. However, no unified measure
-exists for mixed data. In this section, we introduce a novel measure of
-conditional association for mixed data by extending the idea of partial
-correlation coefficient to mixed data. Specifically, our approach combines a
-mixed data residualization method \citep{Ankan2023} with Pillai's Trace
+exists for mixed data. Here, we introduce a measure of conditional association
+for mixed data by extending the concept of partial correlation coefficient for
+continuous variables to mixed data. Specifically, our approach combines a mixed
+data residualization method \citep{Ankan2023} with Pillai's Trace
 \citep{Pillai1955}, a multivariate measure of association based on canonical
@@
 Given a dataset $ D = (x, y, \bm{z}) $ on variables $ X $, $ Y $, and $ \bm{Z}
-$, we aim to estimate the conditional association $ \rho(X, Y; \bm{Z}) $
-between $ X $ and $ Y $ conditioned on $ \bm{Z} $. We first use a mixed data
-residualization approach \citep{Ankan2023} to compute residuals $R_X$ and $R_Y$
-for $ X $ and $ Y $. The method for computing these residuals depend on the 
-type of variable:
+$, we goal to estimate the conditional association $ \rho(X, Y; \bm{Z}) $ . We
+achieve this by first computing the residuals $ R_X $ and $ R_Y $ for $ X $ 
+and $ Y $ respectively, using the following residualization approach:
 
@@
 
-We repeat the above steps for both the variables $ X $ and  $ Y $ to get
-residuals $ R_X $ and $ R_Y $. Here, we have the option to choose the
-estimators based on the properties (distribution, type of relationship, and so
-on) of the data. Non-parametric ensemble estimators such as Random Forest and
-XGBoost can handle any combination of data types while performing reasonably
-well on a wide range of distributions and relationship types. For linear
-relationships, we can also choose simpler models such as linear regression.
+Here, we have the option to choose the estimators based on the properties
+(distribution, type of relationship, and so on) of the data. Non-parametric
+ensemble estimators such as Random Forest and XGBoost can handle any
+combination of data types while performing reasonably well on a wide range of
+distributions and relationship types. For linear relationships, we can also
+choose simpler models such as linear regression.
 
-After the above residualization steps are repeated for both $ X $ and $ Y $, we
-end up with two residual matrices: $ R_X $ and $ R_Y $. The type of the
-variable determines the shape for these matrices. If the variable is continuous
-or ordinal, we get a residual matrix of shape $ ( n \times 1 ) $ and if the
-variable is categorical, we have a residual matrix of shape $ ( x \times (k-1))
-$ where one of the dummy encoded variables is dropped to avoid
-multicollinearity.
+Repeating the above residualization method for both $ X $ and  $ Y $, we obtain
+residual matrices $ R_X $ and $ R_Y $. The type of the variable determines the
+shape for these matrices. If the variable is continuous or ordinal, the
+residual matrix is of shape $ ( n \times 1 ) $ and if the variable is
+categorical, we have a residual matrix of shape $ ( x \times (k-1)) $ where $ k
+$ is the number of categories of the variable.
 
-We now have two continuous-valued residual matrices. The next step is
-to quantify the association between these residuals. For this purpose
-we use canonical correlations that have been used to measure the association
-between two sets of random variables.
+The next step is to quantify the association between these residual matrices.
+For this purpose we use canonical correlations that have been widely used to
+measure the association between sets of random variables.
 

### 2025-02-05 paper/main.tex
Updates

@@
 
-This measure of association has many desired properties for our application:
+Our measure of conditional association has several key properties that make it
+well-suited for our application:
 
 \begin{enumerate}
-	\item \textbf{Bounded: } $ \rho $ is bounded between $ 0 $ and $ 1 $,
-		with $ 0 $ representing no association while $ 1 $ represents
-		perfect linear association. This helps in interpretation and
-		comparing values across multiple partial associations.
+	\item \textbf{Bounded: } $ \rho $ is bounded between $ 0 $ (no
+		association) to $ 1 $ (perfect linear association). This bounded
+		nature simplifies the interpretation and allows for easy 
+		comparisons across multiple values.
 	\item \textbf{Independent of Sample Size: } The measure remains consistent
@@
 		studies with different sample sizes.
-	\item \textbf{Invariant to the Dummy Encoding Scheme: } Since canonical
-		correlations search for linear combinations over the columns,
-		the measure is invariant to the specific dummy encoding
-		scheme used for categorical variables.
+	\item \textbf{Invariant to the Dummy Encoding: } We use dummy encoding 
+		to compute the residual matrix for categorical variables. Since
+		canonical correlations identify linear combinations across
+		columns, this measure is invariant to the specific dummy
+		encoding scheme used.
 	\item \textbf{Equivalent to Partial Correlations for Continuous variables}
@@
 		partial correlations to mixed data types.
-	\item \textbf{Equivalent to polychoric and polyserial correlation for ordinal and continuous variables: }
+	\item \textbf{Equivalent to polychoric and polyserial correlation for
+			ordinal and continuous variables: }
 		Under the assumption that the ordinal variable is generated by
@@
 		coefficient.
-	\item \textbf{Related to Cram\'ers V: } 
+	\item \textbf{Related to Cram\'er's V: } 
 \end{enumerate}

### 2025-02-05 paper/main.tex
Updates

@@
 		categorical variable and then compute the residuals as follows: 
-		$$ R_X = x_i - \hat{p}(\bm{z}_i) $$
+		$$ R_{x_i} = x_i - \hat{p}(\bm{z}_i) $$
 \end{enumerate}
@@
 		encoding scheme used.
-	\item \textbf{Equivalent to Partial Correlations for Continuous variables}
-		Canonical correlations is equivalent to Pearson's correlation coefficient
-		when both $ X $ and $ Y $ are continuous. Hence, $ \rho $ generalizes
-		partial correlations to mixed data types.
+	\item \textbf{Equivalent to Partial Correlation}
+		When both $ X $ and $ Y $ are continuous variables, our measure
+		is equivalent to the absolute value of partial correlation.
 	\item \textbf{Equivalent to polychoric and polyserial correlation for
@@
 
-Using this mixed data partial association measure, we now propose a method for
-construcing DAGs while incorporating expert knowledge iteratively. Given a
-potentially empty DAG $ G $ and a dataset $ D $, a p-value threshold of $
-\alpha $ and a measure of association threshold of $ \beta $, the construction
+Using this mixed data measure of association, we now propose a method for
+iteratively constructing DAGs while incorporating expert knowledge. This
+approach is based on implied CI based testing where we use the measure of
+association to rank the failing CI tests. This ranking allows us to prioritize
+adding edges between pair of variables that have the most severe problems.
+
+Given a DAG $ G $ (potentially empty), a dataset $ D $, a p-value threshold $
+\alpha $ and a measure of association threshold $ \beta $, our construction
 procedure works as follows:
@@
 \begin{enumerate}
-	\item For each pair of variables $ X $ and $ Y $ in the model:
+	\item For each pair of variables $ X $ and $ Y $ in $ G $:
 		\begin{itemize}
-			\item If there is an edge between $ X $ and $ Y $, remove the edge to obtain the DAG $ \underline{G} $, and 
-				compute the measure of association $ \phi(X, Y \rvert \underline{G}_{pa}(X) \cup \underline{G}_{pa}(Y)) $
+			\item If there is an edge between $ X $ and $ Y $, remove the edge to obtain the DAG $ G' $, and compute the measure of association $ \phi(X, Y \rvert G'_{pa}(X) \cup G'_{pa}(Y)) $
 			\item If there is no edge between $ X $ and $ Y $, compute the measure of association $ \phi(X, Y \rvert G_{pa}(X) \cup \underline{G}_{pa}(Y) $.
@@
 		the model using undirected red edges with the strength of
-		association represented using edge width. Users can then add
-		edges to the model (shown in green). Unnecessary edges are shown in 
-		black.}
+		association represented using edge width. Users can then
+		iteratively add edges to the model (shown in green) while
+		deciding the edge orientation based on domain knowledge.
+		Unnecessary edges are shown in black.}
 	\label{fig:web}
@@
 \section{Conclusions}
+
 Points:

### 2025-02-05 paper/main.tex
Updates

@@
 distributions and relationship types. For linear relationships, we can also
-choose simpler models such as linear regression.
+choose simpler models such as linear regression and its variants.
 
-Repeating the above residualization method for both $ X $ and  $ Y $, we obtain
-residual matrices $ R_X $ and $ R_Y $. The type of the variable determines the
-shape for these matrices. If the variable is continuous or ordinal, the
-residual matrix is of shape $ ( n \times 1 ) $ and if the variable is
-categorical, we have a residual matrix of shape $ ( x \times (k-1)) $ where $ k
-$ is the number of categories of the variable.
+Repeating the above residualization process for both $ X $ and  $ Y $, we
+obtain residual matrices $ R_X $ and $ R_Y $. The type of the variable
+determines the shape for these matrices. If the variable is continuous or
+ordinal, the residual matrix is of shape $ ( n \times 1 ) $ and if the variable
+is categorical, we have a residual matrix of shape $ ( x \times (k-1)) $ where
+$ k $ is the number of categories of the variable.
 
@@
 
-We use Pillai's Trace for our purpose for two main reasons: 1) It uses all the
-canonical correlation values, 2) Its interpretation is similar to correlation
-coefficient, i.e., $ 0 $ signifies no association and $ 1 $ signifies perfect
-linear relationship. This gives us the following partial measure of
-association:
+We use a slightly modified version of Pillai's Trace for our purpose for two
+main reasons: 1) It uses all the canonical correlation values, 2) Its
+interpretation is similar to correlation coefficient, i.e., $ 0 $ signifies no
+association and $ 1 $ signifies perfect linear relationship. This gives us the
+following partial measure of association:
 
@@
 		comparisons across multiple values.
-	\item \textbf{Independent of Sample Size: } The measure remains consistent
-		regardless of the sample size ensuring reliability across 
-		studies with different sample sizes.
+	\item \textbf{Independent of Sample Size: } 
 	\item \textbf{Invariant to the Dummy Encoding: } We use dummy encoding 

### 2025-02-05 paper/main.tex
Updates

@@
 \subsection{Notations}
-\todo[inline]{Improve this section once all the notations are in place}
 We denote random variables with uppercase letters $ X $, and a set of random
@@
 overview of some of the commonly used measures for different data types.
-
-\todo[inline]{Remove or move equations to appendix if no space}
 

### 2025-02-05 paper/main.tex
Updates

@@
 \subsection{Connection to Score Based Methods}
-\begin{itemize}
-	\item The process of iteratively adding and removing edges in very similar to
-		Greedy Equivalence Search (GES).
-	\item But not decomposible like scoring methods. 
-	\item But Because they make only local changes, can get stuck in local minimas.
-	\item Both approach looks for modifications globally.
-	\item The sum of unexplained effect provides an absolute measure of model fit 
-		unlike scoring methods.
-\end{itemize}
+
+Our procedure closely resembles certain score-based automated causal discovery
+methods, such as Greedy Equivalence Search (GES), where the algorithm
+iteratively adds or removes edges that improve the scoring metric the most.
+Similarly, a greedy version of our approach would involve adding an edge
+between the pair of variables with the highest unexplained correlation. While
+both methods aim to find modifications that most improve the model at each
+step, a key distinction lies in the nature of the evaluation criteria: unlike
+standard scoring metrics, our measure of association is not decomposable. A key
+property of scoring metrics is that they should be decomposible, i.e., they can
+be expressed as the sum of scores of nodes given its parents. This means adding
+or removing an edge affects only a localized part of the model. In contrast,
+our measure of association is global, meaning that modifying an edge in one
+part of the model can influence association values elsewhere.
+
+Another major difference is in the interpretability of the evaluation metric.
+Most scoring metrics are based on log-likelihood with a penalty for model
+complexity. These score metrics allow for relative comparisons between models
+but do not provide an absolute measure of model fit. That is, they indicate
+which model is better for a given dataset but do not reveal how well the model
+explains the data in an absolute sense. In contrast, summing our measure of
+association across all variables provides a direct fitness measure—this sum
+approaches zero when the model perfectly accounts for all observed correlations
+in the data. This property enables an absolute assessment of model quality,
+rather than just a relative comparison between models.
 

### 2025-02-05 paper/main.tex
Updates

@@
 domains remain limited, as researchers often prefer to construct DAGs manually
-based on their domain knowledge. This preference arises due to several
-practical issues with automated algorithms, such as they make obvious errors
-and output Markov Equivalence Classes (MECs). To address these challenges, we
-propose an iterative method to assist researchers in manually constructing or
-improving an existing DAG by combining implied Conditional Independence (CI)
-testing with expert knowledge. Our proposed method utilizes a novel measure of
-partial association for mixed data to rank violations of implied CIs in a given
-DAG, enabling users to prioritize modifications while utilizing their domain
-knowledge to decide the orientation of new edges. Empirical results show that
-this manual model construction approach achieves performance comparable to
-automated algorithms if the user is able correctly orientation edges in one out
-of three cases. Additionally, we demonstrate that, in the absence of an expert,
-a Large Language Model (LLM) can be potentially determine edge orientations. We
-provide the implementation of our approach in both a web tool at: <redacted for
-review> and a Python package <redacted for review>.
+based on domain knowledge. This preference arises due to several practical
+challenges with automated algorithms, such as their tendency to make obvious
+errors and output Markov Equivalence Classes (MECs) rather than a DAG. To
+address these challenges, we propose an iterative method that guides
+researchers in manually constructing or improving the fit of an existing DAG.
+Our approach combines implied Conditional Independence (CI) testing with a
+novel measure of conditional association between variables to rank violations
+of implied CIs in a given DAG. This ranking helps prioritize modifications to
+the DAG to improve its fit to the data while utilizing expert knowledge to
+decide the orientation of new edges. Empirical results show that this guided
+manual model construction approach achieves performance comparable to automated
+algorithms if we are able to correctly decide the orientation of new edges in
+one out of three cases. Additionally, we demonstrate that, in the absence of an
+expert, a Large Language Model (LLM) can be potentially used to determine edge
+orientations. We provide the implementation of our approach in both a web tool
+at: <redacted for review> and a Python package <redacted for review>.
 

### 2025-02-05 paper/main.tex
Updates

@@
 \begin{abstract}
-% Causal discovery has received significant attention in the Directed Acyclic
-% Graphs (DAGs) literature, leading to the development of numerous automated
-% algorithms for learning DAGs from data. Despite these advancements, the
-% application of these algorithms in applied domains remains limited, with
-% researchers often preferring to manually construct DAGs based on their expert
-% knowledge. This preference stems due to some practical issues with these
-% algorithms such as making obvious mistakes, and outputting Markov Equivalence
-% Class (MEC). As a result of these issues users typically have to rely on post
-% hoc modifications to the model based on their expertise. Given this reliance on
-% manual model modification and construction, we propose an iterative method to
-% assist researchers in manually constructing DAGs by combining implied
-% Conditional Independence (CI) based model testing and utilizing their expert
-% knowledge. We present a novel measure of partial association for mixed data
-% that allows us to rank violations of implied CIs in the model. Using this
-% ranking users can prioritize modifications to the model while using their
-% expert knowledge to decide orientation of edges. Empirically, we show that
-% a greedy approach achieves performance comparable to automated algorithms, provided
-% the user can correctly determine the edge orientation in $ 1 $ out of $ 3 $ cases.
-% Additionally, we show that in the absence of an expert, a Large Language Model can
-% potentially be used for determining edge orientations.
 
@@
 behind observed phenomena and guide effective interventions or policy
-decisions. \emph{Causal Discovery} methods aim to discover such relationships
+decisions. Causal discovery methods aim to discover such relationships
 among random variables using observational data. Approaches to causal discovery
@@
 algorithms to learn causal structures from datasets. These efforts have led to
-numerous causal discovery algorithms, including constraint-based methods like
-the PC algorithm \citep{Spirtes2001} and Fast Causal Inference (FCI)
-\citep{Spirtes2000}), score-based methods such as Hill-Climb Search and Greedy
-Equivalence Search (GES) \citep{Chickering2002}, and continuous
-optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
-\citep{Bello2022}. While DAG-based methods focus on automated discovery,
-SEM-based methods emphasize expert-driven model specification. This includes
-tools to assist researchers in manually constructing models, enabling them to
-incorporate their domain knowledge in the model building process. Researchers
-typically begin with an initial model based on their expertise and then use
-these tools to guide modifications that improve the model's fit to data. This
-process is commonly known as \emph{Specification Search} \citep{Long1983} and
-uses method such as modification indices, and Wald-based tests
-\citep{Marcoulides2018}.
+numerous causal discovery algorithms, such as constraint-based methods like PC
+algorithm \citep{Spirtes2001} and Fast Causal Inference \citep{Spirtes2000}),
+score-based methods such as Hill-Climb Search and Greedy Equivalence Search
+\citep{Chickering2002}, and continuous optimization-based methods like NO TEARS
+\citep{Zheng2018} and DAGMA \citep{Bello2022}. While DAG-based methods focus on
+automated discovery, SEM-based methods emphasize expert driven model
+specification. This includes tools to assist researchers in manually
+constructing models, enabling them to incorporate their domain knowledge in the
+model building process. Researchers typically begin with an initial model based
+on their domain knowledge and then use these tools to guide modifications that
+improve the model's fit to data. This process is commonly known as
+Specification Search \citep{Long1983} and uses method such as modification
+indices, and Wald-based tests \citep{Marcoulides2018}.
 
@@
 		consistent, their behavior on finite samples is not well
-		understood. Additionally, their output can vary significantly
-		depending on the choice of algorithm and hyperparameters,
-		making it difficult to assess reliability. The absence of
+		understood. Their output can vary significantly depending on
+		the choice of algorithm and hyperparameters, making it
+		difficult to assess reliability. Additionally, the absence of
 		robust performance evaluation methods for any given dataset
@@
 
-Fig.~\ref{fig:intro} highlights some of these issues. When constructing models
-manually it is important to test whether the constructed DAG is representative
-of the dataset or not. Based on this testing and domain knowledge further
-modifications can be made to the DAG. One of the approaches to test and modify
-DAGs is by testing whether the implied CI tests of the DAG are true in the data
-or not \citep{Ankan2021}. If some of the tests are violated in the data,
-researchers can use their domain knowledge to make modifications to the DAG.
+Figure~\ref{fig:intro} highlights some of these issues.
 
-However, this CI testing approach also has a few drawbacks: 1) CIs are implied
-only by missing edges in the model, so we can not test whether the existing
-edges are correct or not. 2) The decision whether a CI holds in the data or not
-is based on a p-value threshold. But we can get significant p-values for even
-very weak relationship if the sample size is high. 3) Lastly, the number of
-implied CIs can be large making it difficult to go through all of them; for 
-example, in the moderately sized alarm network \citep{Beinlich1989} with $ 37 $
-nodes, there are $ 287 $ implied CIs.
+When manually constructing models, it is important to test whether the
+resulting DAG accurately represents the dataset. Based on this evaluation and
+domain knowledge, we can make further modification to the model. One common
+approach for assessing DAGs is to check whether the Conditional Independences
+(CIs) implied by the DAG hold in the data \citep{Ankan2021}. If we find
+violations to these CIs, we can use our domain knowledge to make appropriate
+modifications to the DAG. However, this CI testing approach has a few
+limitations:
 
-% To deal with these issues
-% methods for integrating expert knowledge have been developed. The two main
-% methods to integrate expert knowledge in the model construction process is to:
-% 1. Provide all the expert knowledge to the causal discovery algorithm for
-% example in the form of required and forbidden edges \citep{Meek1995}, or
-% temporal information \citep{Bang2023} and let the algorithm use it to construct
-% the model. 2. Utilize conditional independence (CI) testing to identify issues in
-% the model, and manually fix them based on expert knowledge \citep{Ankan2021}.
-% We focus on this second approach in this paper.
-
-% In this CI testing procedure, users test whether the implied CIs of the DAG
-% hold in the dataset or not using statistical CI tests. Based on the results of
-% these statistical tests, we can accordingly make modifications to the current
-% model to fix the violations. However, this CI based testing also has a few
-% drawbacks: 1) CIs are implied only by missing edges in the model, so we can not
-% test for already existing edges. 2) In high sample size scenarios, p-values can
-% suggest a significant association between variables even if they have a very
-% weak relationship. 3) Number of implied CIs can be large making it difficult to
-% manually go through each of them; for example, in the moderately sized alarm
-% network with $ 37 $ node, there are $287$ implied CIs.
-
-% There has been work on trying to come up with a single test to combine all
-% these test to give a single answer \citep{Eulig2023, Shipley2000}. However,
-% what is missing is the guide on how to how to fix these tests. Additionally
-% there has been a lot of work to integrate expert knowledge into causal
-% discovery algorithms which has shown significant improvement.
+\begin{enumerate}
+\item As CIs are only implied by missing edges in the model, this approach does 
+	not test whether the existing edges are correct.
+\item Determining whether a CI holds in data is based on a p-value threshold. This can 
+	be unreliable, for example, we get a significant p-value for even very weak 
+	relationship if the sample size is high.
+\item The number of implied CIs can be large making it difficult to manually
+	check all of them. For example, a moderately sized alarm network
+	\citep{Beinlich1989} with $ 37 $ nodes, implies $ 287 $ CIs.
+\end{enumerate}
 
 To tackle these issues with CI based testing, we take inspiration from
-\emph{modification indices} in specification search. The modification indices
+modification indices in specification search. The modification indices
 method tries to identify and rank changes in the model such that modifications

### 2025-02-05 paper/main.tex
Udpates

@@
 Our approach combines implied Conditional Independence (CI) testing with a
-novel measure of conditional association between variables to rank violations
+measure of conditional association between variables to rank violations
 of implied CIs in a given DAG. This ranking helps prioritize modifications to
@@
 
-To tackle these issues with CI based testing, we take inspiration from
-modification indices in specification search. The modification indices
-method tries to identify and rank changes in the model such that modifications
-that improve the model fit the most are ranked higher. This ranking allows
-researchers to focus first on the suggestions that can improve the model the
-most and while integrating their expertise to choose the most optimal
-modifications. Analogous to this, we develop a procedure for ranking CI test
-violations allowing researchers to focus on the worst violations in the model.
-This ranking method utilizes both p-values and a measure of conditional
-association. Utilizing this procedure researchers can iteratively construct
-DAGs while integrating their domain knowledge in the model building procedure.
+To tackle these limitations, we draw inspiration from modification indices in
+specification search. Modification indices provide ranking for potential model
+modification based on their impact on the model's fit, allowing us to
+prioritize most meaningful modifications. Similarly, we propose a method for
+ranking CI test violations to help us focus on the most critical
+inconsistencies in the model. To determine this ranking we utilize a measure of
+conditional association between variables in combination with p-values from the
+CI test. This approach helps us in tackling all the three issues outlined
+above. The measure of association allows us to test for existing edges, as well
+as not only rely on p-value to make decision whether a CI test holds. The
+ranking helps us in focusing on the most critical violations.
 
@@
 
-\caption{A comparison of Markov Equivalence Class (MEC) learned using Adult Income Dataset \citep{Becker1996} using (a \& b) PC algorithm using a mutual information based test, (c \& d) Hill Climb Search with Bayesian Information Criterion (BIC) score, (e \& f) PC algorithm with a residualization based test \citep{Ankan2023}.
-The learned model structure varies significantly depending on the algorithm, the CI test, and the sample size used.}
+    \caption{A comparison of Markov Equivalence Class (MEC) learned using Adult
+	     Income Dataset \citep{Becker1996} using (a \& b) PC algorithm
+	     using a mutual information based test, (c \& d) Hill Climb Search
+	     with Bayesian Information Criterion (BIC) score, (e \& f) PC
+	     algorithm with a residualization based test \citep{Ankan2023}. The
+	     learned model structure varies significantly depending on the
+     	     algorithm, the CI test, and the sample size used.}
     \label{fig:intro}
@@
 Our main contributions in this paper are as follows:
+
 \begin{enumerate}
-	\item We propose a novel measure of conditional association for mixed
-		data based on canonical correlations
-		(Section~\ref{sec:mixed_association}. This measure of
-		association generalizes some of the commonly used special case
-		measures of association to mixed data.
-	\item Using this conditional association measure, we present a
-		procedure to manually construct DAGs by ranking violation of
-		implied CIs in the model (Section~\ref{sec:modification}. This
-		iterative procedure allows users to integrate their domain
-		knowledge in the model construction procedure.
-	\item To allow researchers to easily apply this method on their
-		datasets, we present a web-browser based interactive tool as well
-		as implementation in a Python package (Section~\ref{sec:web}).
+    \item We propose a novel measure of conditional association for mixed data
+	    based on canonical correlations
+	    (Section~\ref{sec:mixed_association}). This measure generalizes
+	    several commonly used special case measures of association to mixed
+	    data.
+    \item Using this measure of conditional association, we develop a procedure
+	    to rank violations of implied CIs. Using this ranking helps us
+	    prioritize modifications to DAG to improve its fit while allowing
+	    us to integrate expert knowledge when adding new edges (Section~\ref{sec:modification}).
+    \item We provide a web-based interactive tool and a Python package to allow 
+	    researchers to easily apply this method to their datasets (Section~\ref{sec:web}).
     \item Lastly, we compare the manual DAG construction method with automated
 	    causal discovery algorithms (Section~\ref{sec:empirical}).
-
 \end{enumerate}
@@
 correlations.
-
-% As there are no generalized measures of association for mixed data, in this
-% section, we first introduce a novel measure of partial association for mixed
-% data by extending the concept of partial correlations \citep{Whittaker2009}
-% \todo{Verify this citation} to handle mixed data types. To compute partial
-% correlations, we have two key steps: 1) a residualization method, and 2) a
-% measure of associations between the residuals. When both $ X $ and $ Y $ are
-% continuous variables, residuals are typically computed using a regression-based
-% method, and Pearson's correlation coefficient is used to measure the
-% association between these residuals. To extend this approach to mixed data, we
-% combine a mixed data residualization method \citep{Ankan2023} with a
-% multivariate measure of association based on canonical correlations, Pillai's
-% Trace \citep{Pillai1955}.
 

### 2025-02-05 paper/main.tex
Updates

@@
 conditional association between variables in combination with p-values from the
-CI test. This approach helps us in tackling all the three issues outlined
-above. The measure of association allows us to test for existing edges, as well
-as not only rely on p-value to make decision whether a CI test holds. The
-ranking helps us in focusing on the most critical violations.
+CI test. This approach addresses all three of the issues outlined above. The
+measure of association allows us to evaluate the validity of existing edges. It
+reduces our reliance on p-values alone for deciding whether the CI test holds.
+The ranking helps us prioritize the most critical violations, making it feasible 
+to focus on the areas that most improve the model's fit.
 

### 2025-02-05 paper/main.tex
Updates

@@
 \subsection{Measures of Conditional Association}
-Our proposed method for iterative DAG construction relies on a measure of
-conditional association, also commonly known as partial association, between
-variables. Specifically, we are interested in quantifying the association
-between variables $ X $ and $ Y $ when conditioned on a set of variables $
-\bm{Z} $, which may be empty ($ \bm{Z} = \emptyset $). Various measures of
-conditional association have been used in the literature, depending on the type
+Our approach for ranking CI tests is based on a measure of conditional
+association -- also known as partial association -- between variables.
+Specifically, we are interested in quantifying the association between
+variables $ X $ and $ Y $ when conditioned on a set of variables $ \bm{Z} $
+(which may be empty, i.e., $ \bm{Z} = \emptyset $). Various measures of
+conditional association have been used for this purpose depending on the type
 of $ X $, $ Y $, and $ \bm{Z} $. These measures are often based on the effect
-size of the CI test $ X \ci Y \rvert \bm{Z} $. In this section, we provide an
+size of the CI test $ X \ci Y \rvert \bm{Z} $. In this section, we give an
 overview of some of the commonly used measures for different data types.

### 2025-02-05 paper/main.tex
Updates

@@
 observed variables, and linear measure of association. 
+where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, and $
+\sigma_X $ and $ \sigma_Y $ are the standard deviations of $ X $ and $ Y $
+respectively.
 
@@
 coefficient is typically used. In the case when $ \bm{Z} = \emptyset $, the
-correlation coefficient is given as:
+correlation coefficient is defined as:
 
@@
 
-where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, and $
-\sigma_X $ and $ \sigma_Y $ are the standard deviations of $ X $ and $ Y $
-respectively. When $ \bm{Z} \neq \emptyset $, partial correlation coefficient
+When $ \bm{Z} \neq \emptyset $, partial correlation coefficient
 can be used. One of the ways to estimate it is to use two regression
@@
 likelihood of data given polychoric or polyserial correlation coefficient. This 
-method gives us values that are Pearson's correlation coefficient.
+method gives us the covariance matrix, $\Sigma$ on the variables $ X $, $ Y $, and $ \bm{Z} $.
+
+\begin{equation}
+	\rho_{X, Y} = \Sigma_{XY}
+\end{equation}
+
+When $ \bm{Z} \neq \emptyset $,
+
+\begin{equation}
+	\rho_{X, Y; \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
+\end{equation}
 

### 2025-02-06 paper/main.tex
Updates

@@
 \subsection{Notations} 
+\todo[inline]{Needs improvement}
 We denote random variables with uppercase letters $ X $, and a set of random
@@
 In the previous section, we discussed how different types of variables require
-different conditional measures of association. However, no unified measure
-exists for mixed data. Here, we introduce a measure of conditional association
-for mixed data by extending the concept of partial correlation coefficient for
-continuous variables to mixed data. Specifically, our approach combines a mixed
-data residualization method \citep{Ankan2023} with Pillai's Trace
+different measures of conditional association. However, there is no unified
+measure that handles mixed data. Here, we introduce a measure of conditional
+association for mixed data by extending the concept of partial correlation
+coefficient (commonly used for continuous variables) to mixed data. Similar to
+partial correlation coefficient, our method integrates a mixed data
+residualization method \citep{Ankan2023} with Pillai's Trace
 \citep{Pillai1955}, a multivariate measure of association based on canonical
@@
 Given a dataset $ D = (x, y, \bm{z}) $ on variables $ X $, $ Y $, and $ \bm{Z}
-$, we goal to estimate the conditional association $ \rho(X, Y; \bm{Z}) $ . We
-achieve this by first computing the residuals $ R_X $ and $ R_Y $ for $ X $ 
-and $ Y $ respectively, using the following residualization approach:
+$, our goal is to estimate the conditional association $ \rho(X, Y; \bm{Z}) $. 
+In the first step, we compute the residuals $ R_X $ and $ R_Y $ for variables
+$ X $ and $ Y $ respectively. Depending on the type of variable the residual
+is computed as follows:
 
@@
 
-Here, we have the option to choose the estimators based on the properties
-(distribution, type of relationship, and so on) of the data. Non-parametric
-ensemble estimators such as Random Forest and XGBoost can handle any
-combination of data types while performing reasonably well on a wide range of
-distributions and relationship types. For linear relationships, we can also
-choose simpler models such as linear regression and its variants.
+We have the option here to choose the estimators based on the characteristics
+such as distribution, type of relationship, and so on of our dataset.
+Non-parametric ensemble estimators such as Random Forest and XGBoost are robust 
+for diverse data types and complex relationships. For linear relationships,
+simpler models such as linear regression and its variants may suffice.
 
-Repeating the above residualization process for both $ X $ and  $ Y $, we
-obtain residual matrices $ R_X $ and $ R_Y $. The type of the variable
-determines the shape for these matrices. If the variable is continuous or
-ordinal, the residual matrix is of shape $ ( n \times 1 ) $ and if the variable
-is categorical, we have a residual matrix of shape $ ( x \times (k-1)) $ where
-$ k $ is the number of categories of the variable.
+We repeat the above residualization step for both the variables $ X $ and $ Y $
+obtaining residual matrices $ R_X $ and $ R_Y $. The type of variable determines
+the shape of these matrices. If the variable is continuous or ordinal, the 
+residual matrix is of shape $ (n \times 1 ) $ and if the variable is categorical,
+we get a residual matrix of shape $ (x \times (k - 1)) $, where $ k $ is the number
+of categories of the variable.
 
-The next step is to quantify the association between these residual matrices.
+The second step is to quantify the association between these residual matrices.
 For this purpose we use canonical correlations that have been widely used to
@@
 \begin{definition}
-
 	Given two sets of random variables $ \bm{U} = (U_1, U_2, \cdots, U_p) $
@@
 
-Canonical correlations generalize correlation coefficients to multi-dimensional
-variables. They find orthogonal linear transformations $ a $ and $ b $ of $
-\bm{U} $ and $ \bm{V} $ such that the correlation coefficient between the
-transformed variables $ a^T \bm{U} $ and $ b^T \bm{V} $ is maximized. This
-yields a vector of correlation coefficient values of size $ \min(p, q) $,
-representing the correlation coefficient of each pair of transformed variables.
-Pearson's correlation coefficient is a special case of canonical correlations
-when $ p = q = 1 $.
+Canonical correlations generalize the concept of correlation coefficients to
+multi-dimensional variables. It finds orthogonal linear transformations $ a $
+and $ b $ that maximized the correlation between the transformed variables $
+a^T \bm{U} $ and $ b^T \bm{V} $. This yields a vector of correlation
+coefficient values of size $ \min(p, q) $, representing the correlation
+coefficient of each pair of transformed variables. Notably, Pearson's
+correlation coefficient is a special case of canonical correlations when $ p =
+q = 1 $.
 
-Many measures of association based on canonical correlation have been used in
-multivariate statistics, each with slightly different properties.
+Several measures of association have been derived from canonical correlations, such as:
 \begin{itemize}
@@
 We use a slightly modified version of Pillai's Trace for our purpose for two
-main reasons: 1) It uses all the canonical correlation values, 2) Its
-interpretation is similar to correlation coefficient, i.e., $ 0 $ signifies no
-association and $ 1 $ signifies perfect linear relationship. This gives us the
-following partial measure of association:
+main reasons: 1) It uses all the canonical correlation values, capturing the
+full extent of association, 2) Its interpretation is similar to Pearson's
+correlation coefficient, i.e., $ 0 $ signifies no association and $ 1 $
+signifies perfect linear relationship. This leads to our proposed measure of
+conditional association
 
@@
 
-Our measure of conditional association has several key properties that make it
-well-suited for our application:
+Our proposed measure has several desirable properties that make it well-suited
+for our application:
 
 \begin{enumerate}
-	\item \textbf{Bounded: } $ \rho $ is bounded between $ 0 $ (no
-		association) to $ 1 $ (perfect linear association). This bounded
-		nature simplifies the interpretation and allows for easy 
-		comparisons across multiple values.
+	\item \textbf{Bounded: } The measure is bounded between $ 0 $ (no
+		association) to $ 1 $ (perfect linear association), simplifying
+		interpretation.
 	\item \textbf{Independent of Sample Size: } 
-	\item \textbf{Invariant to the Dummy Encoding: } We use dummy encoding 
-		to compute the residual matrix for categorical variables. Since
-		canonical correlations identify linear combinations across
+	\item \textbf{Invariant to the Dummy Encoding: } For categorical
+		variables, residuals are computed using dummy encoding. Since
+		canonical correlations identifies linear combinations across
 		columns, this measure is invariant to the specific dummy
 		encoding scheme used.
-	\item \textbf{Equivalent to Partial Correlation}
+	\item \textbf{Equivalent to Partial Correlation for Continuous Variables: }
 		When both $ X $ and $ Y $ are continuous variables, our measure
-		is equivalent to the absolute value of partial correlation.
+		is equivalent to the absolute value of partial correlation coefficient.
 	\item \textbf{Equivalent to polychoric and polyserial correlation for
@@
 
-So, in a sense this measure generalized currently used measures of associations
-to mixed data.
+In essence, our measure of conditional association extends existing metrics to mixed data, providing 
+a unified measure that is interpretable.
 

### 2025-02-06 paper/appendix.tex
Uupdates
udpates

@@
 \section{Prompts Used for LLM}
+\label{section:llms}
 
@@
 	\centering
-	\begin{Verbatim}
+	\begin{verbatim}
 You are an expert in Social Science. Following are the descriptions of two variables:
@@
 in the answer; Do not add any text formatting to the answer.
-	\end{Verbatim}
+	\end{verbatim}
 	\caption{Prompt used for the LLM. Here the variable descriptions are replaced with description provided in Fig.~\ref{fig:var_description}}
@@
 \begin{figure}[ht!]
-	\begin{Verbatim}
+	\begin{verbatim}
           Age: The age of a person
@@
        Income: The income i.e. amount of money the person makes.
-	\end{Verbatim}
+	\end{verbatim}
 	\caption{Variable descriptions used for prompting the LLM}

### 2025-02-06 paper/main.tex
udpates
Updates

@@
 \usepackage{subcaption}
-\usepackage{fancyvrb}
+% \usepackage{fancyvrb}
 \usepackage[ruled]{algorithm2e}
@@
 
+Using the mixed data measure of association presented in the last section, we 
+now present a method to rank modifications to a given DAG.
+
 Using this mixed data measure of association, we now propose a method for
@@
 	\includegraphics[page=1]{fig5.pdf}
-	\caption{DAG learned using the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert.}
+	\caption{DAG learned using the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert. The p-value threshold used is $ 0.05 $ and the measure of association threshold is $ 0. 1 $}
 	\label{fig:adult_llm}

### 2025-02-06 paper/main.tex
Updates

@@
 observed variables, and linear measure of association. 
-where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, and $
+where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, $ \mathrm{corr}(X, Y) $ is the correlation between $ X $ and $ Y $, and $
 \sigma_X $ and $ \sigma_Y $ are the standard deviations of $ X $ and $ Y $
-respectively.
+respectively.  Covariance matrix using $ \Sigma $ and the entry corresponding to
+covariance between $ X $ and $ Y $ in the matrix using $ \Sigma_{XY} $.
 
@@
 \subsubsection{Both $ X $ and $ Y $ are continuous}
-When both $ X $ and $ Y $ are continuous variables, Pearson's correlation
-coefficient is typically used. In the case when $ \bm{Z} = \emptyset $, the
-correlation coefficient is defined as:
+When both $ X $ and $ Y $ are continuous, Pearson's correlation coefficient is
+typically used. When $ \bm{Z} = \emptyset $, the correlation
+coefficient is defined as:
 
@@
 
-When $ \bm{Z} \neq \emptyset $, partial correlation coefficient
-can be used. One of the ways to estimate it is to use two regression
-models $ E_X: X \sim \bm{Z} $, $ E_Y: Y \sim \bm{Z} $, and compute the Pearson's 
-correlation coefficient between the residuals of these regression models:
+When $ \bm{Z} \neq \emptyset $, partial correlation coefficient can be used.
+This is estimated by fitting two regression models $ E_X: X \sim \bm{Z} $ and $
+E_Y: Y \sim \bm{Z} $, calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $
+R_Y = Y - E_Y(\bm{Y}) $, and then computing the correlation between the
+residuals:
 
 \begin{equation}
-	\begin{split}
-		R_X &= X - E_X(\bm{Z})  \;\;\;\;\;\;\;\;  R_Y = Y - E_Y(\bm{Z}) \\
-		\rho_{X, Y; \bm{Z}} &= \frac{\mathrm{cov}(R_X, R_Y)}{\sigma_{R_X} \sigma_{R_Y}}
-	\end{split}
+	\rho_{X, Y; \bm{Z}} = \rho_{R_X, R_Y}
 \end{equation}
@@
 
-In the case of all discrete variables, Cram\'er's V can used as a measure of
-association. When $ \bm{Z} = \emptyset $, Cram\'er's V can be computed from the 
-chi-squared test statistic as:
+When all variables are discrete, Cram\'er's V can used as a measure of
+association. When $ \bm{Z} = \emptyset $, Cram\'er's V is derived from the 
+chi-squared test statistic:
 
 \begin{equation}
-	\mathbb{V_{X, Y}} = \sqrt{\frac{\chi^2 / n}{\mathrm{min}(k-1, r-1)}}
+	\mathbf{V}_{X, Y} = \sqrt{\frac{\chi^2 / n}{\mathrm{min}(k-1, r-1)}}
 \end{equation}
@@
 where $ \chi^2 $ is the chi-squared test statistic for $ X $ and $ Y $, $ n $
-is the sample size and $ k $ and $ r $ are the number of categories of $ X $ and 
-$ Y $.
-
-However, when $ \bm{Z} \neq \emptyset $, we first split the dataset by unique
-values of $ \bm{Z} $, $ D = \{D_{\bm{Z} = \bm{Z_1}}, D_{\bm{Z} = \bm{Z_2}},
-\cdots, \} $  and then combine them as:
+is the sample size, and $ k $ and $ r $ are the number of categories in $ X $
+and $ Y $ respectively. When $ \bm{Z} \neq \emptyset $, we start by splitting
+the dataset by each unique combination of $ \bm{Z} $ into subsets,  $ \bm{Z} $,
+$ D = \{ D_{\bm{Z} = \bm{Z}_1}, D_{\bm{Z} = \bm{Z}_2}, \cdots, D_{\bm{Z} =
+\bm{Z}_k} \} $. Then we compute the Cram\'er's V for each of these smaller datasets
+and combine them as follows:
 
 \begin{equation}
-	\mathbb{V_{X, Y; Z}} = \sum_{D_{\bm{Z} = \bm{Z_i}}} \mathbb{V}_{X_{D_i}, Y_{D_i}}
+	\mathbf{V}_{X, Y; \bm{Z}} = \sum_{i=1}^{k} \left[ \mathbf{V}_{X, Y} \right]_{D_{\bm{Z} = \bm{Z}_i}} 
 \end{equation}
 
-\paragraph{$ X $ is ordinal and $ Y $ is continuous or ordinal}
-When one of the variables is ordinal and the other is continuous polyserial
-correlation can be used. When both the variables are ordinal polychoric
-correlation can be used. Both polychoric and polyserial correlation make the
-assumption that the ordinal variable has an underlying normal distributed
-continuous variable, and the ordinal variable has been generated by
-thresholding this latent variable. The method then tries to maximize the
-likelihood of data given polychoric or polyserial correlation coefficient. This 
-method gives us the covariance matrix, $\Sigma$ on the variables $ X $, $ Y $, and $ \bm{Z} $.
+\paragraph{$ X $ is ordinal, and $ Y $ and $ Z $  are continuous or ordinal}
+
+Polyserial (when one is ordinal and the other is continuous) and Polychoric
+(when both are ordinal) correlation have been used to estimate the covariance
+matrix between them \citep{Poon1987}. Both methods make the assumption that the
+observed ordinal variable is a result of thresholding a latent normally
+distributed continuous variable. Under this assumption, the methods then try to
+estimate the covariance matrix while maximizing the likelihood of the dataset.
+Using the estimated covariance matrix, $ \Sigma $ we can the compute Pearson's
+correlation coefficient. When $ \bm{Z} = \emptyset $,
 
 \begin{equation}
-	\rho_{X, Y} = \Sigma_{XY}
+	\rho_{X, Y} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}
 \end{equation}
@@
 different measures of conditional association. However, there is no unified
-measure that handles mixed data. Here, we introduce a measure of conditional
+measure that handles mixed data. In this section, we introduce a measure of conditional
 association for mixed data by extending the concept of partial correlation
@@
 Given a dataset $ D = (x, y, \bm{z}) $ on variables $ X $, $ Y $, and $ \bm{Z}
-$, our goal is to estimate the conditional association $ \rho(X, Y; \bm{Z}) $. 
+$, our goal is to estimate the conditional association $ \phi_{X, Y; \bm{Z}} $. 
 In the first step, we compute the residuals $ R_X $ and $ R_Y $ for variables
@@
 The second step is to quantify the association between these residual matrices.
-For this purpose we use canonical correlations that have been widely used to
+For this purpose we use canonical correlations \citep{Hotelling1936} that have been widely used to
 measure the association between sets of random variables.
@@
 	and $ \bm{V} = (V_1, V_2, \cdots, V_q) $, canonical correlation between
-	them, $\rho(\bm{U}, \bm{V}) $ is defined as:
+	them, $\upsilon(\bm{U}, \bm{V}) $ is defined as:
 
 	\begin{equation}
-		\rho(\bm{U}, \bm{V}) = \max_{a, b} Corr(a^T \bm{U}, b^T \bm{V})
+		% \nu_{\bm{U}, \bm{V}}= \max_{a, b} \mathrm{corr}(a^T \bm{U}, b^T \bm{V})
+		\rho_{\bm{U}, \bm{V}} = \max_{a, b} \frac{a^T \Sigma_{\bm{UV}} b}{\sqrt{a^T \Sigma_{\bm{UU}} a \cdot b^T \Sigma_{\bm{VV}} b}}
 	\end{equation}
@@
 \begin{itemize}
-	\item Wilks' Lambda: $\Lambda = \prod_{i}^{\min(p, q)} (1 - \phi_i^2) $
-	\item Roy's Largest Root: $ \theta = \max_i(\phi_i^2) $
-	\item Pillai's Trace: $ \tau = \sum_{i=1}^{\min(p, q)} \phi_i^2 $
+	\item Wilks' Lambda: $\Lambda = \prod_{i}^{\min(p, q)} (1 - \nu_i^2) $
+	\item Roy's Largest Root: $ \theta = \max_i(\nu_i^2) $
+	\item Pillai's Trace: $ \tau = \sum_{i=1}^{\min(p, q)} \nu_i^2 $
 \end{itemize}
@@
 signifies perfect linear relationship. This leads to our proposed measure of
-conditional association
+conditional association: We divide to make the measure normalized when comparing
+between pair of variables with different number of categories.
 
@@
 	\rho(X, Y; \bm{Z}) = \frac{1}{\min(\rvert R_X \rvert, \rvert R_Y \rvert)}
-	\sum_{i=1}^{\min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\phi_i(R_X, R_Y))^2
+	\sum_{i=1}^{\min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\nu_i(R_X, R_Y))^2
 \end{equation}

### 2025-02-06 paper/main.tex
Updates

@@
 	\item \textbf{Categorical:} We again start by training a probability
-		estimator $ p_X: X \sim \bm{Z} $, and obtain probability
-		estimates $ \hat{p}_X: p_X(\bm{Z}) $. Next, we dummy encode the
+		estimator $ p_X: x \sim \bm{z} $, and obtain probability
+		estimates $ \hat{p}_X: p_X(\bm{z}) $. Next, we dummy encode the
 		categorical variable and then compute the residuals as follows: 

### 2025-02-06 paper/main.tex
Updates

@@
 
-We use a slightly modified version of Pillai's Trace for our purpose for two
-main reasons: 1) It uses all the canonical correlation values, capturing the
-full extent of association, 2) Its interpretation is similar to Pearson's
-correlation coefficient, i.e., $ 0 $ signifies no association and $ 1 $
-signifies perfect linear relationship. This leads to our proposed measure of
-conditional association: We divide to make the measure normalized when comparing
-between pair of variables with different number of categories.
+We use a normalized version of Pillai's Trace for our purpose, given as:
 
@@
 \end{equation}
+
+We use Pillai's Trace because: 1) It uses all the canonical correlation values,
+capturing the full extent of association, 2) Its interpretation is similar to
+Pearson's correlation coefficient, i.e., $ 0 $ signifies no association and $ 1
+$ signifies perfect linear relationship. We use a normalized version because
+when comparing between pair of variables with different number of categories.
 

### 2025-02-06 paper/main.tex
Updates

@@
 \subsection{Notations} 
-\todo[inline]{Needs improvement}
+\todo[inline]{Need to rewrite this}
 We denote random variables with uppercase letters $ X $, and a set of random
@@
 when comparing between pair of variables with different number of categories.
-
 Our proposed measure has several desirable properties that make it well-suited
@@
 		interpretation.
-	\item \textbf{Independent of Sample Size: } 
+	% \item \textbf{Independent of Sample Size: } 
 	\item \textbf{Invariant to the Dummy Encoding: } For categorical

### 2025-02-06 paper/figures_v2.tex
Updates

\documentclass[tikz,margin=2mm]{standalone}
\pagestyle{empty}

%\usepackage{aistats2020}
\usepackage{amsmath}
\usepackage{bm}
\usetikzlibrary{positioning,calc,arrows,arrows.meta}

\begin{document}

	% PC algorithm MI test 400 samples
	\begin{tikzpicture}
	\begin{scope}[xscale=1.2, yscale=1.5]
		\tikzstyle{every node}=[align=center, inner sep=1pt]
		\node (ms)   at (0, 0)    {\footnotesize{Marital Status}};
		\node (rln)  at (0, -0.8) {\footnotesize{Relationship}};
		\node (sex)  at (0, -1.6) {\footnotesize{Sex}};
		\node (hrpw) at (0, -2.4) {\footnotesize{Hours Per Week}};

		\node (age) at (2.5, 0)    {\footnotesize{Age}};
		\node (inc) at (2.5, -0.8) {\footnotesize{Income}};
		\node (occ) at (2.5, -1.6) {\footnotesize{Occupation}};
		\node (wc)  at (2.5, -2.4) {\footnotesize{Work Class}};

		\node (nc)   at (5, 0)    {\footnotesize{Native Country}};
		\node (race) at (5, -0.8) {\footnotesize{Race}};
		\node (edu)  at (5, -1.6) {\footnotesize{Education}};

		\draw [red, transform canvas={xshift=0.15em}] (ms) to (rln);
		\draw [blue, transform canvas={xshift=-0.15em} ] (ms) to (rln);

		\draw [red] (rln.-70) to (sex.70);
		\draw [red] (inc) to (edu);
		\draw [red] (occ.-5) to (edu.185);

		\draw [blue, ->, >=stealth] (sex.110) to (rln.-110);
		\draw [blue, ->, >=stealth] (sex) to (occ);
		\draw [blue, ->, >=stealth] (occ) to (wc);
		\draw [blue, ->, >=stealth] (occ.5) to (edu.175);
		\draw [blue, ->, >=stealth] (wc) to (edu);
		\draw [blue, ->, >=stealth] (wc) to[bend right=40] (inc);
	\end{scope}
	\end{tikzpicture}

	% PC algorithm Q3 test 400 samples
	\begin{tikzpicture}
	\begin{scope}[xscale=1.2, yscale=1.5]
		\tikzstyle{every node}=[align=center, inner sep=0.2em]

		\node (ms)   at (0, 0)    {\footnotesize{Marital Status}};
		\node (rln)  at (0, -0.8) {\footnotesize{Relationship}};
		\node (sex)  at (0, -1.6) {\footnotesize{Sex}};
		\node (hrpw) at (0, -2.4) {\footnotesize{Hours Per Week}};

		\node (age) at (2.5, 0)    {\footnotesize{Age}};
		\node (inc) at (2.5, -0.8) {\footnotesize{Income}};
		\node (occ) at (2.5, -1.6) {\footnotesize{Occupation}};
		\node (wc)  at (2.5, -2.4) {\footnotesize{Work Class}};

		\node (nc)   at (5, 0)    {\footnotesize{Native Country}};
		\node (race) at (5, -0.8) {\footnotesize{Race}};
		\node (edu)  at (5, -1.6) {\footnotesize{Education}};

		\draw [red, transform canvas={xshift=0.15em}] (ms) to (rln);
		\draw [blue, transform canvas={xshift=-0.15em}] (ms) to (rln);

		\draw [red, ->, >=stealth, transform canvas={xshift=0.15em}] (sex) to (rln);
		\draw [blue,->, >=stealth, transform canvas={xshift=-0.15em}] (rln) to (sex);

		\draw [red, ->, >=stealth, transform canvas={xshift=0.15em}] (sex) to (hrpw);
		\draw [blue,->, >=stealth, transform canvas={xshift=-0.15em}] (hrpw) to (sex);

		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (ms) to (age);
		\draw [blue,->, >=stealth, transform canvas={yshift=-0.15em}] (age) to (ms);

		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (rln) to (age);
		\draw [blue,->, >=stealth, transform canvas={yshift=-0.15em}] (rln) to (age);

		\draw [red, ->, >=stealth, transform canvas={xshift=0.15em}] (inc) to (age);
		\draw [blue,->, >=stealth, transform canvas={xshift=-0.15em}] (inc) to (age);
		\draw [red, ->, >=stealth, transform canvas={xshift=0.15em, yshift=0.15em}] (inc) to (edu);
		\draw [blue,->, >=stealth, transform canvas={xshift=-0.15em, yshift=-0.15em}] (inc) to (edu);

		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (occ) to (hrpw);
		\draw [blue,->, >=stealth, transform canvas={yshift=-0.15em}] (occ) to (hrpw);

		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (occ) to (sex);
		\draw [blue,->, >=stealth, transform canvas={yshift=-0.15em}] (occ) to (sex);

		\draw [red, ->, >=stealth] (sex) to (wc);
		\draw [red, ->, >=stealth] (occ) to (ms);
		\draw [red] (nc) to (race);

		\draw [blue,->, >=stealth] (rln) to[bend right=50] (hrpw);
		\draw [blue,->, >=stealth] (hrpw) to (edu);

		\draw [blue,->, >=stealth] (inc) to (occ);
		\draw [blue,->, >=stealth] (occ) to (rln);
		\draw [blue,->, >=stealth] (occ) to (wc);
		\draw [blue,->, >=stealth] (occ) to (edu);

	\end{scope}
	\end{tikzpicture}

	% Hill Climb with BIC Score Q3 test 400 samples
	\begin{tikzpicture}
	\begin{scope}[xscale=0.1]
		\tikzstyle{every node}=[align=center, inner sep=1pt]

		\node (ms)   at (0, 0)    {\footnotesize{Marital Status}};
		\node (rln)  at (0, -0.8) {\footnotesize{Relationship}};
		\node (sex)  at (0, -1.6) {\footnotesize{Sex}};
		\node (hrpw) at (0, -2.4) {\footnotesize{Hours Per Week}};

		\node (age) at (1.5, 0)    {\footnotesize{Age}};
		\node (inc) at (1.5, -0.8) {\footnotesize{Income}};
		\node (occ) at (1.5, -1.6) {\footnotesize{Occupation}};
		\node (wc)  at (1.5, -2.4) {\footnotesize{Work Class}};

		\node (nc)   at (3, 0)    {\footnotesize{Native Country}};
		\node (race) at (3, -0.8) {\footnotesize{Race}};
		\node (edu)  at (3, -1.6) {\footnotesize{Education}};

		\draw [] (ms) to (rln);
		\draw [] (rln) to (sex);
		\draw [->, >=stealth] (sex) to (hrpw);

		\draw [->, >=stealth] (ms) to (age);
		\draw [] (rln) to (inc);
		\draw [] (sex) to (occ);
	\end{scope}
	\end{tikzpicture}

	% Hill Climb with BIC Score Q3 test 800 samples
	\begin{tikzpicture}
	\begin{scope}[xscale=0.9]
		\tikzstyle{every node}=[align=center, inner sep=1pt]

		\node (ms)   at (0, 0)    {\footnotesize{MS}};
		\node (rln)  at (0, -0.8) {\footnotesize{Rln.}};
		\node (sex)  at (0, -1.6) {\footnotesize{Sex}};
		\node (hrpw) at (0, -2.4) {\footnotesize{HrPW}};

		\node (age) at (1.5, 0)    {\footnotesize{Age}};
		\node (inc) at (1.5, -0.8) {\footnotesize{Inc.}};
		\node (occ) at (1.5, -1.6) {\footnotesize{Occ.}};
		\node (wc)  at (1.5, -2.4) {\footnotesize{WC}};

		\node (nc)   at (3, 0)    {\footnotesize{NC}};
		\node (race) at (3, -0.8) {\footnotesize{Race}};
		\node (edu)  at (3, -1.6) {\footnotesize{Edu.}};

		\draw [] (ms) to (rln);
		\draw [] (rln) to (sex);
		\draw [->, >=stealth] (rln) to[bend right=40] (hrpw);

		\draw [->, >=stealth] (ms) to (age);
		\draw [] (rln) to (inc);
		\draw [] (sex) to (occ);
	\end{scope}
	\end{tikzpicture}

\end{document}

### 2025-02-06 code/fig4/make_plots.R
Fixes bug

@@ -36,7 +36,7 @@ p_ll <- ggplot(d_results, aes(x=x, y=ll, color=algo)) +
 	labs(y="Log-Likelihood") +
 	theme(legend.position='top',
 	      legend.title=element_blank(),
-	      legned.text=element_blank())
+	      legend.text=element_blank())
 
 ggsave('unexplained_effect.pdf', p_unexplained_effect, height=2.5, width=1.6, units='in')
 ggsave('ll.pdf', p_ll, height=2.5, width=1.6, units='in')

### 2025-02-06 paper/main.tex
Updates
updates

@@
     \begin{subfigure}{0.25 \textwidth}
-    	\includegraphics[page=1]{figures.pdf}
+	\centering
+    	\includegraphics[page=1]{figures_v2.pdf}
     	\caption{$ N = 400 $}
-    \end{subfigure}%
-    \begin{subfigure}{0.25 \textwidth}
-    	\includegraphics[page=2]{figures.pdf}
-    	\caption{$ N = 800$ }
     \end{subfigure}
-    \begin{subfigure}{0.25 \textwidth}
-    	\includegraphics[page=5]{figures.pdf}
-    	\caption{$ N = 400 $ }
-    \end{subfigure}%
     \begin{subfigure}{0.25\textwidth}
-    	\includegraphics[page=6]{figures.pdf}
+    	\includegraphics[page=2]{figures_v2.pdf}
     	\caption{ $ N = 800$ }
@@
     \begin{subfigure}{0.25 \textwidth}
-    	\includegraphics[page=3]{figures.pdf}
+    	\includegraphics[page=3]{figures_v2.pdf}
     	\caption{ $ N = 400$ }
@@
     \begin{subfigure}{0.25\textwidth}
-    	\includegraphics[page=4]{figures.pdf}
+    	\includegraphics[page=4]{figures_v2.pdf}
     	\caption{$ N = 800$ }
@@
 \label{sec:background}
-\subsection{Notations} 
 \todo[inline]{Need to rewrite this}
@@
 		estimates $ \hat{p}_X: p_X(\bm{z}) $. Next, we dummy encode the
-		categorical variable and then compute the residuals as follows: 
-		$$ R_{x_i} = x_i - \hat{p}(\bm{z}_i) $$
+		categorical variable, resulting in a binary vector and then
+		compute the residuals as follows: 
+		$$ R_{x_i} = x_i - \hat{p}_X(\bm{z}_i) $$
 \end{enumerate}
@@
 We repeat the above residualization step for both the variables $ X $ and $ Y $
-obtaining residual matrices $ R_X $ and $ R_Y $. The type of variable determines
+obtaining residual matrices $ R_x $ and $ R_y $. The type of variable determines
 the shape of these matrices. If the variable is continuous or ordinal, the 
@@
 \begin{equation}
-	\rho(X, Y; \bm{Z}) = \frac{1}{\min(\rvert R_X \rvert, \rvert R_Y \rvert)}
-	\sum_{i=1}^{\min(\rvert R_X \rvert, \rvert R_Y \rvert)} (\nu_i(R_X, R_Y))^2
+	\rho(X, Y; \bm{Z}) = \frac{1}{\min(\rvert R_x \rvert, \rvert R_y \rvert)}
+	\sum_{i=1}^{\min(\rvert R_x \rvert, \rvert R_y \rvert)} (\nu_i(R_x, R_y))^2
 \end{equation}

### 2025-02-06 paper/main.tex
Updates

@@
 \begin{equation}
-	\rho_{X, Y} = \frac{\mathrm{cov}(X, Y)}{\sigma_X \sigma_Y}
+	r_{X, Y} = \frac{\mathrm{cov}(X, Y)}{\sigma_X \sigma_Y}
 \end{equation}
@@
 \begin{equation}
-	\rho_{X, Y; \bm{Z}} = \rho_{R_X, R_Y}
+	r_{X, Y; \bm{Z}} = r_{R_X, R_Y}
 \end{equation}
@@
 \begin{equation}
-	\rho_{X, Y} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}
+	r_{X, Y} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}
 \end{equation}
@@
 \begin{equation}
-	\rho_{X, Y; \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
+	r_{X, Y; \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
 \end{equation}
@@
 	and $ \bm{V} = (V_1, V_2, \cdots, V_q) $, canonical correlation between
-	them, $\upsilon(\bm{U}, \bm{V}) $ is defined as:
+	them, $\rho_{\bm{U}, \bm{V}} $ is defined as:
 
@@
 \begin{itemize}
-	\item Wilks' Lambda: $\Lambda = \prod_{i}^{\min(p, q)} (1 - \nu_i^2) $
-	\item Roy's Largest Root: $ \theta = \max_i(\nu_i^2) $
-	\item Pillai's Trace: $ \tau = \sum_{i=1}^{\min(p, q)} \nu_i^2 $
+	\item Wilks' Lambda: $\Lambda = \prod_{i}^{\min(p, q)} (1 - \rho_i^2) $
+	\item Roy's Largest Root: $ \theta = \max_i(\rho_i^2) $
+	\item Pillai's Trace: $ \tau = \sum_{i=1}^{\min(p, q)} \rho_i^2 $
 \end{itemize}
@@
 \begin{equation}
-	\rho(X, Y; \bm{Z}) = \frac{1}{\min(\rvert R_x \rvert, \rvert R_y \rvert)}
-	\sum_{i=1}^{\min(\rvert R_x \rvert, \rvert R_y \rvert)} (\nu_i(R_x, R_y))^2
+	\tau_{X, Y; \bm{Z}} = \frac{1}{\min(\rvert R_x \rvert, \rvert R_y \rvert)}
+	\sum_{i=1}^{\min(\rvert R_x \rvert, \rvert R_y \rvert)} (\rho_{R_x, R_y})_i^2
 \end{equation}

### 2025-02-06 paper/main.tex
Update
Updates

@@
 \begin{figure}[t!]
-    \begin{subfigure}{0.25 \textwidth}
+    \begin{subfigure}{0.5 \textwidth}
 	\centering
     	\includegraphics[page=1]{figures_v2.pdf}
-    	\caption{$ N = 400 $}
+    	\caption{PC algorithm with mutual information test. Red edges: $ N=400 $, Blue edges: $ N=800 $}
     \end{subfigure}
-    \begin{subfigure}{0.25\textwidth}
+    \begin{subfigure}{0.5\textwidth}
     	\includegraphics[page=2]{figures_v2.pdf}
-    	\caption{ $ N = 800$ }
+    	\caption{PC algorithm with a residualization based test. Red edges $N=400$, Blue edges: $ N=800 $}
     \end{subfigure}
-    \begin{subfigure}{0.25 \textwidth}
+    \begin{subfigure}{0.5\textwidth}
     	\includegraphics[page=3]{figures_v2.pdf}
-    	\caption{ $ N = 400$ }
-    \end{subfigure}%
-    \begin{subfigure}{0.25\textwidth}
-    	\includegraphics[page=4]{figures_v2.pdf}
-    	\caption{$ N = 800$ }
+    	\caption{PC algorithm with Hill Climb Search with BIC score}
     \end{subfigure}
@@
 		coefficient.
+		\todo[inline]{Verify if this is correct}
 	\item \textbf{Related to Cram\'er's V: } 
@@
 
-\section{Using Measure of Association for DAG Construction}
+\section{Using Measure of Association for Ranking and DAG Construction}
 \label{sec:modification}

### 2025-02-06 paper/appendix.tex
Move section 3 to appendix

@@
 \end{figure}
+
+\section{A Generalized Measure of Conditional Association}
+\label{sec:mixed_association}
+
+In the previous section, we discussed how different types of variables require
+different measures of conditional association. However, there is no unified
+measure that handles mixed data. In this section, we introduce a measure of conditional
+association for mixed data by extending the concept of partial correlation
+coefficient (commonly used for continuous variables) to mixed data. Similar to
+partial correlation coefficient, our method integrates a mixed data
+residualization method \citep{Ankan2023} with Pillai's Trace
+\citep{Pillai1955}, a multivariate measure of association based on canonical
+correlations.
+
+Given a dataset $ D = (x, y, \bm{z}) $ on variables $ X $, $ Y $, and $ \bm{Z}
+$, our goal is to estimate the conditional association $ \phi_{X, Y; \bm{Z}} $. 
+In the first step, we compute the residuals $ R_X $ and $ R_Y $ for variables
+$ X $ and $ Y $ respectively. Depending on the type of variable the residual
+is computed as follows:
+
+\begin{enumerate}
+	\item \textbf{Continuous:} We train a model, $ E_X: x \sim
+		\bm{z} $. The residuals are then computed by taking the difference
+		between the true and the predicted values using $ E_X $. 
+		$$ R_{x_i} = x_i - E_X(\bm{z}_i) $$
+	\item \textbf{Ordinal:} We start by training a probability estimator, $
+		p_X: x \sim \bm{z} $, and then use the estimated probabilities, 
+		$ \hat{p}_X(x) $ to compute the residuals:
+		$$ R_{x_i} = \hat{p}_X(X < x_i) - \hat{p}_X(X > x_i) $$
+	\item \textbf{Categorical:} We again start by training a probability
+		estimator $ p_X: x \sim \bm{z} $, and obtain probability
+		estimates $ \hat{p}_X: p_X(\bm{z}) $. Next, we dummy encode the
+		categorical variable, resulting in a binary vector and then
+		compute the residuals as follows: 
+		$$ R_{x_i} = x_i - \hat{p}_X(\bm{z}_i) $$
+\end{enumerate}
+
+We have the option here to choose the estimators based on the characteristics
+such as distribution, type of relationship, and so on of our dataset.
+Non-parametric ensemble estimators such as Random Forest and XGBoost are robust 
+for diverse data types and complex relationships. For linear relationships,
+simpler models such as linear regression and its variants may suffice.
+
+We repeat the above residualization step for both the variables $ X $ and $ Y $
+obtaining residual matrices $ R_x $ and $ R_y $. The type of variable determines
+the shape of these matrices. If the variable is continuous or ordinal, the 
+residual matrix is of shape $ (n \times 1 ) $ and if the variable is categorical,
+we get a residual matrix of shape $ (x \times (k - 1)) $, where $ k $ is the number
+of categories of the variable.
+
+The second step is to quantify the association between these residual matrices.
+For this purpose we use canonical correlations \citep{Hotelling1936} that have been widely used to
+measure the association between sets of random variables.
+
+\begin{definition}
+	Given two sets of random variables $ \bm{U} = (U_1, U_2, \cdots, U_p) $
+	and $ \bm{V} = (V_1, V_2, \cdots, V_q) $, canonical correlation between
+	them, $\rho_{\bm{U}, \bm{V}} $ is defined as:
+
+	\begin{equation}
+		% \nu_{\bm{U}, \bm{V}}= \max_{a, b} \mathrm{corr}(a^T \bm{U}, b^T \bm{V})
+		\rho_{\bm{U}, \bm{V}} = \max_{a, b} \frac{a^T \Sigma_{\bm{UV}} b}{\sqrt{a^T \Sigma_{\bm{UU}} a \cdot b^T \Sigma_{\bm{VV}} b}}
+	\end{equation}
+
+	where $ a $ and $ b $ are vectors of coefficients that maximize the correlations
+	between the linear combinations of $ a^T \bm{U} $ and $ b^T \bm{V} $.
+\end{definition}
+
+Canonical correlations generalize the concept of correlation coefficients to
+multi-dimensional variables. It finds orthogonal linear transformations $ a $
+and $ b $ that maximized the correlation between the transformed variables $
+a^T \bm{U} $ and $ b^T \bm{V} $. This yields a vector of correlation
+coefficient values of size $ \min(p, q) $, representing the correlation
+coefficient of each pair of transformed variables. Notably, Pearson's
+correlation coefficient is a special case of canonical correlations when $ p =
+q = 1 $.
+
+Several measures of association have been derived from canonical correlations, such as:
+\begin{itemize}
+	\item Wilks' Lambda: $\Lambda = \prod_{i}^{\min(p, q)} (1 - \rho_i^2) $
+	\item Roy's Largest Root: $ \theta = \max_i(\rho_i^2) $
+	\item Pillai's Trace: $ \tau = \sum_{i=1}^{\min(p, q)} \rho_i^2 $
+\end{itemize}
+
+We use a normalized version of Pillai's Trace for our purpose, given as:
+
+\begin{equation}
+	\tau_{X, Y; \bm{Z}} = \frac{1}{\min(\rvert R_x \rvert, \rvert R_y \rvert)}
+	\sum_{i=1}^{\min(\rvert R_x \rvert, \rvert R_y \rvert)} (\rho_{R_x, R_y})_i^2
+\end{equation}
+
+We use Pillai's Trace because: 1) It uses all the canonical correlation values,
+capturing the full extent of association, 2) Its interpretation is similar to
+Pearson's correlation coefficient, i.e., $ 0 $ signifies no association and $ 1
+$ signifies perfect linear relationship. We use a normalized version because
+when comparing between pair of variables with different number of categories.
+Our proposed measure has several desirable properties that make it well-suited
+for our application:
+
+\begin{enumerate}
+	\item \textbf{Bounded: } The measure is bounded between $ 0 $ (no
+		association) to $ 1 $ (perfect linear association), simplifying
+		interpretation.
+	% \item \textbf{Independent of Sample Size: } 
+	\item \textbf{Invariant to the Dummy Encoding: } For categorical
+		variables, residuals are computed using dummy encoding. Since
+		canonical correlations identifies linear combinations across
+		columns, this measure is invariant to the specific dummy
+		encoding scheme used.
+	\item \textbf{Equivalent to Partial Correlation for Continuous Variables: }
+		When both $ X $ and $ Y $ are continuous variables, our measure
+		is equivalent to the absolute value of partial correlation coefficient.
+	\item \textbf{Equivalent to polychoric and polyserial correlation for
+			ordinal and continuous variables: }
+		Under the assumption that the ordinal variable is generated by
+		discretizing an underlying gaussian continuous variable, this
+		effect size is equivalent to polychoric and polyserial
+		correlation. As both of them recover the Pearson correlation
+		coefficient.
+		\todo[inline]{Verify if this is correct}
+	\item \textbf{Related to Cram\'er's V: } 
+\end{enumerate}
+
+In essence, our measure of conditional association extends existing metrics to mixed data, providing 
+a unified measure that is interpretable.

### 2025-02-06 paper/main.tex
Move section 3 to appendix

@@
 
-\section{A Generalized Measure of Conditional Association}
-\label{sec:mixed_association}
-
-In the previous section, we discussed how different types of variables require
-different measures of conditional association. However, there is no unified
-measure that handles mixed data. In this section, we introduce a measure of conditional
-association for mixed data by extending the concept of partial correlation
-coefficient (commonly used for continuous variables) to mixed data. Similar to
-partial correlation coefficient, our method integrates a mixed data
-residualization method \citep{Ankan2023} with Pillai's Trace
-\citep{Pillai1955}, a multivariate measure of association based on canonical
-correlations.
-
-Given a dataset $ D = (x, y, \bm{z}) $ on variables $ X $, $ Y $, and $ \bm{Z}
-$, our goal is to estimate the conditional association $ \phi_{X, Y; \bm{Z}} $. 
-In the first step, we compute the residuals $ R_X $ and $ R_Y $ for variables
-$ X $ and $ Y $ respectively. Depending on the type of variable the residual
-is computed as follows:
-
-\begin{enumerate}
-	\item \textbf{Continuous:} We train a model, $ E_X: x \sim
-		\bm{z} $. The residuals are then computed by taking the difference
-		between the true and the predicted values using $ E_X $. 
-		$$ R_{x_i} = x_i - E_X(\bm{z}_i) $$
-	\item \textbf{Ordinal:} We start by training a probability estimator, $
-		p_X: x \sim \bm{z} $, and then use the estimated probabilities, 
-		$ \hat{p}_X(x) $ to compute the residuals:
-		$$ R_{x_i} = \hat{p}_X(X < x_i) - \hat{p}_X(X > x_i) $$
-	\item \textbf{Categorical:} We again start by training a probability
-		estimator $ p_X: x \sim \bm{z} $, and obtain probability
-		estimates $ \hat{p}_X: p_X(\bm{z}) $. Next, we dummy encode the
-		categorical variable, resulting in a binary vector and then
-		compute the residuals as follows: 
-		$$ R_{x_i} = x_i - \hat{p}_X(\bm{z}_i) $$
-\end{enumerate}
-
-We have the option here to choose the estimators based on the characteristics
-such as distribution, type of relationship, and so on of our dataset.
-Non-parametric ensemble estimators such as Random Forest and XGBoost are robust 
-for diverse data types and complex relationships. For linear relationships,
-simpler models such as linear regression and its variants may suffice.
-
-We repeat the above residualization step for both the variables $ X $ and $ Y $
-obtaining residual matrices $ R_x $ and $ R_y $. The type of variable determines
-the shape of these matrices. If the variable is continuous or ordinal, the 
-residual matrix is of shape $ (n \times 1 ) $ and if the variable is categorical,
-we get a residual matrix of shape $ (x \times (k - 1)) $, where $ k $ is the number
-of categories of the variable.
-
-The second step is to quantify the association between these residual matrices.
-For this purpose we use canonical correlations \citep{Hotelling1936} that have been widely used to
-measure the association between sets of random variables.
-
-\begin{definition}
-	Given two sets of random variables $ \bm{U} = (U_1, U_2, \cdots, U_p) $
-	and $ \bm{V} = (V_1, V_2, \cdots, V_q) $, canonical correlation between
-	them, $\rho_{\bm{U}, \bm{V}} $ is defined as:
-
-	\begin{equation}
-		% \nu_{\bm{U}, \bm{V}}= \max_{a, b} \mathrm{corr}(a^T \bm{U}, b^T \bm{V})
-		\rho_{\bm{U}, \bm{V}} = \max_{a, b} \frac{a^T \Sigma_{\bm{UV}} b}{\sqrt{a^T \Sigma_{\bm{UU}} a \cdot b^T \Sigma_{\bm{VV}} b}}
-	\end{equation}
-
-	where $ a $ and $ b $ are vectors of coefficients that maximize the correlations
-	between the linear combinations of $ a^T \bm{U} $ and $ b^T \bm{V} $.
-\end{definition}
-
-Canonical correlations generalize the concept of correlation coefficients to
-multi-dimensional variables. It finds orthogonal linear transformations $ a $
-and $ b $ that maximized the correlation between the transformed variables $
-a^T \bm{U} $ and $ b^T \bm{V} $. This yields a vector of correlation
-coefficient values of size $ \min(p, q) $, representing the correlation
-coefficient of each pair of transformed variables. Notably, Pearson's
-correlation coefficient is a special case of canonical correlations when $ p =
-q = 1 $.
-
-Several measures of association have been derived from canonical correlations, such as:
-\begin{itemize}
-	\item Wilks' Lambda: $\Lambda = \prod_{i}^{\min(p, q)} (1 - \rho_i^2) $
-	\item Roy's Largest Root: $ \theta = \max_i(\rho_i^2) $
-	\item Pillai's Trace: $ \tau = \sum_{i=1}^{\min(p, q)} \rho_i^2 $
-\end{itemize}
-
-We use a normalized version of Pillai's Trace for our purpose, given as:
-
-\begin{equation}
-	\tau_{X, Y; \bm{Z}} = \frac{1}{\min(\rvert R_x \rvert, \rvert R_y \rvert)}
-	\sum_{i=1}^{\min(\rvert R_x \rvert, \rvert R_y \rvert)} (\rho_{R_x, R_y})_i^2
-\end{equation}
-
-We use Pillai's Trace because: 1) It uses all the canonical correlation values,
-capturing the full extent of association, 2) Its interpretation is similar to
-Pearson's correlation coefficient, i.e., $ 0 $ signifies no association and $ 1
-$ signifies perfect linear relationship. We use a normalized version because
-when comparing between pair of variables with different number of categories.
-Our proposed measure has several desirable properties that make it well-suited
-for our application:
-
-\begin{enumerate}
-	\item \textbf{Bounded: } The measure is bounded between $ 0 $ (no
-		association) to $ 1 $ (perfect linear association), simplifying
-		interpretation.
-	% \item \textbf{Independent of Sample Size: } 
-	\item \textbf{Invariant to the Dummy Encoding: } For categorical
-		variables, residuals are computed using dummy encoding. Since
-		canonical correlations identifies linear combinations across
-		columns, this measure is invariant to the specific dummy
-		encoding scheme used.
-	\item \textbf{Equivalent to Partial Correlation for Continuous Variables: }
-		When both $ X $ and $ Y $ are continuous variables, our measure
-		is equivalent to the absolute value of partial correlation coefficient.
-	\item \textbf{Equivalent to polychoric and polyserial correlation for
-			ordinal and continuous variables: }
-		Under the assumption that the ordinal variable is generated by
-		discretizing an underlying gaussian continuous variable, this
-		effect size is equivalent to polychoric and polyserial
-		correlation. As both of them recover the Pearson correlation
-		coefficient.
-		\todo[inline]{Verify if this is correct}
-	\item \textbf{Related to Cram\'er's V: } 
-\end{enumerate}
-
-In essence, our measure of conditional association extends existing metrics to mixed data, providing 
-a unified measure that is interpretable.
-
 \section{Using Measure of Association for Ranking and DAG Construction}
@@
 rather than just a relative comparison between models.
-
-\section{Web Tool}
-\label{sec:web}
-To enable users to apply this method to their own datasets, we developed an
-interactive web-based tool (shown in Fig.~\ref{fig:web}) for constructing
-models. Users begin by uploading their dataset, which initializes an empty DAG
-with nodes corresponding to the dataset\'s variables. They can then specify a
-p-value threshold and a measure of association threshold.
-
-Using the specified thresholds, the tool visually highlights unexplained
-correlations by displaying red edges between variables where correlation exists
-in the data but the current DAG is not able to explain. The thickness of these
-edges represents the strength of the correlation, helping users prioritize
-which edges to add. Similarly, if an edge in the graph is detected to be
-unnecessary, it is highlighted in black. Based on this information, users can
-select to remove unnecessary edge or select a potential edge to add and specify
-the orientation of the edge. The tool computes Shipley’s C \citep{Shipley2000}
-value at each change to determine the overall fit of the model to the data.
-Once satisfied with the constructed DAG, users can export the model for further
-analysis.
-
-\begin{figure}[t!]
-	\centering
-	\includegraphics[scale=0.4]{../code/plots/web_tool_full.png}
-	\caption{A screenshot of the web tool for constructing the model. Users
-		can upload their dataset after which the tool creates an empty
-		graph and shows all pair of variables which are associated in
-		the model using undirected red edges with the strength of
-		association represented using edge width. Users can then
-		iteratively add edges to the model (shown in green) while
-		deciding the edge orientation based on domain knowledge.
-		Unnecessary edges are shown in black.}
-	\label{fig:web}
-\end{figure}
 
@@
 
+\section{Web Tool}
+\label{sec:web}
+To enable users to apply this method to their own datasets, we developed an
+interactive web-based tool (shown in Fig.~\ref{fig:web}) for constructing
+models. Users begin by uploading their dataset, which initializes an empty DAG
+with nodes corresponding to the dataset\'s variables. They can then specify a
+p-value threshold and a measure of association threshold.
+
+Using the specified thresholds, the tool visually highlights unexplained
+correlations by displaying red edges between variables where correlation exists
+in the data but the current DAG is not able to explain. The thickness of these
+edges represents the strength of the correlation, helping users prioritize
+which edges to add. Similarly, if an edge in the graph is detected to be
+unnecessary, it is highlighted in black. Based on this information, users can
+select to remove unnecessary edge or select a potential edge to add and specify
+the orientation of the edge. The tool computes Shipley’s C \citep{Shipley2000}
+value at each change to determine the overall fit of the model to the data.
+Once satisfied with the constructed DAG, users can export the model for further
+analysis.
+
+\begin{figure}[t!]
+	\centering
+	\includegraphics[scale=0.4]{../code/plots/web_tool_full.png}
+	\caption{A screenshot of the web tool for constructing the model. Users
+		can upload their dataset after which the tool creates an empty
+		graph and shows all pair of variables which are associated in
+		the model using undirected red edges with the strength of
+		association represented using edge width. Users can then
+		iteratively add edges to the model (shown in green) while
+		deciding the edge orientation based on domain knowledge.
+		Unnecessary edges are shown in black.}
+	\label{fig:web}
+\end{figure}
+
 \begin{figure}[t!]

### 2025-02-07 paper/main.tex
Merge branch

@@
 \usepackage{subcaption}
-% \usepackage{fancyvrb}
-\usepackage[ruled]{algorithm2e}
+\usepackage[linesnumbered, ruled]{algorithm2e}
 
@@
 
-\section{Using Measure of Association for Ranking and DAG Construction}
+\section{DAG Structure Learning using Ancestral Oracles}
 \label{sec:modification}
+
+Our approach to causal structure learning combines domain knowledge with data-driven
+insights in a manner that is based on the following base assumptions: 
+(1) Domain experts are good at determining causal directions between variables if 
+there is a clear causal direction between them. (2) Domain experts are less good
+at identifying cases where there is no causal relationship between the variables,
+since \emph{potential} causal relationships can often be argued for anyway.
+(3) Domain experts are even less good at distinguishing direct from indirect 
+effects, since the question whether a direct effect exists between two variables
+depends on all other variables present in the graph.
+
+We therefore consider the following two models of a domain experts: Given two variables
+\emph{X} and \emph{Y} that are part of an unknown causal DAG $G$, an \emph{strong ancestral oracle} 
+$\mathcal{A}_G$ is defined as:
+$$\mathcal{A}_G(X,Y)=\begin{cases}
+ X \to Y & \textrm{if } X \in \textrm{An}_G(Y) \\
+ X \gets Y & \textrm{if } Y \in \textrm{An}_G(X) \\
+ \textrm{None} & \textrm{otherwise} \\
+\end{cases}$$
+
+whereas a \emph{weak ancestral oracle} is defined as:
+$$
+\mathcal{A}_G(X,Y)=\begin{cases}
+ X \to Y & \textrm{if } X \in \textrm{An}_G(Y) \\
+ X \gets Y & \textrm{if } Y \in \textrm{An}_G(X) \\
+ \textrm{rand}(X \gets Y, X \to Y) & \textrm{otherwise} \\
+\end{cases}
+$$
+
+That is, our ancestral oracles never distinguish whether the relationships between variables is
+direct or indirect; and the weak ancestral oracle is even unable to detect cases where there is no
+cause-effect relationship between two variables. Using such oracles for DAG construction therefore
+inherently risks that superfluous edges are created. In our approach, such superfluous edges will
+be detected and removed by utilizing the data.
+
+To model how our learning procedure interacts with the data, we
+assume that we have access to a second oracle $\mathcal{D}_G$ than can answer
+d-separation queries with respect to the unknown target graph: 
+$\mathcal{D}_G(X,Y,\mathbf{Z})=1$ iff $X \ci_G Y \mid \mathbf{Z}$. 
+These are the standard oracles considered in constraint-based structure learning algorithms,
+such as the PC algorithm \cite{Kalisch2012}. Based on this oracle, we now define two procedures
+that iteratively change a current DAG structure.
+
+Our structure learning algorithm is based on two core procedures. \textsc{Expand} 
+(Algorithm~\ref{algo:expand}) uses conditional independence information to search for
+missing edges in the graph and then uses domain knowledge to orient them. 
+\textsc{Prune} (Algorithm~\ref{algo:prune}) uses conditional independence information
+to remove superfluous edges from the graph.
+
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+\SetKwFunction{Expand}{Expand}
+\SetKw{KwGoTo}{go to}
+\SetKwProg{Fn}{Function}{:}{}
+\Fn{{\sc Expand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$B=\emptyset$,$k=\infty$)}{
+    let $L \gets \{\}$\;
+    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
+	$\{X,Y\} \notin B$}{
+        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
+        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
+            $L \gets L \cup \mathcal{A}_G(X, Y)$ \;
+        }
+	\If{$|L|\geq k$}{\KwGoTo 12\;}
+    }
+    add all edges in $L$ to $E$\;
+    \Return{$(V,E)$}\;
+}
+\caption{Adding edges based on data an domain knowledge}
+\label{algo:expand}
+\end{algorithm}
+
+The function $\textsc{Expand}$ takes an initial list of edges and searches for 
+any vertex pairs that are not connected but where the d-separation oracle 
+indicates  that there is still a residual association not explained by the other variables.
+An optional parameter $B$, the empty set by default, allows to specify a ``black list''
+of vertex pairs that must not be connected, and another optional parameter $k$
+specifies the maximum amount of edges to be added by this procedure.
+
+\begin{proposition}
+For a conditional independence oracle
+ $\mathcal{D}_G$ and a strong ancestral oracle $\mathcal{A}_G$, 
+$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_G)=G^+$.
+\end{proposition}
+
+\begin{proposition}
+For a conditional independence oracle
+ $\mathcal{D}_G$ and a weak ancestral oracle $\mathcal{A}_G$, 
+$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_G)\supseteq G^+$.
+\end{proposition}
+
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+
+\SetKwProg{Fn}{Function}{:}{}
+\Fn{{\sc Prune}($V$,$E$,$\mathcal{D}_G$)}{
+    let $L \gets \{\}$\;
+    \ForEach{$X \to Y \in E$}{
+        remove $X \to Y$ from $E$\;
+        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
+        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 1$}{
+            $L \gets L \cup  \{X \to Y\}$ \;
+        }
+    }
+    remove all edges in $L$ from $G$\;
+    \Return{$(V,E)$}\;
+}
+\caption{Pruning superfluous edges}
+\label{algo:prune}
+\end{algorithm}
+
+\begin{proposition}
+Consider two DAGs $G=(V,E)$ and $G'=(V,E')$ where $E \subseteq E'$. 
+Then $\textsc{Prune}(G',\mathcal{D}_G)=G$.
+\label{prop:prune}
+\end{proposition}
+
+By combining the previous three propositions, we immediately obtain the following:
+
+\begin{theorem}
+Given a d-separation oracle $\mathcal{D}_G$ and a strong or weak 
+ancestral oracle $\mathcal{A}_G$, 
+$\textsc{Prune}(\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_G),\mathcal{D}_G)=G.$
+\end{theorem}
+
+Since we require expert knowledge only in the $\textsc{Expand}$ operation, we may try
+to be more economical by asking fewer questions at a time and interleaving expansion and
+pruning steps. This leads us to the following, more incremental DAG construction algorithm.
+
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+\SetKwProg{Fn}{Function}{:}{}
+
+\Fn{{\sc ExpertInLoop}($V,\mathcal{D}_G,\mathcal{A}_G$)}{
+$E_p \gets \emptyset$ \tcc{Current edges} 
+$B \gets \emptyset$ \tcc{Edges that were pruned} 
+\Repeat{ $E=E_p$ }{
+	$E \gets E_p$ \;
+	$(V,E) \gets \textsc{Expand}(V,E,\mathcal{D}_G,\mathcal{A}_G,B,1)$ \;
+	$(V,E_p) \gets \textsc{Prune}(V,E,\mathcal{D}_G)$ \;
+	$B \gets B \cup \{ E_p \setminus E \} $
+}
+\Return{(V,E)}
+}
+
+\caption{Iterative structure learning with expert in the loop}
+\label{algo:expert}
+\end{algorithm}
+
+\begin{theorem}
+Let $G=(V,^*E)$ be a DAG, $\mathcal{D}_G$ a d-separation oracle for $G$, and 
+$\mathcal{A}_G$ a strong or weak ancestral oracle for $G$. Then 
+$\textsc{ExpertInLoop}(V,\mathcal{D}_G,\mathcal{A}_G)=G$.
+\end{theorem}
+
+\begin{proof}
+The loop in Algorithm~\ref{algo:expert} terminates if and only if $E=E^*$. If the loop does not 
+terminate, then a new edge has been added in line 6, or one or more edges were pruned in line 7.
+Every edge can be added at 
+most once and pruned at most once. Therefore, the algorithm always terminates
+after at most $|V|(|V|-1)+1$ iterations of the loop.
+For every edge $e=X\to Y $ in the skeleton of $G$, $\mathcal{D}_G(X,Y,\mathbf{Z})=0$ irrespective of $\mathbf{Z}$.
+Therefore, $e$ must be added to $E$ in some iteration, and can never be pruned afterwards.
+Therefore, after some iteration, $(V,E)$ must be a supergraph of $G$ after executing line 6, and this will be pruned to the real graph $G$ in line 7 (Proposition~\ref{prop:prune}).
+In the next iteration, no further changes are made, and the loop terminates.
+\end{proof}
 

### 2025-02-07 paper/main.tex
Remove older text

@@
 
-Using the mixed data measure of association presented in the last section, we
-now present a method to rank modifications to a given DAG.
-
-Using this mixed data measure of association, we now propose a method for
-iteratively constructing DAGs while incorporating expert knowledge. This
-approach is based on implied CI based testing where we use the measure of
-association to rank the failing CI tests. This ranking allows us to prioritize
-adding edges between pair of variables that have the most severe problems.
-
-Given a DAG $ G $ (potentially empty), a dataset $ D $, a p-value threshold $
-\alpha $ and a measure of association threshold $ \beta $, our construction
-procedure works as follows:
-
-\begin{enumerate}
-	\item For each pair of variables $ X $ and $ Y $ in $ G $:
-		\begin{itemize}
-			\item If there is an edge between $ X $ and $ Y $, remove the edge to obtain the DAG $ G' $, and compute the measure of association $ \phi(X, Y \rvert G'_{pa}(X) \cup G'_{pa}(Y)) $
-			\item If there is no edge between $ X $ and $ Y $, compute the measure of association $ \phi(X, Y \rvert G_{pa}(X) \cup \underline{G}_{pa}(Y) $.
-		\end{itemize}
-	\item If for any pair $ X $ and $ Y $, if there is an edge between them with p-value $ > \alpha $ and $ \phi < \beta $, remove the edge between them.
-	\item For the remaining $ X $ and $ Y $ pairs with p-value $ \geq \alpha $, sort them by $ \beta $.
-Select the pair with highest $ \beta $, ask the user to specify the direction between $ X $ and $ Y $ and add the edge to $ G $.
-\end{enumerate}
-
-For a pair of variables $ X $ and $ Y $, if there is an edge present between
-them, the interpretation of this partial association measure is similar to path
-coefficients in SEMs, i.e. how strong is the edge between the variables. Hence,
-we can use this to identify potential edges that can be removed from the model.
-In the case when there is no edge between the variables, the value of this
-partial measure of association can be interpreted as - given the current
-structure of the model how much of the observed correlation between the
-variables $ X $ and $ Y $ in data is not explained by the model. To modify the
-model based on this, we can look at the pair of variables that have an edge
-between them. If any of these pair have a p-value greater than partial
-association below our chosen threshold, we can remove that edge as it is
-redundant.
-
-The measure of association also gives us a way to rank the potential
-modifications. The smallest values (below the threshold) of the association
-measure are the worst offending ones and for when the edges are not present the
-largest values would add the most amount of explainability to the model. Using
-this ranking the user can decide which pair of variables to focus on.
-
-We can then use this to rank pair of variables which have very high unexplained
-association between them. Based on the domain expertise, we can then choose one
-of these suggested pair of variables, decide the direction of the edge between
-the variables, and add it to the model. Once the new edge is added we can
-recompute the association of all other variables and $ X $ and $ Y $.
-
-Additionally, in each iteration of modification, for each pair of variable in $
-G $, that has an edge between them in we compute the partial association
-measure: $ \rho(X, Y; pa_{\underline{G}}(X) \cup pa_{\underline{G}}(Y) $. This
-measure of association can be interpreted as the strength of the edge between
-the variables. If the value of this association below a certain threshold, we
-can remove the edge.
-
-\begin{theorem}
-	If we have access to an oracle that can give correct edge directions, taking a
-	greedy approach with the procedure above results in the correct graph
-	being recovered. \todo[inline]{Reword and add proof}
-\end{theorem}
-
 \begin{figure}[t!]

### 2025-02-07 paper/figures_v2.tex
Updates
Udpates figure

@@
 	\begin{tikzpicture}
-	\begin{scope}[xscale=1.2, yscale=1.5]
+	\begin{scope}[xscale=1, yscale=1]
 		\tikzstyle{every node}=[align=center, inner sep=1pt]
-		\node (ms)   at (0, 0)    {\footnotesize{Marital Status}};
-		\node (rln)  at (0, -0.8) {\footnotesize{Relationship}};
-		\node (sex)  at (0, -1.6) {\footnotesize{Sex}};
-		\node (hrpw) at (0, -2.4) {\footnotesize{Hours Per Week}};
+		\node (ms)   at (0, 0)    {{Marital Status}};
+		\node (rln)  at (0, -0.8) {{Relationship}};
+		\node (sex)  at (0, -1.6) {{Sex}};
+		\node (hrpw) at (0, -2.4) {{Hours Per Week}};
 
-		\node (age) at (2.5, 0)    {\footnotesize{Age}};
-		\node (inc) at (2.5, -0.8) {\footnotesize{Income}};
-		\node (occ) at (2.5, -1.6) {\footnotesize{Occupation}};
-		\node (wc)  at (2.5, -2.4) {\footnotesize{Work Class}};
+		\node (age) at (2.5, 0)    {{Age}};
+		\node (inc) at (2.5, -0.8) {{Income}};
+		\node (occ) at (2.5, -1.6) {{Occupation}};
+		\node (wc)  at (2.5, -2.4) {{Work Class}};
 
-		\node (nc)   at (5, 0)    {\footnotesize{Native Country}};
-		\node (race) at (5, -0.8) {\footnotesize{Race}};
-		\node (edu)  at (5, -1.6) {\footnotesize{Education}};
+		\node (nc)   at (5, 0)    {{Native Country}};
+		\node (race) at (5, -0.8) {{Race}};
+		\node (edu)  at (5, -1.6) {{Education}};
 
@@
 
-		\draw [red] (rln.-70) to (sex.70);
+		\draw [red, transform canvas={xshift=0.15em}] (rln) to (sex);
+		\draw [blue, ->, >=stealth, transform canvas={xshift=-0.15em}] (sex) to (rln);
+
+		\draw [red, transform canvas={xshift=0.15em}] (occ) to (edu);
+		\draw [blue, ->, >=stealth, transform canvas={xshift=-0.15em}] (occ) to (edu);
+
 		\draw [red] (inc) to (edu);
-		\draw [red] (occ.-5) to (edu.185);
 
-		\draw [blue, ->, >=stealth] (sex.110) to (rln.-110);
 		\draw [blue, ->, >=stealth] (sex) to (occ);
 		\draw [blue, ->, >=stealth] (occ) to (wc);
-		\draw [blue, ->, >=stealth] (occ.5) to (edu.175);
 		\draw [blue, ->, >=stealth] (wc) to (edu);
-		\draw [blue, ->, >=stealth] (wc) to[bend right=40] (inc);
+		\draw [blue, ->, >=stealth] (wc) to[bend left=80] (inc);
 	\end{scope}
@@
 	\begin{tikzpicture}
-	\begin{scope}[xscale=1.2, yscale=1.5]
+	\begin{scope}
 		\tikzstyle{every node}=[align=center, inner sep=0.2em]
 
-		\node (ms)   at (0, 0)    {\footnotesize{Marital Status}};
-		\node (rln)  at (0, -0.8) {\footnotesize{Relationship}};
-		\node (sex)  at (0, -1.6) {\footnotesize{Sex}};
-		\node (hrpw) at (0, -2.4) {\footnotesize{Hours Per Week}};
+		\node (ms)   at (0, 0)    {{Marital Status}};
+		\node (rln)  at (0, -0.8) {{Relationship}};
+		\node (sex)  at (0, -1.6) {{Sex}};
+		\node (hrpw) at (0, -2.4) {{Hours Per Week}};
 
-		\node (age) at (2.5, 0)    {\footnotesize{Age}};
-		\node (inc) at (2.5, -0.8) {\footnotesize{Income}};
-		\node (occ) at (2.5, -1.6) {\footnotesize{Occupation}};
-		\node (wc)  at (2.5, -2.4) {\footnotesize{Work Class}};
+		\node (age) at (2.5, 0)    {{Age}};
+		\node (inc) at (2.5, -0.8) {{Income}};
+		\node (occ) at (2.5, -1.6) {{Occupation}};
+		\node (wc)  at (2.5, -2.4) {{Work Class}};
 
-		\node (nc)   at (5, 0)    {\footnotesize{Native Country}};
-		\node (race) at (5, -0.8) {\footnotesize{Race}};
-		\node (edu)  at (5, -1.6) {\footnotesize{Education}};
+		\node (nc)   at (5, 0)    {{Native Country}};
+		\node (race) at (5, -0.8) {{Race}};
+		\node (edu)  at (5, -1.6) {{Education}};
 
@@
 
-		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (rln) to (age);
-		\draw [blue,->, >=stealth, transform canvas={yshift=-0.15em}] (rln) to (age);
+		\draw [red, ->, >=stealth, transform canvas={yshift=0em}] (rln) to (age);
+		\draw [blue,->, >=stealth, transform canvas={yshift=-0.3em}] (rln) to (age);
 
@@
 		\draw [red, ->, >=stealth] (sex) to (wc);
-		\draw [red, ->, >=stealth] (occ) to (ms);
+		\draw [red, ->, >=stealth] (occ) to (ms.-10);
 		\draw [red] (nc) to (race);
@@
 		\draw [blue,->, >=stealth] (rln) to[bend right=50] (hrpw);
-		\draw [blue,->, >=stealth] (hrpw) to (edu);
+		\draw [blue,->, >=stealth] (hrpw) to[bend right=25] (edu);
 
@@
 	\begin{tikzpicture}
-	\begin{scope}[xscale=0.1]
-		\tikzstyle{every node}=[align=center, inner sep=1pt]
+	\begin{scope}
+		\tikzstyle{every node}=[align=center, inner sep=0.2em]
 
-		\node (ms)   at (0, 0)    {\footnotesize{Marital Status}};
-		\node (rln)  at (0, -0.8) {\footnotesize{Relationship}};
-		\node (sex)  at (0, -1.6) {\footnotesize{Sex}};
-		\node (hrpw) at (0, -2.4) {\footnotesize{Hours Per Week}};
+		\node (ms)   at (0, 0)    {{Marital Status}};
+		\node (rln)  at (0, -0.8) {{Relationship}};
+		\node (sex)  at (0, -1.6) {{Sex}};
+		\node (hrpw) at (0, -2.4) {{Hours Per Week}};
 
-		\node (age) at (1.5, 0)    {\footnotesize{Age}};
-		\node (inc) at (1.5, -0.8) {\footnotesize{Income}};
-		\node (occ) at (1.5, -1.6) {\footnotesize{Occupation}};
-		\node (wc)  at (1.5, -2.4) {\footnotesize{Work Class}};
+		\node (age) at (2.5, 0)    {{Age}};
+		\node (inc) at (2.5, -0.8) {{Income}};
+		\node (occ) at (2.5, -1.6) {{Occupation}};
+		\node (wc)  at (2.5, -2.4) {{Work Class}};
 
-		\node (nc)   at (3, 0)    {\footnotesize{Native Country}};
-		\node (race) at (3, -0.8) {\footnotesize{Race}};
-		\node (edu)  at (3, -1.6) {\footnotesize{Education}};
+		\node (nc)   at (5, 0)    {{Native Country}};
+		\node (race) at (5, -0.8) {{Race}};
+		\node (edu)  at (5, -1.6) {{Education}};
 
-		\draw [] (ms) to (rln);
-		\draw [] (rln) to (sex);
-		\draw [->, >=stealth] (sex) to (hrpw);
+		\draw [red, transform canvas={xshift=0.15em}] (ms) to (rln);
+		\draw [blue, transform canvas={xshift=-0.15em}] (ms) to (rln);
 
-		\draw [->, >=stealth] (ms) to (age);
-		\draw [] (rln) to (inc);
-		\draw [] (sex) to (occ);
+		\draw [red, transform canvas={xshift=0.15em}] (rln) to (sex);
+		\draw [blue, transform canvas={xshift=-0.15em}] (rln) to (sex);
+
+		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (ms) to (age);
+		\draw [blue, ->, >=stealth, transform canvas={yshift=-0.15em}] (ms) to (age);
+
+		\draw [red, transform canvas={yshift=0.15em}] (rln) to (inc);
+		\draw [blue, transform canvas={yshift=-0.15em}] (rln) to (inc);
+
+		\draw [red, transform canvas={yshift=0.15em}] (sex) to (occ);
+		\draw [blue, transform canvas={yshift=-0.15em}] (sex) to (occ);
+
+		\draw [red, ->, >=stealth] (sex) to (hrpw);
+
+		\draw [blue, ->, >=stealth] (rln) to[bend right=40] (hrpw);
+
 	\end{scope}
 	\end{tikzpicture}
-
-	% Hill Climb with BIC Score Q3 test 800 samples
-	\begin{tikzpicture}
-	\begin{scope}[xscale=0.9]
-		\tikzstyle{every node}=[align=center, inner sep=1pt]
-
-		\node (ms)   at (0, 0)    {\footnotesize{MS}};
-		\node (rln)  at (0, -0.8) {\footnotesize{Rln.}};
-		\node (sex)  at (0, -1.6) {\footnotesize{Sex}};
-		\node (hrpw) at (0, -2.4) {\footnotesize{HrPW}};
-
-		\node (age) at (1.5, 0)    {\footnotesize{Age}};
-		\node (inc) at (1.5, -0.8) {\footnotesize{Inc.}};
-		\node (occ) at (1.5, -1.6) {\footnotesize{Occ.}};
-		\node (wc)  at (1.5, -2.4) {\footnotesize{WC}};
-
-		\node (nc)   at (3, 0)    {\footnotesize{NC}};
-		\node (race) at (3, -0.8) {\footnotesize{Race}};
-		\node (edu)  at (3, -1.6) {\footnotesize{Edu.}};
-
-		\draw [] (ms) to (rln);
-		\draw [] (rln) to (sex);
-		\draw [->, >=stealth] (rln) to[bend right=40] (hrpw);
-
-		\draw [->, >=stealth] (ms) to (age);
-		\draw [] (rln) to (inc);
-		\draw [] (sex) to (occ);
-	\end{scope}
-	\end{tikzpicture}
-
 \end{document}

### 2025-02-07 paper/main.tex
Udpates figure

@@
     	\includegraphics[page=1]{figures_v2.pdf}
-    	\caption{PC algorithm with mutual information test. Red edges: $ N=400 $, Blue edges: $ N=800 $}
+    	\caption{}
     \end{subfigure}
@@
     	\includegraphics[page=2]{figures_v2.pdf}
-    	\caption{PC algorithm with a residualization based test. Red edges $N=400$, Blue edges: $ N=800 $}
+    	\caption{}
     \end{subfigure}
@@
     	\includegraphics[page=3]{figures_v2.pdf}
-    	\caption{PC algorithm with Hill Climb Search with BIC score}
+    	\caption{}
     \end{subfigure}
@@
     \caption{A comparison of Markov Equivalence Class (MEC) learned using Adult
-	     Income Dataset \citep{Becker1996} using (a \& b) PC algorithm
-	     using a mutual information based test, (c \& d) Hill Climb Search
-	     with Bayesian Information Criterion (BIC) score, (e \& f) PC
-	     algorithm with a residualization based test \citep{Ankan2023}. The
-	     learned model structure varies significantly depending on the
-     	     algorithm, the CI test, and the sample size used.}
+	    Income Dataset \citep{Becker1996} using different algorithms and
+	    sample sizes. Edge colour represents the sample size used: Red
+	    ($N=400$), and blue ($N=800$). (a) PC algorithm with a mutual
+	    information based CI test, (b) PC algorithm with a residualization
+	    based test \citep{Ankan2023}, (c) Hill Climb Search with Bayesian
+	    Information Criterion (BIC) score. The learned model structure
+	    varies significantly in each case.}
     \label{fig:intro}

### 2025-02-07 paper/figures_v2.tex
Updates

@@
 		\draw [red, ->, >=stealth, transform canvas={yshift=0em}] (rln) to (age);
-		\draw [blue,->, >=stealth, transform canvas={yshift=-0.3em}] (rln) to (age);
+		\draw [blue,->, >=stealth, transform canvas={yshift=-0.3em, xshift=0.1em}] (rln) to (age);
 
@@
 		\draw [blue,->, >=stealth, transform canvas={xshift=-0.15em}] (inc) to (age);
-		\draw [red, ->, >=stealth, transform canvas={xshift=0.15em, yshift=0.15em}] (inc) to (edu);
+		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (inc) to (edu);
 		\draw [blue,->, >=stealth, transform canvas={xshift=-0.15em, yshift=-0.15em}] (inc) to (edu);
 
-		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (occ) to (hrpw);
-		\draw [blue,->, >=stealth, transform canvas={yshift=-0.15em}] (occ) to (hrpw);
+		\draw [red, ->, >=stealth, transform canvas={yshift=0.15em}] (occ.-170) to (hrpw.15);
+		\draw [blue,->, >=stealth, transform canvas={yshift=-0.15em, xshift=0.1em}] (occ.-170) to (hrpw.15);
 
@@
 		\draw [blue,->, >=stealth] (rln) to[bend right=50] (hrpw);
-		\draw [blue,->, >=stealth] (hrpw) to[bend right=25] (edu);
+		\draw [blue,->, >=stealth] (hrpw) to[bend right=30] (edu);
 

### 2025-02-07 paper/main.tex
Updates

@@
     \begin{subfigure}{0.5\textwidth}
+	\centering
     	\includegraphics[page=2]{figures_v2.pdf}
@@
     \begin{subfigure}{0.5\textwidth}
+	\centering
     	\includegraphics[page=3]{figures_v2.pdf}
@@
 
-\subsubsection{Both $ X $ and $ Y $ are continuous}
+\paragraph{Both $ X $ and $ Y $ are continuous: }
 When both $ X $ and $ Y $ are continuous, Pearson's correlation coefficient is
@@
 
-\paragraph{All $ X $, $ Y $, and $ \bm{Z} $ are discrete}
+\paragraph{All $ X $, $ Y $, and $ \bm{Z} $ are discrete: }
 
@@
 
-\paragraph{$ X $ is ordinal, and $ Y $ and $ Z $  are continuous or ordinal}
+\paragraph{$ X $ is ordinal, and $ Y $ and $ Z $  are continuous or ordinal: }
 

### 2025-02-07 paper/main.tex
Updates

@@
 
-That is, our ancestral oracles never distinguish whether the relationships between variables is
-direct or indirect; and the weak ancestral oracle is even unable to detect cases where there is no
-cause-effect relationship between two variables. Using such oracles for DAG construction therefore
-inherently risks that superfluous edges are created. In our approach, such superfluous edges will
-be detected and removed by utilizing the data.
+That is, our ancestral oracles never distinguish whether the relationships
+between variables is direct or indirect; and the weak ancestral oracle is even
+unable to detect cases where there is no cause-effect relationship between two
+variables. Using such oracles for DAG construction therefore inherently risks
+that superfluous edges are created. In our approach, such superfluous edges
+will be detected and removed by utilizing the data.
 
-To model how our learning procedure interacts with the data, we
-assume that we have access to a second oracle $\mathcal{D}_G$ than can answer
-d-separation queries with respect to the unknown target graph: 
-$\mathcal{D}_G(X,Y,\mathbf{Z})=1$ iff $X \ci_G Y \mid \mathbf{Z}$. 
-These are the standard oracles considered in constraint-based structure learning algorithms,
-such as the PC algorithm \cite{Kalisch2012}. Based on this oracle, we now define two procedures
-that iteratively change a current DAG structure.
+To model how our learning procedure interacts with the data, we assume that we
+have access to a second oracle $\mathcal{D}_G$ than can answer d-separation
+queries with respect to the unknown target graph:
+$\mathcal{D}_G(X,Y,\mathbf{Z})=1$ iff $X \ci_G Y \mid \mathbf{Z}$. These are
+the standard oracles considered in constraint-based structure learning
+algorithms, such as the PC algorithm \cite{Spirtes2001}. Based on this oracle,
+we now define two procedures that iteratively change a current DAG structure.
 
-Our structure learning algorithm is based on two core procedures. \textsc{Expand} 
-(Algorithm~\ref{algo:expand}) uses conditional independence information to search for
-missing edges in the graph and then uses domain knowledge to orient them. 
-\textsc{Prune} (Algorithm~\ref{algo:prune}) uses conditional independence information
-to remove superfluous edges from the graph.
+Our structure learning algorithm is based on two core procedures.
+\textsc{Expand} (Algorithm~\ref{algo:expand}) uses conditional independence
+information to search for missing edges in the graph and then uses domain
+knowledge to orient them. \textsc{Prune} (Algorithm~\ref{algo:prune}) uses
+conditional independence information to remove superfluous edges from the
+graph.
 
@@
 }
-\caption{Adding edges based on data an domain knowledge}
+\caption{Adding edges based on data and domain knowledge}
 \label{algo:expand}

### 2025-02-07 web/api.R
Updates Fisher C to RMSEA

@@ -104,8 +104,8 @@ run_citests <- function( req, res, dag, threshold, pval ){
 }
 
 #* @param dag
-#* @get /getfisher
-compute_fisher <- function( dag ){
+#* @get /rmsea
+rmsea <- function( dag ){
 	g <- dagitty::dagitty(dag)
 	pvalues <- c()
 	nodes <- names(g)
@@ -123,4 +123,5 @@ compute_fisher <- function( dag ){
 	}
 	fisherc <- -2 * sum(log(pvalues))
 	return (pchisq(fisherc, 2*length(pvalues), lower.tail=F))
+	return(sqrt( (fisherc**2 - 2*length(pvalues)) / (2*length(pvalues)) / nrow(dataset) ))
 }

### 2025-02-07 web/utils.js

@@ -64,7 +64,7 @@ async function send(){
 	const a = await fetch('http://127.0.0.1:8000/getassoc?dag='+encodeURIComponent(g.toString())+'&threshold='+document.getElementById('thres_txt').value+'&pval='+document.getElementById('pval_txt').value)
 	const b = await a.json()
 
-	const fisher = await fetch('http://127.0.0.1:8000/getfisher?dag='+encodeURIComponent(g.toString()))
+	const fisher = await fetch('http://127.0.0.1:8000/rmsea?dag='+encodeURIComponent(g.toString()))
 	const pval = await fisher.json()
 	document.getElementById('fisherc').innerHTML = pval;
 

### 2025-02-07 web/api.R
Fixes bug

@@ -122,6 +122,6 @@ rmsea <- function( dag ){
 		}
 	}
 	fisherc <- -2 * sum(log(pvalues))
-	return (pchisq(fisherc, 2*length(pvalues), lower.tail=F))
+	# return (pchisq(fisherc, 2*length(pvalues), lower.tail=F))
 	return(sqrt( (fisherc**2 - 2*length(pvalues)) / (2*length(pvalues)) / nrow(dataset) ))
 }

### 2025-02-07 paper/main.tex
Updates
Update example

@@
 		\includegraphics[page=2]{example.pdf}
-		\caption{No edges}
+		\caption{No edges.}
 	\end{subfigure}%
@@
 		\includegraphics[page=3]{example.pdf}
-		\caption{Orient $ X_1 \rightarrow X_3 $}
+		\caption{Orient $ X_1 \rightarrow X_4 $}
 	\end{subfigure}
@@
 		\includegraphics[page=4]{example.pdf}
-		\caption{Orient $ X_2 \rightarrow X_3 $}
+		\caption{Orient $ X_1 \rightarrow X_2 $}
 	\end{subfigure}%
@@
 		\includegraphics[page=5]{example.pdf}
-		\caption{Orient $ X_3 \rightarrow X_4 $}
+		\caption{Orient $ X_1 \rightarrow X_3 $}
 	\end{subfigure}%
@@
 		\includegraphics[page=6]{example.pdf}
-		\caption{Orient $ X_3 \rightarrow X_5 $}
+		\caption{Orient $ X_2 \rightarrow X_4 $}
 	\end{subfigure}
-	\caption{An example showing a greedy approach for causal discovery using the mixed data measure of association.
-We start with an empty graph (b) where all the variables in the model are associated with each other.
-The strength of association is denoted using the thickness of the dotted edges.
-We select the pair of variables with the highest association, for example, $X_1$ and $X_3$, and orient it based on our domain knowledge.
-We continue the process till we have a DAG.
+	\begin{subfigure}{0.17 \textwidth}
+		\includegraphics[page=7]{example.pdf}
+		\caption{Expand $ X_3 \rightarrow X_4 $. Prune $ X_1 \rightarrow X_4 $.}
+	\end{subfigure}
+	\caption{An example showing the steps of \textit{ExpertInLoop} algorithm. The dashed edges show all the potential edges that \textit{Expand} algorithm could add in each step.
 	}
 \end{figure}
-
-\subsection{Connection to Score Based Methods}
-
-Our procedure closely resembles certain score-based automated causal discovery
-methods, such as Greedy Equivalence Search (GES), where the algorithm
-iteratively adds or removes edges that improve the scoring metric the most.
-Similarly, a greedy version of our approach would involve adding an edge
-between the pair of variables with the highest unexplained correlation. While
-both methods aim to find modifications that most improve the model at each
-step, a key distinction lies in the nature of the evaluation criteria: unlike
-standard scoring metrics, our measure of association is not decomposable. A key
-property of scoring metrics is that they should be decomposible, i.e., they can
-be expressed as the sum of scores of nodes given its parents. This means adding
-or removing an edge affects only a localized part of the model. In contrast,
-our measure of association is global, meaning that modifying an edge in one
-part of the model can influence association values elsewhere.
-
-Another major difference is in the interpretability of the evaluation metric.
-Most scoring metrics are based on log-likelihood with a penalty for model
-complexity. These score metrics allow for relative comparisons between models
-but do not provide an absolute measure of model fit. That is, they indicate
-which model is better for a given dataset but do not reveal how well the model
-explains the data in an absolute sense. In contrast, summing our measure of
-association across all variables provides a direct fitness measure—this sum
-approaches zero when the model perfectly accounts for all observed correlations
-in the data. This property enables an absolute assessment of model quality,
-rather than just a relative comparison between models.
 
@@
 
+\subsection{Connection to Score Based Methods}
+
+Our procedure closely resembles certain score-based automated causal discovery
+methods, such as Greedy Equivalence Search (GES), where the algorithm
+iteratively adds or removes edges that improve the scoring metric the most.
+Similarly, a greedy version of our approach would involve adding an edge
+between the pair of variables with the highest unexplained correlation. While
+both methods aim to find modifications that most improve the model at each
+step, a key distinction lies in the nature of the evaluation criteria: unlike
+standard scoring metrics, our measure of association is not decomposable. A key
+property of scoring metrics is that they should be decomposible, i.e., they can
+be expressed as the sum of scores of nodes given its parents. This means adding
+or removing an edge affects only a localized part of the model. In contrast,
+our measure of association is global, meaning that modifying an edge in one
+part of the model can influence association values elsewhere.
+
+Another major difference is in the interpretability of the evaluation metric.
+Most scoring metrics are based on log-likelihood with a penalty for model
+complexity. These score metrics allow for relative comparisons between models
+but do not provide an absolute measure of model fit. That is, they indicate
+which model is better for a given dataset but do not reveal how well the model
+explains the data in an absolute sense. In contrast, summing our measure of
+association across all variables provides a direct fitness measure—this sum
+approaches zero when the model perfectly accounts for all observed correlations
+in the data. This property enables an absolute assessment of model quality,
+rather than just a relative comparison between models.
+
 \subsection{Using LLMs as Experts}

### 2025-02-07 code/fig4/make_plots.R
Fix plots

@@ -25,8 +25,7 @@ p_unexplained_effect <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
 	labs(x="Iteration No.") +
 	labs(y="Total Unexplained Effect") + 
 	theme(legend.position='top',
-	      legend.title=element_blank(),
-	      legned.text=element_blank())
+	      legend.title=element_blank())
 
 p_ll <- ggplot(d_results, aes(x=x, y=ll, color=algo)) +
 	geom_line(alpha=0.6) +
@@ -35,8 +34,7 @@ p_ll <- ggplot(d_results, aes(x=x, y=ll, color=algo)) +
 	labs(x="Iteration No.") +
 	labs(y="Log-Likelihood") +
 	theme(legend.position='top',
-	      legend.title=element_blank(),
-	      legend.text=element_blank())
+	      legend.title=element_blank())
 
 ggsave('unexplained_effect.pdf', p_unexplained_effect, height=2.5, width=1.6, units='in')
 ggsave('ll.pdf', p_ll, height=2.5, width=1.6, units='in')

### 2025-02-07 paper/main.tex
Updates

@@
 \begin{figure}[t!]
-	\begin{subfigure}{0.17 \textwidth}
+	\begin{subfigure}{0.125 \textwidth}
+		\centering
 		\includegraphics[page=1]{example.pdf}
@@
 	\end{subfigure}%
-	\begin{subfigure}{0.17 \textwidth}
+	\begin{subfigure}{0.125 \textwidth}
+		\centering
 		\includegraphics[page=2]{example.pdf}
@@
 	\end{subfigure}%
-	\begin{subfigure}{0.17 \textwidth}
+	\begin{subfigure}{0.125 \textwidth}
+		\centering
 		\includegraphics[page=3]{example.pdf}
-		\caption{Orient $ X_1 \rightarrow X_4 $}
+		\caption{$ X_1 \rightarrow X_4 $}
+	\end{subfigure}%
+	\begin{subfigure}{0.125 \textwidth}
+		\centering
+		\includegraphics[page=4]{example.pdf}
+		\caption{$ X_1 \rightarrow X_2 $}
 	\end{subfigure}
-	\begin{subfigure}{0.17 \textwidth}
-		\includegraphics[page=4]{example.pdf}
-		\caption{Orient $ X_1 \rightarrow X_2 $}
+	\begin{subfigure}{0.125 \textwidth}
+		\centering
+		\includegraphics[page=5]{example.pdf}
+		\caption{$ X_1 \rightarrow X_3 $}
 	\end{subfigure}%
-	\begin{subfigure}{0.17 \textwidth}
-		\includegraphics[page=5]{example.pdf}
-		\caption{Orient $ X_1 \rightarrow X_3 $}
+	\begin{subfigure}{0.125 \textwidth}
+		\centering
+		\includegraphics[page=6]{example.pdf}
+		\caption{$ X_2 \rightarrow X_4 $}
 	\end{subfigure}%
-	\begin{subfigure}{0.17 \textwidth}
-		\includegraphics[page=6]{example.pdf}
-		\caption{Orient $ X_2 \rightarrow X_4 $}
-	\end{subfigure}
-	\begin{subfigure}{0.17 \textwidth}
+	\begin{subfigure}{0.250 \textwidth}
+		\centering
 		\includegraphics[page=7]{example.pdf}
-		\caption{Expand $ X_3 \rightarrow X_4 $. Prune $ X_1 \rightarrow X_4 $.}
+		\caption{$ X_3 \rightarrow X_4 $. $ X_1 \not \rightarrow X_4 $.}
 	\end{subfigure}
@@
 
-We analyzed the performance of this manual approach by comparing it with two
-automated algorithms: PC and Greedy Equivalence Search (GES). For the
-comparison we use simulated data from a known DAG and compare how well the
-algorithm is able to recover the original model using two metrics: Structural
-Hamming Distance (SHD) and Structural Intervention Distance (SID). To simulate
-the data, we start with a randomly generated DAG on $ 10 $ nodes and use linear
-models with random effects to generate the data.
+We analyzed the performance of \textit{ExpertInLoop} algorithm by comparing it
+with three commonly used automated algorithms: PC, Hill-Climb Search, and
+Greedy Equivalence Search (GES). For the analysis, we simulated data from known
+DAG structures and
+
+we use simulated data from
+a known DAG and compare how well the algorithm is able to recover the original
+model using two metrics: Structural Hamming Distance (SHD) and Structural
+Intervention Distance (SID). To simulate the data, we start with a randomly
+generated DAG on $ 10 $ nodes and use linear models with random effects to
+generate the data.
 

### 2025-02-07 paper/main.tex
Updates

@@
 We analyzed the performance of \textit{ExpertInLoop} algorithm by comparing it
-with three commonly used automated algorithms: PC, Hill-Climb Search, and
-Greedy Equivalence Search (GES). For the analysis, we simulated data from known
-DAG structures and
+with three other automated algorithms: PC, Hill-Climb Search, and Greedy
+Equivalence Search (GES). For the analysis, we simulated data from known DAG
+structures and use the algorithms to recover the original DAG. We start by
+generating a random DAG on $10$ nodes and use linear models with random effects
+to generate the data. We vary the edge probability of the DAG to generate DAGs
+with varying density. We repeat the data generation $30$ times for each density
+value for the DAG. We compare the learned DAG to the original DAG using two
+metrics: Structural Hamming Distance (SHD) and Structural Intervention Distance
+(SID) \citep{Peters2015}.
+
+As both PC and Hill-Climb Search can only give us a CPDAG, we
 

### 2025-02-07 code/fig3/sl.R
Modify figure 3 for using linear continuous data

@@ -97,7 +97,7 @@ simulate_human_sl <- function(sim_data, true_dag, oracle_acc, max_iter=1e4){
 }
 
 run_single_exp_hc <- function(n_nodes, edge_prob){
-	d <- gen_mixed_data(n_nodes=n_nodes, edge_prob=edge_prob)
+	d <- gen_linear_data(n_nodes=n_nodes, edge_prob=edge_prob)
 	true_adj <- d$true_adj
 	sim_data <- d$sim_data
 	var_types <- d$var_types
@@ -140,7 +140,7 @@ run_single_exp_hc <- function(n_nodes, edge_prob){
 }
 
 run_single_exp_pc <- function(n_nodes, edge_prob){
-	d <- gen_mixed_data(n_nodes=n_nodes, edge_prob=edge_prob)
+	d <- gen_linear_data(n_nodes=n_nodes, edge_prob=edge_prob)
 	true_adj <- d$true_adj
 	sim_data <- d$sim_data
 
@@ -179,15 +179,19 @@ run_single_exp_pc <- function(n_nodes, edge_prob){
 }
 
 run_single_exp_ges <- function(n_nodes, edge_prob){
-	d <- gen_mixed_data(n_nodes=n_nodes, edge_prob=edge_prob)
+	d <- gen_linear_data(n_nodes=n_nodes, edge_prob=edge_prob)
 	true_adj <- d$true_adj
 	sim_data <- d$sim_data
 
-	rand_str <- stringi::stri_rand_strings(1, 5)
-	write.csv(sim_data, paste0("temp/", rand_str,".csv"), row.names=F)
-	system(paste0("python sl_ges.py temp/", rand_str, ".csv"), intern=T)
-	
-	ges_adj <- as.matrix(read.csv(paste0("temp/adj_", rand_str, ".csv"), row.names=1))
+	# rand_str <- stringi::stri_rand_strings(1, 5)
+	# write.csv(sim_data, paste0("temp/", rand_str,".csv"), row.names=F)
+	# system(paste0("python sl_ges.py temp/", rand_str, ".csv"), intern=T)
+	# 
+	# ges_adj <- as.matrix(read.csv(paste0("temp/adj_", rand_str, ".csv"), row.names=1))
+
+	score <- new("GaussL0penObsScore", sim_data)
+	ges.cpdag <- pcalg::ges(score)$essgraph
+	ges_adj <- as(ges.cpdag, 'matrix')
 
[…]

### 2025-02-07 code/fig3/utils.R

@@ -16,6 +16,24 @@ gen_mixed_data <- function(n_nodes, edge_prob){
 	return (list(true_dag=dag, true_adj=true_adj, sim_data=sim_data$d, var_types=sim_data$var_types))
 }
 
+gen_linear_data <- function(n_nodes, edge_prob){
+	varnames <- sapply(1:n_nodes, function(v) {
+	    paste0("x", v)
+	})
+	
+	dag <- pcalg::randomDAG(n=n_nodes, prob=edge_prob, lB=1, uB=1, V=varnames)
+	dag <- pcalg::pcalg2dagitty(as(dag, "matrix"), labels = varnames, type = "dag")
+	sim_data <- dagitty::simulateSEM(dag, N=500)
+	true_adj <- dag_to_adjmatrix(dag)
+
+	var_types <- list()
+	for (variable in varnames){
+		var_types[[variable]] <- "cont"
+	}
+	return (list(true_dag=dag, true_adj=true_adj, sim_data=sim_data, var_types=var_types))
+}
+
+
 # Takes a dagitty DAG and returns an adjacency matrix representation.
 dag_to_adjmatrix <- function(daggity_obj) {
   	edg <- dagitty:::edges(daggity_obj)

### 2025-02-07 paper/main.tex
updates

@@
 
-As both PC and Hill-Climb Search can only give us a CPDAG, we
-
-we use simulated data from
-a known DAG and compare how well the algorithm is able to recover the original
-model using two metrics: Structural Hamming Distance (SHD) and Structural
-Intervention Distance (SID). To simulate the data, we start with a randomly
-generated DAG on $ 10 $ nodes and use linear models with random effects to
-generate the data.
-
-We vary the edge probability of the DAG to generate DAGs with varying density.
-We repeat the data generation $ 30 $ times for each density value for the DAG.
-As both PC and Hill-Climb Search can only give us a CPDAG, we compare the
-orientation of these CPDAG which result in the best and worst case scores on
-SHD and SID.
+As the automated algorithms can only recover the CPDAG from data, and our method
+recovers the DAG. To compare them fairly, we computed the SHD and SID for all
+possible orientations of the CPDAG and show them in the Figure.
 

### 2025-02-07 code/fig3/sl.R
Updates

@@ -26,7 +26,7 @@ oracle <- function(u, v, true_dag, accuracy=1){
 		}
 	}
 	else{
-		return(list(c(u, v), c(v, u), NULL)[[sample(1:3, 1)]]) } }
+		return(list(c(u, v), c(v, u))[[sample(1:2, 1)]]) } }
 
 
 # Simulate a human structure learning based on CI tests.

### 2025-02-07 code/fig3/utils.R

@@ -23,7 +23,14 @@ gen_linear_data <- function(n_nodes, edge_prob){
 	
 	dag <- pcalg::randomDAG(n=n_nodes, prob=edge_prob, lB=1, uB=1, V=varnames)
 	dag <- pcalg::pcalg2dagitty(as(dag, "matrix"), labels = varnames, type = "dag")
-	sim_data <- dagitty::simulateSEM(dag, N=500)
+
+	repeat{
+		sim_data <- try(dagitty::simulateSEM(dag, N=500))
+		if (!(inherits(sim_data, "try-error"))){
+			break
+		}
+	}
+
 	true_adj <- dag_to_adjmatrix(dag)
 
 	var_types <- list()

### 2025-02-07 web/api.R
Fixes RMSEA

@@ -117,11 +117,13 @@ rmsea <- function( dag ){
 	
 			if (!(nodes[i] %in% pa_j) & !(nodes[j] %in% pa_i)){
 				tst <- dagitty::ciTest(X=nodes[i], Y=nodes[j], Z=union(pa_i, pa_j), dataset, type="cis.pillai")
-				pvalues <- c(pvalues, tst[,"p.value"])
+				pvalues <- c(pvalues, max(tst[,"p.value"], 1e-20))
 			}
 		}
 	}
+
 	fisherc <- -2 * sum(log(pvalues))
 	# return (pchisq(fisherc, 2*length(pvalues), lower.tail=F))
-	return(sqrt( (fisherc**2 - 2*length(pvalues)) / (2*length(pvalues)) / nrow(dataset) ))
+	print(pvalues)
+	return(sqrt( max(fisherc**2 - 2*length(pvalues), 0) / (2*length(pvalues) * nrow(dataset) )))
 }

### 2025-02-08 paper/main.tex
Updates

@@
 
-As the automated algorithms can only recover the CPDAG from data, and our method
-recovers the DAG. To compare them fairly, we computed the SHD and SID for all
-possible orientations of the CPDAG and show them in the Figure.
+\todo[inline]{Add equation for data generation method once the plot is finalized}
 
-To use our proposed approach, we need an expert to tell us the direction of the
-edges. We simulated an expert using this assisted model construction approach,
-we start with an empty model and take a greedy approach to add edges to this
-model. At any point, we select the pair of variables that has significant
-correlation and the highest unexplained correlation, i.e., highest conditional
-association. We then use an oracle to decide the direction of the edge between
-the variables. Given an accuracy of the oracle, $ \alpha $, the oracle either
-gives the correct direction, incorrect direction, or returns None representing
-that it does not know the direction.
+As the automated algorithms can only recover the CPDAG from data, and our
+method recovers the DAG. To compare them fairly, we computed the SHD and SID
+for all possible orientations of the CPDAG to show the best and worst case scenarios.
+
+To use the \textit{ExpertInLoop} algorithm, we need two oracles: a d-separation
+oracle and an ancestral oracle. We use a conditional independence test (partial
+correlation if using linear data) for determining the d-separation with a
+p-value threshold of $ 0.05 $. For the ancestral oracle, we simulated the
+strong ancestral oracle. We vary the accuracy of our oracle to test the
+performance for varying oracle accuracies. Given an accuracy level $ \alpha $
+of the oracle, the oracle returns:
 
@@
 only edges which do not form a cyclic in the graph are used to select the next
-potential edge. After adding each edge, we also check if the p-value of any
-existing edge shows independence. If that happens we remove that edge and
-blacklist that edge, i.e., that edge is not prompted again during the run of
-the algorithm. We repeat this procedure till the model explains all
-correlations between the variables. There is a possibility that this greedy
-expert gets stuck at incorrect graph structures as it does not do any
-backtracking to fix its earlier mistakes. This procedure is outlined in the
-Algorithm 1.
+potential edge. There is a possibility that this greedy expert gets stuck at
+incorrect graph structures as it does not do any backtracking to fix its
+earlier mistakes. This procedure is outlined in the Algorithm 1.
 \todo[inline]{Show an algorithm of how this expert method is working}
 
-\begin{verbatim}
-	Input: Dataset $ D $ on variables $ \bm{X} $
-	Output: Learned DAG G.
-
-	function compute_effects(G):
-		pvalues <- dict()
-		effects <- dict()
-		for X_i, X_j in X:
-			if edge between X_i, X_j:
-				p_values[(X_i, X_j)] <- pvalue(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
-				effects[(X_i, X_j)] <- \phi(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
-			else:
-				p_values[(X_i, X_j)] <- pvalue(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
-				effects[(X_i, X_j)] <- \phi(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
-		return p_values, effects
-
-	$ G $ <- Empty Graph
-	p_values, effects <- compute_effects(G)
-	while (p_values > p_thres, 
-\end{verbatim}
-
-\begin{algorithm}
-	\KwData{Data set $ D $ on variables $ \bm{X} $, pthres, effectthres}
-	\KwResult{The learned DAG: $ G $}
-	\BlankLine
-	$ G \leftarrow $ Empty Graph
-\end{algorithm}
+% \begin{verbatim}
+% 	Input: Dataset $ D $ on variables $ \bm{X} $
+% 	Output: Learned DAG G.
+% 	
+% 	function compute_effects(G):
+% 		pvalues <- dict()
+% 		effects <- dict()
+% 		for X_i, X_j in X:
+% 			if edge between X_i, X_j:
+% 				p_values[(X_i, X_j)] <- pvalue(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
+% 				effects[(X_i, X_j)] <- \phi(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
+% 			else:
+% 				p_values[(X_i, X_j)] <- pvalue(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
+% 				effects[(X_i, X_j)] <- \phi(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
+% 		return p_values, effects
+% 	
+% 	$ G $ <- Empty Graph
+% 	p_values, effects <- compute_effects(G)
+% 	while (p_values > p_thres, 
+% \end{verbatim}
+% 
+% \begin{algorithm}
+% 	\KwData{Data set $ D $ on variables $ \bm{X} $, pthres, effectthres}
+% 	\KwResult{The learned DAG: $ G $}
+% 	\BlankLine
+% 	$ G \leftarrow $ Empty Graph
+% \end{algorithm}
 

### 2025-02-08 web/api.R
Fixes RMSEA

@@ -117,13 +117,12 @@ rmsea <- function( dag ){
 	
 			if (!(nodes[i] %in% pa_j) & !(nodes[j] %in% pa_i)){
 				tst <- dagitty::ciTest(X=nodes[i], Y=nodes[j], Z=union(pa_i, pa_j), dataset, type="cis.pillai")
-				pvalues <- c(pvalues, max(tst[,"p.value"], 1e-20))
+				pvalues <- c(pvalues, max(tst[,"p.value"], 1e-40))
 			}
 		}
 	}
 
 	fisherc <- -2 * sum(log(pvalues))
 	# return (pchisq(fisherc, 2*length(pvalues), lower.tail=F))
-	print(pvalues)
-	return(sqrt( max(fisherc**2 - 2*length(pvalues), 0) / (2*length(pvalues) * nrow(dataset) )))
+	return(sqrt( max(fisherc**2 - 2*length(pvalues), 0) / (2*length(pvalues) * (nrow(dataset) - 1) )))
 }

### 2025-02-08 web/api.R
Fixes RMSEA computation

@@ -124,5 +124,5 @@ rmsea <- function( dag ){
 
 	fisherc <- -2 * sum(log(pvalues))
 	# return (pchisq(fisherc, 2*length(pvalues), lower.tail=F))
-	return(sqrt( max(fisherc**2 - 2*length(pvalues), 0) / (2*length(pvalues) * (nrow(dataset) - 1) )))
+	return(sqrt( max(fisherc - 2*length(pvalues), 0) / (2*length(pvalues) * (nrow(dataset) - 1) )))
 }

### 2025-02-08 code/fig3/sl.R
Updated results

@@ -26,7 +26,7 @@ oracle <- function(u, v, true_dag, accuracy=1){
 		}
 	}
 	else{
-		return(list(c(u, v), c(v, u))[[sample(1:2, 1)]]) } }
+		return(list(c(u, v), c(v, u), NULL)[[sample(1:3, 1)]]) } }
 
 
 # Simulate a human structure learning based on CI tests.

### 2025-02-08 code/fig3/make_plots.R
Linear data fig3

@@ -256,5 +256,5 @@ make_plots_v2 <- function(filename, plot_type){
 	ggsave("sid_ribbon.pdf", p_sid, height=2, width=3.2, units='in')
 }
 
-make_plots_v2(filename='results/sl_results_mixed.csv', plot_type='ribbon')
+make_plots_v2(filename='results/sl_results_linear.csv', plot_type='ribbon')
 

### 2025-02-08 code/fig3/make_plots.R
Use alpha effective

@@ -176,6 +176,7 @@ make_plots_v2 <- function(filename, plot_type){
 	d_shd$lower <- as.double(d_shd$lower)
 	d_shd$upper <- as.double(d_shd$upper)
 
+	d_shd$line_labels <- round(as.numeric(d_shd$line_labels) + ((1 - as.numeric(d_shd$line_labels))/3), 2)
 	p_shd <- ggplot(d_shd, aes(x=edge_prob, y=mean_val, 
 				   ymin=lower, ymax=upper, 
 				   group=grp, color=algo, 
@@ -195,7 +196,7 @@ make_plots_v2 <- function(filename, plot_type){
 		ylim(0, 40) +
 		guides(linetype='none', fill='none')
 
-	ggsave("shd_ribbon.pdf", p_shd, height=2, width=3.2, units='in')
+	ggsave("shd_ribbon.pdf", p_shd, height=2, width=3.4, units='in')
 
 
 	d_sid <- d %>% select(edge_prob, oracle_acc, ges_lower_sid_mean, ges_upper_sid_mean, ges_lower_sid_sd, ges_upper_sid_sd, hc_lower_sid_mean, hc_upper_sid_mean, hc_lower_sid_sd, hc_upper_sid_sd, pc_lower_sid_mean, pc_upper_sid_mean, pc_lower_sid_sd, pc_upper_sid_sd, human_sid_mean, human_sid_sd)
@@ -235,6 +236,7 @@ make_plots_v2 <- function(filename, plot_type){
 	d_sid$lower <- as.double(d_sid$lower)
 	d_sid$upper <- as.double(d_sid$upper)
 
+	d_sid$line_labels <- round(as.numeric(d_sid$line_labels) + ((1 - as.numeric(d_sid$line_labels))/3), 2)
 	p_sid <- ggplot(d_sid, aes(x=edge_prob, y=mean_val, 
 				   ymin=lower, ymax=upper, 
 				   group=grp, color=algo, 
@@ -253,7 +255,7 @@ make_plots_v2 <- function(filename, plot_type){
 		ylim(0, 90) +
 		guides(linetype='none', fill='none')
 
-	ggsave("sid_ribbon.pdf", p_sid, height=2, width=3.2, units='in')
+	ggsave("sid_ribbon.pdf", p_sid, height=2, width=3.4, units='in')
 }
 
 make_plots_v2(filename='results/sl_results_linear.csv', plot_type='ribbon')

### 2025-02-08 paper/main.tex
Updates screenshot
Updates

@@
 	\end{subfigure}
-	\caption{Comparison of PC and Hill-Climb Search algorithms against
-		manually drawn DAGs using the assistance method. As PC and
-		Hill-Climb Search return the Markov Equivalence Class (MEC), we
-		use the best and worst scoring orientation of the MEC to get
-		the range of values. The human drawn values are done for
-		multiple accuracy values of the oracle.}
+	\caption{Comparison of PC, Hill-Climb Search, and GES algorithms against
+		\textit{ExpertInLoop} algorithm. As automated algorithms only recover
+		the CPDAG, we use the best and worst scoring orientation of the CPDAG 
+		to get the range. The \textit{ExpertInLoop} shows values for various
+		$ \alpha_{\textit{eff}} $ values corresponding to $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $.}
+	\label{fig:shd_sid}
 \end{figure}
@@
 	}
+	\label{fig:unexplained_ll}
 \end{figure}
@@
 	\end{split}
+\end{equation}
+
+\begin{equation}
+	\alpha_{\mathrm{eff}} = \alpha + (1 - \alpha) / 3
 \end{equation}
@@
 in the data. This property enables an absolute assessment of model quality,
-rather than just a relative comparison between models.
+rather than just a relative comparison between models. Figure~\ref{fig:unexplained_ll}
 
@@
 	\centering
-	\includegraphics[scale=0.4]{../code/plots/web_tool_full.png}
+	\includegraphics[scale=0.4]{../code/plots/web_tool_full_new.png}
 	\caption{A screenshot of the web tool for constructing the model. Users

### 2025-02-08 paper/main.tex
Updated

@@
 	\end{subfigure}
-	\caption{An example showing the steps of \textit{ExpertInLoop} algorithm. The dashed edges show all the potential edges that \textit{Expand} algorithm could add in each step.
-	}
+	\caption{An example showing each iteration of the \emph{ExpertInLoop} algorithm. The dashed edges show all the potential edges that \emph{Expand} step could add in each iteration.}
 \end{figure}
@@
 	\caption{Comparison of PC, Hill-Climb Search, and GES algorithms against
-		\textit{ExpertInLoop} algorithm. As automated algorithms only recover
-		the CPDAG, we use the best and worst scoring orientation of the CPDAG 
-		to get the range. The \textit{ExpertInLoop} shows values for various
-		$ \alpha_{\textit{eff}} $ values corresponding to $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $.}
+		\emph{ExpertInLoop} algorithm. As automated algorithms only
+		recover the CPDAG, we use the best and worst scoring
+		orientation of the CPDAG to get the range. We test
+		\emph{ExpertInLoop} with varying values of expert accuracy, $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $. The corresponding
+		$\alpha_{\textrm{eff}} $ is shown in the plot.}
 	\label{fig:shd_sid}
@@
 
-We analyzed the performance of \textit{ExpertInLoop} algorithm by comparing it
+The \emph{ExpertInLoop} algorithm requires two oracles, one that can answer
+d-separation queries and an ancestral oracle that can give the orientation or
+non-existence of edges between pairs of variables. For answering the
+d-separation queries, we use a CI test, the partial correlation test with a
+p-value threshold of $ 0.05 $. Secondly, to tell the direction between edges or
+non-existence of edges, we simulate an expert with accuracy $ \alpha $ as
+follows:
+
+\begin{equation}
+	\begin{split}
+		x &= \textrm{rand}([0, 1]) \\
+		E(\alpha) &= \begin{cases} 
+			\mathrm{TrueDir}(X, Y),  & \textrm{if  } x <= \alpha \\
+			\textrm{rand}(X \rightarrow Y, Y \leftarrow X, \textrm{None}) & \textrm{otherwise} \\
+				\end{cases} \\
+	\end{split}
+\end{equation}
+
+where given the true DAG, $ G $:
+
+\begin{equation}
+	\begin{split}
+	\mathrm{TrueDir}(X, Y) &= \begin{cases}
+					X \rightarrow Y, & \textrm{if } X \rightarrow Y \in G \\
+					Y \rightarrow X, & \textrm{if } Y \rightarrow X \in G \\
+					\textrm{None}, & \textrm{otherwise }
+				  \end{cases}
+	\end{split}
+\end{equation}
+
+An important point to note here is that the effective accuracy of $ E $ is
+higher than $ \alpha $ as even when giving a random answer, there is a $ 1 $ in
+$ 3 $ chance that the answer is correct. The effective accuracy is:
+
+\begin{equation}
+	\alpha_{\mathrm{eff}} = \alpha + (1 - \alpha) / 3
+\end{equation}
+
+We analyzed the performance of \emph{ExpertInLoop} algorithm by comparing it
 with three other automated algorithms: PC, Hill-Climb Search, and Greedy
-Equivalence Search (GES). For the analysis, we simulated data from known DAG
-structures and use the algorithms to recover the original DAG. We start by
+Equivalence Search (GES). For the analysis, we simulated $ 500 $ samples from
+known DAGs and use the algorithms to recover the original DAG. We start by
 generating a random DAG on $10$ nodes and use linear models with random effects
@@
 method recovers the DAG. To compare them fairly, we computed the SHD and SID
-for all possible orientations of the CPDAG to show the best and worst case scenarios.
-
-To use the \textit{ExpertInLoop} algorithm, we need two oracles: a d-separation
-oracle and an ancestral oracle. We use a conditional independence test (partial
-correlation if using linear data) for determining the d-separation with a
-p-value threshold of $ 0.05 $. For the ancestral oracle, we simulated the
-strong ancestral oracle. We vary the accuracy of our oracle to test the
-performance for varying oracle accuracies. Given an accuracy level $ \alpha $
-of the oracle, the oracle returns:
-
-\begin{equation}
-	\begin{split}
-		x &= \textnormal{rand}([0, 1]) \\
-		O(\alpha) &= \begin{cases} 
-			M \rightarrow Y, & \textnormal{if  } x <= \alpha \\
-			\textnormal{rand}(M \rightarrow Y, M \leftarrow Y, None) & \textnormal{otherwise} \\
-				\end{cases} \\
-	\end{split}
-\end{equation}
-
-\begin{equation}
-	\alpha_{\mathrm{eff}} = \alpha + (1 - \alpha) / 3
-\end{equation}
-
-If the oracle returns None for any pair of variables, an edge between this pair
-is not suggested to the oracle in future iterations. We also make sure that
-only edges which do not form a cyclic in the graph are used to select the next
-potential edge. There is a possibility that this greedy expert gets stuck at
-incorrect graph structures as it does not do any backtracking to fix its
-earlier mistakes. This procedure is outlined in the Algorithm 1.
-\todo[inline]{Show an algorithm of how this expert method is working}
+for all possible orientations of the CPDAG to show the best and worst case
+scenarios. The results are shown in Figure~\ref{shd_sid}. For SHD, the
+performance of \emph{ExpertInLoop} is comparable to the automated algorithms
+for $ \alpha = 0.3 (\alpha_{\textrm{eff}} = 0.53 $, and performs better for
+higher $ \alpha $ values. For SID, the \emph{ExpertInLoop} performs better in
+denser DAGs.
 

### 2025-02-08 paper/main.tex
Updates

@@
 \subsection{Using LLMs as Experts}
+
+\begin{figure}[t!]
+	\centering
+	\includegraphics[page=1]{fig5.pdf}
+	\caption{DAG learned from the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert. The p-value threshold used is $ 0.05 $ and the measure of association threshold is $ 0. 1 $}
+	\label{fig:adult_llm}
+\end{figure}
 Recently, there has been a lot of work towards using LLMs for causal discovery,
@@
 \label{sec:web}
+
+\begin{figure}[t!]
+	\centering
+	\includegraphics[scale=0.4]{../code/plots/web_tool_full_new.png}
+	\caption{A screenshot of the web tool for constructing the model. Users
+		can upload their dataset after which the tool creates an empty
+		graph and shows all pair of variables which are associated in
+		the model using undirected red edges with the strength of
+		association represented using edge width. Users can then
+		iteratively add edges to the model (shown in green) while
+		deciding the edge orientation based on domain knowledge.
+		Unnecessary edges are shown in black.}
+	\label{fig:web}
+\end{figure}
+
 To enable users to apply this method to their own datasets, we developed an
@@
 analysis.
-
-\begin{figure}[t!]
-	\centering
-	\includegraphics[scale=0.4]{../code/plots/web_tool_full_new.png}
-	\caption{A screenshot of the web tool for constructing the model. Users
-		can upload their dataset after which the tool creates an empty
-		graph and shows all pair of variables which are associated in
-		the model using undirected red edges with the strength of
-		association represented using edge width. Users can then
-		iteratively add edges to the model (shown in green) while
-		deciding the edge orientation based on domain knowledge.
-		Unnecessary edges are shown in black.}
-	\label{fig:web}
-\end{figure}
-
-\begin{figure}[t!]
-	\centering
-	\includegraphics[page=1]{fig5.pdf}
-	\caption{DAG learned using the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert. The p-value threshold used is $ 0.05 $ and the measure of association threshold is $ 0. 1 $}
-	\label{fig:adult_llm}
-\end{figure}
 

### 2025-02-08 paper/appendix.tex
Updates

@@
 
-We use Pillai's Trace because: 1) It uses all the canonical correlation values,
-capturing the full extent of association, 2) Its interpretation is similar to
-Pearson's correlation coefficient, i.e., $ 0 $ signifies no association and $ 1
-$ signifies perfect linear relationship. We use a normalized version because
-when comparing between pair of variables with different number of categories.
-Our proposed measure has several desirable properties that make it well-suited
-for our application:
+We use a normalized version of Pillai's Trace as it is bounded between $ 0 $
+and $ 1 $ and easier to interpret.
 
-\begin{enumerate}
-	\item \textbf{Bounded: } The measure is bounded between $ 0 $ (no
-		association) to $ 1 $ (perfect linear association), simplifying
-		interpretation.
-	% \item \textbf{Independent of Sample Size: } 
-	\item \textbf{Invariant to the Dummy Encoding: } For categorical
-		variables, residuals are computed using dummy encoding. Since
-		canonical correlations identifies linear combinations across
-		columns, this measure is invariant to the specific dummy
-		encoding scheme used.
-	\item \textbf{Equivalent to Partial Correlation for Continuous Variables: }
-		When both $ X $ and $ Y $ are continuous variables, our measure
-		is equivalent to the absolute value of partial correlation coefficient.
-	\item \textbf{Equivalent to polychoric and polyserial correlation for
-			ordinal and continuous variables: }
-		Under the assumption that the ordinal variable is generated by
-		discretizing an underlying gaussian continuous variable, this
-		effect size is equivalent to polychoric and polyserial
-		correlation. As both of them recover the Pearson correlation
-		coefficient.
-		\todo[inline]{Verify if this is correct}
-	\item \textbf{Related to Cram\'er's V: } 
-\end{enumerate}
+% We use Pillai's Trace because: 1) It uses all the canonical correlation values,
+% capturing the full extent of association, 2) Its interpretation is similar to
+% Pearson's correlation coefficient, i.e., $ 0 $ signifies no association and $ 1
+% $ signifies perfect linear relationship. We use a normalized version because
+% when comparing between pair of variables with different number of categories.
+% Our proposed measure has several desirable properties that make it well-suited
+% for our application:
 
-In essence, our measure of conditional association extends existing metrics to mixed data, providing 
-a unified measure that is interpretable.
+% \begin{enumerate}
+% 	\item \textbf{Bounded: } The measure is bounded between $ 0 $ (no
+% 		association) to $ 1 $ (perfect linear association), simplifying
+% 		interpretation.
+% 	% \item \textbf{Independent of Sample Size: } 
+% 	\item \textbf{Invariant to the Dummy Encoding: } For categorical
+% 		variables, residuals are computed using dummy encoding. Since
+% 		canonical correlations identifies linear combinations across
+% 		columns, this measure is invariant to the specific dummy
+% 		encoding scheme used.
+% 	\item \textbf{Equivalent to Partial Correlation for Continuous Variables: }
+% 		When both $ X $ and $ Y $ are continuous variables, our measure
+% 		is equivalent to the absolute value of partial correlation coefficient.
+% 	\item \textbf{Equivalent to polychoric and polyserial correlation for
+% 			ordinal and continuous variables: }
+% 		Under the assumption that the ordinal variable is generated by
+% 		discretizing an underlying gaussian continuous variable, this
+% 		effect size is equivalent to polychoric and polyserial
+% 		correlation. As both of them recover the Pearson correlation
+% 		coefficient.
+% 		\todo[inline]{Verify if this is correct}
+% 	\item \textbf{Related to Cram\'er's V: } 
+% \end{enumerate}
+% 
+% In essence, our measure of conditional association extends existing metrics to mixed data, providing 
+% a unified measure that is interpretable.

### 2025-02-08 paper/main.tex
Updates
Udpates

@@
 address these challenges, we propose an iterative method that guides
-researchers in manually constructing or improving the fit of an existing DAG.
-Our approach combines implied Conditional Independence (CI) testing with a
-measure of conditional association between variables to rank violations
-of implied CIs in a given DAG. This ranking helps prioritize modifications to
-the DAG to improve its fit to the data while utilizing expert knowledge to
-decide the orientation of new edges. Empirical results show that this guided
-manual model construction approach achieves performance comparable to automated
-algorithms if we are able to correctly decide the orientation of new edges in
-one out of three cases. Additionally, we demonstrate that, in the absence of an
-expert, a Large Language Model (LLM) can be potentially used to determine edge
-orientations. We provide the implementation of our approach in both a web tool
-at: <redacted for review> and a Python package <redacted for review>.
+researchers in manually constructing or improving the fit of an existing DAG
+while incorporating their expert knowledge. Compared to other expert knowledge
+specification methods, this method explicitly asks for the causal ordering
+between two specific variables in each step. Additionally, we provide a way to
+rank the potential new edge addition that improves the model fit the most in
+each step. Empirical results show that this guided manual model construction
+approach achieves performance comparable to automated algorithms if we are able
+to correctly decide the orientation of new edges in one out of three cases.
+Additionally, we demonstrate that, in the absence of an expert, a Large
+Language Model (LLM) can be potentially used to determine edge orientations. We
+provide the implementation of our approach in both a web tool at: <redacted for
+review> and a Python package <redacted for review>.
+
+% Our approach combines implied Conditional Independence (CI) testing with a
+% measure of conditional association between variables to rank violations
+% of implied CIs in a given DAG. This ranking helps prioritize modifications to
+% the DAG to improve its fit to the data while utilizing expert knowledge to
+% decide the orientation of new edges.
 
@@
 
+\subsection{Ranking Potential New Edges}
+
+\subsection{Comparison to Automated Algorithms}
 \begin{figure}[t!]

### 2025-02-08 paper/main.tex
Adds the ranked expand algorithm
Updates

@@
 
+In this section, we give some of the practical aspects of the
+\emph{ExpertInLoop} algorithm, and then compare the algorithm with automated
+causal discovery algorithms.
+
 \subsection{Ranking Potential New Edges}
+In Algorithm~\ref{algo:expert}, in each step we add a single edge to the DAG
+and remove any edges that can be pruned. In practice, we use a slightly
+modified version of the Algorithm~\ref{algo:expand} for adding new edges as
+given in Algorithm~\ref{algo:ranked_expand}. Rather than choosing the first
+potential edge addition, this algorithm ranks all potential new edges based on
+the conditional association between the variables. The edge with the highest
+conditional association is then added to the graph. This modification allows us
+to prioritize edges that would help in the explaining the observed correlation
+in the data the highest. For this purpose, we use a measure of conditional
+association measure, $\phi$ between variables.
+
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+\SetKwFunction{RankedExpand}{RankedExpand}
+\SetKw{KwGoTo}{go to}
+\SetKwProg{Fn}{Function}{:}{}
+\Fn{{\sc RankedExpand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$\phi$,$B=\emptyset$)}{
+    let $\phi_{\max} \gets 0 $\;
+    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
+	$\{X,Y\} \notin B$}{
+        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
+        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
+		\If{$\lvert \phi(X, Y, \mathbf{Z}) \rvert > \phi_{\max}$}{
+			$ \phi_{\max} \gets \lvert \phi(X, Y, \mathbf{Z}) \rvert $ \;
+			$ L_{\max} \gets \mathcal{A}_G(X, Y)$ \;
+		}
+        }
+    }
+    add $L_{\max}$ to $E$\;
+    \Return{$(V,E)$}\;
+}
+\caption{Adding an edge between variables with the highest correlation}
+\label{algo:ranked_expand}
+\end{algorithm}
 
@@
 for all possible orientations of the CPDAG to show the best and worst case
-scenarios. The results are shown in Figure~\ref{shd_sid}. For SHD, the
+scenarios. The results are shown in Figure~\ref{fig:shd_sid}. For SHD, the
 performance of \emph{ExpertInLoop} is comparable to the automated algorithms

### 2025-02-08 paper/main.tex
Updates

--- 2025-02-08 paper/main.tex
@@
 
+In this section, we give some of the practical aspects of the
+\emph{ExpertInLoop} algorithm, and then compare the algorithm with automated
+causal discovery algorithms.
+
 \subsection{Ranking Potential New Edges}
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+\SetKwFunction{RankedExpand}{RankedExpand}
+\SetKw{KwGoTo}{go to}
+\SetKwProg{Fn}{Function}{:}{}
+\Fn{{\sc RankedExpand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$\phi$,$B=\emptyset$)}{
+    let $\phi_{\max} \gets 0 $\;
+    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
+	$\{X,Y\} \notin B$}{
+        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
+        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
+		\If{$\lvert \phi(X, Y, \mathbf{Z}) \rvert > \phi_{\max}$}{
+			$ \phi_{\max} \gets \lvert \phi(X, Y, \mathbf{Z}) \rvert $ \;
+			$ L_{\max} \gets \mathcal{A}_G(X, Y)$ \;
+		}
+        }
+    }
+    add $L_{\max}$ to $E$\;
+    \Return{$(V,E)$}\;
+}
+\caption{Adding an edge between variables with the highest correlation}
+\label{algo:ranked_expand}
+\end{algorithm}
+
+In Algorithm~\ref{algo:expert}, in each iteration we add a new edge to the DAG
+chosen by Algorithm~\ref{algo:expand}. This edge is chosen at random depending
+on the order of iteration in Algoirthm~\ref{algo:expand}. A more optimal way to
+add these edges would be to choose the pairs of variables that could help
+improve the fit the most in each step. One potential greedy approach to do this
+would be to select the variable pair $ X $ and $ Y $ that have the highest
+amount of observed conditional correlation in the data. In
+Algorithm~\ref{algo:ranked_expand}, we slightly modify
+Algorithm~\ref{algo:expand} to show this approach. In this case we first go
+through all potential edges that can be added to the DAG in each step, and rank
+them based on the conditional association between them given the current DAG.
+This gives us a measure of how much of the observed correlation between them is
+not explained by the current model. We then choose the pair of variables the
+highest conditional association. This modification allows us to prioritize pair
+of variables to quickly improve the fit of the model. For the conditional
+measure of association, we use a measures outlined in
+Section~\ref{sec:background} depending on the data type of the variables.
 

### 2025-02-08 paper/main.tex
Updates
Improve abstract

@@
 errors and output Markov Equivalence Classes (MECs) rather than a DAG. To
-address these challenges, we propose an iterative method that guides
-researchers in manually constructing or improving the fit of an existing DAG
-while incorporating their expert knowledge. Compared to other expert knowledge
-specification methods, this method explicitly asks for the causal ordering
-between two specific variables in each step. Additionally, we provide a way to
-rank the potential new edge addition that improves the model fit the most in
-each step. Empirical results show that this guided manual model construction
-approach achieves performance comparable to automated algorithms if we are able
-to correctly decide the orientation of new edges in one out of three cases.
-Additionally, we demonstrate that, in the absence of an expert, a Large
-Language Model (LLM) can be potentially used to determine edge orientations. We
-provide the implementation of our approach in both a web tool at: <redacted for
-review> and a Python package <redacted for review>.
+address these challenges, we propose an iterative DAG construction that
+combines domain knowledge with data-driven insights. Our approach iteratively
+identifies pair of variables with a missing edge in the DAG, and asks for the
+orientation of the edge from an expert. We also propose a method to rank these
+potential new edges based on their impact on the model fit. Empirical results
+show that this iterative model construction approach achieves performance
+comparable to automated algorithms if we are able to correctly decide the
+orientation of new edges in one out of three cases. Additionally, we
+demonstrate that, in the absence of an expert, a Large Language Model (LLM) can
+be used to determine edge orientations. We provide the implementation of our
+approach in both a web tool at: <redacted for review> and a Python package
+<redacted for review>.
 
@@
 
+Given a DAG $ G = (V, E) $, the total unexplained correlation, $\tau$ in define as:
+\begin{equation}
+	\tau = \sum_{\{ (X, Y) \in V | X \rightarrow Y \not \in E \wedge Y \rightarrow X \not \in E \} } \phi(X, Y, \mathrm{pa}_G(X) \cup \mathrm{pa}_G(Y)))
+\end{equation}
+
 Another major difference is in the interpretability of the evaluation metric.
@@
 with nodes corresponding to the dataset\'s variables. They can then specify a
-p-value threshold and a measure of association threshold.
+p-value threshold and a measure of association threshold. We use the measure of 
+conditional association shown in Appendix~\ref{sec:mixed_association}.
 

### 2025-02-08 paper/main.tex
Updatse

@@
 errors and output Markov Equivalence Classes (MECs) rather than a DAG. To
-address these challenges, we propose an iterative DAG construction that
+address these challenges, we propose an iterative DAG construction method that
 combines domain knowledge with data-driven insights. Our approach iteratively
@@
 
-Figure~\ref{fig:intro} highlights some of these issues.
+Figure~\ref{fig:intro} highlights some of these issues. To assist researchers
+in constructing DAGs manually, we propose an iterative model construction
+method that combines domain knowledge with data-driven insights. Our approach 
+iteratively identifies pairs of variables with a missing edge in the DAG and
 
-When manually constructing models, it is important to test whether the
-resulting DAG accurately represents the dataset. Based on this evaluation and
-domain knowledge, we can make further modification to the model. One common
-approach for assessing DAGs is to check whether the Conditional Independences
-(CIs) implied by the DAG hold in the data \citep{Ankan2021}. If we find
-violations to these CIs, we can use our domain knowledge to make appropriate
-modifications to the DAG. However, this CI testing approach has a few
-limitations:
-
-\begin{enumerate}
-\item As CIs are only implied by missing edges in the model, this approach does 
-	not test whether the existing edges are correct.
-\item Determining whether a CI holds in data is based on a p-value threshold. This can 
-	be unreliable, for example, we get a significant p-value for even very weak 
-	relationship if the sample size is high.
-\item The number of implied CIs can be large making it difficult to manually
-	check all of them. For example, a moderately sized alarm network
-	\citep{Beinlich1989} with $ 37 $ nodes, implies $ 287 $ CIs.
-\end{enumerate}
-
-To tackle these limitations, we draw inspiration from modification indices in
-specification search. Modification indices provide ranking for potential model
-modification based on their impact on the model's fit, allowing us to
-prioritize most meaningful modifications.  Similarly, we propose a method for
-ranking CI test violations to help us focus on the most critical
-inconsistencies in the model. To determine this ranking we utilize a measure of
-conditional association between variables in combination with p-values from the
-CI test. This approach addresses all three of the issues outlined above. The
-measure of association allows us to evaluate the validity of existing edges. It
-reduces our reliance on p-values alone for deciding whether the CI test holds.
-The ranking helps us prioritize the most critical violations, making it feasible 
-to focus on the areas that most improve the model's fit.
+% When manually constructing models, it is important to test whether the
+% resulting DAG accurately represents the dataset. Based on this evaluation and
+% domain knowledge, we can make further modification to the model. One common
+% approach for assessing DAGs is to check whether the Conditional Independences
+% (CIs) implied by the DAG hold in the data \citep{Ankan2021}. If we find
+% violations to these CIs, we can use our domain knowledge to make appropriate
+% modifications to the DAG. However, this CI testing approach has a few
+% limitations:
+% 
+% \begin{enumerate}
+% \item As CIs are only implied by missing edges in the model, this approach does 
+% 	not test whether the existing edges are correct.
+% \item Determining whether a CI holds in data is based on a p-value threshold. This can 
+% 	be unreliable, for example, we get a significant p-value for even very weak 
+% 	relationship if the sample size is high.
+% \item The number of implied CIs can be large making it difficult to manually
+% 	check all of them. For example, a moderately sized alarm network
+% 	\citep{Beinlich1989} with $ 37 $ nodes, implies $ 287 $ CIs.
+% \end{enumerate}
+% 
+% To tackle these limitations, we draw inspiration from modification indices in
+% specification search. Modification indices provide ranking for potential model
+% modification based on their impact on the model's fit, allowing us to
+% prioritize most meaningful modifications.  Similarly, we propose a method for
+% ranking CI test violations to help us focus on the most critical
+% inconsistencies in the model. To determine this ranking we utilize a measure of
+% conditional association between variables in combination with p-values from the
+% CI test. This approach addresses all three of the issues outlined above. The
+% measure of association allows us to evaluate the validity of existing edges. It
+% reduces our reliance on p-values alone for deciding whether the CI test holds.
+% The ranking helps us prioritize the most critical violations, making it feasible 
+% to focus on the areas that most improve the model's fit.
 
@@
 covariance between $ X $ and $ Y $ in the matrix using $ \Sigma_{XY} $.
+
+\subsection{Integrating domain knowledge in causal discovery}
 

### 2025-02-08 paper/main.tex
Updates

@@
 \subsection{Integrating domain knowledge in causal discovery}
+Multiple ways to integrate, originally using forbidden and required edges. Or
+temporal algorithms by specifying causal tiers.
 
@@
 We therefore consider the following two models of a domain experts: Given two variables
-\emph{X} and \emph{Y} that are part of an unknown causal DAG $G$, an \emph{strong ancestral oracle} 
+$ X $ and $ Y $ that are part of an unknown causal DAG $G$, an \emph{strong ancestral oracle} 
 $\mathcal{A}_G$ is defined as:
@@
 	\end{subfigure}
-	\caption{An example showing each iteration of the \emph{ExpertInLoop} algorithm. The dashed edges show all the potential edges that \emph{Expand} step could add in each iteration.}
+	\caption{An example showing each iteration of the \textsc{ExpertInLoop} algorithm. The dashed edges show all the potential edges that \textsc{Expand} step could add in each iteration.}
 \end{figure}
@@
 In this section, we give some of the practical aspects of the
-\emph{ExpertInLoop} algorithm, and then compare the algorithm with automated
+\textsc{ExpertInLoop} algorithm, and then compare the algorithm with automated
 causal discovery algorithms.
@@
 	\caption{Comparison of PC, Hill-Climb Search, and GES algorithms against
-		\emph{ExpertInLoop} algorithm. As automated algorithms only
+		\textsc{ExpertInLoop} algorithm. As automated algorithms only
 		recover the CPDAG, we use the best and worst scoring
 		orientation of the CPDAG to get the range. We test
-		\emph{ExpertInLoop} with varying values of expert accuracy, $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $. The corresponding
+		\textsc{ExpertInLoop} with varying values of expert accuracy, $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $. The corresponding
 		$\alpha_{\textrm{eff}} $ is shown in the plot.}
@@
 
-The \emph{ExpertInLoop} algorithm requires two oracles, one that can answer
+The \textsc{ExpertInLoop} algorithm requires two oracles, one that can answer
 d-separation queries and an ancestral oracle that can give the orientation or
@@
 
-We analyzed the performance of \emph{ExpertInLoop} algorithm by comparing it
+We analyzed the performance of \textsc{ExpertInLoop} algorithm by comparing it
 with three other automated algorithms: PC, Hill-Climb Search, and Greedy
@@
 scenarios. The results are shown in Figure~\ref{fig:shd_sid}. For SHD, the
-performance of \emph{ExpertInLoop} is comparable to the automated algorithms
+performance of \textsc{ExpertInLoop} is comparable to the automated algorithms
 for $ \alpha = 0.3 (\alpha_{\textrm{eff}} = 0.53 $, and performs better for
-higher $ \alpha $ values. For SID, the \emph{ExpertInLoop} performs better in
+higher $ \alpha $ values. For SID, the \textsc{ExpertInLoop} performs better in
 denser DAGs.

### 2025-02-08 paper/main.tex
Updates

@@
 method that combines domain knowledge with data-driven insights. Our approach
-iteratively identifies pairs of variables with a missing edge in the DAG and
+utilizes conditional independence testing to identify pair of variables in the
+DAG that are missing an edge or have a superfluous edge. Based on this
+information, in each iteration we present one such pair of variables to an
+expert who then decides the orientation of the edge between the variables based
+on their domain knowledge. The method also iteratively prunes all superfluous
+edges in each iteration. As a result this approach guides the expert in
+constructing their model while ensuring that the model is consistent with the
+data.
 

### 2025-02-08 paper/main.tex
Update

@@
 
-Figure~\ref{fig:intro} highlights some of these issues. To assist researchers
-in constructing DAGs manually, we propose an iterative model construction
-method that combines domain knowledge with data-driven insights. Our approach
-utilizes conditional independence testing to identify pair of variables in the
-DAG that are missing an edge or have a superfluous edge. Based on this
-information, in each iteration we present one such pair of variables to an
-expert who then decides the orientation of the edge between the variables based
-on their domain knowledge. The method also iteratively prunes all superfluous
-edges in each iteration. As a result this approach guides the expert in
-constructing their model while ensuring that the model is consistent with the
-data.
+Figure~\ref{fig:intro} highlights some of these issues.   When manually
+constructing models, it is important to test whether the resulting DAG
+accurately represents the dataset. Based on this evaluation and domain
+knowledge, we can make further modification to the model. One common approach
+for assessing DAGs is to check whether the Conditional Independences (CIs)
+implied by the DAG hold in the data \citep{Ankan2021}. If we find violations to
+these CIs, we can use our domain knowledge to make appropriate changes to the
+model.
+
+In this paper we present a method that utilizes this idea of CI testing to
+assist researchers during the DAG construction instead of testing the final
+model. We propose an iterative model construction method that combines domain
+knowledge with data-driven insights. Our approach utilizes conditional
+independence testing to identify pair of variables in the DAG that are missing
+an edge or have a superfluous edge. Based on this information, in each
+iteration we present one such pair of variables to an expert who then decides
+the orientation of the edge between the variables based on their domain
+knowledge. The method also iteratively prunes all superfluous edges in each
+iteration. As a result this approach guides the expert in constructing their
+model while ensuring that the model is consistent with the data.
 
@@
 \begin{enumerate}
-    \item We propose a novel measure of conditional association for mixed data
-	    based on canonical correlations
-	    (Section~\ref{sec:mixed_association}). This measure generalizes
-	    several commonly used special case measures of association to mixed
-	    data.
-    \item Using this measure of conditional association, we develop a procedure
-	    to rank violations of implied CIs. Using this ranking helps us
-	    prioritize modifications to DAG to improve its fit while allowing
-	    us to integrate expert knowledge when adding new edges (Section~\ref{sec:modification}).
+%     \item We propose a novel measure of conditional association for mixed data
+% 	    based on canonical correlations
+% 	    (Section~\ref{sec:mixed_association}). This measure generalizes
+% 	    several commonly used special case measures of association to mixed
+% 	    data.
+    \item We develop an iterative method based on CI testing to assist researchers in 
+	    constructing DAGs using their domain knowledge (Section~\ref{sec:modification}).
+    \item We develop a ranking method to prioritize potential modifications in the DAG that 
+    	    can help in improving the model's fit quickly (Section~\ref{sec:ranking}).
     \item We provide a web-based interactive tool and a Python package to allow 
@@
 \subsection{Ranking Potential New Edges}
+\label{sec:ranking}
+
 \begin{algorithm}[h]

### 2025-02-09 paper/main.tex
Updates

@@
 
-Causal discovery has received significant attention in the Directed Acyclic
-Graphs (DAGs) literature, leading to the development of numerous automated
-algorithms for learning DAGs from data. However, their adoption in applied
-domains remain limited, as researchers often prefer to construct DAGs manually
-based on domain knowledge. This preference arises due to several practical
-challenges with automated algorithms, such as their tendency to make obvious
-errors and output Markov Equivalence Classes (MECs) rather than a DAG. To
-address these challenges, we propose an iterative DAG construction method that
-combines domain knowledge with data-driven insights. Our approach iteratively
-identifies pair of variables with a missing edge in the DAG, and asks for the
-orientation of the edge from an expert. We also propose a method to rank these
-potential new edges based on their impact on the model fit. Empirical results
-show that this iterative model construction approach achieves performance
-comparable to automated algorithms if we are able to correctly decide the
-orientation of new edges in one out of three cases. Additionally, we
-demonstrate that, in the absence of an expert, a Large Language Model (LLM) can
-be used to determine edge orientations. We provide the implementation of our
-approach in both a web tool at: <redacted for review> and a Python package
-<redacted for review>.
+	Causal discovery has received significant attention in the Directed
+	Acyclic Graphs (DAGs) literature, leading to the development of
+	numerous automated algorithms for learning DAGs from data. However,
+	their adoption in applied domains remain limited, as researchers often
+	prefer to construct DAGs manually based on domain knowledge. This
+	preference arises due to several practical challenges with automated
+	algorithms, such as their tendency to make obvious errors and output
+	Markov Equivalence Classes (MECs) rather than a DAG. To assist
+	researchers in constructing DAGs manually, we propose an iterative DAG
+	construction approach that combines domain knowledge with data-driven
+	insights. Our method leverages Conditional Independence (CI) testing to
+	iteratively identify variable pairs where an edge is either missing or
+	superfluous. Based on this information, we can choose to add missing
+	edges with appropriate orientation based on domain knowledge or remove
+	unnecessary ones. We also give a method to rank potential new edges
+	based on their impact on the overall model fit. Empirical results show
+	that this iterative approach achieves performance comparable to
+	automated algorithms if the orientation of new edges is correctly
+	determined in at least one out of two cases. Additionally, we show
+	that, in the absence of an expert, a Large Language Model (LLM) can be
+	used to determine edge orientations. We provide an implementation of
+	our approach as both a webtool: <redacted for review> and a Python
+	package <redacted for review>.
 

### 2025-02-09 paper/main.tex
Updates
udpates

@@
 model while ensuring that the model is consistent with the data.
+
+As at any iteration, multiple changes to the DAG are possible. We also give a
+way to rank potential new edges additions to the DAG using effect size measures
+for CI tests.
 
@@
 \begin{itemize}
-	\item As researchers prefer to anyways build models by hand, this is like a companion tool to help them in deciding the method.
-	\item Solves all the $ 3 $ issues outlined in the introduction.
-	\item Compared to other expert knowledge specification methods, users don't need to guess what information then need to provide as exact questions are asked. Also enables us to ask an LLM for it.
+	\item As researchers prefer to build models by hand, we present a tool
+		to assist them in construction using CI testing.
+	\item In practice as many edges are possible for any given DAG, we provide
+		a ranking method to find the best edge addition.
+	\item Compared to other expert knowledge specification methods, users
+		don't need to guess what information then need to provide as
+		exact questions are asked. Also enables us to ask an LLM for
+		it.
 \end{itemize}

### 2025-02-09 paper/main.tex
Updates

@@
 
-% Our approach combines implied Conditional Independence (CI) testing with a
-% measure of conditional association between variables to rank violations
-% of implied CIs in a given DAG. This ranking helps prioritize modifications to
-% the DAG to improve its fit to the data while utilizing expert knowledge to
-% decide the orientation of new edges.
-
 \end{abstract}
@@
 decisions. Causal discovery methods aim to discover such relationships
-among random variables using observational data. Approaches to causal discovery
-have been developed within both the Directed Acyclic Graphs (DAGs) and
-Structural Equation Models (SEMs) frameworks, each adapting a different
-approach.
+among random variables using observational data.
+
+% Approaches to causal discovery
+% have been developed within both the Directed Acyclic Graphs (DAGs) and
+% Structural Equation Models (SEMs) frameworks, each adapting a different
+% approach.
 
@@
 \citep{Chickering2002}, and continuous optimization-based methods like NO TEARS
-\citep{Zheng2018} and DAGMA \citep{Bello2022}. While DAG-based methods focus on
-automated discovery, SEM-based methods emphasize expert driven model
-specification. This includes tools to assist researchers in manually
-constructing models, enabling them to incorporate their domain knowledge in the
-model building process. Researchers typically begin with an initial model based
-on their domain knowledge and then use these tools to guide modifications that
-improve the model's fit to data. This process is commonly known as
-Specification Search \citep{Long1983} and uses method such as modification
-indices, and Wald-based tests \citep{Marcoulides2018}.
+\citep{Zheng2018} and DAGMA \citep{Bello2022}.
+
+% While DAG-based methods focus on
+% automated discovery, SEM-based methods emphasize expert driven model
+% specification. This includes tools to assist researchers in manually
+% constructing models, enabling them to incorporate their domain knowledge in the
+% model building process. Researchers typically begin with an initial model based
+% on their domain knowledge and then use these tools to guide modifications that
+% improve the model's fit to data. This process is commonly known as
+% Specification Search \citep{Long1983} and uses method such as modification
+% indices, and Wald-based tests \citep{Marcoulides2018}.
 

### 2025-02-09 paper/main.tex
fixes in proofs, related to cycles

--- 2025-02-09 paper/main.tex
@@
 
-	Causal discovery has received significant attention in the Directed
-	Acyclic Graphs (DAGs) literature, leading to the development of
-	numerous automated algorithms for learning DAGs from data. However,
-	their adoption in applied domains remain limited, as researchers often
-	prefer to construct DAGs manually based on domain knowledge. This
-	preference arises due to several practical challenges with automated
-	algorithms, such as their tendency to make obvious errors and output
-	Markov Equivalence Classes (MECs) rather than a DAG. To assist
-	researchers in constructing DAGs manually, we propose an iterative DAG
-	construction approach that combines domain knowledge with data-driven
-	insights. Our method leverages Conditional Independence (CI) testing to
-	iteratively identify variable pairs where an edge is either missing or
-	superfluous. Based on this information, we can choose to add missing
-	edges with appropriate orientation based on domain knowledge or remove
-	unnecessary ones. We also give a method to rank potential new edges
-	based on their impact on the overall model fit. Empirical results show
-	that this iterative approach achieves performance comparable to
-	automated algorithms if the orientation of new edges is correctly
-	determined in at least one out of two cases. Additionally, we show
-	that, in the absence of an expert, a Large Language Model (LLM) can be
-	used to determine edge orientations. We provide an implementation of
-	our approach as both a webtool: <redacted for review> and a Python
-	package <redacted for review>.
+Causal discovery has received significant attention in the Directed Acyclic
+Graphs (DAGs) literature, leading to the development of numerous automated
+algorithms for learning DAGs from data. However, their adoption in applied
+domains remain limited, as researchers often prefer to construct DAGs manually
+based on domain knowledge. This preference arises due to several practical
+challenges with automated algorithms, such as their tendency to make obvious
+errors and output Markov Equivalence Classes (MECs) rather than a DAG. To
+address these challenges, we propose an iterative DAG construction method that
+combines domain knowledge with data-driven insights. Our approach iteratively
+identifies pair of variables with a missing edge in the DAG, and asks for the
+orientation of the edge from an expert. We also propose a method to rank these
+potential new edges based on their impact on the model fit. Empirical results
+show that this iterative model construction approach achieves performance
+comparable to automated algorithms if we are able to correctly decide the
+orientation of new edges in one out of three cases. Additionally, we
+demonstrate that, in the absence of an expert, a Large Language Model (LLM) can
+be used to determine edge orientations. We provide the implementation of our
+approach in both a web tool at: <redacted for review> and a Python package
+<redacted for review>.
 
@@
 Our approach to causal structure learning combines domain knowledge with data-driven
-insights in a manner that is based on the following base assumptions: 
-(1) Domain experts are good at determining causal directions between variables if 
-there is a clear causal direction between them. (2) Domain experts are less good
+insights in a manner that is based on the following considerations: 
+(1) A domain experts is possibly good at determining causal directions between variables if 
+there is a clear causal direction between them. (2) A domain expert may be less good
 at identifying cases where there is no causal relationship between the variables,
 since \emph{potential} causal relationships can often be argued for anyway.
-(3) Domain experts are even less good at distinguishing direct from indirect 
-effects, since the question whether a direct effect exists between two variables
-depends on all other variables present in the graph.
+(3) Many domain experts could struggle to distinguish direct from indirect 
+effects, since the presence of a direct effect between two variables
+of interest depends on all other variables present in the graph.
 
@@
 
-That is, our ancestral oracles never distinguish whether the relationships
-between variables is direct or indirect; and the weak ancestral oracle is even
-unable to detect cases where there is no cause-effect relationship between two
-variables. Using such oracles for DAG construction therefore inherently risks
-that superfluous edges are created. In our approach, such superfluous edges
+That is, our ancestral oracles only provide information on the direction of cause-effect
+relationships, and do not distinguish between direct and indirect effects. 
+The weak ancestral oracle is even unable to detect cases where there is no cause-effect
+relationship between two variables. Using such oracles for DAG construction therefore 
+inherently risks that superfluous edges are created. In our approach, such superfluous edges
 will be detected and removed by utilizing the data.
@@
 To model how our learning procedure interacts with the data, we assume that we
-have access to a second oracle $\mathcal{D}_G$ than can answer d-separation
+additionally have access to a second oracle $\mathcal{D}_G$ than can answer d-separation
 queries with respect to the unknown target graph:
-$\mathcal{D}_G(X,Y,\mathbf{Z})=1$ iff $X \ci_G Y \mid \mathbf{Z}$. These are
-the standard oracles considered in constraint-based structure learning
-algorithms, such as the PC algorithm \cite{Spirtes2001}. Based on this oracle,
-we now define two procedures that iteratively change a current DAG structure.
+$\mathcal{D}_G(X,Y,\mathbf{Z})=1$ iff $X$ and $Y$ are $d$-separated by $\mathbf{Z}$ in 
+$G$. These are the standard oracles considered in constraint-based structure learning
+algorithms, such as the PC algorithm \cite{Spirtes2001}.
 
-Our structure learning algorithm is based on two core procedures.
+Based on these two oracles, we now define two core procedures that we will use to 
+iteratively change a current DAG structure. The procedure 
 \textsc{Expand} (Algorithm~\ref{algo:expand}) uses conditional independence
 information to search for missing edges in the graph and then uses domain
-knowledge to orient them. \textsc{Prune} (Algorithm~\ref{algo:prune}) uses
+knowledge to orient them. The procedure \textsc{Prune} (Algorithm~\ref{algo:prune}) uses
 conditional independence information to remove superfluous edges from the
@@
 \SetKwProg{Fn}{Function}{:}{}
-\Fn{{\sc Expand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$B=\emptyset$,$k=\infty$)}{
+\Fn{{\sc Expand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$B$,$k$)}{
     let $L \gets \{\}$\;
@@
         \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
-            $L \gets L \cup \mathcal{A}_G(X, Y)$ \;
+	    $e \gets \mathcal{A}_G(X, Y)$ \;
+	    \uIf{ $(V,E \cup L \cup e)$ is acyclic }{
+            	$L \gets L \cup  \{ e \} $\;
+	    } \Else {
+            	$L \gets L \cup  \{ \textrm{reverse}(e) \} $\;
+	    }
         }
@@
 }
-\caption{Adding edges based on data and domain knowledge}
+\caption{Adding edges based on data and domain knowledge.}
 \label{algo:expand}
@@
 
+Let $G^+$ denote the transitive closure of $G$, that is, the graph in which 
+there is a direct edge $X \to Y$ whenever $X \in \textrm{An}_G(Y)$. The transitive
+closure is related to the result of $\textsc{Expand}$ by the following two 
+easily proved facts.
+
 \begin{proposition}
@@
  $\mathcal{D}_G$ and a strong ancestral oracle $\mathcal{A}_G$, 
-$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_G)=G^+$.
+$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_G, \emptyset, \infty)=G^+$.
 \end{proposition}
+
+\begin{proof}
+Since we start from an empty graph, every pair of vertices $X, Y$ where
+ $X \in \tetxrm{An}_G(Y)$ is d-separated by the empty set and is
+ connected by the edge $X \to Y$. No other edges are added. Therefore, 
+the result is $G^+$. 
+\end{proof}
 
@@
  $\mathcal{D}_G$ and a weak ancestral oracle $\mathcal{A}_G$, 
-$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_G)\supseteq G^+$.
+$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_G, \emptyset, \infty)
+\supseteq G^+$.
 \end{proposition}
+
+\begin{proof}
+Every pair of vertices $X, Y$ that are $d$-connected in $G$ are now connected by an edge 
+during $\textsc{Expand}$. For $X \in \tetxrm{An}_G(Y)$, the edge is 
+$X \to Y$ just like when using a strong ancestral oracle; since only correctly oriented
+edges are added between ancestor-descendant pairs, such additions never create cycles.
+For $X, Y$ that are only $d$-connected throug common ancestors, an arbitrary 
+edge is added in a manner that preserves acyclicity. The result is a supergraph
+of $G^+$.
+\end{proof}
 
@@
 \end{algorithm}
+
+Unlike our ancestral oracles, the pruning operation is quite effective at distinguishing
+direct and indirect effects, as shown by the following result.
 

### 2025-02-09 paper/main.tex

@@
 
-Causal discovery has received significant attention in the Directed Acyclic
-Graphs (DAGs) literature, leading to the development of numerous automated
-algorithms for learning DAGs from data. However, their adoption in applied
-domains remain limited, as researchers often prefer to construct DAGs manually
-based on domain knowledge. This preference arises due to several practical
-challenges with automated algorithms, such as their tendency to make obvious
-errors and output Markov Equivalence Classes (MECs) rather than a DAG. To
-address these challenges, we propose an iterative DAG construction method that
-combines domain knowledge with data-driven insights. Our approach iteratively
-identifies pair of variables with a missing edge in the DAG, and asks for the
-orientation of the edge from an expert. We also propose a method to rank these
-potential new edges based on their impact on the model fit. Empirical results
-show that this iterative model construction approach achieves performance
-comparable to automated algorithms if we are able to correctly decide the
-orientation of new edges in one out of three cases. Additionally, we
-demonstrate that, in the absence of an expert, a Large Language Model (LLM) can
-be used to determine edge orientations. We provide the implementation of our
-approach in both a web tool at: <redacted for review> and a Python package
-<redacted for review>.
-
-% Our approach combines implied Conditional Independence (CI) testing with a
-% measure of conditional association between variables to rank violations
-% of implied CIs in a given DAG. This ranking helps prioritize modifications to
-% the DAG to improve its fit to the data while utilizing expert knowledge to
-% decide the orientation of new edges.
+	Causal discovery has received significant attention in the Directed
+	Acyclic Graphs (DAGs) literature, leading to the development of
+	numerous automated algorithms for learning DAGs from data. However,
+	their adoption in applied domains remain limited, as researchers often
+	prefer to construct DAGs manually based on domain knowledge. This
+	preference arises due to several practical challenges with automated
+	algorithms, such as their tendency to make obvious errors and output
+	Markov Equivalence Classes (MECs) rather than a DAG. To assist
+	researchers in constructing DAGs manually, we propose an iterative DAG
+	construction approach that combines domain knowledge with data-driven
+	insights. Our method leverages Conditional Independence (CI) testing to
+	iteratively identify variable pairs where an edge is either missing or
+	superfluous. Based on this information, we can choose to add missing
+	edges with appropriate orientation based on domain knowledge or remove
+	unnecessary ones. We also give a method to rank potential new edges
+	based on their impact on the overall model fit. Empirical results show
+	that this iterative approach achieves performance comparable to
+	automated algorithms if the orientation of new edges is correctly
+	determined in at least one out of two cases. Additionally, we show
+	that, in the absence of an expert, a Large Language Model (LLM) can be
+	used to determine edge orientations. We provide an implementation of
+	our approach as both a webtool: <redacted for review> and a Python
+	package <redacted for review>.
 

### 2025-02-09 paper/main.tex
Updates

--- 2025-02-09 paper/main.tex
@@
 \citep{Chickering2002}, and continuous optimization-based methods like NO TEARS
-\citep{Zheng2018} and DAGMA \citep{Bello2022}.
-
-% While DAG-based methods focus on
-% automated discovery, SEM-based methods emphasize expert driven model
-% specification. This includes tools to assist researchers in manually
-% constructing models, enabling them to incorporate their domain knowledge in the
-% model building process. Researchers typically begin with an initial model based
-% on their domain knowledge and then use these tools to guide modifications that
-% improve the model's fit to data. This process is commonly known as
-% Specification Search \citep{Long1983} and uses method such as modification
-% indices, and Wald-based tests \citep{Marcoulides2018}.
-
-Despite significant progress in automated causal discovery, their adoption in
-applied research has been limited. Researchers often prefer to manually
-construct DAGs based on their domain expertise \citep{Tennant2020,
-Petersen2021}. We attribute this preference to several challenges with existing
-causal discovery algorithms in practical settings:
+\citep{Zheng2018} and DAGMA \citep{Bello2022}. Despite significant progress in
+automated causal discovery, their adoption in applied research has been
+limited. Researchers often prefer to manually construct DAGs based on their
+domain expertise \citep{Tennant2020, Petersen2021}. We attribute this
+preference to several challenges with existing causal discovery algorithms in
+practical settings:
 
@@
 
-Figure~\ref{fig:intro} highlights some of these issues.   When manually
-constructing models, it is important to test whether the resulting DAG
-accurately represents the dataset. Based on this evaluation and domain
-knowledge, we can make further modification to the model. One common approach
-for assessing DAGs is to check whether the Conditional Independences (CIs)
-implied by the DAG hold in the data \citep{Ankan2021}. If we find violations to
-these CIs, we can use our domain knowledge to make appropriate changes to the
-model.
+Figure~\ref{fig:intro} highlights some of these issues. When constructing DAGs
+manually, it is important to test how consistent our DAG is to our dataset. One
+of the methods for checking the consistency is by testing whether the implied
+CIs of the DAG hold in the dataset or not \citep{Ankan2021}. As each pair of
+variables that do not have an edge between them in the DAG implies a CI
+statement, we can use a statistical test to check whether that statement holds
+in our dataset or not. If we find any violations we can also modify our DAG
+based on this and utilizing our domain knowledge to find a suitable orientation
+for the edge between the pair of variables.
+
+% While DAG-based methods focus on
+% automated discovery, SEM-based methods emphasize expert driven model
+% specification. This includes tools to assist researchers in manually
+% constructing models, enabling them to incorporate their domain knowledge in the
+% model building process. Researchers typically begin with an initial model based
+% on their domain knowledge and then use these tools to guide modifications that
+% improve the model's fit to data. This process is commonly known as
+% Specification Search \citep{Long1983} and uses method such as modification
+% indices, and Wald-based tests \citep{Marcoulides2018}.
 

### 2025-02-09 paper/main.tex
Updates

@@
     \caption{A comparison of Markov Equivalence Class (MEC) learned using Adult
-	    Income Dataset \citep{Becker1996} using different algorithms and
-	    sample sizes. Edge colour represents the sample size used: Red
-	    ($N=400$), and blue ($N=800$). (a) PC algorithm with a mutual
-	    information based CI test, (b) PC algorithm with a residualization
-	    based test \citep{Ankan2023}, (c) Hill Climb Search with Bayesian
-	    Information Criterion (BIC) score. The learned model structure
-	    varies significantly in each case.}
+	     Income Dataset \citep{Becker1996} using different causal discovery
+	     algorithms and sample sizes. Edge colour represents the sample size
+	     used: red ($N=400$), and blue ($N=800$). (a) PC algorithm with a
+	     mutual information based CI test, (b) PC algorithm with a
+    	     residualization based test \citep{Ankan2023}, (c) Hill Climb Search with
+	     Bayesian Information Criterion (BIC) score. The learned model structure varies
+             significantly in each case.}
     \label{fig:intro}

### 2025-02-09 paper/main.tex
Updates

@@
 
-Figure~\ref{fig:intro} highlights some of these issues. When constructing DAGs
-manually, it is important to test how consistent our DAG is to our dataset. One
-of the methods for checking the consistency is by testing whether the implied
-CIs of the DAG hold in the dataset or not \citep{Ankan2021}. As each pair of
-variables that do not have an edge between them in the DAG implies a CI
-statement, we can use a statistical test to check whether that statement holds
+Figure~\ref{fig:intro} highlights some of these issues. Constructing DAG
+manually, on the other hand, can also be error prone as it can be difficult to
+distinguish various scenarios such as direct or indirect effects, mediation,
+and so on. This makes it important to test the consistency of our DAG against
+our dataset. One of the methods for testing this is to check whether the
+implied CIs of the DAG hold in the dataset or not \citep{Ankan2021}. Each pair
+of variables with a missing edge between them in the DAG implies a CI
+statement, we can use a statistical test to check whether the statement holds
 in our dataset or not. If we find any violations we can also modify our DAG

### 2025-02-09 paper/main.tex
UPdates

--- 2025-02-09 paper/main.tex
@@
 
-Figure~\ref{fig:intro} highlights some of these issues. When constructing DAGs
-manually, it is important to test how consistent our DAG is to our dataset. One
-of the methods for checking the consistency is by testing whether the implied
-CIs of the DAG hold in the dataset or not \citep{Ankan2021}. As each pair of
-variables that do not have an edge between them in the DAG implies a CI
-statement, we can use a statistical test to check whether that statement holds
-in our dataset or not. If we find any violations we can also modify our DAG
-based on this and utilizing our domain knowledge to find a suitable orientation
-for the edge between the pair of variables.
+Figure~\ref{fig:intro} highlights some of these issues. On the other hand,
+constructing DAG based solely on domain expertise can be error prone as it
+requires us to distinguish between different causal structures, such as direct
+or indirect effects or causal mediation. Given these complexities, it is
+important to validate the consistency of a DAG against our dataset. One way to
+test this consistency is by testing whether the CI statements implied by the
+DAG hold in the data \citep{Ankan2021}. Specifically, each missing edge between
+a pair of variables in the DAG implies a CI statement, which can be tested in
+the dataset using a statistical test. If we find any violation to a CI
+statement, we can fix it by adding an edge between the variables in the DAG. We
+also need to decide the orientation of this edge based on domain expertise.
 
@@
 
-In this paper we present a method that utilizes this idea of CI testing to
-assist researchers during the DAG construction instead of testing the final
-model. We propose an iterative model construction method that combines domain
-knowledge with data-driven insights. Our approach utilizes conditional
-independence testing to identify pair of variables in the DAG that are missing
-an edge or have a superfluous edge. Based on this information, in each
-iteration we present one such pair of variables to an expert who then decides
-the orientation of the edge between the variables based on their domain
-knowledge. The method also iteratively prunes all superfluous edges in each
-iteration. As a result this approach guides the expert in constructing their
-model while ensuring that the model is consistent with the data.
-
-As at any iteration, multiple changes to the DAG are possible. We also give a
-way to rank potential new edges additions to the DAG using effect size measures
-for CI tests.
+In this paper, we propose a method that leverages this CI testing approach to
+assist researchers during the DAG construction process, rather than just
+validating the model. Our approach follows an iterative model-building strategy
+that integrates domain expertise with CI testing on data. At each iterations,
+we use CI testing to identify pairs of variables where an edge is either
+missing or superfluous. These pairs are then presented to the expert, who
+determines whether to add or remove these edges. The orientation of the edge
+when adding is determined based on the domain expertise. As at any iteration,
+multiple changes to the DAG are possible. We also give a way to rank potential
+new edges additions to the DAG using effect size measures for CI tests to
+address the most severe violations of CI tests first. As a result this approach
+acts like a guide to the expert when manually constructing their model while
+ensuring that the model is consistent with the data.
 

### 2025-02-09 paper/main.tex
Updates

@@
 
-% Approaches to causal discovery
-% have been developed within both the Directed Acyclic Graphs (DAGs) and
-% Structural Equation Models (SEMs) frameworks, each adapting a different
-% approach.
-
 In the DAG literature, the primary focus has been on developing automated
@@
 
+\begin{figure}[t!]
+    \begin{subfigure}{0.5 \textwidth}
+	\centering
+    	\includegraphics[page=1]{figures_v2.pdf}
+    	\caption{}
+    \end{subfigure}
+    \begin{subfigure}{0.5\textwidth}
+	\centering
+    	\includegraphics[page=2]{figures_v2.pdf}
+    	\caption{}
+    \end{subfigure}
+    \begin{subfigure}{0.5\textwidth}
+	\centering
+    	\includegraphics[page=3]{figures_v2.pdf}
+    	\caption{}
+    \end{subfigure}
+
+    \caption{A comparison of Markov Equivalence Class (MEC) learned using Adult
+	     Income Dataset \citep{Becker1996} using different causal discovery
+	     algorithms and sample sizes. Edge colour represents the sample size
+	     used: red ($N=400$), and blue ($N=800$). (a) PC algorithm with a
+	     mutual information based CI test, (b) PC algorithm with a
+    	     residualization based test \citep{Ankan2023}, (c) Hill Climb Search with
+	     Bayesian Information Criterion (BIC) score. The learned model structure varies
+             significantly in each case.}
+    \label{fig:intro}
+\end{figure}
+
 In this paper, we propose a method that leverages this CI testing approach to
@@
 
-\begin{figure}[t!]
-    \begin{subfigure}{0.5 \textwidth}
-	\centering
-    	\includegraphics[page=1]{figures_v2.pdf}
-    	\caption{}
-    \end{subfigure}
-    \begin{subfigure}{0.5\textwidth}
-	\centering
-    	\includegraphics[page=2]{figures_v2.pdf}
-    	\caption{}
-    \end{subfigure}
-    \begin{subfigure}{0.5\textwidth}
-	\centering
-    	\includegraphics[page=3]{figures_v2.pdf}
-    	\caption{}
-    \end{subfigure}
-
-    \caption{A comparison of Markov Equivalence Class (MEC) learned using Adult
-	     Income Dataset \citep{Becker1996} using different causal discovery
-	     algorithms and sample sizes. Edge colour represents the sample size
-	     used: red ($N=400$), and blue ($N=800$). (a) PC algorithm with a
-	     mutual information based CI test, (b) PC algorithm with a
-    	     residualization based test \citep{Ankan2023}, (c) Hill Climb Search with
-	     Bayesian Information Criterion (BIC) score. The learned model structure varies
-             significantly in each case.}
-    \label{fig:intro}
-\end{figure}
-
 Our main contributions in this paper are as follows:
@@
 \begin{enumerate}
-%     \item We propose a novel measure of conditional association for mixed data
-% 	    based on canonical correlations
-% 	    (Section~\ref{sec:mixed_association}). This measure generalizes
-% 	    several commonly used special case measures of association to mixed
-% 	    data.
     \item We develop an iterative method based on CI testing to assist researchers in 
@@
 \end{enumerate}
-
-% The rest of the paper is structured as follows. In
-% Section~\ref{sec:background}, we give a background on the commonly used
-% measures of association for various data types. In
-% Section~\ref{sec:mixed_association}, we present our generalized measure of
-% conditional association for mixed data. Section~\ref{sec:modification} provides
-% details on the procedure for using the measure of association for constructing
-% DAGs. Section~\ref{sec:web} presents our web-browser based tool and lastly in
-% Section~\ref{sec:empirical}, we show some empirical results to compare this
-% manual DAG construction approach to automated algorithms.
 
@@
 \subsection{Measures of Conditional Association}
-Our approach for ranking CI tests is based on a measure of conditional
-association -- also known as partial association -- between variables.
-Specifically, we are interested in quantifying the association between
-variables $ X $ and $ Y $ when conditioned on a set of variables $ \bm{Z} $
-(which may be empty, i.e., $ \bm{Z} = \emptyset $). Various measures of
-conditional association have been used for this purpose depending on the type
-of $ X $, $ Y $, and $ \bm{Z} $. These measures are often based on the effect
-size of the CI test $ X \ci Y \rvert \bm{Z} $. In this section, we give an
-overview of some of the commonly used measures for different data types.
+Our DAG structure learning approach relies on using CI tests and a measure of
+conditional association between variables -- also known as partial association
+-- between variables. Specifically, firstly we are interested in testing a CI
+test: $ X \ci Y \bm{Z} $, i.e., whether the variables $ X $ and $ Y $ are
+independent when conditioned on $ \bm{Z} $. Secondly, we are interested in
+quantifying the strength of relationship between $ X $ and $ Y $ conditioned on
+$ \bm{Z} $. We can quantify this using the effect size measures of the CI
+tests. In this section, we give an overview of some of the commonly used CI
+tests for various data types and their effect size measures.
 

### 2025-02-09 paper/main.tex
improvements in theorems

--- 2025-02-09 paper/main.tex
@@
 $\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_G, \emptyset, \infty)=G^+$.
+\label{prop:strongexpand}
 \end{proposition}
@@
 Since we start from an empty graph, every pair of vertices $X, Y$ where
- $X \in \tetxrm{An}_G(Y)$ is d-separated by the empty set and is
+ $X \in \textrm{An}_G(Y)$ is d-separated by the empty set and is
  connected by the edge $X \to Y$. No other edges are added. Therefore, 
@@
 \supseteq G^+$.
+\label{prop:weakexpand}
 \end{proposition}
@@
 Every pair of vertices $X, Y$ that are $d$-connected in $G$ are now connected by an edge 
-during $\textsc{Expand}$. For $X \in \tetxrm{An}_G(Y)$, the edge is 
+during $\textsc{Expand}$. For $X \in \textrm{An}_G(Y)$, the edge is 
 $X \to Y$ just like when using a strong ancestral oracle; since only correctly oriented
@@
 
-By combining the previous three propositions, we immediately obtain the following:
+\begin{proof}
+For an edge $X \to Y$ in the skeleton of $G$, $\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$ regardless 
+of $\mathbf{Z}$, so all edges in the skeleton are retained after {\sc Prune}. Conversely, if 
+ $X \to Y$ is not in the skeleton of $G$, let $\mathbf{Z}$ be the $d$-separating set chosen 
+in line~5 of {\sc Prune} (there is at least one such $\mathbf{Z}$, the union of the parents of 
+$X$ and $Y$). Since  $\mathbf{Z}$ $d$-separates all paths from $X$ to $Y$ in $G'$, it does the same 
+in $G$ which contains a subset of these paths. Therefore, all edges not in the skeleton of $G$ are 
+removed.
+\end{proof}
+
+By combining Propositions~\ref{prop:strongexpand}, \ref{prop:weakexpand} and 
+\ref{prop:prune}, we immediately obtain the following:
 
@@
 ancestral oracle $\mathcal{A}_G$, 
-$\textsc{Prune}(\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_G),\mathcal{D}_G)=G.$
+$\textsc{Prune}(\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_G,\emptyset,\infty),\mathcal{D}_G)=G.$
 \end{theorem}
@@
 The loop in Algorithm~\ref{algo:expert} terminates if and only if $E=E^*$. If the loop does not 
-terminate, then a new edge has been added in line 6, or one or more edges were pruned in line 7.
+terminate, a new edge has been added in line 6, and/or one or more edges were pruned in line 7; the pruned 
+edges never include the edge added in line 6 in the same iteration.
 Every edge can be added at 
@@
 after at most $|V|(|V|-1)+1$ iterations of the loop.
-For every edge $e=X\to Y $ in the skeleton of $G$, $\mathcal{D}_G(X,Y,\mathbf{Z})=0$ irrespective of $\mathbf{Z}$.
-Therefore, $e$ must be added to $E$ in some iteration, and can never be pruned afterwards.
+For every edge $e=X\to Y $ in the skeleton of $G$, $\mathcal{D}_G(X,Y,\mathbf{Z})=0$ irrespective of $\mathbf{Z}$, and $e$ must be added to $E$ in some iteration, and can never be pruned afterwards.
 Therefore, after some iteration, $(V,E)$ must be a supergraph of $G$ after executing line 6, and this will be pruned to the real graph $G$ in line 7 (Proposition~\ref{prop:prune}).

### 2025-02-09 paper/main.tex
updates

--- 2025-02-09 paper/main.tex
@@
     \item We develop an iterative method based on CI testing to assist researchers in 
-	    constructing DAGs using their domain knowledge (Section~\ref{sec:modification}).
-    \item We develop a ranking method to prioritize potential modifications in the DAG that 
-    	    can help in improving the model's fit quickly (Section~\ref{sec:ranking}).
+	    constructing DAGs using their domain expertise (Section~\ref{sec:modification}).
+    \item We show a ranking method to prioritize potential modifications to the DAG that 
+    	    help in fixing the most severe violations first (Section~\ref{sec:ranking}).
     \item We provide a web-based interactive tool and a Python package to allow 
-	    researchers to easily apply this method to their datasets (Section~\ref{sec:web}).
+	    researchers to easily apply this method to their own datasets (Section~\ref{sec:web}).
     \item Lastly, we compare the manual DAG construction method with automated

### 2025-02-09 paper/main.tex
Udpates

@@
 
-\todo[inline]{Add equation for data generation method once the plot is finalized}
+\begin{equation}
+	Y = \bm{\beta} \mathrm{Pa}_G(Y) + \mathcal{N}(0, 1)
+\end{equation}
 

### 2025-02-09 paper/main.tex
Updates

@@
 	Markov Equivalence Classes (MECs) rather than a DAG. To assist
-	researchers in constructing DAGs manually, we propose an iterative DAG
-	construction approach that combines domain knowledge with data-driven
+	researchers in constructing DAGs manually, we propose an iterative 
+	structure learning approach that combines domain knowledge with data-driven
 	insights. Our method leverages Conditional Independence (CI) testing to
@@
 
-In this paper, we propose a method that leverages this CI testing approach to
-assist researchers during the DAG construction process, rather than just
-validating the model. Our approach follows an iterative model-building strategy
-that integrates domain expertise with CI testing on data. At each iterations,
-we use CI testing to identify pairs of variables where an edge is either
-missing or superfluous. These pairs are then presented to the expert, who
-determines whether to add or remove these edges. The orientation of the edge
-when adding is determined based on the domain expertise. As at any iteration,
-multiple changes to the DAG are possible. We also give a way to rank potential
-new edges additions to the DAG using effect size measures for CI tests to
-address the most severe violations of CI tests first. As a result this approach
-acts like a guide to the expert when manually constructing their model while
-ensuring that the model is consistent with the data.
+In this paper, we propose a structure learning method that leverages this CI
+testing approach to assist researchers during the DAG construction process,
+rather than just validating the model. Our approach follows an iterative
+model-building strategy that integrates domain expertise with CI testing on
+data. At each iterations, we use CI testing to identify pairs of variables
+where an edge is either missing or superfluous. These pairs are then presented
+to the expert, who determines whether to add or remove these edges. The
+orientation of the edge when adding is determined based on the domain
+expertise. As at any iteration, multiple changes to the DAG are possible. We
+also give a way to rank potential new edges additions to the DAG using effect
+size measures for CI tests to address the most severe violations of CI tests
+first. As a result this approach acts like a guide to the expert when manually
+constructing their model while ensuring that the model is consistent with the
+data.
 
@@
 
-\subsection{Integrating domain knowledge in causal discovery}
-Multiple ways to integrate, originally using forbidden and required edges. Or
-temporal algorithms by specifying causal tiers.
-
-\subsection{Measures of Conditional Association}
+\subsection{CI Tests and Measures of Conditional Association}
 Our DAG structure learning approach relies on using CI tests and a measure of
@@
 \end{equation}
+
+\subsection{Improving DAGs using CI Testing}
 
@@
 \end{figure}
-
-% \begin{figure}
-% 	\centering
-% 	\begin{subfigure}{0.5\textwidth}
-% 		\includegraphics[scale=0.25]{../../presentations/2024_05_das/2.png}
-% 	\end{subfigure}%
-% 	\begin{subfigure}{0.5\textwidth}
-% 		\includegraphics[scale=0.25]{../../presentations/2024_05_das/5.png}
-% 	\end{subfigure}
-% 	\caption{Screenshots of the web-tool. \todo[inline]{Insert screenshots of the web-tool}}
-% \end{figure}
 

### 2025-02-10 paper/main.tex
Updates

@@
 
-In Algorithm~\ref{algo:expert}, in each iteration we add a new edge to the DAG
-chosen by Algorithm~\ref{algo:expand}. This edge is chosen at random depending
-on the order of iteration in Algoirthm~\ref{algo:expand}. A more optimal way to
-add these edges would be to choose the pairs of variables that could help
-improve the fit the most in each step. One potential greedy approach to do this
-would be to select the variable pair $ X $ and $ Y $ that have the highest
-amount of observed conditional correlation in the data. In
-Algorithm~\ref{algo:ranked_expand}, we slightly modify
-Algorithm~\ref{algo:expand} to show this approach. In this case we first go
-through all potential edges that can be added to the DAG in each step, and rank
-them based on the conditional association between them given the current DAG.
-This gives us a measure of how much of the observed correlation between them is
-not explained by the current model. We then choose the pair of variables the
-highest conditional association. This modification allows us to prioritize pair
-of variables to quickly improve the fit of the model. For the conditional
-measure of association, we use a measures outlined in
-Section~\ref{sec:background} depending on the data type of the variables.
+The \textsc{ExpertInLoop} algorithm adds a new edge in each iteration. The edge
+to be added is selected using the \textsc{Expand} algorithm, which returns a
+potential edge at random based on the iteration order in the algorithm.
+However, if the residual association between the selected variables is low,
+adding the edge may result in only a marginal improvement in the model's
+overall fit. As the algorithm requires manual intervention in each iteration in
+terms of domain knowledge, it would be beneficial to prioritize the edges that
+improve the model's fit the most.
+
+Therefore, an alternative to selecting an edge at random is to rank potential
+edges based on the residual association between the variables given the DAG.
+For quantifying the residual association between variables, we can use the
+effect size measures of the CI test. So, for a d-separation query $ D_G(X, Y,
+\mathbf{Z}) $, we can quantify the residual association between $ X $ and $ Y $
+conditioned on $ \mathbf{Z} $ as the effect size of the CI test $ X \ci Y |
+\mathbf{Z} $. A greedy approach could be to choose the edge that has the
+highest residual association between the variables.
+Algorithm~\ref{algo:ranked_expand} outlines this approach. This approach has
+the benefit that it prioritizes edges that improve the model fit the most.
 

### 2025-02-10 paper/main.tex
Updates

@@
 			$ \phi_{\max} \gets \lvert \phi(X, Y, \mathbf{Z}) \rvert $ \;
-			$ L_{\max} \gets \mathcal{A}_G(X, Y)$ \;
+			$ e_{\max} \gets \mathcal{A}_G(X, Y)$ \;
 		}
@@
     }
-    add $L_{\max}$ to $E$\;
+
+    \uIf{ $(V,E \cup e_{\max})$ is acyclic }{
+	add $ e_{\max}$ to $E$ \;
+    } \Else {
+	add $ \textrm{reverse}(e_{\max})$ to $E$ \;
+    }
     \Return{$(V,E)$}\;
@@
 \mathbf{Z}) $, we can quantify the residual association between $ X $ and $ Y $
-conditioned on $ \mathbf{Z} $ as the effect size of the CI test $ X \ci Y |
-\mathbf{Z} $. A greedy approach could be to choose the edge that has the
-highest residual association between the variables.
+conditioned on $ \mathbf{Z} $, $ \phi(X, Y, \bm{Z}) $ as the effect size of the
+CI test $ X \ci Y | \mathbf{Z} $. A greedy approach could be to choose the edge
+that has the highest absolute residual association between the variables.
 Algorithm~\ref{algo:ranked_expand} outlines this approach. This approach has
-the benefit that it prioritizes edges that improve the model fit the most.
+the benefit that it prioritizes edges that improve the model fit the most. We use
+the \textsc{RankedExpand} in all of our experiments in the next section.
 

### 2025-02-10 paper/main.tex
Updates

@@
 	on the Adult income dataset. For the expert simulation an LLM is used
-	for the edge orientations. (a) Shows using total unexplained effect that is the
+	for the edge orientations. (a) Shows using total residual association that is the
 	sum of effect size between variables that don't have an edge between them. (b)
@@
 Similarly, a greedy version of our approach would involve adding an edge
-between the pair of variables with the highest unexplained correlation. While
+between the pair of variables with the highest residual association. While
 both methods aim to find modifications that most improve the model at each
@@
 
-Given a DAG $ G = (V, E) $, the total unexplained correlation, $\tau$ in define as:
+Given a DAG $ G = (V, E) $, the total residual association, $\tau$ in define as:
 \begin{equation}
-	\tau = \sum_{\{ (X, Y) \in V | X \rightarrow Y \not \in E \wedge Y \rightarrow X \not \in E \} } \phi(X, Y, \mathrm{pa}_G(X) \cup \mathrm{pa}_G(Y)))
+	\tau = \sum_{\substack{X, Y \in V \\ X \rightarrow Y, Y \rightarrow X \not \in E}}   \phi(X, Y, \mathrm{pa}_G(X) \cup \mathrm{pa}_G(Y))
 \end{equation}

### 2025-02-10 paper/main.tex
updates
Updates

@@
 
-    \caption{A comparison of Markov Equivalence Class (MEC) learned using Adult
+    \caption{A comparison of Markov Equivalence Classes (MECs) learned from the Adult
 	     Income Dataset \citep{Becker1996} using different causal discovery
-	     algorithms and sample sizes. Edge colour represents the sample size
-	     used: red ($N=400$), and blue ($N=800$). (a) PC algorithm with a
+	     algorithms and sample sizes. Edge colors represents the sample size
+	     used: red for $N=400$, and blue for $N=800$. (a) PC algorithm with a
 	     mutual information based CI test, (b) PC algorithm with a
@@
 	     Bayesian Information Criterion (BIC) score. The learned model structure varies
-             significantly in each case.}
+             significantly across different algorithms and sample sizes.}
     \label{fig:intro}
@@
 with nodes corresponding to the dataset\'s variables. They can then specify a
-p-value threshold and a measure of association threshold. We use the measure of 
-conditional association shown in Appendix~\ref{sec:mixed_association}.
+p-value threshold and a measure of association threshold. To make the method
+work for mixed data types, we use a measure based on canonical correlations
+(Appendix~\ref{sec:mixed_association}).
 

### 2025-02-10 paper/main.tex
Udpates

@@
 to be added is selected using the \textsc{Expand} algorithm, which returns a
-potential edge at random based on the iteration order in the algorithm.
-However, if the residual association between the selected variables is low,
-adding the edge may result in only a marginal improvement in the model's
-overall fit. As the algorithm requires manual intervention in each iteration in
-terms of domain knowledge, it would be beneficial to prioritize the edges that
-improve the model's fit the most.
+potential edge at random based on the iteration order. However, if the residual
+association between the selected variables is low, adding the edge may result
+in only a marginal improvement in the overall model fit. Given that the
+algorithm requires expert intervention in each iteration to specify the edge
+orientation, it is beneficial to prioritize edges that contribute the most to
+improving the model.
 
-Therefore, an alternative to selecting an edge at random is to rank potential
-edges based on the residual association between the variables given the DAG.
-For quantifying the residual association between variables, we can use the
-effect size measures of the CI test. So, for a d-separation query $ D_G(X, Y,
-\mathbf{Z}) $, we can quantify the residual association between $ X $ and $ Y $
-conditioned on $ \mathbf{Z} $, $ \phi(X, Y, \bm{Z}) $ as the effect size of the
-CI test $ X \ci Y | \mathbf{Z} $. A greedy approach could be to choose the edge
-that has the highest absolute residual association between the variables.
-Algorithm~\ref{algo:ranked_expand} outlines this approach. This approach has
-the benefit that it prioritizes edges that improve the model fit the most. We use
-the \textsc{RankedExpand} in all of our experiments in the next section.
+To achieve this, instead of selecting edges randomly, we propose ranking
+potential edges based on their residual association given the current DAG. This
+residual association can be quantified using the effect size measures from the
+CI tests used for querying the d-separation in Algorithm~\ref{algo:expand}.
+Specifically, for a d-separation query $ D_G(X, Y, \mathbf{Z}) $, we quantify
+the residual association, $ \phi(X, Y, \bm{Z}) $, as the effect size of the CI
+test $ X \ci Y \rvert \bm{Z} $. In Algorithm~\ref{algo:ranked_expand}, we
+outline a greedy approach to edge selection using the effect size measure. The
+algorithm selects the edge that has the highest absolute residual association,
+resulting in prioritization of edges that contribute the most to improving the
+model. We use this \textsc{RankedExpand} in all of our experiments in the
+following sections.
 
@@
 The \textsc{ExpertInLoop} algorithm requires two oracles, one that can answer
-d-separation queries and an ancestral oracle that can give the orientation or
-non-existence of edges between pairs of variables. For answering the
-d-separation queries, we use a CI test, the partial correlation test with a
-p-value threshold of $ 0.05 $. Secondly, to tell the direction between edges or
-non-existence of edges, we simulate an expert with accuracy $ \alpha $ as
-follows:
+d-separation queries and an ancestral oracle that can give the causal direction
+between variable pairs. For the experiments, we simulate the d-separation
+oracle using a CI test with a p-value threshold of $ 0.05 $. To simulate the
+edge orientation oracle, we simulate a domain expert that can give the correct
+causal relationship with some accuracy $ \alpha $.
 
@@
 		x &= \textrm{rand}([0, 1]) \\
-		E(\alpha) &= \begin{cases} 
-			\mathrm{TrueDir}(X, Y),  & \textrm{if  } x <= \alpha \\
+		\mathrm{Expert}(\alpha) &= \begin{cases} 
+			\mathrm{TrueReln}(X, Y),  & \textrm{if  } x <= \alpha \\
 			\textrm{rand}(X \rightarrow Y, Y \leftarrow X, \textrm{None}) & \textrm{otherwise} \\
@@
 	\begin{split}
-	\mathrm{TrueDir}(X, Y) &= \begin{cases}
+	\mathrm{TrueReln}(X, Y) &= \begin{cases}
 					X \rightarrow Y, & \textrm{if } X \rightarrow Y \in G \\
@@
 
-An important point to note here is that the effective accuracy of $ E $ is
-higher than $ \alpha $ as even when giving a random answer, there is a $ 1 $ in
-$ 3 $ chance that the answer is correct. The effective accuracy is:
+An important point to note here is that the effective accuracy of $
+\mathrm{Expert} $ is higher than $ \alpha $ as even when giving a random
+answer, there is a $ 1 $ in $ 3 $ chance that the answer is correct by chance.
+Hence, for an $ \mathrm{Expert} $ with accuracy $ \alpha $, the effective
+accuracy becomes:
 
@@
 with three other automated algorithms: PC, Hill-Climb Search, and Greedy
-Equivalence Search (GES). For the analysis, we simulated $ 500 $ samples from
-known DAGs and use the algorithms to recover the original DAG. We start by
-generating a random DAG on $10$ nodes and use linear models with random effects
-to generate the data. We vary the edge probability of the DAG to generate DAGs
-with varying density. We repeat the data generation $30$ times for each density
-value for the DAG. We compare the learned DAG to the original DAG using two
-metrics: Structural Hamming Distance (SHD) and Structural Intervention Distance
-(SID) \citep{Peters2015}.
+Equivalence Search (GES). For the analysis, we simulated data from known DAGs
+and use the algorithms to recover the original DAG. We start by generating a
+random DAG on $10$ nodes and use linear models with random effects to simulate
+$ 500 $ samples from it. We simulate each variable $ Y_i $ in the DAG as:
 
 \begin{equation}
-	Y = \bm{\beta} \mathrm{Pa}_G(Y) + \mathcal{N}(0, 1)
+	\begin{split}
+		\beta &= \mathcal{N}(\bm{0}, \bm{I}) \\
+		Y_i &= \bm{\beta} \mathrm{Pa}_G(Y_i) + \mathcal{N}(0, 1)
+	\end{split}
 \end{equation}
 
-As the automated algorithms can only recover the CPDAG from data, and our
-method recovers the DAG. To compare them fairly, we computed the SHD and SID
-for all possible orientations of the CPDAG to show the best and worst case
-scenarios. The results are shown in Figure~\ref{fig:shd_sid}. For SHD, the
-performance of \textsc{ExpertInLoop} is comparable to the automated algorithms
-for $ \alpha = 0.3 (\alpha_{\textrm{eff}} = 0.53 $, and performs better for
-higher $ \alpha $ values. For SID, the \textsc{ExpertInLoop} performs better in
-denser DAGs.
+We perform the experiment for both varying densities of the DAG and accuracy of
+$\mathrm{Expert}$. We also repeat each experiment $ 30 $ times and report the
+mean and standard error of the results.
 
-% \begin{verbatim}
-% 	Input: Dataset $ D $ on variables $ \bm{X} $
-% 	Output: Learned DAG G.
-% 	
-% 	function compute_effects(G):
-% 		pvalues <- dict()
-% 		effects <- dict()
-% 		for X_i, X_j in X:
-% 			if edge between X_i, X_j:
-% 				p_values[(X_i, X_j)] <- pvalue(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
-% 				effects[(X_i, X_j)] <- \phi(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
-% 			else:
-% 				p_values[(X_i, X_j)] <- pvalue(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
-% 				effects[(X_i, X_j)] <- \phi(X_i, X_j, pa_{\bar{G}}(X_i) \cup pa_{\bar{G}}(X_j))
-% 		return p_values, effects
-% 	
-% 	$ G $ <- Empty Graph
-% 	p_values, effects <- compute_effects(G)
-% 	while (p_values > p_thres, 
-% \end{verbatim}
-% 
-% \begin{algorithm}
-% 	\KwData{Data set $ D $ on variables $ \bm{X} $, pthres, effectthres}
-% 	\KwResult{The learned DAG: $ G $}
-% 	\BlankLine
-% 	$ G \leftarrow $ Empty Graph
-% \end{algorithm}
+To compare the learned DAG to the original DAG, we use two metrics: Structural
+Hamming Distance (SHD) and Structural Intervention Distance (SID)
+\citep{Peters2015}. Unlike \textrm{ExpertInLoop}, the automated algorithms can
+only recover the CPDAG, we consider all possible orientations of the CPDAG to
+compute the range of SHD and SID values to show the best and worst case
+results. The results of this analysis are shown in Figure~\ref{fig:shd_sid}.
+For SHD, the performance of \textsc{ExpertInLoop} is comparable to the
+automated algorithms for $ \alpha = 0.3 (\alpha_{\textrm{eff}} = 0.53 $, which
+is equivalent to the expert getting the orientation correct $ 1 $ out of $ 2 $
+case. For SID, the performance is comparable to the automated algorithms for $
+\alpha = 0.5 (\alpha_{\textrm{eff}} = 0.67) $, which is equivalent to the
+expert getting the orientation correct in $ 2 $ out of $ 3 $ cases. For denser
+DAGs, the \textsc{ExpertInLoop} is able to perform better for lower $ \alpha $
+values.
 

### 2025-02-10 paper/main.tex
Updates

@@
 	\end{subfigure}
-	\caption{Plots showing improvements in fit of the model over iterations
-	on the Adult income dataset. For the expert simulation an LLM is used
-	for the edge orientations. (a) Shows using total residual association that is the
-	sum of effect size between variables that don't have an edge between them. (b)
-	Shows the log-likelihood of the overall model.
+	\caption{Model fit improvement over iterations on the Adult Income dataset. In the expert simulation, an 
+		LLM is used to determine the edge orientations. (a) Shows the total residual association, $\tau$. 
+		(b) Shows the log-likelihood of the overall model.
 	}
@@
 	\begin{split}
-		\beta &= \mathcal{N}(\bm{0}, \bm{I}) \\
+		\beta &= \mathrm{Uniform}([-0.6, 0.6]) \\
 		Y_i &= \bm{\beta} \mathrm{Pa}_G(Y_i) + \mathcal{N}(0, 1)
@@
 
-Our procedure closely resembles certain score-based automated causal discovery
+Our procedure shares similarities with score-based automated causal discovery
 methods, such as Greedy Equivalence Search (GES), where the algorithm
-iteratively adds or removes edges that improve the scoring metric the most.
-Similarly, a greedy version of our approach would involve adding an edge
-between the pair of variables with the highest residual association. While
-both methods aim to find modifications that most improve the model at each
-step, a key distinction lies in the nature of the evaluation criteria: unlike
-standard scoring metrics, our measure of association is not decomposable. A key
-property of scoring metrics is that they should be decomposible, i.e., they can
-be expressed as the sum of scores of nodes given its parents. This means adding
-or removing an edge affects only a localized part of the model. In contrast,
-our measure of association is global, meaning that modifying an edge in one
-part of the model can influence association values elsewhere.
+iteratively adds or removes edges that maximize improvements in a specified
+scoring metric. Given a scoring metric, --total residual association, $ \tau $ for 
+a DAG $ G = (V, E) $, defined as:
 
-Given a DAG $ G = (V, E) $, the total residual association, $\tau$ in define as:
 \begin{equation}
@@
 
-Another major difference is in the interpretability of the evaluation metric.
-Most scoring metrics are based on log-likelihood with a penalty for model
-complexity. These score metrics allow for relative comparisons between models
-but do not provide an absolute measure of model fit. That is, they indicate
-which model is better for a given dataset but do not reveal how well the model
-explains the data in an absolute sense. In contrast, summing our measure of
-association across all variables provides a direct fitness measure—this sum
-approaches zero when the model perfectly accounts for all observed correlations
-in the data. This property enables an absolute assessment of model quality,
-rather than just a relative comparison between models. Figure~\ref{fig:unexplained_ll}
+our \textsc{RankedExpand} approach behaves similarly to GES. In each iteration,
+we add the edge between the variable pair that results in the lowest $ \tau $
+and remove edges that have no impact on $ \tau $.
+
+While both methods aim to make iterative modifications to the DAG, a key
+distinction lies in the nature of the evaluation criteria: unlike standard
+scoring metrics, $ \tau $ is not decomposable. Decomposibility is a key
+property of scoring metrics, allowing them to be expressed as the sum of
+individual variables in the model. This property ensures that adding or
+removing an edge only affects a localized part of the model. In contrast, $
+\tau $ can not be decomposed, as modifying an edge in one part of the model can
+affect the residual associations in other parts of the model.
+
+Another key difference lies in the interpretability of the evaluation metric.
+Scoring metrics are based on usually based on the log-likelihood with a penalty
+for model complexity. They can be used to make relative comparisons between
+models but their values do not have any interpretation in an absolute sense.
+That is, they indicate which model is better for a given dataset but do not
+quantify how well the model explains the data in an absolute sense. In
+contrast, $ \tau $ provides an absolute measure of model fit -- its value
+approaches $ 0 $ as the model perfectly explains the observed data.
+Figure~\ref{fig:unexplained_ll} illustrates this distinction by comparing GES
+to \textsc{ExpertInLoop} on the Adult income dataset across iterations of a
+single run of the algorithms.
 

### 2025-02-10 code/fig4/make_plots.R

@@ -23,7 +23,7 @@ p_unexplained_effect <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
 	geom_point(alpha=0.6, show.legend=F) +
 	theme_minimal(base_size=8) +
 	labs(x="Iteration No.") +
-	labs(y="Total Unexplained Effect") + 
+	labs(y="Total Residual Association") + 
 	theme(legend.position='top',
 	      legend.title=element_blank())
 

### 2025-02-10 paper/main.tex
Updates

@@
 \end{figure}
-Recently, there has been a lot of work towards using LLMs for causal discovery,
-with tasks ranging from determining pairwise edge orientation
-\citep{Kiciman2023, Jin2024} to full causal structure learning \citep{Naik2023,
-Vashishtha2023} and counterfactual reasoning tasks\citep{Kiciman2023} (see
-\citet{Liu2024} for an overview).
 
-Since our approach utilizes expert knowledge to determine edge orientations, we
-investigated the potential of using an LLM for this task. We applied our causal
-discovery procedure with using Gemini 1.5 flash model as the expert. Our method
-uses a greedy approach where the pair of variables with the highest unexplained
-correlation is oriented first. The prompt used to determine the orientation is
-provided in Appendix~\ref{section:llms}. Our LLM prompt utilizes variable
-description to determine the edge orientation. We applied this approach to the
-adult income dataset and the output is shown in Fig.~\ref{fig:adult_llm}.
+Recently, there has been significant interest in leveraging Large Language
+Models (LLMs) for causal discovery. These applications range from determining
+pairwise edge orientations \citep{Kiciman2023, Jin2024} to full causal
+structure learning \citep{Naik2023, Vashishtha2023} and counterfactual
+reasoning \citep{Kiciman2023} (see \citet{Liu2024} for a comprehensive
+overview).
 
-Unlike the other approaches to utilize LLM for doing full causal discovery, we 
-used a combination of statistical tests and LLMs. This greatly reduces the
-amount of information that we need from the LLM. Our approach is also able to 
-ask exact questions about the variables unlike some of the other approaches.
+Since our approach relies on expert knowledge to determine edge orientations,
+we explored the potential of using an LLM for this task. We applied our causal
+discovery procedure using the Gemini 1.5 Flash model as the expert. We simulate
+the \emph{weak ancestral oracle} using the LLM by asking it to choose the
+causal direction between a pair of variables. The LLM is provided a description
+of each of the variables for the task. We have provided the prompt used in
+Appendix~\ref{section:llms}. Figure~\ref{fig:adult_llm} shows the DAG learned
+on the Adult Income dataset using the LLM as the expert. Compared to other
+methods that use LLMs for causal discovery, our approach combines a data-driven
+method with LLMs. We believe this should help in reducing the amount of
+mistakes made by the LLM as well as reduce the amount of information needed
+from the LLM.
 

### 2025-02-10 paper/main.tex
Udpates

@@
 	\includegraphics[page=1]{fig5.pdf}
-	\caption{DAG learned from the adult income dataset using an LLM (GEMINI-1.5-flash) as the expert. The p-value threshold used is $ 0.05 $ and the measure of association threshold is $ 0. 1 $}
+	\caption{DAG learned from the Adult Income dataset using an LLM
+		(Gemini 1.5 Flash) as the expert. The p-value threshold used is $ 0.05 $ 
+		and the measure of association threshold is $ 0. 1 $}
 	\label{fig:adult_llm}
@@
 
-To enable users to apply this method to their own datasets, we developed an
-interactive web-based tool (shown in Fig.~\ref{fig:web}) for constructing
-models. Users begin by uploading their dataset, which initializes an empty DAG
-with nodes corresponding to the dataset\'s variables. They can then specify a
-p-value threshold and a measure of association threshold. To make the method
+To enable researchers to easily apply our approach to their own datasets, we
+also developed an interactive web-based tool (shown in Fig.~\ref{fig:web}) for
+constructing models. Users can upload their dataset, which initializes an empty
+DAG with nodes corresponding to the dataset's variables. They can then specify
+a p-value threshold and a measure of association threshold. To make the method
 work for mixed data types, we use a measure based on canonical correlations
@@
 
-Using the specified thresholds, the tool visually highlights unexplained
-correlations by displaying red edges between variables where correlation exists
-in the data but the current DAG is not able to explain. The thickness of these
-edges represents the strength of the correlation, helping users prioritize
-which edges to add. Similarly, if an edge in the graph is detected to be
-unnecessary, it is highlighted in black. Based on this information, users can
-select to remove unnecessary edge or select a potential edge to add and specify
-the orientation of the edge. The tool computes Shipley’s C \citep{Shipley2000}
-value at each change to determine the overall fit of the model to the data.
-Once satisfied with the constructed DAG, users can export the model for further
-analysis.
+Using the specified thresholds, the tool visually highlights variable pairs
+(using red edges) with a residual association higher than the threshold. The
+thickness of these edges represents the strength of association, helping users
+prioritize which variable pairs to address first. Similarly, if an edge in the
+graph is detected to be unnecessary, it is highlighted in black. Based on this
+information, make modifications to the model by adding or removing edges. The
+tool also provides the Room Mean Square Error of Approximation (RMSEA) of the
+overall model fit to the data. Once satisfied with the constructed DAG, users
+can export the model for further analysis.
 

### 2025-02-10 paper/main.tex
Updates

@@
 To enable researchers to easily apply our approach to their own datasets, we
-also developed an interactive web-based tool (shown in Fig.~\ref{fig:web}) for
-constructing models. Users can upload their dataset, which initializes an empty
-DAG with nodes corresponding to the dataset's variables. They can then specify
-a p-value threshold and a measure of association threshold. To make the method
-work for mixed data types, we use a measure based on canonical correlations
-(Appendix~\ref{sec:mixed_association}).
+developed an interactive web-based tool (Figure~\ref{fig:web}) for constructing
+DAGs. Users can upload their dataset, which initializes an empty DAG with nodes
+corresponding to the dataset's variables. They can then specify a p-value
+threshold and a measure of association threshold. To accommodate mixed data
+types, we use a measure of association based on canonical correlations
+(detailed in Appendix~\ref{sec:mixed_association}).
 
 Using the specified thresholds, the tool visually highlights variable pairs
-(using red edges) with a residual association higher than the threshold. The
-thickness of these edges represents the strength of association, helping users
-prioritize which variable pairs to address first. Similarly, if an edge in the
-graph is detected to be unnecessary, it is highlighted in black. Based on this
-information, make modifications to the model by adding or removing edges. The
-tool also provides the Room Mean Square Error of Approximation (RMSEA) of the
-overall model fit to the data. Once satisfied with the constructed DAG, users
-can export the model for further analysis.
+with a residual association greater than the threshold, marking them with red
+edges. The thickness of these edges represents the strength of association,
+helping users prioritize which variable pairs to address first. Similarly, if
+an existing edge is superfluous, it is highlighted in black, signaling
+potentially unnecessary edge. Using this information, users can iteratively
+modify the model by adding or removing edges. The tool also computes the Root
+Mean Square Error of Approximation (RMSEA) to provide an estimate of the
+overall fit of the model. Once satisfied with the constructed DAG, users can
+export the model for further analysis.
 

### 2025-02-10 paper/main.tex
Updates

@@
 random DAG on $10$ nodes and use linear models with random effects to simulate
-$ 500 $ samples from it. We simulate each variable $ Y_i $ in the DAG as:
+$ 500 $ samples from it. We simulate each variable $ V_i $ in the DAG as:
 
@@
 		\beta &= \mathrm{Uniform}([-0.6, 0.6]) \\
-		Y_i &= \bm{\beta} \mathrm{Pa}_G(Y_i) + \mathcal{N}(0, 1)
+		V_i &= \bm{\beta} \mathrm{Pa}_G(V_i) + \mathcal{N}(0, 1)
 	\end{split}
@@
 \section{Conclusions}
+Researchers often prefer to construct DAGs manually rather than using automated
+algorithms due to practical challenges in applying automated methods. To assist
+in the process, we developed an iterative structure learning method that
+integrates manual construction with data-driven insights. This approach bridges
+the gap between fully manually and fully automated methods by providing a
+systematic way to construct DAGs while utilizing data.
 
-Points:
-\begin{itemize}
-	\item As researchers prefer to build models by hand, we present a tool
-		to assist them in construction using CI testing.
-	\item In practice as many edges are possible for any given DAG, we provide
-		a ranking method to find the best edge addition.
-	\item Compared to other expert knowledge specification methods, users
-		don't need to guess what information then need to provide as
-		exact questions are asked. Also enables us to ask an LLM for
-		it.
-\end{itemize}
-
-Future work:
-\begin{itemize}
-	\item Add latent variables; extend to ADMGs.
-	\item Can potentially be combined with pairwise orientation methods.
-	\item Improvements in the web-interface.
-\end{itemize}
-
+In future we would like to extend our approach to latent variables and Acyclic
+Directed Mixed Graphs (ADMGs), explore the potential of combining our approach
+with pairwise orientation methods. We would also improve the web interface for
+better usability. 
 \bibliography{references}

### 2025-02-10 paper/main.tex
Udpate
Updates

@@
 \end{equation}
-
-\subsection{Improving DAGs using CI Testing}
 
@@
 \section{Conclusions}
+\todo[inline]{Rewrite}
 Researchers often prefer to construct DAGs manually rather than using automated

### 2025-02-10 paper/main.tex
Updates

@@
 	algorithms, such as their tendency to make obvious errors and output
-	Markov Equivalence Classes (MECs) rather than a DAG. To assist
+	Markov Equivalence Classes (MECs) instead of a DAG. To assist
 	researchers in constructing DAGs manually, we propose an iterative
-	structure learning approach that combines domain knowledge with data-driven
-	insights. Our method leverages Conditional Independence (CI) testing to
-	iteratively identify variable pairs where an edge is either missing or
-	superfluous. Based on this information, we can choose to add missing
-	edges with appropriate orientation based on domain knowledge or remove
-	unnecessary ones. We also give a method to rank potential new edges
-	based on their impact on the overall model fit. Empirical results show
-	that this iterative approach achieves performance comparable to
-	automated algorithms if the orientation of new edges is correctly
-	determined in at least one out of two cases. Additionally, we show
-	that, in the absence of an expert, a Large Language Model (LLM) can be
-	used to determine edge orientations. We provide an implementation of
-	our approach as both a webtool: <redacted for review> and a Python
-	package <redacted for review>.
+	structure learning approach that combines domain knowledge with
+	data-driven insights. Our method leverages Conditional Independence
+	(CI) testing to iteratively identify variable pairs where an edge is
+	either missing or superfluous. Based on this information, we can choose
+	to add missing edges with appropriate orientation based on domain
+	knowledge or remove unnecessary ones. We also give a method to rank
+	these missing edges based on their impact on the overall model fit.
+	Empirical results show that this iterative approach achieves
+	performance comparable to automated algorithms if the orientation of
+	new edges is correctly determined in at least one out of two cases.
+	Additionally, we show that, in the absence of an expert, a Large
+	Language Model (LLM) can be used to determine edge orientations. We
+	provide an implementation of our approach as both a web tool: <redacted
+	for review> and a Python package <redacted for review>.
 
@@
 behind observed phenomena and guide effective interventions or policy
-decisions. Causal discovery methods aim to discover such relationships
-among random variables using observational data.
-
-In the DAG literature, the primary focus has been on developing automated
-algorithms to learn causal structures from datasets. These efforts have led to
-numerous causal discovery algorithms, such as constraint-based methods like PC
-algorithm \citep{Spirtes2001} and Fast Causal Inference \citep{Spirtes2000}),
-score-based methods such as Hill-Climb Search and Greedy Equivalence Search
-\citep{Chickering2002}, and continuous optimization-based methods like NO TEARS
-\citep{Zheng2018} and DAGMA \citep{Bello2022}. Despite significant progress in
-automated causal discovery, their adoption in applied research has been
-limited. Researchers often prefer to manually construct DAGs based on their
-domain expertise \citep{Tennant2020, Petersen2021}. We attribute this
-preference to several challenges with existing causal discovery algorithms in
-practical settings:
+decisions. Causal discovery methods aim to discover such relationships among
+random variables using observational data. In the DAG literature, the primary
+focus has been on developing automated algorithms to learn causal structures
+from datasets. These efforts have led to numerous causal discovery algorithms,
+such as constraint-based methods like PC algorithm \citep{Spirtes2001} and Fast
+Causal Inference \citep{Spirtes2000}), score-based methods such as Hill-Climb
+Search and Greedy Equivalence Search \citep{Chickering2002}, and continuous
+optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
+\citep{Bello2022}. Despite significant progress in automated causal discovery,
+their adoption in applied research has been limited. Researchers often prefer
+to manually construct DAGs based on their domain expertise \citep{Tennant2020,
+Petersen2021}. We attribute this preference to several challenges with existing
+causal discovery algorithms in practical settings:
 
@@
 a pair of variables in the DAG implies a CI statement, which can be tested in
-the dataset using a statistical test. If we find any violation to a CI
-statement, we can fix it by adding an edge between the variables in the DAG. We
-also need to decide the orientation of this edge based on domain expertise.
+the dataset using a statistical test. If we find any violation to CI
+statements, we can make modifications to our model accordingly to fix the
+violations.
 
@@
 In this paper, we propose a structure learning method that leverages this CI
-testing approach to assist researchers during the DAG construction process,
-rather than just validating the model. Our approach follows an iterative
+testing approach to assist researchers during manual DAG construction, rather
+than just validating the model. Our approach follows an iterative
 model-building strategy that integrates domain expertise with CI testing on
-data. At each iterations, we use CI testing to identify pairs of variables
-where an edge is either missing or superfluous. These pairs are then presented
-to the expert, who determines whether to add or remove these edges. The
-orientation of the edge when adding is determined based on the domain
+data. At each iteration, we use CI testing to identify pairs of variables where
+an edge is either missing or superfluous. We can then use this information
+along with domain expertise to determines whether to add or remove these edges.
+The orientation of the edge when adding is determined based on the domain
 expertise. As at any iteration, multiple changes to the DAG are possible. We
@@
 data.
-
-% When manually constructing models, it is important to test whether the
-% resulting DAG accurately represents the dataset. Based on this evaluation and
-% domain knowledge, we can make further modification to the model. One common
-% approach for assessing DAGs is to check whether the Conditional Independences
-% (CIs) implied by the DAG hold in the data \citep{Ankan2021}. If we find
-% violations to these CIs, we can use our domain knowledge to make appropriate
-% modifications to the DAG. However, this CI testing approach has a few
-% limitations:
-% 
-% \begin{enumerate}
-% \item As CIs are only implied by missing edges in the model, this approach does 
-% 	not test whether the existing edges are correct.
-% \item Determining whether a CI holds in data is based on a p-value threshold. This can 
-% 	be unreliable, for example, we get a significant p-value for even very weak 
-% 	relationship if the sample size is high.
-% \item The number of implied CIs can be large making it difficult to manually
-% 	check all of them. For example, a moderately sized alarm network
-% 	\citep{Beinlich1989} with $ 37 $ nodes, implies $ 287 $ CIs.
-% \end{enumerate}
-% 
-% To tackle these limitations, we draw inspiration from modification indices in
-% specification search. Modification indices provide ranking for potential model
-% modification based on their impact on the model's fit, allowing us to
-% prioritize most meaningful modifications.  Similarly, we propose a method for
-% ranking CI test violations to help us focus on the most critical
-% inconsistencies in the model. To determine this ranking we utilize a measure of
-% conditional association between variables in combination with p-values from the
-% CI test. This approach addresses all three of the issues outlined above. The
-% measure of association allows us to evaluate the validity of existing edges. It
-% reduces our reliance on p-values alone for deciding whether the CI test holds.
-% The ranking helps us prioritize the most critical violations, making it feasible 
-% to focus on the areas that most improve the model's fit.
 
@@
 \end{equation}
+
+\subsection{Model Testing Using CI Tests}
 

### 2025-02-10 paper/main.tex
Updates

@@
 \subsection{CI Tests and Measures of Conditional Association}
-Our DAG structure learning approach relies on using CI tests and a measure of
-conditional association between variables -- also known as partial association
--- between variables. Specifically, firstly we are interested in testing a CI
-test: $ X \ci Y \bm{Z} $, i.e., whether the variables $ X $ and $ Y $ are
-independent when conditioned on $ \bm{Z} $. Secondly, we are interested in
-quantifying the strength of relationship between $ X $ and $ Y $ conditioned on
-$ \bm{Z} $. We can quantify this using the effect size measures of the CI
-tests. In this section, we give an overview of some of the commonly used CI
-tests for various data types and their effect size measures.
+Our structure learning approach relies on using CI tests and a measure of
+conditional association between variables -- also known as partial association.
+Specifically, we are interested in testing CI tests: $ X \ci Y \bm{Z} $, i.e.,
+whether the variables $ X $ and $ Y $ are independent when conditioned on $
+\bm{Z} $. Secondly, we are interested in quantifying the strength of
+relationship between $ X $ and $ Y $ conditioned on $ \bm{Z} $. We can use
+the effect size measures of the CI tests for this quantification. In this 
+section, we give an overview of some of the commonly used CI tests for 
+various data types and their effect size measures.
 
@@
 \subsection{Model Testing Using CI Tests}
+Our structure learning algorithm relies on suggesting modifications to the
+expert based on CI testing. In this section, we show an example of how this
+works. Our structure learning algorithm can be thought of an iterative
+application of this model testing approach.
 

### 2025-02-10 paper/main.tex
Udpates

@@
 \subsection{Model Testing Using CI Tests}
-Our structure learning algorithm relies on suggesting modifications to the
-expert based on CI testing. In this section, we show an example of how this
-works. Our structure learning algorithm can be thought of an iterative
-application of this model testing approach.
+Our structure learning algorithm relies on the idea of testing DAGs using CI
+tests. In this section, we give an example of how this model testing and
+corresponding modification work. We take the example of the MEC learned by the
+PC algorithm in Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all
+the undirected edges in the model based on our domain knowledge. In this case
+we only have one undirected edge between Marital Status and Relationship. Let's 
+orient the edge as Marital Status to Relationship.
+
+ The implied CIs of the 
+model along with the test results are shown in Figure.
 
@@
 
-\begin{figure}[t!]
-	\begin{subfigure}{0.25\textwidth}
-		\centering
-		\includegraphics{../code/fig4/unexplained_effect.pdf}
-		\caption{}
-	\end{subfigure}%
-	\begin{subfigure}{0.25\textwidth}
-		\centering
-		\includegraphics{../code/fig4/ll.pdf}
-		\caption{}
-	\end{subfigure}
-	\caption{Model fit improvement over iterations on the Adult Income dataset. In the expert simulation, an 
-		LLM is used to determine the edge orientations. (a) Shows the total residual association, $\tau$. 
-		(b) Shows the log-likelihood of the overall model.
-	}
-	\label{fig:unexplained_ll}
-\end{figure}
-
 The \textsc{ExpertInLoop} algorithm requires two oracles, one that can answer
@@
 \subsection{Connection to Score Based Methods}
+
+\begin{figure}[t!]
+	\begin{subfigure}{0.25\textwidth}
+		\centering
+		\includegraphics{../code/fig4/unexplained_effect.pdf}
+		\caption{}
+	\end{subfigure}%
+	\begin{subfigure}{0.25\textwidth}
+		\centering
+		\includegraphics{../code/fig4/ll.pdf}
+		\caption{}
+	\end{subfigure}
+	\caption{Model fit improvement over iterations on the Adult Income dataset. In the expert simulation, an 
+		LLM is used to determine the edge orientations. (a) Shows the total residual association, $\tau$. 
+		(b) Shows the log-likelihood of the overall model.
+	}
+	\label{fig:unexplained_ll}
+\end{figure}
 

### 2025-02-10 code/fig1/fig2.R
Updates

@@ -1,0 +1,49 @@
+library(dagitty)
+
+d <- read.csv("../utils/adult_unproc.csv")
+
+# Preprocess the data
+d$Fnlwgt <- NULL
+d$EducationNum <- NULL
+d$CapitalGain <- NULL
+d$CaptialLoss <- NULL
+
+# Categorical
+d$Workclass <- factor(d$Workclass, ordered = F)
+d$MaritalStatus <- factor(d$MaritalStatus, ordered = F)
+d$Occupation <- factor(d$Occupation, ordered = F)
+d$Relationship <- factor(d$Relationship, ordered = F)
+d$Race <- factor(d$Race, ordered = F)
+d$Sex <- factor(d$Sex, ordered = F)
+d$NativeCountry <- factor(d$NativeCountry, ordered = F)
+d$Income <- factor(d$Income, ordered = F)
+
+# Continuous
+d$Age <- as.double(d$Age)
+d$HoursPerWeek <- as.double(d$HoursPerWeek)
+
+# Ordinal
+education_levels <- c("Preschool", "1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th", "12th", "HS-grad", "Some-college", "Assoc-voc", "Assoc-acdm", "Bachelors", "Masters", "Prof-school", "Doctorate")
+d$Education <- factor(d$Education, levels = education_levels, ordered = T)
+
+set.seed(42)
+d <- d[complete.cases(d), ]
+d <- d[sample(nrow(d), 1000), ]
+
+dag <- dagitty("dag{MaritalStatus -> Relationship
+	       	    Relationship -> Sex 
+		    HoursPerWeek -> Sex
+		    Relationship -> HoursPerWeek
+		    Relationship -> Age
+		    HoursPerWeek -> Education
+		    Age -> MaritalStatus
+		    Occupation -> Relationship
+		    Occupation -> HoursPerWeek
+		    Occupation -> Sex
+		    Income -> Age
+		    Income -> Occupation
+		    Occupation -> Workclass
+		    Income -> Education
+		    Occupation -> Education}")
+
+print(localTests(x=dag, data=d, type="cis.pillai"))

### 2025-02-10 paper/main.tex
Updates
restructuring

@@
 
-\title{Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge}
+\title{Expert-In-The-Loop Causal Discovery: \\ Iterative Model Refinement Using Expert Knowledge}
 
@@
     	    help in fixing the most severe violations first (Section~\ref{sec:ranking}).
-    \item We provide a web-based interactive tool and a Python package to allow 
-	    researchers to easily apply this method to their own datasets (Section~\ref{sec:web}).
-    \item Lastly, we compare the manual DAG construction method with automated
+    \item We show empirically that even experts that make mistakes can outperform purely data-driven
 	causal discovery algorithms (Section~\ref{sec:empirical}).
+    \item We provide two proof-of-concept implementations of our approach: one geared towards
+	LLMs as experts, and a graphical user interface designed 
+	for human experts (Section~\ref{sec:web}).
 \end{enumerate}
 
-\section{Background and Related Work}
+\section{Background}
 \label{sec:background}
@@
 
-\subsection{CI Tests and Measures of Conditional Association}
+\subsection{Model Testing Using CI Tests}
+Our structure learning algorithm relies on the idea of testing DAGs using CI
+tests. In this section, we give an example of how this model testing and
+corresponding modification work. We take the example of the MEC learned by the
+PC algorithm in Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all
+the undirected edges in the model based on our domain knowledge. In this case
+we only have one undirected edge between Marital Status and Relationship. Let's 
+orient the edge as Marital Status to Relationship.
+
+ The implied CIs of the 
+model along with the test results are shown in the Figure.
+
+We can think of our proposed approach to be an iterative application of this model
+testing approach.
+
+\subsection{Effect Sizes for CI Tests}
+
 Our structure learning approach relies on using CI tests and a measure of
@@
 the effect size measures of the CI tests for this quantification. In this 
-section, we give an overview of some of the commonly used CI tests for 
-various data types and their effect size measures.
+section, we give an overview of some of the effect size measures for 
+different types of (possibly mixed) data.
 
@@
 	r_{X, Y; \bm{Z}} = r_{R_X, R_Y}
-\end{equation}
-
-\paragraph{All $ X $, $ Y $, and $ \bm{Z} $ are discrete: }
-
-When all variables are discrete, Cram\'er's V can used as a measure of
-association. When $ \bm{Z} = \emptyset $, Cram\'er's V is derived from the 
-chi-squared test statistic:
-
-\begin{equation}
-	\mathbf{V}_{X, Y} = \sqrt{\frac{\chi^2 / n}{\mathrm{min}(k-1, r-1)}}
-\end{equation}
-
-where $ \chi^2 $ is the chi-squared test statistic for $ X $ and $ Y $, $ n $
-is the sample size, and $ k $ and $ r $ are the number of categories in $ X $
-and $ Y $ respectively. When $ \bm{Z} \neq \emptyset $, we start by splitting
-the dataset by each unique combination of $ \bm{Z} $ into subsets,  $ \bm{Z} $,
-$ D = \{ D_{\bm{Z} = \bm{Z}_1}, D_{\bm{Z} = \bm{Z}_2}, \cdots, D_{\bm{Z} =
-\bm{Z}_k} \} $. Then we compute the Cram\'er's V for each of these smaller datasets
-and combine them as follows:
-
-\begin{equation}
-	\mathbf{V}_{X, Y; \bm{Z}} = \sum_{i=1}^{k} \left[ \mathbf{V}_{X, Y} \right]_{D_{\bm{Z} = \bm{Z}_i}} 
 \end{equation}
@@
 
-\subsection{Model Testing Using CI Tests}
-Our structure learning algorithm relies on the idea of testing DAGs using CI
-tests. In this section, we give an example of how this model testing and
-corresponding modification work. We take the example of the MEC learned by the
-PC algorithm in Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all
-the undirected edges in the model based on our domain knowledge. In this case
-we only have one undirected edge between Marital Status and Relationship. Let's 
-orient the edge as Marital Status to Relationship.
+\paragraph{$ X $, $ Y $, and $ \bm{Z} $ are discrete (ordinal or caterical): }
 
- The implied CIs of the 
-model along with the test results are shown in Figure.
+\todo[inline]{Here we use an effect size RMSEA based on the previous Hotelling's test paper.}
 
-\section{DAG Structure Learning using Ancestral Oracles}
-\label{sec:modification}
+\section{Expert-In-The-Loop Causal Discovery}
 
@@
 of interest depends on all other variables present in the graph.
+
+We will now first present theoretical results showing how a structure
+learning algorithm using such experts can uncover the true DAG structure.
+Afterwards, we will present a heuristic for deciding which changes should
+be prioritized.
+
+\subsection{DAG Structure Learning using Ancestral Oracles}
+
+\label{sec:modification}
 
@@
 
-\section{Empirical Analysis}
-\label{sec:empirical}
-
-In this section, we give some of the practical aspects of the
-\textsc{ExpertInLoop} algorithm, and then compare the algorithm with automated
-causal discovery algorithms.
-
 \subsection{Ranking Potential New Edges}
@@
 test $ X \ci Y \rvert \bm{Z} $. In Algorithm~\ref{algo:ranked_expand}, we
-outline a greedy approach to edge selection using the effect size measure. The
+outline a greedy approach to edge selection using the effect size measure.
+
+To implement Algorithm~\ref{algo:ranked_expand}, we need to have an effect size 
+metric by which we can rank effects. For some simple cases, such as 
+linear-Gaussian models, such effect sizes exist. However, a challenge arises
+when using mixed data. We here present a standardized effect size hat 
+is useable for datasets containing mixtures of continuous, ordinal and
+
+The
 algorithm selects the edge that has the highest absolute residual association,
 resulting in prioritization of edges that contribute the most to improving the
-model. We use this \textsc{RankedExpand} in all of our experiments in the
-following sections.
+model.
 
-\subsection{Comparison to Automated Algorithms}
+\subsection{Connection to Score-Based Methods}
+
+Our procedure shares similarities with score-based automated causal discovery
+methods, such as Greedy Equivalence Search (GES), where the algorithm
+iteratively adds or removes edges that maximize improvements in a specified
+scoring metric. Given a scoring metric, --total residual association, $ \tau $ for 
+a DAG $ G = (V, E) $, defined as:
+
+\begin{equation}
+	\tau = \sum_{\substack{X, Y \in V \\ X \rightarrow Y, Y \rightarrow X \not \in E}}   \phi(X, Y, \mathrm{pa}_G(X) \cup \mathrm{pa}_G(Y))
+\end{equation}
+
+our \textsc{RankedExpand} approach behaves similarly to GES. In each iteration,
+we add the edge between the variable pair that results in the lowest $ \tau $
+and remove edges that have no impact on $ \tau $.
+
+While both methods aim to make iterative modifications to the DAG, a key
+distinction lies in the nature of the evaluation criteria: unlike standard
+scoring metrics, $ \tau $ is not decomposable. Decomposibility is a key
+property of scoring metrics, allowing them to be expressed as the sum of
+individual variables in the model. This property ensures that adding or
+removing an edge only affects a localized part of the model. In contrast, $
+\tau $ can not be decomposed, as modifying an edge in one part of the model can
+affect the residual associations in other parts of the model.
+
+Another key difference lies in the interpretability of the evaluation metric.
+Scoring metrics are based on usually based on the log-likelihood with a penalty
+for model complexity. They can be used to make relative comparisons between
+models but their values do not have any interpretation in an absolute sense.
+That is, they indicate which model is better for a given dataset but do not
+quantify how well the model explains the data in an absolute sense. In
+contrast, $ \tau $ provides an absolute measure of model fit -- its value
+approaches $ 0 $ as the model perfectly explains the observed data.
+
+\section{Empirical Analysis}
+\label{sec:empirical}
+
+In this section, we give some of the practical aspects of the
+\textsc{ExpertInLoop} algorithm, and then compare the algorithm with automated
+causal discovery algorithms.
+
+In all experiments, we use the \textsc{RankedExpand} method together
+with a version of the strong ancestral oracle that can make mistakes.
+
+\todo{Want to know: what if my experts are not perfect?}
+
 \begin{figure}[t!]
@@
 
-\subsection{Connection to Score Based Methods}
-
 \begin{figure}[t!]
@@
 
-Our procedure shares similarities with score-based automated causal discovery
-methods, such as Greedy Equivalence Search (GES), where the algorithm
-iteratively adds or removes edges that maximize improvements in a specified
-scoring metric. Given a scoring metric, --total residual association, $ \tau $ for 
-a DAG $ G = (V, E) $, defined as:
+\section{Practical Implementation}
 
-\begin{equation}
-	\tau = \sum_{\substack{X, Y \in V \\ X \rightarrow Y, Y \rightarrow X \not \in E}}   \phi(X, Y, \mathrm{pa}_G(X) \cup \mathrm{pa}_G(Y))
-\end{equation}
+\label{sec:web}
 
-our \textsc{RankedExpand} approach behaves similarly to GES. In each iteration,
-we add the edge between the variable pair that results in the lowest $ \tau $
-and remove edges that have no impact on $ \tau $.
-
-While both methods aim to make iterative modifications to the DAG, a key
-distinction lies in the nature of the evaluation criteria: unlike standard
-scoring metrics, $ \tau $ is not decomposable. Decomposibility is a key
-property of scoring metrics, allowing them to be expressed as the sum of
-individual variables in the model. This property ensures that adding or
-removing an edge only affects a localized part of the model. In contrast, $
-\tau $ can not be decomposed, as modifying an edge in one part of the model can
-affect the residual associations in other parts of the model.
-
-Another key difference lies in the interpretability of the evaluation metric.
-Scoring metrics are based on usually based on the log-likelihood with a penalty
-for model complexity. They can be used to make relative comparisons between
-models but their values do not have any interpretation in an absolute sense.
-That is, they indicate which model is better for a given dataset but do not
-quantify how well the model explains the data in an absolute sense. In
-contrast, $ \tau $ provides an absolute measure of model fit -- its value
-approaches $ 0 $ as the model perfectly explains the observed data.
-Figure~\ref{fig:unexplained_ll} illustrates this distinction by comparing GES
-to \textsc{ExpertInLoop} on the Adult income dataset across iterations of a
-single run of the algorithms.
+We now present two proof-of-concept implementations of our approach, 
+geared towards using  large language models (LLMs) and humans as experts
+to decide on edge directions.
 
@@
 of each of the variables for the task. We have provided the prompt used in
-Appendix~\ref{section:llms}. Figure~\ref{fig:adult_llm} shows the DAG learned
+Appendix~\ref{section:llms}. 
+Figure~\ref{fig:unexplained_ll} illustrates this distinction by comparing GES
+to \textsc{ExpertInLoop} on the Adult income dataset across iterations of a
+single run of the algorithms.
+Figure~\ref{fig:adult_llm} shows the DAG learned
 on the Adult Income dataset using the LLM as the expert. Compared to other
@@
 
-\section{Web Tool}
-\label{sec:web}
+\subsection{Using Humans as Experts}
 

### 2025-02-10 code/fig4/make_plots.R
Updates likelihood plot

@@ -16,7 +16,8 @@ rownames(d_ges) <- seq(1, nrow(d_ges))
 d_ges <- cbind(d_ges, x=seq(1, nrow(d_ges)))
 d_ges$algo <- 'GES'
 
-d_results = rbind(d_expert, d_ges)
+d_results <- rbind(d_expert, d_ges)
+d_results <- d_results[d_results$x < 31, ]
 
 p_unexplained_effect <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
 	geom_line(alpha=0.6) +

### 2025-02-10 paper/main.tex
abstract edits done

@@
 \begin{abstract}
-
-	Causal discovery has received significant attention in the Directed
-	Acyclic Graphs (DAGs) literature, leading to the development of
-	numerous automated algorithms for learning DAGs from data. However,
-	their adoption in applied domains remain limited, as researchers often
+	Numerous causal discovery algorithms were developed to automatically learn
+	directed acyclic graphs (DAGs) and other causal models from data. However,
+	their adoption in applied domains remains limited, as researchers often
 	prefer to construct DAGs manually based on domain knowledge. This
 	preference arises due to several practical challenges with automated
-	algorithms, such as their tendency to make obvious errors and output
-	Markov Equivalence Classes (MECs) instead of a DAG. To assist
-	researchers in constructing DAGs manually, we propose an iterative
+	algorithms, such as their tendency to produce results that contradict
+	obvious domain knowledge and their inability to distinguish Markov equivalent
+	models. To assist researchers in constructing DAGs manually, we propose an iterative
 	structure learning approach that combines domain knowledge with
-	data-driven insights. Our method leverages Conditional Independence
-	(CI) testing to iteratively identify variable pairs where an edge is
+	data-driven insights. Our method leverages conditional independence
+	testing to iteratively identify variable pairs where an edge is
 	either missing or superfluous. Based on this information, we can choose
@@
 	these missing edges based on their impact on the overall model fit.
-	Empirical results show that this iterative approach achieves
-	performance comparable to automated algorithms if the orientation of
-	new edges is correctly determined in at least one out of two cases.
-	Additionally, we show that, in the absence of an expert, a Large
-	Language Model (LLM) can be used to determine edge orientations. We
-	provide an implementation of our approach as both a web tool: <redacted
-	for review> and a Python package <redacted for review>.
-
+	In a simulation study, we find that this iterative approach to leverage domain 
+	knowledge already starts outperforming purely data-driven structure learning if 
+	the orientation of new edge is correctly determined in at least one out of two cases.
+	We present a proof-of-concept implementation using a large language 
+	model as a domain expert and a graphical user interface designed to 
+	assist human experts with DAG construction.
 \end{abstract}

### 2025-02-10 paper/main.tex
Adds notation

--- 2025-02-10 paper/main.tex
@@
 \label{sec:background}
-\todo[inline]{Need to rewrite this}
-We denote random variables with uppercase letters $ X $, and a set of random
-variables as $ \bm{X} = \{X_1, \cdots, X_k\} $ with $ \rvert \bm{X} \rvert = k
-$. A sample from the random variable $ X $ is denoted as $ x $ and from a set
-of random variables $ \bm{X} $ is denoted as $ \bm{x} $. In this paper, we
-consider random variables in the mixed data setting where each of the random
-variables can be either of continuous, categorical, or ordinal unless
-specified. We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $,
-conditional expectations as $ \mathbb{E}[X | \bm{Z}] $. We denote a graph $ G =
-(V, E) $ which nodes $ V$ and edges $ E $. In this paper we focus on all
-observed variables, and linear measure of association. 
-where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, $ \mathrm{corr}(X, Y) $ is the correlation between $ X $ and $ Y $, and $
-\sigma_X $ and $ \sigma_Y $ are the standard deviations of $ X $ and $ Y $
-respectively.  Covariance matrix using $ \Sigma $ and the entry corresponding to
-covariance between $ X $ and $ Y $ in the matrix using $ \Sigma_{XY} $.
+We denote random variables using uppercase letters, such as $ X $, and a set of
+random variables by $ \bm{X} = \{X_1, \cdots, X_k\} $, where $ \rvert \bm{X}
+\rvert = k $ indicates the number of variables in the set. We denote the
+covariance between variables $ X $ and $ Y $ as $ \mathrm{cov}(X, Y) $, the
+covariance matrix as $ \Sigma $, and the entry corresponding to covariance
+between $ X $ and $ Y $ using $ \Sigma_{XY} $. In the context of this paper, we
+consider causal DAGs that contain only observed variables. The structure of the
+DAG is represented using a graph $ G = (V, E) $, where $ V $ are the nodes
+corresponding to the random variables, and $ E $ are the edges denoting the
+causal relationship between the variables. The set of parents of $ X $ in $ G $
+are denoted as $ \textrm{Pa}_G(X) $, and its ancestors and descendants as $
+\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $ respectively. We denote a path, $
+\pi(X, Y) = \{ X, V_0, V_1, \cdots, V_k, Y \} $ between $ X $ and $ Y $ in $ G
+$ where each consecutive pair of variables in $ \pi $ are connected through an
+edge. We say $ X $ and $ Y $ are d-connected in $ G $ given a conditioning set
+$ \bm{Z} \subseteq V - \{X, Y\} $, if for a path $ \pi(X, Y) $ (i) every collider 
+structure ($ V_1 \rightarrow V_2 \leftarrow V_3 $) on $ \pi(X, Y) $, $ \bm{Z} \cap \{ V_2, \textrm{De}_G(V_2) \} \ne \emptyset $, and (ii) for every non-collider structure
+($ V_1 \rightarrow V_2 \rightarrow V_3 $, $ V_1 \leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are d-separated given $ G $ if no d-connecting
+path exists between them.
 
-\subsection{Model Testing Using CI Tests}
-Our structure learning algorithm relies on the idea of testing DAGs using CI
-tests. In this section, we give an example of how this model testing and
-corresponding modification work. We take the example of the MEC learned by the
-PC algorithm in Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all
-the undirected edges in the model based on our domain knowledge. In this case
+\subsection{Testing DAGs Using CI Tests}
+Our structure learning algorithm uses the idea of testing DAGs using CI tests.
+In this section, we give an example of how this model testing works in
+practice. We take the example of the MEC learned by the PC algorithm in
+Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all the undirected
+edges in the model based on our domain knowledge to obtain a DAG. In this case
 we only have one undirected edge between Marital Status and Relationship. Let's
@@
 
- The implied CIs of the 
-model along with the test results are shown in the Figure.
-
-We can think of our proposed approach to be an iterative application of this model
-testing approach.
+Using these statistical tests we can determine the pair of variables that
+should have edge between them. We can think of our proposed approach to be an
+iterative application of this model testing approach.
 
@@
 conditional association between variables -- also known as partial association.
-Specifically, we are interested in testing CI tests: $ X \ci Y \bm{Z} $, i.e.,
-whether the variables $ X $ and $ Y $ are independent when conditioned on $
-\bm{Z} $. Secondly, we are interested in quantifying the strength of
-relationship between $ X $ and $ Y $ conditioned on $ \bm{Z} $. We can use
-the effect size measures of the CI tests for this quantification. In this 
-section, we give an overview of some of the effect size measures for 
-different types of (possibly mixed) data.
+Specifically, we are interested in testing CI tests: $ X \ci Y \rvert \bm{Z} $
+(with potentially $ \bm{Z} = \emptyset $. Secondly, we are interested in
+quantifying the strength of the relationship between $ X $ and $ Y $
+conditioned on $ \bm{Z} $. We can use the effect size measures of the CI tests
+for this quantification. In this section, we give an overview of some of the
+effect size measures for different types of (possibly mixed) data.
 
@@
 
-\paragraph{$ X $ is ordinal, and $ Y $ and $ Z $  are continuous or ordinal: }
+\paragraph{$ X $ is ordinal, and $ Y $ and $ \bm{Z} $  are continuous or ordinal: }
 
@@
 
-\paragraph{$ X $, $ Y $, and $ \bm{Z} $ are discrete (ordinal or caterical): }
+\paragraph{All $ X $, $ Y $, and $ \bm{Z} $ are discrete (ordinal or categorical): }
+
+In the case when we have a combination of ordinal and categorical variables, we use
+the
 

### 2025-02-10 paper/main.tex
Notation fixes

@@
 edge. We say $ X $ and $ Y $ are d-connected in $ G $ given a conditioning set
-$ \bm{Z} \subseteq V - \{X, Y\} $, if for a path $ \pi(X, Y) $ (i) every collider 
-structure ($ V_1 \rightarrow V_2 \leftarrow V_3 $) on $ \pi(X, Y) $, $ \bm{Z} \cap \{ V_2, \textrm{De}_G(V_2) \} \ne \emptyset $, and (ii) for every non-collider structure
-($ V_1 \rightarrow V_2 \rightarrow V_3 $, $ V_1 \leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are d-separated given $ G $ if no d-connecting
-path exists between them.
+$ \bm{Z} \subseteq V - \{X, Y\} $, if for a path $ \pi(X, Y) $ (i) for every
+collider structure ($ V_1 \rightarrow V_2 \leftarrow V_3 $) on $ \pi(X, Y) $, $
+\bm{Z} \cap \{ V_2, \textrm{De}_G(V_2) \} \ne \emptyset $, and (ii) for every
+non-collider structure ($ V_1 \rightarrow V_2 \rightarrow V_3 $, $ V_1
+\leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
+d-separated given $ \bm{Z} $ if no d-connecting path exists between them.
 

### 2025-02-10 paper/main.tex
done with intro

--- 2025-02-10 paper/main.tex
@@
 from datasets. These efforts have led to numerous causal discovery algorithms,
-such as constraint-based methods like PC algorithm \citep{Spirtes2001} and Fast
-Causal Inference \citep{Spirtes2000}), score-based methods such as Hill-Climb
+such as constraint-based methods like PC algorithm \citep{Spirtes2001,KalischB07} 
+and Fast Causal Inference \citep{Spirtes2000}), score-based methods such as Hill-Climb
 Search and Greedy Equivalence Search \citep{Chickering2002}, and continuous
-optimization-based methods like NO TEARS \citep{Zheng2018} and DAGMA
+optimization-based methods like NOTEARS \citep{Zheng2018} and DAGMA
 \citep{Bello2022}. Despite significant progress in automated causal discovery,
-their adoption in applied research has been limited. Researchers often prefer
-to manually construct DAGs based on their domain expertise \citep{Tennant2020,
-Petersen2021}. We attribute this preference to several challenges with existing
-causal discovery algorithms in practical settings:
+their adoption in applied research has been limited. Challenges encountered with existing
+causal discovery algorithms in practice include but are not limited to:
 
 \begin{enumerate}
-	\item \textbf{Lack of Trust:} While most algorithms are asymptotically
-		consistent, their behavior on finite samples is not well
-		understood. Their output can vary significantly depending on
-		the choice of algorithm and hyperparameters, making it
+	\item \textbf{Lack of Trust:} While constraint-based algorithms are 
+		often asymptotically
+		consistent \citep{KalischB07}, they can and do make mistakes
+		on  finite samples. These mistakes can be severe and contradict
+		obvious domain knowledge (think of edges going into unmodifiable
+		attributes such as Age). The choice of algorithm and hyperparameters
+		significantly aa=ffect the output, making it
 		difficult to assess reliability. Additionally, the absence of
 		robust performance evaluation methods for any given dataset
-		further reduce the confidence in their outputs. 
+		further reduce the confidence in their outputs. A recent paper
+		therefore advised Epidemiologists to not attempt using structure
+		learning algorithms without help of an expert \citep{Gururaghavendran_2024}.
 	\item \textbf{Outputs Markov Equivalence Class (MEC):} As multiple
@@
 		algorithms can only recover the MECs. These MECs can contain a
-		combination of directed and undirected edges. However, most
-		methods for downstream tasks, such as identification or causal
-		effect estimation, assume knowledge of a fully oriented DAG. 
+		combination of directed and undirected edges. This structural uncertainty
+		can make difficult or impossible to apply the learned model for downstream
+		tasks, such as identification or causal
+		effect estimation \citep{Maathuis_2009,PerkovicTKM17}.
 \end{enumerate}
 
-Figure~\ref{fig:intro} highlights some of these issues. On the other hand,
-constructing DAG based solely on domain expertise can be error prone as it
-requires us to distinguish between different causal structures, such as direct
-or indirect effects or causal mediation. Given these complexities, it is
-important to validate the consistency of a DAG against our dataset. One way to
-test this consistency is by testing whether the CI statements implied by the
-DAG hold in the data \citep{Ankan2021}. Specifically, each missing edge between
-a pair of variables in the DAG implies a CI statement, which can be tested in
-the dataset using a statistical test. If we find any violation to CI
-statements, we can make modifications to our model accordingly to fix the
-violations.
+Figure~\ref{fig:intro} highlights some of these issues. In practice, DAGs 
+are still largely constructed from domain knowledge alone 
+\citep{Tennant2020,Petersen2021}. This can, however, be equally problematic.
+Constructing DAGs requires us to distinguish between different causal structures, 
+such as direct or indirect effects or causal mediation. Specifically, there is often a lack
+of theoretical support for the \emph{absence} of direct causal effects, which are 
+the key assumptions that graphical models make to enable downstream causal inferences.
+Given the likelihood of making mistakes, it is therefore
+important to at least validate the consistency of a DAG against our dataset. One way to
+test this consistency is by testing whether the conditional independence 
+(CI) statements implied by the DAG hold in the data \citep{Ankan2021}. Specifically, 
+each missing edge between
+a pair of variables in the DAG leads to one or more CI statements, 
+which can be checked using statistical tests. Violations to CI
+statements can point to the erroneous omission of important causal effects from a model
+or to the presence of latent confounders.
 
@@
 	     Income Dataset \citep{Becker1996} using different causal discovery
-	     algorithms and sample sizes. Edge colors represents the sample size
+	     algorithms and sample sizes. Edge colors represent the sample size
 	     used: red for $N=400$, and blue for $N=800$. (a) PC algorithm with a
 	     mutual information based CI test, (b) PC algorithm with a
-    	     residualization based test \citep{Ankan2023}, (c) Hill Climb Search with
+    	     residualization-based test \citep{Ankan2023}, (c) Hill Climb Search with
 	     Bayesian Information Criterion (BIC) score. The learned model structure varies
@@
 model-building strategy that integrates domain expertise with CI testing on
-data. At each iteration, we use CI testing to identify pairs of variables where
-an edge is either missing or superfluous. We can then use this information
-along with domain expertise to determines whether to add or remove these edges.
+data. At each iteration, we use CI testing to identify pairs of variables that
+lack (direct or indirect) connections or are connected by a superfluous edge.
+ We then use this information
+along with domain expertise to determine which changes.
 The orientation of the edge when adding is determined based on the domain
-expertise. As at any iteration, multiple changes to the DAG are possible. We
-also give a way to rank potential new edges additions to the DAG using effect
-size measures for CI tests to address the most severe violations of CI tests
-first. As a result this approach acts like a guide to the expert when manually
-constructing their model while ensuring that the model is consistent with the
-data.
+expertise; critically, this does not require the domain expert to distinguish 
+between direct and indirect effects. Our resulting approach can be 
+thought of as a guidance during manual model construction that ensures the 
+model remains consistent to the data.
 
-Our main contributions in this paper are as follows:
+Our contributions are organized as follows:
 
 \begin{enumerate}
-    \item We develop an iterative method based on CI testing to assist researchers in 
-	    constructing DAGs using their domain expertise (Section~\ref{sec:modification}).
-    \item We show a ranking method to prioritize potential modifications to the DAG that 
+    \item In Section~\ref{sec:modification}, we present our iterative method for DAG construction
+	based on domain knowledge and data, and prove its correctness in an ``oracle setting''.
+    \item We then present a ranking method to prioritize potential modifications to the DAG that 
     	    help in fixing the most severe violations first (Section~\ref{sec:ranking}).
-    \item We show empirically that even experts that make mistakes can outperform purely data-driven
+    \item We show empirically that even experts that make mistakes can outperform purely 
+	data-driven
 	causal discovery algorithms (Section~\ref{sec:empirical}).

### 2025-02-10 paper/main.tex

@@
 \label{sec:background}
-\todo[inline]{Need to rewrite this}
-We denote random variables with uppercase letters $ X $, and a set of random
-variables as $ \bm{X} = \{X_1, \cdots, X_k\} $ with $ \rvert \bm{X} \rvert = k
-$. A sample from the random variable $ X $ is denoted as $ x $ and from a set
-of random variables $ \bm{X} $ is denoted as $ \bm{x} $. In this paper, we
-consider random variables in the mixed data setting where each of the random
-variables can be either of continuous, categorical, or ordinal unless
-specified. We write the expectation of a variable $ X $ as $ \mathbb{E}[X] $,
-conditional expectations as $ \mathbb{E}[X | \bm{Z}] $. We denote a graph $ G =
-(V, E) $ which nodes $ V$ and edges $ E $. In this paper we focus on all
-observed variables, and linear measure of association. 
-where $ \mathrm{cov}(X, Y) $ is the covariance between $ X $ and $ Y $, $ \mathrm{corr}(X, Y) $ is the correlation between $ X $ and $ Y $, and $
-\sigma_X $ and $ \sigma_Y $ are the standard deviations of $ X $ and $ Y $
-respectively.  Covariance matrix using $ \Sigma $ and the entry corresponding to
-covariance between $ X $ and $ Y $ in the matrix using $ \Sigma_{XY} $.
+We denote random variables using uppercase letters, such as $ X $, and a set of
+random variables by $ \bm{X} = \{X_1, \cdots, X_k\} $, where $ \rvert \bm{X}
+\rvert = k $ indicates the number of variables in the set. We denote the
+covariance between variables $ X $ and $ Y $ as $ \mathrm{cov}(X, Y) $, the
+covariance matrix as $ \Sigma $, and the entry corresponding to covariance
+between $ X $ and $ Y $ using $ \Sigma_{XY} $. In the context of this paper, we
+consider causal DAGs that contain only observed variables. The structure of the
+DAG is represented using a graph $ G = (V, E) $, where $ V $ are the nodes
+corresponding to the random variables, and $ E $ are the edges denoting the
+causal relationship between the variables. The set of parents of $ X $ in $ G $
+are denoted as $ \textrm{Pa}_G(X) $, and its ancestors and descendants as $
+\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $ respectively. We denote a path, $
+\pi(X, Y) = \{ X, V_0, V_1, \cdots, V_k, Y \} $ between $ X $ and $ Y $ in $ G
+$ where each consecutive pair of variables in $ \pi $ are connected through an
+edge. We say $ X $ and $ Y $ are d-connected in $ G $ given a conditioning set
+$ \bm{Z} \subseteq V - \{X, Y\} $, if for a path $ \pi(X, Y) $ (i) for every
+collider structure ($ V_1 \rightarrow V_2 \leftarrow V_3 $) on $ \pi(X, Y) $, $
+\bm{Z} \cap \{ V_2, \textrm{De}_G(V_2) \} \ne \emptyset $, and (ii) for every
+non-collider structure ($ V_1 \rightarrow V_2 \rightarrow V_3 $, $ V_1
+\leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
+d-separated given $ \bm{Z} $ if no d-connecting path exists between them.
 
-\subsection{Model Testing Using CI Tests}
-Our structure learning algorithm relies on the idea of testing DAGs using CI
-tests. In this section, we give an example of how this model testing and
-corresponding modification work. We take the example of the MEC learned by the
-PC algorithm in Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all
-the undirected edges in the model based on our domain knowledge. In this case
+\subsection{Testing DAGs Using CI Tests}
+Our structure learning algorithm uses the idea of testing DAGs using CI tests.
+In this section, we give an example of how this model testing works in
+practice. We take the example of the MEC learned by the PC algorithm in
+Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all the undirected
+edges in the model based on our domain knowledge to obtain a DAG. In this case
 we only have one undirected edge between Marital Status and Relationship. Let's

### 2025-02-10 paper/main.tex
Updates background section

--- 2025-02-10 paper/main.tex
@@
 \leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
-d-separated given $ \bm{Z} $ if no d-connecting path exists between them.
+d-separated given $ \bm{Z} $ if no d-connecting path exist between them.
 
@@
 practice. We take the example of the MEC learned by the PC algorithm in
-Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all the undirected
+Figure~\ref{fig:intro}(a) using $ N=800 $. We first orient all the undirected
 edges in the model based on our domain knowledge to obtain a DAG. In this case
-we only have one undirected edge between Marital Status and Relationship. Let's
-orient the edge as Marital Status to Relationship.
+we only have one undirected edge between Marital Status and Relationship. We 
+orient the edge as Marital Status to Relationship. We then proceed to test
+each of the implied CI of the model using a statistical test. We show the result
+of some of these tests in Figure.
 
@@
 conditional association between variables -- also known as partial association.
-Specifically, we are interested in testing CI tests: $ X \ci Y \rvert \bm{Z} $
-(with potentially $ \bm{Z} = \emptyset $. Secondly, we are interested in
-quantifying the strength of the relationship between $ X $ and $ Y $
-conditioned on $ \bm{Z} $. We can use the effect size measures of the CI tests
-for this quantification. In this section, we give an overview of some of the
-effect size measures for different types of (possibly mixed) data.
+We are interested in testing CI tests: $ X \ci Y \rvert \bm{Z} $ (with
+potentially $ \bm{Z} = \emptyset $) in data and quantifying the strength of the
+relationship between $ X $ and $ Y $ conditioned on $ \bm{Z} $. For the
+quantification we use effect size measures based on the CI test. In this
+section, we give an overview of some of the CI tests and effect size measures
+for different types of (possibly mixed) data.
 
 \paragraph{Both $ X $ and $ Y $ are continuous: }
-When both $ X $ and $ Y $ are continuous, Pearson's correlation coefficient is
-typically used. When $ \bm{Z} = \emptyset $, the correlation
-coefficient is defined as:
+When both $ X $ and $ Y $ are continuous, Pearson's correlation test can be
+used and Pearson's correlation coefficient can be used at the effect size. When
+$ \bm{Z} = \emptyset $, the correlation coefficient is defined as:
 
@@
 E_Y: Y \sim \bm{Z} $, calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $
-R_Y = Y - E_Y(\bm{Y}) $, and then computing the correlation between the
-residuals:
+R_Y = Y - E_Y(\bm{Y}) $, and then computing the correlation coefficient between
+the residuals:
 
 \begin{equation}
-	r_{X, Y; \bm{Z}} = r_{R_X, R_Y}
+	r_{X, Y \rvert \bm{Z}} = r_{R_X, R_Y}
 \end{equation}
@@
 Using the estimated covariance matrix, $ \Sigma $ we can the compute Pearson's
-correlation coefficient. When $ \bm{Z} = \emptyset $,
+correlation coefficient to use as the effect size as well as for testing. When
+$ \bm{Z} = \emptyset $,
 
@@
 \begin{equation}
-	r_{X, Y; \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
+	r_{X, Y \rvert \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
 \end{equation}
@@
 
-In the case when we have a combination of ordinal and categorical variables, we use
-the
+In the case when we have a combination of ordinal and categorical variables, we
+can use the residualization based CI test from \citep{Ankan2023}. This test
+gives a chi-square distributed test statistic, that can be used for computing
+an effect size such as Root Mean Squared Error of Approximation (RMSEA). Given
+a chi-squared test statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
+freedom, the RMSEA is defined as:
 
-\todo[inline]{Here we use an effect size RMSEA based on the previous Hotelling's test paper.}
+\begin{equation}
+	\textrm{RMSEA}_{X, Y \rvert \bm{Z}} = \sqrt{\frac{\chi^2 - \textrm{df}}{\textrm{df} (N-1)}}
+\end{equation}
+
+where $ N $ is the sample size.
 

### 2025-02-10 paper/main.tex
Updates background

@@
 
-\subsection{Testing DAGs Using CI Tests}
+\subsection{Testing DAG Using CI Tests}
 Our structure learning algorithm uses the idea of testing DAGs using CI tests.
@@
 of some of these tests in Figure.
+
+\todo[inline]{Add a table here}
+% \begin{figure}
+% 	\centering
+% 	\begin{tabular}{llrr}
+% 		\multicolumn{4}{c}{Z=Age, Sex} \\
+% 		$X$ & $Y$ & \bm{Z} & p \\
+% 		\hline
+% 		Edct & Wrkc & 1.00  \\
+% 		Occp & Wrkc & 1.00  \\
+% 		Rltn & HrPW &  .99  \\
+% 		Incm & Occp & .08   \\
+% 		Incm & Wrkc &  .50  \\
+% 		Incm & HrPW & .003  \\
+% 		\hline
+% 	\end{tabular}
+% \end{figure}
 
@@
 estimate the covariance matrix while maximizing the likelihood of the dataset.
-Using the estimated covariance matrix, $ \Sigma $ we can the compute Pearson's
-correlation coefficient to use as the effect size as well as for testing. When
-$ \bm{Z} = \emptyset $,
+Using the estimated covariance matrix, $ \Sigma $ we can perform the CI test and compute 
+the Pearson's correlation coefficient as the effect size:
 
 \begin{equation}
-	r_{X, Y} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}
-\end{equation}
-
-When $ \bm{Z} \neq \emptyset $,
-
-\begin{equation}
-	r_{X, Y \rvert \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
+	r_{X, Y \rvert \bm{Z} = \emptyset} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}; \;\;\; \; r_{X, Y \rvert \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
 \end{equation}
@@
 In the case when we have a combination of ordinal and categorical variables, we
-can use the residualization based CI test from \citep{Ankan2023}. This test
-gives a chi-square distributed test statistic, that can be used for computing
-an effect size such as Root Mean Squared Error of Approximation (RMSEA). Given
-a chi-squared test statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
+can use a residualization based CI test \citep{Ankan2023}. This test gives a
+chi-square distributed test statistic, that can be used for computing an effect
+size such as the Root Mean Squared Error of Approximation (RMSEA). Given a
+chi-squared test statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
 freedom, the RMSEA is defined as:

### 2025-02-10 paper/main.tex
edits sections 5 & 6

--- 2025-02-10 paper/main.tex
@@
 	\end{subfigure}
-	\caption{Model fit improvement over iterations on the Adult Income dataset. In the expert simulation, an 
-		LLM is used to determine the edge orientations. (a) Shows the total residual association, $\tau$. 
-		(b) Shows the log-likelihood of the overall model.
+	\caption{LLM-in-the-loop causal discovery on the Adult Income datases. As the ground 
+truth is unknown, two measures of fit are shown over 30 iterative modifications: 
+total residual association $\tau$ (a, lower is bettter); log-likelihood (a, higher is better).
 	}
@@
 geared towards using  large language models (LLMs) and humans as experts
-to decide on edge directions.
+to decide on edge directions. We use the Adult Income dataset from the 
+introduction as an example. There is no known ground truth for this dataset,
+and it contains mixed types of data; here, we use a version that contains
+ordinal and nominal variables and use the RMSEA as a measure of association
+to prioritize modifications (Section~\ref{sec:background}).
 
@@
 causal direction between a pair of variables. The LLM is provided a description
-of each of the variables for the task. We have provided the prompt used in
+of each of the variables and given the prompt shown in
 Appendix~\ref{section:llms}. 
-Figure~\ref{fig:unexplained_ll} illustrates this distinction by comparing GES
-to \textsc{ExpertInLoop} on the Adult income dataset across iterations of a
-single run of the algorithms.
+Figure~\ref{fig:unexplained_ll} shows how the model fit to the data improves 
+across 30 iterations compared to GES for a run of the algorithm.
 Figure~\ref{fig:adult_llm} shows the DAG learned
-on the Adult Income dataset using the LLM as the expert. Compared to other
-methods that use LLMs for causal discovery, our approach combines a data-driven
-method with LLMs. We believe this should help in reducing the amount of
-mistakes made by the LLM as well as reduce the amount of information needed
-from the LLM.
+on the Adult Income dataset using the LLM as the expert. We can see 
+that the model fit improves comparably to a greedy approach and 
+the resulting DAG contains largely sensible edge directions.
 
@@
 corresponding to the dataset's variables. They can then specify a p-value
-threshold and a measure of association threshold. To accommodate mixed data
-types, we use a measure of association based on canonical correlations
-(detailed in Appendix~\ref{sec:mixed_association}).
-
-Using the specified thresholds, the tool visually highlights variable pairs
+threshold and a threshold for a minimal association strength. The tool 
+then visually highlights variable pairs
 with a residual association greater than the threshold, marking them with red
-edges. The thickness of these edges represents the strength of association,
-helping users prioritize which variable pairs to address first. Similarly, if
-an existing edge is superfluous, it is highlighted in black, signaling
-potentially unnecessary edge. Using this information, users can iteratively
+edges; to avoid cluttering, no lines are drawn for residual associations that
+are statistically insignificanct or below the specified threshold.
+The thickness of these edges represents the strength of association,
+helping users prioritize which variable pairs to address first; however, note that
+users are free to choose which variables to connect. Similarly, if
+an existing edge is found to be potentially superfluous (statistically insignificant),
+it is highlighted in black. Using this information, users can iteratively
 modify the model by adding or removing edges. The tool also computes the Root
-Mean Square Error of Approximation (RMSEA) to provide an estimate of the
+Mean Square Error of Approximation (RMSEA) based on Shipley's C test, a global test 
+of model fit based on the implied CIs \citep{Shipley2000}, to provide an estimate of the
 overall fit of the model. Once satisfied with the constructed DAG, users can
@@
 \section{Conclusions}
-\todo[inline]{Rewrite}
+
 Researchers often prefer to construct DAGs manually rather than using automated
 algorithms due to practical challenges in applying automated methods. To assist
-in the process, we developed an iterative structure learning method that
-integrates manual construction with data-driven insights. This approach bridges
-the gap between fully manually and fully automated methods by providing a
-systematic way to construct DAGs while utilizing data.
+this process, we developed an iterative structure learning method that
+integrates manual construction with data-driven feedback, bridging
+the gap between fully manually and fully automated methods.
+The idea of augmenting structure learning by domain knowledge is of course not 
+new. Common ways to provide such background knowledge are "whitelists" or "blacklists"
+of edges that must or must not be present in the resulting DAG, or specification
+of a temporal structure between the variables. 
+Compared to such approaches, where domain knowledge 
+is specified and supplied beforehand, the novelty of our method lies in the interactive
+back-and-forth between domain knowledge and data, which provably leads to the correct
+result in a polynomial amount of steps (if the expert answers all
+ancestral queries correctly).
 
-In future we would like to extend our approach to latent variables and Acyclic
-Directed Mixed Graphs (ADMGs), explore the potential of combining our approach
-with pairwise orientation methods. We would also improve the web interface for
-better usability. 
+Avenues for future work include extending this approach to latent variables 
+as encoded by acyclic directed mixed graphs (ADMGs). Further, if experts make 
+mistakes that lead to inconsistencies during model construction (such as cycles),
+this information could be leveraged in a more systematic way to allow backtracking
+and recovery.
 \bibliography{references}

### 2025-02-10 paper/main.tex

--- 2025-02-10 paper/main.tex
@@
 \leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
-d-separated given $ \bm{Z} $ if no d-connecting path exists between them.
+d-separated given $ \bm{Z} $ if no d-connecting path exist between them.
 
-\subsection{Testing DAGs Using CI Tests}
+\subsection{Testing DAG Using CI Tests}
 Our structure learning algorithm uses the idea of testing DAGs using CI tests.
@@
 practice. We take the example of the MEC learned by the PC algorithm in
-Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all the undirected
+Figure~\ref{fig:intro}(a) using $ N=800 $. We first orient all the undirected
 edges in the model based on our domain knowledge to obtain a DAG. In this case
-we only have one undirected edge between Marital Status and Relationship. Let's
-orient the edge as Marital Status to Relationship.
+we only have one undirected edge between Marital Status and Relationship. We 
+orient the edge as Marital Status to Relationship. We then proceed to test
+each of the implied CI of the model using a statistical test. We show the result
+of some of these tests in Figure.
+
+\todo[inline]{Add a table here}
+% \begin{figure}
+% 	\centering
+% 	\begin{tabular}{llrr}
+% 		\multicolumn{4}{c}{Z=Age, Sex} \\
+% 		$X$ & $Y$ & \bm{Z} & p \\
+% 		\hline
+% 		Edct & Wrkc & 1.00  \\
+% 		Occp & Wrkc & 1.00  \\
+% 		Rltn & HrPW &  .99  \\
+% 		Incm & Occp & .08   \\
+% 		Incm & Wrkc &  .50  \\
+% 		Incm & HrPW & .003  \\
+% 		\hline
+% 	\end{tabular}
+% \end{figure}
 
@@
 estimate the covariance matrix while maximizing the likelihood of the dataset.
-Using the estimated covariance matrix, $ \Sigma $ we can the compute Pearson's
-correlation coefficient. When $ \bm{Z} = \emptyset $,
+Using the estimated covariance matrix, $ \Sigma $ we can perform the CI test and compute 
+the Pearson's correlation coefficient as the effect size:
 
 \begin{equation}
-	r_{X, Y} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}
-\end{equation}
-
-When $ \bm{Z} \neq \emptyset $,
-
-\begin{equation}
-	r_{X, Y; \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
+	r_{X, Y \rvert \bm{Z} = \emptyset} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}; \;\;\; \; r_{X, Y \rvert \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
 \end{equation}
@@
 
-In the case when we have a combination of ordinal and categorical variables, we use
-the
+In the case when we have a combination of ordinal and categorical variables, we
+can use a residualization based CI test \citep{Ankan2023}. This test gives a
+chi-square distributed test statistic, that can be used for computing an effect
+size such as the Root Mean Squared Error of Approximation (RMSEA). Given a
+chi-squared test statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
+freedom, the RMSEA is defined as:
 
-\todo[inline]{Here we use an effect size RMSEA based on the previous Hotelling's test paper.}
+\begin{equation}
+	\textrm{RMSEA}_{X, Y \rvert \bm{Z}} = \sqrt{\frac{\chi^2 - \textrm{df}}{\textrm{df} (N-1)}}
+\end{equation}
+
+where $ N $ is the sample size.
 

### 2025-02-10 code/fig4/make_plots.R
Compacts figures

@@ -37,5 +37,5 @@ p_ll <- ggplot(d_results, aes(x=x, y=ll, color=algo)) +
 	theme(legend.position='top',
 	      legend.title=element_blank())
 
-ggsave('unexplained_effect.pdf', p_unexplained_effect, height=2.5, width=1.6, units='in')
-ggsave('ll.pdf', p_ll, height=2.5, width=1.6, units='in')
+ggsave('unexplained_effect.pdf', p_unexplained_effect, height=2.0, width=1.6, units='in')
+ggsave('ll.pdf', p_ll, height=2.0, width=1.6, units='in')

### 2025-02-10 paper/main.tex
Empirical section

@@
 \subsection{Effect Sizes for CI Tests}
+\label{sec:ci_tests}
 
@@
 
-In this section, we give some of the practical aspects of the
-\textsc{ExpertInLoop} algorithm, and then compare the algorithm with automated
-causal discovery algorithms.
-
-In all experiments, we use the \textsc{RankedExpand} method together
-with a version of the strong ancestral oracle that can make mistakes.
-
-\todo{Want to know: what if my experts are not perfect?}
-
-\begin{figure}[t!]
-	\centering
-	\begin{subfigure}{0.5\textwidth}
-		\centering
-		\includegraphics{../code/fig3/shd_ribbon.pdf}
-		\caption{}
-	\end{subfigure}
-	\begin{subfigure}{0.5\textwidth}
-		\centering
-		\includegraphics{../code/fig3/sid_ribbon.pdf}
-		\caption{}
-	\end{subfigure}
-	\caption{Comparison of PC, Hill-Climb Search, and GES algorithms against
-		\textsc{ExpertInLoop} algorithm. As automated algorithms only
-		recover the CPDAG, we use the best and worst scoring
-		orientation of the CPDAG to get the range. We test
-		\textsc{ExpertInLoop} with varying values of expert accuracy, $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $. The corresponding
-		$\alpha_{\textrm{eff}} $ is shown in the plot.}
-	\label{fig:shd_sid}
-\end{figure}
-
-The \textsc{ExpertInLoop} algorithm requires two oracles, one that can answer
-d-separation queries and an ancestral oracle that can give the causal direction
-between variable pairs. For the experiments, we simulate the d-separation
-oracle using a CI test with a p-value threshold of $ 0.05 $. To simulate the
-edge orientation oracle, we simulate a domain expert that can give the correct
-causal relationship with some accuracy $ \alpha $.
+In this section, we compare our \textsc{ExpertInLoop} algorithm with automated
+causal discovery algorithms. Our goal with the empirical analysis is to
+understand the behaviour of our method when the experts are not perfect. To
+simulate this imperfect expert, we use a version of the strong ancestral oracle
+that has an accuracy of $ \alpha $:
 
@@
 \mathrm{Expert} $ is higher than $ \alpha $ as even when giving a random
-answer, there is a $ 1 $ in $ 3 $ chance that the answer is correct by chance.
-Hence, for an $ \mathrm{Expert} $ with accuracy $ \alpha $, the effective
-accuracy becomes:
+answer, there is a $ 1 $ in $ 3 $ chance that the answer is correct. Therefore,
+an $ \mathrm{Expert} $ with accuracy $ \alpha $, has an effective accuracy:
 
@@
 
+\begin{figure}[t!]
+	\centering
+	\begin{subfigure}{0.5\textwidth}
+		\centering
+		\includegraphics{../code/fig3/shd_ribbon.pdf}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}{0.5\textwidth}
+		\centering
+		\includegraphics{../code/fig3/sid_ribbon.pdf}
+		\caption{}
+	\end{subfigure}
+	\caption{Comparison of PC, Hill-Climb Search, and GES algorithms against
+		\textsc{ExpertInLoop} algorithm. As automated algorithms only
+		recover the CPDAG, we use the best and worst scoring
+		orientation of the CPDAG to get the range. We test
+		\textsc{ExpertInLoop} with varying values of expert accuracy, $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $. The corresponding
+		$\alpha_{\textrm{eff}} $ is shown in the plot.}
+	\label{fig:shd_sid}
+\end{figure}
+
+, we use the
+\textsc{RankedExpand} method together with a version of the strong ancestral
+oracle that can make mistakes.
+
 We analyzed the performance of \textsc{ExpertInLoop} algorithm by comparing it
 with three other automated algorithms: PC, Hill-Climb Search, and Greedy
-Equivalence Search (GES). For the analysis, we simulated data from known DAGs
-and use the algorithms to recover the original DAG. We start by generating a
-random DAG on $10$ nodes and use linear models with random effects to simulate
-$ 500 $ samples from it. We simulate each variable $ V_i $ in the DAG as:
+Equivalence Search (GES) on linear Gaussian data. For \textsc{ExpertInLoop}, we
+use the \textsc{RankedExpand} method with Pearson's correlation coefficient as
+the effect size measure (Section~\ref{ref:ci_tests}). For the analysis, we
+simulated data from known DAGs and use the algorithms to recover the original
+DAG. We start by generating a random DAG on $10$ nodes and use linear models
+with random effects to simulate $ 500 $ samples from it. We simulate each
+variable $ V_i $ in the DAG as:
 
@@
 	\begin{split}
-		\beta &= \mathrm{Uniform}([-0.6, 0.6]) \\
+		\bm{\beta} &= \mathrm{Uniform}([-0.6, 0.6]) \\
 		V_i &= \bm{\beta} \mathrm{Pa}_G(V_i) + \mathcal{N}(0, 1)
@@
 \citep{Peters2015}. Unlike \textrm{ExpertInLoop}, the automated algorithms can
-only recover the CPDAG, we consider all possible orientations of the CPDAG to
-compute the range of SHD and SID values to show the best and worst case
-results. The results of this analysis are shown in Figure~\ref{fig:shd_sid}.
-For SHD, the performance of \textsc{ExpertInLoop} is comparable to the
-automated algorithms for $ \alpha = 0.3 (\alpha_{\textrm{eff}} = 0.53 $, which
-is equivalent to the expert getting the orientation correct $ 1 $ out of $ 2 $
-case. For SID, the performance is comparable to the automated algorithms for $
-\alpha = 0.5 (\alpha_{\textrm{eff}} = 0.67) $, which is equivalent to the
-expert getting the orientation correct in $ 2 $ out of $ 3 $ cases. For denser
-DAGs, the \textsc{ExpertInLoop} is able to perform better for lower $ \alpha $
-values.
+only recover the MEC, therefore, we considered all possible orientations of the
+MEC to compute a range of SHD and SID values representing the best and worst
+case scenarios. The results of this analysis are shown in
+Figure~\ref{fig:shd_sid}. For SHD, the performance of \textsc{ExpertInLoop} is
+comparable to the automated algorithms for $ \alpha = 0.3
+(\alpha_{\textrm{eff}} = 0.53 $, implying that if an expert is able to get the
+edge orientation correctly in more than $ 1 $ out of $ 2 $ cases, they can
+outperform the automated algorithms. Similarly for SID, an expert can outperform 
+the automated algorithms if they are able to get the orientation correct in $ 2 $ 
+out of $ 3 $ cases. Another thing to note is that for denser DAGs, the 
+\textsc{ExpertInLoop} is able to perform better for lower $ \alpha $ values.
 

### 2025-02-10 paper/main.tex
Make equations compacter

@@
 				\end{cases} \\
-	\end{split}
-\end{equation}
-
-where given the true DAG, $ G $:
-
-\begin{equation}
-	\begin{split}
 		\mathrm{TrueReln}(X, Y) &= \begin{cases}
-					X \rightarrow Y, & \textrm{if } X \rightarrow Y \in G \\
-					Y \rightarrow X, & \textrm{if } Y \rightarrow X \in G \\
+				X \rightarrow Y, & \textrm{if } X \rightarrow Y \in G_{\textrm{true}} \\
+				Y \rightarrow X, & \textrm{if } Y \rightarrow X \in G_{\textrm{true}} \\
 					\textrm{None}, & \textrm{otherwise }
@@
 
+where $G_{\textrm{true}}$ is the true data generating DAG.
+
 An important point to note here is that the effective accuracy of $
@@
 answer, there is a $ 1 $ in $ 3 $ chance that the answer is correct. Therefore,
-an $ \mathrm{Expert} $ with accuracy $ \alpha $, has an effective accuracy:
-
-\begin{equation}
-	\alpha_{\mathrm{eff}} = \alpha + (1 - \alpha) / 3
-\end{equation}
+an $ \mathrm{Expert} $ with accuracy $ \alpha $, has an effective accuracy, $ \alpha_{\mathrm{eff}} = \alpha + (1 - \alpha) / 3 $.
 
@@
 \end{figure}
-
-, we use the
-\textsc{RankedExpand} method together with a version of the strong ancestral
-oracle that can make mistakes.
 

### 2025-02-10 paper/main.tex
Updates

@@
 To implement Algorithm~\ref{algo:ranked_expand}, we need to have an effect size
-metric by which we can rank effects. For some simple cases, such as 
-linear-Gaussian models, such effect sizes exist. However, a challenge arises
-when using mixed data. We here present a standardized effect size hat 
-is useable for datasets containing mixtures of continuous, ordinal and
-
-The
-algorithm selects the edge that has the highest absolute residual association,
-resulting in prioritization of edges that contribute the most to improving the
-model.
+metric by which we can rank effects. Depending on the type of variables, we can
+use the effect size measures shown in Section~\ref{sec:ci_tests}. The algorithm
+selects the edge that has the highest absolute residual association, resulting
+in prioritization of edges that contribute the most to improving the model.
 
@@
 use the \textsc{RankedExpand} method with Pearson's correlation coefficient as
-the effect size measure (Section~\ref{ref:ci_tests}). For the analysis, we
+the effect size measure (Section~\ref{sec:ci_tests}). For the analysis, we
 simulated data from known DAGs and use the algorithms to recover the original

### 2025-02-10 paper/main.tex
Fixes
Compact

@@
 used and Pearson's correlation coefficient can be used at the effect size. When
-$ \bm{Z} = \emptyset $, the correlation coefficient is defined as:
-
-\begin{equation}
-	r_{X, Y} = \frac{\mathrm{cov}(X, Y)}{\sigma_X \sigma_Y}
-\end{equation}
-
-When $ \bm{Z} \neq \emptyset $, partial correlation coefficient can be used.
-This is estimated by fitting two regression models $ E_X: X \sim \bm{Z} $ and $
-E_Y: Y \sim \bm{Z} $, calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $
-R_Y = Y - E_Y(\bm{Y}) $, and then computing the correlation coefficient between
-the residuals:
-
-\begin{equation}
-	r_{X, Y \rvert \bm{Z}} = r_{R_X, R_Y}
-\end{equation}
+$ \bm{Z} = \emptyset $, the correlation coefficient is defined as: $ r_{X, Y} =
+\mathrm{cov}(X, Y) / (\sigma_X \sigma_Y) $. When $ \bm{Z} \neq \emptyset $,
+partial correlation coefficient can be used. This is estimated by fitting two
+regression models $ E_X: X \sim \bm{Z} $ and $ E_Y: Y \sim \bm{Z} $,
+calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $ R_Y = Y - E_Y(\bm{Y})
+$, and then computing the correlation coefficient between the residuals: $r_{X,
+Y \rvert \bm{Z}} = r_{R_X, R_Y}$
 
@@
 estimate the covariance matrix while maximizing the likelihood of the dataset.
-Using the estimated covariance matrix, $ \Sigma $ we can perform the CI test and compute 
-the Pearson's correlation coefficient as the effect size:
-
-\begin{equation}
-	r_{X, Y \rvert \bm{Z} = \emptyset} = \frac{\Sigma_{XY}}{\sqrt{\Sigma_{XX} \Sigma_{YY}}}; \;\;\; \; r_{X, Y \rvert \bm{Z}} = - \frac{\Sigma^{-1}_{XY}}{\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}}
-\end{equation}
+Using the estimated covariance matrix, $ \Sigma $ we can perform the CI test
+and compute the Pearson's correlation coefficient as the effect size. When 
+$\bm{Z} = \emptyset $, $ r_{X, Y} = \Sigma_{XY} / (\sqrt{\Sigma_{XX} \Sigma_{YY}}) $, 
+	and when $ \bm{Z} \ne \emptyset $,
+	$ r_{X, Y \rvert \bm{Z}} = - \Sigma^{-1}_{XY}/ (\sqrt{\Sigma^{-1}_{XX} \Sigma^{-1}_{YY}}) $.
 
@@
 chi-squared test statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
-freedom, the RMSEA is defined as:
-
-\begin{equation}
-	\textrm{RMSEA}_{X, Y \rvert \bm{Z}} = \sqrt{\frac{\chi^2 - \textrm{df}}{\textrm{df} (N-1)}}
-\end{equation}
-
-where $ N $ is the sample size.
+freedom, the RMSEA is defined as: $ \textrm{RMSEA}_{X, Y \rvert \bm{Z}} =
+\sqrt{\chi^2 - \textrm{df}/ (\textrm{df} (N-1))} $, where $ N $ is the sample
+size.
 
@@
 		\mathrm{Expert}(\alpha) &= \begin{cases} 
-			\mathrm{TrueReln}(X, Y),  & \textrm{if  } x <= \alpha \\
+			\mathcal{O}(X, Y),  & \textrm{if  } x <= \alpha \\
 			\textrm{rand}(X \rightarrow Y, Y \leftarrow X, \textrm{None}) & \textrm{otherwise} \\
 				\end{cases} \\
-		\mathrm{TrueReln}(X, Y) &= \begin{cases}
+			\mathcal{O}(X, Y) &= \begin{cases}
 				X \rightarrow Y, & \textrm{if } X \rightarrow Y \in G_{\textrm{true}} \\

### 2025-02-10 paper/main.tex
Adds acknowledgements for use of Chat gpt
Improves plots

@@
 practice. We take the example of the MEC learned by the PC algorithm in
-Figure~\ref{fig:intro}(a) using $ N=800 $. We first orient all the undirected
+Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all the undirected
 edges in the model based on our domain knowledge to obtain a DAG. In this case
-we only have one undirected edge between Marital Status and Relationship. We 
-orient the edge as Marital Status to Relationship. We then proceed to test
-each of the implied CI of the model using a statistical test. We show the result
-of some of these tests in Figure.
+we only have one undirected edge between \emph{Marital Status} and
+\emph{Relationship} that we orient as $ \textrm{Marital Status} \rightarrow
+\textrm{Relationship} $. We can now test the implied CIs of this DAG using
+data. Figure~\ref{fig:ci_table} shows the result of this testing on some of the
+implied CIs. If we use a p-value threshold of $ 0.05 $, the first $ 3 $ implied
+CIs hold in the data whereas the remaining tests fail. One of the ways to fix
+these failing tests is to add an edge between the variables $ X $ and $ Y $.
+While this method tells up which variable pairs need to have an edge between
+them, it does not tell anything about the orientation of the edge, and we would
+need to orient it based on our domain knowledge.
 
-\todo[inline]{Add a table here}
-% \begin{figure}
-% 	\centering
-% 	\begin{tabular}{llrr}
-% 		\multicolumn{4}{c}{Z=Age, Sex} \\
-% 		$X$ & $Y$ & \bm{Z} & p \\
-% 		\hline
-% 		Edct & Wrkc & 1.00  \\
-% 		Occp & Wrkc & 1.00  \\
-% 		Rltn & HrPW &  .99  \\
-% 		Incm & Occp & .08   \\
-% 		Incm & Wrkc &  .50  \\
-% 		Incm & HrPW & .003  \\
-% 		\hline
-% 	\end{tabular}
-% \end{figure}
-
-Using these statistical tests we can determine the pair of variables that
-should have edge between them. We can think of our proposed approach to be an
-iterative application of this model testing approach.
+\begin{figure}
+	\centering
+	\begin{tabular}{llrr}
+		$X$ & $Y$ & $ \bm{Z} $ & p-value \\
+		\hline
+		Incm & MrtS &  Age, Occp, Rltn & $ 0.29 $     \\
+		Incm & Sex  &  Occp, Rltn      & $ 0.21 $     \\
+		Incm & Wrkc &  Occp 	       & $ 1.5e^{-3} $ \\
+		Edct & MrtS & Incm, Occp, Rltn & $ 0.02 $       \\
+		\hline
+	\end{tabular}
+	\caption{Results of testing some of implied CIs of the DAG in
+		 Figure~\ref{fig:intro}(b) using the same CI test that was used for
+		 learning it.}
+	\label{fig:ci_table}
+\end{figure}
 
@@
 and recovery.
+
+\begin{acknowledgements} 
+	We acknowledge the use of ChatGPT (\url{https://chatgpt.com/}) for
+	proofreading parts the manuscript.
+\end{acknowledgements}
+
 \bibliography{references}

### 2025-02-10 code/fig1/fig2.R

@@ -1,4 +1,5 @@
 library(dagitty)
+source('ci_tests.R')
 
 d <- read.csv("../utils/adult_unproc.csv")
 
@@ -26,11 +27,30 @@ d$HoursPerWeek <- as.double(d$HoursPerWeek)
 education_levels <- c("Preschool", "1st-4th", "5th-6th", "7th-8th", "9th", "10th", "11th", "12th", "HS-grad", "Some-college", "Assoc-voc", "Assoc-acdm", "Bachelors", "Masters", "Prof-school", "Doctorate")
 d$Education <- factor(d$Education, levels = education_levels, ordered = T)
 
+d$AgeStr <- rep("<21", nrow(d))
+d$AgeStr[d$Age >= 21 & d$Age <= 30] <- "21-30"
+d$AgeStr[d$Age >= 31 & d$Age <= 40] <- "31-40"
+d$AgeStr[d$Age >= 41 & d$Age <= 50] <- "41-50"
+d$AgeStr[d$Age >= 51 & d$Age <= 60] <- "51-60"
+d$AgeStr[d$Age >= 61 & d$Age <= 70] <- "61-70"
+d$AgeStr[d$Age > 70] <- ">70"
+d$Age <- as.factor(d$AgeStr)
+d$AgeStr <- NULL
+
+d$HoursPerWeekStr <- as.character(d$HoursPerWeek)
+d$HoursPerWeekStr[d$HoursPerWeek <= 20] <- "<=20" 
+d$HoursPerWeekStr[d$HoursPerWeek > 40] <- ">40" 
+d$HoursPerWeekStr[d$HoursPerWeek >= 20 & d$HoursPerWeek < 30] <- "21-30"
+d$HoursPerWeekStr[d$HoursPerWeek >= 30 & d$HoursPerWeek <= 40] <- "31-40" 
+d$HoursPerWeek <- as.factor(d$HoursPerWeekStr)
+d$HoursPerWeekStr <- NULL
+
+
 set.seed(42)
 d <- d[complete.cases(d), ]
-d <- d[sample(nrow(d), 1000), ]
+d <- d[sample(nrow(d), 800), ]
 
-dag <- dagitty("dag{MaritalStatus -> Relationship
+dag1 <- dagitty("dag{MaritalStatus -> Relationship
 	       	    Relationship -> Sex 
 		    HoursPerWeek -> Sex
 		    Relationship -> HoursPerWeek
[…]

### 2025-02-10 code/fig3/make_plots.R

@@ -164,7 +164,7 @@ make_plots_v2 <- function(filename, plot_type){
 	      	       c(d_shd$ges_mean, d_shd$hc_mean, d_shd$pc_mean, d_shd$human_mean),
 	      	       c(d_shd$ges_lower, d_shd$hc_lower, d_shd$pc_lower, d_shd$human_lower),
 	      	       c(d_shd$ges_upper, d_shd$hc_upper, d_shd$pc_upper, d_shd$human_upper),
-		       c(rep('GES', nrow(d_shd)), rep('Hill-Climb Search', nrow(d_shd)), rep('PC', nrow(d_shd)), rep('Expert', nrow(d_shd))),
+		       c(rep('GES', nrow(d_shd)), rep('Hill-Climb', nrow(d_shd)), rep('PC', nrow(d_shd)), rep('Expert', nrow(d_shd))),
 		       c(rep('ges', nrow(d_shd)), rep('hc', nrow(d_shd)), rep('pc', nrow(d_shd)), paste('human', d_shd$oracle_acc)),
 		       c(rep(NA, nrow(d_shd)), rep(NA, nrow(d_shd)), rep(NA, nrow(d_shd)), d_shd$oracle_acc)
 		       ))
@@ -181,9 +181,9 @@ make_plots_v2 <- function(filename, plot_type){
 				   ymin=lower, ymax=upper, 
 				   group=grp, color=algo, 
 				   fill=algo,
-				   linetype=ifelse(algo %in% c('GES', 'PC', 'Hill-Climb Search'), NA, "solid"))) + 
+				   linetype=ifelse(algo %in% c('GES', 'PC', 'Hill-Climb'), NA, "solid"))) + 
 		geom_line(alpha=0.4, show.legend=T) +
-		geom_point(data=subset(d_shd, !(algo %in% c('GES', 'PC', 'Hill-Climb Search'))), pch=19, size=1, alpha=0.4, show.legend=F) +
+		geom_point(data=subset(d_shd, !(algo %in% c('GES', 'PC', 'Hill-Climb'))), pch=19, size=1, alpha=0.4, show.legend=F) +
 		geom_ribbon(linetype=0, alpha=0.4, show.legend=F) + 
 		geom_dl(aes(label=line_labels), color='black', method=list(dl.trans(x=x+.1), "last.points", cex=0.5)) + 
       		theme_minimal(base_size = 8) + 
[…]

### 2025-02-10 code/fig4/make_plots.R

@@ -1,4 +1,5 @@
 library(ggplot2)
+library(latex2exp)
 
 
 d_expert <- read.csv('results/unexplained_effect.txt', header=F)
@@ -18,6 +19,7 @@ d_ges$algo <- 'GES'
 
 d_results <- rbind(d_expert, d_ges)
 d_results <- d_results[d_results$x < 31, ]
+d_results$ll <- d_results$ll/10000
 
 p_unexplained_effect <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
 	geom_line(alpha=0.6) +
@@ -26,16 +28,21 @@ p_unexplained_effect <- ggplot(d_results, aes(x=x, y=pillai, color=algo)) +
 	labs(x="Iteration No.") +
 	labs(y="Total Residual Association") + 
 	theme(legend.position='top',
-	      legend.title=element_blank())
+	      legend.title=element_blank(),
+	      legend.text=element_text(size=8)
+	)
 
+browser()
 p_ll <- ggplot(d_results, aes(x=x, y=ll, color=algo)) +
 	geom_line(alpha=0.6) +
 	geom_point(alpha=0.6, show.legend=F) +
 	theme_minimal(base_size=8) +
 	labs(x="Iteration No.") +
-	labs(y="Log-Likelihood") +
+	labs(y=TeX("Log-Likelihood $ (\\times 1e^{4}) $")) +
 	theme(legend.position='top',
-	      legend.title=element_blank())
+	      legend.title=element_blank(),
+	      legend.text=element_text(size=8)
+	)
 
 ggsave('unexplained_effect.pdf', p_unexplained_effect, height=2.0, width=1.6, units='in')
 ggsave('ll.pdf', p_ll, height=2.0, width=1.6, units='in')

### 2025-02-10 paper/main.tex
Fixes

@@
 we only have one undirected edge between \emph{Marital Status} and
-\emph{Relationship} that we orient as $ \textrm{Marital Status} \rightarrow
-\textrm{Relationship} $. We can now test the implied CIs of this DAG using
-data. Figure~\ref{fig:ci_table} shows the result of this testing on some of the
-implied CIs. If we use a p-value threshold of $ 0.05 $, the first $ 3 $ implied
-CIs hold in the data whereas the remaining tests fail. One of the ways to fix
-these failing tests is to add an edge between the variables $ X $ and $ Y $.
-While this method tells up which variable pairs need to have an edge between
-them, it does not tell anything about the orientation of the edge, and we would
-need to orient it based on our domain knowledge.
+\emph{Relationship} that we orient as \emph{Marital Status} $ \rightarrow $
+\emph{Relationship} . We can then test the implied CIs of the DAG using our
+dataset. Figure~\ref{fig:ci_table} shows the result of this testing on some of
+the implied CIs. If we use a p-value threshold of $ 0.05 $, the first $ 2 $
+implied CIs hold in the data whereas the remaining tests fail. One of the ways
+to fix these failing tests is to add an edge between the variable pair $ X $
+and $ Y $. While this method helps us in finding variable pairs where we need
+to add an edge, it does not give us any information about the orientation of
+this edge. We would have to rely on domain knowledge to determine this
+orientation.
 

### 2025-02-10 paper/main.tex
udpates

@@
 \paragraph{Both $ X $ and $ Y $ are continuous: }
-When both $ X $ and $ Y $ are continuous, Pearson's correlation test can be
-used and Pearson's correlation coefficient can be used at the effect size. When
-$ \bm{Z} = \emptyset $, the correlation coefficient is defined as: $ r_{X, Y} =
-\mathrm{cov}(X, Y) / (\sigma_X \sigma_Y) $. When $ \bm{Z} \neq \emptyset $,
-partial correlation coefficient can be used. This is estimated by fitting two
-regression models $ E_X: X \sim \bm{Z} $ and $ E_Y: Y \sim \bm{Z} $,
-calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $ R_Y = Y - E_Y(\bm{Y})
-$, and then computing the correlation coefficient between the residuals: $r_{X,
-Y \rvert \bm{Z}} = r_{R_X, R_Y}$
+When both $ X $ and $ Y $ are continuous, we can use a (partial) correlation
+test for CI testing and Pearson's correlation coefficient can be used at the
+effect size. When $ \bm{Z} = \emptyset $, the correlation coefficient is
+defined as: $ r_{X, Y} = \mathrm{cov}(X, Y) / (\sigma_X \sigma_Y) $. When $
+\bm{Z} \neq \emptyset $, partial correlation coefficient can be used. This is
+estimated by fitting two regression models $ E_X: X \sim \bm{Z} $ and $ E_Y: Y
+\sim \bm{Z} $, calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $ R_Y =
+Y - E_Y(\bm{Y}) $, and then computing the correlation coefficient between the
+residuals: $r_{X, Y \rvert \bm{Z}} = r_{R_X, R_Y}$
 
@@
 estimate the covariance matrix while maximizing the likelihood of the dataset.
-Using the estimated covariance matrix, $ \Sigma $ we can perform the CI test
-and compute the Pearson's correlation coefficient as the effect size. When 
+Using the estimated covariance matrix, $ \Sigma $ we can perform a correlation
+test for CI and compute the Pearson's correlation coefficient as the effect
+size. When 
 $\bm{Z} = \emptyset $, $ r_{X, Y} = \Sigma_{XY} / (\sqrt{\Sigma_{XX} \Sigma_{YY}}) $, 

### 2025-02-10 paper/main.tex
Typos

@@
 	knowledge already starts outperforming purely data-driven structure learning if 
-	the orientation of new edge is correctly determined in at least one out of two cases.
+	the orientation of new edge is correctly determined in at least two out of three cases.
 	We present a proof-of-concept implementation using a large language 
@@
 		attributes such as Age). The choice of algorithm and hyperparameters
-		significantly aa=ffect the output, making it
+		significantly affects the output, making it
 		difficult to assess reliability. Additionally, the absence of
@@
 		combination of directed and undirected edges. This structural uncertainty
-		can make difficult or impossible to apply the learned model for downstream
+		can make it difficult or impossible to apply the learned model for downstream
 		tasks, such as identification or causal
@@
  We then use this information
-along with domain expertise to determine which changes.
+along with domain expertise to determine which changes to make.
 The orientation of the edge when adding is determined based on the domain

### 2025-02-10 paper/main.tex
Minor changes

@@
 covariance between variables $ X $ and $ Y $ as $ \mathrm{cov}(X, Y) $, the
-covariance matrix as $ \Sigma $, and the entry corresponding to covariance
-between $ X $ and $ Y $ using $ \Sigma_{XY} $. In the context of this paper, we
+covariance matrix as $ \Sigma $, and the entry corresponding to the covariance
+between $ X $ and $ Y $ using $ \Sigma_{XY} $. In this paper, we
 consider causal DAGs that contain only observed variables. The structure of the
@@
 corresponding to the random variables, and $ E $ are the edges denoting the
-causal relationship between the variables. The set of parents of $ X $ in $ G $
-are denoted as $ \textrm{Pa}_G(X) $, and its ancestors and descendants as $
-\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $ respectively. We denote a path, $
+causal relationship between the variables. The set of \emph{parents} of $ X $ in $ G $
+are denoted as $ \textrm{Pa}_G(X) $, and its \emph{ancestors} and \emph{descendants} as $
+\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $ respectively. We denote a \emph{path}, $
 \pi(X, Y) = \{ X, V_0, V_1, \cdots, V_k, Y \} $ between $ X $ and $ Y $ in $ G
 $ where each consecutive pair of variables in $ \pi $ are connected through an
-edge. We say $ X $ and $ Y $ are d-connected in $ G $ given a conditioning set
+edge. We say $ X $ and $ Y $ are \emph{d-connected} in $ G $ given a conditioning set
 $ \bm{Z} \subseteq V - \{X, Y\} $, if for a path $ \pi(X, Y) $ (i) for every
@@
 \leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
-d-separated given $ \bm{Z} $ if no d-connecting path exist between them.
+\emph{d-separated} given $ \bm{Z} $ if no d-connecting path exist between them.
 

### 2025-02-10 paper/main.tex
Improves testing section

@@
 Our structure learning algorithm uses the idea of testing DAGs using CI tests.
-In this section, we give an example of how this model testing works in
+In this section, we show an example of how this model testing works in
 practice. We take the example of the MEC learned by the PC algorithm in
@@
 \emph{Relationship} that we orient as \emph{Marital Status} $ \rightarrow $
-\emph{Relationship} . We can then test the implied CIs of the DAG using our
-dataset. Figure~\ref{fig:ci_table} shows the result of this testing on some of
-the implied CIs. If we use a p-value threshold of $ 0.05 $, the first $ 2 $
-implied CIs hold in the data whereas the remaining tests fail. One of the ways
-to fix these failing tests is to add an edge between the variable pair $ X $
-and $ Y $. While this method helps us in finding variable pairs where we need
-to add an edge, it does not give us any information about the orientation of
-this edge. We would have to rely on domain knowledge to determine this
-orientation.
+\emph{Relationship}. We can then test each of the implied CIs of the DAG
+against our dataset using a statistical test. Figure~\ref{fig:ci_table} shows
+some of the implied CIs of the model and the corresponding p-value of
+statistical test. Using a p-value threshold of $ 0.05 $, we can determine that
+the first $2$ implied CIs hold in the data while the remaining two do not.
+These failing tests tell us that there is residual association between these
+variable pairs that is currently not being explained by the DAG. This method,
+however, does not give us any information on the reason, such as there is a
+direct or indirect effect, or a confounding variable, behind this observed
+association. To fix these failing tests, we still have to rely on domain
+knowledge to determine the type of relationship and the causal directions
+between these variables.
 

### 2025-02-10 paper/main.tex
tightening language in (2)

--- 2025-02-10 paper/main.tex
@@
 
-\def\ci{\perp\!\!\!\!\!\perp}
+\def\ci{\perp\!\!\!\!\perp}
 
@@
 	knowledge already starts outperforming purely data-driven structure learning if 
-	the orientation of new edge is correctly determined in at least two out of three cases.
+	the orientation of new edge is correctly determined in at least one out of two cases.
 	We present a proof-of-concept implementation using a large language 
@@
 		attributes such as Age). The choice of algorithm and hyperparameters
-		significantly affects the output, making it
+		significantly aa=ffect the output, making it
 		difficult to assess reliability. Additionally, the absence of
@@
 		combination of directed and undirected edges. This structural uncertainty
-		can make it difficult or impossible to apply the learned model for downstream
+		can make difficult or impossible to apply the learned model for downstream
 		tasks, such as identification or causal
@@
  We then use this information
-along with domain expertise to determine which changes to make.
+along with domain expertise to determine which changes.
 The orientation of the edge when adding is determined based on the domain
@@
 \label{sec:background}
-We denote random variables using uppercase letters, such as $ X $, and a set of
+We denote random variables using uppercase letters like $X$, and a set of
 random variables by $ \bm{X} = \{X_1, \cdots, X_k\} $, where $ \rvert \bm{X}
-\rvert = k $ indicates the number of variables in the set. We denote the
-covariance between variables $ X $ and $ Y $ as $ \mathrm{cov}(X, Y) $, the
-covariance matrix as $ \Sigma $, and the entry corresponding to covariance
-between $ X $ and $ Y $ using $ \Sigma_{XY} $. In the context of this paper, we
-consider causal DAGs that contain only observed variables. The structure of the
-DAG is represented using a graph $ G = (V, E) $, where $ V $ are the nodes
-corresponding to the random variables, and $ E $ are the edges denoting the
-causal relationship between the variables. The set of parents of $ X $ in $ G $
-are denoted as $ \textrm{Pa}_G(X) $, and its ancestors and descendants as $
-\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $ respectively. We denote a path, $
+\rvert = k $ is the number of variables in the set. We denote the
+covariance between variables $ X $ and $ Y $ as $ \mathrm{cov}(X, Y) $, a 
+covariance matrix as $ \Sigma $, and the entry corresponding to the covariance
+between $ X $ and $ Y $ as $ \Sigma_{XY} $. We consider causal DAGs that contain 
+only observed variables. A DAG $ G = (V, E) $ is an acyclic directed graph
+whose nodes $ V $ corresponding to random variables and whose edges $ E $ represent
+direct causal relationships \cite{Pearl2009}. For an edge $e=X \to Y$, 
+$\textrm{reverse}(e)=Y \to X$. The set of parents of $ X $ in $ G $
+is denoted as $ \textrm{Pa}_G(X) $, and its ancestors and descendants as $
+\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $, respectively; we use the convention
+that $X \notin \textrm{An}_G(X)$ and $X \notin \textrm{De}_G(X)$. We denote a path, $
 \pi(X, Y) = \{ X, V_0, V_1, \cdots, V_k, Y \} $ between $ X $ and $ Y $ in $ G
-$ where each consecutive pair of variables in $ \pi $ are connected through an
+$ where consecutive pairs of variables in $ \pi $ are connected by an
 edge. We say $ X $ and $ Y $ are d-connected in $ G $ given a conditioning set
@@
 
-\subsection{Testing DAG Using CI Tests}
-Our structure learning algorithm uses the idea of testing DAGs using CI tests.
-In this section, we give an example of how this model testing works in
-practice. We take the example of the MEC learned by the PC algorithm in
-Figure~\ref{fig:intro}(b) using $ N=800 $. We first orient all the undirected
-edges in the model based on our domain knowledge to obtain a DAG. In this case
-we only have one undirected edge between \emph{Marital Status} and
-\emph{Relationship} that we orient as \emph{Marital Status} $ \rightarrow $
-\emph{Relationship} . We can then test the implied CIs of the DAG using our
-dataset. Figure~\ref{fig:ci_table} shows the result of this testing on some of
-the implied CIs. If we use a p-value threshold of $ 0.05 $, the first $ 2 $
+\subsection{Conditional Independence Tests}
+\label{sec:ci_tests}
+Our structure learning algorithm uses the idea of testing DAGs using CI statements
+like $ X \ci Y \rvert \bm{Z} $ (with
+potentially $ \bm{Z} = \emptyset $).
+For example, take the MEC learned by the PC algorithm in
+Figure~\ref{fig:intro}(b) using $ N=800 $, and suppose that we orient 
+the undirected edge between \emph{Marital Status} and
+\emph{Relationship}  as \emph{Marital Status} $ \rightarrow $
+\emph{Relationship}. We can then read implied CIs of the DAG using $d$-separation,
+and test them in our dataset. Figure~\ref{fig:ci_table} shows the result of some tests. 
+Using a significance threshold  $\alpha=0.05 $, the first $ 2 $
 implied CIs hold in the data whereas the remaining tests fail. One of the ways
@@
 to add an edge, it does not give us any information about the orientation of
-this edge. We would have to rely on domain knowledge to determine this
+this edge. In this paper, we use domain knowledge to determine this
 orientation.
@@
 
-\subsection{Effect Sizes for CI Tests}
-\label{sec:ci_tests}
-
-Our structure learning approach relies on using CI tests and a measure of
-conditional association between variables -- also known as partial association.
-We are interested in testing CI tests: $ X \ci Y \rvert \bm{Z} $ (with
-potentially $ \bm{Z} = \emptyset $) in data and quantifying the strength of the
-relationship between $ X $ and $ Y $ conditioned on $ \bm{Z} $. For the
-quantification we use effect size measures based on the CI test. In this
-section, we give an overview of some of the CI tests and effect size measures
+In addition to determining sigificance, we are interested in quantifying the 
+\emph{strength} of any partial association between $X$ and $Y$ after conditioning 
+on $\mathbf{Z}$. This metric depends on the type of CI test used. In the following,
+we briefly overview some CI tests and effect size measures
 for different types of (possibly mixed) data.
@@
 When both $ X $ and $ Y $ are continuous, we can use a (partial) correlation
-test for CI testing and Pearson's correlation coefficient can be used at the
+test for CI testing and Pearson's correlation coefficient can be used as the
 effect size. When $ \bm{Z} = \emptyset $, the correlation coefficient is
 defined as: $ r_{X, Y} = \mathrm{cov}(X, Y) / (\sigma_X \sigma_Y) $. When $
-\bm{Z} \neq \emptyset $, partial correlation coefficient can be used. This is
-estimated by fitting two regression models $ E_X: X \sim \bm{Z} $ and $ E_Y: Y
+\bm{Z} \neq \emptyset $, the partial correlation coefficient can be used instead.
+This is estimated by fitting two regression models $ E_X: X \sim \bm{Z} $ and $ E_Y: Y
 \sim \bm{Z} $, calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $ R_Y =
@@
 
-Polyserial (when one is ordinal and the other is continuous) and Polychoric
-(when both are ordinal) correlation have been used to estimate the covariance
-matrix between them \citep{Poon1987}. Both methods make the assumption that the
+Polyserial (for continuous $Y$) and polychoric
+(ordinal $Y$) correlations are used to estimate correlations 
+involving ordinal variables \citep{Poon1987}. Both methods make the assumption that the
 observed ordinal variable is a result of thresholding a latent normally
-distributed continuous variable. Under this assumption, the methods then try to
-estimate the covariance matrix while maximizing the likelihood of the dataset.
+distributed continuous variable. Under this assumption, the methods then 
+estimate the threshold values and covariance matrix using maximum likelihood.
 Using the estimated covariance matrix, $ \Sigma $ we can perform a correlation
 test for CI and compute the Pearson's correlation coefficient as the effect
-size. When 
+size in the same way as for continuous $ X, Y $ and $ \bm{Z} $, i.e., 
 $\bm{Z} = \emptyset $, $ r_{X, Y} = \Sigma_{XY} / (\sqrt{\Sigma_{XX} \Sigma_{YY}}) $, 
@@
 
-\paragraph{All $ X $, $ Y $, and $ \bm{Z} $ are discrete (ordinal or categorical): }
+\paragraph{$ X $, $ Y $, and $ \bm{Z} $ are all discrete (ordinal or categorical): }
 
-In the case when we have a combination of ordinal and categorical variables, we
-can use a residualization based CI test \citep{Ankan2023}. This test gives a
-chi-square distributed test statistic, that can be used for computing an effect
-size such as the Root Mean Squared Error of Approximation (RMSEA). Given a
-chi-squared test statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
-freedom, the RMSEA is defined as: $ \textrm{RMSEA}_{X, Y \rvert \bm{Z}} =
-\sqrt{\chi^2 - \textrm{df}/ (\textrm{df} (N-1))} $, where $ N $ is the sample
-size.
+For combinations of ordinal and categorical variables, we
+can use a residualization-based CI test \citep{Ankan2023} that returns a 
+chi-square distributed test statistic. Given the statistic, $ \chi^2 $, with $ \textrm{df} $ degrees of
+freedom, we use the Root Mean Squared Error of Approximation (RMSEA) defined as
+ $ \textrm{RMSEA}_{X, Y \rvert \bm{Z}} =
+\sqrt{\max(0,\chi^2 - \textrm{df})/ (\textrm{df} (N-1))} $, where $ N $ is the sample
+size. This effect size can be used for any statistical test with a chi-square 
+distributed test statistic.
 

### 2025-02-10 paper/main.tex
edits section 3

@@
 insights in a manner that is based on the following considerations: 
-(1) A domain experts is possibly good at determining causal directions between variables if 
-there is a clear causal direction between them. (2) A domain expert may be less good
+(1) A domain expert is possibly good at determining causal directions between variables if 
+there \emph{is} a clear causal direction between them. (2) A domain expert may be less good
 at identifying cases where there is no causal relationship between the variables,
@@
 
-We therefore consider the following two models of a domain experts: Given two variables
+We consider two models of a domain experts. Given two variables
 $ X $ and $ Y $ that are part of an unknown causal DAG $G$, an \emph{strong ancestral oracle} 
@@
  X \gets Y & \textrm{if } Y \in \textrm{An}_G(X) \\
- \textrm{rand}(X \gets Y, X \to Y) & \textrm{otherwise} \\
+ \textrm{rand}(\}X \gets Y, X \to Y\}) & \textrm{otherwise} \\
 \end{cases}
 $$
-
+where $\textrm{rand}(S)$ denotes a uniformly chosen element of $S$.
 That is, our ancestral oracles only provide information on the direction of cause-effect
 relationships, and do not distinguish between direct and indirect effects. 
-The weak ancestral oracle is even unable to detect cases where there is no cause-effect
-relationship between two variables. Using such oracles for DAG construction therefore 
-inherently risks that superfluous edges are created. In our approach, such superfluous edges
-will be detected and removed by utilizing the data.
+The weak ancestral oracle is even unable to detect the absence of cause-effect
+relationships between two variables. Using such oracles for DAG construction  
+inherently risks that superfluous edges are created.
 
-To model how our learning procedure interacts with the data, we assume that we
-additionally have access to a second oracle $\mathcal{D}_G$ than can answer d-separation
+Our approach uses the data to decide where to add edges and which edges 
+are superfluous and can be removed.
+As a useful abstraction, let us assume that we
+have access to a second oracle $\mathcal{D}_G$ than can answer d-separation
 queries with respect to the unknown target graph:
 $\mathcal{D}_G(X,Y,\mathbf{Z})=1$ iff $X$ and $Y$ are $d$-separated by $\mathbf{Z}$ in 
-$G$. These are the standard oracles considered in constraint-based structure learning
-algorithms, such as the PC algorithm \cite{Spirtes2001}.
+$G$. These are the standard oracles considered in constraint-based structure
+learning algorithms, such as the PC algorithm \cite{Spirtes2001}.
 
-Based on these two oracles, we now define two core procedures that we will use to 
+Based on these two oracles, we now define two core procedures that
 iteratively change a current DAG structure. The procedure 
-\textsc{Expand} (Algorithm~\ref{algo:expand}) uses conditional independence
-information to search for missing edges in the graph and then uses domain
-knowledge to orient them. The procedure \textsc{Prune} (Algorithm~\ref{algo:prune}) uses
-conditional independence information to remove superfluous edges from the
+\textsc{Expand} (Algorithm~\ref{algo:expand}) uses CI  
+information (supplied in this case by the d-separation oracle) 
+to search for unexplained associations in the graph and 
+then uses domain knowledge to determine ancestry. The 
+procedure \textsc{Prune} (Algorithm~\ref{algo:prune}) uses
+CI information to remove superfluous edges from the
 graph.
@@
 
-The function $\textsc{Expand}$ takes an initial list of edges and searches for 
-any vertex pairs that are not connected but where the d-separation oracle 
-indicates  that there is still a residual association not explained by the other variables.
-An optional parameter $B$, the empty set by default, allows to specify a ``black list''
-of vertex pairs that must not be connected, and another optional parameter $k$
-specifies the maximum amount of edges to be added by this procedure.
+$\textsc{Expand}$ takes an initial list of edges and searches for 
+any unconnected vertex pairs that are not connected but where the
+d-separation oracle indicates a residual association not explained 
+by other paths.
+The parameter $B$ specifies a ``black list''
+of vertex pairs that must not be connected, 
+whereas $k$
+can be used to limit the maximum amount of edges 
+to be added by this procedure; the usefulness of the $B$ and $k$
+parameters will become clear soon.
 
-Let $G^+$ denote the transitive closure of $G$, that is, the graph in which 
-there is a direct edge $X \to Y$ whenever $X \in \textrm{An}_G(Y)$. The transitive
-closure is related to the result of $\textsc{Expand}$ by the following two 
-easily proved facts.
+Let $G^+$ denote the transitive closure of $G$ (the graph in which 
+there is a direct edge $X \to Y$ whenever $X \in \textrm{An}_G(Y)$). 
+The transitive closure is related to the result of $\textsc{Expand}$ 
+by the following two easily proved facts.
 
@@
 \begin{proof}
-Every pair of vertices $X, Y$ that are $d$-connected in $G$ are now connected by an edge 
+Every pair of vertices $X, Y$ that are $d$-connected 
+in $G$ are now connected by an edge 
 during $\textsc{Expand}$. For $X \in \textrm{An}_G(Y)$, the edge is 
-$X \to Y$ just like when using a strong ancestral oracle; since only correctly oriented
-edges are added between ancestor-descendant pairs, such additions never create cycles.
-For $X, Y$ that are only $d$-connected throug common ancestors, an arbitrary 
-edge is added in a manner that preserves acyclicity. The result is a supergraph
-of $G^+$.
+$X \to Y$ like when using a strong ancestral oracle; 
+since only correctly oriented
+edges are added between ancestor-descendant pairs, 
+such additions never create cycles.
+For $X, Y$ that are only $d$-connected throug common ancestors, 
+an arbitrary edge is added in a manner that preserves acyclicity.
+The result is a supergraph of $G^+$.
 \end{proof}
@@
 
-Unlike our ancestral oracles, the pruning operation is quite effective at distinguishing
+Unlike our ancestral oracles, the pruning operation is 
+quite effective at distinguishing
 direct and indirect effects, as shown by the following result.
@@
 \begin{proof}
-For an edge $X \to Y$ in the skeleton of $G$, $\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$ regardless 
-of $\mathbf{Z}$, so all edges in the skeleton are retained after {\sc Prune}. Conversely, if 
- $X \to Y$ is not in the skeleton of $G$, let $\mathbf{Z}$ be the $d$-separating set chosen 
-in line~5 of {\sc Prune} (there is at least one such $\mathbf{Z}$, the union of the parents of 
-$X$ and $Y$). Since  $\mathbf{Z}$ $d$-separates all paths from $X$ to $Y$ in $G'$, it does the same 
-in $G$ which contains a subset of these paths. Therefore, all edges not in the skeleton of $G$ are 
+For an edge $X \to Y$ in the skeleton of $G$, 
+$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$ regardless 
+of $\mathbf{Z}$, so all edges in the skeleton are 
+retained after {\sc Prune}. Conversely, if 
+ $X \to Y$ is not in the skeleton of $G$, let 
+$\mathbf{Z}$ be the $d$-separating set chosen 
+in line~5 of {\sc Prune} (there is at least one 
+such $\mathbf{Z}$, the union of the parents of 
+$X$ and $Y$). Since  $\mathbf{Z}$ $d$-separates 
+all paths from $X$ to $Y$ in $G'$, it does the same 
+in $G$ which contains a subset of these paths. 
+Therefore, all edges not in the skeleton of $G$ are 
 removed.
@@
 
-Since we require expert knowledge only in the $\textsc{Expand}$ operation, we may try
-to be more economical by asking fewer questions at a time and interleaving expansion and
-pruning steps. This leads us to the following, more incremental DAG construction algorithm.
+Since we require expert knowledge only in the $\textsc{Expand}$ 
+operation, we may try to be more economical by asking fewer 
+questions at a time and interleaving expansion and
+pruning steps. This leads us to the following, more iterative 
+DAG construction algorithm.
 
@@
 \begin{proof}
-The loop in Algorithm~\ref{algo:expert} terminates if and only if $E=E^*$. If the loop does not 
-terminate, a new edge has been added in line 6, and/or one or more edges were pruned in line 7; the pruned 
-edges never include the edge added in line 6 in the same iteration.
-Every edge can be added at 
-most once and pruned at most once. Therefore, the algorithm always terminates
-after at most $|V|(|V|-1)+1$ iterations of the loop.
-For every edge $e=X\to Y $ in the skeleton of $G$, $\mathcal{D}_G(X,Y,\mathbf{Z})=0$ irrespective of $\mathbf{Z}$, and $e$ must be added to $E$ in some iteration, and can never be pruned afterwards.
-Therefore, after some iteration, $(V,E)$ must be a supergraph of $G$ after executing line 6, and this will be pruned to the real graph $G$ in line 7 (Proposition~\ref{prop:prune}).
-In the next iteration, no further changes are made, and the loop terminates.
+The loop in Algorithm~\ref{algo:expert} terminates if and 
+only if $E=E^*$. If the loop does not  terminate, a new
+ edge has been added in line 6, and/or one or more edges were 
+pruned in line 7; the pruned edges never include the edge added 
+in line 6 in the same iteration. Every edge can be added at 
+most once and pruned at most once. Therefore, the algorithm 
+always terminates after at most $|V|(|V|-1)+1$ iterations of
+the loop. For every edge $e=X\to Y $ in the skeleton of $G$, 
+$\mathcal{D}_G(X,Y,\mathbf{Z})=0$ irrespective of $\mathbf{Z}$, 
+and $e$ must be added to $E$ in some iteration, and can never 
+be pruned afterwards. Therefore, after some iteration, 
+$(V,E)$ must be a supergraph of $G$ after executing line~6, 
+and this will be pruned to the real graph $G$ in line~7 
+(Proposition~\ref{prop:prune}). In the next iteration, 
+no further changes are made, and the loop terminates.
+\label{thm:itworks}
 \end{proof}
@@
 	\end{subfigure}
-	\caption{An example showing each iteration of the \textsc{ExpertInLoop} algorithm. The dashed edges show all the potential edges that \textsc{Expand} step could add in each iteration.}
+	\caption{An example showing each iteration of \textsc{ExpertInLoop}. The dashed edges show all the potential edges that \textsc{Expand} could add in each iteration.}
 \end{figure}
@@
 
-The \textsc{ExpertInLoop} algorithm adds a new edge in each iteration. The edge
-to be added is selected using the \textsc{Expand} algorithm, which returns a
-potential edge at random based on the iteration order. However, if the residual
-association between the selected variables is low, adding the edge may result
-in only a marginal improvement in the overall model fit. Given that the
+\textsc{ExpertInLoop} adds a new edge in each iteration. The edge
+to be added is selected using the \textsc{Expand} algorithm, which 
+returns a potential edge at random based on the iteration order. 
+However, if the residual association between the selected variables 
+is low, adding the edge may result in only a marginal improvement 
+in the overall model fit. Given that the
 algorithm requires expert intervention in each iteration to specify the edge
-orientation, it is beneficial to prioritize edges that contribute the most to
-improving the model.
+orientation, it can be beneficial to prioritize edges that contribute 
+the most to improving the model.
 
@@
 potential edges based on their residual association given the current DAG. This
-residual association can be quantified using the effect size measures from the
-CI tests used for querying the d-separation in Algorithm~\ref{algo:expand}.
-Specifically, for a d-separation query $ D_G(X, Y, \mathbf{Z}) $, we quantify
+residual association can be quantified using the effect sizes from the
+CI tests used for deciding d-separation in Algorithm~\ref{algo:expand}.
+Specifically, for a d-separation query $D_G(X, Y, \mathbf{Z}) $, we quantify
 the residual association, $ \phi(X, Y, \bm{Z}) $, as the effect size of the CI
-test $ X \ci Y \rvert \bm{Z} $. In Algorithm~\ref{algo:ranked_expand}, we
-outline a greedy approach to edge selection using the effect size measure.
+test $X \ci Y \rvert \bm{Z} $.
 
-To implement Algorithm~\ref{algo:ranked_expand}, we need to have an effect size
-metric by which we can rank effects. Depending on the type of variables, we can
-use the effect size measures shown in Section~\ref{sec:ci_tests}. The algorithm
-selects the edge that has the highest absolute residual association, resulting
+Algorithm~\ref{algo:ranked_expand} shows a 
+version of {\sc Expand} where we use the largest unexplained association to
+determine where to next add an edge. 
+To implement this, we need an effect size
+metric by which we can rank effects. Depending on the type of variables, 
+we can use the effect size measures shown in Section~\ref{sec:ci_tests}. 
+The algorithm selects the edge that has the highest absolute residual 
+association, resulting
 in prioritization of edges that contribute the most to improving the model.
 
-\subsection{Connection to Score-Based Methods}
+\subsection{Comparison to Score-Based Methods}
 
-Our procedure shares similarities with score-based automated causal discovery
-methods, such as Greedy Equivalence Search (GES), where the algorithm
-iteratively adds or removes edges that maximize improvements in a specified
-scoring metric. Given a scoring metric, --total residual association, $ \tau $ for 
-a DAG $ G = (V, E) $, defined as:
+Superficially, the data-driven aspects of our procedure appear
+ similar to score-based automated causal discovery methods like 
+Greedy Equivalence Search (GES), which 
+iteratively add or remove edges that maximize improvements in a specified
+scoring metric. Indeed, we can define the \emph{total residual association}
+ $ \tau $ for a DAG $ G = (V, E) $ as:
 
@@
 
-our \textsc{RankedExpand} approach behaves similarly to GES. In each iteration,
-we add the edge between the variable pair that results in the lowest $ \tau $
-and remove edges that have no impact on $ \tau $.
-
-While both methods aim to make iterative modifications to the DAG, a key
-distinction lies in the nature of the evaluation criteria: unlike standard
-scoring metrics, $ \tau $ is not decomposable. Decomposibility is a key
-property of scoring metrics, allowing them to be expressed as the sum of
-individual variables in the model. This property ensures that adding or
-removing an edge only affects a localized part of the model. In contrast, $
-\tau $ can not be decomposed, as modifying an edge in one part of the model can
-affect the residual associations in other parts of the model.
+Our \textsc{RankedExpand} approach behaves similarly to GES in that it 
+tries to prioritize modifications that lead to the largest improvements 
+in $\tau$. However, adding an edge between the two vertices that contribute
+the most to $\tau$ does not necessarily lead to the largest 
+possible decrease in $\tau$ because the new edge can also affect other 
+residual associations. In other words, $\tau$ is not a \emph{decomposable}
+metric that can be expressed as a sum over individual variables in the model
+and their parents.
 
@@
 quantify how well the model explains the data in an absolute sense. In
-contrast, $ \tau $ provides an absolute measure of model fit -- its value
-approaches $ 0 $ as the model perfectly explains the observed data.
+contrast, $ \tau $ could be seen as an absolute measure of model fit -- 
+its value approaches $ 0 $ as the model perfectly explains the observed data.
 

### 2025-02-10 code/fig3/sl.R
Fixes oracle to return ancestral relations

@@ -15,10 +15,10 @@ PVALUE_THRESHOLD <- 0.05
 # Oracle function. Returns either true or (random or don't know) direction of the edge.
 oracle <- function(u, v, true_dag, accuracy=1){
 	if (runif(1) <= accuracy){
-		if (v %in% dagitty::children(true_dag, u)){
+		if (v %in% dagitty::descendants(true_dag, u)){
 			return(c(u, v))
 		}
-		else if (u %in% dagitty::children(true_dag, v)){
+		else if (u %in% dagitty::descendants(true_dag, v)){
 			return(c(v, u))
 		}
 		else{

### 2025-02-10 code/fig3/sl.R
Adds the edge reversal

@@ -87,8 +87,14 @@ simulate_human_sl <- function(sim_data, true_dag, oracle_acc, max_iter=1e4){
 					new_dag <- temp_dag
 				}
 				else{
-					print("Not adding edge to avoid cycle.")
-					blacklist_edges <- rbind(blacklist_edges, oracle_edge)
+					temp_dag <- add_edge(new_dag, c(oracle_edge[2], oracle_edge[1]))
+					if (dagitty::isAcyclic(temp_dag)){
+						new_dag <- temp_dag
+					}
+					else{
+						print("Not adding edge to avoid cycle.")
+						blacklist_edges <- rbind(blacklist_edges, oracle_edge)
+					}
 				}
 			}
 		}

### 2025-02-10 paper/main.tex

@@
 metric by which we can rank effects. Depending on the type of variables, 
-we can use the effect size measures shown in Section~\ref{sec:ci_tests}. 
+we can use the effect sizes shown in Section~\ref{sec:ci_tests}. 
 The algorithm selects the edge that has the highest absolute residual 
@@
 understand the behaviour of our method when the experts are not perfect. To
-simulate this imperfect expert, we use a version of the strong ancestral oracle
+simulate an imperfect domain expert, we use a version of the strong ancestral oracle
 that has an accuracy of $ \alpha $:
-
 \begin{equation}
@@
 \end{equation}
-
 where $G_{\textrm{true}}$ is the true data generating DAG.
 
-An important point to note here is that the effective accuracy of $
+It is important to note that the effective accuracy of $
 \mathrm{Expert} $ is higher than $ \alpha $ as even when giving a random
@@
 use the \textsc{RankedExpand} method with Pearson's correlation coefficient as
-the effect size measure (Section~\ref{sec:ci_tests}). For the analysis, we
-simulated data from known DAGs and use the algorithms to recover the original
+the effect size (Section~\ref{sec:ci_tests}) and significance threshold
+$\alpha=0.05$; note that this means that also the d-separation ``oracle''
+makes both type~I and type~II errors. For the analysis, we
+simulate data from known DAGs and use the algorithms to recover the original
 DAG. We start by generating a random DAG on $10$ nodes and use linear models
 with random effects to simulate $ 500 $ samples from it. We simulate each
-variable $ V_i $ in the DAG as:
-
-\begin{equation}
-	\begin{split}
-		\bm{\beta} &= \mathrm{Uniform}([-0.6, 0.6]) \\
-		V_i &= \bm{\beta} \mathrm{Pa}_G(V_i) + \mathcal{N}(0, 1)
-	\end{split}
-\end{equation}
+variable $ V_i $ in the DAG as
+$$
+	V_i = \bm{\beta} \cdot \mathrm{Pa}_G(V_i) + \mathcal{N}(0, 1)
+$$
+where each coefficient $\beta_i$ of the vector $\bm{\beta} \in \mathbb{R}^{|\bm{\beta} |}$ is randomly drawn as
+$$
+\beta_i \sim \mathrm{Uniform}([-0.6, 0.6])
+$$
 
@@
 new. Common ways to provide such background knowledge are "whitelists" or "blacklists"
-of edges that must or must not be present in the resulting DAG, or specification
-of a temporal structure between the variables. 
+of edges 
+\citep{scutari2010}, or specification
+of a temporal structure between the variables \cite{Bang2023}.
 Compared to such approaches, where domain knowledge 
-is specified and supplied beforehand, the novelty of our method lies in the interactive
-back-and-forth between domain knowledge and data, which provably leads to the correct
+is specified and supplied beforehand, the novelty 
+of our method lies in the interactive
+back-and-forth between domain knowledge and data, 
+which provably leads to the correct
 result in a polynomial amount of steps (if the expert answers all

### 2025-02-10 paper/main.tex
UPdates

@@
 \leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
-\emph{d-separated} given $ \bm{Z} $ if no d-connecting path exist between them.
+d-separated given $ \bm{Z} $ if no d-connecting path exist between them.
 
@@
 	\centering
-	\begin{tabular}{llrr}
+	\begin{tabular}{llrl}
 		$X$ & $Y$ & $ \bm{Z} $ & p-value \\
@@
  X \gets Y & \textrm{if } Y \in \textrm{An}_G(X) \\
- \textrm{rand}(\}X \gets Y, X \to Y\}) & \textrm{otherwise} \\
+ \textrm{rand}(\{X \gets Y, X \to Y\}) & \textrm{otherwise} \\
 \end{cases}
@@
 			\mathcal{O}(X, Y) &= \begin{cases}
-				X \rightarrow Y, & \textrm{if } X \rightarrow Y \in G_{\textrm{true}} \\
-				Y \rightarrow X, & \textrm{if } Y \rightarrow X \in G_{\textrm{true}} \\
+				X \rightarrow Y, & \textrm{if } X \in \textrm{An}_{G_{\textrm{true}}}(Y) \\
+				X \leftarrow Y, & \textrm{if } Y \in \textrm{An}_{G_{\textrm{true}}}(X) \\
 				\textrm{None}, & \textrm{otherwise }
@@
 
-It is important to note that the effective accuracy of $
-\mathrm{Expert} $ is higher than $ \alpha $ as even when giving a random
-answer, there is a $ 1 $ in $ 3 $ chance that the answer is correct. Therefore,
-an $ \mathrm{Expert} $ with accuracy $ \alpha $, has an effective accuracy, $ \alpha_{\mathrm{eff}} = \alpha + (1 - \alpha) / 3 $.
+It is important to note that the effective accuracy of $ \mathrm{Expert} $ is
+higher than $ \alpha $ as even when $ x > \alpha $, there is a $ 1 $ in
+$ 3 $ chance that edge orientation is correct. Therefore, an $ \mathrm{Expert} $ with
+accuracy $ \alpha $, has an effective accuracy, $ \alpha_{\mathrm{eff}} =
+\alpha + (1 - \alpha) / 3 $.
 
@@
 comparable to the automated algorithms for $ \alpha = 0.3
-(\alpha_{\textrm{eff}} = 0.53 $, implying that if an expert is able to get the
+(\alpha_{\textrm{eff}} = 0.53 $), implying that if an expert is able to get the
 edge orientation correctly in more than $ 1 $ out of $ 2 $ cases, they can

### 2025-02-10 code/fig3/make_plots.R
Updated plots

@@ -258,5 +258,5 @@ make_plots_v2 <- function(filename, plot_type){
 	ggsave("sid_ribbon.pdf", p_sid, height=2, width=3.37, units='in')
 }
 
-make_plots_v2(filename='results/sl_results_linear.csv', plot_type='ribbon')
+make_plots_v2(filename='results/sl_results_lin_reverse.csv', plot_type='ribbon')
 

### 2025-02-11 paper/main.tex
final edits

@@
 	\tau = \sum_{\substack{X, Y \in V \\ X \rightarrow Y, Y \rightarrow X \not \in E}}   \phi(X, Y, \mathrm{pa}_G(X) \cup \mathrm{pa}_G(Y))
+\label{eqn:tau}
 \end{equation}
@@
 That is, they indicate which model is better for a given dataset but do not
-quantify how well the model explains the data in an absolute sense. In
+quantify how well the model explains the data. In
 contrast, $ \tau $ could be seen as an absolute measure of model fit -- 
@@
 causal discovery algorithms. Our goal with the empirical analysis is to
-understand the behaviour of our method when the experts are not perfect. To
-simulate an imperfect domain expert, we use a version of the strong ancestral oracle
-that has an accuracy of $ \alpha $:
-\begin{equation}
+understand the behaviour of our method when both experts are not perfect.
+Specifically, how good does the domain expert have to be so we actually benefit
+from their expertise, compared to a purely data-driven approach?
+
+In our analysis, we simulate data from a ``true'' DAG $G$ and use the
+{\sc ExpertInLoop} algorithm with the {\sc RankedExpand} heuristic 
+to recover $G$. We implement the $d$-separation oracle $\mathcal{D}_G(X,Y,\mathbf{Z})$
+by conducting conditional independence tests of $X \ci Y \mid \mathbf{Z}$, which on finite
+data inevitably make type~I and type~II errors. To
+simulate an imperfect domain expert, we use a version of a strong 
+ancestral oracle $\mathcal{A}_G$ (Section~\ref{sec:modification})
+that with probability $\alpha$ knows the correct answer and randomly
+guesses otherwise, i.e.,
+\begin{equation*}
 	\begin{split}
@@
 		\mathrm{Expert}(\alpha) &= \begin{cases} 
-			\mathcal{O}(X, Y),  & \textrm{if  } x <= \alpha \\
+			\mathcal{A}_G(X, Y),  & \textrm{if  } x <= \alpha \\
 			\textrm{rand}(X \rightarrow Y, Y \leftarrow X, \textrm{None}) & \textrm{otherwise} \\
-				\end{cases} \\
-			\mathcal{O}(X, Y) &= \begin{cases}
-				X \rightarrow Y, & \textrm{if } X \in \textrm{An}_{G_{\textrm{true}}}(Y) \\
-				X \leftarrow Y, & \textrm{if } Y \in \textrm{An}_{G_{\textrm{true}}}(X) \\
-				\textrm{None}, & \textrm{otherwise }
 				\end{cases} 
 	\end{split}
-\end{equation}
-where $G_{\textrm{true}}$ is the true data generating DAG.
+\end{equation*}
 
 It is important to note that the effective accuracy of $ \mathrm{Expert} $ is
-higher than $ \alpha $ as even when $ x > \alpha $, there is a $ 1 $ in
-$ 3 $ chance that edge orientation is correct. Therefore, an $ \mathrm{Expert} $ with
+higher than $ \alpha $ as even when $ x > \alpha $, there is a $ 1/3 $
+that edge orientation is correct. Therefore, an $ \mathrm{Expert} $ with
 accuracy $ \alpha $, has an effective accuracy, $ \alpha_{\mathrm{eff}} =
@@
 
-We analyzed the performance of \textsc{ExpertInLoop} algorithm by comparing it
-with three other automated algorithms: PC, Hill-Climb Search, and Greedy
-Equivalence Search (GES) on linear Gaussian data. For \textsc{ExpertInLoop}, we
+We compare the performance of the \textsc{ExpertInLoop} algorithm by
+to three other automated algorithms: PC \citep{Spirtes2001,KalischB07}, 
+Hill-Climb Search \citep{scutari2010}, and Greedy Equivalence Search (GES) \citep{Chickering2002}
+on linear-Gaussian data. For \textsc{ExpertInLoop}, we
 use the \textsc{RankedExpand} method with Pearson's correlation coefficient as
 the effect size (Section~\ref{sec:ci_tests}) and significance threshold
-$\alpha=0.05$; note that this means that also the d-separation ``oracle''
-makes both type~I and type~II errors. For the analysis, we
-simulate data from known DAGs and use the algorithms to recover the original
-DAG. We start by generating a random DAG on $10$ nodes and use linear models
+(and type~I error rate) $\alpha=0.05$. We start by generating a random DAG 
+on $10$ nodes and use linear models
 with random effects to simulate $ 500 $ samples from it. We simulate each
@@
 $$
-where each coefficient $\beta_i$ of the vector $\bm{\beta} \in \mathbb{R}^{|\bm{\beta} |}$ is randomly drawn as
+where each coefficient $\beta_i$ of the vector $\bm{\beta} \in \mathbb{R}^{|\mathrm{Pa}_G(V_i)|}$ is randomly drawn as
 $$
@@
 
-We perform the experiment for both varying densities of the DAG and accuracy of
-$\mathrm{Expert}$. We also repeat each experiment $ 30 $ times and report the
+We perform the experiment for varying densities of the DAG and accuracies of
+$\mathrm{Expert}$. We repeat each experiment $ 30 $ times and report the
 mean and standard error of the results.
 
-To compare the learned DAG to the original DAG, we use two metrics: Structural
-Hamming Distance (SHD) and Structural Intervention Distance (SID)
-\citep{Peters2015}. Unlike \textrm{ExpertInLoop}, the automated algorithms can
+To compare the learned DAG to the original DAG, we use two metrics: the Structural
+Hamming Distance (SHD) and the Structural Intervention Distance (SID)
+\citep{Peters2015}. Unlike \textsc{ExpertInLoop}, the automated algorithms can
 only recover the MEC, therefore, we considered all possible orientations of the
@@
 Figure~\ref{fig:shd_sid}. For SHD, the performance of \textsc{ExpertInLoop} is
-comparable to the automated algorithms for $ \alpha = 0.3
-(\alpha_{\textrm{eff}} = 0.53 $), implying that if an expert is able to get the
+comparable to the automated algorithms for $ \alpha = 0.3$
+($\alpha_{\textrm{eff}} = 0.53 $), implying that if an expert is able to get the
 edge orientation correctly in more than $ 1 $ out of $ 2 $ cases, they can
@@
 truth is unknown, two measures of fit are shown over 30 iterative modifications: 
-total residual association $\tau$ (a, lower is bettter); log-likelihood (a, higher is better).
+total residual association $\tau$ (a, see Equation~\ref{eqn:tau}; lower is bettter); log-likelihood (b, higher is better).
 	}
@@
 and it contains mixed types of data; here, we use a version that contains
-ordinal and nominal variables and use the RMSEA as a measure of association
+ordinal and nominal variables and use the RMSEA of a residualization-based
+conditional independence test \citep{Ankan2023} as a measure of association
 to prioritize modifications (Section~\ref{sec:background}).
@@
 		(Gemini 1.5 Flash) as the expert. The p-value threshold used is $ 0.05 $ 
-		and the measure of association threshold is $ 0. 1 $}
+		and the measure of association threshold is $ 0. 1 $, meaning
+		that associations below $0.1$ are not considered by 
+		\textsc{RankedExpand}.}
 	\label{fig:adult_llm}
@@
 
-Since our approach relies on expert knowledge to determine edge orientations,
+Since our approach relies on expert knowledge to determine ancestral relationships,
 we explored the potential of using an LLM for this task. We applied our causal
@@
 that the model fit improves comparably to a greedy approach and 
-the resulting DAG contains largely sensible edge directions.
+the resulting DAG appears to contain largely sensible edge directions.
 
@@
 edges; to avoid cluttering, no lines are drawn for residual associations that
-are statistically insignificanct or below the specified threshold.
+are statistically insignificant or below the specified threshold.
 The thickness of these edges represents the strength of association,
@@
 integrates manual construction with data-driven feedback, bridging
-the gap between fully manually and fully automated methods.
+the gap between fully manual and fully automated methods.
 The idea of augmenting structure learning by domain knowledge is of course not 
-new. Common ways to provide such background knowledge are "whitelists" or "blacklists"
-of edges 
+new. Common ways to provide such background knowledge are ``whitelists'' or 
+``blacklists'' of edges 
 \citep{scutari2010}, or specification
-of a temporal structure between the variables \cite{Bang2023}.
+of a ``tiered'' (e.g., temporal) structure between the variables \citep{Bang2023}.
 Compared to such approaches, where domain knowledge 
@@
 which provably leads to the correct
-result in a polynomial amount of steps (if the expert answers all
-ancestral queries correctly).
+result in a polynomial amount of steps in the oracle settings we considered.
 

### 2025-02-11 code/fig3/make_plots.R
UPdates plot and caption

@@ -191,7 +191,7 @@ make_plots_v2 <- function(filename, plot_type){
 		      legend.title=element_blank(),
 		      legend.text=element_text(size=8)) +
 		labs(x = "Edge Probability") +
-		labs(y = "Mean SHD") + 
+		labs(y = "SHD") + 
 		labs(color = "Algorithm") +
 		ylim(0, 40) +
 		guides(linetype='none', fill='none')
@@ -250,7 +250,7 @@ make_plots_v2 <- function(filename, plot_type){
 		      legend.title=element_blank(),
 		      legend.text=element_text(size=8)) +
 		labs(x = "Edge Probability") +
-		labs(y = "Mean SID") + 
+		labs(y = "SID") + 
 		labs(color = "Algorithm") +
 		ylim(0, 90) +
 		guides(linetype='none', fill='none')

### 2025-02-11 paper/main.tex
UPdates plot and caption
Updates

@@
 random variables by $ \bm{X} = \{X_1, \cdots, X_k\} $, where $ \rvert \bm{X}
-\rvert = k $ is the number of variables in the set. We denote the
-covariance between variables $ X $ and $ Y $ as $ \mathrm{cov}(X, Y) $, a 
-covariance matrix as $ \Sigma $, and the entry corresponding to the covariance
-between $ X $ and $ Y $ as $ \Sigma_{XY} $. We consider causal DAGs that contain 
-only observed variables. A DAG $ G = (V, E) $ is an acyclic directed graph
-whose nodes $ V $ corresponding to random variables and whose edges $ E $ represent
-direct causal relationships \cite{Pearl2009}. For an edge $e=X \to Y$, 
-$\textrm{reverse}(e)=Y \to X$. The set of parents of $ X $ in $ G $
-is denoted as $ \textrm{Pa}_G(X) $, and its ancestors and descendants as $
-\textrm{An}_G(X) $ and $ \textrm{De}_G(X) $, respectively; we use the convention
-that $X \notin \textrm{An}_G(X)$ and $X \notin \textrm{De}_G(X)$. We denote a path, $
-\pi(X, Y) = \{ X, V_0, V_1, \cdots, V_k, Y \} $ between $ X $ and $ Y $ in $ G
-$ where consecutive pairs of variables in $ \pi $ are connected by an
-edge. We say $ X $ and $ Y $ are d-connected in $ G $ given a conditioning set
-$ \bm{Z} \subseteq V - \{X, Y\} $, if for a path $ \pi(X, Y) $ (i) for every
-collider structure ($ V_1 \rightarrow V_2 \leftarrow V_3 $) on $ \pi(X, Y) $, $
-\bm{Z} \cap \{ V_2, \textrm{De}_G(V_2) \} \ne \emptyset $, and (ii) for every
-non-collider structure ($ V_1 \rightarrow V_2 \rightarrow V_3 $, $ V_1
-\leftarrow V_2 \rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are
-d-separated given $ \bm{Z} $ if no d-connecting path exist between them.
+\rvert = k $ is the number of variables in the set. We denote the standard
+deviation of $ X $ as $ \sigma_X $, the covariance between variables $ X $ and
+$ Y $ as $ \mathrm{cov}(X, Y) $, a covariance matrix as $ \Sigma $, and the
+entry corresponding to the covariance between $ X $ and $ Y $ as $ \Sigma_{XY}
+$. We consider causal DAGs that contain only observed variables. A DAG $ G =
+(V, E) $ is an acyclic directed graph whose nodes $ V $ corresponding to random
+variables and whose edges $ E $ represent direct causal relationships
+\cite{Pearl2009}. For an edge $e=X \to Y$, $\textrm{reverse}(e)=Y \to X$. The
+set of parents of $ X $ in $ G $ is denoted as $ \textrm{Pa}_G(X) $, and its
+ancestors and descendants as $ \textrm{An}_G(X) $ and $ \textrm{De}_G(X) $,
+respectively; we use the convention that $X \notin \textrm{An}_G(X)$ and $X
+\notin \textrm{De}_G(X)$. We denote a path, $ \pi(X, Y) = \{ X, V_0, V_1,
+\cdots, V_k, Y \} $ between $ X $ and $ Y $ in $ G $ where consecutive pairs of
+variables in $ \pi $ are connected by an edge. We say $ X $ and $ Y $ are
+d-connected in $ G $ given a conditioning set $ \bm{Z} \subseteq V - \{X, Y\}
+$, if for a path $ \pi(X, Y) $ (i) for every collider structure ($ V_1
+\rightarrow V_2 \leftarrow V_3 $) on $ \pi(X, Y) $, $ \bm{Z} \cap \{ V_2,
+\textrm{De}_G(V_2) \} \ne \emptyset $, and (ii) for every non-collider
+structure ($ V_1 \rightarrow V_2 \rightarrow V_3 $, $ V_1 \leftarrow V_2
+\rightarrow V_3 $), $ V_2 \not \in \bm{Z} $. $ X $ and $ Y $ are d-separated
+given $ \bm{Z} $ if no d-connecting path exist between them.
 
@@
 \sim \bm{Z} $, calculating the residuals $ R_X = X - E_X(\bm{Z}) $ and $ R_Y =
-Y - E_Y(\bm{Y}) $, and then computing the correlation coefficient between the
+Y - E_Y(\bm{Y}) $, and computing the Pearson's correlation coefficient between the
 residuals: $r_{X, Y \rvert \bm{Z}} = r_{R_X, R_Y}$
@@
 		\textsc{ExpertInLoop} with varying values of expert accuracy, $ \alpha = \{0.1, 0.3, 0.5, 0.7, 0.9\} $. The corresponding
-		$\alpha_{\textrm{eff}} $ is shown in the plot.}
+	$\alpha_{\textrm{eff}} $ is shown in the plot. Expert shading: mean $\pm$ standard error. Others shading: [mean min. - standard error min., mean max. + standard error max.]}
 	\label{fig:shd_sid}

### 2025-02-11 paper/main.tex
Updates

@@
 	\end{tabular}
-	\caption{Results of testing some of implied CIs of the DAG in
-		 Figure~\ref{fig:intro}(b) using the same CI test that was used for
-		 learning it.}
+	\caption{Results of testing some of the implied CIs of the DAG in
+		 Figure~\ref{fig:intro}(b) using the same residualization based
+	 	 CI test that was used for learning it.}
 	\label{fig:ci_table}

### 2025-02-11 paper/main.tex
UPdates

@@
 test for CI and compute the Pearson's correlation coefficient as the effect
-size in the same way as for continuous $ X, Y $ and $ \bm{Z} $, i.e., 
+size (same as for continuous $ X $ and $ Y $), i.e., 
 $\bm{Z} = \emptyset $, $ r_{X, Y} = \Sigma_{XY} / (\sqrt{\Sigma_{XX} \Sigma_{YY}}) $, 

### 2025-04-10 web/api.R
cd7v

@@ -1,6 +1,7 @@
 library(jsonlite)
 
 preprocess_adult <- function(d){
+	print(d)
 	d$X <- NULL
 	d$Fnlwgt <- NULL
 	d$EducationNum <- NULL

### 2025-05-26 code/fig3/make_plots.R
Updates make plot

@@ -258,5 +258,5 @@ make_plots_v2 <- function(filename, plot_type){
 	ggsave("sid_ribbon.pdf", p_sid, height=2, width=3.37, units='in')
 }
 
-make_plots_v2(filename='results/sl_results_lin_reverse.csv', plot_type='ribbon')
+make_plots_v2(filename='results/sl_results_mixed.csv', plot_type='ribbon')
 

### 2025-06-05 paper/main.tex
cd7v
fixes theory section

@@
-\documentclass{uai2025} % for initial submission
+\documentclass[accepted]{uai2025} % for initial submission
 %\documentclass[accepted]{uai2025} % after acceptance, for a revised version; 
@@
 We consider two models of a domain experts. Given two variables
-$ X $ and $ Y $ that are part of an unknown causal DAG $G$, an \emph{strong ancestral oracle} 
+$ X $ and $ Y $ that are part of a directed graph $G$, an \emph{strong ancestral oracle} 
 $\mathcal{A}_G$ is defined as:
@@
 \end{cases}$$
+Note that the ancestral oracle does not consider differences between direct and 
+indirect relationships: for any $G, H$ where $G^+=H^+$, we have $\mathcal{A}_G=\mathcal{A}_H$.
 
-whereas a \emph{weak ancestral oracle} is defined as:
-$$
-\mathcal{A}_G(X,Y)=\begin{cases}
- X \to Y & \textrm{if } X \in \textrm{An}_G(Y) \\
- X \gets Y & \textrm{if } Y \in \textrm{An}_G(X) \\
- \textrm{rand}(\{X \gets Y, X \to Y\}) & \textrm{otherwise} \\
-\end{cases}
-$$
-where $\textrm{rand}(S)$ denotes a uniformly chosen element of $S$.
-That is, our ancestral oracles only provide information on the direction of cause-effect
-relationships, and do not distinguish between direct and indirect effects. 
-The weak ancestral oracle is even unable to detect the absence of cause-effect
-relationships between two variables. Using such oracles for DAG construction  
-inherently risks that superfluous edges are created.
+Experts can make mistakes. In our theoretical analysis, we will model two different types
+of mistakes that the algorithm can recover from.
+
+\begin{definition}
+Let $G=(V,E_G)$ be a DAG let and $H=(V,E_H)$ be a directed graph (possibly cyclic) 
+such that $E_G \subseteq E_H$. Then $\mathcal{A}_H$
+is called a \emph{$G$-consistent ancestral oracle}. Further, 
+if $H=(V,E_H)$ is a supergraph of $G$ that contains at least one cycle, then 
+$\mathcal{A}_H$ is called a \emph{$G$-consistent ancestral oracle} with respect to $G$.
+\end{definition}
+
+These two oracles still provide correct information on all ancestral relationships that are 
+present in $G$, but cannot reliably detect the \emph{absence} of ancestral relationships.
+Further, the inconsistent ancestral oracle sometimes contradicts itself: there are at least two
+variables $X,Y$ where its information implies both a causal effect of $X$ on $Y$ and a reverse 
+causal effect of $Y$ on $X$. Therefore, if we use such oracles for DAG construction, we will
+introduce superfluous edges into our graph.
 
@@
 $\mathcal{D}_G(X,Y,\mathbf{Z})=1$ iff $X$ and $Y$ are $d$-separated by $\mathbf{Z}$ in 
-$G$. These are the standard oracles considered in constraint-based structure
+$G$, and $0$ otherwise. These are the standard oracles considered in constraint-based structure
 learning algorithms, such as the PC algorithm \cite{Spirtes2001}.
@@
 	    $e \gets \mathcal{A}_G(X, Y)$ \;
-	    \uIf{ $(V,E \cup L \cup e)$ is acyclic }{
             $L \gets L \cup  \{ e \} $\;
-	    } \Else {
-            	$L \gets L \cup  \{ \textrm{reverse}(e) \} $\;
         }
-        }
-	\If{$|L|\geq k$}{\KwGoTo 12\;}
+	\If{$|L|\geq k$}{\KwGoTo 13\;}
     }
@@
 
+The following is immediately obtained as a corollary.
+
 \begin{proposition}
 For a conditional independence oracle
- $\mathcal{D}_G$ and a weak ancestral oracle $\mathcal{A}_G$, 
-$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_G, \emptyset, \infty)
-\supseteq G^+$.
+ $\mathcal{D}_G$ and a $G$-consistent ancestral oracle $\mathcal{A}_H$, 
+$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)
+=H^+ \supseteq G^+$.
 \label{prop:weakexpand}
 \end{proposition}
-
-\begin{proof}
-Every pair of vertices $X, Y$ that are $d$-connected 
-in $G$ are now connected by an edge 
-during $\textsc{Expand}$. For $X \in \textrm{An}_G(Y)$, the edge is 
-$X \to Y$ like when using a strong ancestral oracle; 
-since only correctly oriented
-edges are added between ancestor-descendant pairs, 
-such additions never create cycles.
-For $X, Y$ that are only $d$-connected throug common ancestors, 
-an arbitrary edge is added in a manner that preserves acyclicity.
-The result is a supergraph of $G^+$.
-\end{proof}
 
@@
 \begin{theorem}
-Given a d-separation oracle $\mathcal{D}_G$ and a strong or weak 
-ancestral oracle $\mathcal{A}_G$, 
-$\textsc{Prune}(\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_G,\emptyset,\infty),\mathcal{D}_G)=G.$
+Given a d-separation oracle $\mathcal{D}_G$ and a $G$-consistent 
+ancestral oracle $\mathcal{A}_H$, 
+$\textsc{Prune}(\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_H,\emptyset,\infty),\mathcal{D}_G)=G.$
 \end{theorem}

### 2025-06-08 code/fig3/sl.R
Deal with cycles

@@ -88,11 +88,12 @@ simulate_human_sl <- function(sim_data, true_dag, oracle_acc, max_iter=1e4){
 					for (i in 1:(length(cycle)-1)){
 						edge = c(cycle[i], cycle[i+1])
 						if ((edge[1] == oracle_edge[1]) & (edge[2] == oracle_edge[2])){
-							invisible()	
+							invisible()
 						}
 						else{
-							Z = setdiff(cycle, c(edge[1], edge[2]))
-							p.value <- dagitty::ciTest(X=edge[1], Y=edge[2], Z=Z, sim_data, type='cis.pillai')
+							adj_set <- setdiff(parents(temp_dag, edge[2]), edge[1])
+							Z = setdiff(cycle, c(edge[1], edge[2]))[1] # Take any variable on the cycle
+							p.value <- dagitty::ciTest(X=edge[1], Y=edge[2], Z=union(adj_set, Z), sim_data, type='cis.pillai')
 							if (p.value["p.value"] > PVALUE_THRESHOLD){
 								temp_dag <- remove_edges(temp_dag, as.data.frame(list(X=edge[1], A='->', Y=edge[2])))
 								new_dag <- temp_dag
@@ -269,7 +270,7 @@ run_sim <- function(R, n_nodes, edge_probs, oracle_accs){
 		print("Done with PC")
 
 		for (oracle_acc in oracle_accs){
-			human_dist <- t(replicate(R, run_single_exp_human(n_nodes=n_nodes, edge_prob=edge_prob, oracle_acc=oracle_acc)))
+			human_dist <- t(future_replicate(R, run_single_exp_human(n_nodes=n_nodes, edge_prob=edge_prob, oracle_acc=oracle_acc)))
 
 			human_mean <- apply(human_dist, mean, MARGIN=2)
 			human_sd <- apply(human_dist, sd, MARGIN=2)/sqrt(R)

### 2025-06-08 paper/main.tex

--- 2025-02-11 paper/main.tex
@@
 	    $e \gets \mathcal{A}_G(X, Y)$ \;
-	    \uIf{ $(V,E \cup L \cup e)$ is acyclic }{
-            	$L \gets L \cup  \{ e \} $\;
-	    } \Else {
-            	$L \gets L \cup  \{ \textrm{reverse}(e) \} $\;
+	    $L \gets L \cup \{ e \} $ \;
+	    \If{ $ (V, E \cup L) $ is cyclic }{
+		    $ L \gets \textsc{FixCycles}(V, E, L, e, \mathcal{D}_G) $ \;
 	    }
+	    % \If{ $(V,E \cup L \cup e)$ is acyclic }{
+            % 	$L \gets L \cup  \{ e \} $\;
+	    % } \Else {
+            % 	$L \gets L \cup  \{ \textrm{reverse}(e) \} $\;
+	    % }
         }
-	\If{$|L|\geq k$}{\KwGoTo 12\;}
+	\If{$|L|\geq k$}{\KwGoTo 16\;}
     }
@@
 \label{algo:expand}
+\end{algorithm}
+
+\begin{algorithm}[h]
+\DontPrintSemicolon
+\SetAlgoLined
+\SetKwFunction{FixCycle}{FixCycle}
+\SetKw{KwGoTo}{go to}
+\SetKwProg{Fn}{Function}{:}{}
+\Fn{{\sc FixCycles}($V, E, L, e, \mathcal{D}_G $)}{
+	$ u, v \gets e $ \;
+	\ForEach {$ P = \{ v \rightarrow x_1, x_1 \rightarrow x_2, \cdots,  x_{k-1} \rightarrow x_k, x_k \rightarrow u \} $ in ($V, E \cup L$) }{
+		\ForEach {$ x_i \rightarrow x_j $ in $ P $}{
+        	let $\mathbf{Z}$ be a set that $d$-separates $x_i$ and $x_j$ in $(V, E \cup L)$\;
+		\If {$ \mathcal{D}_G(X, Y, \mathbf{Z}) = 0 $}{
+			$ L \gets L - \{ x_i \rightarrow x_j \} $ \;
+
+		}
+	}
+	}
+	\Return $ L $ \;
+}
+\caption{Fixing cycles by removing incorrect edges.}
+\label{algo:fix_cycles}
 \end{algorithm}
@@
 such additions never create cycles.
-For $X, Y$ that are only $d$-connected throug common ancestors, 
+For $X, Y$ that are only $d$-connected through common ancestors, 
 an arbitrary edge is added in a manner that preserves acyclicity.
@@
     remove all edges in $L$ from $G$\;
-    \Return{$(V,E)$}\;
+    \Return{$(V, E)$}\;
 }

### 2025-06-09 paper/main.tex
Updates paper
Add example with cycles
changes

--- 2025-06-05 paper/main.tex
@@
 Let $G=(V,E_G)$ be a DAG let and $H=(V,E_H)$ be a directed graph (possibly cyclic) 
-such that $E_G \subseteq E_H$. Then $\mathcal{A}_H$
-is called a \emph{$G$-consistent ancestral oracle}. Further, 
-if $H=(V,E_H)$ is a supergraph of $G$ that contains at least one cycle, then 
-$\mathcal{A}_H$ is called a \emph{$G$-consistent ancestral oracle} with respect to $G$.
+such that $E_{G^+} \subseteq E_{H^+}$. Then $\mathcal{A}_H$
+is called a \emph{$G$-compatible ancestral oracle}. Further, 
+if $H=(V,E_H)$ is an acyclic supergraph of $G$, then 
+$\mathcal{A}_H$ is called a \emph{acyclic $G$-compatible ancestral oracle} with respect to $G$.
 \end{definition}
@@
 	    $e \gets \mathcal{A}_G(X, Y)$ \;
-            $L \gets L \cup  \{ e \} $\;
+	    $L \gets L \cup \{ e \} $ \;
+	    \If{ $ (V, E \cup L) $ is cyclic }{
+		    $ L \gets \textsc{FixCycles}(V, E, L, e, \mathcal{D}_G) $ \;
 	    }
-	\If{$|L|\geq k$}{\KwGoTo 13\;}
+	    % \If{ $(V,E \cup L \cup e)$ is acyclic }{
+            % 	$L \gets L \cup  \{ e \} $\;
+	    % } \Else {
+            % 	$L \gets L \cup  \{ \textrm{reverse}(e) \} $\;
+	    % }
+        }
+	\If{$|L|\geq k$}{\KwGoTo 16\;}
     }
@@
 For a conditional independence oracle
- $\mathcal{D}_G$ and a $G$-consistent ancestral oracle $\mathcal{A}_H$, 
+ $\mathcal{D}_G$ and an acyclic $G$-compatible ancestral oracle $\mathcal{A}_H$, 
 $\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)

### 2025-06-10 paper/main.tex
definitions and pseudocode

@@
 \begin{abstract}
-	Numerous causal discovery algorithms were developed to automatically learn
-	directed acyclic graphs (DAGs) and other causal models from data. However,
-	their adoption in applied domains remains limited, as researchers often
-	prefer to construct DAGs manually based on domain knowledge. This
-	preference arises due to several practical challenges with automated
-	algorithms, such as their tendency to produce results that contradict
-	obvious domain knowledge and their inability to distinguish Markov equivalent
-	models. To assist researchers in constructing DAGs manually, we propose an iterative
-	structure learning approach that combines domain knowledge with
-	data-driven insights. Our method leverages conditional independence
+	Many researchers construct directed acyclic graph (DAG) models manually
+	based on domain knowledge. 
+	Although numerous causal discovery algorithms were developed to automatically learn
+	DAGs and other causal models from data, these remain challenging to 
+	use due to their tendency to produce results that contradict
+	domain knowledge, among other issues. Here we propose a hybrid, 
+	iterative structure learning approach that combines domain knowledge with
+	data-driven insights to assist researches in constructing DAGs.
+	Our method leverages conditional independence
 	testing to iteratively identify variable pairs where an edge is
@@
 	assist human experts with DAG construction.
+\todo[inline]{Should we add links here to the implementations? Is the two out of three cases still correct?}
 \end{abstract}
@@
 decisions. Causal discovery methods aim to discover such relationships among
-random variables using observational data. In the DAG literature, the primary
-focus has been on developing automated algorithms to learn causal structures
-from datasets. These efforts have led to numerous causal discovery algorithms,
-such as constraint-based methods like PC algorithm \citep{Spirtes2001,KalischB07} 
+random variables using observational data. These
+include constraint-based methods like PC algorithm \citep{Spirtes2001,KalischB07} 
 and Fast Causal Inference \citep{Spirtes2000}), score-based methods such as Hill-Climb
@@
 optimization-based methods like NOTEARS \citep{Zheng2018} and DAGMA
-\citep{Bello2022}. Despite significant progress in automated causal discovery,
-their adoption in applied research has been limited. Challenges encountered with existing
+\citep{Bello2022}. Despite this significant body of work,
+the adoption of causal discovery methods in observational research has so far 
+been limited. Challenges encountered with existing
 causal discovery algorithms in practice include but are not limited to:
@@
 		further reduce the confidence in their outputs. A recent paper
-		therefore advised Epidemiologists to not attempt using structure
+		 advised Epidemiologists to not attempt using structure
 		learning algorithms without help of an expert \citep{Gururaghavendran_2024}.
@@
 entry corresponding to the covariance between $ X $ and $ Y $ as $ \Sigma_{XY}
-$. We consider causal DAGs that contain only observed variables. A DAG $ G =
+$. A DAG $ G =
 (V, E) $ is an acyclic directed graph whose nodes $ V $ corresponding to random
@@
 given $ \bm{Z} $ if no d-connecting path exist between them.
+
+\subsection{Assumptions} 
+In this paper, we consider the structure learning problem under the widely
+known \emph{causal Markov}, \emph{causal sufficiency} and \emph{faithfulness}
+assumptions \citep{Spirtes1993}. These entail that there exists a 
+DAG $G$ on the variables $\bm{X}$ whose implied $d$-separation statements coincide
+exactly with the conditional independence relationships among the variables 
+in $\bm{X}$, and for which all direct common causes of variables in $\bm{X}$ are 
+also included in $\bm{X}$. These are the same assumptions made by the structure
+learning algorithms we consider in this paper, such as the PC algorithm \citep{Spirtes1993}.
 
@@
 
-We consider two models of a domain experts. Given two variables
-$ X $ and $ Y $ that are part of a directed graph $G$, an \emph{strong ancestral oracle} 
-$\mathcal{A}_G$ is defined as:
+We model domain experts as procedures that take two variables $X$ and $Y$ that are assumed 
+to be part of a DAG $G$ and 
+provide information on the ancestral relationship between them. First, 
+a \emph{strong ancestral oracle} $\mathcal{A}_G$ is defined as:
 $$\mathcal{A}_G(X,Y)=\begin{cases}
@@
 present in $G$, but cannot reliably detect the \emph{absence} of ancestral relationships.
-Further, the inconsistent ancestral oracle sometimes contradicts itself: there are at least two
-variables $X,Y$ where its information implies both a causal effect of $X$ on $Y$ and a reverse 
-causal effect of $Y$ on $X$. Therefore, if we use such oracles for DAG construction, we will
+We even allow our oracles to contradict themselves by giving responses that would lead to a 
+cyclic graph when combined. If we use such oracles for DAG construction, we will
 introduce superfluous edges into our graph.
@@
 $G$, and $0$ otherwise. These are the standard oracles considered in constraint-based structure
-learning algorithms, such as the PC algorithm \cite{Spirtes2001}.
+learning algorithms, such as the PC algorithm \citep{Spirtes2001}.
 
-Based on these two oracles, we now define two core procedures that
-iteratively change a current DAG structure. The procedure 
+We now define two core procedures that use ancestral and $d$-separation queries
+to iteratively change a current DAG structure. The procedure 
 \textsc{Expand} (Algorithm~\ref{algo:expand}) uses CI  
-information (supplied in this case by the d-separation oracle) 
-to search for unexplained associations in the graph and 
+information to search for unexplained associations in the graph and 
 then uses domain knowledge to determine ancestry. The 
@@
 	    \If{ $ (V, E \cup L) $ is cyclic }{
-		    $ L \gets \textsc{FixCycles}(V, E, L, e, \mathcal{D}_G) $ \;
+		    $ L' \gets \textsc{FixCycles}(V, E \cup L, \mathcal{D}_G) $ \;
+		    $ B \gets B \cup (L \setminus L') $ \;
+		    $ L \gets L'$
 	    }
@@
         }
-	\If{$|L|\geq k$}{\KwGoTo 16\;}
+	\If{$|L|\geq k$}{\KwGoTo 18\;}
     }
     add all edges in $L$ to $E$\;
-    \Return{$(V,E)$}\;
+    \Return{$(V,E,B)$}\;
 }
@@
 \SetKwProg{Fn}{Function}{:}{}
-\Fn{{\sc FixCycles}($V, E, L, e, \mathcal{D}_G $)}{
-	$ u, v \gets e $ \;
-	\ForEach {$ P = \{ v \rightarrow x_1, x_1 \rightarrow x_2, \cdots,  x_{k-1} \rightarrow x_k, x_k \rightarrow u \} $ in ($V, E \cup L$) }{
-		\ForEach {$ x_i \rightarrow x_j $ in $ P $}{
-        	let $\mathbf{Z}$ be a set that $d$-separates $x_i$ and $x_j$ in $(V, E \cup L)$\;
-		\If {$ \mathcal{D}_G(X, Y, \mathbf{Z}) = 0 $}{
-			$ L \gets L - \{ x_i \rightarrow x_j \} $ \;
+\Fn{{\sc FixCycles}($V, E, \mathcal{D}_G $)}{
+		$R \gets \emptyset$ \; 
+		\ForEach {$ x_i \rightarrow x_j $ on a cycle in $ (V,E) $}{
+		\If {there exists a set $\mathbf{Z}\subseteq\mathbf{X}$ where $\mathcal{D}_G(x_i, x_j)=1$
+}{
+			$ R \gets R \cup \{ x_i \rightarrow x_j \} $ \;
 
@@
 	}
-	}
-	\Return $ L $ \;
+	\Return $ L \setminus R $ \;
 }

### 2025-06-10 paper/main.tex
definitions and pseudocode

@@
 $E_p \gets \emptyset$ \tcc{Current edges} 
-$B \gets \emptyset$ \tcc{Edges that were pruned} 
+$B \gets \emptyset$ \tcc{Edges that were pruned or removed from cycles} 
 \Repeat{ $E=E_p$ }{
 	$E \gets E_p$ \;
-	$(V,E) \gets \textsc{Expand}(V,E,\mathcal{D}_G,\mathcal{A}_G,B,1)$ \;
+	$(V,E,B<D-s>) \gets \textsc{Expand}(V,E,\mathcal{D}_G,\mathcal{A}_G,B,1)$ \;
 	$(V,E_p) \gets \textsc{Prune}(V,E,\mathcal{D}_G)$ \;

### 2025-06-10 paper/main.tex
fixing everything cycle related

--- 2025-06-10 paper/main.tex
@@
 such that $E_{G^+} \subseteq E_{H^+}$. Then $\mathcal{A}_H$
-is called a \emph{$G$-compatible ancestral oracle}. Further, 
-if $H=(V,E_H)$ is an acyclic supergraph of $G$, then 
+is called a \emph{$G$-compatible ancestral oracle}. If $H$ is furthermore acyclic, then 
 $\mathcal{A}_H$ is called a \emph{acyclic $G$-compatible ancestral oracle} with respect to $G$.
@@
 		$R \gets \emptyset$ \; 
-		\ForEach {$ x_i \rightarrow x_j $ on a cycle in $ (V,E) $}{
-		\If {there exists a set $\mathbf{Z}\subseteq\mathbf{X}$ where $\mathcal{D}_G(x_i, x_j)=1$
+		\ForEach {$ X \rightarrow Y $ on a cycle in $ (V,E) $}{
+		\If {there exists a set $\mathbf{Z}\subseteq\mathbf{X}$ where $\mathcal{D}_G(X, Y, \mathbf{Z})=1$
 }{
-			$ R \gets R \cup \{ x_i \rightarrow x_j \} $ \;
+			$ R \gets R \cup \{ X \rightarrow Y \} $ \;
 
@@
 The parameter $B$ specifies a ``black list''
-of vertex pairs that must not be connected, 
-whereas $k$
+of vertex pairs that must not be connected, which is important
+to prevent edges that were removed from cycles to be added again 
+and will have another important role in the overall algorithm.
+In addition, $k$
 can be used to limit the maximum amount of edges 
-to be added by this procedure; the usefulness of the $B$ and $k$
-parameters will become clear soon.
+to be added by this procedure; it will become clear soon why
+this is useful.
 
@@
 \end{proposition}
+
+When using experts that generate cyclic graphs, edges are pruned to 
+break these cycles, which can lead to a more sparse result.
+
+\begin{proposition}
+For a conditional independence oracle
+ $\mathcal{D}_G$ and a $G$-compatible ancestral oracle $\mathcal{A}_H$ where
+$H$ contains at least one cycle,
+let $\tilde{G}=\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)$
+Then $H^+ \supsetneq \tilde{G} \supseteq G$. 
+\label{prop:weakexpand:cyclic}
+\end{proposition}
+
+\begin{proof}
+All edges in the skeleton of $G$ are added during $\textsc{Expand}$ with the 
+correct orientation and cannot be removed by $\textsc{FixCycles}$. Therefore, the result
+is a supergraph of $G$. If cycles are created during $\textsc{Expand}$, then every 
+cycle contains at least one edge that is not the skeleton of $G$, otherwise $G$ itself
+would be cyclic. At least one edge of every cycle is therefore removed by
+ $\textsc{FixCycles}$, making the result acyclic again.
+\end{proof}
 
@@
 
-By combining Propositions~\ref{prop:strongexpand}, \ref{prop:weakexpand} and 
+By combining Propositions~\ref{prop:strongexpand}, \ref{prop:weakexpand}, 
+\ref{prop:weakexpand:cyclic} and 
 \ref{prop:prune}, we immediately obtain the following:
@@
 ancestral oracle $\mathcal{A}_H$, 
-$\textsc{Prune}(\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_H,\emptyset,\infty),\mathcal{D}_G)=G.$
+let $(V',E',B)=\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_H,\emptyset,\infty)$.
+Then $\textsc{Prune}(V',E',\mathcal{D}_G)=G.$
 \end{theorem}
@@
 $E_p \gets \emptyset$ \tcc{Current edges} 
-$B \gets \emptyset$ \tcc{Edges that were pruned} 
+$B \gets \emptyset$ \tcc{Edges that were pruned or removed from cycle} 
 \Repeat{ $E=E_p$ }{
 	$E \gets E_p$ \;
-	$(V,E) \gets \textsc{Expand}(V,E,\mathcal{D}_G,\mathcal{A}_G,B,1)$ \;
+	$(V,E,B) \gets \textsc{Expand}(V,E,\mathcal{D}_G,\mathcal{A}_G,B,1)$ \;
 	$(V,E_p) \gets \textsc{Prune}(V,E,\mathcal{D}_G)$ \;
@@
 \end{proof}
+
+We note that the general property of Algorithm~\ref{algo:expert} that 
+each possible edge can be added and removed at most once, which guarantees
+that the algorithm eventually terminates, remains valid even if the oracles make
+arbitrary mistakes. This is crucial when using the algorithm in practice.
 
@@
 in prioritization of edges that contribute the most to improving the model.
+
+In addition to prioritizing edges to add or remove, the ranking could also
+be used to break cycles when CI tests fail to identify the edges that are not
+in the skeleton during the procedure  $\textsc{FixCycle}$, which could happen
+due to finite sample effects. In that case, removing edges that correspond to 
+weak associations may be preferable to keeping the cycles in the model.
 

### 2025-06-11 paper/main.tex
fixes to expand and fixcycles

@@
 include constraint-based methods like PC algorithm \citep{Spirtes2001,KalischB07} 
-and Fast Causal Inference \citep{Spirtes2000}), score-based methods such as Hill-Climb
+and Fast Causal Inference \citep{Spirtes2000}, score-based methods such as Hill-Climb
 Search and Greedy Equivalence Search \citep{Chickering2002}, and continuous
@@
 variables and whose edges $ E $ represent direct causal relationships
-\cite{Pearl2009}. For an edge $e=X \to Y$, $\textrm{reverse}(e)=Y \to X$. The
+\citep{Pearl2009}.  The
 set of parents of $ X $ in $ G $ is denoted as $ \textrm{Pa}_G(X) $, and its
@@
     let $L \gets \{\}$\;
-    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
-	$\{X,Y\} \notin B$}{
+    \ForEach{$X, Y$ where $X \to Y \notin E \cup B $ and $Y \to X \notin E \cup B$}{
         let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
         \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
-	    $e \gets \mathcal{A}_G(X, Y)$ \;
-	    $L \gets L \cup \{ e \} $ \;
-	    \If{ $ (V, E \cup L) $ is cyclic }{
-		    $ L' \gets \textsc{FixCycles}(V, E \cup L, \mathcal{D}_G) $ \;
-		    $ B \gets B \cup (L \setminus L') $ \;
-		    $ L \gets L'$
+	    $L \gets L \cup \mathcal{A}_G(X, Y) $ \;
         }
-	    % \If{ $(V,E \cup L \cup e)$ is acyclic }{
-            % 	$L \gets L \cup  \{ e \} $\;
-	    % } \Else {
-            % 	$L \gets L \cup  \{ \textrm{reverse}(e) \} $\;
-	    % }
+	\If{$|L|\geq k$}{\KwGoTo 12\;}
     }
-	\If{$|L|\geq k$}{\KwGoTo 18\;}
-    }
-    add all edges in $L$ to $E$\;
+    $ R \gets \textsc{FixCycles}(V, E \cup L, \mathcal{D}_G) $ \; 
+    $ B \gets B \cup R \; ; \; L \gets L \setminus R \; ; \;  E \gets (E \cup L) \setminus B$ \;
     \Return{$(V,E,B)$}\;
@@
 \SetAlgoLined
-\SetKwFunction{FixCycle}{FixCycle}
+\SetKwFunction{FixCycles}{FixCycles}
 \SetKw{KwGoTo}{go to}
@@
 }{
-			$ R \gets R \cup \{ X \rightarrow Y \} $ \;
+			$ R \gets R \cup \{ X \to Y \} \cup \{ Y \to X \} $ \;
 
@@
 	}
-	\Return $ L \setminus R $ \;
+	\Return $ R $ \;
 }
@@
     let $\phi_{\max} \gets 0 $\;
-    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
-	$\{X,Y\} \notin B$}{
+    \ForEach{$X, Y$ where $X \to Y \notin E \setminus B$ and $Y \to X \notin E \setminus B$}{
         let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
@@
     }
+    $L' \gets \textsc{FixCycles}(V, E \cup L, \mathcal{D}_G)$ \;
 

### 2025-06-11 paper/main.tex
fixed propositions

@@
 
-The following is immediately obtained as a corollary.
+When using experts that do not always correctly detect the absence of 
+causal relationships, the resulting graph can get larger but also smaller,
+if this leads to the occurrence of cycles that need to be broken.
 
@@
 For a conditional independence oracle
- $\mathcal{D}_G$ and an acyclic $G$-compatible ancestral oracle $\mathcal{A}_H$, 
-$\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)
-=H^+ \supseteq G^+$.
-\label{prop:weakexpand}
-\end{proposition}
-
-When using experts that generate cyclic graphs, edges are pruned to 
-break these cycles, which can lead to a more sparse result.
-
-\begin{proposition}
-For a conditional independence oracle
- $\mathcal{D}_G$ and a $G$-compatible ancestral oracle $\mathcal{A}_H$ where
-$H$ contains at least one cycle,
+ $\mathcal{D}_G$ and a $G$-compatible ancestral oracle $\mathcal{A}_H$, 
 let $\tilde{G}=\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)$
-Then $H^+ \supsetneq \tilde{G} \supseteq G$. 
+Then $H^+ \subseteq \tilde{G} \supseteq G$. If $H$ contains at least one cycle, then
+$H^+ \subsetneq \tilde{G}$.
 \label{prop:weakexpand:cyclic}
@@
 is a supergraph of $G$. If cycles are created during $\textsc{Expand}$, then every 
-cycle contains at least one edge that is not the skeleton of $G$, otherwise $G$ itself
+cycle contains at least one edge that is not in the skeleton of $G$, otherwise $G$ itself
 would be cyclic. At least one edge of every cycle is therefore removed by
- $\textsc{FixCycles}$, making the result acyclic again.
+ $\textsc{FixCycles}$, making the result acyclic again. Since the removed edges need
+to be in $H^+$, the first inclusion is strict for cyclic $H$.
 \end{proof}

### 2025-06-11 paper/main.tex
Decent version of theory part

@@
 
-Experts can make mistakes. In our theoretical analysis, we will model two different types
-of mistakes that the algorithm can recover from.
+Experts can make mistakes. In our analysis, we will consider experts that
+essentially ``make up'' non-existing causal relationships, but do provide 
+correct answers on the directionality of existing ones.
 
 \begin{definition}
-Let $G=(V,E_G)$ be a DAG let and $H=(V,E_H)$ be a directed graph (possibly cyclic) 
-such that $E_{G^+} \subseteq E_{H^+}$. Then $\mathcal{A}_H$
-is called a \emph{$G$-compatible ancestral oracle}. If $H$ is furthermore acyclic, then 
-$\mathcal{A}_H$ is called a \emph{acyclic $G$-compatible ancestral oracle} with respect to $G$.
+Let $G=(V,E_G)$ and $H=(V,E_H)$ be two DAGs. 
+The \emph{$G$-compatible ancestral oracle} $\mathcal{A}_{G\mid H}$ is defined by
+$$\mathcal{A}_{G\mid H}(X,Y)=\begin{cases}
+ X \to Y & \textrm{if } X \in \textrm{An}_G(Y) \\
+ X \gets Y & \textrm{if } Y \in \textrm{An}_G(X) \\
+ \mathcal{A}_{H}(X,Y) & \textrm{otherwise} \\
+\end{cases}$$
+Consider the graph $G\mid H$ containing the edges $ \mathcal{A}_{G\mid H}(X,Y)$ for all 
+pairs $(X, Y) \in V \times V$.   
+If $G\mid H$ is acyclic, then 
+$\mathcal{A}_H$ is called an \emph{acyclic $G$-compatible ancestral oracle}.
 \end{definition}
 
-These two oracles still provide correct information on all ancestral relationships that are 
-present in $G$, but cannot reliably detect the \emph{absence} of ancestral relationships.
-We even allow our oracles to contradict themselves by giving responses that would lead to a 
-cyclic graph when combined. If we use such oracles for DAG construction, we will
-introduce superfluous edges into our graph.
+Note that a $G$-compatible oracle can contradict itself. Consider 
+$V=(X,Y,Z)$, $G=(V,\{X \to Y\})$ and $H=(V,\{Y \to Z, Z \to X\})$.
+This gives $\mathcal{A}_{G \mid H}(Y,Z) = Y \to Z$  
+ and $\mathcal{A}_{G \mid H}(X,Z) = Z \to X$ which imply that $Y$ causes $X$, 
+in contradiction to
+ $\mathcal{A}_{G \mid H}(X,Y) = X \to Y$. Still, it will turn out that we can recover
+from such errors.
 
+Generally, using ancestral oracles for DAG construction will introduce superfluous edges.
 Our approach uses the data to decide where to add edges and which edges 
@@
 \SetKwProg{Fn}{Function}{:}{}
-\Fn{{\sc Expand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$B$,$k$)}{
+\Fn{{\sc Expand}($V,E,\mathcal{D},\mathcal{A}$,$B$,$k$)}{
     let $L \gets \{\}$\;
@@
         let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
-        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
-	    $L \gets L \cup \mathcal{A}_G(X, Y) $ \;
+        \If{$\mathcal{D}(X, Y, \mathbf{Z}) = 0$}{
+	    $L \gets L \cup \mathcal{A}(X, Y) $ \;
         }
@@
     }
-    $ R \gets \textsc{FixCycles}(V, E \cup L, \mathcal{D}_G) $ \; 
-    $ B \gets B \cup R \; ; \; L \gets L \setminus R \; ; \;  E \gets (E \cup L) \setminus B$ \;
+    $ R \gets \textsc{FixCycles}(V, E \cup L, \mathcal{D}) $ \; 
+    $ B \gets B \cup R \; ; \;  E \gets (E \cup L) \setminus R$ \;
     \Return{$(V,E,B)$}\;
@@
 \SetKwProg{Fn}{Function}{:}{}
-\Fn{{\sc FixCycles}($V, E, \mathcal{D}_G $)}{
+\Fn{{\sc FixCycles}($V, E, \mathcal{D} $)}{
 		$R \gets \emptyset$ \; 
 		\ForEach {$ X \rightarrow Y $ on a cycle in $ (V,E) $}{
-		\If {there exists a set $\mathbf{Z}\subseteq\mathbf{X}$ where $\mathcal{D}_G(X, Y, \mathbf{Z})=1$
+		\If {there exists a set $\mathbf{Z}\subseteq\mathbf{X}$ where $\mathcal{D}(X, Y, \mathbf{Z})=1$
 }{
@@
 let $\tilde{G}=\textsc{Expand}(V,\emptyset,\mathcal{D}_G,\mathcal{A}_H, \emptyset, \infty)$
-Then $H^+ \subseteq \tilde{G} \supseteq G$. If $H$ contains at least one cycle, then
-$H^+ \subsetneq \tilde{G}$.
+Then $G \subseteq \tilde{G} \subseteq H^+$. If $H$ contains at least one cycle, then
+$\tilde{G} \subsetneq H^+$.
 \label{prop:weakexpand:cyclic}
@@
 would be cyclic. At least one edge of every cycle is therefore removed by
- $\textsc{FixCycles}$, making the result acyclic again. Since the removed edges need
-to be in $H^+$, the first inclusion is strict for cyclic $H$.
+ $\textsc{FixCycles}$, making the result acyclic again. Since any edges removed from cycles need
+to be in $H^+$, the result must be a proper subgraph of $H^+$.
 \end{proof}
@@
 \SetKwProg{Fn}{Function}{:}{}
-\Fn{{\sc Prune}($V$,$E$,$\mathcal{D}_G$)}{
-    let $L \gets \{\}$\;
+\Fn{{\sc Prune}($V$,$E$,$\mathcal{D}$)}{
+    let $R \gets \{\}$\;
     \ForEach{$X \to Y \in E$}{
-        remove $X \to Y$ from $E$\;
-        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
-        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 1$}{
+        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ 
+		in $(V,E \setminus \{ X \to Y \} )$\;
+        \If{$\mathcal{D}(X, Y, \mathbf{Z}) = 1$}{
             $L \gets L \cup  \{X \to Y\}$ \;
@@
     }
-    remove all edges in $L$ from $G$\;
+    remove all edges in $R$ from $G$\;
     \Return{$(V, E)$}\;
@@
 
-By combining Propositions~\ref{prop:strongexpand}, \ref{prop:weakexpand}, 
+By combining Propositions~\ref{prop:strongexpand}, 
 \ref{prop:weakexpand:cyclic} and 
@@
 \begin{theorem}
-Given a d-separation oracle $\mathcal{D}_G$ and a $G$-consistent 
-ancestral oracle $\mathcal{A}_H$, 
-let $(V',E',B)=\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_H,\emptyset,\infty)$.
+Given a d-separation oracle $\mathcal{D}_G$ and a $G$-compatible 
+ancestral oracle $\mathcal{A}_{G\mid H}$, 
+let $(V',E',B)=\textsc{Expand}((V,\emptyset),\mathcal{D}_G,\mathcal{A}_{G\mid H},\emptyset,\infty)$.
 Then $\textsc{Prune}(V',E',\mathcal{D}_G)=G.$
@@
 
-\Fn{{\sc ExpertInLoop}($V,\mathcal{D}_G,\mathcal{A}_G$)}{
+\Fn{{\sc ExpertInLoop}($V,E,\mathcal{D},\mathcal{A}$)}{
 $E_p \gets \emptyset$ \tcc{Current edges} 
@@
 	$E \gets E_p$ \;
-	$(V,E,B) \gets \textsc{Expand}(V,E,\mathcal{D}_G,\mathcal{A}_G,B,1)$ \;
-	$(V,E_p) \gets \textsc{Prune}(V,E,\mathcal{D}_G)$ \;
+	$(V,E,B) \gets \textsc{Expand}(V,E,\mathcal{D},\mathcal{A},B,1)$ \;
+	$(V,E_p) \gets \textsc{Prune}(V,E,\mathcal{D})$ \;
 	$B \gets B \cup \{ E_p \setminus E \} $
@@
 \SetKwProg{Fn}{Function}{:}{}
-\Fn{{\sc RankedExpand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$\phi$,$B=\emptyset$)}{
+\Fn{{\sc RankedExpand}($V,E,\mathcal{D},\mathcal{A}$,$\phi$,$B$)}{
     let $\phi_{\max} \gets 0 $\;
-    \ForEach{$X, Y$ where $X \to Y \notin E \setminus B$ and $Y \to X \notin E \setminus B$}{
+    \ForEach{$X, Y$ where $X \to Y \notin E \cup B$ and $Y \to X \notin E \cup B$}{
         let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
-        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
+        \If{$\mathcal{D}(X, Y, \mathbf{Z}) = 0$}{
 		\If{$\lvert \phi(X, Y, \mathbf{Z}) \rvert > \phi_{\max}$}{
 			$ \phi_{\max} \gets \lvert \phi(X, Y, \mathbf{Z}) \rvert $ \;
-			$ e_{\max} \gets \mathcal{A}_G(X, Y)$ \;
+			$ e_{\max} \gets \mathcal{A}(X, Y)$ \;
 		}
@@
     }
-    $L' \gets \textsc{FixCycles}(V, E \cup L, \mathcal{D}_G)$ \;
-
-    \uIf{ $(V,E \cup e_{\max})$ is acyclic }{
-	add $ e_{\max}$ to $E$ \;
-    } \Else {
-	add $ \textrm{reverse}(e_{\max})$ to $E$ \;
-    }
-    \Return{$(V,E)$}\;
+    $ R \gets \textsc{FixCycles}(V, E \cup \{e_{\max}\}, \mathcal{D}) $ \; 
+    $ B \gets B \cup R \; ; \;  E \gets (E \cup  \{e_{\max}\}) \setminus R$ \;
+    \Return{$(V,E,B)$}\;
 }
 \caption{Adding an edge between variables with the highest correlation}
-\label{algo:ranked_expand}
+\label{algo:rankedexpand}
 \end{algorithm}
@@
 
-Algorithm~\ref{algo:ranked_expand} shows a 
+Algorithm~\ref{algo:rankedexpand} shows a 
 version of {\sc Expand} where we use the largest unexplained association to

### 2025-06-11 paper/main.tex
small fixes

@@
 \Fn{{\sc Expand}($V,E,\mathcal{D},\mathcal{A}$,$B$,$k$)}{
-    let $L \gets \{\}$\;
+    $L \gets \{\}$\;
     \ForEach{$X, Y$ where $X \to Y \notin E \cup B $ and $Y \to X \notin E \cup B$}{
@@
 $\mathbf{Z}$ be the $d$-separating set chosen 
-in line~5 of {\sc Prune} (there is at least one 
+in line~4 of {\sc Prune} (there is at least one 
 such $\mathbf{Z}$, the union of the parents of 

### 2025-06-11 paper/main.tex
discussion draft

@@
 
+It is possible to integrate domain knowledge into score-based methods
+in a similar way as we do here for constraint-based structure learning.
+\citet{Kitson_2025} recently extended the Tabu search algorithm by a procedure
+that asks a domain expert for advice before making changes that are classified as 
+``uncertain'' such as in cases of tied score improvement or low-confidence modifications. 
+In contrast, our approach places greater emphasis on expert input as no edges are oriented
+without consulting the expert. Further, the approach by \citet{Kitson_2025} 
+requires experts that are able to distinguish between direct and indirect effects.
+
 \section{Empirical Analysis}
@@
 Researchers often prefer to construct DAGs manually rather than using automated
-algorithms due to practical challenges in applying automated methods. To assist
-this process, we developed an iterative structure learning method that
+algorithms, which could be partly due to practical challenges in applying automated methods.
+To assist this process, we developed an iterative structure learning method that
 integrates manual construction with data-driven feedback, bridging
@@
 ``blacklists'' of edges 
-\citep{scutari2010}, or specification
-of a ``tiered'' (e.g., temporal) structure between the variables \citep{Bang2023}.
+\citep{scutari2010}, or specification of a ``tiered'' (e.g., temporal) 
+structure between the variables \citep{Bang2023}.
 Compared to such approaches, where domain knowledge 
@@
 result in a polynomial amount of steps in the oracle settings we considered.
+
+Nevertheless, the algorithm presented in this paper has 
+significant limitations. First of all, the causal Markov and causal sufficiency assumptions are
+widely seen as too restrictive, since they do not allow for latent confounders.
+It would therefore be interesting to extend our approach to a more general 
+algorithm such as FCI \cite{Spirtes1995}. Further, while it is theoretically possible
+to recover from some expert mistakes that introduce cycles, this requires searching for a 
+separating set for each edge on each cycle, which in the worst case can take exponential time 
+and would then essentially amount to running significant parts of the PC algorithm. 
+In practice, we may be better off breaking cycles by other means -- such as the heuristics
+implemented in our experiments -- even if the theoretical correctness guarantee is lost. 
+Lastly, any iterative model improvement procedure runs the risk of overfitting to the 
+given dataset. While this is also the case for purely data-driven structure learning
+(which may even require a vastly larger amount of model improvement steps), the 
+inclusion of human judgement in this process means that it will be seen as less reliable.
+Models generated by this approach should certainly be seen as preliminary and in need
+of further validation using independent data, and ideally a full log of all human decisions 
+made during the process should be kept. In this respect, it will be interesting to see how
+the research community will feel about the use of LLMs instead of humans for model refinement.
 

### 2025-06-12 code/fig3/sl.R
Expert results

@@ -292,9 +292,31 @@ run_sim <- function(R, n_nodes, edge_probs, oracle_accs){
 }
 
 
+run_sim_only_expert <- function(R, n_nodes, edge_probs, oracle_accs){
+	results <- data.frame()
+	
+	pb <- progress_bar$new(total=length(oracle_accs) * length(edge_probs))
+	for (edge_prob in edge_probs){
+		for (oracle_acc in oracle_accs){
+			human_dist <- t(future_replicate(R, run_single_exp_human(n_nodes=n_nodes, edge_prob=edge_prob, oracle_acc=oracle_acc)))
+
+			human_mean <- apply(human_dist, mean, MARGIN=2)
+			human_sd <- apply(human_dist, sd, MARGIN=2)/sqrt(R)
+			
+			results <- rbind(results, c(oracle_acc, edge_prob, human_mean, human_sd))
+			pb$tick()
+		}
+	}
+	colnames(results) <- c(
+		'oracle_acc', 'edge_prob', 
+		'human_shd_mean', 'human_sid_mean', 'human_shd_sd', 'human_sid_sd')
+	write.csv(results, 'results/sl_results_mixed_only_expert.csv')
+}
+
+
 edge_probs <- seq(0.1, 0.9, 0.1)
 oracle_accs <- seq(0, 0.9, 0.2)
 n_nodes <- 10
 R <- 30
 
-run_sim(R, n_nodes, edge_probs, oracle_accs)
+run_sim_only_expert(R, n_nodes, edge_probs, oracle_accs)

### 2025-06-12 code/fig3/make_plots.R
Fixed older plot

@@ -3,139 +3,16 @@ library(dplyr)
 library(tibble)
 library(directlabels)
 
-
-# make_plots <- function(filename, plot_type){
-# 	d <- read.csv(filename, row.names=1)
-# 	# d <- d[d$edge_prob %in% c(0.1, 0.3, 0.5, 0.7, 0.9), ]
-# 	d <- d[d$oracle_acc %in% c(0.1, 0.3, 0.5, 0.7, 0.9), ]
-# 
-# 	d$edge_prob <- sapply(d$edge_prob, as.character)
-# 	col_names <- c('oracle_acc', 'edge_prob', 'shd_mean', 'sid_mean', 'shd_sd', 'sid_sd', 'alg')
-# 
-# 	d_ges_lower <- d %>% select(oracle_acc, edge_prob, ges_lower_shd_mean, ges_lower_sid_mean, ges_lower_shd_sd, ges_lower_sid_sd) %>% add_column(alg='ges_lower')
-# 	colnames(d_ges_lower) <- col_names
-# 	d_ges_upper <- d %>% select(oracle_acc, edge_prob, ges_upper_shd_mean, ges_upper_sid_mean, ges_upper_shd_sd, ges_upper_sid_sd) %>% add_column(alg='ges_upper')
-# 	colnames(d_ges_upper) <- col_names
-# 
-# 	d_hc_lower <- d %>% select(oracle_acc, edge_prob, hc_lower_shd_mean, hc_lower_sid_mean, hc_lower_shd_sd, hc_lower_sid_sd) %>% add_column(alg='hc_lower')
-# 	colnames(d_hc_lower) <- col_names
-# 	d_hc_upper <- d %>% select(oracle_acc, edge_prob, hc_upper_shd_mean, hc_upper_sid_mean, hc_upper_shd_sd, hc_upper_sid_sd) %>% add_column(alg='hc_upper')
-# 	colnames(d_hc_upper) <- col_names
-# 	
-# 	d_pc_lower <- d %>% select(oracle_acc, edge_prob, pc_lower_shd_mean, pc_lower_sid_mean, pc_lower_shd_sd, pc_lower_sid_sd) %>% add_column(alg='pc_lower')
-# 	colnames(d_pc_lower) <- col_names
-# 	d_pc_upper <- d %>% select(oracle_acc, edge_prob, pc_upper_shd_mean, pc_upper_sid_mean, pc_upper_shd_sd, pc_upper_sid_sd) %>% add_column(alg='pc_upper')
[…]
